### 2017-05-24 HinesHetlandGPR.tex
initial commit

\documentclass[10pt,a4paper]{article}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
\usepackage{arydshln}
\title{Revealing Transient Strain in Geodetic Data with the Radial Basis Function Finite Difference Method}
\author{Trever T. Hines and Eric A. Hetland}

% OUTLINE
% INTRODUCTION
% - Why we care about strain in the crust
%   - Knowing long term strain rates tells us where we can expect seismicity
%
% - How people have gone about estimating strain rates
%   - Draw distinction between parametric and nonparametric methods
%   - Draw distinction between model based and data based parametric methods 
%   - Talk about Kato
%
% - Another, more challenging problem is how to calculate time-dependent strain
%   - Time dependent strain rates could illuminate transient geophysical signals such as postseismic deformation, 
%	  slow slip events, or volcanic deformation. 	  
%	- Since numerous megathrust earthquakes were preceded by slow slip, it is important to detect these features.
%	- We can also use estimates of transient strain to calibrate strain meters.	
%	- People who have tackled this problem before include Ohtani and Holt.	
%	- A major deficiency with these methods is that they are parametric giving poor uncertainties. Without good 
%	  uncertainties we cannot objectively discern signal from noise and we cannot calibrate BSM.	
% - Here we propose GPR and we tackle several two major difficulties	
%   - The method is similar to an extension of Kato to the time domain. We address two debilitating difficulties that 
%     arise. We also dont do that autocovariance bullshit.  Instead we use REML.  
%	- outliers  
%	- dense matrix
%	
% - GPR can be used for more than just estimating strain and it is a powerful geophysical tool. We can use it to quantify
%   spatial noise in GPS data. (this should be a side blurb when talking about quantifying signal.   
%
% - We discuss the method, form prior with REML, and demonstrate utility in detecting SSE by comparing it to tremor.	  
%
% METHODS
% - define model
% - discuss known deficiencies with the model
% - describe covariance as a time-space separable function 
% - describe geophysically relevant gaussian processese
% - REML for determining optimal hyperparameters
% - conditioning the prior with observations
% - differentiating the Gaussian process
% - note on addition, subtraction, and scaling
% - REML
%
% APPLICATION TO CASCADIA SSE
% - SSE background
% - REML results
%   - noise 
%   - signal for IBM, SE, and WEN

% - Strain results
%   - map view plot
%   - time series plot

% CONCLUSION
% - future work
% - additional applications other than sse. Discuss potential use for noise modeling

\begin{document}

\maketitle

\section{Introduction}\label{sec:Introduction}

Crustal strain rates are fundamentally important quantities for assessing seismic hazard, since knowing where and how quickly strain is accumulating gives insight into where we can expect stored elastic energy to be released seismically.  It is then important to develop and improve upon methods for mapping strain in tectonically active regions because such maps could conceivably feed into seismic hazard models such as UCERF3 \citep{Field2014}.

Methods for estimating strain from geodetic data fall in one of two categories.  There are model-based approaches which assume that strain is the result of loading on faults which have a known geometry, and there are data-based approaches which make no assumptions about the source of deformation.  We will exclusively consider data-based approaches in this paper.  The classic and simplest method for estimating strain is to assume that the strain rate is constant in time and spatially uniform within subnetworks of the geodetic data.  Linear least squares is then used to find the components of the strain rate tensor for each subnetwork \citep[e.g][]{Frank1966,Prescott1976,Savage1986,Feigl1990,Murray2000}. Several algorithms have been developed to improve upon this procedure. \citet{Shen1996} and \citet{Shen2015} discuss an algorithm where, instead of using the immediately adjacent stations to calculate strain at a position, the strain is computed with a weighted average over the entire network where the weighting is smaller for more distant stations.  Another strategy is to fit a set of interpolating basis functions to the deformation field and then compute the strain from the analytical derivative of the interpolant \citep[e.g.][]{Beavan2001,Tape2009,Sandwell2016}.

The aforementioned studies have all been concerned with estimating long term strain rates. Time dependent strain would be useful for studying geophysical processes which occur over timescales of days to years such as slow slip events, postseismic relaxation, or volcanic deformation.  \citet{Ohtani2010} identified transient strain events by fitting a set of spatial wavelet basis functions to the deformation field at discrete time epochs, and a Kalman filtering strategy was used to ensure that the coefficients for each basis function varied smoothly in time. \citet{Holt2013} calculated time dependent strain by differentiating a bicubic interpolant that was fit to each epoch of a temporally smoothed deformation field.

Each of the methods described above are designed to overcome two complications that arise in estimating deformation gradients: (1) geodetic data are noisy and differentiation will only amplify the noise and (2) geodetic data are not observed on a regular grid, which prevents the use of standard finite difference methods for computing derivatives. In this paper we demonstrate that both of these complications can be elegantly handled with the Radial Basis Function-Finite Difference (RBF-FD) method.

The RBF-FD method was developed simultaneously and independently by \citet{Tolstykh2003}, \citet{Shu2003}, \citet{Cecil2004}, and \citet{Wright2006} as a computationally efficient way to solve large scale partial differential equations over irregular, multi-dimensional domains.  The RBF-FD method can be thought of as a generalization of the traditional finite difference method, where the node layout is no longer restricted to regular grids. Indeed, the RBF-FD method can be used to estimate derivates of discrete data located at arbitrary scattered positions in multi-dimensional space.  The RBF-FD method is particularly appealing because it is algorithmically simple, regardless of the domain shape or node layout, and also because the method has performed well in numerous benchmark tests \citep[and references therein]{Fornberg2015}.

In this paper, we do not use the RBF-FD method to solve a partial differential equation, but rather we use it to spatially smooth geodetic data and to compute deformation gradients. Our smoothing strategy can be viewed as a non-parametric, low-pass filter for scattered data where the degree of smoothness is controlled by a user specifies cutoff frequency. This can be contrasted with interpolation based methods where the resulting interpolant can be largely and perhaps unpredictably controlled by the choice of basis function. This process of smoothing and differentiating can be extended to estimate time dependent strain rates.  In that case, we first temporally smooth and differentiate GPS displacement time series to get time dependent velocities.  We then spatially smooth and differentiate the resulting velocities for each time epoch to get time dependent strain rates.

The method proposed in this paper has numerous advantages which set it apart from other methods for computing strain rates. The method is computationally efficient and stable (there is no inversion of an ill-conditioned matrix).  There are no hyper parameters or penalty parameters that need to be tuned for each application.  As opposed to interpolation strategies such as \citet{Beavan2001}, \citet{Tape2009}, or \citet{Ohtani2010}, our method assumes that velocities are locally rather than globally continuous, which allows us to easily handle discontinuities resulting from, for example, a creeping fault.

We begin this paper by summarizing the RBF-FD scheme and explaining how we construct differentiation matrices for scattered data. We then introduce the RBF-FD filter, which is used to smooth the observed geodetic data prior to differentiation.  We provide two real world demonstraitions of our method for calculating strain rates.  First we calculate the long term strain rates in Southern California from the CMM3 velocity data set \citep{Shen2011}, and we verify that our results are consistent with other studies. We then calculate time dependent strain rates in Cascadia from the GPS data provided by UNAVCO.  In Cascadia, we analyze strain resulting from slow slip events and compare it to the long term tectonic strain accumulation. Slow slip events are found to produce compression in the Olympic Peninsula, which is in addition to the compression resulting from tectonic loading.  Further south in Oregon, the slow slip events tend to release the compressional strain that is accumulated tectonically.  While similar conclusions have been drawn from fault slip inversions for slow slip events, it is important to recognize that slip inversion are the product of inverting an ill-conditioned matrix making it difficult to determine whether slip inferences are real or just an artifact of the inversion.  The strain rates presented in this paper are more direct observations and can be interpretted with a higher degree of confidence.

\section{Method}\label{sec:Method}

\subsection{Time Depenent Strain Rate in Cascadia}\label{sec:ApplicationsCascadia}

BACKGROUND

\citet{Dragert2001} first discovered slow slip events

Slow slip event depths from \citet{Dragert2001}, \citet{Wech2009}, \citet{Schmidt2010}, \citet{Bartlow2011} show slip concentrated at depths from 30 to 50 km detph.

Interseismic locking depths from \citet{Fluck1997}, \citet{Murray2000}, \citet{McCaffrey2007} and \citet{McCaffrey2013}, \citet{Burgette2009}, \citet{schmalzle2014} are consistent with a full locking down to about 20 km.

%Studies which simultaneously looked at intersesimic and ets are \citet{Holtkamp2010} and \citet{Schmalzle2014}
Studies which simultaneously modeled interseismic and ets are \citet{Holtkamp2010} and \citet{schmalzle2014}

METHOD

We use GPS displacement time series to make daily estimates of strain rates from 2010-01-01 to 2016-07-01 in the Pacific Northwest.  Our method consists of two distinct steps; we temporally smooth and differentiate the displacement time series for each station, and then we spatially smooth and differentiate the velocity field for each day.  We temporally smooth the GPS displacements with the method described in \ref{sec:Filter}, and we treat days with missing data as observations with infinite uncertainty. As noted above, this is effectively equivalent to fitting a smoothing spline to each time series, which can also be efficiently done with a Kalman filtering strategy \citep{Kohn1987}.  We chose a cutoff frequency of ZZZ, which is high enough to acurately describe deformation during a slow slip event.  Spatial smoothing is done with the RBF-FD filter and differentiation to get the strain rates was done with the RBF-FD scheme.  We picked a cutoff frequency of XXX for smoothing.

\section{Discussion and Conclusion}\label{sec:Discussion}
% discuss potential applications for slip models

% discuss how we remove errors like reference frame and seasonals

% discuss how this can be used for tikhonov regularization

% we do not interpolate the strain field because it may introduce spurious artifacts Baxter2011

% Discuss similarities with PCA used in Dong2006 and common mode error by Wdowinski1997
% note that these errors are due to long wavelength features which get removed in strain calculations!!!
% see strain method by almendinger2007

% mention the scec geodetic transient exercise Lohman2013 and references therein

% Note that segalls 2016 paper mentions the gap between interseismic and sse

% mention howell2016, who said some crap about verticals, mention Hammond2016 so said more sensible crap about verticals

% note that this can be used for smoothing verticals, and compare it with burgette2009

% change the notation for stencil size or smoothness order because they are both n. use p for smoothness order

The material presented in this paper is primarily focused on GPS data; however, we speculate that the RBF-FD scheme can be of particular use in denoising borehole strain meter (BSM) data.  The Plate Boundary Observatory has deployed X BSMs along the Western United States.  BSM data contains low frequency drift resulting from relaxation of the borehole \citep{Gladwin1987}, which can obscure the geophysical signal of interest.  We suggest that the RBF-FD scheme may be useful in denoising BSM data.  Since the RBF-FD scheme provides a straight-forward mapping from GPS displacements to strain at any target locations, it is possible to use GPS derived strains as \textit{a priori} information for strain at BSM sites. GPS derived strains could then aid in discerning tectonic signal from noise in BSM data.

Additional uses for strain: transient detection, prior for strain meters,

Additional potential applications of the RBF-FD method: regularizing inverse problems,

% cite eric calais for triangulation strain

\bibliographystyle{apalike}
\bibliography{mybib}

\end{document}

### 2017-05-24 HinesHetlandGPR.tex
added more

@@
 % - describe covariance as a time-space separable function 
-% - describe geophysically relevant gaussian processese
+% - describe geophysically relevant gaussian processes. (include seasonal and spherical)
 % - REML for determining optimal hyperparameters
@@
 % - REML
+% - Numerical notes
+%   - partition inverse and solve with CHOLMOD
 %
@@
 %   - signal for IBM, SE, and WEN
+%   - Plot all covariances functions
 
@@
 
+In this section, we describe how GNSS displacement observations are used to infer transient strain rates.
+We consider transient strain rates to be any deviation from the steady rate of strain accumulation resulting from plate tectonics.
+We model transient deformation, $u(x,t)$, as a three-dimensional (two spatial dimensions and time) Gaussian process,
+\begin{equation}\label{eq:TransientDeformation}
+u(x,t) \sim \mathcal{N}\left(0,C_u\right),
+\end{equation}
+where the covariance function, $C_u$, can be separated into positive definite spatial and temporal functions as 
+\begin{equation}\label{eq:TransientCovariance}
+C_u(x,t;x',t') = X(x,x')T(t,t').
+\end{equation}  
+The appropriate choice for $X$ and $T$ may vary depending on the geophysical signal we are trying to describe (e.g.
+postseismic deformation or deformation from slow slip events), and we discuss this matter in the next section.
+GNSS observations record transient deformation as well as other physical and non-physical processes which we are not interested in.
+We describe GNSS data that is observed at position $x_i$ and time $t_j$ as
+
+\begin{equation}\label{eq:Data}
+d_{ij} = a_{i} + b_{i} t_j + c_{1i}\sin(2 \pi t) + c_{2i}\cos(2 \pi t) + c_{3i}\sin(4 \pi t) + c_{4i}\cos(4 \pi t) + u(x_i,t_j) + \epsilon(x_i,t_j),
+\end{equation}
+where $a_i$ is an offset, $b_i$ is the steady rate of tectonic deformation at $x_i$, the sinusoids describe seasonal deformation, and $\epsilon_{ij}$ is a Gaussian process for noise which cannot be described parametrically.
+$\epsilon$ can consist of white noise, temporally correlated noise describing benchmark wobble \citep[e.g.,][]{Wyatt1982,Wyatt1989}, or spatially correlated noise describing common mode error \citep[e.g.,][]{Wdowinski1997}.
+For now, we just describe $\epsilon$ as $\epsilon \sim \mathcal{N}(0,C_d)$.
+In this study, we consider the coefficients $a_{i}$, $b_{i}$, and $c_{kj}$, to be unknown and uncorrelated between stations.
+Of course, the tectonic deformation, $b_i$, is in reality spatially correlated and we could invoke tectonic model to form a prior on $b_i$.
+However, in our application to Cascadia, we will be using displacement time series which are long enough to sufficiently constrain $b_i$ for each station and there is no need to incorporate a prior.
+Likewise, the seasonal coefficients may be spatially correlated as suggested by Langbein2008?.
+It may be worth exploring and exploiting such a correlation in a future study.
+
 \subsection{Time Depenent Strain Rate in Cascadia}\label{sec:ApplicationsCascadia}

### 2017-05-25 HinesHetlandGPR.tex
added more

--- 2017-05-24 HinesHetlandGPR.tex
@@
 \usepackage{arydshln}
+\usepackage{bm}
 \title{Revealing Transient Strain in Geodetic Data with the Radial Basis Function Finite Difference Method}
@@
 
+In this section, we describe how GNSS displacement observations are used to identify transient crustal strain rates, which we denote as $\dot{\bm{\varepsilon}}(x,t)$.
+We consider $\dot{\bm{\varepsilon}}$ to be spatially and temporally coherent deviations from the steady rate of strain accumulation from plate tectonics.
+We determine $\dot{\bm{\varepsilon}}$ by first identifying transient displacements, $\bm{u}(x,t)$, which we then spatially and temporally differentiate.
+As we will show in Section X, estimates of $\dot{\bm{\varepsilon}}$ turn out to be more effective at illuminating geophysical signal than estimates of $\bm{u}$ or $\dot{\bm{u}}$.
+We make a prior assumption that each component of $\mathbf{u}$ is a three-dimensional (two spatial dimensions and time) Gaussian process,
+\begin{equation}\label{eq:TransientDeformation}
+u_i(x,t) \sim \mathcal{N}\left(0,C_{u_i}\right),
+\end{equation}
+where $C_{u_i}(x,t;x',t')$ is a covariance function indicating how we expect $u_i(x,t)$ to covary with $u_i(x',t')$.
+For simplicity, we treat each component of displacement independently and ignore any potential covariance.
+We then drop the component subscripts with the understanding that the same analysis is being repeated to estimate the east, north, and vertical components of $\bm{u}$.
+We further assume that $C_u$ can be separated into positive definite spatial and temporal functions as
+\begin{equation}\label{eq:TransientCovariance}
+C_{u}(x,t;x',t') = X(x,x')T(t,t').
+\end{equation}  
+The appropriate choice for $X$ and $T$ may vary depending on the geophysical signal we are trying to describe (e.g.
+postseismic deformation or deformation from slow slip events), and we discuss this matter in the next section.
+
+\subsection{Spatial and temporal covariance functions}
+In this section, we list several covariance functions which are considered in this paper or could be of use in future studies. For see \citet{Rasmussen2006} for additional information.
+
+The squared exponential covariance function,
+\begin{equation}\label{eq:SE}
+C(x,x') = \alpha^2 \exp\left(\frac{-||x - x'||_2^2}{2 \beta^2}\right),
+\end{equation}
+is commonly used in Kriging \citep[e.g,][]{Cressie1992} and Gaussian process regression \citep[e.g.,][]{Rasmussen2006}.
+The squared exponential is a valid (i.e.
+positive definite) covariance function for any number of spatial dimensions, and it describes an isotropic Gaussian processes with realizations that are infinitely differentiable.
+\citet{Kato1998} demonstrated that eq.
+(\ref{eq:SE}) is an appropriate covariance model for describing long-term tectonic strain rates.
+
+The Wendland class of covariance functions \citep{Wendland2005} are positive definite in $\mathbb{R}^d$, and they describes an isotropic Gaussian process with realizations that can be differentiated $k$ times.
+The form of the covariance function depends on the choice of $d$ and $k$.
+Wendland covariance functions have compact support, which is a particularly useful property because we can exploit the sparsity of the the corresponding covariance matrices.
+In this study we use Wendland covariance functions to describe the temporal component of transient displacements.
+Therefore, we only require that $d=1$ and $k=1$.
+The corresponding Wendland covariance function is
+\begin{equation}\label{eq:Wendland}
+C(x,x') = \alpha^2\left(1 - \frac{|x - x'|}{\beta}\right)^3_+ \left(\frac{3|x - x'|}{\beta} + 1\right), \ \ \ x \in \mathbb{R}^1,
+\end{equation}
+where
+\begin{equation}
+(x)_+ = 
+\begin{cases}
+x, \ \ \ x > 0 \\
+0, \ \ \ \mathrm{otherwise}.
+\end{cases}
+\end{equation}
+
+They are positive definite for a finite dimensional space and are finitely differentiable.
+If the length-scale is sufficiently small then we can take advantage of the sparsity of the covariance matrices in Section X.
+This turns out to significantly increase scale of the problems which can be addressed in this analysis.
+
+FOGM solves the differential equation .... Has been used by Langbein.
+
+Valid in ND
+
+Brownian Motion is the solution to the differential equation .... Has been used by Langbein to describe noise
+Valid in 1D
+
+Integrated Brownian motion solves ... . It has been used by Segall to describe slip, Ohatani, Murray, 
+Valid in 1D
+
+We also want to give notable mention to the Periodic covariance function
+
+\subsection{Constraining displacements and strain with GNSS data}
+We constrain $u$ with GNSS data, which records $u$ as well as other physical and non-physical processes which we are not interested in.
+We describe GNSS data, $\bm{d}_*$, as the realization of a random vector $\bm{d} = [d_{ij}]^T$, where $d_{ij}$ describes observations at position $x_i$ and time $t_j$.
+We assume that the observations can be described as
+\begin{align}\label{eq:Data}
+d_{ij} = &u(x_i,t_j) + \epsilon(x_i,t_j) + a_{i} + b_{i} t_j + \\
+         &c_{1i}\sin(2 \pi t_j) + c_{2i}\cos(2 \pi t_j) + c_{3i}\sin(4 \pi t_j) + c_{4i}\cos(4 \pi t_j), 
+\end{align}
+where $a_i$ is an offset that is unique for each GNSS monument, $b_i$ is the steady rate of tectonic deformation at $x_i$, the sinusoids describe seasonal deformation (using units of years for $t$), and $\epsilon$ is a Gaussian process for noise that cannot be described parametrically.
+$\epsilon$ can consist of white noise, temporally correlated noise describing benchmark wobble \citep[e.g.,][]{Wyatt1982,Wyatt1989}, or spatially correlated noise describing common mode error \citep[e.g.,][]{Wdowinski1997}.
+For now, we describe $\epsilon$ generally as $\epsilon \sim \mathcal{N}(0,C_d)$.
+We consider the coefficients $a_{i}$, $b_{i}$, and $c_{kj}$, to be unknown and uncorrelated random variables.
+Of course, the tectonic deformation, $b_i$, is in reality spatially correlated and we could invoke tectonic model to form a prior on $b_i$.
+However, in our application to Cascadia, we will be using displacement time series which are long enough to sufficiently constrain $b_i$ for each station and there is no need to incorporate a prior.
+Likewise, the seasonal coefficients may be spatially correlated as suggested by Langbein2008?.
+It may be worth exploring and exploiting such a correlation in a future study.
+
+Optimal Hyperparameters
+
+Conditioning the GP
+
+Differentiation, scaling and addition to get strain
+
 \subsection{Time Depenent Strain Rate in Cascadia}\label{sec:ApplicationsCascadia}

### 2017-05-27 HinesHetlandGPR.tex
added more

@@
 \subsection{Spatial and temporal covariance functions}
-In this section, we list several covariance functions which are considered in this paper or could be of use in future studies. For see \citet{Rasmussen2006} for additional information.
+In this section, we list several covariance functions which are considered in this paper or could be of use in future studies.
 
@@
 The squared exponential is a valid (i.e.
-positive definite) covariance function for any number of spatial dimensions, and it describes an isotropic Gaussian processes with realizations that are infinitely differentiable.
-\citet{Kato1998} demonstrated that eq.
-(\ref{eq:SE}) is an appropriate covariance model for describing long-term tectonic strain rates.
+positive definite) covariance function for any number of spatial dimensions, and it describes an isotropic (i.e.
+$C(x,x')$ can be expressed as $\hat{C}(||x - x'||_2)$) Gaussian processes with realizations that are infinitely differentiable.
+\citet{Kato1998} and \cite{El-Fiky1999} demonstrated that eq.
+(\ref{eq:SE}) is an appropriate covariance model for describing long-term tectonic strain rates for Japan.
 
@@
 \begin{equation}\label{eq:Wendland}
-C(x,x') = \alpha^2\left(1 - \frac{|x - x'|}{\beta}\right)^3_+ \left(\frac{3|x - x'|}{\beta} + 1\right), \ \ \ x \in \mathbb{R}^1,
+C(x,x') = \alpha^2\left(1 - \frac{|x - x'|}{\beta}\right)^3_+ \left(\frac{3|x - x'|}{\beta} + 1\right), \ \ \ x,x' \in \mathbb{R}^1,
 \end{equation}
@@
 
-They are positive definite for a finite dimensional space and are finitely differentiable.
-If the length-scale is sufficiently small then we can take advantage of the sparsity of the covariance matrices in Section X.
-This turns out to significantly increase scale of the problems which can be addressed in this analysis.
+We next consider the covariance function which describes a first-order Gauss-Markov (FOGM) process. A FOGM process is a solution to the stochastic differential equation
+\begin{equation}
+\dot{u}(x) + \alpha u(x) = \beta w(x), \ \ \ x,x' \in \mathbb{R}^1,
+\end{equation}
+where $w(x)$ is white noise with unit variance.  The solution is a Gaussian process with zero mean and covariance
+% This needs to be scaled to match my work
+\begin{equation}\label{eq:FOGM}
+C(x,x') = \frac{\beta^2}{2\alpha}\exp\left(-\alpha|x - x'|\right), \ \ \ x,x' \in \mathbb{R}^1. 
+\end{equation}
+FOGM processes are isotropic and realizations are not differentiable.
+The FOGM model has been used to describe temporally correlated noise in EDM data \citep{Langbein2004} and GNSS data \citep[e.g.,][]{Langbein2008}.
 
-FOGM solves the differential equation .... Has been used by Langbein.
+One last covariance function which is worth mentioning, even though we do not use it in this paper, is the periodic covariance function \citep{Mackay1998}, 
+\begin{equation}\label{eq:Periodic}
+C(x,x') = \alpha^2 \exp\left(\frac{-\sin(\frac{\pi}{\lambda} |x - x'|)^2}{2 \beta^2} \right), \ \ \ x,x' \in \mathbb{R}^1.
+\end{equation}
+The periodic covariance function describes an isotropic Gaussian process with realizations that have period $\lambda$.
+The length-scale of features within each period is controlled by $\beta$.
+In terms of GNSS data, a periodic Gaussian process can be used to describe seasonal deformation.
+The more common method of treating seasonal deformation is to model it as a linear combination of annual and semi-annual sinusoids \citep[e.g.,][]{Blewitt2002}.
+We believe there is potential to use periodic Gaussian processes to explore the statistical properties of seasonal signals; however, that is outside the scope of this paper.
 
-Valid in ND
-
-Brownian Motion is the solution to the differential equation .... Has been used by Langbein to describe noise
-Valid in 1D
-
-Integrated Brownian motion solves ... . It has been used by Segall to describe slip, Ohatani, Murray, 
-Valid in 1D
-
-We also want to give notable mention to the Periodic covariance function
+Brownian motion and integrated Brownian motion are two Gaussian processes which are frequently used in the geodetic literature.  Brownian motion, which has zero mean and covariance function
+\begin{equation}\label{eq:BM}
+C(x,x') =\alpha^2 \min(x,x'), \ \ \ x,x' \in \mathbb{R}^1 \ \mathrm{and} \ \ x,x' \geq 0, 
+\end{equation} 
+is commonly used to describe noise resulting from the instability of geodetic monuments \citep[e.g.,][]{Wyatt1982,Wyatt1989,Langbein1997}.
+Integrated Brownian motion, which has zero mean and covariance function
+\begin{align}\label{eq:IBM}
+C(x,x') &= \int_0^x \int_0^{x'} \alpha^2 \min(u,u') \,du'\,du \\
+        &= \frac{\alpha^2}{2}\min(x,x')^2 \left(\max(x,x') - \frac{1}{3}\min(x,x')\right), \ \ \ x,x' \in \mathbb{R}^1 \ \mathrm{and} \ \ x,x' \geq 0,
+\end{align}
+has been used as a nonparametric model for the time dependence of geophysical signals \citep[e.g.,][]{Segall1997,McGuire2003,Ohtani2010,Hines2016}.
+It should be noted that Brownian motion and integrated Brownian motion have a reference time at which the Gaussian process is exactly zero.
+For some geophysical signals, such as postseismic deformation, it is reasonable to have such a reference time.
+On the other hand, if there is no sense of a start time for the geophysical process, then Brownian motion or integrated Brownian motion may not be appropriate, and one of the isoptropic Gaussian processes described above should be used.
 

### 2017-05-29 HinesHetlandGPR.tex
added more

@@
 \end{equation}
-where $C_{u_i}(x,t;x',t')$ is a covariance function indicating how we expect $u_i(x,t)$ to covary with $u_i(x',t')$.
+where $C_{u_i}\left((x,t),(x',t')\right)$ is a covariance function indicating how we expect $u_i(x,t)$ to covary with $u_i(x',t')$.
 For simplicity, we treat each component of displacement independently and ignore any potential covariance.
@@
 \begin{equation}\label{eq:TransientCovariance}
-C_{u}(x,t;x',t') = X(x,x')T(t,t').
+C_{u}\left((x,t),(x',t')\right) = X(x,x')T(t,t').
 \end{equation}  
@@
 
+We constrain $u$ with GNSS data, which records $u$ as well as other physical and non-physical processes which we are not interested in.
+We describe GNSS observations at position $x_i$ and time $t_j$ as a realization of the random variable
+\begin{align}\label{eq:Data}
+d_{ij} = &u(x_i,t_j) + \epsilon(x_i,t_j) + m_{1i} + m_{2i} t_j + \\
+         &m_{3i}\sin(2 \pi t_j) + m_{4i}\cos(2 \pi t_j) + m_{5i}\sin(4 \pi t_j) + m_{6i}\cos(4 \pi t_j), 
+\end{align}
+where $m_1i$ is an offset that is unique for each GNSS monument, $m_2i$ is the steady rate of tectonic deformation at $x_i$, the sinusoids describe seasonal deformation (using units of years for $t_j$), and $\epsilon$ is a Gaussian process for noise that cannot be described parametrically.
+$\epsilon$ can consist of white noise, temporally correlated noise describing benchmark wobble \citep[e.g.,][]{Wyatt1982,Wyatt1989}, or spatially correlated noise describing common mode error \citep[e.g.,][]{Wdowinski1997}.
+For now, we describe $\epsilon$ generally as $\epsilon \sim \mathcal{N}(0,C_\epsilon)$.
+We consider the coefficients $m_{kj}$, to be uncorrelated random variables with a diffuse prior.
+Of course, the tectonic deformation, $b_i$, is in reality spatially correlated and we could invoke tectonic model to form a prior on $b_i$.
+However, in our application to Cascadia, we will be using displacement time series which are long enough to sufficiently constrain $b_i$ for each station and there is no need to incorporate a prior.
+Likewise, the seasonal coefficients may be spatially correlated as suggested by Langbein2008?.
+It may be worth exploring and exploiting such a correlation in a future study.
+
 \subsection{Spatial and temporal covariance functions}
-In this section, we list several covariance functions which are considered in this paper or could be of use in future studies.
+We must choose an appropriate Gaussian process describing transient displacements $u$ and GNSS noise $\epsilon$.
+We assume that $u$ and $\epsilon$ both have zero mean, which leaves for use to select a parameterization for the covariance functions $X$, $T$, and $C_d$.
+In this section, we list the covariance functions which are considered in this paper and those which could be of use in future studies.
+In the following discussion $\alpha$ and $\beta$ are hyperparameters.
+There are numerous strategies for selecting these hyperparameters (see \citet{Cressie1992}) and we save a discussion on choosing the hyperparameters for Section X.
 
@@
 In terms of GNSS data, a periodic Gaussian process can be used to describe seasonal deformation.
-The more common method of treating seasonal deformation is to model it as a linear combination of annual and semi-annual sinusoids \citep[e.g.,][]{Blewitt2002}.
+The more common method of treating seasonal deformation is to model it as a linear combination of annual and semi-annual sinusoids \citep[e.g.,][]{Dong2002}.
 We believe there is potential to use periodic Gaussian processes to explore the statistical properties of seasonal signals; however, that is outside the scope of this paper.
@@
 
-\subsection{Constraining displacements and strain with GNSS data}
-We constrain $u$ with GNSS data, which records $u$ as well as other physical and non-physical processes which we are not interested in.
-We describe GNSS data, $\bm{d}_*$, as the realization of a random vector $\bm{d} = [d_{ij}]^T$, where $d_{ij}$ describes observations at position $x_i$ and time $t_j$.
-We assume that the observations can be described as
-\begin{align}\label{eq:Data}
-d_{ij} = &u(x_i,t_j) + \epsilon(x_i,t_j) + a_{i} + b_{i} t_j + \\
-         &c_{1i}\sin(2 \pi t_j) + c_{2i}\cos(2 \pi t_j) + c_{3i}\sin(4 \pi t_j) + c_{4i}\cos(4 \pi t_j), 
-\end{align}
-where $a_i$ is an offset that is unique for each GNSS monument, $b_i$ is the steady rate of tectonic deformation at $x_i$, the sinusoids describe seasonal deformation (using units of years for $t$), and $\epsilon$ is a Gaussian process for noise that cannot be described parametrically.
-$\epsilon$ can consist of white noise, temporally correlated noise describing benchmark wobble \citep[e.g.,][]{Wyatt1982,Wyatt1989}, or spatially correlated noise describing common mode error \citep[e.g.,][]{Wdowinski1997}.
-For now, we describe $\epsilon$ generally as $\epsilon \sim \mathcal{N}(0,C_d)$.
-We consider the coefficients $a_{i}$, $b_{i}$, and $c_{kj}$, to be unknown and uncorrelated random variables.
-Of course, the tectonic deformation, $b_i$, is in reality spatially correlated and we could invoke tectonic model to form a prior on $b_i$.
-However, in our application to Cascadia, we will be using displacement time series which are long enough to sufficiently constrain $b_i$ for each station and there is no need to incorporate a prior.
-Likewise, the seasonal coefficients may be spatially correlated as suggested by Langbein2008?.
-It may be worth exploring and exploiting such a correlation in a future study.
+\subsection{Conditioning}
+Assuming that a covariance function and corresponding hyperparameters for $X$, $T$, and $C_d$ have been chosen,
+
+In this section we describe how to condition the prior for transient displacements, $u$, with GNSS observations.
+We will assume that a appropriate covariance functions and corresponding hyperparameters for $X$, $T$, and $C_d$ have already been chosen.
+Let $\mathbf{y}$ be the set of all $(x_i, t_j)$ pairs where a GNSS observation has been made, and let $\mathbf{d} = [d_{ij}]^T_{(x_i,t_j) \in \mathbf{y}}$.
+The random vector $\mathbf{d}$ is distributed as $\mathbf{d} \sim \mathcal{N}(\bar{\mathbf{d}},\mathbf{C_d})$, where
+\begin{equation}
+\mathbf{d}
+\end{equation}
+
+We consider the column vector of GNSS observations, $\mathbf{d}_*$, to be realizations of $d_{ij}$ at the positions and times in $\mathbf{y}$ (i.e, $\mathbf{d}_* = [d_{ij}(\omega_*)]^T_{(x_i,t_j) \in \mathbf{y}}$).
+We want to find the posterior estimate of transient displacements, $u | \mathbf{d}_*$, which incorporates our prior and the GNSS observations.
+
+Once we have chosen covariance functions and corresponding hyperparameters for $X$, $T$, and $C_d$, we then condition our prior Gaussian process, $u$, with the observations $d_*$ to form a posterior estimate of transient displacements, $u|d_*$.
+$d_*$ is a realization of the
+
+Whichever the Gaussian processes describing transient displacements, $u$, and noise in the GNSS data, $\epsilon$, have covariance functions which
 

### 2017-05-29 HinesHetlandGPR.tex
added more

@@
 
-In this section, we describe how GNSS displacement observations are used to identify transient crustal strain rates, which we denote as $\dot{\bm{\varepsilon}}(x,t)$.
+In this section, we describe how GNSS displacement observations are used to identify transient crustal strain rates, which we denote as $\dot{\bm{\varepsilon}}(z)$, where $z$ is the ordered pair $(x,t)$, $x$ are spatial coordinates in $\mathbb{R}^2$, and $t$ is time.
 We consider $\dot{\bm{\varepsilon}}$ to be spatially and temporally coherent deviations from the steady rate of strain accumulation from plate tectonics.
-We determine $\dot{\bm{\varepsilon}}$ by first identifying transient displacements, $\bm{u}(x,t)$, which we then spatially and temporally differentiate.
+We determine $\dot{\bm{\varepsilon}}$ by first identifying transient displacements, $\bm{u}(z)$, which we then spatially and temporally differentiate.
 As we will show in Section X, estimates of $\dot{\bm{\varepsilon}}$ turn out to be more effective at illuminating geophysical signal than estimates of $\bm{u}$ or $\dot{\bm{u}}$.
-We make a prior assumption that each component of $\mathbf{u}$ is a three-dimensional (two spatial dimensions and time) Gaussian process,
+We make a prior assumption that each component of $\bm{u}$ is a Gaussian process,
 \begin{equation}\label{eq:TransientDeformation}
-u_i(x,t) \sim \mathcal{N}\left(0,C_{u_i}\right),
+u_i(z) \sim \mathcal{N}\left(0,C_{u_i}\right),
 \end{equation}
-where $C_{u_i}\left((x,t),(x',t')\right)$ is a covariance function indicating how we expect $u_i(x,t)$ to covary with $u_i(x',t')$.
+where $C_{u_i}(z,z')$ is a covariance function indicating how we expect $u_i(z)$ to covary with $u_i(z')$.
 For simplicity, we treat each component of displacement independently and ignore any potential covariance.
@@
 We constrain $u$ with GNSS data, which records $u$ as well as other physical and non-physical processes which we are not interested in.
-We describe GNSS observations at position $x_i$ and time $t_j$ as a realization of the random variable
+We describe GNSS observations at position $x_i$ and time $t$ as a realization of the random variable
 \begin{align}\label{eq:Data}
-d_{ij} = &u(x_i,t_j) + \epsilon(x_i,t_j) + m_{1i} + m_{2i} t_j + \\
-         &m_{3i}\sin(2 \pi t_j) + m_{4i}\cos(2 \pi t_j) + m_{5i}\sin(4 \pi t_j) + m_{6i}\cos(4 \pi t_j), 
+\begin{split}
+d_{i}(t) = &u(x_i,t) + \epsilon(x_i,t) + a_{i} + b_{i}t + \\
+         &c_{i}\sin(2 \pi t) + q_{i}\cos(2 \pi t) + r_{i}\sin(4 \pi t) + s_{i}\cos(4 \pi t), 
+\end{split}
 \end{align}
-where $m_1i$ is an offset that is unique for each GNSS monument, $m_2i$ is the steady rate of tectonic deformation at $x_i$, the sinusoids describe seasonal deformation (using units of years for $t_j$), and $\epsilon$ is a Gaussian process for noise that cannot be described parametrically.
-$\epsilon$ can consist of white noise, temporally correlated noise describing benchmark wobble \citep[e.g.,][]{Wyatt1982,Wyatt1989}, or spatially correlated noise describing common mode error \citep[e.g.,][]{Wdowinski1997}.
+where $a_{i}$ is an offset that is unique for each GNSS monument, $b_{i}$ is the steady rate of tectonic deformation at $x_i$, the sinusoids describe seasonal deformation (using units of years for $t$), and $\epsilon$ is a Gaussian process for noise that cannot be described parametrically.
+$\epsilon$ can consist of white noise, temporally correlated noise describing, for example, benchmark wobble \citep[e.g.,][]{Wyatt1982,Wyatt1989}, and/or spatially correlated noise describing, for example, common mode error \citep[e.g.,][]{Wdowinski1997}.
 For now, we describe $\epsilon$ generally as $\epsilon \sim \mathcal{N}(0,C_\epsilon)$.
-We consider the coefficients $m_{kj}$, to be uncorrelated random variables with a diffuse prior.
-Of course, the tectonic deformation, $b_i$, is in reality spatially correlated and we could invoke tectonic model to form a prior on $b_i$.
-However, in our application to Cascadia, we will be using displacement time series which are long enough to sufficiently constrain $b_i$ for each station and there is no need to incorporate a prior.
+We consider the six coefficients in eq.
+(\ref{eq:Data}) to be uncorrelated random variables distributed as $\mathcal{N}(0,\lambda^2)$ in the limit as $\lambda \to \infty$ (i.e., the coefficients have diffuse priors).
+Of course, the tectonic deformation, $b_{i}$, is in reality spatially correlated and we could invoke tectonic model to form a prior on $b_{i}$.
+However, in our application to Cascadia, we will be using displacement time series which are long enough to sufficiently constrain $b_{i}$ for each station and there is no need to incorporate a prior.
 Likewise, the seasonal coefficients may be spatially correlated as suggested by Langbein2008?.
 It may be worth exploring and exploiting such a correlation in a future study.
+
+Let $\bm{y}$ be the set of $(x_i, t)$ pairs describing where and when each of the $n$ GNSS observations have been made.
+Furthermore, let $\bm{a}$ be the vector of coefficients from eq.
+(\ref{eq:Data}) for each of the $m$ GNSS monuments, and let $\bm{P}$ be the $n \times 6m$ matrix of corresponding basis functions evaluated at each point in $\bm{y}$.
+The $n$ GNSS observations, $\bm{d}_*$, are a realization of the random vector
+\begin{equation}
+\bm{d} = u(\bm{y}) + \epsilon(\bm{y}) + \bm{P}\bm{a},
+\end{equation}
+which is distributed as $\mathcal{N}(\bm{0},\bm{C_d})$, where
+\begin{equation}\label{eq:Cd}
+\bm{C_d} = C_u(\bm{y},\bm{y}) + C_\epsilon(\bm{y},\bm{y}) + 
+           \lambda^2\bm{P}\bm{P}^T.
+\end{equation}
+It should be understood that the notation $f(\bm{x})$ and $f(\bm{x},\bm{y})$ is used as shorthand for $[f(x_i)]_{x_i \in \bm{x}}$ and $[f(x_i,y_j)]_{(x_i,y_j) \in \bm{x} \times \bm{y}}$, respectively.
+In section X, we describe how the observations $\bm{d}_*$ are used to constrain transient displacements $u$ and then for estimates of transient strain rates.
 
@@
 \subsection{Conditioning}
-Assuming that a covariance function and corresponding hyperparameters for $X$, $T$, and $C_d$ have been chosen,
-
-In this section we describe how to condition the prior for transient displacements, $u$, with GNSS observations.
-We will assume that a appropriate covariance functions and corresponding hyperparameters for $X$, $T$, and $C_d$ have already been chosen.
-Let $\mathbf{y}$ be the set of all $(x_i, t_j)$ pairs where a GNSS observation has been made, and let $\mathbf{d} = [d_{ij}]^T_{(x_i,t_j) \in \mathbf{y}}$.
-The random vector $\mathbf{d}$ is distributed as $\mathbf{d} \sim \mathcal{N}(\bar{\mathbf{d}},\mathbf{C_d})$, where
+In this section we describe how to condition the prior for transient displacements, $u$, with GNSS observations, $\bm{d}_*$, to form a posterior estimate of transient displacements, $\hat{u} = u | \bm{d}_*$.
+We will assume that an appropriate covariance functions and corresponding hyperparameters for $X$, $T$, and $C_\epsilon$ have already been chosen.
+If $\lambda$ is kept finite, then we can follow \citet{Rasmussen2006} to find that $\hat{u}$ is distributed as $\mathcal{N}(\mu_{\hat{u}},C_{\hat{u}})$, where
+\begin{equation}\label{eq:PosteriorMean}
+\mu_{\hat{u}}(z) = C_u(z,\bm{y})\bm{C_d}^{-1}\bm{d}_*
+\end{equation}    
+and
+\begin{equation}\label{eq:PosteriorCov}
+C_{\hat{u}}(z,z') = C_u(z,\bm{y})\bm{C_d}^{-1}C_u(\bm{y},z').
+\end{equation}
+However, we are interested in the limit as $\lambda \to \infty$, and the form for eq.
+(\ref{eq:PosteriorMean}) and eq.
+(\ref{eq:PosteriorCov}) is not suitable for evaluating this limit.
+We can use the partitioned matrix inversion identity \citep[e.g.,][]{Press2007} to rewrite eq.
+(\ref{eq:PosteriorMean}) and eq.
+(\ref{eq:PosteriorCov}) as
+ \begin{equation}\label{eq:PosteriorMean2}
+\mu_{\hat{u}}(z) =
+\left[ 
+\begin{array}{cc}
+C_u(z,\bm{y}) & \bm{0} \\
+\end{array}
+\right]
+\left[
+\begin{array}{cc}
+C_u(\bm{y},\bm{y}) + C_\epsilon(\bm{y},\bm{y}) & \bm{P} \\
+\bm{P}^T  & -\lambda^{-2} \bm{I} \\
+\end{array}
+\right]^{-1}
+\left[
+\begin{array}{c}
+\bm{d}_* \\
+\bm{0} \\
+\end{array}
+\right]
+\end{equation}    
+and
+\begin{equation}\label{eq:PosteriorCov2}
+C_{\hat{u}}(z,z') = 
+C_u(z,z') - 
+\left[ 
+\begin{array}{cc}
+C_u(z,\bm{y}) & \bm{0} \\
+\end{array}
+\right]
+\left[
+\begin{array}{cc}
+C_u(\bm{y},\bm{y}) + C_\epsilon(\bm{y},\bm{y}) & \bm{P} \\
+\bm{P}^T  & -\lambda^{-2} \bm{I} \\
+\end{array}
+\right]^{-1}
+\left[
+\begin{array}{c}
+C_u(\bm{y},z') \\
+\bm{0} \\
+\end{array}
+\right].
+\end{equation}
+Taking the limit as $\lambda \to \infty$, we get the solution for the mean and covariance of $\hat{u}$,
+ \begin{equation}\label{eq:PosteriorMean3}
+\mu_{\hat{u}}(z) =
+\left[ 
+\begin{array}{cc}
+C_u(z,\bm{y}) & \bm{0} \\
+\end{array}
+\right]
+\left[
+\begin{array}{cc}
+C_u(\bm{y},\bm{y}) + C_\epsilon(\bm{y},\bm{y}) & \bm{P} \\
+\bm{P}^T  & \bm{0} \\
+\end{array}
+\right]^{-1}
+\left[
+\begin{array}{c}
+\bm{d}_* \\
+\bm{0} \\
+\end{array}
+\right]
+\end{equation}    
+and
+\begin{equation}\label{eq:PosteriorCov3}
+C_{\hat{u}}(z,z') = 
+C_u(z,z') - 
+\left[ 
+\begin{array}{cc}
+C_u(z,\bm{y}) & \bm{0} \\
+\end{array}
+\right]
+\left[
+\begin{array}{cc}
+C_u(\bm{y},\bm{y}) + C_\epsilon(\bm{y},\bm{y}) & \bm{P} \\
+\bm{P}^T  & \bm{0} \\
+\end{array}
+\right]^{-1}
+\left[
+\begin{array}{c}
+C_u(\bm{y},z') \\
+\bm{0} \\
+\end{array}
+\right].
+\end{equation}
+which allows us to evaluate the limit  evaluate the limit 
+and we consider the GNSS observations $\bm{d}_*$ to be a realization of $\bm{d}$. Furthermore, we The random vector $\bm{d}$ is distributed as $\bm{d} \sim \mathcal{N}(\bar{\bm{d}},\bm{C_d})$, where 
 \begin{equation}
-\mathbf{d}
+\bm{d}
 \end{equation}
 
-We consider the column vector of GNSS observations, $\mathbf{d}_*$, to be realizations of $d_{ij}$ at the positions and times in $\mathbf{y}$ (i.e, $\mathbf{d}_* = [d_{ij}(\omega_*)]^T_{(x_i,t_j) \in \mathbf{y}}$).
-We want to find the posterior estimate of transient displacements, $u | \mathbf{d}_*$, which incorporates our prior and the GNSS observations.
+We consider the column vector of GNSS observations, $\bm{d}_*$, to be realizations of $d_{ij}$ at the positions and times in $\bm{y}$ (i.e, $\bm{d}_* = [d_{ij}(\omega_*)]^T_{(x_i,t_j) \in \bm{y}}$).
+We want to find the posterior estimate of transient displacements, $u | \bm{d}_*$, which incorporates our prior and the GNSS observations.
 

### 2017-05-29 HinesHetlandGPR.tex
did more

@@
 %   - partition inverse and solve with CHOLMOD
-%
+% - outlier detection
+
 % APPLICATION TO CASCADIA SSE
@@
 \end{equation}
+
+The posterior transients displacements, $\hat{\bm{u}}$, are found by evaluating eq.
+(\ref{eq:PosteriorMean3}) and (\ref{eq:PosteriorCov3}) for each displacement component.
+We differentiate the Gaussian process $\hat{\bm{u}}$ (see \cite{Abrahamsen1997}) to form an estimate of transient strain rate, $\dot{\bm{\epsilon}}$, which is itself a spatially and temporally continuous Gaussian process.
+The component of the transient strain rate tensor, $\dot{\epsilon}_{ij}$, are given by
+\begin{equation}\label{eq:StrainRate}
+\dot{\epsilon}_{ij}(z) = \frac{1}{2} \frac{\partial}{\partial t} \left(
+                                     \frac{\partial \hat{u}_i(z)}{\partial x_j} +  
+                                     \frac{\partial \hat{u}_j(z)}{\partial x_i}\right).
+\end{equation}
+Component $\dot{\epsilon}_{ij}$ of the strain rate tensor is distributed as $\mathcal{N}(\mu_{\dot{\epsilon}_{ij}},C_{\dot{\epsilon}_{ij}})$, where
+\begin{equation}\label{eq:StrainMean}
+\mu_{\dot{\epsilon}_{ij}}(z) = \frac{1}{2}\frac{\partial}{\partial t}\left(
+                               \frac{\partial \mu_{\hat{u}_i}(z)}{\partial x_j} + 
+                               \frac{\partial \mu_{\hat{u}_j}(z)}{\partial x_i} \right)
+\end{equation} 
+and  
+\begin{equation}\label{eq:StrainCov}
+C_{\dot{\epsilon}_{ij}}(z,z') = \frac{1}{4} \frac{\partial^2}{\partial t \, \partial t'}\left(
+                           \frac{\partial^2 C_{\hat{u}_i}(z,z')}{\partial x_j \, \partial x'_j} + 
+                           \frac{\partial^2 C_{\hat{u}_j}(z,z')}{\partial x_i \, \partial x'_i} \right).
+\end{equation}
+
 which allows us to evaluate the limit  evaluate the limit 

### 2017-05-29 HinesHetlandGPR.tex
added more

@@
 \end{equation}
-It should be understood that the notation $f(\bm{x})$ and $f(\bm{x},\bm{y})$ is used as shorthand for $[f(x_i)]_{x_i \in \bm{x}}$ and $[f(x_i,y_j)]_{(x_i,y_j) \in \bm{x} \times \bm{y}}$, respectively.
+It should be understood that the notation $f(\bm{x})$ and $f(\bm{x},\bm{y})$ is shorthand for $[f(x_i)]_{x_i \in \bm{x}}$ and $[f(x_i,y_j)]_{(x_i,y_j) \in \bm{x} \times \bm{y}}$, respectively.
 In section X, we describe how the observations $\bm{d}_*$ are used to constrain transient displacements $u$ and then for estimates of transient strain rates.
 
-\subsection{Spatial and temporal covariance functions}
-We must choose an appropriate Gaussian process describing transient displacements $u$ and GNSS noise $\epsilon$.
-We assume that $u$ and $\epsilon$ both have zero mean, which leaves for use to select a parameterization for the covariance functions $X$, $T$, and $C_d$.
-In this section, we list the covariance functions which are considered in this paper and those which could be of use in future studies.
-In the following discussion $\alpha$ and $\beta$ are hyperparameters.
-There are numerous strategies for selecting these hyperparameters (see \citet{Cressie1992}) and we save a discussion on choosing the hyperparameters for Section X.
-
-The squared exponential covariance function,
-\begin{equation}\label{eq:SE}
-C(x,x') = \alpha^2 \exp\left(\frac{-||x - x'||_2^2}{2 \beta^2}\right),
-\end{equation}
-is commonly used in Kriging \citep[e.g,][]{Cressie1992} and Gaussian process regression \citep[e.g.,][]{Rasmussen2006}.
-The squared exponential is a valid (i.e.
-positive definite) covariance function for any number of spatial dimensions, and it describes an isotropic (i.e.
-$C(x,x')$ can be expressed as $\hat{C}(||x - x'||_2)$) Gaussian processes with realizations that are infinitely differentiable.
-\citet{Kato1998} and \cite{El-Fiky1999} demonstrated that eq.
-(\ref{eq:SE}) is an appropriate covariance model for describing long-term tectonic strain rates for Japan.
-
-The Wendland class of covariance functions \citep{Wendland2005} are positive definite in $\mathbb{R}^d$, and they describes an isotropic Gaussian process with realizations that can be differentiated $k$ times.
-The form of the covariance function depends on the choice of $d$ and $k$.
-Wendland covariance functions have compact support, which is a particularly useful property because we can exploit the sparsity of the the corresponding covariance matrices.
-In this study we use Wendland covariance functions to describe the temporal component of transient displacements.
-Therefore, we only require that $d=1$ and $k=1$.
-The corresponding Wendland covariance function is
-\begin{equation}\label{eq:Wendland}
-C(x,x') = \alpha^2\left(1 - \frac{|x - x'|}{\beta}\right)^3_+ \left(\frac{3|x - x'|}{\beta} + 1\right), \ \ \ x,x' \in \mathbb{R}^1,
-\end{equation}
-where
-\begin{equation}
-(x)_+ = 
-\begin{cases}
-x, \ \ \ x > 0 \\
-0, \ \ \ \mathrm{otherwise}.
-\end{cases}
-\end{equation}
-
-We next consider the covariance function which describes a first-order Gauss-Markov (FOGM) process. A FOGM process is a solution to the stochastic differential equation
-\begin{equation}
-\dot{u}(x) + \alpha u(x) = \beta w(x), \ \ \ x,x' \in \mathbb{R}^1,
-\end{equation}
-where $w(x)$ is white noise with unit variance.  The solution is a Gaussian process with zero mean and covariance
-% This needs to be scaled to match my work
-\begin{equation}\label{eq:FOGM}
-C(x,x') = \frac{\beta^2}{2\alpha}\exp\left(-\alpha|x - x'|\right), \ \ \ x,x' \in \mathbb{R}^1. 
-\end{equation}
-FOGM processes are isotropic and realizations are not differentiable.
-The FOGM model has been used to describe temporally correlated noise in EDM data \citep{Langbein2004} and GNSS data \citep[e.g.,][]{Langbein2008}.
-
-One last covariance function which is worth mentioning, even though we do not use it in this paper, is the periodic covariance function \citep{Mackay1998}, 
-\begin{equation}\label{eq:Periodic}
-C(x,x') = \alpha^2 \exp\left(\frac{-\sin(\frac{\pi}{\lambda} |x - x'|)^2}{2 \beta^2} \right), \ \ \ x,x' \in \mathbb{R}^1.
-\end{equation}
-The periodic covariance function describes an isotropic Gaussian process with realizations that have period $\lambda$.
-The length-scale of features within each period is controlled by $\beta$.
-In terms of GNSS data, a periodic Gaussian process can be used to describe seasonal deformation.
-The more common method of treating seasonal deformation is to model it as a linear combination of annual and semi-annual sinusoids \citep[e.g.,][]{Dong2002}.
-We believe there is potential to use periodic Gaussian processes to explore the statistical properties of seasonal signals; however, that is outside the scope of this paper.
-
-Brownian motion and integrated Brownian motion are two Gaussian processes which are frequently used in the geodetic literature.  Brownian motion, which has zero mean and covariance function
-\begin{equation}\label{eq:BM}
-C(x,x') =\alpha^2 \min(x,x'), \ \ \ x,x' \in \mathbb{R}^1 \ \mathrm{and} \ \ x,x' \geq 0, 
-\end{equation} 
-is commonly used to describe noise resulting from the instability of geodetic monuments \citep[e.g.,][]{Wyatt1982,Wyatt1989,Langbein1997}.
-Integrated Brownian motion, which has zero mean and covariance function
-\begin{align}\label{eq:IBM}
-C(x,x') &= \int_0^x \int_0^{x'} \alpha^2 \min(u,u') \,du'\,du \\
-        &= \frac{\alpha^2}{2}\min(x,x')^2 \left(\max(x,x') - \frac{1}{3}\min(x,x')\right), \ \ \ x,x' \in \mathbb{R}^1 \ \mathrm{and} \ \ x,x' \geq 0,
-\end{align}
-has been used as a nonparametric model for the time dependence of geophysical signals \citep[e.g.,][]{Segall1997,McGuire2003,Ohtani2010,Hines2016}.
-It should be noted that Brownian motion and integrated Brownian motion have a reference time at which the Gaussian process is exactly zero.
-For some geophysical signals, such as postseismic deformation, it is reasonable to have such a reference time.
-On the other hand, if there is no sense of a start time for the geophysical process, then Brownian motion or integrated Brownian motion may not be appropriate, and one of the isoptropic Gaussian processes described above should be used.
-
-\subsection{Conditioning}
-In this section we describe how to condition the prior for transient displacements, $u$, with GNSS observations, $\bm{d}_*$, to form a posterior estimate of transient displacements, $\hat{u} = u | \bm{d}_*$.
+We now describe how to condition the prior for transient displacements, $u$, with GNSS observations, $\bm{d}_*$, to form a posterior estimate of transient displacements, $\hat{u} = u | \bm{d}_*$.
 We will assume that an appropriate covariance functions and corresponding hyperparameters for $X$, $T$, and $C_\epsilon$ have already been chosen.
@@
 
-which allows us to evaluate the limit  evaluate the limit 
-and we consider the GNSS observations $\bm{d}_*$ to be a realization of $\bm{d}$. Furthermore, we The random vector $\bm{d}$ is distributed as $\bm{d} \sim \mathcal{N}(\bar{\bm{d}},\bm{C_d})$, where 
+\subsection{Spatial and temporal covariance functions}
+We must choose an appropriate Gaussian process describing transient displacements $u$ and GNSS noise $\epsilon$.
+We assume that $u$ and $\epsilon$ both have zero mean, which leaves for use to select a parameterization for the covariance functions $X$, $T$, and $C_d$.
+In this section, we list the covariance functions which are considered in this paper and those which could be of use in future studies.
+In the following discussion $\alpha$ and $\beta$ are hyperparameters.
+There are numerous strategies for selecting these hyperparameters (see \citet{Cressie1992}) and we save a discussion on choosing the hyperparameters for Section X.
+
+The squared exponential covariance function,
+\begin{equation}\label{eq:SE}
+C(x,x') = \alpha^2 \exp\left(\frac{-||x - x'||_2^2}{2 \beta^2}\right),
+\end{equation}
+is commonly used in Kriging \citep[e.g,][]{Cressie1992} and Gaussian process regression \citep[e.g.,][]{Rasmussen2006}.
+The squared exponential is a valid (i.e.
+positive definite) covariance function for any number of spatial dimensions, and it describes an isotropic (i.e.
+$C(x,x')$ can be expressed as $\hat{C}(||x - x'||_2)$) Gaussian processes with realizations that are infinitely differentiable.
+\citet{Kato1998} and \cite{El-Fiky1999} demonstrated that eq.
+(\ref{eq:SE}) is an appropriate covariance model for describing long-term tectonic strain rates for Japan.
+
+The Wendland class of covariance functions \citep{Wendland2005} are positive definite in $\mathbb{R}^d$, and they describes an isotropic Gaussian process with realizations that can be differentiated $k$ times.
+The form of the covariance function depends on the choice of $d$ and $k$.
+Wendland covariance functions have compact support, which is a particularly useful property because we can exploit the sparsity of the the corresponding covariance matrices.
+In this study we use Wendland covariance functions to describe the temporal component of transient displacements.
+Therefore, we only require that $d=1$ and $k=1$.
+The corresponding Wendland covariance function is
+\begin{equation}\label{eq:Wendland}
+C(x,x') = \alpha^2\left(1 - \frac{|x - x'|}{\beta}\right)^3_+ \left(\frac{3|x - x'|}{\beta} + 1\right), \ \ \ x,x' \in \mathbb{R}^1,
+\end{equation}
+where
 \begin{equation}
-\bm{d}
+(x)_+ = 
+\begin{cases}
+x, \ \ \ x > 0 \\
+0, \ \ \ \mathrm{otherwise}.
+\end{cases}
 \end{equation}
 
-We consider the column vector of GNSS observations, $\bm{d}_*$, to be realizations of $d_{ij}$ at the positions and times in $\bm{y}$ (i.e, $\bm{d}_* = [d_{ij}(\omega_*)]^T_{(x_i,t_j) \in \bm{y}}$).
-We want to find the posterior estimate of transient displacements, $u | \bm{d}_*$, which incorporates our prior and the GNSS observations.
+We next consider the covariance function which describes a first-order Gauss-Markov (FOGM) process. A FOGM process is a solution to the stochastic differential equation
+\begin{equation}
+\dot{u}(x) + \alpha u(x) = \beta w(x), \ \ \ x,x' \in \mathbb{R}^1,
+\end{equation}
+where $w(x)$ is white noise with unit variance.  The solution is a Gaussian process with zero mean and covariance
+% This needs to be scaled to match my work
+\begin{equation}\label{eq:FOGM}
+C(x,x') = \frac{\beta^2}{2\alpha}\exp\left(-\alpha|x - x'|\right), \ \ \ x,x' \in \mathbb{R}^1. 
+\end{equation}
+FOGM processes are isotropic and realizations are not differentiable.
+The FOGM model has been used to describe temporally correlated noise in EDM data \citep{Langbein2004} and GNSS data \citep[e.g.,][]{Langbein2008}.
 
-Once we have chosen covariance functions and corresponding hyperparameters for $X$, $T$, and $C_d$, we then condition our prior Gaussian process, $u$, with the observations $d_*$ to form a posterior estimate of transient displacements, $u|d_*$.
-$d_*$ is a realization of the
+One last covariance function which is worth mentioning, even though we do not use it in this paper, is the periodic covariance function \citep{Mackay1998}, 
+\begin{equation}\label{eq:Periodic}
+C(x,x') = \alpha^2 \exp\left(\frac{-\sin(\frac{\pi}{\lambda} |x - x'|)^2}{2 \beta^2} \right), \ \ \ x,x' \in \mathbb{R}^1.
+\end{equation}
+The periodic covariance function describes an isotropic Gaussian process with realizations that have period $\lambda$.
+The length-scale of features within each period is controlled by $\beta$.
+In terms of GNSS data, a periodic Gaussian process can be used to describe seasonal deformation.
+The more common method of treating seasonal deformation is to model it as a linear combination of annual and semi-annual sinusoids \citep[e.g.,][]{Dong2002}.
+We believe there is potential to use periodic Gaussian processes to explore the statistical properties of seasonal signals; however, that is outside the scope of this paper.
 
-Whichever the Gaussian processes describing transient displacements, $u$, and noise in the GNSS data, $\epsilon$, have covariance functions which
-
-Optimal Hyperparameters
-
-Conditioning the GP
-
-Differentiation, scaling and addition to get strain
+Brownian motion and integrated Brownian motion are two Gaussian processes which are frequently used in the geodetic literature.  Brownian motion, which has zero mean and covariance function
+\begin{equation}\label{eq:BM}
+C(x,x') =\alpha^2 \min(x,x'), \ \ \ x,x' \in \mathbb{R}^1 \ \mathrm{and} \ \ x,x' \geq 0, 
+\end{equation} 
+is commonly used to describe noise resulting from the instability of geodetic monuments \citep[e.g.,][]{Wyatt1982,Wyatt1989,Langbein1997}.
+Integrated Brownian motion, which has zero mean and covariance function
+\begin{align}\label{eq:IBM}
+C(x,x') &= \int_0^x \int_0^{x'} \alpha^2 \min(u,u') \,du'\,du \\
+        &= \frac{\alpha^2}{2}\min(x,x')^2 \left(\max(x,x') - \frac{1}{3}\min(x,x')\right), \ \ \ x,x' \in \mathbb{R}^1 \ \mathrm{and} \ \ x,x' \geq 0,
+\end{align}
+has been used as a nonparametric model for the time dependence of geophysical signals \citep[e.g.,][]{Segall1997,McGuire2003,Ohtani2010,Hines2016}.
+It should be noted that Brownian motion and integrated Brownian motion have a reference time at which the Gaussian process is exactly zero.
+For some geophysical signals, such as postseismic deformation, it is reasonable to have such a reference time.
+On the other hand, if there is no sense of a start time for the geophysical process, then Brownian motion or integrated Brownian motion may not be appropriate, and one of the isoptropic Gaussian processes described above should be used.
 

### 2017-05-30 HinesHetlandGPR.tex
added more

@@
 \section{Method}\label{sec:Method}
-
 In this section, we describe how GNSS displacement observations are used to identify transient crustal strain rates, which we denote as $\dot{\bm{\varepsilon}}(z)$, where $z$ is the ordered pair $(x,t)$, $x$ are spatial coordinates in $\mathbb{R}^2$, and $t$ is time.
@@
 where $a_{i}$ is an offset that is unique for each GNSS monument, $b_{i}$ is the steady rate of tectonic deformation at $x_i$, the sinusoids describe seasonal deformation (using units of years for $t$), and $\epsilon$ is a Gaussian process for noise that cannot be described parametrically.
-$\epsilon$ can consist of white noise, temporally correlated noise describing, for example, benchmark wobble \citep[e.g.,][]{Wyatt1982,Wyatt1989}, and/or spatially correlated noise describing, for example, common mode error \citep[e.g.,][]{Wdowinski1997}.
-For now, we describe $\epsilon$ generally as $\epsilon \sim \mathcal{N}(0,C_\epsilon)$.
+For example, $\epsilon$ can consist of white noise, temporally correlated noise describing benchmark wobble \citep[e.g.,][]{Wyatt1982,Wyatt1989}, and/or spatially correlated noise describing common mode error \citep[e.g.,][]{Wdowinski1997}.
+For now, $\epsilon$ will simply be described as $\epsilon \sim \mathcal{N}(0,C_\epsilon)$.
 We consider the six coefficients in eq.
 (\ref{eq:Data}) to be uncorrelated random variables distributed as $\mathcal{N}(0,\lambda^2)$ in the limit as $\lambda \to \infty$ (i.e., the coefficients have diffuse priors).
-Of course, the tectonic deformation, $b_{i}$, is in reality spatially correlated and we could invoke tectonic model to form a prior on $b_{i}$.
-However, in our application to Cascadia, we will be using displacement time series which are long enough to sufficiently constrain $b_{i}$ for each station and there is no need to incorporate a prior.
-Likewise, the seasonal coefficients may be spatially correlated as suggested by Langbein2008?.
-It may be worth exploring and exploiting such a correlation in a future study.
+Of course, the tectonic deformation, $b_{i}$, is spatially correlated and we could invoke a tectonic model to form a prior on $b_{i}$.
+However, in our application to Cascadia, we will be using displacement time series which are long enough to sufficiently constrain $b_{i}$ for each station, and there is no need to incorporate a prior.
+Likewise, the seasonal coefficients may be spatially correlated \citep{Langbein2008}, and it may be worth exploring and exploiting such a correlation in a future study.
 
-Let $\bm{y}$ be the set of $(x_i, t)$ pairs describing where and when each of the $n$ GNSS observations have been made.
+We now consider the column vector of $n$ GNSS observations, $\bm{d}_*$.
+Let $\bm{y}$ be the set of $(x_i, t)$ pairs describing where and when each of the GNSS observations have been made.
 Furthermore, let $\bm{a}$ be the vector of coefficients from eq.
 (\ref{eq:Data}) for each of the $m$ GNSS monuments, and let $\bm{P}$ be the $n \times 6m$ matrix of corresponding basis functions evaluated at each point in $\bm{y}$.
-The $n$ GNSS observations, $\bm{d}_*$, are a realization of the random vector
+The observations $\bm{d}_*$ are then a realization of the random vector
 \begin{equation}
@@
 It should be understood that the notation $f(\bm{x})$ and $f(\bm{x},\bm{y})$ is shorthand for $[f(x_i)]_{x_i \in \bm{x}}$ and $[f(x_i,y_j)]_{(x_i,y_j) \in \bm{x} \times \bm{y}}$, respectively.
-In section X, we describe how the observations $\bm{d}_*$ are used to constrain transient displacements $u$ and then for estimates of transient strain rates.
 
-We now describe how to condition the prior for transient displacements, $u$, with GNSS observations, $\bm{d}_*$, to form a posterior estimate of transient displacements, $\hat{u} = u | \bm{d}_*$.
+We now describe how to condition the prior for each component of transient displacements, $u$, with $\bm{d}_*$ to form a posterior estimate of transient displacements, $\hat{u} = u | \bm{d}_*$.
 We will assume that an appropriate covariance functions and corresponding hyperparameters for $X$, $T$, and $C_\epsilon$ have already been chosen.
@@
 We differentiate the Gaussian process $\hat{\bm{u}}$ (see \cite{Abrahamsen1997}) to form an estimate of transient strain rate, $\dot{\bm{\epsilon}}$, which is itself a spatially and temporally continuous Gaussian process.
-The component of the transient strain rate tensor, $\dot{\epsilon}_{ij}$, are given by
+The components of the transient strain rate tensor, $\dot{\epsilon}_{ij}$, are given by
 \begin{equation}\label{eq:StrainRate}
@@
 \end{equation}
+
+\subsection{Outlier detection}
+In our formulation for estimating transient strain rates, we have assumed that noise in the data vector is normally distributed.
+This is not an appropriate assumption for GNSS displacement time series which are prone to more outliers than would be predicted for normally distributed noise.
+It follows that proposed methods for analyzing GNSS timeseries should be robust against outliers \citep[e.g.,][]{Blewitt2016}.
+In order to make our estimates of transient strain more robust, we automatically identify and remove outliers in the GNSS data as a pre-processing step.
+Our strategy for identifying outliers is based on the data editing algorithms described in \citet{Gibbs2011}.
+In essence, outliers are considered to be observations which are abnormally inconsistent with $\bm{d}$.
+This algorithm is particularly effective because $\bm{d}$ has a spatial and temporal dependence, which allows us to identify anomalous observations which may not be apparent by inspection of timeseries alone.
+
+We now describe the data editing algorithm. Let $\Omega$ be the index set identifying non-outliers in $\bm{d}_*$,  which initially contains all $n$ indices.
+
+We again treat $\bm{d}_*$ as a sample of $\bm{d}$ and we assume some prior covariance function for $u$.
+We let $\epsilon$ be uncorrelated noise which has standard deviation given by the formal data uncertainties, $\bm{\sigma}$.
+Outliers are considered to be the observations which are abnormally inconsistent with the Gaussian process $u(\bm{y}) + \bm{P}\bm{a}$.
+We identify outliers in an iterative algorithm that is based on the data editing algorithm described in \citet{Gibbs2011}, which we now describe.
+
+Let $\Omega$ be the index set of non-outliers in $\bm{d}_*$, which we initiate with all $n$ indices. For each iteration we calculate the residual vector
+\begin{align}
+\bm{r} &= \frac{\bm{d}_* - \mathrm{E}\left[(u(\bm{y}) + \bm{P}\bm{a})|\tilde{\bm{d}}_* \right]}{\bm{\sigma}} \\
+       &= \frac{1}{\bm{\sigma}}\left(\bm{d}_*  - 
+\left[ 
+\begin{array}{cc}
+C_u(\bm{y},\tilde{\bm{y}}) & \bm{P} \\
+\end{array}
+\right]
+\left[
+\begin{array}{cc}
+C_u(\tilde{\bm{y}},\tilde{\bm{y}}) + \mathrm{diag}(\tilde{\bm{\sigma}}^2) & \tilde{\bm{P}} \\
+\tilde{\bm{P}}^T  & \bm{0} \\
+\end{array}
+\right]^{-1}
+\left[
+\begin{array}{c}
+\tilde{\bm{d}}_* \\
+\bm{0} \\
+\end{array}
+\right] 
+\right).
+\end{align}
+
+as an approximate realization of $\epsilon$.
+
+in addition to spatially and/or temporally correlated noise.
+
+  be Furthermore, we consider $\epsilon$ to be composed of uncorrelated noise with standard deviations given by the formal uncertainties on $\bm{d}_*$, $\bm{\sigma}$, as well as  as well as s composed of
 

### 2017-05-31 HinesHetlandGPR.tex
added more

@@
 \end{equation}
-Component $\dot{\epsilon}_{ij}$ of the strain rate tensor is distributed as $\mathcal{N}(\mu_{\dot{\epsilon}_{ij}},C_{\dot{\epsilon}_{ij}})$, where
+which are distributed as $\mathcal{N}(\mu_{\dot{\epsilon}_{ij}},C_{\dot{\epsilon}_{ij}})$, where
 \begin{equation}\label{eq:StrainMean}
@@
 In our formulation for estimating transient strain rates, we have assumed that noise in the data vector is normally distributed.
-This is not an appropriate assumption for GNSS displacement time series which are prone to more outliers than would be predicted for normally distributed noise.
+This is not an appropriate assumption for GNSS displacement timeseries which are prone to more outliers than would be predicted for normally distributed noise.
 It follows that proposed methods for analyzing GNSS timeseries should be robust against outliers \citep[e.g.,][]{Blewitt2016}.
 In order to make our estimates of transient strain more robust, we automatically identify and remove outliers in the GNSS data as a pre-processing step.
-Our strategy for identifying outliers is based on the data editing algorithms described in \citet{Gibbs2011}.
-In essence, outliers are considered to be observations which are abnormally inconsistent with $\bm{d}$.
-This algorithm is particularly effective because $\bm{d}$ has a spatial and temporal dependence, which allows us to identify anomalous observations which may not be apparent by inspection of timeseries alone.
 
-We now describe the data editing algorithm. Let $\Omega$ be the index set identifying non-outliers in $\bm{d}_*$,  which initially contains all $n$ indices.
-
-We again treat $\bm{d}_*$ as a sample of $\bm{d}$ and we assume some prior covariance function for $u$.
+Our method for detecting outliers is based on the data editing algorithm described in \citet{Gibbs2011}.
+We again treat $\bm{d}_*$ as a sample of $\bm{d}$, and we assume some prior covariance function for $u$.
+The choice for $u$ here does not need to be the same as that used in Section X, and a conservative approach would be to choose a $u$ that is more flexible (e.g., shorter characteristic length-scales and higher variance).
 We let $\epsilon$ be uncorrelated noise which has standard deviation given by the formal data uncertainties, $\bm{\sigma}$.
-Outliers are considered to be the observations which are abnormally inconsistent with the Gaussian process $u(\bm{y}) + \bm{P}\bm{a}$.
-We identify outliers in an iterative algorithm that is based on the data editing algorithm described in \citet{Gibbs2011}, which we now describe.
-
-Let $\Omega$ be the index set of non-outliers in $\bm{d}_*$, which we initiate with all $n$ indices. For each iteration we calculate the residual vector
-\begin{align}
+We consider outliers to be observations which are abnormally inconsistent with the random vector $u(\bm{y}) + \bm{P}\bm{a}$.
+With this criterion, we takes into account the spatial and temporal relationship between the data, allowing us to identify anomalous observations which may not be immediately apparent based on inspection of individual timeseries.
+To begin the algorithm, we let $\Omega$ be the index set of non-outliers in $\bm{d}_*$ and initiate it with all $n$ indices.
+This algorithm is iterative, and for each iteration we calculate the residual vector
+\begin{align}\label{eq:Residual}
 \bm{r} &= \frac{\bm{d}_* - \mathrm{E}\left[(u(\bm{y}) + \bm{P}\bm{a})|\tilde{\bm{d}}_* \right]}{\bm{\sigma}} \\
@@
 \right] 
-\right).
+\right),
 \end{align}
+where the tilde indicates that only elements corresponding to indices in $\Omega$ are retained (e.g., $\tilde{\bm{y}} = (y_i)_{i\in\Omega}$). We then update $\Omega$ to be
+\begin{equation}\label{eq:Update}
+\Omega = \{i : |r_i| < \eta \tilde{\bm{r}}_{rms}\}
+\end{equation} 
+where $\tilde{\bm{r}}_{rms}$ is the root-mean-square of $\tilde{\bm{r}}$ and $\lambda$ is an outlier tolerance.
+We use $\lambda=4$ in this study, which seems to accurately identify outliers without unnecessarily decimating the data.
+Iterations continue until the new $\Omega$ is equal to the previous $\Omega$ (i.e., no new outliers are detected).
 
-as an approximate realization of $\epsilon$.
-
-in addition to spatially and/or temporally correlated noise.
-
-  be Furthermore, we consider $\epsilon$ to be composed of uncorrelated noise with standard deviations given by the formal uncertainties on $\bm{d}_*$, $\bm{\sigma}$, as well as  as well as s composed of
+It should be noted that this algorithm does not identify jumps in GNSS time series, which are another common defect.
+Some, but not all, jumps can be automatically removed by parsing station logs and identifying dates of antenna changes.
+However, it is still necessary to manually inspect and remove jumps of unknown origin.
+That being said, this outlier detection algorithm significantly reduces the effort needed to manually clean GNSS data.
 
@@
 
-\subsection{Time Depenent Strain Rate in Cascadia}\label{sec:ApplicationsCascadia}
+\section{Application Cascadia Slow Slip Events}\label{sec:Cascadia}
+
+We demonstrate the method for estimating transient strain described in the previous section by applying it to GNSS data in Cascadia.
+We are mainly interested in identifying transient strain resulting from slow slip events (SSEs) \citep[e.g.,][]{Dragert2001}.
+We show that Gaussian processes are an effective tool for (1) detecting SSEs and (2) understanding how the strain from SSEs are contributing to the strain accumulated over the seismic cycle.
+Our primary motivation for detecting SSEs is because there have been several documented cases of SSEs triggering megathrust earthquakes \citep{Roeloffs2006}.
+In Cascadia, there is less of a need to detect SSEs with GNSS data because the SSEs are associated with a spike in seismic tremor, which is actively monitored by the Pacific Northwest Seismic Network (PNSN) \citep{Wech2010}.
+However, SSEs have been observed in other subduction zones without any associated tremor \citep{Schwartz2007}, and so there is still utility in developing method for detecting SSEs with GNSS data.
+Furthermore, by comparing our estimated transient strain in Cascadia with the tremor catalog, we can verify the accuracy of method for detecting SSEs.
+
+\subsection{Noise model}
+In this section we discuss our choice for the noise covariance function $C_\epsilon$.
+We will assume that the noise consists of uncorrelated noise with standard deviation given by the formal data uncertainties and potentially spatially and/or temporally correlated noise.
+There have been numerous studies on temporally correlated noise in GNSS data \citep[e.g.,]{Zhang1997,Mao1999,Williams2004,Langbein2008}.
+These studies have found some combination of Brownian motion or flicker noise best describes the temporally correlated noise in GNSS data.
+Brownian motion has also been used to describe error resulting from an unstable geodetic monument \citep[e.g.,][]{Wyatt1982,Wyatt1989}.
+We assume that temporally correlated noise can be described by a FOGM stochastic model, noting that the Brownian motion is a degenerate case of a FOGM process.
+
+To determine the appropriate hyperparameters for the FOGM model, $\alpha$ and $\beta$, we first select 38 stations in in Washington and Oregon which are east of $121^\circ$ W, where we can reasonably expect no transient signal from SSEs (i.e.
+$u = 0$).
+We then find $\alpha$ and $\beta$ for each station timeseries with the Restricted Maximum Likelihood (REML) method \cite[e.g.,][]{Harville1974,Cressie1992} and Hines 2017.
+The REML method finds the hyperparameters which maximize the probability of drawing $\bm{d}_*$ from $\bm{d}$.
+We use the REML method over the maximum likelihood (ML) method \citep[e.g.,][]{Langbein1997} because the REML method properly accounts for the improper prior that we assigned to $\bm{a}$.
+The distribution of inferred $\alpha$ and $\beta$ are shown in Figure X.
+The amplitude of FOGM noise for the east and north components is notable low and are unimodally clustered around 0.5 mm/yr$^{0.5}$.
+The corresponding estimates of $\alpha$ tend to cluster around 0 yr$^{-1}$, which is the degenerate case where FOGM behaves likes Brownian motion.
+The amplitude of FOGM noise for the vertical component is significantly larger with a median value of 13.8 mm/yr$^{0.5}$.
+The inferred values for alpha are higher for the vertical component with a median value of 1.87 yr$^{-1}$.
+In Figure X, we use the median values of $\alpha$ and $\beta$ to generate two random samples of FOGM noise.
+The samples span seven years and over these seven years the east and north samples drift by about 1 mm.
+In the context of detecting SSEs, which produce several mm's of surface displacement on the timescale of weeks to months, the estimated FOGM noise for the east and north component is negligible.
+In contrast, the estimated FOGM noise for the vertical component is significantly larger than the signal we would expect from SSEs.
+We suspect that the higher amplitude for the FOGM model in the vertical component is accommodating for deficiencies in our rather simple seasonal model.
+Based on this analysis, we henceforth ignore temporally correlated noise in the east and north component because of its low amplitude, and we do not estimate $u$ for vertical deformation because of the low signal to noise ratio.
+
+Another significant source of noise in GNSS data is common mode error \citep[e.g.,][]{Wdowinski1997}, which is noise that is highly spatially correlated.
+When not accounted for, common mode error manifests itself as spatially uniform undulations in $\hat{u}$.
+However, we are primarily interested in estimating strain which is insensitive to common mode error.
+We therefore do not include common mode error in our noise model.
+
+Our noise is just white.
+
+\subsection{Transient displacement model}
+
+model exhibits the same relationship  which can described
+
+Both of which have clear implications for assessing seismic hazard.
+
+, which can have applications for seismic hazard assessment,Identifying strain resulting from ETS events serves two purposes: (1) ETS events have  events which are  were first discovered by \citet{Dragert2001}.
+
+We demonstrate that Gaussian processes are effective tools for detecting transient strain by using them to detect Cascadia slow slip events.
 
@@
 
+Explain strain concentration with this:
 Slow slip event depths from \citet{Dragert2001}, \citet{Wech2009}, \citet{Schmidt2010}, \citet{Bartlow2011} show slip concentrated at depths from 30 to 50 km detph.
@@
 Studies which simultaneously modeled interseismic and ets are \citet{Holtkamp2010} and \citet{schmalzle2014}
-
-METHOD
-
-We use GPS displacement time series to make daily estimates of strain rates from 2010-01-01 to 2016-07-01 in the Pacific Northwest.
-Our method consists of two distinct steps; we temporally smooth and differentiate the displacement time series for each station, and then we spatially smooth and differentiate the velocity field for each day.
-We temporally smooth the GPS displacements with the method described in \ref{sec:Filter}, and we treat days with missing data as observations with infinite uncertainty.
-As noted above, this is effectively equivalent to fitting a smoothing spline to each time series, which can also be efficiently done with a Kalman filtering strategy \citep{Kohn1987}.
-We chose a cutoff frequency of ZZZ, which is high enough to acurately describe deformation during a slow slip event.
-Spatial smoothing is done with the RBF-FD filter and differentiation to get the strain rates was done with the RBF-FD scheme.
-We picked a cutoff frequency of XXX for smoothing.
 

### 2017-06-01 HinesHetlandGPR.tex
added more crap

@@
 
-\subsection{Spatial and temporal covariance functions}
-We must choose an appropriate Gaussian process describing transient displacements $u$ and GNSS noise $\epsilon$.
-We assume that $u$ and $\epsilon$ both have zero mean, which leaves for use to select a parameterization for the covariance functions $X$, $T$, and $C_d$.
-In this section, we list the covariance functions which are considered in this paper and those which could be of use in future studies.
-In the following discussion $\alpha$ and $\beta$ are hyperparameters.
-There are numerous strategies for selecting these hyperparameters (see \citet{Cressie1992}) and we save a discussion on choosing the hyperparameters for Section X.
-
-The squared exponential covariance function,
-\begin{equation}\label{eq:SE}
-C(x,x') = \alpha^2 \exp\left(\frac{-||x - x'||_2^2}{2 \beta^2}\right),
-\end{equation}
-is commonly used in Kriging \citep[e.g,][]{Cressie1992} and Gaussian process regression \citep[e.g.,][]{Rasmussen2006}.
-The squared exponential is a valid (i.e.
-positive definite) covariance function for any number of spatial dimensions, and it describes an isotropic (i.e.
-$C(x,x')$ can be expressed as $\hat{C}(||x - x'||_2)$) Gaussian processes with realizations that are infinitely differentiable.
-\citet{Kato1998} and \cite{El-Fiky1999} demonstrated that eq.
-(\ref{eq:SE}) is an appropriate covariance model for describing long-term tectonic strain rates for Japan.
-
-The Wendland class of covariance functions \citep{Wendland2005} are positive definite in $\mathbb{R}^d$, and they describes an isotropic Gaussian process with realizations that can be differentiated $k$ times.
-The form of the covariance function depends on the choice of $d$ and $k$.
-Wendland covariance functions have compact support, which is a particularly useful property because we can exploit the sparsity of the the corresponding covariance matrices.
-In this study we use Wendland covariance functions to describe the temporal component of transient displacements.
-Therefore, we only require that $d=1$ and $k=1$.
-The corresponding Wendland covariance function is
-\begin{equation}\label{eq:Wendland}
-C(x,x') = \alpha^2\left(1 - \frac{|x - x'|}{\beta}\right)^3_+ \left(\frac{3|x - x'|}{\beta} + 1\right), \ \ \ x,x' \in \mathbb{R}^1,
-\end{equation}
-where
-\begin{equation}
-(x)_+ = 
-\begin{cases}
-x, \ \ \ x > 0 \\
-0, \ \ \ \mathrm{otherwise}.
-\end{cases}
-\end{equation}
-
-We next consider the covariance function which describes a first-order Gauss-Markov (FOGM) process. A FOGM process is a solution to the stochastic differential equation
-\begin{equation}
-\dot{u}(x) + \alpha u(x) = \beta w(x), \ \ \ x,x' \in \mathbb{R}^1,
-\end{equation}
-where $w(x)$ is white noise with unit variance.  The solution is a Gaussian process with zero mean and covariance
-% This needs to be scaled to match my work
-\begin{equation}\label{eq:FOGM}
-C(x,x') = \frac{\beta^2}{2\alpha}\exp\left(-\alpha|x - x'|\right), \ \ \ x,x' \in \mathbb{R}^1. 
-\end{equation}
-FOGM processes are isotropic and realizations are not differentiable.
-The FOGM model has been used to describe temporally correlated noise in EDM data \citep{Langbein2004} and GNSS data \citep[e.g.,][]{Langbein2008}.
-
-One last covariance function which is worth mentioning, even though we do not use it in this paper, is the periodic covariance function \citep{Mackay1998}, 
-\begin{equation}\label{eq:Periodic}
-C(x,x') = \alpha^2 \exp\left(\frac{-\sin(\frac{\pi}{\lambda} |x - x'|)^2}{2 \beta^2} \right), \ \ \ x,x' \in \mathbb{R}^1.
-\end{equation}
-The periodic covariance function describes an isotropic Gaussian process with realizations that have period $\lambda$.
-The length-scale of features within each period is controlled by $\beta$.
-In terms of GNSS data, a periodic Gaussian process can be used to describe seasonal deformation.
-The more common method of treating seasonal deformation is to model it as a linear combination of annual and semi-annual sinusoids \citep[e.g.,][]{Dong2002}.
-We believe there is potential to use periodic Gaussian processes to explore the statistical properties of seasonal signals; however, that is outside the scope of this paper.
-
-Brownian motion and integrated Brownian motion are two Gaussian processes which are frequently used in the geodetic literature.  Brownian motion, which has zero mean and covariance function
-\begin{equation}\label{eq:BM}
-C(x,x') =\alpha^2 \min(x,x'), \ \ \ x,x' \in \mathbb{R}^1 \ \mathrm{and} \ \ x,x' \geq 0, 
-\end{equation} 
-is commonly used to describe noise resulting from the instability of geodetic monuments \citep[e.g.,][]{Wyatt1982,Wyatt1989,Langbein1997}.
-Integrated Brownian motion, which has zero mean and covariance function
-\begin{align}\label{eq:IBM}
-C(x,x') &= \int_0^x \int_0^{x'} \alpha^2 \min(u,u') \,du'\,du \\
-        &= \frac{\alpha^2}{2}\min(x,x')^2 \left(\max(x,x') - \frac{1}{3}\min(x,x')\right), \ \ \ x,x' \in \mathbb{R}^1 \ \mathrm{and} \ \ x,x' \geq 0,
-\end{align}
-has been used as a nonparametric model for the time dependence of geophysical signals \citep[e.g.,][]{Segall1997,McGuire2003,Ohtani2010,Hines2016}.
-It should be noted that Brownian motion and integrated Brownian motion have a reference time at which the Gaussian process is exactly zero.
-For some geophysical signals, such as postseismic deformation, it is reasonable to have such a reference time.
-On the other hand, if there is no sense of a start time for the geophysical process, then Brownian motion or integrated Brownian motion may not be appropriate, and one of the isoptropic Gaussian processes described above should be used.
+GIVE A DEMONSTRATION HERE
 
 \section{Application Cascadia Slow Slip Events}\label{sec:Cascadia}
-
 We demonstrate the method for estimating transient strain described in the previous section by applying it to GNSS data in Cascadia.
@@
 In this section we discuss our choice for the noise covariance function $C_\epsilon$.
-We will assume that the noise consists of uncorrelated noise with standard deviation given by the formal data uncertainties and potentially spatially and/or temporally correlated noise.
+We will assume that the noise consists of an uncorrelated component with standard deviation given by the formal data uncertainties, $\bm{\sigma}$, and a spatially and/or temporally correlated component.
 There have been numerous studies on temporally correlated noise in GNSS data \citep[e.g.,]{Zhang1997,Mao1999,Williams2004,Langbein2008}.
-These studies have found some combination of Brownian motion or flicker noise best describes the temporally correlated noise in GNSS data.
-Brownian motion has also been used to describe error resulting from an unstable geodetic monument \citep[e.g.,][]{Wyatt1982,Wyatt1989}.
-We assume that temporally correlated noise can be described by a FOGM stochastic model, noting that the Brownian motion is a degenerate case of a FOGM process.
+In these studies, temporally correlated noise was described with some combination of Brownian motion, first-order Gauss-Markov (FOGM) processes, and/or flicker noise.
+There is some physical justification for using Brownian motion as a noise model because it accurately describes the power spectrum of motion resulting from instability in geodetic monuments \citep[e.g.,][]{Wyatt1982,Wyatt1989,Langbein1997}.
+We describe temporally correlated noise as a FOGM process, noting that Brownian motion is a degenerate case of a FOGM process.
+A FOGM process is a solution to the stochastic differential equation
+\begin{equation}
+\dot{v}(t) + \alpha v(t) = \beta w(t),
+\end{equation}
+where $w(t)$ is white noise with unit variance. The solution is an isotropic Gaussian process with zero mean and covariance function
+\begin{equation}\label{eq:FOGM}
+C_v(t,t') = \frac{\beta^2}{2\alpha}\exp\left(-\alpha|t - t'|\right). 
+\end{equation}
+The FOGM model degenerates to Brownian motion under the condition that $\alpha = 0$ and $v(0) = 0$.
 
-To determine the appropriate hyperparameters for the FOGM model, $\alpha$ and $\beta$, we first select 38 stations in in Washington and Oregon which are east of $121^\circ$ W, where we can reasonably expect no transient signal from SSEs (i.e.
+To determine the appropriate hyperparameters for the FOGM model, $\alpha$ and $\beta$, we first select 38 stations in Washington and Oregon which are east of $121^\circ$ W, where we can reasonably expect no transient signal from SSEs (i.e.
 $u = 0$).
@@
 The REML method finds the hyperparameters which maximize the probability of drawing $\bm{d}_*$ from $\bm{d}$.
-We use the REML method over the maximum likelihood (ML) method \citep[e.g.,][]{Langbein1997} because the REML method properly accounts for the improper prior that we assigned to $\bm{a}$.
+We use the REML method over the maximum likelihood (ML) method \citep[e.g.,][]{Langbein1997} because the REML method accounts for the improper prior that we assigned to $\bm{a}$ without biasing the hyperparameter estimates.
 The distribution of inferred $\alpha$ and $\beta$ are shown in Figure X.
-The amplitude of FOGM noise for the east and north components is notable low and are unimodally clustered around 0.5 mm/yr$^{0.5}$.
-The corresponding estimates of $\alpha$ tend to cluster around 0 yr$^{-1}$, which is the degenerate case where FOGM behaves likes Brownian motion.
+The amplitude of FOGM noise for the east and north components is notable low and are clustered around 0.5 mm/yr$^{0.5}$.
+The corresponding estimates of $\alpha$ tend to cluster around 0 yr$^{-1}$, suggesting that Brownian motion would have also been an appropriate noise model.
 The amplitude of FOGM noise for the vertical component is significantly larger with a median value of 13.8 mm/yr$^{0.5}$.
@@
 The samples span seven years and over these seven years the east and north samples drift by about 1 mm.
-In the context of detecting SSEs, which produce several mm's of surface displacement on the timescale of weeks to months, the estimated FOGM noise for the east and north component is negligible.
+In the context of detecting SSEs, which produce several mm's of surface displacement on the timescale of weeks, the estimated FOGM noise for the east and north component is negligible.
 In contrast, the estimated FOGM noise for the vertical component is significantly larger than the signal we would expect from SSEs.
@@
 
-Another significant source of noise in GNSS data is common mode error \citep[e.g.,][]{Wdowinski1997}, which is noise that is highly spatially correlated.
+Another significant source of noise in GNSS data is common mode error \citep[e.g.,][]{Wdowinski1997,Dong2006}, which is noise that is highly spatially correlated.
 When not accounted for, common mode error manifests itself as spatially uniform undulations in $\hat{u}$.
@@
 
-Our noise is just white.
+For our estimation of of transient strain resulting from SSEs, we then simply assume that $\epsilon$ consists of only uncorrelated noise with covariance $C_\epsilon(\bm{y},\bm{y}') = \mathrm{diag}(\sigma^2)$.
 
 \subsection{Transient displacement model}
+We next discuss our prior model for transient displacements, and, specifically, our choice for the covariance functions $X(x,x')$ and $T(t,t')$.
+For $X$, we use the squared exponential (SE) covariance function,
+\begin{equation}\label{eq:SE}
+X(x,x') = \alpha^2 \exp\left(\frac{-||x - x'||_2^2}{2 \beta^2}\right).
+\end{equation}
+The SE covariance function is commonly used in Kriging \citep[e.g,][]{Cressie1992} and Gaussian process regression \citep[e.g.,][]{Rasmussen2006}.
+The SE is a valid (i.e.
+positive definite) covariance function for any number of spatial dimensions, and it describes an isotropic Gaussian processes with realizations that are infinitely differentiable.
+In terms of geodetic applications, \citet{Kato1998} and \cite{El-Fiky1999} demonstrated that the SE is an appropriate covariance model for describing long-term tectonic strain rates in Japan.
+
+We consider three potential covariance functions to describe the temporal covariance of $u$.
+First, we consider the one-dimensional SE covariance function, simply because the SE describes a good general purpose Gaussian process.
+Second, we consider integrated Brownian motion (IBM).
+IBM has zero mean and its covariance function can be found by integrating the covariance function for Brownian motion as
+\begin{align}\label{eq:IBM}
+T(t,t') &= \int_0^t \int_0^{t'} \alpha^2 \min(\tau,\tau') \,d\tau'\,d\tau \\
+        &= \frac{\alpha^2}{2}\min(t,t')^2 \left(\max(t,t') - \frac{1}{3}\min(t,t')\right), \ \ \ t,t' \geq 0.
+\end{align}
+IBM has been used, in the context of Kalman filtering, as a nonparametric model for the time dependence of geophysical signals \citep[e.g.,][]{Segall1997,McGuire2003,Ohtani2010,Hines2016}.
+It should be noted that IBM has a reference time at which the Gaussian process is exactly zero.
+For some geophysical signals, such as postseismic deformation, it is appropriate to have this reference time.
+However, if we are interesting in detecting transient events, where there is no known start time, then the IBM model may not be appropriate and an isotropic Gaussian process should be preferred.
+Alternatively we can assume that the start date for the IBM process is long before the first data obvservation; however this can introduce numerical difficulties because the covariance matrix will become ill-conditioned.
+Third, we consider the Wendland class of covariance functions \citep{Wendland2005}.
+Wendland functions are positive definite in $\mathbb{R}^d$, and they describes an isotropic Gaussian process with realizations that can be differentiated $k$ times.
+The form of the covariance function depends on the choice of $d$ and $k$.
+Wendland covariance functions have compact support, which is a particularly useful property because we can exploit the sparsity of the the corresponding covariance matrices.
+To describe the temporal covariance of $u$, we only require that $d=1$ and $k=1$.
+The corresponding Wendland covariance function is
+\begin{equation}\label{eq:Wendland}
+T(t,t') = \alpha^2\left(1 - \frac{|t - t'|}{\beta}\right)^3_+ \left(\frac{3|t - t'|}{\beta} + 1\right), 
+\end{equation}
+where
+\begin{equation}
+(t)_+ = 
+\begin{cases}
+t, \ \ \ t > 0 \\
+0, \ \ \ \mathrm{otherwise}.
+\end{cases}
+\end{equation}
+
+We next find the optimal hyperparameters for $X$ and each of the three candidate functions for $T$.
+We want to find the hyperparameters which best describes the spatial and temporal statistics of ground deformation resulting from SSEs.
+We focus on describing the SSEs in the Puget Sound region, which regularly occur with a frequency of about 14 months \citep{Brudzinski2007}.
+
+To describe these events we use the set of GNSS station shown in Figure X.
+We use the GNSS data that has been made publicly available through UNAVCO \citep{Herring2016}.
+We first remove timeseries jumps based on the compiled GNSS offsets from blewitt.
+We then use the data editing algorithm to remove outliers.
+We find the data editing algorithm is particularly useful at removing outliers for stations at higher altitudes, which can be adversely affected by snow \citep{Lisowski2008}.
+For our data editing algorithm we use X for the prior and we give an example in figure X.
+
+We use the REML method to find the hyperparameters that best describe SSE events which occurred since 2010, corresponding to the time that coverage improved due to the PBO network.
+We identify the times of seven SSEs using tremor records from \cite{Wech2010}, and we break the spatio-temporal GNSS data into four month long subsets roughly centered on the time of each SSE.
+For each data subset use the REML method to find the hyperparameters for $T$ and $X$ that best describe deformation from the SSE.
+Our motivation analyzing subsets of the data is two-fold: (1) it keeps the analysis computationally tractible and (2) the distribution of estimated hyperparameters will indicates whether we can use a single covariance model to describe all SSEs.
+Our hyperparameter estimates are summarized in Table X.
+
+We next want to identify which covariance function for $T$ best describes the SSEs.
+We could use the REML likelihoods to compare different models, similar to \citep{Langbein2004}.
+However, the maximum likelihoods do not vary by more than a factor of X between the models, which does not definitively identify the best model.
+We then compare the posterior estimates $\hat{u}$ for each model to the observations to see if any model is noticeably better at describing the data.
+We show $\hat{u}$ using the three different function for $T$ in Figure X.
+It is clear that neither model offer any distinct improvement over the others.
+For most stations $\hat{u}$ appears to accurately capture the deformation from SSEs; however, for some stations which are closer to Puget Sound such as X and X, the rate of deformation is underestimated by the best fitting model.
+We could have manually specified a prior.
+We choose the Wendland function to describe $T$ because evaluating $T$ will yield sparse covariance matrices.
+We utilize the sparsity of the covariance matrices with the sparse matrix library CHOLMOD \citep{Chen2008}, which significantly expands the amount of data which we are able to tractably analyze with the method described in Section X.
+
+\subsection{Transient Strain Rates}
+
+Having established noise model and a prior for transient displacements, we calculate the transient strain rates in the Puget Sound region from eq.
+(\ref{eq:StrainRate}) over the time period 2010-01-01 to 2017-05-15.
+Maybe this is the best place to mention the prior for editing the data.
+Our solution is shown in Figures X through X.
+Figures X through X show the estimates $\dot{\bm{\epsilon}}$ complete with uncertainties at select times over the course of the Winter 2016 SSE.
+Figure X, which shows strain rate at the height of the SSE, makes the implications of SSEs for seismic hazard clear.
+The SSE causes compression in the Olympic Peninsula and extension in East of Puget Sound.
+The SSE are thus concentrating strain energy trench-ward, and pushing the subduction zone closer to failure.
+This is not novel insight because the same conclusions has been drawn from fault slip models which show that SSEs tend to be just at the transition between locked and decoupled.
+Slow slip event depths from \citet{Dragert2001}, \citet{Wech2009}, \citet{Schmidt2010}, \citet{Bartlow2011} show slip concentrated at depths from 30 to 50 km detph.
+What is novel is that our estimates of strain rate from SSEs are nonparametric and less likely to be biased by systematic errors such as errors in the fault model.
+Are uncertainties are also likely to be more conservative since we do not make any dubious assumptions.
+Thus we can be more confident in our assessment of how much strain is accumulating from SSEs.
+
+To demonstrate the effectiveness of our method for detecting SSEs, we show a timeseries of estimated strain components at the position indicated in figure X.
+We also compare the strain rate magnitude and the frequency of tremors in this region.
+The strain rate magnitudes have been normalized by its estimated uncertainty and so strain magnitudes less than ${\sim}$2 can be considered to be part of the background noise.
+The only time that the normalized strain rate magnitude exceeds 2 is when there is a corresponding peak in tremor.
+We are able to accurately detect the seven SSEs in over this period in addition to the inter-SSE events such as the one at X.
+
+made in ,  is accumulating to the trench perpendicular compression releases At the height of the SSE, the implications of
 
@@
 Explain strain concentration with this:
-Slow slip event depths from \citet{Dragert2001}, \citet{Wech2009}, \citet{Schmidt2010}, \citet{Bartlow2011} show slip concentrated at depths from 30 to 50 km detph.
 

### 2017-06-06 figures/outliers/make_plot.py
added more

@@ -1,0 +1,59 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import pygeons
+
+def trend(u1,t1):
+  idx = np.isfinite(u1)
+  u2 = u1[idx]
+  t2 = t1[idx]
+  G = np.array([np.ones(t2.shape[0]),t2]).T
+  m = np.linalg.inv(G.T.dot(G)).dot(G.T).dot(u2)
+  return m
+  
+
+
+xtick_dates = ['2011-01-01','2012-01-01','2013-01-01','2014-01-01','2015-01-01','2016-01-01','2017-01-01']
+xticks = [pygeons.mjd.mjd(i,'%Y-%m-%d') for i in xtick_dates]
+xtick_labels = ['2011-01-01','2012-01-01','2013-01-01','2014-01-01',' ','2016-01-01',' ']
+
+
+
+
+# plot raw data
+fig,ax = plt.subplots(figsize=(7,3.5))
+pygeons.plot.plot._setup_ts_ax([ax])
+ax.set_xticks(xticks)
+ax.set_xticklabels(xtick_labels)
+ax.tick_params(labelsize=10)
+ax.set_xlim((55378.0,56839.0))
+ax.set_ylim((-10.0,15.0))
+ax.grid()
+
+data = pygeons.io.io.dict_from_hdf5('data-2017-05-17.h5')
+idx = (data['id'] == 'SC03').nonzero()[0][0]
+lon = data['longitude'][idx]
+lat = data['latitude'][idx]
+
+t = data['time']
+u = 1000*data['east'][:,idx]
+us = 1000*data['east_std_dev'][:,idx]
+m = trend(u,t)
+u = u - m[0] - m[1]*t
+ax.errorbar(t,u,us,marker='None',linestyle='None',color=(1.0,0.7,0.7))
+ax.plot(t,u,'r.',ms=4.99)
+
+data = pygeons.io.io.dict_from_hdf5('data.final.h5')
+idx = (data['id'] == 'SC03').nonzero()[0][0]
+t = data['time']
+u = 1000*data['east'][:,idx]
+us = 1000*data['east_std_dev'][:,idx]
+u = u - m[0] - m[1]*t
+ax.errorbar(t,u,us,marker='None',linestyle='None',color=(0.7,0.7,0.7))
+ax.plot(t,u,'k.',ms=5.0)
+
[…]

### 2017-06-06 HinesHetlandGPR.tex
added more

@@
 \begin{equation}\label{eq:TransientCovariance}
-C_{u}\left((x,t),(x',t')\right) = X(x,x')T(t,t').
+C_{u}\left((x,t),(x',t')\right) = \alpha^2 X(x,x')T(t,t'),
 \end{equation}  
+where $\alpha$ is a hyperparameter.
 The appropriate choice for $X$ and $T$ may vary depending on the geophysical signal we are trying to describe (e.g.
@@
 Our method for detecting outliers is based on the data editing algorithm described in \citet{Gibbs2011}.
-We again treat $\bm{d}_*$ as a sample of $\bm{d}$, and we assume some prior covariance function for $u$.
-The choice for $u$ here does not need to be the same as that used in Section X, and a conservative approach would be to choose a $u$ that is more flexible (e.g., shorter characteristic length-scales and higher variance).
+We calculate the residuals between the observations and a best fitting model, and data with residuals that are anomalously large are identified as outliers.
+We again treat $\bm{d}_*$ as a sample of $\bm{d}$.
+The best fitting model for $\bm{d}_*$ is considered to be the expected value of the random vector $u(\bm{y}) + \bm{P}\bm{a}$ after conditioning it with non-outlier observations.
 We let $\epsilon$ be uncorrelated noise which has standard deviation given by the formal data uncertainties, $\bm{\sigma}$.
-We consider outliers to be observations which are abnormally inconsistent with the random vector $u(\bm{y}) + \bm{P}\bm{a}$.
-With this criterion, we takes into account the spatial and temporal relationship between the data, allowing us to identify anomalous observations which may not be immediately apparent based on inspection of individual timeseries.
+We still consider $u$ to have a separable covariance function as in eq.
+(\ref{eq:TransientCovariance}), and the choice for $X$ and $T$ does not need to be the same as that used in Section X.
+Since outliers are determined based on how well a spatially and temporally dependent model fits the data, we are able to identify anomalous observations which may not be immediately apparent based on inspection of individual timeseries.
+
 To begin the algorithm, we let $\Omega$ be the index set of non-outliers in $\bm{d}_*$ and initiate it with all $n$ indices.
@@
 \end{align}
-where the tilde indicates that only elements corresponding to indices in $\Omega$ are retained (e.g., $\tilde{\bm{y}} = (y_i)_{i\in\Omega}$). We then update $\Omega$ to be
+where the tilde indicates that only elements corresponding to indices in $\Omega$ are retained (e.g., $\tilde{\bm{y}} = \{y_i\}_{i\in\Omega}$). We then update $\Omega$ to be
 \begin{equation}\label{eq:Update}
-\Omega = \{i : |r_i| < \eta \tilde{\bm{r}}_{rms}\}
+\Omega = \{i : |r_i| < \eta \tilde{\bm{r}}_{rms}\}, \ \ \ r_i \in \bm{r}
 \end{equation} 
-where $\tilde{\bm{r}}_{rms}$ is the root-mean-square of $\tilde{\bm{r}}$ and $\lambda$ is an outlier tolerance.
-We use $\lambda=4$ in this study, which seems to accurately identify outliers without unnecessarily decimating the data.
-Iterations continue until the new $\Omega$ is equal to the previous $\Omega$ (i.e., no new outliers are detected).
+where $\tilde{\bm{r}}_{rms}$ is the root-mean-square of $\tilde{\bm{r}}$ and $\eta$ is an outlier tolerance.
+We use $\eta=4$ in this study, which seems to accurately identify outliers without unnecessarily decimating the data.
+Iterations continue until the new $\Omega$ is equal to the previous $\Omega$.
 
@@
 
-GIVE A DEMONSTRATION HERE
-
 \section{Application Cascadia Slow Slip Events}\label{sec:Cascadia}
-We demonstrate the method for estimating transient strain described in the previous section by applying it to GNSS data in Cascadia.
+In this section we estimate transient strain from GNSS data in Cascadia.
 We are mainly interested in identifying transient strain resulting from slow slip events (SSEs) \citep[e.g.,][]{Dragert2001}.
 We show that Gaussian processes are an effective tool for (1) detecting SSEs and (2) understanding how the strain from SSEs are contributing to the strain accumulated over the seismic cycle.
-Our primary motivation for detecting SSEs is because there have been several documented cases of SSEs triggering megathrust earthquakes \citep{Roeloffs2006}.
-In Cascadia, there is less of a need to detect SSEs with GNSS data because the SSEs are associated with a spike in seismic tremor, which is actively monitored by the Pacific Northwest Seismic Network (PNSN) \citep{Wech2010}.
-However, SSEs have been observed in other subduction zones without any associated tremor \citep{Schwartz2007}, and so there is still utility in developing method for detecting SSEs with GNSS data.
-Furthermore, by comparing our estimated transient strain in Cascadia with the tremor catalog, we can verify the accuracy of method for detecting SSEs.
+Our motivation for detecting SSEs is, in part, because there have been several instances where SSEs immediately preceded megathrust earthquakes \citep{Roeloffs2006}.
+In Cascadia, SSEs can be detected by monitoring for associated seismic tremor, which is actively being done by the Pacific Northwest Seismic Network (PNSN) \citep{Wech2010}.
+Here we demonstrate that strain rates derived from GNSS data can also be an effective tool for detecting SSEs.
+Furthermore, estimated strain rates provide a clear picture of how SSEs transfer elastic strain energy through the crust, which is also important information for assessing seismic hazard.
+We envision that GNSS derived transient strain rates could be useful for monitoring subduction zones which have SSEs but no associated tremor \citep{Schwartz2007}.
+
+In this study, we use GNSS data which has been made publicly available by the University Navstar Consortium (UNAVCO) \citep{Herring2016} and can be found at www.unavco.org.
+We limit the dataset to the stations and time ranges which are pertinent to seven of the most recent SSEs in the Puget Sound region; the earliest SSE began in the August 2010, and the most recent SSE began in the February of 2017.
+We consider these most recent SSEs because the station coverage is sufficiently dense for us to make well constrained inferences of optimal prior models.
+The distribution of stations considered in this station is shown in Figure X.
 
@@
 $u = 0$).
+We remove jumps and clean the data with the outlier detection algorithm described in section X.
 We then find $\alpha$ and $\beta$ for each station timeseries with the Restricted Maximum Likelihood (REML) method \cite[e.g.,][]{Harville1974,Cressie1992} and Hines 2017.
@@
 
+\begin{figure*}
+\includegraphics{figures/noise/noise-params.pdf}
+\caption{Distribution of estimated FOGM hyperparameters from eq.
+(\ref{eq:FOGM}).
+Hyperparameters are independently estimated for each of the stations shown in Figure X and for each displacement component.
+``IQR'' is the inter-quartile range.}
+\label{fig:NoiseParams}
+\end{figure*}
+
+\begin{figure*}
+\includegraphics{figures/noise/noise-samples.pdf}
+\caption{Two FOGM noise samples, where the hyperparameters have been set to the median values from Figure \ref{fig:NoiseParams}.}   
+\label{fig:NoiseSamples}
+\end{figure*}
+
 Another significant source of noise in GNSS data is common mode error \citep[e.g.,][]{Wdowinski1997,Dong2006}, which is noise that is highly spatially correlated.
@@
 \begin{equation}\label{eq:SE}
-X(x,x') = \alpha^2 \exp\left(\frac{-||x - x'||_2^2}{2 \beta^2}\right).
+X(x,x') = \exp\left(\frac{-||x - x'||_2^2}{2 \beta^2}\right).
 \end{equation}
@@
 \begin{align}\label{eq:IBM}
-T(t,t') &= \int_0^t \int_0^{t'} \alpha^2 \min(\tau,\tau') \,d\tau'\,d\tau \\
-        &= \frac{\alpha^2}{2}\min(t,t')^2 \left(\max(t,t') - \frac{1}{3}\min(t,t')\right), \ \ \ t,t' \geq 0.
+T(t,t') &= \int_0^t \int_0^{t'} \min(\tau,\tau') \,d\tau'\,d\tau \\
+        &= \frac{1}{2}\min(t,t')^2 \left(\max(t,t') - \frac{1}{3}\min(t,t')\right), \ \ \ t,t' \geq 0.
 \end{align}
@@
 \begin{equation}\label{eq:Wendland}
-T(t,t') = \alpha^2\left(1 - \frac{|t - t'|}{\beta}\right)^3_+ \left(\frac{3|t - t'|}{\beta} + 1\right), 
+T(t,t') = \left(1 - \frac{|t - t'|}{\beta}\right)^3_+ \left(\frac{3|t - t'|}{\beta} + 1\right), 
 \end{equation}
@@
 
-We next find the optimal hyperparameters for $X$ and each of the three candidate functions for $T$.
-We want to find the hyperparameters which best describes the spatial and temporal statistics of ground deformation resulting from SSEs.
-We focus on describing the SSEs in the Puget Sound region, which regularly occur with a frequency of about 14 months \citep{Brudzinski2007}.
-
-To describe these events we use the set of GNSS station shown in Figure X.
-We use the GNSS data that has been made publicly available through UNAVCO \citep{Herring2016}.
-We first remove timeseries jumps based on the compiled GNSS offsets from blewitt.
-We then use the data editing algorithm to remove outliers.
-We find the data editing algorithm is particularly useful at removing outliers for stations at higher altitudes, which can be adversely affected by snow \citep{Lisowski2008}.
-For our data editing algorithm we use X for the prior and we give an example in figure X.
-
-We use the REML method to find the hyperparameters that best describe SSE events which occurred since 2010, corresponding to the time that coverage improved due to the PBO network.
-We identify the times of seven SSEs using tremor records from \cite{Wech2010}, and we break the spatio-temporal GNSS data into four month long subsets roughly centered on the time of each SSE.
-For each data subset use the REML method to find the hyperparameters for $T$ and $X$ that best describe deformation from the SSE.
-Our motivation analyzing subsets of the data is two-fold: (1) it keeps the analysis computationally tractible and (2) the distribution of estimated hyperparameters will indicates whether we can use a single covariance model to describe all SSEs.
+We next want to determine appropriate hyperparameters for $X$ and each of the three candidate functions for $T$.
+We use the REML method to find the optimal hyperparameters for describing each of the seven SSEs.
+Before using the REML method, we clean the GNSS datasets by removing offsets at times of equipment changes and removing outliers with the method describe in Section X.
+For the outlier detection algorithm, our prior model, $u$, is chosen to be have a length-scale and time-scale which is sufficiently short to describe SSE displacements.
+We use the Squared Exponential covariance function for $X$ with $\beta = 100$ km, and we use the Wendland covariance function for $T$, due to its computational efficiency, with $\beta = 0.1$ yr$^{-1}$.
+The product of the two covariance functions is then scaled by $\alpha^2 = 1$ mm$^{2}$.
+The outlier detection algorithm is particularly effective at removing outliers for stations at high elevation, which can be adversely affected by ice or snow \citep{Lisowski2008}.
+As an example, Figure \ref{fig:Outliers} shows the outliers which were automatically detected for station SC03, which is located on Mount Olympus in Washington.
+Note that most of the detected outliers occur during the winter and can likely be attributed to ice or snow.
+After cleaning the dataset, we divide it into four month long subsets, where each subset is roughly centered on the time of an SSE.
+We identify the times of seven SSEs using tremor records from \cite{Wech2010}.
+For each data subset, we use the REML method to find the hyperparameters for $T$ and $X$ that best describe deformation from the SSE.
+Our motivation analyzing subsets of the data is two-fold: (1) it keeps the analysis computationally tractable and (2) the distribution of estimated hyperparameters will indicates whether we can use a single covariance model to describe all SSEs.
 Our hyperparameter estimates are summarized in Table X.
 
+\begin{figure*}
+\includegraphics{figures/outliers/outliers.pdf}
+\caption{Detrended easting component of GNSS displacement data for station SC03, which is located on Mt.
+Olympus in Washington.
+The orange markers indicate outliers which were automatically detected using the algorithm from Section \ref{sec:Method}.
+The error bars show one standard deviation uncertainty.}
+\label{fig:Outliers}
+\end{figure*}
+
 We next want to identify which covariance function for $T$ best describes the SSEs.
-We could use the REML likelihoods to compare different models, similar to \citep{Langbein2004}.
+We could use the REML likelihoods to compare different models, similar to \citet{Langbein2004}.
 However, the maximum likelihoods do not vary by more than a factor of X between the models, which does not definitively identify the best model.
@@
 \subsection{Transient Strain Rates} 
-
-Having established noise model and a prior for transient displacements, we calculate the transient strain rates in the Puget Sound region from eq.
-(\ref{eq:StrainRate}) over the time period 2010-01-01 to 2017-05-15.
-Maybe this is the best place to mention the prior for editing the data.
+Having established noise model and a prior for transient displacements, we use cleaned GNSS dataset described in the previous section to calculate transient strain rates in the Puget Sound region from eq.
+(\ref{eq:StrainRate}).
 Our solution is shown in Figures X through X.

### 2017-06-06 figures/noise/make-plot.py

@@ -1,0 +1,162 @@
+#!/usr/bin/env python
+import numpy as np
+from pygeons.io.parser import _get_field
+from pygeons.main.gpstation import fogm
+from pygeons.basemap import make_basemap
+from pygeons.plot.plot import _setup_map_ax
+import os
+import matplotlib.pyplot as plt
+from mpl_toolkits.basemap import Basemap
+np.random.seed(0)
+
+def get_meridians_and_parallels(bm,ticks):
+  ''' 
+  returns the meridians and parallels that should be plotted.
+  '''
+  diff_lon = (bm.urcrnrlon-bm.llcrnrlon)
+  round_digit = int(np.ceil(np.log10(ticks/diff_lon)))
+  dlon = np.round(diff_lon/ticks,round_digit)
+
+  diff_lat = (bm.urcrnrlat-bm.llcrnrlat)
+  round_digit = int(np.ceil(np.log10(ticks/diff_lat)))
+  dlat = np.round(diff_lat/ticks,round_digit)
+
+  meridians = np.arange(np.floor(bm.llcrnrlon),
+                        np.ceil(bm.urcrnrlon),dlon)
+  parallels = np.arange(np.floor(bm.llcrnrlat),
+                        np.ceil(bm.urcrnrlat),dlat)
+
+  return meridians,parallels
+
+bins = 20
+files = os.listdir('results')
+
+params0 = {'east':[],'north':[],'vertical':[]}
+params1 = {'east':[],'north':[],'vertical':[]}
+lons = []
+lats = []
+for f in files:
+  buff = open('results/%s' % f,'r')
+  content = buff.read()
+  buff.close()
+  lat = _get_field('latitude range',content)
+  lat = float(lat.strip().split(',')[0])
+  lon = _get_field('longitude range',content)
+  lon = float(lon.strip().split(',')[0])
+  lons += [lon]
+  lats += [lat]
+  content = content[content.find('REML RESULTS'):]
+  content = content[content.find('station :'):]
+  for dir in ['east','north','vertical']:
+    params = _get_field('optimal %s parameters' % dir,content).strip().split(',')
[…]

### 2017-06-06 figures/outliers/make_plot.py

@@ -1,6 +1,7 @@
 import numpy as np
 import matplotlib.pyplot as plt
 import pygeons
+from matplotlib import colors
 
 def trend(u1,t1):
   idx = np.isfinite(u1)
@@ -17,7 +18,10 @@ xticks = [pygeons.mjd.mjd(i,'%Y-%m-%d') for i in xtick_dates]
 xtick_labels = ['2011-01-01','2012-01-01','2013-01-01','2014-01-01',' ','2016-01-01',' ']
 
 
-
+c1 = colors.to_rgb('C0')
+c1t = tuple(0.6 + 0.4*np.array(colors.to_rgb('C0')))
+c2 = colors.to_rgb('C1')
+c2t = tuple(0.6 + 0.4*np.array(colors.to_rgb('C1')))
 
 # plot raw data
 fig,ax = plt.subplots(figsize=(7,3.5))
@@ -39,8 +43,8 @@ u = 1000*data['east'][:,idx]
 us = 1000*data['east_std_dev'][:,idx]
 m = trend(u,t)
 u = u - m[0] - m[1]*t
-ax.errorbar(t,u,us,marker='None',linestyle='None',color=(1.0,0.7,0.7))
-ax.plot(t,u,'r.',ms=4.99)
+ax.errorbar(t,u,us,marker='None',linestyle='None',color=c2t)
+ax.plot(t,u,'.',ms=4.99,color=c2)
 
 data = pygeons.io.io.dict_from_hdf5('data.final.h5')
 idx = (data['id'] == 'SC03').nonzero()[0][0]
@@ -48,8 +52,8 @@ t = data['time']
 u = 1000*data['east'][:,idx]
 us = 1000*data['east_std_dev'][:,idx]
 u = u - m[0] - m[1]*t
-ax.errorbar(t,u,us,marker='None',linestyle='None',color=(0.7,0.7,0.7))
-ax.plot(t,u,'k.',ms=5.0)
+ax.errorbar(t,u,us,marker='None',linestyle='None',color=c1t)
+ax.plot(t,u,'.',ms=5.0,color=c1)
 
 ax.set_title('Station SC03 (%.2f$^\mathregular{\circ}$ W, %.2f$^\mathregular{\circ}$ N)' % 
                  (-lon,lat),fontsize=10)

### 2017-06-07 figures/noise/make-plot.py
more

@@ -120,7 +120,7 @@ plt.savefig('noise-params.pdf',format='pdf')
 fig,ax = plt.subplots(3,1,figsize=(7,4),sharex=True)
 
 dt = 1.0/365.25
-time = np.arange(0.0,7.0,dt)
+time = np.arange(0.0,5.0,dt)
 
 gp = fogm(np.median(east_p0),np.median(east_p1),convert=False)
 sample = gp.sample(time[:,None],use_cholesky=True)
@@ -153,7 +153,7 @@ ax[2].grid(c='0.5',alpha=0.5)
 ax[2].set_xlabel('time [yr]',fontsize=10)
 ax[2].set_ylabel('vertical [mm]',fontsize=10)
 ax[2].tick_params(labelsize=10)
-ax[2].set_xlim((0.0,7.0))
+ax[2].set_xlim((0.0,5.0))
 fig.tight_layout()
 plt.savefig('noise-samples.pdf',format='pdf')
 

### 2017-06-07 HinesHetlandGPR.tex
added more

@@
 \begin{equation}\label{eq:TransientCovariance}
-C_{u}\left((x,t),(x',t')\right) = \alpha^2 X(x,x')T(t,t'),
+C_{u}\left((x,t),(x',t')\right) = X(x,x')T(t,t'),
 \end{equation}  
-where $\alpha$ is a hyperparameter.
 The appropriate choice for $X$ and $T$ may vary depending on the geophysical signal we are trying to describe (e.g.
@@
 We consider the six coefficients in eq.
-(\ref{eq:Data}) to be uncorrelated random variables distributed as $\mathcal{N}(0,\lambda^2)$ in the limit as $\lambda \to \infty$ (i.e., the coefficients have diffuse priors).
+(\ref{eq:Data}) to be uncorrelated random variables distributed as $\mathcal{N}(0,\kappa^2)$ in the limit as $\kappa \to \infty$ (i.e., the coefficients have diffuse priors).
 Of course, the tectonic deformation, $b_{i}$, is spatially correlated and we could invoke a tectonic model to form a prior on $b_{i}$.
@@
 \bm{C_d} = C_u(\bm{y},\bm{y}) + C_\epsilon(\bm{y},\bm{y}) + 
-           \lambda^2\bm{P}\bm{P}^T.
+           \kappa^2\bm{P}\bm{P}^T.
 \end{equation}
@@
 We will assume that an appropriate covariance functions and corresponding hyperparameters for $X$, $T$, and $C_\epsilon$ have already been chosen.
-If $\lambda$ is kept finite, then we can follow \citet{Rasmussen2006} to find that $\hat{u}$ is distributed as $\mathcal{N}(\mu_{\hat{u}},C_{\hat{u}})$, where
+If $\kappa$ is kept finite, then we can follow \citet{Rasmussen2006} to find that $\hat{u}$ is distributed as $\mathcal{N}(\mu_{\hat{u}},C_{\hat{u}})$, where
 \begin{equation}\label{eq:PosteriorMean}
@@
 \end{equation}
-However, we are interested in the limit as $\lambda \to \infty$, and the form for eq.
+However, we are interested in the limit as $\kappa \to \infty$, and the form for eq.
 (\ref{eq:PosteriorMean}) and eq.
@@
 C_u(\bm{y},\bm{y}) + C_\epsilon(\bm{y},\bm{y}) & \bm{P} \\
-\bm{P}^T  & -\lambda^{-2} \bm{I} \\
+\bm{P}^T  & -\kappa^{-2} \bm{I} \\
 \end{array}
@@
 \end{equation}
-Taking the limit as $\lambda \to \infty$, we get the solution for the mean and covariance of $\hat{u}$,
+Taking the limit as $\kappa \to \infty$, we get the solution for the mean and covariance of $\hat{u}$,
  \begin{equation}\label{eq:PosteriorMean3}
@@
 
-\subsection{Outlier detection}
+\subsection{Outlier detection}\label{sec:Outlier}
 In our formulation for estimating transient strain rates, we have assumed that noise in the data vector is normally distributed.
@@
 
-\subsection{Noise model}
-In this section we discuss our choice for the noise covariance function $C_\epsilon$.
+\subsection{Noise model}\label{sec:NoiseModel}
+In this section we discuss our choice for the noise covariance function $C_\epsilon$ which will be used when estimating transient strain.
 We will assume that the noise consists of an uncorrelated component with standard deviation given by the formal data uncertainties, $\bm{\sigma}$, and a spatially and/or temporally correlated component.
-There have been numerous studies on temporally correlated noise in GNSS data \citep[e.g.,]{Zhang1997,Mao1999,Williams2004,Langbein2008}.
+There have been numerous studies on temporally correlated noise in GNSS data \citep[e.g.,][]{Zhang1997,Mao1999,Williams2004,Langbein2008}.
 In these studies, temporally correlated noise was described with some combination of Brownian motion, first-order Gauss-Markov (FOGM) processes, and/or flicker noise.
@@
 \end{equation}
-The FOGM model degenerates to Brownian motion under the condition that $\alpha = 0$ and $v(0) = 0$.
+A FOGM process degenerates to Brownian motion under the condition that $\alpha = 0$ and $v(0) = 0$.
 
-To determine the appropriate hyperparameters for the FOGM model, $\alpha$ and $\beta$, we first select 38 stations in Washington and Oregon which are east of $121^\circ$ W, where we can reasonably expect no transient signal from SSEs (i.e.
-$u = 0$).
-We remove jumps and clean the data with the outlier detection algorithm described in section X.
+We use a set of 38 inland stations, shown in Figure X, to constrain the FOGM hyperparameters, $\alpha$ and $\beta$.
+These stations are sufficiently far from the subduction zone that they are unlikely to contain transient signal, and thus we assume that $u$ is zero.
+We clean the data for these stations by removing jumps at times of equipment changes, and we remove outliers that have been detected with the algorithm described in Section \ref{sec:Outlier}.
 We then find $\alpha$ and $\beta$ for each station timeseries with the Restricted Maximum Likelihood (REML) method \cite[e.g.,][]{Harville1974,Cressie1992} and Hines 2017.
 The REML method finds the hyperparameters which maximize the probability of drawing $\bm{d}_*$ from $\bm{d}$.
-We use the REML method over the maximum likelihood (ML) method \citep[e.g.,][]{Langbein1997} because the REML method accounts for the improper prior that we assigned to $\bm{a}$ without biasing the hyperparameter estimates.
-The distribution of inferred $\alpha$ and $\beta$ are shown in Figure X.
+We use the REML method over the maximum likelihood (ML) method \citep[e.g.,][]{Langbein1997} because the REML method properly accounts for the improper prior that we assigned to $\bm{a}$.
+The distribution of inferred $\alpha$ and $\beta$ are shown in Figure \ref{fig:NoiseParams}.
 The amplitude of FOGM noise for the east and north components is notable low and are clustered around 0.5 mm/yr$^{0.5}$.
@@
 
-\subsection{Transient displacement model}
+\subsection{Transient displacement model}\label{sec:SignalModel}
 We next discuss our prior model for transient displacements, and, specifically, our choice for the covariance functions $X(x,x')$ and $T(t,t')$.
@@
 \begin{equation}\label{eq:SE}
-X(x,x') = \exp\left(\frac{-||x - x'||_2^2}{2 \beta^2}\right).
+X(x,x') = \exp\left(\frac{-||x - x'||_2^2}{2 \lambda^2}\right).
 \end{equation}
@@
 
-We consider three potential covariance functions to describe the temporal covariance of $u$.
-First, we consider the one-dimensional SE covariance function, simply because the SE describes a good general purpose Gaussian process.
+We consider three potential covariance functions to describe the temporal covariance of $u$.  First, we consider the one-dimensional SE covariance function, 
+\begin{equation}\label{eq:TimeSE}
+T(t,t') = \phi^2\exp\left(\frac{-|t - t'|^2}{\theta^2}\right),
+\end{equation}
+simply because the SE covariance function is versatile and commonly used for non-parametric modeling.
+Note that $T$ includes the hyperparameter $\phi$, which serves to scale the covariance function $C_u$.
 Second, we consider integrated Brownian motion (IBM).
@@
 \begin{align}\label{eq:IBM}
-T(t,t') &= \int_0^t \int_0^{t'} \min(\tau,\tau') \,d\tau'\,d\tau \\
-        &= \frac{1}{2}\min(t,t')^2 \left(\max(t,t') - \frac{1}{3}\min(t,t')\right), \ \ \ t,t' \geq 0.
+T(t,t') &= \int_0^t \int_0^{t'} \phi^2 \min(\tau,\tau') \,d\tau'\,d\tau \\
+        &= \frac{\phi^2}{2}\min(t,t')^2 \left(\max(t,t') - \frac{1}{3}\min(t,t')\right), \ \ \ t,t' \geq 0.
 \end{align}
-IBM has been used, in the context of Kalman filtering, as a nonparametric model for the time dependence of geophysical signals \citep[e.g.,][]{Segall1997,McGuire2003,Ohtani2010,Hines2016}.
+IBM has been used, in the context of Kalman filtering, as a non-parametric model for the time dependence of geophysical signals \citep[e.g.,][]{Segall1997,McGuire2003,Ohtani2010,Hines2016}.
 It should be noted that IBM has a reference time at which the Gaussian process is exactly zero.
@@
 \begin{equation}\label{eq:Wendland}
-T(t,t') = \left(1 - \frac{|t - t'|}{\beta}\right)^3_+ \left(\frac{3|t - t'|}{\beta} + 1\right), 
+T(t,t') = \phi^2\left(1 - \frac{|t - t'|}{\theta}\right)^3_+ \left(\frac{3|t - t'|}{\theta} + 1\right), 
 \end{equation}
@@
 We next want to determine appropriate hyperparameters for $X$ and each of the three candidate functions for $T$.
-We use the REML method to find the optimal hyperparameters for describing each of the seven SSEs.
-Before using the REML method, we clean the GNSS datasets by removing offsets at times of equipment changes and removing outliers with the method describe in Section X.
-For the outlier detection algorithm, our prior model, $u$, is chosen to be have a length-scale and time-scale which is sufficiently short to describe SSE displacements.
+First, we again clean the GNSS datasets by removing offsets at times of equipment changes and removing outliers with the method describe in Section X.
+For the outlier detection algorithm, our prior model, $u$, is chosen to have a length-scale and time-scale which is able to approximately describe SSE displacements.
 We use the Squared Exponential covariance function for $X$ with $\beta = 100$ km, and we use the Wendland covariance function for $T$, due to its computational efficiency, with $\beta = 0.1$ yr$^{-1}$.
 The product of the two covariance functions is then scaled by $\alpha^2 = 1$ mm$^{2}$.
-The outlier detection algorithm is particularly effective at removing outliers for stations at high elevation, which can be adversely affected by ice or snow \citep{Lisowski2008}.
+The outlier detection algorithm is particularly effective at removing outliers for stations at high elevation, which can be adversely affected by ice or snow during the winter \citep{Lisowski2008}.
 As an example, Figure \ref{fig:Outliers} shows the outliers which were automatically detected for station SC03, which is located on Mount Olympus in Washington.
-Note that most of the detected outliers occur during the winter and can likely be attributed to ice or snow.
-After cleaning the dataset, we divide it into four month long subsets, where each subset is roughly centered on the time of an SSE.
-We identify the times of seven SSEs using tremor records from \cite{Wech2010}.
-For each data subset, we use the REML method to find the hyperparameters for $T$ and $X$ that best describe deformation from the SSE.
-Our motivation analyzing subsets of the data is two-fold: (1) it keeps the analysis computationally tractable and (2) the distribution of estimated hyperparameters will indicates whether we can use a single covariance model to describe all SSEs.
-Our hyperparameter estimates are summarized in Table X.
+After cleaning the data, we divide it into subsets which are four months long and centered on the time of an SSE.
+The times of seven SSEs are determined using tremor records from \cite{Wech2010}.
+We find the optimal hyperparameters for $T$ and $X$ for each subset of data.
+The REML method is once again used to determine the hyperparameters for $T$ and $X$ which maximize the probability of drawing $\bm{d}_*$ from $\bm{d}$, where $\bm{d}_*$ now refers to a data subset.
+We chose to make each data subsets four months long because it is long enough to encompass the SSE, and it is short enough to still be computationally tractable.
+However, four months is too short to resolve the sinusoids in $\bm{d}$, and they are omitted from $\bm{d}$ for the REML analysis.
+The estimated hyperparameters for $u$ are summarized in Table X.
+When using SE or the Wendland function for $T$, the optimal hyperparameters appear have a unimodal distribution.
+When using IBM for $T$ there were several anomalously large inferences, hence the larger IQR.
 
@@
 \end{figure*}
+
+\begin{table}\label{tab:MaterialProperties}
+\begin{tabular} {l l l l l l}
+cov. function & direction & $\lambda$  & $\phi$   & $\theta$  & diff. REML \\ \hline
+SE-SE & east   & 92 $\pm$ 25 km  & 0.62 $\pm$ 0.11 mm  & 0.026 $\pm$ 0.011 yr  &  - \\
+SE-SE & north  & 91 $\pm$ 53 km  & 0.43 $\pm$ 0.05 mm  & 0.030 $\pm$ 0.017 yr  &  - \\
+SE-WEN & east   & 95 $\pm$ 30 km  & 0.66 $\pm$ 0.15 mm  & 0.093 $\pm$ 0.044 yr &  0.78 $\pm$ 0.87 \\
+SE-WEN & north  & 92 $\pm$ 57 km  & 0.46 $\pm$ 0.10 mm  & 0.116 $\pm$ 0.057 yr &  0.08 $\pm$ 0.58 \\
+SE-IBM & east   & 110 $\pm$ 130 km & 290 $\pm$ 420 mm/yr$^{0.5}$  & -          & -16.4 $\pm$ 7.8 \\
+SE-IBM & north  & 150 $\pm$ 560 km & 110 $\pm$ 250 mm/yr$^{0.5}$ & -           & -10.1 $\pm$ 2.3 \\
+\end{tabular}
+\caption{Hyperparameters. The shown values are the medians and interquartile ranges.} 
+\end{table}
 
@@
 \subsection{Transient Strain Rates} 
-Having established noise model and a prior for transient displacements, we use cleaned GNSS dataset described in the previous section to calculate transient strain rates in the Puget Sound region from eq.
+Having established noise model and a prior for transient displacements, we use cleaned GNSS dataset described in Section \ref{sec:SignalModel} to calculate transient strain rates in the Puget Sound region from eq.
 (\ref{eq:StrainRate}).

### 2017-06-07 figures/noise/make-plot.py

@@ -2,12 +2,12 @@
 import numpy as np
 from pygeons.io.parser import _get_field
 from pygeons.main.gpstation import fogm
-from pygeons.basemap import make_basemap
-from pygeons.plot.plot import _setup_map_ax
 import os
 import matplotlib.pyplot as plt
-from mpl_toolkits.basemap import Basemap
 np.random.seed(0)
+
+def iqr(x):
+  return np.subtract(*np.percentile(x,[75.0,25.0]))
 
 def get_meridians_and_parallels(bm,ticks):
   ''' 
@@ -36,6 +36,7 @@ params1 = {'east':[],'north':[],'vertical':[]}
 lons = []
 lats = []
 for f in files:
+  print(f)
   buff = open('results/%s' % f,'r')
   content = buff.read()
   buff.close()
@@ -64,53 +65,53 @@ stats_box = {'facecolor':'w','alpha':0.8,'boxstyle':'round','edgecolor':'0.8'}
 fig,axs = plt.subplots(3,2,figsize=(7,6))
 
 axs[0][0].hist(east_p0,np.linspace(0.0,3.0,bins),color='C0',alpha=0.8)
-text = 'median : %.2f\nIQR : %.2f - %.2f' % ((np.median(east_p0),) + tuple(np.percentile(east_p0,[25,75])))
+text = 'median : %.2f\nIQR : %.2f' % (np.median(east_p0),iqr(east_p0))
 axs[0][0].text(0.55,0.68,text,transform=axs[0][0].transAxes,bbox=stats_box,fontsize=10)
 axs[0][0].grid(c='0.5',alpha=0.5)
 axs[0][0].set_ylabel('count',fontsize=10)
-axs[0][0].set_xlabel(r'east $\alpha$ [mm/yr^0.5]',fontsize=10)
+axs[0][0].set_xlabel(r'east $\mathregular{\alpha}$ [mm/yr$^{0.5}$]',fontsize=10)
 axs[0][0].tick_params(labelsize=10)
 
 
 axs[1][0].hist(north_p0,np.linspace(0.0,3.0,bins),color='C0',alpha=0.8)
-text = 'median : %.2f\nIQR : %.2f - %.2f' % ((np.median(north_p0),) + tuple(np.percentile(north_p0,[25,75])))
[…]

### 2017-06-07 HinesHetlandGPR.tex
did more

@@
 We use a set of 38 inland stations, shown in Figure X, to constrain the FOGM hyperparameters, $\alpha$ and $\beta$.
-These stations are sufficiently far from the subduction zone that they are unlikely to contain transient signal, and thus we assume that $u$ is zero.
+These stations are sufficiently far from the subduction zone that they are unlikely to contain transient signal.
 We clean the data for these stations by removing jumps at times of equipment changes, and we remove outliers that have been detected with the algorithm described in Section \ref{sec:Outlier}.
 We then find $\alpha$ and $\beta$ for each station timeseries with the Restricted Maximum Likelihood (REML) method \cite[e.g.,][]{Harville1974,Cressie1992} and Hines 2017.
-The REML method finds the hyperparameters which maximize the probability of drawing $\bm{d}_*$ from $\bm{d}$.
+The REML method finds the hyperparameters which maximize the expression
+\begin{equation}\label{eq:REML}
+\left(\frac{\left|\bm{P}^T\bm{P}\right|}
+           {(2\pi)^{n-m}
+            \left| \bm{\Sigma} \right| 
+            \left| \bm{P}^T\bm{\Sigma}\bm{P} \right|}\right)^{\frac{1}{2}} 
+e^{-\tfrac{1}{2}\mathbf{d}_*^T\mathbf{K}\mathbf{d}_*},
+\end{equation}
+where
+\begin{equation}
+\bm{\Sigma} = C_u(\bm{y},\bm{y}) + C_\epsilon(\bm{y},\bm{y}),
+\end{equation}
+\begin{equation}
+\bm{K} = \bm{\Sigma}^{-1} - 
+         \bm{\Sigma}^{-1}\bm{P}
+         \left(\bm{P}^T\bm{\Sigma}^{-1}\bm{P}\right)^{-1}
+         \bm{P}^T\bm{\Sigma}^{-1},
+\end{equation}
+and $m$ is the number of columns in $\bm{P}$.
+For this section, $\bm{d}_*$ is used to denote the displacement time series for just one station, and thus $m = 6$.
+We are also assuming $C_u$ is zero when estimating the noise hyperparameters for this section.
+\citet{Harville1974} showed that choosing the hyperparameters which maximize eq.
+(\ref{eq:REML}) is equivalent to chosing the hyperparameters which maximize the probability of drawing $\bm{d}_*$ from $\bm{d}$.
 We use the REML method over the maximum likelihood (ML) method \citep[e.g.,][]{Langbein1997} because the REML method properly accounts for the improper prior that we assigned to $\bm{a}$.
@@
 \begin{tabular} {l l l l l l}
-cov. function & direction & $\lambda$  & $\phi$   & $\theta$  & diff. REML \\ \hline
-SE-SE & east   & 92 $\pm$ 25 km  & 0.62 $\pm$ 0.11 mm  & 0.026 $\pm$ 0.011 yr  &  - \\
-SE-SE & north  & 91 $\pm$ 53 km  & 0.43 $\pm$ 0.05 mm  & 0.030 $\pm$ 0.017 yr  &  - \\
-SE-WEN & east   & 95 $\pm$ 30 km  & 0.66 $\pm$ 0.15 mm  & 0.093 $\pm$ 0.044 yr &  0.78 $\pm$ 0.87 \\
-SE-WEN & north  & 92 $\pm$ 57 km  & 0.46 $\pm$ 0.10 mm  & 0.116 $\pm$ 0.057 yr &  0.08 $\pm$ 0.58 \\
-SE-IBM & east   & 110 $\pm$ 130 km & 290 $\pm$ 420 mm/yr$^{0.5}$  & -          & -16.4 $\pm$ 7.8 \\
-SE-IBM & north  & 150 $\pm$ 560 km & 110 $\pm$ 250 mm/yr$^{0.5}$ & -           & -10.1 $\pm$ 2.3 \\
+$T$ & direction & $\lambda$  & $\phi$   & $\theta$  & diff. REML \\ \hline
+SE & east   & 92 $\pm$ 25 km  & 0.62 $\pm$ 0.11 mm  & 0.026 $\pm$ 0.011 yr  &  - \\
+SE & north  & 91 $\pm$ 53 km  & 0.43 $\pm$ 0.05 mm  & 0.030 $\pm$ 0.017 yr  &  - \\
+Wendland & east   & 95 $\pm$ 30 km  & 0.66 $\pm$ 0.15 mm  & 0.093 $\pm$ 0.044 yr &  0.78 $\pm$ 0.87 \\
+Wendland & north  & 92 $\pm$ 57 km  & 0.46 $\pm$ 0.10 mm  & 0.116 $\pm$ 0.057 yr &  0.08 $\pm$ 0.58 \\
+IBM & east   & 110 $\pm$ 130 km & 290 $\pm$ 420 mm/yr$^{0.5}$  & -          & -16.4 $\pm$ 7.8 \\
+IBM & north  & 150 $\pm$ 560 km & 110 $\pm$ 250 mm/yr$^{0.5}$ & -           & -10.1 $\pm$ 2.3 \\
 \end{tabular}

### 2017-06-07 figures/signal_reml/make_plot.py

@@ -96,6 +96,10 @@ vertical_p2 = params2['vertical']
 east_like = np.array(like['east']) - np.array(sese_like['east'])
 north_like = np.array(like['north']) - np.array(sese_like['north'])
 
+plt.hist(east_like,50)
+#plt.hist(north_like)
+plt.show()
+
 msg += 'SE-WEN & east   & %.5g $\pm$ %.5g km  & %.5g $\pm$ %.5g mm  & %.5g $\pm$ %.5g yr & %.5g $\pm$ %.5g \\\\\n' % (
                                                                                       np.median(east_p2),iqr(east_p2),
                                                                                       np.median(east_p0),iqr(east_p0),

### 2017-06-08 HinesHetlandGPR.tex
did more

@@
 We constrain $u$ with GNSS data, which records $u$ as well as other physical and non-physical processes which we are not interested in.
-We describe GNSS observations at position $x_i$ and time $t$ as a realization of the random variable
+We describe GNSS observations at position $x_i$ and time $t_j$ as a realization of the random variable
 \begin{align}\label{eq:Data}
 \begin{split}
-d_{i}(t) = &u(x_i,t) + \epsilon(x_i,t) + a_{i} + b_{i}t + \\
-         &c_{i}\sin(2 \pi t) + q_{i}\cos(2 \pi t) + r_{i}\sin(4 \pi t) + s_{i}\cos(4 \pi t), 
+d_{ij} = &u(x_i,t_j) + \epsilon(x_i,t_j) + w_{ij} + a^{(1)}_i + a^{(2)}_it_j + \\
+         &a^{(3)}_i\sin(2 \pi t_j) + a^{(4)}_i\cos(2 \pi t_j) + a^{(5)}_i\sin(4 \pi t_j) + a^{(6)}_i\cos(4 \pi t_j), 
 \end{split}
 \end{align}
-where $a_{i}$ is an offset that is unique for each GNSS monument, $b_{i}$ is the steady rate of tectonic deformation at $x_i$, the sinusoids describe seasonal deformation (using units of years for $t$), and $\epsilon$ is a Gaussian process for noise that cannot be described parametrically.
-For example, $\epsilon$ can consist of white noise, temporally correlated noise describing benchmark wobble \citep[e.g.,][]{Wyatt1982,Wyatt1989}, and/or spatially correlated noise describing common mode error \citep[e.g.,][]{Wdowinski1997}.
-For now, $\epsilon$ will simply be described as $\epsilon \sim \mathcal{N}(0,C_\epsilon)$.
+where $a^{(1)}_{i}$ is an offset that is unique for each GNSS monument, $a^{(2)}_{i}$ is the steady rate of tectonic deformation at $x_i$, and the sinusoids describe seasonal deformation (using units of years for $t$).
+We use $w_{ij}$ to denote normally distributed, uncorrelated noise.
+Correlated noise which does not have a parametric representation is denoted by $\epsilon$.
+For example, $\epsilon$ can consist of temporally correlated noise describing benchmark wobble \citep[e.g.,][]{Wyatt1982,Wyatt1989}, and/or spatially correlated noise describing common mode error \citep[e.g.,][]{Wdowinski1997}.
+For now, we will only assume that $\epsilon \sim \mathcal{N}(0,C_\epsilon)$.
 We consider the six coefficients in eq.
 (\ref{eq:Data}) to be uncorrelated random variables distributed as $\mathcal{N}(0,\kappa^2)$ in the limit as $\kappa \to \infty$ (i.e., the coefficients have diffuse priors).
-Of course, the tectonic deformation, $b_{i}$, is spatially correlated and we could invoke a tectonic model to form a prior on $b_{i}$.
-However, in our application to Cascadia, we will be using displacement time series which are long enough to sufficiently constrain $b_{i}$ for each station, and there is no need to incorporate a prior.
+Of course, the tectonic deformation, $a^{(2)}_{i}$, is spatially correlated and we could invoke a tectonic model to form a prior on $a^{(2)}_{i}$.
+However, in our application to Cascadia, we will be using displacement time series which are long enough to sufficiently constrain $a^{(2)}_{i}$ for each station, and there is no need to incorporate a prior.
 Likewise, the seasonal coefficients may be spatially correlated \citep{Langbein2008}, and it may be worth exploring and exploiting such a correlation in a future study.
@@
 We now consider the column vector of $n$ GNSS observations, $\bm{d}_*$.
-Let $\bm{y}$ be the set of $(x_i, t)$ pairs describing where and when each of the GNSS observations have been made.
-Furthermore, let $\bm{a}$ be the vector of coefficients from eq.
-(\ref{eq:Data}) for each of the $m$ GNSS monuments, and let $\bm{P}$ be the $n \times 6m$ matrix of corresponding basis functions evaluated at each point in $\bm{y}$.
-The observations $\bm{d}_*$ are then a realization of the random vector
+Let $\bm{y}$ be the set of $(x_i, t_j)$ pairs describing where and when each of the GNSS observations have been made.
+Let $\bm{a}$ be the vector of coefficients from eq.
+(\ref{eq:Data}) for each of the $m$ GNSS stations.
+We use $\bm{P}$ to represent the $n \times 6m$ matrix of corresponding basis functions evaluated at each point in $\bm{y}$.
+We  also denote the vector of uncorrelated noise for each observation as $\bm{w}$, whose standard deviations are given by the formal data uncertainty $\bm{\sigma}$.
+The observations $\bm{d}_*$ can then be viewed as a realization of the random vector
 \begin{equation}
-\bm{d} = u(\bm{y}) + \epsilon(\bm{y}) + \bm{P}\bm{a},
+\bm{d} = u(\bm{y}) + \epsilon(\bm{y}) + \bm{w} + \bm{P}\bm{a},
 \end{equation}
-which is distributed as $\mathcal{N}(\bm{0},\bm{C_d})$, where
+which is distributed as $\mathcal{N}(\bm{0},\bm{\Sigma} + \kappa^2\bm{P}\bm{P}^T)$, where
 \begin{equation}\label{eq:Cd}
-\bm{C_d} = C_u(\bm{y},\bm{y}) + C_\epsilon(\bm{y},\bm{y}) + 
-           \kappa^2\bm{P}\bm{P}^T.
+\bm{\Sigma} = C_u(\bm{y},\bm{y}) + C_\epsilon(\bm{y},\bm{y}) + 
+              \mathrm{diag}\left(\bm{\sigma}^2\right).  
 \end{equation}
@@
 \begin{equation}\label{eq:PosteriorMean}
-\mu_{\hat{u}}(z) = C_u(z,\bm{y})\bm{C_d}^{-1}\bm{d}_*
+\mu_{\hat{u}}(z) = C_u(z,\bm{y})\left(\bm{\Sigma} + \kappa^2\bm{P}\bm{P}^T\right)^{-1}\bm{d}_*
 \end{equation}    
@@
 \begin{equation}\label{eq:PosteriorCov}
-C_{\hat{u}}(z,z') = C_u(z,\bm{y})\bm{C_d}^{-1}C_u(\bm{y},z').
+C_{\hat{u}}(z,z') = C_u(z,z') - C_u(z,\bm{y})\left(\bm{\Sigma} + \kappa^2\bm{P}\bm{P}^T\right)^{-1}C_u(\bm{y},z').
 \end{equation}
@@
 \begin{array}{cc}
-C_u(\bm{y},\bm{y}) + C_\epsilon(\bm{y},\bm{y}) & \bm{P} \\
+\bm{\Sigma} & \bm{P} \\
 \bm{P}^T  & -\kappa^{-2} \bm{I} \\
@@
 \begin{array}{cc}
-C_u(\bm{y},\bm{y}) + C_\epsilon(\bm{y},\bm{y}) & \bm{P} \\
+\bm{\Sigma} & \bm{P} \\
 \bm{P}^T  & \bm{0} \\
@@
 We calculate the residuals between the observations and a best fitting model, and data with residuals that are anomalously large are identified as outliers.
-We again treat $\bm{d}_*$ as a sample of $\bm{d}$.
+We treat $\bm{d}_*$ as a sample of $\bm{d}$ and assume that there is no temporally correlated noise (i.e., $\epsilon = 0$).
 The best fitting model for $\bm{d}_*$ is considered to be the expected value of the random vector $u(\bm{y}) + \bm{P}\bm{a}$ after conditioning it with non-outlier observations.
-We let $\epsilon$ be uncorrelated noise which has standard deviation given by the formal data uncertainties, $\bm{\sigma}$.
 We still consider $u$ to have a separable covariance function as in eq.
@@
 In this section we discuss our choice for the noise covariance function $C_\epsilon$ which will be used when estimating transient strain.
-We will assume that the noise consists of an uncorrelated component with standard deviation given by the formal data uncertainties, $\bm{\sigma}$, and a spatially and/or temporally correlated component.
 There have been numerous studies on temporally correlated noise in GNSS data \citep[e.g.,][]{Zhang1997,Mao1999,Williams2004,Langbein2008}.
@@
 There is some physical justification for using Brownian motion as a noise model because it accurately describes the power spectrum of motion resulting from instability in geodetic monuments \citep[e.g.,][]{Wyatt1982,Wyatt1989,Langbein1997}.
-We describe temporally correlated noise as a FOGM process, noting that Brownian motion is a degenerate case of a FOGM process.
+Here we describe the time dependence of $\epsilon$ as a FOGM process and consider $\epsilon$ to be spatially uncorrelated.
 A FOGM process is a solution to the stochastic differential equation
 \begin{equation}
-\dot{v}(t) + \alpha v(t) = \beta w(t),
+\dot{\epsilon}(t) + \alpha \epsilon(t) = \beta w(t),
 \end{equation}
-where $w(t)$ is white noise with unit variance. The solution is an isotropic Gaussian process with zero mean and covariance function
+where $w(t)$ is white noise with unit variance. The noise model is then a Gaussian process with zero mean and covariance function
 \begin{equation}\label{eq:FOGM}
-C_v(t,t') = \frac{\beta^2}{2\alpha}\exp\left(-\alpha|t - t'|\right). 
+C_\epsilon\left((x,t),(x',t')\right) = \frac{\beta^2}{2\alpha}\exp\left(-\alpha|t - t'|\right) \delta(||x - x'||_2). 
 \end{equation}
-A FOGM process degenerates to Brownian motion under the condition that $\alpha = 0$ and $v(0) = 0$.
+Under the condition that $\alpha = 0$ and $v(0) = 0$, $\epsilon$ degenerates to the more commonly used Brownian motion noise model.
 
-We use a set of 38 inland stations, shown in Figure X, to constrain the FOGM hyperparameters, $\alpha$ and $\beta$.
+We constrain the hyperparameters for $\epsilon$, $\alpha$ and $\beta$, with a set of 38 inland stations, which are shown in Figure X.
 These stations are sufficiently far from the subduction zone that they are unlikely to contain transient signal.
@@
 \left(\frac{\left|\bm{P}^T\bm{P}\right|}
-           {(2\pi)^{n-m}
+           {(2\pi)^{n-6m}
             \left| \bm{\Sigma} \right| 
@@
 \begin{equation}
-\bm{\Sigma} = C_u(\bm{y},\bm{y}) + C_\epsilon(\bm{y},\bm{y}),
-\end{equation}
-\begin{equation}
 \bm{K} = \bm{\Sigma}^{-1} - 
@@
          \left(\bm{P}^T\bm{\Sigma}^{-1}\bm{P}\right)^{-1}
-         \bm{P}^T\bm{\Sigma}^{-1},
+         \bm{P}^T\bm{\Sigma}^{-1}.
 \end{equation}
-and $m$ is the number of columns in $\bm{P}$.
-For this section, $\bm{d}_*$ is used to denote the displacement time series for just one station, and thus $m = 6$.
+For this section, $\bm{d}_*$ is used to denote the displacement time series for just one station, and thus $m = 1$.
 We are also assuming $C_u$ is zero when estimating the noise hyperparameters for this section.
@@
 
-For our estimation of of transient strain resulting from SSEs, we then simply assume that $\epsilon$ consists of only uncorrelated noise with covariance $C_\epsilon(\bm{y},\bm{y}') = \mathrm{diag}(\sigma^2)$.
+For these reasons, we henceforth make the simplifying assumption that temporally and spatially correlated noise is negligible and $\epsilon = 0$ for the easting and northing component of GNSS data.
 
@@
 IBM has been used, in the context of Kalman filtering, as a non-parametric model for the time dependence of geophysical signals \citep[e.g.,][]{Segall1997,McGuire2003,Ohtani2010,Hines2016}.
-It should be noted that IBM has a reference time at which the Gaussian process is exactly zero.
-For some geophysical signals, such as postseismic deformation, it is appropriate to have this reference time.
-However, if we are interesting in detecting transient events, where there is no known start time, then the IBM model may not be appropriate and an isotropic Gaussian process should be preferred.
-Alternatively we can assume that the start date for the IBM process is long before the first data obvservation; however this can introduce numerical difficulties because the covariance matrix will become ill-conditioned.
-Third, we consider the Wendland class of covariance functions \citep{Wendland2005}.
+It should be emphasized $t=0$ is a reference time at which the Gaussian process is exactly zero.
+For some geophysical signals, it is appropriate to have this reference time.
+For example, if we are trying to identify postseismic deformation then $t$ should be zero at the time of the earthquake.
+However, if we are interesting in detecting transient events, where there is no known start time, then IBM may not be an appropriate prior and an isotropic Gaussian process should be preferred.
+In the following analysis, we make the quite arbitrary choice that $t$ is zero on the first epoch of $\bm{d}_*$.
+Our third option for $T$ is the Wendland class of covariance functions \citep{Wendland2005}.
 Wendland functions are positive definite in $\mathbb{R}^d$, and they describes an isotropic Gaussian process with realizations that can be differentiated $k$ times.
@@
 We next want to determine appropriate hyperparameters for $X$ and each of the three candidate functions for $T$.
-First, we again clean the GNSS datasets by removing offsets at times of equipment changes and removing outliers with the method describe in Section X.
+First, we again clean the GNSS datasets by removing offsets at times of equipment changes and removing outliers with the method describe in Section \ref{sec:Outlier}.
 For the outlier detection algorithm, our prior model, $u$, is chosen to have a length-scale and time-scale which is able to approximately describe SSE displacements.
-We use the Squared Exponential covariance function for $X$ with $\beta = 100$ km, and we use the Wendland covariance function for $T$, due to its computational efficiency, with $\beta = 0.1$ yr$^{-1}$.
-The product of the two covariance functions is then scaled by $\alpha^2 = 1$ mm$^{2}$.
+We use the SE covariance function for $X$ with $\lambda = 100$ km, and we use the Wendland covariance function for $T$, due to its computational efficiency, with $\theta = 0.1$ yr and $\phi = 1$ mm.
 The outlier detection algorithm is particularly effective at removing outliers for stations at high elevation, which can be adversely affected by ice or snow during the winter \citep{Lisowski2008}.
@@
 However, four months is too short to resolve the sinusoids in $\bm{d}$, and they are omitted from $\bm{d}$ for the REML analysis.
-The estimated hyperparameters for $u$ are summarized in Table X.
+The estimated hyperparameters for $u$ are summarized in Table 1.
 When using SE or the Wendland function for $T$, the optimal hyperparameters appear have a unimodal distribution.
@@
 
-\begin{table}\label{tab:MaterialProperties}
+\begin{table}\label{tab:Parameters}
 \begin{tabular} {l l l l l l}

### 2017-06-08 figures/noise/make-plot.py

@@ -36,7 +36,6 @@ params1 = {'east':[],'north':[],'vertical':[]}
 lons = []
 lats = []
 for f in files:
-  print(f)
   buff = open('results/%s' % f,'r')
   content = buff.read()
   buff.close()
@@ -64,7 +63,7 @@ stats_box = {'facecolor':'w','alpha':0.8,'boxstyle':'round','edgecolor':'0.8'}
 
 fig,axs = plt.subplots(3,2,figsize=(7,6))
 
-axs[0][0].hist(east_p0,np.linspace(0.0,3.0,bins),color='C0',alpha=0.8)
+axs[0][0].hist(east_p0,np.linspace(0.0,2.0,bins),color='C0')
 text = 'median : %.2f\nIQR : %.2f' % (np.median(east_p0),iqr(east_p0))
 axs[0][0].text(0.55,0.68,text,transform=axs[0][0].transAxes,bbox=stats_box,fontsize=10)
 axs[0][0].grid(c='0.5',alpha=0.5)
@@ -73,7 +72,7 @@ axs[0][0].set_xlabel(r'east $\mathregular{\alpha}$ [mm/yr$^{0.5}$]',fontsize=10)
 axs[0][0].tick_params(labelsize=10)
 
 
-axs[1][0].hist(north_p0,np.linspace(0.0,3.0,bins),color='C0',alpha=0.8)
+axs[1][0].hist(north_p0,np.linspace(0.0,2.0,bins),color='C0')
 text = 'median : %.2f\nIQR : %.2f' % (np.median(north_p0),iqr(north_p0))
 axs[1][0].text(0.55,0.68,text,transform=axs[1][0].transAxes,bbox=stats_box,fontsize=10)
 axs[1][0].grid(c='0.5',alpha=0.5)
@@ -81,7 +80,7 @@ axs[1][0].set_ylabel('count',fontsize=10)
 axs[1][0].set_xlabel(r'north $\mathregular{\alpha}$ [mm/yr$^{0.5}$]',fontsize=10)
 axs[1][0].tick_params(labelsize=10)
 
-axs[2][0].hist(vertical_p0,np.linspace(0.0,50.0,bins),color='C0',alpha=0.8)
+axs[2][0].hist(vertical_p0,np.linspace(0.0,50.0,bins),color='C0')
 text = 'median : %.1f\nIQR : %.1f' % (np.median(vertical_p0),iqr(vertical_p0))
[…]

### 2017-06-09 HinesHetlandGPR.tex
started adding fit stuff

@@
             \left| \bm{P}^T\bm{\Sigma}\bm{P} \right|}\right)^{\frac{1}{2}} 
-e^{-\tfrac{1}{2}\mathbf{d}_*^T\mathbf{K}\mathbf{d}_*},
+e^{-\tfrac{1}{2}\bm{d}_*^T\bm{K}\bm{d}_*},
 \end{equation}
@@
 Our third option for $T$ is the Wendland class of covariance functions \citep{Wendland2005}.
+Wendland covariance functions have compact support and hence their corresponding covariance matrices will be sparse.
+In our analysis, we exploit this sparsity with the CHOLMOD software package \citep{Chen2008}.
 Wendland functions are positive definite in $\mathbb{R}^d$, and they describes an isotropic Gaussian process with realizations that can be differentiated $k$ times.
 The form of the covariance function depends on the choice of $d$ and $k$.
-Wendland covariance functions have compact support, which is a particularly useful property because we can exploit the sparsity of the the corresponding covariance matrices.
 To describe the temporal covariance of $u$, we only require that $d=1$ and $k=1$.
@@
 \begin{tabular} {l l l l l l}
-$T$ & direction & $\lambda$  & $\phi$   & $\theta$  & diff. REML \\ \hline
+$T$ & direction & $\lambda$  & $\phi$   & $\theta$  & diff. $\log$(REML) \\ \hline
 SE & east   & 92 $\pm$ 25 km  & 0.62 $\pm$ 0.11 mm  & 0.026 $\pm$ 0.011 yr  &  - \\
@@
 Wendland & north  & 92 $\pm$ 57 km  & 0.46 $\pm$ 0.10 mm  & 0.116 $\pm$ 0.057 yr &  0.08 $\pm$ 0.58 \\
-IBM & east   & 110 $\pm$ 130 km & 290 $\pm$ 420 mm/yr$^{0.5}$  & -          & -16.4 $\pm$ 7.8 \\
-IBM & north  & 150 $\pm$ 560 km & 110 $\pm$ 250 mm/yr$^{0.5}$ & -           & -10.1 $\pm$ 2.3 \\
+IBM & east   & 110 $\pm$ 130 km & 290 $\pm$ 420 mm/yr$^{1.5}$  & -          & -16.4 $\pm$ 7.8 \\
+IBM & north  & 150 $\pm$ 560 km & 110 $\pm$ 250 mm/yr$^{1.5}$ & -           & -10.1 $\pm$ 2.3 \\
 \end{tabular}
@@
 
-We next want to identify which covariance function for $T$ best describes the SSEs.
-We could use the REML likelihoods to compare different models, similar to \citet{Langbein2004}.
-However, the maximum likelihoods do not vary by more than a factor of X between the models, which does not definitively identify the best model.
-We then compare the posterior estimates $\hat{u}$ for each model to the observations to see if any model is noticeably better at describing the data.
-We show $\hat{u}$ using the three different function for $T$ in Figure X.
-It is clear that neither model offer any distinct improvement over the others.
-For most stations $\hat{u}$ appears to accurately capture the deformation from SSEs; however, for some stations which are closer to Puget Sound such as X and X, the rate of deformation is underestimated by the best fitting model.
-We could have manually specified a prior.
-We choose the Wendland function to describe $T$ because evaluating $T$ will yield sparse covariance matrices.
-We utilize the sparsity of the covariance matrices with the sparse matrix library CHOLMOD \citep{Chen2008}, which significantly expands the amount of data which we are able to tractably analyze with the method described in Section X.
+Next we identify which covariance function for $T$ best describes the SSEs.
+One approach is to compare the REML likelihood for each covariance function, similar to the analysis in \citet{Langbein2004}.
+In Table 1, we summarize how the log REML likelihoods for the Wendland and IBM covariance functions compare to the SE covariance function.
+The tabulated values show the median and IQR differences in log REML likelihoods over each of the seven SSEs considered in this study.
+Negative values indicate that the observed data is less likely to have been sampled from the covariance function.
+Based on the differences in log REML likelihoods, the data is substantially more likely to come form the SE or Wendland covariance functions than from the IBM covariance function.
+The REML likelihoods do not definitively indicate whether the SE or Wendland covariance function is preferable.
+
+As another test to determine the most appropriate covariance function for $T$, we compare the observations to the predicted displacements for each covariance function.
+let $\hat{\bm{d}} = \left(u(\bm{y}) + \bm{P}\bm{a}\right)|\bm{d}_*$ be the data prediction vector.
+It can be shown that $\bm{d}_*$ is normally distributed with mean
+\begin{equation}\label{eq:DataPredMean}
+\bm{\mu}_{\hat{d}} =
+\left[ 
+\begin{array}{cc}
+C_u(\bm{y},\bm{y}) & \bm{P} \\
+\end{array}
+\right]
+\left[
+\begin{array}{cc}
+\bm{\Sigma} & \bm{P} \\
+\bm{P}^T  & \bm{0} \\
+\end{array}
+\right]^{-1}
+\left[
+\begin{array}{c}
+\bm{d}_* \\
+\bm{0} \\
+\end{array}
+\right]
+\end{equation}  
+and covariance
+\begin{equation}\label{eq:DataPredCov}
+\bm{C}_{\hat{\bm{d}}} = 
+C_u(\bm{y},\bm{y}) - 
+\left[ 
+\begin{array}{cc}
+C_u(\bm{y},\bm{y}) & \bm{P} \\
+\end{array}
+\right]
+\left[
+\begin{array}{cc}
+\bm{\Sigma} & \bm{P} \\
+\bm{P}^T  & \bm{0} \\
+\end{array}
+\right]^{-1}
+\left[
+\begin{array}{c}
+C_u(\bm{y},\bm{y}) \\
+\bm{P}^T \\
+\end{array}
+\right].
+\end{equation}
+We compute $\hat{\bm{d}}$ using SE, Wendland, and IBM covariance functions for $T$ and the median hyperparameters from Table 1.
+Figure 1 compares the easting component of $\bm{d}_*$ to $\hat{\bm{d}}$ for the Winter 2016 SSE.
+For most stations, the data prediction vector appears to appropriately describe displacements throughout the SSE, regardless of the choice of $T$.
+The prediction for the IBM model contains slightly more high frequency, and perhaps spurious, features.
+The predictions for the Wendland and SE covariance functions are nearly indistinguishable.
+We ultimately settle on the Wendland covariance function for $T$.
+We choose the Wendland covariance function over the SE covariance function because of its computational advantages noted above.
 

### 2017-06-10 figures/outliers/make_plot.py
added more

@@ -43,8 +43,7 @@ u = 1000*data['east'][:,idx]
 us = 1000*data['east_std_dev'][:,idx]
 m = trend(u,t)
 u = u - m[0] - m[1]*t
-ax.errorbar(t,u,us,marker='None',linestyle='None',color=c2t)
-ax.plot(t,u,'.',ms=4.99,color=c2)
+ax.errorbar(t,u,us,marker='.',linestyle='None',color=c2,ecolor=c2t,ms=4.99)
 
 data = pygeons.io.io.dict_from_hdf5('data.final.h5')
 idx = (data['id'] == 'SC03').nonzero()[0][0]
@@ -52,8 +51,7 @@ t = data['time']
 u = 1000*data['east'][:,idx]
 us = 1000*data['east_std_dev'][:,idx]
 u = u - m[0] - m[1]*t
-ax.errorbar(t,u,us,marker='None',linestyle='None',color=c1t)
-ax.plot(t,u,'.',ms=5.0,color=c1)
+ax.errorbar(t,u,us,marker='.',linestyle='None',color=c1,ecolor=c1t,ms=5.0)
 
 ax.set_title('Station SC03 (%.2f$^\mathregular{\circ}$ W, %.2f$^\mathregular{\circ}$ N)' % 
                  (-lon,lat),fontsize=10)

### 2017-06-10 figures/signal_fit/make-plot.py

@@ -1,0 +1,73 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import pygeons
+from matplotlib import colors
+
+xtick_dates = ['2015-11-01','2015-12-01','2016-01-01','2016-02-01','2016-03-01']
+xticks = [pygeons.mjd.mjd(i,'%Y-%m-%d') for i in xtick_dates]
+xtick_labels = ['2015-11-01','2015-12-01','2016-01-01','2016-02-01','2016-03-01']
+
+c0 = colors.to_rgb('C0')
+c0t = tuple(0.6 + 0.4*np.array(colors.to_rgb('C0')))
+c1 = colors.to_rgb('C1')
+c1t = tuple(0.6 + 0.4*np.array(colors.to_rgb('C1')))
+c2 = colors.to_rgb('C2')
+c2t = tuple(0.6 + 0.4*np.array(colors.to_rgb('C2')))
+c3 = colors.to_rgb('C3')
+c3t = tuple(0.6 + 0.4*np.array(colors.to_rgb('C3')))
+
+fig,ax = plt.subplots(figsize=(7,3.5))
+pygeons.plot.plot._setup_ts_ax([ax])
+ax.set_xticks(xticks)
+ax.set_xticklabels(xtick_labels)
+ax.tick_params(labelsize=10)
+ax.set_xlim((57327.0,57448.0))
+ax.set_ylabel('displacement [mm]')
+#ax.set_ylim((-10.0,15.0))
+ax.grid()
+
+# OBSERVED
+data = pygeons.io.io.dict_from_hdf5('data/data.h5')
+idx = (data['id'] == 'P436').nonzero()[0][0]
+lon = data['longitude'][idx]
+lat = data['latitude'][idx]
+t = data['time']
+u = 1000*data['east'][:,idx]
+us = 1000*data['east_std_dev'][:,idx]
+ax.errorbar(t,u,us,marker='.',linestyle='None',color=c0,ecolor=c0t,ms=5.0,zorder=1,label='observed')
+
+# SE
+data = pygeons.io.io.dict_from_hdf5('data/se.per.fit.h5')
+idx = (data['id'] == 'P436').nonzero()[0][0]
+lon = data['longitude'][idx]
+lat = data['latitude'][idx]
+t = data['time']
+u = 1000*data['east'][:,idx]
+us = 1000*data['east_std_dev'][:,idx]
[…]

### 2017-06-12 HinesHetlandGPR.tex
added more

@@
 let $\hat{\bm{d}} = \left(u(\bm{y}) + \bm{P}\bm{a}\right)|\bm{d}_*$ be the data prediction vector.
-It can be shown that $\bm{d}_*$ is normally distributed with mean
+It can be shown that $\hat{\bm{d}}$ is normally distributed with mean
 \begin{equation}\label{eq:DataPredMean}
@@
 
+\begin{figure*}
+\includegraphics{figures/signal_fit/signal-fit.pdf}
+\caption{Fit.}   
+\label{fig:Fit}
+\end{figure*}
+
 \subsection{Transient Strain Rates} 
@@
 
+\begin{figure*}
+\includegraphics{figures/strain_map/strain-map.pdf}
+\caption{Strain map.}   
+\label{fig:Fit}
+\end{figure*}
+
 To demonstrate the effectiveness of our method for detecting SSEs, we show a timeseries of estimated strain components at the position indicated in figure X.
@@
 We are able to accurately detect the seven SSEs in over this period in addition to the inter-SSE events such as the one at X.
+
+\begin{figure*}
+\includegraphics{figures/strain_ts/strain-ts.pdf}
+\caption{Strain time series.}   
+\label{fig:Fit}
+\end{figure*}
+
+\begin{figure*}
+\includegraphics{figures/strain_ts/mag-ts.pdf}
+\caption{Strain magnitude time series.}   
+\label{fig:Fit}
+\end{figure*}
+
+A video can be found as supp material
 

### 2017-06-12 figures/strain_ts/make-plot.py

@@ -1,0 +1,126 @@
+import numpy as np
+import pygeons
+import matplotlib.pyplot as plt
+import h5py
+from matplotlib import colors
+from mpl_toolkits.basemap import Basemap
+from pygeons.plot.plot import _setup_ts_ax
+import rbf
+import sympy
+
+# Parameters
+station = 'I0049'
+radius = 150000.0
+
+r,eps = rbf.basis.get_r(),rbf.basis.get_eps()
+expr = 1.0/sympy.sqrt((2*sympy.pi)*eps**2)*sympy.exp(-r**2/(2*eps**2))
+normal1d = rbf.basis.RBF(expr)
+
+bm = Basemap(projection='tmerc',
+          resolution='i',
+          lon_0 = -123.0,lat_0 = 45.0,
+          llcrnrlon = -127.5,llcrnrlat = 45.5,
+          urcrnrlon = -120.0,urcrnrlat = 50.0)
+                                        
+c0 = colors.to_rgb('C0')
+c0t = tuple(0.6 + 0.4*np.array(colors.to_rgb('C0')))
+c1 = colors.to_rgb('C1')
+c1t = tuple(0.6 + 0.4*np.array(colors.to_rgb('C1')))
+
+
+start_day = pygeons.mjd.mjd('2010-01-01','%Y-%m-%d')
+end_day = pygeons.mjd.mjd('2017-05-15','%Y-%m-%d')
+total_days = end_day - start_day
+xtick_dates = ['2010-01-01','2011-01-01','2012-01-01','2013-01-01','2014-01-01','2015-01-01','2016-01-01','2017-01-01']
+xticks = [pygeons.mjd.mjd(i,'%Y-%m-%d') for i in xtick_dates]
+xtick_labels = [' ','2011-01-01',' ','2013-01-01',' ','2015-01-01',' ','2017-01-01']
+
+fdx = h5py.File('merged.dudx.h5')
+fdy = h5py.File('merged.dudy.h5')
+time = fdx['time'][...]
+tidx = np.nonzero((time > start_day) & (time < end_day))[0]
+time = time[tidx]
+
+xidx = np.nonzero(fdx['id'][...] == station)[0][0]
+sta_lon = fdx['longitude'][xidx]
+sta_lat = fdx['latitude'][xidx]
[…]

### 2017-06-12 figures/context_map/make-plot.py
added context map

@@ -1,0 +1,127 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import pygeons
+from mpl_toolkits.basemap import Basemap
+import h5py
+from matplotlib.colors import ListedColormap
+from matplotlib.patches import Ellipse
+from matplotlib.cm import viridis
+from myplot.colorbar import pseudo_transparent_cmap,transparent_colorbar
+from sympy import sqrt,pi,exp
+import rbf
+
+r,eps = rbf.basis.get_r(),rbf.basis.get_eps()
+expr = 1.0/sqrt((2*pi)**2*eps**4)*exp(-r**2/(2*eps**2))
+normal2d = rbf.basis.RBF(expr)
+
+def get_meridians_and_parallels(bm,ticks):
+  ''' 
+  returns the meridians and parallels that should be plotted.
+  '''
+  diff_lon = (bm.urcrnrlon-bm.llcrnrlon)
+  round_digit = int(np.ceil(np.log10(ticks/diff_lon)))
+  dlon = np.round(diff_lon/ticks,round_digit)
+
+  diff_lat = (bm.urcrnrlat-bm.llcrnrlat)
+  round_digit = int(np.ceil(np.log10(ticks/diff_lat)))
+  dlat = np.round(diff_lat/ticks,round_digit)
+
+  meridians = np.arange(np.floor(bm.llcrnrlon),
+                        np.ceil(bm.urcrnrlon),dlon)
+  parallels = np.arange(np.floor(bm.llcrnrlat),
+                        np.ceil(bm.urcrnrlat),dlat)
+
+  return meridians,parallels
+
+def setup_ax(bm,ax):
+  # function which prints out the coordinates on the bottom left 
+  # corner of the figure
+  def coord_string(x,y):
+    str = 'x : %g  y : %g  ' % (x,y)
+    str += '(lon : %g E  lat : %g N)' % bm(x,y,inverse=True)
+    return str
+
+  ax.format_coord = coord_string
+  bm.drawcountries(ax=ax,zorder=1)
+  bm.drawstates(ax=ax,zorder=1)
+  bm.drawcoastlines(ax=ax,zorder=1)
+  bm.drawmapboundary(ax=ax,fill_color=(0.8,0.8,0.8),zorder=0)
[…]

### 2017-06-12 figures/noise/make-plot.py

@@ -63,58 +63,57 @@ stats_box = {'facecolor':'w','alpha':0.8,'boxstyle':'round','edgecolor':'0.8'}
 
 fig,axs = plt.subplots(3,2,figsize=(7,6))
 
-axs[0][0].hist(east_p0,np.linspace(0.0,2.0,bins),color='C0')
+axs[0][0].hist(east_p0,np.linspace(0.0,2.0,bins),color='C0',zorder=2)
 text = 'median : %.2f\nIQR : %.2f' % (np.median(east_p0),iqr(east_p0))
 axs[0][0].text(0.55,0.68,text,transform=axs[0][0].transAxes,bbox=stats_box,fontsize=10)
-axs[0][0].grid(c='0.5',alpha=0.5)
+axs[0][0].grid(zorder=0)
 axs[0][0].set_ylabel('count',fontsize=10)
 axs[0][0].set_xlabel(r'east $\mathregular{\alpha}$ [mm/yr$^{0.5}$]',fontsize=10)
 axs[0][0].tick_params(labelsize=10)
 
 
-axs[1][0].hist(north_p0,np.linspace(0.0,2.0,bins),color='C0')
+axs[1][0].hist(north_p0,np.linspace(0.0,2.0,bins),color='C0',zorder=2)
 text = 'median : %.2f\nIQR : %.2f' % (np.median(north_p0),iqr(north_p0))
 axs[1][0].text(0.55,0.68,text,transform=axs[1][0].transAxes,bbox=stats_box,fontsize=10)
-axs[1][0].grid(c='0.5',alpha=0.5)
+axs[1][0].grid(zorder=0)
 axs[1][0].set_ylabel('count',fontsize=10)
 axs[1][0].set_xlabel(r'north $\mathregular{\alpha}$ [mm/yr$^{0.5}$]',fontsize=10)
 axs[1][0].tick_params(labelsize=10)
 
-axs[2][0].hist(vertical_p0,np.linspace(0.0,50.0,bins),color='C0')
+axs[2][0].hist(vertical_p0,np.linspace(0.0,50.0,bins),color='C0',zorder=2)
 text = 'median : %.1f\nIQR : %.1f' % (np.median(vertical_p0),iqr(vertical_p0))
 axs[2][0].text(0.55,0.68,text,transform=axs[2][0].transAxes,bbox=stats_box,fontsize=10)
-axs[2][0].grid(c='0.5',alpha=0.5)
+axs[2][0].grid(zorder=0)
[…]

### 2017-06-12 figures/signal_fit/make-plot.py

@@ -9,12 +9,6 @@ xtick_labels = ['2015-11-01','2015-12-01','2016-01-01','2016-02-01','2016-03-01'
 
 c0 = colors.to_rgb('C0')
 c0t = tuple(0.6 + 0.4*np.array(colors.to_rgb('C0')))
-c1 = colors.to_rgb('C1')
-c1t = tuple(0.6 + 0.4*np.array(colors.to_rgb('C1')))
-c2 = colors.to_rgb('C2')
-c2t = tuple(0.6 + 0.4*np.array(colors.to_rgb('C2')))
-c3 = colors.to_rgb('C3')
-c3t = tuple(0.6 + 0.4*np.array(colors.to_rgb('C3')))
 
 fig,ax = plt.subplots(figsize=(7,3.5))
 pygeons.plot.plot._setup_ts_ax([ax])
@@ -44,8 +38,8 @@ lat = data['latitude'][idx]
 t = data['time']
 u = 1000*data['east'][:,idx]
 us = 1000*data['east_std_dev'][:,idx]
-ax.plot(t,u,'-',color=c1,zorder=2,lw=1.5,label='SE')
-ax.fill_between(t,u-us,u+us,color=c1,zorder=2,alpha=0.4)
+ax.plot(t,u,'-',color='C1',zorder=2,lw=1.5,label='SE')
+ax.fill_between(t,u-us,u+us,edgecolor='none',facecolor='C1',zorder=2,alpha=0.4)
 
 # WEN
 data = pygeons.io.io.dict_from_hdf5('data/wen12.per.fit.h5')
@@ -55,7 +49,7 @@ lat = data['latitude'][idx]
 t = data['time']
 u = 1000*data['east'][:,idx]
 us = 1000*data['east_std_dev'][:,idx]
-ax.plot(t,u,'-',color=c2,zorder=3,lw=1.5,label='Wendland')
+ax.plot(t,u,'-',color='C2',zorder=3,lw=1.5,label='Wendland')
 
 # IBM
 data = pygeons.io.io.dict_from_hdf5('data/ibm.per.fit.h5')
@@ -65,7 +59,7 @@ lat = data['latitude'][idx]
 t = data['time']
 u = 1000*data['east'][:,idx]
 us = 1000*data['east_std_dev'][:,idx]
-ax.plot(t,u,'-',color=c3,zorder=4,lw=1.5,label='IBM')
[…]

### 2017-06-12 figures/strain_ts/make-plot.py

@@ -2,7 +2,6 @@ import numpy as np
 import pygeons
 import matplotlib.pyplot as plt
 import h5py
-from matplotlib import colors
 from mpl_toolkits.basemap import Basemap
 from pygeons.plot.plot import _setup_ts_ax
 import rbf
@@ -22,12 +21,6 @@ bm = Basemap(projection='tmerc',
           llcrnrlon = -127.5,llcrnrlat = 45.5,
           urcrnrlon = -120.0,urcrnrlat = 50.0)
                                         
-c0 = colors.to_rgb('C0')
-c0t = tuple(0.6 + 0.4*np.array(colors.to_rgb('C0')))
-c1 = colors.to_rgb('C1')
-c1t = tuple(0.6 + 0.4*np.array(colors.to_rgb('C1')))
-
-
 start_day = pygeons.mjd.mjd('2010-01-01','%Y-%m-%d')
 end_day = pygeons.mjd.mjd('2017-05-15','%Y-%m-%d')
 total_days = end_day - start_day
@@ -45,6 +38,8 @@ xidx = np.nonzero(fdx['id'][...] == station)[0][0]
 sta_lon = fdx['longitude'][xidx]
 sta_lat = fdx['latitude'][xidx]
 sta_x,sta_y = bm(sta_lon,sta_lat)
+print(sta_lon)
+print(sta_lat)
 
 # convert to microstrain/year
 dudx = fdx['east'][tidx,xidx] * 1.0e6 * 365.25
@@ -69,17 +64,17 @@ mag_std = np.sqrt((exx_std*exx/mag)**2 + (eyy_std*eyy/mag)**2 + (exy_std*2*exy/m
 fig,axs = plt.subplots(3,1,sharex=True,figsize=(7,5))
 _setup_ts_ax(axs)
 axs[0].plot(time,exx,'C0-',lw=1.5,zorder=2)
-axs[0].fill_between(time,exx-exx_std,exx+exx_std,color='C0',alpha=0.4,zorder=2)
+axs[0].fill_between(time,exx-exx_std,exx+exx_std,edgecolor='none',facecolor='C0',alpha=0.4,zorder=2)
 axs[0].grid()
 axs[0].set_ylabel('east normal\n[$\mathregular{\mu}$strain/yr]',fontsize=10)
 
 axs[1].plot(time,eyy,'C0-',lw=1.5,zorder=2)
[…]

### 2017-06-13 HinesHetlandGPR.tex
added context map
added more

@@
 \section{Introduction}\label{sec:Introduction}
-
 Crustal strain rates are fundamentally important quantities for assessing seismic hazard, since knowing where and how quickly strain is accumulating gives insight into where we can expect stored elastic energy to be released seismically.
@@
 The strain rates presented in this paper are more direct observations and can be interpretted with a higher degree of confidence.
+
+The Southern California Earthquake Center (SCEC) community has recently shown interest in developing methods for detecting transient signals.
+SCEC supported a transient detection exercise, where several research groups tested their transient detection methods with a synthetic GNSS dataset.
+supported the transient Our motivation for detecting SSEs is, in part, because there have been several instances where SSEs immediately preceded megathrust earthquakes \citep{Roeloffs2006}.
 
@@
  \begin{equation}\label{eq:PosteriorMean2}
-\mu_{\hat{u}}(z) =
-\left[ 
-\begin{array}{cc}
+\mu_{\hat{u}}(z) = \left[\begin{array}{cc}
                          C_u(z,\bm{y}) & \bm{0} \\
-\end{array}
-\right]
-\left[
-\begin{array}{cc}
+                         \end{array}\right]
+                   \left[\begin{array}{cc}
                          \bm{\Sigma} & \bm{P} \\
                          \bm{P}^T  & -\kappa^{-2} \bm{I} \\
-\end{array}
-\right]^{-1}
-\left[
-\begin{array}{c}
+                         \end{array}\right]^{-1}
+                   \left[\begin{array}{c}
                          \bm{d}_* \\
                          \bm{0} \\
-\end{array}
-\right]
+                         \end{array}\right]
 \end{equation}    
@@
 \begin{equation}\label{eq:PosteriorCov2}
-C_{\hat{u}}(z,z') = 
-C_u(z,z') - 
-\left[ 
-\begin{array}{cc}
+C_{\hat{u}}(z,z') = C_u(z,z') - 
+                    \left[\begin{array}{cc}
                           C_u(z,\bm{y}) & \bm{0} \\
-\end{array}
-\right]
-\left[
-\begin{array}{cc}
+                          \end{array}\right]
+                    \left[\begin{array}{cc}
                           \bm{\Sigma} & \bm{P} \\
                           \bm{P}^T  & -\kappa^{-2} \bm{I} \\
-\end{array}
-\right]^{-1}
-\left[
-\begin{array}{c}
+                          \end{array}\right]^{-1}
+                    \left[\begin{array}{c}
                           C_u(\bm{y},z') \\
                           \bm{0} \\
-\end{array}
-\right].
+                          \end{array}\right].
 \end{equation}
@@
  \begin{equation}\label{eq:PosteriorMean3}
-\mu_{\hat{u}}(z) =
-\left[ 
-\begin{array}{cc}
+\mu_{\hat{u}}(z) = \left[\begin{array}{cc}
                          C_u(z,\bm{y}) & \bm{0} \\
-\end{array}
-\right]
-\left[
-\begin{array}{cc}
+                         \end{array}\right]
+                   \left[\begin{array}{cc}
                          \bm{\Sigma} & \bm{P} \\
                          \bm{P}^T  & \bm{0} \\
-\end{array}
-\right]^{-1}
-\left[
-\begin{array}{c}
+                         \end{array}\right]^{-1}
+                   \left[\begin{array}{c}
                          \bm{d}_* \\
                          \bm{0} \\
-\end{array}
-\right]
+                         \end{array}\right]
 \end{equation}    
@@
 \begin{equation}\label{eq:PosteriorCov3}
-C_{\hat{u}}(z,z') = 
-C_u(z,z') - 
-\left[ 
-\begin{array}{cc}
+C_{\hat{u}}(z,z') = C_u(z,z') - 
+                    \left[\begin{array}{cc}
                           C_u(z,\bm{y}) & \bm{0} \\
-\end{array}
-\right]
-\left[
-\begin{array}{cc}
+                          \end{array}\right]
+                    \left[\begin{array}{cc}
                           \bm{\Sigma} & \bm{P} \\
                           \bm{P}^T  & \bm{0} \\
-\end{array}
-\right]^{-1}
-\left[
-\begin{array}{c}
+                          \end{array}\right]^{-1}
+                    \left[\begin{array}{c}
                           C_u(\bm{y},z') \\
                           \bm{0} \\
-\end{array}
-\right].
+                          \end{array}\right].
 \end{equation}
@@
        &= \frac{1}{\bm{\sigma}}\left(\bm{d}_*  - 
-\left[ 
-\begin{array}{cc}
+          \left[\begin{array}{cc}
                 C_u(\bm{y},\tilde{\bm{y}}) & \bm{P} \\
-\end{array}
-\right]
-\left[
-\begin{array}{cc}
+                \end{array}\right]
+          \left[\begin{array}{cc}
                 C_u(\tilde{\bm{y}},\tilde{\bm{y}}) + \mathrm{diag}(\tilde{\bm{\sigma}}^2) & \tilde{\bm{P}} \\
                 \tilde{\bm{P}}^T  & \bm{0} \\
-\end{array}
-\right]^{-1}
-\left[
-\begin{array}{c}
+                \end{array}\right]^{-1}
+          \left[\begin{array}{c}
                 \tilde{\bm{d}}_* \\
                 \bm{0} \\
-\end{array}
-\right] 
-\right),
+                \end{array}\right]\right),
 \end{align}
@@
 In this section we estimate transient strain from GNSS data in Cascadia.
-We are mainly interested in identifying transient strain resulting from slow slip events (SSEs) \citep[e.g.,][]{Dragert2001}.
-We show that Gaussian processes are an effective tool for (1) detecting SSEs and (2) understanding how the strain from SSEs are contributing to the strain accumulated over the seismic cycle.
-Our motivation for detecting SSEs is, in part, because there have been several instances where SSEs immediately preceded megathrust earthquakes \citep{Roeloffs2006}.
+In particular, we are interested in identifying transient strain resulting from slow slip events (SSEs) \citep[e.g.,][]{Dragert2001}.
+Our results illuminate where elastic strain energy, which can eventually be released seismically, is accumulating during SSEs.
+The estimated strain rates also reveal how strain from SSEs contributes to strain accumulated over seismic cycles.
+We demonstrate that transient strain rates estimated with the method described above are an effective tool for detecting SSEs and potentially other geophysical signal in GNSS data.
 In Cascadia, SSEs can be detected by monitoring for associated seismic tremor, which is actively being done by the Pacific Northwest Seismic Network (PNSN) \citep{Wech2010}.
-Here we demonstrate that strain rates derived from GNSS data can also be an effective tool for detecting SSEs.
-Furthermore, estimated strain rates provide a clear picture of how SSEs transfer elastic strain energy through the crust, which is also important information for assessing seismic hazard.
+We verify that transient strain rates estimated with the method described above are indeed identifying strain from SSEs by comparing $\bm{\epsilon}$ to the tremor records.
 We envision that GNSS derived transient strain rates could be useful for monitoring subduction zones which have SSEs but no associated tremor \citep{Schwartz2007}.
 
-In this study, we use GNSS data which has been made publicly available by the University Navstar Consortium (UNAVCO) \citep{Herring2016} and can be found at www.unavco.org.
+We use publicly available continuous GNSS data from the University Navstar Consortium (UNAVCO) \citep{Herring2016}, which can be found at www.unavco.org.
 We limit the dataset to the stations and time ranges which are pertinent to seven of the most recent SSEs in the Puget Sound region; the earliest SSE began in the August 2010, and the most recent SSE began in the February of 2017.
 We consider these most recent SSEs because the station coverage is sufficiently dense for us to make well constrained inferences of optimal prior models.
-The distribution of stations considered in this station is shown in Figure X.
+The positions of GNSS stations used to constrain transient strain rates are shown in Figure \ref{fig:Context}.
+
+\begin{figure*}
+\includegraphics{figures/context_map/context-map.pdf}
+\caption{Positions of continuous GNSS stations used to estimate transient strain rates.
+The colored regions indicates the density of detected seismic tremor as determined by \citep{Wech2010}.
+The red dots show the positions of GNSS stations mentioned in this paper.
+The magenta dot indicates the location of the strain rates shown in Figure \ref{fig:StrainTs} and \ref{fig:StrainMag}.
+The magenta circle demarcates the spatial extent of the tremors shown in Figure \ref{fig:StrainMag}}.
+\label{fig:Context}
+\end{figure*}
 
 \subsection{Noise model}\label{sec:NoiseModel}
-In this section we discuss our choice for the noise covariance function $C_\epsilon$ which will be used when estimating transient strain.
+Before we determine the transient strain rates, we must establish a prior for the transient deformation, $u$, and the noise, $\epsilon$.
+In this section we discuss our choice for the noise covariance function $C_\epsilon$.
 There have been numerous studies on temporally correlated noise in GNSS data \citep[e.g.,][]{Zhang1997,Mao1999,Williams2004,Langbein2008}.
-In these studies, temporally correlated noise was described with some combination of Brownian motion, first-order Gauss-Markov (FOGM) processes, and/or flicker noise.
+In these studies, temporally correlated noise was described with some combination of Brownian motion (also referred to as a Weiner process or a random walk), a first-order Gauss-Markov (FOGM) process, and/or flicker noise.
 There is some physical justification for using Brownian motion as a noise model because it accurately describes the power spectrum of motion resulting from instability in geodetic monuments \citep[e.g.,][]{Wyatt1982,Wyatt1989,Langbein1997}.
@@
 A FOGM process is a solution to the stochastic differential equation
-\begin{equation}
+\begin{equation}\label{eq:FOGMdef}
 \dot{\epsilon}(t) + \alpha \epsilon(t) = \beta w(t),
 \end{equation}
-where $w(t)$ is white noise with unit variance. The noise model is then a Gaussian process with zero mean and covariance function
+where $w(t)$ is white noise with unit variance.
+The FOGM process degenerates to the commonly used Brownian motion noise model under the condition that $\alpha=0$ and $\epsilon(0) = 0$, which is evident from the definition of Brownian motion \citep[e.g.,][]{Papoulis1991}.
+Our noise model which satisfies eq.
+(\ref{eq:FOGMdef}) is a Gaussian process with zero mean and covariance function
 \begin{equation}\label{eq:FOGM}
@@
 \end{equation}
-Under the condition that $\alpha = 0$ and $v(0) = 0$, $\epsilon$ degenerates to the more commonly used Brownian motion noise model.
 
-We constrain the hyperparameters for $\epsilon$, $\alpha$ and $\beta$, with a set of 38 inland stations, which are shown in Figure X.
+We constrain the hyperparameters for $\epsilon$, $\alpha$ and $\beta$, with a set of 38 continuous GNSS stations in Washington and Oregon which are east of 121$^\circ$ W.
 These stations are sufficiently far from the subduction zone that they are unlikely to contain transient signal.
 We clean the data for these stations by removing jumps at times of equipment changes, and we remove outliers that have been detected with the algorithm described in Section \ref{sec:Outlier}.
-We then find $\alpha$ and $\beta$ for each station timeseries with the Restricted Maximum Likelihood (REML) method \cite[e.g.,][]{Harville1974,Cressie1992} and Hines 2017.
+We then find $\alpha$ and $\beta$ for each station timeseries with the Restricted Maximum Likelihood (REML) method \cite[e.g.,][]{Harville1974,Cressie1992}.
 The REML method finds the hyperparameters which maximize the expression
@@
 The estimated hyperparameters for $u$ are summarized in Table 1.
-When using SE or the Wendland function for $T$, the optimal hyperparameters appear have a unimodal distribution.
-When using IBM for $T$ there were several anomalously large inferences, hence the larger IQR.
+Based on the interquartile ranges (IQR), the hyperparameters for the SE and Wendland model do not vary significantly between SSEs.
+This suggests that the median of estimated hyperparameters should be an appropriate prior model for all Cascadia SSEs.
+For the IBM model, there are several anomalously large values for $\lambda$ and $\phi$, which is they they have large IQRs.
 
@@
 \end{tabular}
-\caption{Hyperparameters. The shown values are the medians and interquartile ranges.} 
+\caption{Optimal hyperparameters for transient displacements determined with the REML method.
+The temporal covariance function is indicated by the ``$T$'' column, and the spatial covariance function, $X$, is the squared exponential in all cases.
+The hyperparameters are estimated for each of the seven SSEs considered in this study, and the tabulated values indicate the median and interquartile ranges of estimates.
+The ``diff $\log$(REML)'' column compares the log REML likelihood to the log REML likelihood for the SE model. Positive values indicate that observations are more consistent with the SE model.} 
 \end{table}
@@
 \begin{equation}\label{eq:DataPredMean}
-\bm{\mu}_{\hat{d}} =
-\left[ 
-\begin{array}{cc}
+\bm{\mu}_{\hat{d}} = \left[\begin{array}{cc}
                            C_u(\bm{y},\bm{y}) & \bm{P} \\
-\end{array}
-\right]
-\left[
-\begin{array}{cc}
+                           \end{array}\right]
+                     \left[\begin{array}{cc}
                            \bm{\Sigma} & \bm{P} \\
                            \bm{P}^T  & \bm{0} \\
-\end{array}
-\right]^{-1}
-\left[
-\begin{array}{c}
+                           \end{array}\right]^{-1}
+                     \left[\begin{array}{c}
                            \bm{d}_* \\
                            \bm{0} \\
-\end{array}
-\right]
+                           \end{array}\right]
 \end{equation}  
@@
 \begin{equation}\label{eq:DataPredCov}
-\bm{C}_{\hat{\bm{d}}} = 
-C_u(\bm{y},\bm{y}) - 
-\left[ 
-\begin{array}{cc}
+\bm{C}_{\hat{\bm{d}}} = C_u(\bm{y},\bm{y}) - 
+                        \left[\begin{array}{cc}
                               C_u(\bm{y},\bm{y}) & \bm{P} \\
-\end{array}
-\right]
-\left[
-\begin{array}{cc}
+                              \end{array}\right]
+                        \left[\begin{array}{cc}
                               \bm{\Sigma} & \bm{P} \\
                               \bm{P}^T  & \bm{0} \\
-\end{array}
-\right]^{-1}
-\left[
-\begin{array}{c}
+                              \end{array}\right]^{-1}
+                        \left[\begin{array}{c}
                               C_u(\bm{y},\bm{y}) \\
                               \bm{P}^T \\
-\end{array}
-\right].
+                              \end{array}\right].
 \end{equation}
 We compute $\hat{\bm{d}}$ using SE, Wendland, and IBM covariance functions for $T$ and the median hyperparameters from Table 1.
-Figure 1 compares the easting component of $\bm{d}_*$ to $\hat{\bm{d}}$ for the Winter 2016 SSE.
+Figure \ref{fig:Fit} compares the easting component of $\bm{d}_*$ to $\hat{\bm{d}}$ for the Winter 2016 SSE.
 For most stations, the data prediction vector appears to appropriately describe displacements throughout the SSE, regardless of the choice of $T$.
@@
 \caption{Strain map.}   
-\label{fig:Fit}
+\label{fig:StrainMap}
 \end{figure*}
@@
 \caption{Strain time series.}   
-\label{fig:Fit}
+\label{fig:StrainTs}
 \end{figure*}
@@
 \caption{Strain magnitude time series.}   
-\label{fig:Fit}
+\label{fig:StrainMag}
 \end{figure*}

### 2017-06-13 HinesHetlandGPR.tex
did more

@@
 In Table 1, we summarize how the log REML likelihoods for the Wendland and IBM covariance functions compare to the SE covariance function.
-The tabulated values show the median and IQR differences in log REML likelihoods over each of the seven SSEs considered in this study.
-Negative values indicate that the observed data is less likely to have been sampled from the covariance function.
 Based on the differences in log REML likelihoods, the data is substantially more likely to come form the SE or Wendland covariance functions than from the IBM covariance function.
 The REML likelihoods do not definitively indicate whether the SE or Wendland covariance function is preferable.
-
 As another test to determine the most appropriate covariance function for $T$, we compare the observations to the predicted displacements for each covariance function.
@@
 The predictions for the Wendland and SE covariance functions are nearly indistinguishable.
-We ultimately settle on the Wendland covariance function for $T$.
+In our estimates of transient strain discussed in the next section, we ultimately settle on the Wendland covariance function for $T$ and use the median values from Table 1 for the hyperparameters.
 We choose the Wendland covariance function over the SE covariance function because of its computational advantages noted above.
@@
 \subsection{Transient Strain Rates} 
-Having established noise model and a prior for transient displacements, we use cleaned GNSS dataset described in Section \ref{sec:SignalModel} to calculate transient strain rates in the Puget Sound region from eq.
-(\ref{eq:StrainRate}).
-Our solution is shown in Figures X through X.
-Figures X through X show the estimates $\dot{\bm{\epsilon}}$ complete with uncertainties at select times over the course of the Winter 2016 SSE.
-Figure X, which shows strain rate at the height of the SSE, makes the implications of SSEs for seismic hazard clear.
-The SSE causes compression in the Olympic Peninsula and extension in East of Puget Sound.
-The SSE are thus concentrating strain energy trench-ward, and pushing the subduction zone closer to failure.
-This is not novel insight because the same conclusions has been drawn from fault slip models which show that SSEs tend to be just at the transition between locked and decoupled.
-Slow slip event depths from \citet{Dragert2001}, \citet{Wech2009}, \citet{Schmidt2010}, \citet{Bartlow2011} show slip concentrated at depths from 30 to 50 km detph.
-What is novel is that our estimates of strain rate from SSEs are nonparametric and less likely to be biased by systematic errors such as errors in the fault model.
-Are uncertainties are also likely to be more conservative since we do not make any dubious assumptions.
-Thus we can be more confident in our assessment of how much strain is accumulating from SSEs.
+Having established a noise model and a prior for transient displacements, we use the cleaned GNSS dataset described in Section \ref{sec:SignalModel} to calculate transient strain rates in the Puget Sound region.
+We calculate transient strain rate on each day from January 1, 2010 to March 15, 2017.
+The strain rates are estimates on a grid of points spanning the study area.
+In Figure \ref{fig:StrainMap}, we show the transient strain rates on January 1, 2016, which coincides with the height of an SSE.
+We have included a video showing the map view of strain rates over time as supplementary material.
+The strain rates shown in Figure \ref{fig:StrainMap} are generally similar to the strain rates for the other six SSEs considered in this study.
+The SSEs cause compression in the Olympic Peninsula and extension east of Puget Sound.
+For comparison, long-term strain rates indicate trench perpendicular compression throughout this study region \citep{Murray2000,McCaffrey2007,McCaffrey2013}.
+The SSE are thus concentrating tectonically accumulated strain energy trench-ward, and pushing the subduction zone closer to failure.
+Similar conclusions have been drawn based on finite fault modeling \citep[e.g.,][]{Dragert2001,Wech2009,Schmidt2010,Bartlow2011}, which reveal that SSEs are occurring just below the portion of the subduction zone which is inferred to be locked.
+Thus, the SSEs are concentrating strain up dip and pushing the locked fault closer to failure.
+A main difference between the strain inferred here and strain derived from finite fault models is that our estimated strain rates are not based on an assumed physical model.
+In contrast, finite fault models can be biased by errors in the assumed fault geometry or lithospheric rheology.
+Since our estimated strain rates lack such systematic errors, we can be more confident that our solution is unbiased and has meaningful uncertainties.
 
@@
 
-To demonstrate the effectiveness of our method for detecting SSEs, we show a timeseries of estimated strain components at the position indicated in figure X.
-We also compare the strain rate magnitude and the frequency of tremors in this region.
-The strain rate magnitudes have been normalized by its estimated uncertainty and so strain magnitudes less than ${\sim}$2 can be considered to be part of the background noise.
-The only time that the normalized strain rate magnitude exceeds 2 is when there is a corresponding peak in tremor.
-We are able to accurately detect the seven SSEs in over this period in addition to the inter-SSE events such as the one at X.
+In Figure \ref{fig:StrainTs} we show the time dependence of estimated transient strain rates at a position on the Olympic Peninsula, where strain rate magnitudes from SSEs are largest.
+To verify that the estimated strain rates are accurately identifying geophysical signal, we compare the strain rate magnitude to the frequency of seismic tremor (Figure \ref{fig:StrainMag}).
+The shown strain rate magnitude has been normalized by the estimated uncertainty, and it can be interpreted as a signal-to-noise ratio.
+Estimated strain rate magnitudes greater than ${\sim}3$ can then be interpreted as a detected transient event.
+For each detected transient event there is a corresponding peak in seismic tremor.
+We are also able to clearly identify transient strain associated with a more subtle inter-SSE event in August 2014.
+In between SSEs the normalized strain rate magnitude is consistently between 0 and 2, indicating that no anomalous non-SSE events are being detected, at least at this location.
 
@@
 \end{figure*}
-
-A video can be found as supp material
-
-made in ,  is accumulating to the trench perpendicular compression releases At the height of the SSE, the implications of
-
-model exhibits the same relationship  which can described
-
-Both of which have clear implications for assessing seismic hazard.
-
-, which can have applications for seismic hazard assessment,Identifying strain resulting from ETS events serves two purposes: (1) ETS events have  events which are  were first discovered by \citet{Dragert2001}.
-
-We demonstrate that Gaussian processes are effective tools for detecting transient strain by using them to detect Cascadia slow slip events.
-
-BACKGROUND
-
-\citet{Dragert2001} first discovered slow slip events
-
-Explain strain concentration with this:
-
-Interseismic locking depths from \citet{Fluck1997}, \citet{Murray2000}, \citet{McCaffrey2007} and \citet{McCaffrey2013}, \citet{Burgette2009}, \citet{schmalzle2014} are consistent with a full locking down to about 20 km.
-
-%Studies which simultaneously looked at intersesimic and ets are \citet{Holtkamp2010} and \citet{Schmalzle2014}
-Studies which simultaneously modeled interseismic and ets are \citet{Holtkamp2010} and \citet{schmalzle2014}
 

### 2017-06-15 HinesHetlandGPR.tex
added more

@@
 
+GPR is a bayesian approach which does not rely on dubious parametric assumptions.
+
 \section{Method}\label{sec:Method}
@@
 \begin{split}
-d_{ij} = &u(x_i,t_j) + \epsilon(x_i,t_j) + w_{ij} + a^{(1)}_i + a^{(2)}_it_j + \\
+d_{ij} = &u(x_i,t_j) + v(x_i,t_j) + w_{ij} + a^{(1)}_i + a^{(2)}_it_j + \\
          &a^{(3)}_i\sin(2 \pi t_j) + a^{(4)}_i\cos(2 \pi t_j) + a^{(5)}_i\sin(4 \pi t_j) + a^{(6)}_i\cos(4 \pi t_j), 
@@
 We use $w_{ij}$ to denote normally distributed, uncorrelated noise.
-Correlated noise which does not have a parametric representation is denoted by $\epsilon$.
-For example, $\epsilon$ can consist of temporally correlated noise describing benchmark wobble \citep[e.g.,][]{Wyatt1982,Wyatt1989}, and/or spatially correlated noise describing common mode error \citep[e.g.,][]{Wdowinski1997}.
-For now, we will only assume that $\epsilon \sim \mathcal{N}(0,C_\epsilon)$.
+Correlated noise which does not have a parametric representation is denoted by $v$.
+For example, $v$ can consist of temporally correlated noise describing benchmark wobble \citep[e.g.,][]{Wyatt1982,Wyatt1989}, and/or spatially correlated noise describing common mode error \citep[e.g.,][]{Wdowinski1997}.
+For now, we will only assume that $v \sim \mathcal{N}(0,C_v)$.
 We consider the six coefficients in eq.
@@
 \begin{equation}
-\bm{d} = u(\bm{y}) + \epsilon(\bm{y}) + \bm{w} + \bm{P}\bm{a},
+\bm{d} = u(\bm{y}) + v(\bm{y}) + \bm{w} + \bm{P}\bm{a},
 \end{equation}
@@
 \begin{equation}\label{eq:Cd}
-\bm{\Sigma} = C_u(\bm{y},\bm{y}) + C_\epsilon(\bm{y},\bm{y}) + 
+\bm{\Sigma} = C_u(\bm{y},\bm{y}) + C_v(\bm{y},\bm{y}) + 
               \mathrm{diag}\left(\bm{\sigma}^2\right).  
@@
 We now describe how to condition the prior for each component of transient displacements, $u$, with $\bm{d}_*$ to form a posterior estimate of transient displacements, $\hat{u} = u | \bm{d}_*$.
-We will assume that an appropriate covariance functions and corresponding hyperparameters for $X$, $T$, and $C_\epsilon$ have already been chosen.
+We will assume that an appropriate covariance functions and corresponding hyperparameters for $X$, $T$, and $C_v$ have already been chosen.
 If $\kappa$ is kept finite, then we can follow \citet{Rasmussen2006} to find that $\hat{u}$ is distributed as $\mathcal{N}(\mu_{\hat{u}},C_{\hat{u}})$, where
@@
 (\ref{eq:PosteriorMean3}) and (\ref{eq:PosteriorCov3}) for each displacement component.
-We differentiate the Gaussian process $\hat{\bm{u}}$ (see \cite{Abrahamsen1997}) to form an estimate of transient strain rate, $\dot{\bm{\epsilon}}$, which is itself a spatially and temporally continuous Gaussian process.
-The components of the transient strain rate tensor, $\dot{\epsilon}_{ij}$, are given by
+We differentiate the Gaussian process $\hat{\bm{u}}$ (see \cite{Abrahamsen1997}) to form an estimate of transient strain rate, $\dot{\bm{\varepsilon}}$, which is itself a spatially and temporally continuous Gaussian process.
+The components of the transient strain rate tensor, $\dot{\varepsilon}_{ij}$, are given by
 \begin{equation}\label{eq:StrainRate}
-\dot{\epsilon}_{ij}(z) = \frac{1}{2} \frac{\partial}{\partial t} \left(
+\dot{\varepsilon}_{ij}(z) = \frac{1}{2} \frac{\partial}{\partial t} \left(
                                         \frac{\partial \hat{u}_i(z)}{\partial x_j} +  
@@
 \end{equation}
-which are distributed as $\mathcal{N}(\mu_{\dot{\epsilon}_{ij}},C_{\dot{\epsilon}_{ij}})$, where
+which are distributed as $\mathcal{N}(\mu_{\dot{\varepsilon}_{ij}},C_{\dot{\varepsilon}_{ij}})$, where
 \begin{equation}\label{eq:StrainMean}
-\mu_{\dot{\epsilon}_{ij}}(z) = \frac{1}{2}\frac{\partial}{\partial t}\left(
+\mu_{\dot{\varepsilon}_{ij}}(z) = \frac{1}{2}\frac{\partial}{\partial t}\left(
                                   \frac{\partial \mu_{\hat{u}_i}(z)}{\partial x_j} + 
@@
 \begin{equation}\label{eq:StrainCov}
-C_{\dot{\epsilon}_{ij}}(z,z') = \frac{1}{4} \frac{\partial^2}{\partial t \, \partial t'}\left(
+C_{\dot{\varepsilon}_{ij}}(z,z') = \frac{1}{4} \frac{\partial^2}{\partial t \, \partial t'}\left(
                                    \frac{\partial^2 C_{\hat{u}_i}(z,z')}{\partial x_j \, \partial x'_j} + 
@@
 We calculate the residuals between the observations and a best fitting model, and data with residuals that are anomalously large are identified as outliers.
-We treat $\bm{d}_*$ as a sample of $\bm{d}$ and assume that there is no temporally correlated noise (i.e., $\epsilon = 0$).
+We treat $\bm{d}_*$ as a sample of $\bm{d}$ and assume that there is no temporally correlated noise (i.e., $v = 0$).
 The best fitting model for $\bm{d}_*$ is considered to be the expected value of the random vector $u(\bm{y}) + \bm{P}\bm{a}$ after conditioning it with non-outlier observations.
@@
 
-\section{Application Cascadia Slow Slip Events}\label{sec:Cascadia}
+\section{Application to Cascadia Slow Slip Events}\label{sec:Cascadia}
 In this section we estimate transient strain from GNSS data in Cascadia.
@@
 We demonstrate that transient strain rates estimated with the method described above are an effective tool for detecting SSEs and potentially other geophysical signal in GNSS data.
-In Cascadia, SSEs can be detected by monitoring for associated seismic tremor, which is actively being done by the Pacific Northwest Seismic Network (PNSN) \citep{Wech2010}.
-We verify that transient strain rates estimated with the method described above are indeed identifying strain from SSEs by comparing $\bm{\epsilon}$ to the tremor records.
+In Cascadia, SSEs can be detected by monitoring for associated seismic tremor \citep{Rogers2003}, which is actively being done by the Pacific Northwest Seismic Network (PNSN) \citep{Wech2010}.
+We verify that transient strain rates estimated with the method described above are indeed identifying strain from SSEs by comparing $\dot{\bm{\varepsilon}}$ to the tremor records.
 We envision that GNSS derived transient strain rates could be useful for monitoring subduction zones which have SSEs but no associated tremor \citep{Schwartz2007}.
@@
 \subsection{Noise model}\label{sec:NoiseModel}
-Before we determine the transient strain rates, we must establish a prior for the transient deformation, $u$, and the noise, $\epsilon$.
-In this section we discuss our choice for the noise covariance function $C_\epsilon$.
+Before we determine the transient strain rates, we must establish a prior for the transient deformation, $u$, and the noise, $v$.
+In this section we discuss our choice for the noise covariance function $C_v$.
 There have been numerous studies on temporally correlated noise in GNSS data \citep[e.g.,][]{Zhang1997,Mao1999,Williams2004,Langbein2008}.
@@
 There is some physical justification for using Brownian motion as a noise model because it accurately describes the power spectrum of motion resulting from instability in geodetic monuments \citep[e.g.,][]{Wyatt1982,Wyatt1989,Langbein1997}.
-Here we describe the time dependence of $\epsilon$ as a FOGM process and consider $\epsilon$ to be spatially uncorrelated.
+Here we describe the time dependence of $v$ as a FOGM process and consider $v$ to be spatially uncorrelated.
 A FOGM process is a solution to the stochastic differential equation
 \begin{equation}\label{eq:FOGMdef}
-\dot{\epsilon}(t) + \alpha \epsilon(t) = \beta w(t),
+\dot{v}(t) + \alpha v(t) = \beta w(t),
 \end{equation}
 where $w(t)$ is white noise with unit variance.
-The FOGM process degenerates to the commonly used Brownian motion noise model under the condition that $\alpha=0$ and $\epsilon(0) = 0$, which is evident from the definition of Brownian motion \citep[e.g.,][]{Papoulis1991}.
+The FOGM process degenerates to the commonly used Brownian motion noise model under the condition that $\alpha=0$ and $v(0) = 0$, which is evident from the definition of Brownian motion \citep[e.g.,][]{Papoulis1991}.
 Our noise model which satisfies eq.
@@
 \begin{equation}\label{eq:FOGM}
-C_\epsilon\left((x,t),(x',t')\right) = \frac{\beta^2}{2\alpha}\exp\left(-\alpha|t - t'|\right) \delta(||x - x'||_2). 
+C_v\left((x,t),(x',t')\right) = \frac{\beta^2}{2\alpha}\exp\left(-\alpha|t - t'|\right) \delta(||x - x'||_2). 
 \end{equation}
 
-We constrain the hyperparameters for $\epsilon$, $\alpha$ and $\beta$, with a set of 38 continuous GNSS stations in Washington and Oregon which are east of 121$^\circ$ W.
+We constrain the hyperparameters for $v$, $\alpha$ and $\beta$, with a set of 38 continuous GNSS stations in Washington and Oregon which are east of 121$^\circ$ W.
 These stations are sufficiently far from the subduction zone that they are unlikely to contain transient signal.
@@
 We then find $\alpha$ and $\beta$ for each station timeseries with the Restricted Maximum Likelihood (REML) method \cite[e.g.,][]{Harville1974,Cressie1992}.
-The REML method finds the hyperparameters which maximize the expression
+The REML method finds the hyperparameters, which we collectively refer to as $\bm{\theta}$, that maximize the likelihood function
 \begin{equation}\label{eq:REML}
-\left(\frac{\left|\bm{P}^T\bm{P}\right|}
+\mathcal{L}(\bm{\theta}) = \left(\frac{\left|\bm{P}^T\bm{P}\right|}
                            {(2\pi)^{n-6m} 
-            \left| \bm{\Sigma} \right| 
+                           \left| \bm{\Sigma}(\bm{\theta}) \right| 
                            \left| \bm{P}^T\bm{\Sigma}\bm{P} \right|}\right)^{\frac{1}{2}} 
-e^{-\tfrac{1}{2}\bm{d}_*^T\bm{K}\bm{d}_*},
+                           e^{-\tfrac{1}{2}\bm{d}_*^T\bm{K}(\bm{\theta})\bm{d}_*},
 \end{equation}
@@
 \begin{equation}
-\bm{K} = \bm{\Sigma}^{-1} - 
-         \bm{\Sigma}^{-1}\bm{P}
-         \left(\bm{P}^T\bm{\Sigma}^{-1}\bm{P}\right)^{-1}
-         \bm{P}^T\bm{\Sigma}^{-1}.
+\bm{K}(\bm{\theta}) = \bm{\Sigma}(\bm{\theta})^{-1} - 
+                      \bm{\Sigma}(\bm{\theta})^{-1}\bm{P}
+         \left(\bm{P}^T\bm{\Sigma}(\bm{\theta})^{-1}\bm{P}\right)^{-1}
+         \bm{P}^T\bm{\Sigma}(\bm{\theta})^{-1}.
 \end{equation}
@@
 
-For these reasons, we henceforth make the simplifying assumption that temporally and spatially correlated noise is negligible and $\epsilon = 0$ for the easting and northing component of GNSS data.
+For these reasons, we henceforth make the simplifying assumption that temporally and spatially correlated noise is negligible and $v = 0$ for the easting and northing component of GNSS data.
 
@@
 The SSE are thus concentrating tectonically accumulated strain energy trench-ward, and pushing the subduction zone closer to failure.
-Similar conclusions have been drawn based on finite fault modeling \citep[e.g.,][]{Dragert2001,Wech2009,Schmidt2010,Bartlow2011}, which reveal that SSEs are occurring just below the portion of the subduction zone which is inferred to be locked.
-Thus, the SSEs are concentrating strain up dip and pushing the locked fault closer to failure.
-A main difference between the strain inferred here and strain derived from finite fault models is that our estimated strain rates are not based on an assumed physical model.
-In contrast, finite fault models can be biased by errors in the assumed fault geometry or lithospheric rheology.
+Similar conclusions have been drawn based on fault slip models \citep[e.g.,][]{Dragert2001,Wech2009,Schmidt2010}, which reveal that SSEs are occurring downdip of the seismogenic zone.
+The SSEs, which are thrust events, migrate stress updip and push the locked portion of the fault closer to failure.
+A key difference between the strain inferred here and strain derived from fault slip models is that our estimated strain rates are not based on an assumed physical model.
+In contrast, fault slip models can be biased by errors in the assumed fault geometry or lithospheric rheology.
+Moreover, the degrees of freedom in fault slip models usually cannot be constrained by GNSS data alone, and it is necessary to impose regularization which further biases the results.
 Since our estimated strain rates lack such systematic errors, we can be more confident that our solution is unbiased and has meaningful uncertainties.
@@
 To verify that the estimated strain rates are accurately identifying geophysical signal, we compare the strain rate magnitude to the frequency of seismic tremor (Figure \ref{fig:StrainMag}).
-The shown strain rate magnitude has been normalized by the estimated uncertainty, and it can be interpreted as a signal-to-noise ratio.
+The shown strain rate magnitude has been normalized by the estimated uncertainty and can be considered a signal-to-noise ratio.
 Estimated strain rate magnitudes greater than ${\sim}3$ can then be interpreted as a detected transient event.
@@
 
-\section{Discussion and Conclusion}\label{sec:Discussion}
+\section{Other Potential Applications}\label{sec:Other}
+In the previous section we demonstrated that transient strain rates can be used to detect SSEs or other transient geophysical phenomina.
+Another potential application would be to use the GNSS derived transient strain rates to develop noise models for borehole strain meters (BSMs).
+The plate Boundary Observatory maintains 43 BSMs in the Pacific Northwest, and it has been demonstrated that BSMs are able to record transient geophysical events such as SSEs \citep[e.g.,][]{Dragert2011}.
+However, noise in BSM data is not well understood, which complicates the use of BSM data for identifying geophysical signal.
+The noise consists, in part, of a long-term decay resulting from the instrument equilibrating with the surrounding rock \citep{Gladwin1987}.
+Typically, this noise is dealt with in an ad-hoc manner by fitting and removing exponentials and low order polynomials.
+If we use GNSS derived strain rates as an approximation of the signal in BSM data, then we could perform a more rigorous quantification of BSM noise.
+
+We also note that there is potential for a more thorough analysis of the spatio-temporal noise in GNSS data, $v$, than what was performed in Section \ref{sec:NoiseModel}.
+In this study we have not explored the spatial covariance of $v$, since it would not have significantly affected our estimated strain rates.
+If, for example, we were interested in inferring fault slip from GNSS displacements then it may be necessary to account for the spatial covariance of the $v$.
+
+In future studies, we may explore more sophisticated seasonal models than what is used in eq.
+(\ref{eq:Data}).
+As suggested earlier, we may want to quantify the spatial covariance of seasonal deformation.
+We may also want to explore whether the semiannual terms in eq.
+(\ref{eq:Data}) are necessary or if higher frequency terms are warranted.
+The periodic Gaussian process \citep{Mackay1998} is an alternative model for seasonal deformation and is well suited for exploring the roughness of seasonal deformation.
+The periodic Gaussian process has zero mean and covariance function,
+\begin{equation}\label{eq:Periodic}
+T(t,t') = \phi \exp\left(\frac{-\sin(\pi|t - t'|)^2}{2\theta^2}\right).
+\end{equation}
+Realizations have annual periodicity and the roughness is controlled by $\theta$.
+Decreasing $\theta$ has the same effect as including higher frequency sinusoids in the seasonal model.
+The optimal value for $\theta$ can be found with the REML method as described in Section \ref{sec:NoiseModel}.
+
+\section{Conclusion}\label{sec:Conclusion}
+In this paper we describe how Gaussian processes can be used to identifying transient strain from GNSS data.
+Our approach is Bayesian and we use a spatio-temporal stochastic model to describe the prior for transient deformation and the noise in GNSS data.
+Our estimated transient strain rates are complete with meaningful estimates of uncertainties, which allows us to objectively discern geophysical signal from noise.
+Other commonly used methods for estimate long-term and transient strain assume a parametric form for the underlying signal, and errors in this assumption will result in a biased solution.
+In contrast our stochastic prior model is chosen with maximum likelihood methods, which is an objective and data-driven approach.
+The lack of subjectivity in our method gives us confidence that the results are unbiased.
+
+We have demonstrated with GNSS data from Cascadia that Gaussian processes are effective for identifying transient strain.
+We also envision that Gaussian processes can be an effective tool to handle other geophysical things.
+
+used to tackle other several other closely related applications
+
+Since our method lacks subjectivity, the obto be most consistent with the data  maximum likelihood methods.
+that is, constrained by the observed data.
+and is less likely the results and lead to this and We prefer a stochastic prior model over a parametric prior model such
+
+The uncertainties on the estimate of transient strain can be trusted, provided that we have chosen appropriate stochastic models for the prior and noise.
+These models are objectively chosen with on maximum likelihoods methods, and
+
+A stochastic prior model is preferable to a As a result, we obtain sensible
+
+% In the previous section we demonstrate that Gaussian processes can be an effective tool for illuminating 
+% GPR can be used for more than just identifying transient strain rates and
+
 % discuss potential applications for slip models
@@
 
-% note that this can be used for smoothing verticals, and compare it with burgette2009
-
-% change the notation for stencil size or smoothness order because they are both n. use p for smoothness order
-
 The material presented in this paper is primarily focused on GPS data; however, we speculate that the RBF-FD scheme can be of particular use in denoising borehole strain meter (BSM) data.

### 2017-06-15 HinesHetlandGPR.tex
more shit

@@
 \section{Introduction}\label{sec:Introduction}
-Crustal strain rates are fundamentally important quantities for assessing seismic hazard, since knowing where and how quickly strain is accumulating gives insight into where we can expect stored elastic energy to be released seismically.
-It is then important to develop and improve upon methods for mapping strain in tectonically active regions because such maps could conceivably feed into seismic hazard models such as UCERF3 \citep{Field2014}.
+Crustal strain rates are fundamentally important quantities for assessing seismic hazard.
+For one reason, knowing where and how quickly strain is accumulating gives insight into where we can expect stored elastic energy to be released seismically.
+Secular crustal strain rates, based on Geologic and GNSS data, have been used to constrain seismic hazard models such as UCERF3 \citep{Field2014}.
+Dense networks of continuous GNSS stations, such as the Plate Boundary Observatory (PBO), have sufficient spatial and temporal coverage for us to feasibly make estimates of transient crustal strain rates, which is considered to be any deviation from the secular strain rate.
+Transient crustal strain rates can potentially illuminate geophysical signal, which may not be immediately apparent from inspecting the GNSS displacement data.
+This too is relevant for assessing seismic hazard because there are several instances of major earthquakes being triggered by transient geophysical phenomena, such as slow slip events \citep{Roeloffs2006} and postseismic deformation \citep{Freed2001}.
+For these reasons, it is important to develop and improve upon methods for estimating crustal strain rates from GNSS data.
+
+Most methods for estimating crustal strain rates, both secular and time-dependent, from GNSS data assume some parametric form of the deformation signal, and then strain rates are the spatial derivative of the best fitting model.
+The simplest and traditional method for estimating secular crustal strain rates assumes that GNSS derived velocities can be described with a first order polynomial (i.e., having a constant spatial gradient) over some subnetwork of the GNSS stations \citep[e.g][]{Frank1966,Prescott1976,Savage1986,Feigl1990,Murray2000}.
+Linear least squares is then used to find the components of the strain rate tensor for each subnetwork.
+The uniform spatial gradient assumption is obviously not appropriate when subnetworks span a large area, although it is not clear when a subnetwork is too large.
+To help overcome this deficiency, \citet{Shen1996} and \citet{Shen2015} used an inverse distance weighting scheme where the estimated strain rate at some point is primarily controlled by observations at nearby stations.
+However, in formulating the algorithms from \citet{Shen1996} and \citet{Shen2015}, it is still assumed that the underlying displacement gradients is uniform over the entire network.
+The errors in this assumption manifests itself as implausibly low formal uncertainties for the estimated strain rates.
+Other data-based methods for estimating secular strain rates have parameterized velocities with bicubic splines \citep{Beavan2001}, spherical wavelets \citep{Tape2009}, and elastostatic Green's functions \citep{Sandwell2016}.
+The choice of parameterization can be subjective.
+If there are too few degrees of freedom in the parameterization then estimated strain rates will be biased and the uncertainties will be underestimated.
+If there are too many degrees of freedom then there will not be any coherent features in the estimated strain rates.
+One could parameterize deformation with a physically motivated model of interseismic deformation \citep[e.g.,][]{Meade2005,McCaffrey2007}.
+In such models the lithospheric rheology and plate interfaces are assumed to be known.
+Any errors in the assumed physical model could similarly result in biased strain estimates and underestimated formal uncertainties.
+
+The aforementioned studies have all been concerned with estimating secular strain rates.
+In recent years the Southern California Earthquake Center (SCEC) community has shown interest in developing methods for detecting transient strain.
+SCEC supported a transient detection exercise, where several research groups tested their methods for detecting transient geophysical signals with a synthetic GNSS dataset.
+Among the methods tested were the Network Strain Filter (NSF) \citep{Ohtani2010} and the Network Inversion Filter (NIF) \citep{Segall1997}.
+the NSF parameterizes the geophysical signal spatially with wavelets and the NIF, which is intended for detecting slow fault slip, uses the elastic dislocation Green's functions from \citet{Okada1992}.
+For the NSF and NIF, the time dependence of the geophysical signal is modeled as integrated Brownian motion.
+The method described in \citet{Holt2013} was also tested in the transient detection exercise, and they calculate strain rates using a bicubic spatial parameterization of displacements between time epochs.
+\citet{Holt2013} defined a detection threshold based on the strain rate magnitude, and we demonstrate that this is indeed an effective criterion for identifying geophysical signal.
+One main challenge with using these methods for detecting transient geophysical signal is that it is difficult to distinguish signal from noise, since the noise on the inferred deformation can be significantly influenced by the choice of parameterization.
+
+Each of the aforementioned methods have assumed a described above
+
+Each of these methods use a spatial parameterization for the geophysical signal. If the  and deformation spatial  use a parametric model
+
+determine strain rates by spatially parameterizing the and they  described a method for detecting transient geophysical signal based on the strain rate magnitude, which were used in this exercise, and their approach fidentified transient strain events by fitting a set of spatial wavelet basis functions to the deformation field at discrete time epochs, and a Kalman filtering strategy was used to ensure that the coefficients for each basis function varied smoothly in time.
+\citet{Holt2013} calculated time dependent strain by differentiating a bicubic interpolant that was fit to each epoch of a temporally smoothed deformation field.
+
+physical models a physical model that based on an assumption of the earths rheology and the location of plate interfaces, Another approach Model-based estimates of secular strain rate find
+
+Another strategy is to fit a set of interpolating basis functions to the deformation field and then compute the strain from the analytical derivative of the interpolant \citep[e.g.][]{Beavan2001,Tape2009,Sandwell2016}.
 
@@
 The classic and simplest method for estimating strain is to assume that the strain rate is constant in time and spatially uniform within subnetworks of the geodetic data.
-Linear least squares is then used to find the components of the strain rate tensor for each subnetwork \citep[e.g][]{Frank1966,Prescott1976,Savage1986,Feigl1990,Murray2000}.
+Linear least squares is then used to find the components of the strain rate tensor for each subnetwork .
 Several algorithms have been developed to improve upon this procedure.
 \citet{Shen1996} and \citet{Shen2015} discuss an algorithm where, instead of using the immediately adjacent stations to calculate strain at a position, the strain is computed with a weighted average over the entire network where the weighting is smaller for more distant stations.
-Another strategy is to fit a set of interpolating basis functions to the deformation field and then compute the strain from the analytical derivative of the interpolant \citep[e.g.][]{Beavan2001,Tape2009,Sandwell2016}.
 
@@
 The strain rates presented in this paper are more direct observations and can be interpretted with a higher degree of confidence.
-
-The Southern California Earthquake Center (SCEC) community has recently shown interest in developing methods for detecting transient signals.
-SCEC supported a transient detection exercise, where several research groups tested their transient detection methods with a synthetic GNSS dataset.
-supported the transient Our motivation for detecting SSEs is, in part, because there have been several instances where SSEs immediately preceded megathrust earthquakes \citep{Roeloffs2006}.
 
@@
 
+Numerical Notes/scaling. Use partitioned matrix and Cholmod to solve for strain.
+
 \section{Conclusion}\label{sec:Conclusion}

### 2017-06-15 HinesHetlandGPR.tex
added more

--- 2017-06-15 HinesHetlandGPR.tex
@@
 \usepackage{bm}
-\title{Revealing Transient Strain in Geodetic Data with the Radial Basis Function Finite Difference Method}
+\title{Revealing Transient Strain in Geodetic Data with Gaussian Process Regression}
 \author{Trever T. Hines and Eric A. Hetland}
@@
 % INTRODUCTION
-% - Why we care about strain in the crust
-%   - Knowing long term strain rates tells us where we can expect seismicity
-%
+
 % - How people have gone about estimating strain rates
@@
 \section{Introduction}\label{sec:Introduction}
-Crustal strain rates are fundamentally important quantities for assessing seismic hazard, since knowing where and how quickly strain is accumulating gives insight into where we can expect stored elastic energy to be released seismically.
-It is then important to develop and improve upon methods for mapping strain in tectonically active regions because such maps could conceivably feed into seismic hazard models such as UCERF3 \citep{Field2014}.
+Crustal strain rates are fundamentally important quantities for assessing seismic hazard.
+Knowing where and how quickly strain is accumulating gives insight into where we can expect stored elastic energy to be released seismically.
+Consequently, secular crustal strain rates, which are estimated from GNSS data, have been used to constrain seismic hazard models such as UCERF3 \citep{Field2014}.
+Dense networks of continuous GNSS stations, such as the Plate Boundary Observatory (PBO), make it feasible to estimate transient crustal strain, which we consider to be any deviation from the secular strain rate.
+Transient crustal strain can potentially illuminate geophysical signal that may not be immediately apparent from inspecting the GNSS displacement data.
+The ability to detect transient geophysical phenomena is also relevant for seismic hazard because there are instances of major earthquakes being triggered by slow slip events \citep{Roeloffs2006} and postseismic deformation \citep{Freed2001}.
+For these reasons, it is important to develop and improve upon methods for estimating crustal strain rates from GNSS data.
 
-Methods for estimating strain from geodetic data fall in one of two categories.
-There are model-based approaches which assume that strain is the result of loading on faults which have a known geometry, and there are data-based approaches which make no assumptions about the source of deformation.
-We will exclusively consider data-based approaches in this paper.
-The classic and simplest method for estimating strain is to assume that the strain rate is constant in time and spatially uniform within subnetworks of the geodetic data.
-Linear least squares is then used to find the components of the strain rate tensor for each subnetwork \citep[e.g][]{Frank1966,Prescott1976,Savage1986,Feigl1990,Murray2000}.
-Several algorithms have been developed to improve upon this procedure.
-\citet{Shen1996} and \citet{Shen2015} discuss an algorithm where, instead of using the immediately adjacent stations to calculate strain at a position, the strain is computed with a weighted average over the entire network where the weighting is smaller for more distant stations.
-Another strategy is to fit a set of interpolating basis functions to the deformation field and then compute the strain from the analytical derivative of the interpolant \citep[e.g.][]{Beavan2001,Tape2009,Sandwell2016}.
+Most methods for estimating crustal strain rates from GNSS data assume some parametric form of the deformation signal.
+The simplest and traditional method for estimating secular crustal strain rates assumes that GNSS derived velocities can be described with a first order polynomial (i.e., having constant spatial gradients) over some subnetwork of the GNSS stations \citep[e.g][]{Frank1966,Prescott1976,Savage1986,Feigl1990,Murray2000}.
+The components of the strain rate tensor  for each subnetwork are then determined from the least squares fit to the observations.
+The assumption that deformation gradients are spatially uniform is obviously not appropriate when subnetworks span a large area.
+To help overcome this deficiency, \citet{Shen1996} and \citet{Shen2015} used an inverse distance weighting scheme where the estimated strain rate at some point is primarily controlled by observations at nearby stations.
+However, the methods described in \citet{Shen1996} and \citet{Shen2015} are still formulated by assuming that the deformation gradients are uniform over the entire network.
+The errors in this assumption manifest as implausibly low formal uncertainties for the estimated strain rates.
+Other methods for estimating secular strain rates have parameterized velocities with bicubic splines \citep{Beavan2001}, spherical wavelets \citep{Tape2009}, and elastostatic Green's functions \citep{Sandwell2016}.
+The choice of parameterization can be subjective.
+If there are too few degrees of freedom in the parameterization then estimated strain rates will be biased and the uncertainties will be underestimated.
+If there are too many degrees of freedom then there will not be any coherent features in the estimated strain rates.
+One could parameterize deformation with a physically motivated model of interseismic deformation \citep[e.g.,][]{Meade2005,McCaffrey2007}.
+In such models the lithospheric rheology and plate interfaces are assumed to be known.
+Any errors in the assumed physical model could similarly result in biased strain estimates and underestimated formal uncertainties.
 
-The aforementioned studies have all been concerned with estimating long term strain rates.
-Time dependent strain would be useful for studying geophysical processes which occur over timescales of days to years such as slow slip events, postseismic relaxation, or volcanic deformation.
-\citet{Ohtani2010} identified transient strain events by fitting a set of spatial wavelet basis functions to the deformation field at discrete time epochs, and a Kalman filtering strategy was used to ensure that the coefficients for each basis function varied smoothly in time.
-\citet{Holt2013} calculated time dependent strain by differentiating a bicubic interpolant that was fit to each epoch of a temporally smoothed deformation field.
+The aforementioned studies are concerned with estimating secular strain rates.
+In recent years the Southern California Earthquake Center (SCEC) community has shown interest in developing methods for detecting transient strain.
+SCEC supported a transient detection exercise \citep{Lohman2013}, where several research groups tested their methods for detecting transient geophysical signals with a synthetic GNSS dataset.
+Among the methods tested were the Network Strain Filter (NSF) \citep{Ohtani2010} and the Network Inversion Filter (NIF) \citep{Segall1997}.
+the NSF uses a wavelet parameterization to  describe the spatial component of geophysical signal and the NIF, which is intended for detecting slow fault slip, uses the elastic dislocation Green's functions from \citet{Okada1992}.
+For the NSF and NIF, the time dependence of the geophysical signal is modeled as integrated Brownian motion.
+The method described in \citet{Holt2013} was also tested in the transient detection exercise, and they calculate strain rates using a bicubic spatial parameterization of displacements between time epochs.
+\citet{Holt2013} defined a detection threshold based on the strain rate magnitude, and we will demonstrate that this is indeed an effective criterion for identifying geophysical signal.
+One challenge that is common to all these methods is that it is difficult to distinguish signal from noise, since the uncertainty on the inferred deformation can be significantly influenced by the choice of spatial parameterization.
 
-Each of the methods described above are designed to overcome two complications that arise in estimating deformation gradients: (1) geodetic data are noisy and differentiation will only amplify the noise and (2) geodetic data are not observed on a regular grid, which prevents the use of standard finite difference methods for computing derivatives.
-In this paper we demonstrate that both of these complications can be elegantly handled with the Radial Basis Function-Finite Difference (RBF-FD) method.
+In this paper we use Gaussian process regression (GPR) \citep{Rasmussen2006} to infer transient strain from GNSS data.
+GPR is a Bayesian, non-parametric method for inferring a continuous signal from discrete and potentially noisy data.
+GPR is closely related to Kriging \citep{Cressie1992} and Least Squares Collocation \citep{Kato1998}, and it has recently gained popularity as a tool for machine learning.
+GPR is Bayesian in that we describe our prior understanding of the geophysical signal with a Gaussian process.
+A Gaussian process is a normally distributed, continuous, stochastic process that is fully defined in terms of a mean function and a positive-definite covariance function.
+For example, Brownian motion is a well known Gaussian process in $\mathbb{R}^1$ which has zero mean and covariance function $C(t,t')=\min(t,t')$, where $t,t' \ge 0$.
+We use the Restricted Maximum Likelihood (REML) method to guide our choice for the prior Gaussian process.
+We then incorporate the GNSS observations with the prior to form a posterior estimate of transient strain rates.
+The posterior transient strain rate is also a Gaussian process and, provided that our prior is sensible, has well quantified uncertainties.
+With well quantified uncertainties we can more confidently discern geophysical signal from noise in our estimates of transient strain rates.
+We use GPR to infer transient strain rates resulting slow slip events (SSEs) in Cascadia.
+Our results demonstrate that GPR can be an effective tool for detecting SSEs and revealing the spatio-temporal evolution of crustal strain during SSEs.
 
-The RBF-FD method was developed simultaneously and independently by \citet{Tolstykh2003}, \citet{Shu2003}, \citet{Cecil2004}, and \citet{Wright2006} as a computationally efficient way to solve large scale partial differential equations over irregular, multi-dimensional domains.
-The RBF-FD method can be thought of as a generalization of the traditional finite difference method, where the node layout is no longer restricted to regular grids.
-Indeed, the RBF-FD method can be used to estimate derivates of discrete data located at arbitrary scattered positions in multi-dimensional space.
-The RBF-FD method is particularly appealing because it is algorithmically simple, regardless of the domain shape or node layout, and also because the method has performed well in numerous benchmark tests \citep[and references therein]{Fornberg2015}.
-
-In this paper, we do not use the RBF-FD method to solve a partial differential equation, but rather we use it to spatially smooth geodetic data and to compute deformation gradients.
-Our smoothing strategy can be viewed as a non-parametric, low-pass filter for scattered data where the degree of smoothness is controlled by a user specifies cutoff frequency.
-This can be contrasted with interpolation based methods where the resulting interpolant can be largely and perhaps unpredictably controlled by the choice of basis function.
-This process of smoothing and differentiating can be extended to estimate time dependent strain rates.
-In that case, we first temporally smooth and differentiate GPS displacement time series to get time dependent velocities.
-We then spatially smooth and differentiate the resulting velocities for each time epoch to get time dependent strain rates.
-
-The method proposed in this paper has numerous advantages which set it apart from other methods for computing strain rates.
-The method is computationally efficient and stable (there is no inversion of an ill-conditioned matrix).
-There are no hyper parameters or penalty parameters that need to be tuned for each application.
-As opposed to interpolation strategies such as \citet{Beavan2001}, \citet{Tape2009}, or \citet{Ohtani2010}, our method assumes that velocities are locally rather than globally continuous, which allows us to easily handle discontinuities resulting from, for example, a creeping fault.
-
-We begin this paper by summarizing the RBF-FD scheme and explaining how we construct differentiation matrices for scattered data.
-We then introduce the RBF-FD filter, which is used to smooth the observed geodetic data prior to differentiation.
-We provide two real world demonstraitions of our method for calculating strain rates.
-First we calculate the long term strain rates in Southern California from the CMM3 velocity data set \citep{Shen2011}, and we verify that our results are consistent with other studies.
-We then calculate time dependent strain rates in Cascadia from the GPS data provided by UNAVCO.
-In Cascadia, we analyze strain resulting from slow slip events and compare it to the long term tectonic strain accumulation.
-Slow slip events are found to produce compression in the Olympic Peninsula, which is in addition to the compression resulting from tectonic loading.
-Further south in Oregon, the slow slip events tend to release the compressional strain that is accumulated tectonically.
-While similar conclusions have been drawn from fault slip inversions for slow slip events, it is important to recognize that slip inversion are the product of inverting an ill-conditioned matrix making it difficult to determine whether slip inferences are real or just an artifact of the inversion.
-The strain rates presented in this paper are more direct observations and can be interpretted with a higher degree of confidence.
-
-The Southern California Earthquake Center (SCEC) community has recently shown interest in developing methods for detecting transient signals.
-SCEC supported a transient detection exercise, where several research groups tested their transient detection methods with a synthetic GNSS dataset.
-supported the transient Our motivation for detecting SSEs is, in part, because there have been several instances where SSEs immediately preceded megathrust earthquakes \citep{Roeloffs2006}.
-
-GPR is a bayesian approach which does not rely on dubious parametric assumptions.
-
-\section{Method}\label{sec:Method}
-In this section, we describe how GNSS displacement observations are used to identify transient crustal strain rates, which we denote as $\dot{\bm{\varepsilon}}(z)$, where $z$ is the ordered pair $(x,t)$, $x$ are spatial coordinates in $\mathbb{R}^2$, and $t$ is time.
+\section{Estimating Transient Strain Rates}\label{sec:Method}
+We denote transient crustal strain rates as $\dot{\bm{\varepsilon}}(z)$, where $z$ is the ordered pair $(x,t)$, $x$ are spatial coordinates in $\mathbb{R}^2$, and $t$ is time.
 We consider $\dot{\bm{\varepsilon}}$ to be spatially and temporally coherent deviations from the steady rate of strain accumulation from plate tectonics.
-We determine $\dot{\bm{\varepsilon}}$ by first identifying transient displacements, $\bm{u}(z)$, which we then spatially and temporally differentiate.
-As we will show in Section X, estimates of $\dot{\bm{\varepsilon}}$ turn out to be more effective at illuminating geophysical signal than estimates of $\bm{u}$ or $\dot{\bm{u}}$.
+We determine $\dot{\bm{\varepsilon}}$ by spatially and temporally differentiating estimates of transient displacements, $\bm{u}(z)$.
+As we discuss in Section \ref{sec:NoiseModel}, estimates of $\dot{\bm{\varepsilon}}$ turn out to be more effective at illuminating geophysical signal than estimates of $\bm{u}$ or $\dot{\bm{u}}$ because $\dot{\bm{\varepsilon}}$ is insensitive to common mode noise in GNSS data.
 We make a prior assumption that each component of $\bm{u}$ is a Gaussian process,
@@
 The appropriate choice for $X$ and $T$ may vary depending on the geophysical signal we are trying to describe (e.g.
-postseismic deformation or deformation from slow slip events), and we discuss this matter in the next section.
+postseismic deformation or deformation from slow slip events), and we discuss this matter in Section \ref{sec:SignalModel}.
 
@@
 We use $\bm{P}$ to represent the $n \times 6m$ matrix of corresponding basis functions evaluated at each point in $\bm{y}$.
-We  also denote the vector of uncorrelated noise for each observation as $\bm{w}$, whose standard deviations are given by the formal data uncertainty $\bm{\sigma}$.
-The observations $\bm{d}_*$ can then be viewed as a realization of the random vector
+We  also denote the vector of uncorrelated noise for each observation as $\bm{w}$, whose standard deviations are given by the formal data uncertainty, $\bm{\sigma}$.
+The observations can then be viewed as a realization of the random vector
 \begin{equation}
@@
 
-We now describe how to condition the prior for each component of transient displacements, $u$, with $\bm{d}_*$ to form a posterior estimate of transient displacements, $\hat{u} = u | \bm{d}_*$.
+The prior for each component of transient displacements is then conditioned with $\bm{d}_*$ to form a posterior estimate of transient displacements, $\hat{u} = u | \bm{d}_*$.
 We will assume that an appropriate covariance functions and corresponding hyperparameters for $X$, $T$, and $C_v$ have already been chosen.
@@
 
-\subsection{Outlier detection}\label{sec:Outlier}
+\section{Outlier detection}\label{sec:Outlier}
 In our formulation for estimating transient strain rates, we have assumed that noise in the data vector is normally distributed.
@@
 We still consider $u$ to have a separable covariance function as in eq.
-(\ref{eq:TransientCovariance}), and the choice for $X$ and $T$ does not need to be the same as that used in Section X.
+(\ref{eq:TransientCovariance}), and the choice for $X$ and $T$ does not need to be the same as that used in Section \ref{sec:Method}.
 Since outliers are determined based on how well a spatially and temporally dependent model fits the data, we are able to identify anomalous observations which may not be immediately apparent based on inspection of individual timeseries.
@@
 It should be noted that this algorithm does not identify jumps in GNSS time series, which are another common defect.
-Some, but not all, jumps can be automatically removed by parsing station logs and identifying dates of antenna changes.
-However, it is still necessary to manually inspect and remove jumps of unknown origin.
+Some, but not all, jumps can be automatically removed by parsing station logs and identifying dates of equipment changes.
+However, it is still necessary to manually find and remove jumps of unknown origin.
 That being said, this outlier detection algorithm significantly reduces the effort needed to manually clean GNSS data.
@@
 \section{Application to Cascadia Slow Slip Events}\label{sec:Cascadia}
-In this section we estimate transient strain from GNSS data in Cascadia.
-In particular, we are interested in identifying transient strain resulting from slow slip events (SSEs) \citep[e.g.,][]{Dragert2001}.
+We demonstrate our method for estimating transient strain rates with GNSS data from Cascadia.
+We are interested in identifying transient strain resulting from slow slip events (SSEs) \citep[e.g.,][]{Dragert2001}.
 Our results illuminate where elastic strain energy, which can eventually be released seismically, is accumulating during SSEs.
-The estimated strain rates also reveal how strain from SSEs contributes to strain accumulated over seismic cycles.
 We demonstrate that transient strain rates estimated with the method described above are an effective tool for detecting SSEs and potentially other geophysical signal in GNSS data.
@@
 We use publicly available continuous GNSS data from the University Navstar Consortium (UNAVCO) \citep{Herring2016}, which can be found at www.unavco.org.
-We limit the dataset to the stations and time ranges which are pertinent to seven of the most recent SSEs in the Puget Sound region; the earliest SSE began in the August 2010, and the most recent SSE began in the February of 2017.
+We limit the dataset to the stations and time ranges which are pertinent to seven of the most recent SSEs in the Puget Sound region.
+The earliest SSE considered began in August 2010, and the most recent SSE began in February 2017.
 We consider these most recent SSEs because the station coverage is sufficiently dense for us to make well constrained inferences of optimal prior models.
@@
 \caption{Positions of continuous GNSS stations used to estimate transient strain rates.
-The colored regions indicates the density of detected seismic tremor as determined by \citep{Wech2010}.
+The colored regions indicates the density of detected seismic tremor as determined by \citet{Wech2010}.
 The red dots show the positions of GNSS stations mentioned in this paper.
 The magenta dot indicates the location of the strain rates shown in Figure \ref{fig:StrainTs} and \ref{fig:StrainMag}.
-The magenta circle demarcates the spatial extent of the tremors shown in Figure \ref{fig:StrainMag}}.
+The magenta dashed circle demarcates the spatial extent of the tremors shown in Figure \ref{fig:StrainMag}}.
 \label{fig:Context}
@@
 \end{equation}
-For this section, $\bm{d}_*$ is used to denote the displacement time series for just one station, and thus $m = 1$.
+For this section, $\bm{d}_*$ is used to denote the displacement time series for just one station.
 We are also assuming $C_u$ is zero when estimating the noise hyperparameters for this section.
 \citet{Harville1974} showed that choosing the hyperparameters which maximize eq.
-(\ref{eq:REML}) is equivalent to chosing the hyperparameters which maximize the probability of drawing $\bm{d}_*$ from $\bm{d}$.
+(\ref{eq:REML}) is equivalent to choosing the hyperparameters which maximize the probability of drawing $\bm{d}_*$ from $\bm{d}$.
 We use the REML method over the maximum likelihood (ML) method \citep[e.g.,][]{Langbein1997} because the REML method properly accounts for the improper prior that we assigned to $\bm{a}$.
 The distribution of inferred $\alpha$ and $\beta$ are shown in Figure \ref{fig:NoiseParams}.
-The amplitude of FOGM noise for the east and north components is notable low and are clustered around 0.5 mm/yr$^{0.5}$.
-The corresponding estimates of $\alpha$ tend to cluster around 0 yr$^{-1}$, suggesting that Brownian motion would have also been an appropriate noise model.
-The amplitude of FOGM noise for the vertical component is significantly larger with a median value of 13.8 mm/yr$^{0.5}$.
-The inferred values for alpha are higher for the vertical component with a median value of 1.87 yr$^{-1}$.
-In Figure X, we use the median values of $\alpha$ and $\beta$ to generate two random samples of FOGM noise.
-The samples span seven years and over these seven years the east and north samples drift by about 1 mm.
+The amplitude of FOGM noise, $\alpha$, for the east and north components is notable low and are clustered around 0.5 mm/yr$^{0.5}$.
+The corresponding estimates of $\beta$ tend to cluster around 0 yr$^{-1}$, suggesting that Brownian motion would have also been an appropriate noise model.
+The amplitude of FOGM noise for the vertical component is significantly larger with a median value of 13.5 mm/yr$^{0.5}$.
+The inferred values for $\beta$ are higher for the vertical component with a median value of 8.21 yr$^{-1}$.
+In Figure \ref{fig:NoiseSamples}, we use the median values of $\alpha$ and $\beta$ to generate two random samples of FOGM noise.
+The samples span five years and over these five years the east and north samples drift by about 1 mm.
 In the context of detecting SSEs, which produce several mm's of surface displacement on the timescale of weeks, the estimated FOGM noise for the east and north component is negligible.
-In contrast, the estimated FOGM noise for the vertical component is significantly larger than the signal we would expect from SSEs.
+In contrast, the estimated FOGM noise for the vertical component is larger than the signal we would expect from SSEs.
 We suspect that the higher amplitude for the FOGM model in the vertical component is accommodating for deficiencies in our rather simple seasonal model.
-Based on this analysis, we henceforth ignore temporally correlated noise in the east and north component because of its low amplitude, and we do not estimate $u$ for vertical deformation because of the low signal to noise ratio.
+Based on this analysis, we henceforth ignore temporally correlated noise in the east and north component because of its low amplitude.
+We also do not estimate $u$ for vertical deformation because of the low signal to noise ratio.
 
@@
 (\ref{eq:FOGM}).
-Hyperparameters are independently estimated for each of the stations shown in Figure X and for each displacement component.
+Hyperparameters are independently estimated for 38 stations in Washington and Oregon that are east of $121^\circ$ W.
 ``IQR'' is the inter-quartile range.}
@@
 Another significant source of noise in GNSS data is common mode error \citep[e.g.,][]{Wdowinski1997,Dong2006}, which is noise that is highly spatially correlated.
-When not accounted for, common mode error manifests itself as spatially uniform undulations in $\hat{u}$.
+When not accounted for, common mode error manifests as spatially uniform undulations in $\hat{u}$.
 However, we are primarily interested in estimating strain which is insensitive to common mode error.
 We therefore do not include common mode error in our noise model.
-
-For these reasons, we henceforth make the simplifying assumption that temporally and spatially correlated noise is negligible and $v = 0$ for the easting and northing component of GNSS data.
+We then make the simplifying assumption that $v = 0$ for the easting and northing component of GNSS data.
 
 \subsection{Transient displacement model}\label{sec:SignalModel}
-We next discuss our prior model for transient displacements, and, specifically, our choice for the covariance functions $X(x,x')$ and $T(t,t')$.
+We next establish our prior model for transient displacements.
+Specifically, we discuss our choice for the covariance functions $X(x,x')$ and $T(t,t')$.
 For $X$, we use the squared exponential (SE) covariance function,
@@
 The SE covariance function is commonly used in Kriging \citep[e.g,][]{Cressie1992} and Gaussian process regression \citep[e.g.,][]{Rasmussen2006}.
-The SE is a valid (i.e.
-positive definite) covariance function for any number of spatial dimensions, and it describes an isotropic Gaussian processes with realizations that are infinitely differentiable.
-In terms of geodetic applications, \citet{Kato1998} and \cite{El-Fiky1999} demonstrated that the SE is an appropriate covariance model for describing long-term tectonic strain rates in Japan.
+The SE is a valid (i.e., positive definite) covariance function for any number of spatial dimensions, and it describes an isotropic Gaussian processes with realizations that are infinitely differentiable.
+In terms of geodetic applications, \citet{Kato1998} and \cite{El-Fiky1999} demonstrated that the SE covariance function accurately describes secular strain rates in Japan.
 
-We consider three potential covariance functions to describe the temporal covariance of $u$.  First, we consider the one-dimensional SE covariance function, 
+We consider three potential models for the temporal covariance of $u$. First, we consider the one-dimensional SE covariance function, 
 \begin{equation}\label{eq:TimeSE}
@@
 \end{equation}
-simply because the SE covariance function is versatile and commonly used for non-parametric modeling.
+because the SE covariance function is versatile and commonly used for non-parametric modeling.
 Note that $T$ includes the hyperparameter $\phi$, which serves to scale the covariance function $C_u$.
@@
 As an example, Figure \ref{fig:Outliers} shows the outliers which were automatically detected for station SC03, which is located on Mount Olympus in Washington.
-After cleaning the data, we divide it into subsets which are four months long and centered on the time of an SSE.
-The times of seven SSEs are determined using tremor records from \cite{Wech2010}.
+After cleaning the data, we divide it into seven subsets which are four months long and each centered on the time of an SSE.
+The times of seven SSEs are determined with tremor records from \cite{Wech2010}.
 We find the optimal hyperparameters for $T$ and $X$ for each subset of data.
@@
 The estimated hyperparameters for $u$ are summarized in Table 1.
-Based on the interquartile ranges (IQR), the hyperparameters for the SE and Wendland model do not vary significantly between SSEs.
+Based on the interquartile ranges (IQR), the estimated hyperparameters when using the SE and Wendland covariance functions do not vary significantly between SSEs.
 This suggests that the median of estimated hyperparameters should be an appropriate prior model for all Cascadia SSEs.
-For the IBM model, there are several anomalously large values for $\lambda$ and $\phi$, which is they they have large IQRs.
+For the IBM model, there are several anomalously large values for $\lambda$ and $\phi$, which is why they have large IQRs.
 
@@
 \includegraphics{figures/outliers/outliers.pdf}
-\caption{Detrended easting component of GNSS displacement data for station SC03, which is located on Mt.
-Olympus in Washington.
-The orange markers indicate outliers which were automatically detected using the algorithm from Section \ref{sec:Method}.
+\caption{Detrended easting component of GNSS displacement data for station SC03, which is located on Mount Olympus in Washington.
+The orange markers indicate outliers which were automatically detected using the algorithm from Section \ref{sec:Outlier}.
 The error bars show one standard deviation uncertainty.}
@@
 The hyperparameters are estimated for each of the seven SSEs considered in this study, and the tabulated values indicate the median and interquartile ranges of estimates.
-The ``diff $\log$(REML)'' column compares the log REML likelihood to the log REML likelihood for the SE model. Positive values indicate that observations are more consistent with the SE model.} 
+The ``diff $\log$(REML)'' column compares the log REML likelihood to the log REML likelihood for the SE model.
+Positive values indicate that observations are more consistent with the SE model.}
 \end{table}
@@
 In Table 1, we summarize how the log REML likelihoods for the Wendland and IBM covariance functions compare to the SE covariance function.
-Based on the differences in log REML likelihoods, the data is substantially more likely to come form the SE or Wendland covariance functions than from the IBM covariance function.
+Based on the differences in log REML likelihoods, the data is substantially more likely to come from the SE or Wendland covariance functions than from the IBM covariance function.
 The REML likelihoods do not definitively indicate whether the SE or Wendland covariance function is preferable.
-As another test to determine the most appropriate covariance function for $T$, we compare the observations to the predicted displacements for each covariance function.
+To further explore which covariance function for $T$ best describes SSEs, we compare the observations to the predicted displacements for each covariance function.
 let $\hat{\bm{d}} = \left(u(\bm{y}) + \bm{P}\bm{a}\right)|\bm{d}_*$ be the data prediction vector.
@@
 We compute $\hat{\bm{d}}$ using SE, Wendland, and IBM covariance functions for $T$ and the median hyperparameters from Table 1.
-Figure \ref{fig:Fit} compares the easting component of $\bm{d}_*$ to $\hat{\bm{d}}$ for the Winter 2016 SSE.
-For most stations, the data prediction vector appears to appropriately describe displacements throughout the SSE, regardless of the choice of $T$.
+Figure \ref{fig:Fit} compares the easting component of $\bm{d}_*$ to $\hat{\bm{d}}$ for the Winter 2016 SSE at Station P436.
+The data prediction vector appears to accurately describe displacements throughout the SSE, regardless of the choice of $T$.
+Since the SSE signal is strongest at Station P436, it can be assumed that $T$ adequately describes the signal elsewhere.
 The prediction for the IBM model contains slightly more high frequency, and perhaps spurious, features.
@@
 In our estimates of transient strain discussed in the next section, we ultimately settle on the Wendland covariance function for $T$ and use the median values from Table 1 for the hyperparameters.
-We choose the Wendland covariance function over the SE covariance function because of its computational advantages noted above.
+We choose the Wendland covariance function over the SE covariance function because of its computational advantages.
 
@@
 \includegraphics{figures/signal_fit/signal-fit.pdf}
-\caption{Fit.}   
+\caption{Observed displacement at Station P436 and predicted displacements when using different covariance functions for $T$.
+One standard deviation uncertainty is shown for the observations and the predicted displacements when using the SE covariance function.
+For clarity, uncertainties are not shown when using the IBM and Wendland covariance functions.}
 \label{fig:Fit}
@@
 Having established a noise model and a prior for transient displacements, we use the cleaned GNSS dataset described in Section \ref{sec:SignalModel} to calculate transient strain rates in the Puget Sound region.
-We calculate transient strain rate on each day from January 1, 2010 to March 15, 2017.
-The strain rates are estimates on a grid of points spanning the study area.
+We calculate transient strain rate at each day from January 1, 2010 to March 15, 2017.
+The strain rates are estimates at a grid of points spanning the study area.
 In Figure \ref{fig:StrainMap}, we show the transient strain rates on January 1, 2016, which coincides with the height of an SSE.
@@
 The SSEs cause compression in the Olympic Peninsula and extension east of Puget Sound.
-For comparison, long-term strain rates indicate trench perpendicular compression throughout this study region \citep{Murray2000,McCaffrey2007,McCaffrey2013}.
+For comparison, estimated secular strain rates indicate trench perpendicular compression throughout this study region \citep{Murray2000,McCaffrey2007,McCaffrey2013}.
 The SSE are thus concentrating tectonically accumulated strain energy trench-ward, and pushing the subduction zone closer to failure.
@@
 \includegraphics{figures/strain_map/strain-map.pdf}
-\caption{Strain map.}   
+\caption{Estimated transient strain rates during the Winter 2016 SSE.
+Strain glyphs show the normal strain rate along each azimuth, where orange indicates compression and blue indicates extension.
+The semi-transparent regions indicate one standard deviation uncertainties.}
 \label{fig:StrainMap}
@@
 In Figure \ref{fig:StrainTs} we show the time dependence of estimated transient strain rates at a position on the Olympic Peninsula, where strain rate magnitudes from SSEs are largest.
-To verify that the estimated strain rates are accurately identifying geophysical signal, we compare the strain rate magnitude to the frequency of seismic tremor (Figure \ref{fig:StrainMag}).
-The shown strain rate magnitude has been normalized by the estimated uncertainty and can be considered a signal-to-noise ratio.
+To verify that the estimated strain rates are accurately identifying geophysical signal, we compare the strain rate magnitudes to the frequency of seismic tremor (Figure \ref{fig:StrainMag}).
+The shown strain rate magnitudes have been normalized by their uncertainties and can be considered signal-to-noise ratios.
 Estimated strain rate magnitudes greater than ${\sim}3$ can then be interpreted as a detected transient event.
@@
 We are also able to clearly identify transient strain associated with a more subtle inter-SSE event in August 2014.
-In between SSEs the normalized strain rate magnitude is consistently between 0 and 2, indicating that no anomalous non-SSE events are being detected, at least at this location.
+In between SSEs, the normalized strain rate magnitudes are consistently between 0 and 2, indicating that no anomalous non-SSE events are being detected, at least at this location.
 
@@
 \includegraphics{figures/strain_ts/strain-ts.pdf}
-\caption{Strain time series.}   
+\caption{Three components of the transient strain rate tensor estimated at the position shown in Figure \ref{fig:Context}. The semi-transparent regions indicate one standard deviation uncertainty.}   
 \label{fig:StrainTs}
@@
 \includegraphics{figures/strain_ts/mag-ts.pdf}
-\caption{Strain magnitude time series.}   
+\caption{(Top) Strain rate magnitudes at the position shown in Figure \ref{fig:Context} normalized by the one standard deviation uncertainties.
+(bottom) Tremor frequency in the region shown in Figure \ref{fig:Context}.}
 \label{fig:StrainMag}
@@
 
-\section{Other Potential Applications}\label{sec:Other}
-In the previous section we demonstrated that transient strain rates can be used to detect SSEs or other transient geophysical phenomina.
+\section{Discussion}\label{sec:Discussion}
+
+We have demonstrated that transient strain rates estimated with the method described in Section \ref{sec:Method} can be used to detect SSEs or other transient geophysical phenomena.
 Another potential application would be to use the GNSS derived transient strain rates to develop noise models for borehole strain meters (BSMs).
@@
 
+%Numerical Notes/scaling. Use partitioned matrix and Cholmod to solve for strain.
+
 \section{Conclusion}\label{sec:Conclusion}
+In this paper we have demonstrated that Gaussian process regression (GPR) is a powerful tool for estimating transient strain rates from GNSS data.
+Other methods for estimating strain from GNSS rely on parametric representation of deformation, which can bias the estimated strain rates if the parameterization is not chosen carefully.
+In contrast, the GPR method is non-parametric and by the suchoice Our estimated transient strain rates have meaningful uncertainties, which distinguishes it from other methods of estimating strain from GNSS data.
+As a result, we are able to confidently distinguish geophysical signal from noise.
+The key advantage that GPR has over other methods for estimating strain rate, is that GPR
+
 In this paper we describe how Gaussian processes can be used to identifying transient strain from GNSS data.

### 2017-06-16 HinesHetlandGPR.tex
did more shit

@@
 \usepackage{bm}
+\usepackage{setspace}
 \title{Revealing Transient Strain in Geodetic Data with Gaussian Process Regression}
 \author{Trever T. Hines and Eric A. Hetland}
-
-% OUTLINE
-% INTRODUCTION
-
-% - How people have gone about estimating strain rates
-%   - Draw distinction between parametric and nonparametric methods
-%   - Draw distinction between model based and data based parametric methods 
-%   - Talk about Kato
-%
-% - Another, more challenging problem is how to calculate time-dependent strain
-%   - Time dependent strain rates could illuminate transient geophysical signals such as postseismic deformation, 
-%	  slow slip events, or volcanic deformation. 	  
-%	- Since numerous megathrust earthquakes were preceded by slow slip, it is important to detect these features.
-%	- We can also use estimates of transient strain to calibrate strain meters.	
-%	- People who have tackled this problem before include Ohtani and Holt.	
-%	- A major deficiency with these methods is that they are parametric giving poor uncertainties. Without good 
-%	  uncertainties we cannot objectively discern signal from noise and we cannot calibrate BSM.	
-% - Here we propose GPR and we tackle several two major difficulties	
-%   - The method is similar to an extension of Kato to the time domain. We address two debilitating difficulties that 
-%     arise. We also dont do that autocovariance bullshit.  Instead we use REML.  
-%	- outliers  
-%	- dense matrix
-%	
-% - GPR can be used for more than just estimating strain and it is a powerful geophysical tool. We can use it to quantify
-%   spatial noise in GPS data. (this should be a side blurb when talking about quantifying signal.   
-%
-% - We discuss the method, form prior with REML, and demonstrate utility in detecting SSE by comparing it to tremor.	  
-%
-% METHODS
-% - define model
-% - discuss known deficiencies with the model
-% - describe covariance as a time-space separable function 
-% - describe geophysically relevant gaussian processes. (include seasonal and spherical)
-% - REML for determining optimal hyperparameters
-% - conditioning the prior with observations
-% - differentiating the Gaussian process
-% - note on addition, subtraction, and scaling
-% - REML
-% - Numerical notes
-%   - partition inverse and solve with CHOLMOD
-% - outlier detection
-
-% APPLICATION TO CASCADIA SSE
-% - SSE background
-% - REML results
-%   - noise 
-%   - signal for IBM, SE, and WEN
-%   - Plot all covariances functions
-
-% - Strain results
-%   - map view plot
-%   - time series plot
-
-% CONCLUSION
-% - future work
-% - additional applications other than sse. Discuss potential use for noise modeling
 
@@
 \maketitle
+
+\doublespacing
 
@@
 The errors in this assumption manifest as implausibly low formal uncertainties for the estimated strain rates.
-Other methods for estimating secular strain rates have parameterized velocities with bicubic splines \citep{Beavan2001}, spherical wavelets \citep{Tape2009}, and elastostatic Green's functions \citep{Sandwell2016}.
+Other methods for estimating secular strain rates have parameterized GNSS derived velocities with bi-cubic splines \citep{Beavan2001}, spherical wavelets \citep{Tape2009}, and elastostatic Green's functions \citep{Sandwell2016}.
 The choice of parameterization can be subjective.
@@
 For the NSF and NIF, the time dependence of the geophysical signal is modeled as integrated Brownian motion.
-The method described in \citet{Holt2013} was also tested in the transient detection exercise, and they calculate strain rates using a bicubic spatial parameterization of displacements between time epochs.
+The method described in \citet{Holt2013} was also tested in the transient detection exercise, and they calculate strain rates using a bi-cubic spatial parameterization of displacements between time epochs.
 \citet{Holt2013} defined a detection threshold based on the strain rate magnitude, and we will demonstrate that this is indeed an effective criterion for identifying geophysical signal.
@@
 The SSE are thus concentrating tectonically accumulated strain energy trench-ward, and pushing the subduction zone closer to failure.
-Similar conclusions have been drawn based on fault slip models \citep[e.g.,][]{Dragert2001,Wech2009,Schmidt2010}, which reveal that SSEs are occurring downdip of the seismogenic zone.
-The SSEs, which are thrust events, migrate stress updip and push the locked portion of the fault closer to failure.
+Similar conclusions have been drawn based on fault slip models \citep[e.g.,][]{Dragert2001,Wech2009,Schmidt2010}, which reveal that SSEs are occurring down-dip of the seismogenic zone.
+The SSEs, which are thrust events, migrate stress up-dip and push the locked portion of the fault closer to failure.
 A key difference between the strain inferred here and strain derived from fault slip models is that our estimated strain rates are not based on an assumed physical model.
@@
 \section{Conclusion}\label{sec:Conclusion}
-In this paper we have demonstrated that Gaussian process regression (GPR) is a powerful tool for estimating transient strain rates from GNSS data.
-Other methods for estimating strain from GNSS rely on parametric representation of deformation, which can bias the estimated strain rates if the parameterization is not chosen carefully.
-In contrast, the GPR method is non-parametric and by the suchoice Our estimated transient strain rates have meaningful uncertainties, which distinguishes it from other methods of estimating strain from GNSS data.
-As a result, we are able to confidently distinguish geophysical signal from noise.
-The key advantage that GPR has over other methods for estimating strain rate, is that GPR
-
-In this paper we describe how Gaussian processes can be used to identifying transient strain from GNSS data.
-Our approach is Bayesian and we use a spatio-temporal stochastic model to describe the prior for transient deformation and the noise in GNSS data.
-Our estimated transient strain rates are complete with meaningful estimates of uncertainties, which allows us to objectively discern geophysical signal from noise.
-Other commonly used methods for estimate long-term and transient strain assume a parametric form for the underlying signal, and errors in this assumption will result in a biased solution.
-In contrast our stochastic prior model is chosen with maximum likelihood methods, which is an objective and data-driven approach.
-The lack of subjectivity in our method gives us confidence that the results are unbiased.
-
-We have demonstrated with GNSS data from Cascadia that Gaussian processes are effective for identifying transient strain.
-We also envision that Gaussian processes can be an effective tool to handle other geophysical things.
-
-used to tackle other several other closely related applications
-
-Since our method lacks subjectivity, the obto be most consistent with the data  maximum likelihood methods.
-that is, constrained by the observed data.
-and is less likely the results and lead to this and We prefer a stochastic prior model over a parametric prior model such
-
-The uncertainties on the estimate of transient strain can be trusted, provided that we have chosen appropriate stochastic models for the prior and noise.
-These models are objectively chosen with on maximum likelihoods methods, and
-
-A stochastic prior model is preferable to a As a result, we obtain sensible
+In this paper we have discussed how Gaussian process regression (GPR) can be used to estimate transient strain rates from GNSS data.
+Most other methods for estimating strain from GNSS data assume a parametric representation of deformation, which can bias the estimated strain rates if the parameterization is not chosen carefully.
+The method described in this paper assumes a stochastic prior model for displacements, which can be chosen objectively with the Restricted Maximum Likelihood method.
+Because GPR is a Bayesian method, the uncertainties on our estimated transient strain rates are well quantified and we can discern geophysical signal from noise.
+This is demonstrated in our application to Cascadia, where we clearly detect transient strain resulting from slow slip events.
+GPR is not robust against outliers, which are common in GNSS data, and so we have also described an effective pre-processing method for identifying and removing outliers from GNSS data sets.
+Another complication with GPR is that it usually involves inverting a dense covariance matrix with rank equal to the number of observations.
+This can be prohibitive when using years of daily GNSS observations from a network of several hundred stations.
+We overcome this complication by using compact Wendland covariance function for our prior model, which allows us to utilize sparse matrix software.
+We believe that Gaussian process regression is a powerful tool that could be applied to a wider range of geophysical problems than what was considered in this paper.
 
@@
 
-The material presented in this paper is primarily focused on GPS data; however, we speculate that the RBF-FD scheme can be of particular use in denoising borehole strain meter (BSM) data.
+%The material presented in this paper is primarily focused on GPS data; however, we speculate that the RBF-FD scheme can be of particular use in denoising borehole strain meter (BSM) data.
 The Plate Boundary Observatory has deployed X BSMs along the Western United States.
@@
 
-Additional uses for strain: transient detection, prior for strain meters,
+%Additional uses for strain: transient detection, prior for strain meters,
 
-Additional potential applications of the RBF-FD method: regularizing inverse problems,
+%Additional potential applications of the RBF-FD method: regularizing inverse problems,
 

### 2017-06-16 HinesHetlandGPR.tex
added more

@@
-\documentclass[10pt,a4paper]{article}
+\documentclass[10pt,letter]{article}
 \usepackage{amsmath}
@@
 
-\doublespacing
+%\doublespacing
 
@@
 Knowing where and how quickly strain is accumulating gives insight into where we can expect stored elastic energy to be released seismically.
-Consequently, secular crustal strain rates, which are estimated from GNSS data, have been used to constrain seismic hazard models such as UCERF3 \citep{Field2014}.
+Consequently, secular crustal strain rates, which can be estimated from GNSS data, have been used to constrain seismic hazard models such as UCERF3 \citep{Field2014}.
 Dense networks of continuous GNSS stations, such as the Plate Boundary Observatory (PBO), make it feasible to estimate transient crustal strain, which we consider to be any deviation from the secular strain rate.
@@
 Other methods for estimating secular strain rates have parameterized GNSS derived velocities with bi-cubic splines \citep{Beavan2001}, spherical wavelets \citep{Tape2009}, and elastostatic Green's functions \citep{Sandwell2016}.
-The choice of parameterization can be subjective.
+The type of basis functions and the number of degrees of freedom for a parameterization can be subjective.
 If there are too few degrees of freedom in the parameterization then estimated strain rates will be biased and the uncertainties will be underestimated.
 If there are too many degrees of freedom then there will not be any coherent features in the estimated strain rates.
+The methods described by \citet{Beavan2001}, \citet{Tape2009}, and \citet{Sandwell2016} also require the user to specify penalty parameters that control a similar trade-off between bias and variance in the solution.
 One could parameterize deformation with a physically motivated model of interseismic deformation \citep[e.g.,][]{Meade2005,McCaffrey2007}.
@@
 \citet{Holt2013} defined a detection threshold based on the strain rate magnitude, and we will demonstrate that this is indeed an effective criterion for identifying geophysical signal.
-One challenge that is common to all these methods is that it is difficult to distinguish signal from noise, since the uncertainty on the inferred deformation can be significantly influenced by the choice of spatial parameterization.
+For the same reasons descibed above, the transient deformation and corresponding uncertainties estimated by these methods can be biased by the chosen spatial parameterization.
+It is then difficult to distinguish signal from noise with these methods, which limits their utility for transient detection.
 
 In this paper we use Gaussian process regression (GPR) \citep{Rasmussen2006} to infer transient strain from GNSS data.
-GPR is a Bayesian, non-parametric method for inferring a continuous signal from discrete and potentially noisy data.
+GPR is a Bayesian, non-parametric method for inferring a continuous signal from scattered and potentially noisy data.
+GPR is ideal for synthesizing GNSS data into a spatially and temporally continuous represention of surface deformation, since GNSS stations are irregularly spaced and observation times may differ between stations.
 GPR is closely related to Kriging \citep{Cressie1992} and Least Squares Collocation \citep{Kato1998}, and it has recently gained popularity as a tool for machine learning.
@@
 A Gaussian process is a normally distributed, continuous, stochastic process that is fully defined in terms of a mean function and a positive-definite covariance function.
-For example, Brownian motion is a well known Gaussian process in $\mathbb{R}^1$ which has zero mean and covariance function $C(t,t')=\min(t,t')$, where $t,t' \ge 0$.
-We use the Restricted Maximum Likelihood (REML) method to guide our choice for the prior Gaussian process.
-We then incorporate the GNSS observations with the prior to form a posterior estimate of transient strain rates.
-The posterior transient strain rate is also a Gaussian process and, provided that our prior is sensible, has well quantified uncertainties.
-With well quantified uncertainties we can more confidently discern geophysical signal from noise in our estimates of transient strain rates.
-We use GPR to infer transient strain rates resulting slow slip events (SSEs) in Cascadia.
+For example, Brownian motion is a well known Gaussian process in $\mathbb{R}^1$ which has zero mean and covariance function $\mathrm{cov}(t,t')=\min(t,t')$, where $t,t' \ge 0$.
+If no prior information is available for the geophysical signal, then maximum likelihood methods can be used to choose a prior Gaussian process which is most consistent with the observations.
+We incorporate the GNSS observations with the prior to form a posterior estimate of transient strain.
+The posterior transient strain is also a Gaussian process and, provided that our prior is sensible, has well quantified uncertainties.
+With well quantified uncertainties we can more confidently discern geophysical signal from noise in our estimates of transient strain.
+We use GPR to infer transient strain resulting from slow slip events (SSEs) in Cascadia.
 Our results demonstrate that GPR can be an effective tool for detecting SSEs and revealing the spatio-temporal evolution of crustal strain during SSEs.
@@
 In the following analysis, we make the quite arbitrary choice that $t$ is zero on the first epoch of $\bm{d}_*$.
+Using an earlier reference time does not change the results discussed in this section.
 Our third option for $T$ is the Wendland class of covariance functions \citep{Wendland2005}.
@@
 
+% We could have also used a spherical spatial covariance functions
+
 %Numerical Notes/scaling. Use partitioned matrix and Cholmod to solve for strain.

### 2017-06-17 HinesHetlandGPR.tex
did more shit

@@
 \author{Trever T. Hines and Eric A. Hetland}
+
+\newcommand*{\du}[1]{\underline{\underline{#1}}}
 
@@
 \section{Estimating Transient Strain Rates}\label{sec:Method}
-We denote transient crustal strain rates as $\dot{\bm{\varepsilon}}(z)$, where $z$ is the ordered pair $(x,t)$, $x$ are spatial coordinates in $\mathbb{R}^2$, and $t$ is time.
-We consider $\dot{\bm{\varepsilon}}$ to be spatially and temporally coherent deviations from the steady rate of strain accumulation from plate tectonics.
-We determine $\dot{\bm{\varepsilon}}$ by spatially and temporally differentiating estimates of transient displacements, $\bm{u}(z)$.
-As we discuss in Section \ref{sec:NoiseModel}, estimates of $\dot{\bm{\varepsilon}}$ turn out to be more effective at illuminating geophysical signal than estimates of $\bm{u}$ or $\dot{\bm{u}}$ because $\dot{\bm{\varepsilon}}$ is insensitive to common mode noise in GNSS data.
-We make a prior assumption that each component of $\bm{u}$ is a Gaussian process,
+We seek a spatially and temporally dependent estimate of transient crustal strain rate, denoted as $\dot{\du{\varepsilon}}(\zeta)$.
+We use $\zeta$ to represent the ordered pair $(\vec{x},t)$, where $\vec{x}$ are spatial coordinates in $\mathbb{R}^2$, and $t$ is time.
+We consider $\dot{\du{\varepsilon}}(\zeta)$ to be spatially and temporally coherent deviations from the steady rate of strain accumulation from plate tectonics.
+We determine $\dot{\du{\varepsilon}}(\zeta)$ by spatially and temporally differentiating estimates of transient displacements, $\vec{u}(\zeta)$.
+As we discuss in Section \ref{sec:NoiseModel}, estimates of $\dot{\du{\varepsilon}}$ turn out to be more effective at illuminating geophysical signal than estimates of $\vec{u}(\zeta)$ because $\dot{\du{\varepsilon}}$ is insensitive to common mode noise in GNSS data.
+We make a prior assumption that each component of $\vec{u}$ is a Gaussian process,
 \begin{equation}\label{eq:TransientDeformation}
-u_i(z) \sim \mathcal{N}\left(0,C_{u_i}\right),
+u_i(\zeta) \sim \mathcal{N}\left(0,C_{u_i}\right),
 \end{equation}
-where $C_{u_i}(z,z')$ is a covariance function indicating how we expect $u_i(z)$ to covary with $u_i(z')$.
+where $C_{u_i}(\zeta,\zeta')$ is a covariance function indicating how we expect $u_i(\zeta)$ to covary with $u_i(\zeta')$.
 For simplicity, we treat each component of displacement independently and ignore any potential covariance.
-We then drop the component subscripts with the understanding that the same analysis is being repeated to estimate the east, north, and vertical components of $\bm{u}$.
-We further assume that $C_u$ can be separated into positive definite spatial and temporal functions as
+We henceforth drop the component subscripts with the understanding that the same analysis is being repeated to estimate the east, north, and vertical components of $\vec{u}$.
+We assume that $C_u$ can be separated into positive definite spatial and temporal functions as
 \begin{equation}\label{eq:TransientCovariance}
-C_{u}\left((x,t),(x',t')\right) = X(x,x')T(t,t'),
+C_{u}\left((\vec{x},t),(\vec{x}',t')\right) = X(\vec{x},\vec{x}')T(t,t'),
 \end{equation}  
@@
 We constrain $u$ with GNSS data, which records $u$ as well as other physical and non-physical processes which we are not interested in.
-We describe GNSS observations at position $x_i$ and time $t_j$ as a realization of the random variable
+We describe GNSS observations at position $\vec{x}_i$ and time $t_j$ as a realization of the random variable
 \begin{align}\label{eq:Data}
 \begin{split}
-d_{ij} = &u(x_i,t_j) + v(x_i,t_j) + w_{ij} + a^{(1)}_i + a^{(2)}_it_j + \\
+d_{ij} = &u(\vec{x}_i,t_j) + \eta(\vec{x}_i,t_j) + w_{ij} + a^{(1)}_i + a^{(2)}_it_j + \\
          &a^{(3)}_i\sin(2 \pi t_j) + a^{(4)}_i\cos(2 \pi t_j) + a^{(5)}_i\sin(4 \pi t_j) + a^{(6)}_i\cos(4 \pi t_j), 
@@
 \end{align}
-where $a^{(1)}_{i}$ is an offset that is unique for each GNSS monument, $a^{(2)}_{i}$ is the steady rate of tectonic deformation at $x_i$, and the sinusoids describe seasonal deformation (using units of years for $t$).
+where $a^{(1)}_{i}$ is an offset that is unique for each GNSS monument, $a^{(2)}_{i}$ is the steady rate of tectonic deformation at $\vec{x}_i$, and the sinusoids describe seasonal deformation (using units of years for $t$).
 We use $w_{ij}$ to denote normally distributed, uncorrelated noise.
-Correlated noise which does not have a parametric representation is denoted by $v$.
-For example, $v$ can consist of temporally correlated noise describing benchmark wobble \citep[e.g.,][]{Wyatt1982,Wyatt1989}, and/or spatially correlated noise describing common mode error \citep[e.g.,][]{Wdowinski1997}.
-For now, we will only assume that $v \sim \mathcal{N}(0,C_v)$.
+Correlated noise which does not have a parametric representation is denoted by $\eta$.
+For example, $\eta$ can consist of temporally correlated noise describing benchmark wobble \citep[e.g.,][]{Wyatt1982,Wyatt1989}, and/or spatially correlated noise describing common mode error \citep[e.g.,][]{Wdowinski1997}.
+For now, we will only assume that $\eta \sim \mathcal{N}(0,C_\eta)$.
 We consider the six coefficients in eq.
@@
 We now consider the column vector of $n$ GNSS observations, $\bm{d}_*$.
-Let $\bm{y}$ be the set of $(x_i, t_j)$ pairs describing where and when each of the GNSS observations have been made.
+Let $\bm{\xi}$ be the set of $(\vec{x}_i, t_j)$ pairs describing where and when each of the GNSS observations have been made.
 Let $\bm{a}$ be the vector of coefficients from eq.
 (\ref{eq:Data}) for each of the $m$ GNSS stations.
-We use $\bm{P}$ to represent the $n \times 6m$ matrix of corresponding basis functions evaluated at each point in $\bm{y}$.
+We use $\bm{G}$ to represent the $n \times 6m$ matrix of corresponding basis functions evaluated at each point in $\bm{\xi}$.
 We  also denote the vector of uncorrelated noise for each observation as $\bm{w}$, whose standard deviations are given by the formal data uncertainty, $\bm{\sigma}$.
@@
 \begin{equation}
-\bm{d} = u(\bm{y}) + v(\bm{y}) + \bm{w} + \bm{P}\bm{a},
+\bm{d} = u(\bm{\xi}) + \eta(\bm{\xi}) + \bm{w} + \bm{G}\bm{a},
 \end{equation}
-which is distributed as $\mathcal{N}(\bm{0},\bm{\Sigma} + \kappa^2\bm{P}\bm{P}^T)$, where
+which is distributed as $\mathcal{N}(\bm{0},\bm{\Sigma} + \kappa^2\bm{G}\bm{G}^T)$, where
 \begin{equation}\label{eq:Cd}
-\bm{\Sigma} = C_u(\bm{y},\bm{y}) + C_v(\bm{y},\bm{y}) + 
+\bm{\Sigma} = C_u(\bm{\xi},\bm{\xi}) + C_\eta(\bm{\xi},\bm{\xi}) + 
               \mathrm{diag}\left(\bm{\sigma}^2\right).  
 \end{equation}
-It should be understood that the notation $f(\bm{x})$ and $f(\bm{x},\bm{y})$ is shorthand for $[f(x_i)]_{x_i \in \bm{x}}$ and $[f(x_i,y_j)]_{(x_i,y_j) \in \bm{x} \times \bm{y}}$, respectively.
+It should be understood that notation such as $u(\bm{\xi})$ and $C_u(\bm{\xi},\bm{\xi})$ is shorthand for $[u(\xi_i)]_{\xi_i \in \bm{\xi}}$ and $[C_u(\xi_i,\xi_j)]_{(\xi_i,\xi_j) \in \bm{\xi} \times \bm{\xi}}$, respectively.
 
 The prior for each component of transient displacements is then conditioned with $\bm{d}_*$ to form a posterior estimate of transient displacements, $\hat{u} = u | \bm{d}_*$.
-We will assume that an appropriate covariance functions and corresponding hyperparameters for $X$, $T$, and $C_v$ have already been chosen.
+We will assume that an appropriate covariance functions and corresponding hyperparameters for $X$, $T$, and $C_\eta$ have already been chosen.
 If $\kappa$ is kept finite, then we can follow \citet{Rasmussen2006} to find that $\hat{u}$ is distributed as $\mathcal{N}(\mu_{\hat{u}},C_{\hat{u}})$, where
 \begin{equation}\label{eq:PosteriorMean}
-\mu_{\hat{u}}(z) = C_u(z,\bm{y})\left(\bm{\Sigma} + \kappa^2\bm{P}\bm{P}^T\right)^{-1}\bm{d}_*
+\mu_{\hat{u}}(\zeta) = C_u(\zeta,\bm{\xi})\left(\bm{\Sigma} + \kappa^2\bm{G}\bm{G}^T\right)^{-1}\bm{d}_*
 \end{equation}    
@@
 \begin{equation}\label{eq:PosteriorCov}
-C_{\hat{u}}(z,z') = C_u(z,z') - C_u(z,\bm{y})\left(\bm{\Sigma} + \kappa^2\bm{P}\bm{P}^T\right)^{-1}C_u(\bm{y},z').
+C_{\hat{u}}(\zeta,\zeta') = C_u(\zeta,\zeta') - C_u(\zeta,\bm{\xi})\left(\bm{\Sigma} + \kappa^2\bm{G}\bm{G}^T\right)^{-1}C_u(\bm{\xi},\zeta').
 \end{equation}
@@
  \begin{equation}\label{eq:PosteriorMean2}
-\mu_{\hat{u}}(z) = \left[\begin{array}{cc}
-                         C_u(z,\bm{y}) & \bm{0} \\
+\mu_{\hat{u}}(\zeta) = \left[\begin{array}{cc}
+                         C_u(\zeta,\bm{\xi}) & \bm{0} \\
                          \end{array}\right]
                    \left[\begin{array}{cc}
-                         \bm{\Sigma} & \bm{P} \\
-                         \bm{P}^T  & -\kappa^{-2} \bm{I} \\
+                         \bm{\Sigma} & \bm{G} \\
+                         \bm{G}^T  & -\kappa^{-2} \bm{I} \\
                          \end{array}\right]^{-1}
@@
 \begin{equation}\label{eq:PosteriorCov2}
-C_{\hat{u}}(z,z') = C_u(z,z') - 
+C_{\hat{u}}(\zeta,\zeta') = C_u(\zeta,\zeta') - 
                     \left[\begin{array}{cc}
-                          C_u(z,\bm{y}) & \bm{0} \\
+                          C_u(\zeta,\bm{\xi}) & \bm{0} \\
                           \end{array}\right]
                     \left[\begin{array}{cc}
-                          \bm{\Sigma} & \bm{P} \\
-                          \bm{P}^T  & -\kappa^{-2} \bm{I} \\
+                          \bm{\Sigma} & \bm{G} \\
+                          \bm{G}^T  & -\kappa^{-2} \bm{I} \\
                           \end{array}\right]^{-1}
                     \left[\begin{array}{c}
-                          C_u(\bm{y},z') \\
+                          C_u(\bm{\xi},\zeta') \\
                           \bm{0} \\
@@
  \begin{equation}\label{eq:PosteriorMean3}
-\mu_{\hat{u}}(z) = \left[\begin{array}{cc}
-                         C_u(z,\bm{y}) & \bm{0} \\
+\mu_{\hat{u}}(\zeta) = \left[\begin{array}{cc}
+                         C_u(\zeta,\bm{\xi}) & \bm{0} \\
                          \end{array}\right]
                    \left[\begin{array}{cc}
-                         \bm{\Sigma} & \bm{P} \\
-                         \bm{P}^T  & \bm{0} \\
+                         \bm{\Sigma} & \bm{G} \\
+                         \bm{G}^T  & \bm{0} \\
                          \end{array}\right]^{-1}
@@
 \begin{equation}\label{eq:PosteriorCov3}
-C_{\hat{u}}(z,z') = C_u(z,z') - 
+C_{\hat{u}}(\zeta,\zeta') = C_u(\zeta,\zeta') - 
                     \left[\begin{array}{cc}
-                          C_u(z,\bm{y}) & \bm{0} \\
+                          C_u(\zeta,\bm{\xi}) & \bm{0} \\
                           \end{array}\right]
                     \left[\begin{array}{cc}
-                          \bm{\Sigma} & \bm{P} \\
-                          \bm{P}^T  & \bm{0} \\
+                          \bm{\Sigma} & \bm{G} \\
+                          \bm{G}^T  & \bm{0} \\
                           \end{array}\right]^{-1}
                     \left[\begin{array}{c}
-                          C_u(\bm{y},z') \\
+                          C_u(\bm{\xi},\zeta') \\
                           \bm{0} \\
@@
 
-The posterior transients displacements, $\hat{\bm{u}}$, are found by evaluating eq.
-(\ref{eq:PosteriorMean3}) and (\ref{eq:PosteriorCov3}) for each displacement component.
-We differentiate the Gaussian process $\hat{\bm{u}}$ (see \cite{Abrahamsen1997}) to form an estimate of transient strain rate, $\dot{\bm{\varepsilon}}$, which is itself a spatially and temporally continuous Gaussian process.
-The components of the transient strain rate tensor, $\dot{\varepsilon}_{ij}$, are given by
+We use eq.
+(\ref{eq:PosteriorMean3}) and (\ref{eq:PosteriorCov3}) to find the posterior east and north components of transient displacements.
+The posterior transient displacements are differentiated temporally and spatially to form an estimate of $\dot{\du{\varepsilon}}$.
+The components of $\dot{\du{\varepsilon}}$ are given by
 \begin{equation}\label{eq:StrainRate}
-\dot{\varepsilon}_{ij}(z) = \frac{1}{2} \frac{\partial}{\partial t} \left(
-                                        \frac{\partial \hat{u}_i(z)}{\partial x_j} +  
-                                        \frac{\partial \hat{u}_j(z)}{\partial x_i}\right).
+\dot{\varepsilon}_{ij}(\zeta) = \frac{1}{2} \frac{\partial}{\partial t} \left(
+                                        \frac{\partial \hat{u}_i(\zeta)}{\partial x_j} +  
+                                        \frac{\partial \hat{u}_j(\zeta)}{\partial x_i}\right),
 \end{equation}
-which are distributed as $\mathcal{N}(\mu_{\dot{\varepsilon}_{ij}},C_{\dot{\varepsilon}_{ij}})$, where
+which are Gaussian processes with mean functions
 \begin{equation}\label{eq:StrainMean}
-\mu_{\dot{\varepsilon}_{ij}}(z) = \frac{1}{2}\frac{\partial}{\partial t}\left(
-                                  \frac{\partial \mu_{\hat{u}_i}(z)}{\partial x_j} + 
-                                  \frac{\partial \mu_{\hat{u}_j}(z)}{\partial x_i} \right)
+\mu_{\dot{\varepsilon}_{ij}}(\zeta) = \frac{1}{2}\frac{\partial}{\partial t}\left(
+                                  \frac{\partial \mu_{\hat{u}_i}(\zeta)}{\partial x_j} + 
+                                  \frac{\partial \mu_{\hat{u}_j}(\zeta)}{\partial x_i} \right)
 \end{equation} 
-and  
+and covariance functions
 \begin{equation}\label{eq:StrainCov}
-C_{\dot{\varepsilon}_{ij}}(z,z') = \frac{1}{4} \frac{\partial^2}{\partial t \, \partial t'}\left(
-                                   \frac{\partial^2 C_{\hat{u}_i}(z,z')}{\partial x_j \, \partial x'_j} + 
-                                   \frac{\partial^2 C_{\hat{u}_j}(z,z')}{\partial x_i \, \partial x'_i} \right).
+C_{\dot{\varepsilon}_{ij}}(\zeta,\zeta') = \frac{1}{4} \frac{\partial^2}{\partial t \, \partial t'}\left(
+                                   \frac{\partial^2 C_{\hat{u}_i}(\zeta,\zeta')}{\partial x_j \, \partial x'_j} + 
+                                   \frac{\partial^2 C_{\hat{u}_j}(\zeta,\zeta')}{\partial x_i \, \partial x'_i} \right).
 \end{equation} 
+In the above equations we use $\hat{u}_i$ to denote posterior transient displacements along direction $i$ and $x_i$ to represent the components of $\vec{x}$.
 
@@
 We calculate the residuals between the observations and a best fitting model, and data with residuals that are anomalously large are identified as outliers.
-We treat $\bm{d}_*$ as a sample of $\bm{d}$ and assume that there is no temporally correlated noise (i.e., $v = 0$).
-The best fitting model for $\bm{d}_*$ is considered to be the expected value of the random vector $u(\bm{y}) + \bm{P}\bm{a}$ after conditioning it with non-outlier observations.
+We treat $\bm{d}_*$ as a sample of $\bm{d}$ and assume that there is no temporally correlated noise (i.e., $\eta(\zeta) = 0$).
+The best fitting model for $\bm{d}_*$ is considered to be the expected value of the random vector $u(\bm{\xi}) + \bm{G}\bm{a}$ after conditioning it with non-outlier observations.
 We still consider $u$ to have a separable covariance function as in eq.
@@
 \begin{align}\label{eq:Residual}
-\bm{r} &= \frac{\bm{d}_* - \mathrm{E}\left[(u(\bm{y}) + \bm{P}\bm{a})|\tilde{\bm{d}}_* \right]}{\bm{\sigma}} \\
+\bm{r} &= \frac{\bm{d}_* - \mathrm{E}\left[(u(\bm{\xi}) + \bm{G}\bm{a})|\tilde{\bm{d}}_* \right]}{\bm{\sigma}} \\
        &= \frac{1}{\bm{\sigma}}\left(\bm{d}_*  - 
           \left[\begin{array}{cc}
-                C_u(\bm{y},\tilde{\bm{y}}) & \bm{P} \\
+                C_u(\bm{\xi},\tilde{\bm{\xi}}) & \bm{G} \\
                 \end{array}\right]
           \left[\begin{array}{cc}
-                C_u(\tilde{\bm{y}},\tilde{\bm{y}}) + \mathrm{diag}(\tilde{\bm{\sigma}}^2) & \tilde{\bm{P}} \\
-                \tilde{\bm{P}}^T  & \bm{0} \\
+                C_u(\tilde{\bm{\xi}},\tilde{\bm{\xi}}) + \mathrm{diag}(\tilde{\bm{\sigma}}^2) & \tilde{\bm{G}} \\
+                \tilde{\bm{G}}^T  & \bm{0} \\
                 \end{array}\right]^{-1}
@@
 \end{align}
-where the tilde indicates that only elements corresponding to indices in $\Omega$ are retained (e.g., $\tilde{\bm{y}} = \{y_i\}_{i\in\Omega}$). We then update $\Omega$ to be
+where the tilde indicates that only elements corresponding to indices in $\Omega$ are retained (e.g., $\tilde{\bm{\xi}} = \{y_i\}_{i\in\Omega}$). We then update $\Omega$ to be
 \begin{equation}\label{eq:Update}
-\Omega = \{i : |r_i| < \eta \tilde{\bm{r}}_{rms}\}, \ \ \ r_i \in \bm{r}
+\Omega = \{i : |r_i| < \lambda \tilde{\bm{r}}_{rms}\}, \ \ \ r_i \in \bm{r}
 \end{equation} 
-where $\tilde{\bm{r}}_{rms}$ is the root-mean-square of $\tilde{\bm{r}}$ and $\eta$ is an outlier tolerance.
-We use $\eta=4$ in this study, which seems to accurately identify outliers without unnecessarily decimating the data.
+where $\tilde{\bm{r}}_{rms}$ is the root-mean-square of $\tilde{\bm{r}}$ and $\lambda$ is an outlier tolerance.
+We use $\lambda=4$ in this study, which seems to accurately identify outliers without unnecessarily decimating the data.
 Iterations continue until the new $\Omega$ is equal to the previous $\Omega$.
@@
 \subsection{Noise model}\label{sec:NoiseModel}
-Before we determine the transient strain rates, we must establish a prior for the transient deformation, $u$, and the noise, $v$.
-In this section we discuss our choice for the noise covariance function $C_v$.
+Before we determine the transient strain rates, we must establish a prior for the transient deformation, $u$, and the noise, $\eta$.
+In this section we discuss our choice for the noise covariance function $C_\eta$.
 There have been numerous studies on temporally correlated noise in GNSS data \citep[e.g.,][]{Zhang1997,Mao1999,Williams2004,Langbein2008}.
@@
 There is some physical justification for using Brownian motion as a noise model because it accurately describes the power spectrum of motion resulting from instability in geodetic monuments \citep[e.g.,][]{Wyatt1982,Wyatt1989,Langbein1997}.
-Here we describe the time dependence of $v$ as a FOGM process and consider $v$ to be spatially uncorrelated.
+Here we describe the time dependence of $\eta$ as a FOGM process and consider $\eta$ to be spatially uncorrelated.
 A FOGM process is a solution to the stochastic differential equation
 \begin{equation}\label{eq:FOGMdef}
-\dot{v}(t) + \alpha v(t) = \beta w(t),
+\dot{\eta}(t) + \alpha \eta(t) = \beta w(t),
 \end{equation}
 where $w(t)$ is white noise with unit variance.
-The FOGM process degenerates to the commonly used Brownian motion noise model under the condition that $\alpha=0$ and $v(0) = 0$, which is evident from the definition of Brownian motion \citep[e.g.,][]{Papoulis1991}.
+The FOGM process degenerates to the commonly used Brownian motion noise model under the condition that $\alpha=0$ and $\eta(0) = 0$, which is evident from the definition of Brownian motion \citep[e.g.,][]{Papoulis1991}.
 Our noise model which satisfies eq.
@@
 \begin{equation}\label{eq:FOGM}
-C_v\left((x,t),(x',t')\right) = \frac{\beta^2}{2\alpha}\exp\left(-\alpha|t - t'|\right) \delta(||x - x'||_2). 
+C_\eta\left((\vec{x},t),(\vec{x}',t')\right) = \frac{\beta^2}{2\alpha}\exp\left(-\alpha|t - t'|\right) \delta(||\vec{x} - \vec{x}'||_2). 
 \end{equation}
 
-We constrain the hyperparameters for $v$, $\alpha$ and $\beta$, with a set of 38 continuous GNSS stations in Washington and Oregon which are east of 121$^\circ$ W.
+We constrain the hyperparameters for $\eta$, $\alpha$ and $\beta$, with a set of 38 continuous GNSS stations in Washington and Oregon which are east of 121$^\circ$ W.
 These stations are sufficiently far from the subduction zone that they are unlikely to contain transient signal.
@@
 \begin{equation}\label{eq:REML}
-\mathcal{L}(\bm{\theta}) = \left(\frac{\left|\bm{P}^T\bm{P}\right|}
+\mathcal{L}(\bm{\theta}) = \left(\frac{\left|\bm{G}^T\bm{G}\right|}
                            {(2\pi)^{n-6m} 
                            \left| \bm{\Sigma}(\bm{\theta}) \right| 
-                           \left| \bm{P}^T\bm{\Sigma}\bm{P} \right|}\right)^{\frac{1}{2}} 
+                           \left| \bm{G}^T\bm{\Sigma}\bm{G} \right|}\right)^{\frac{1}{2}} 
                            e^{-\tfrac{1}{2}\bm{d}_*^T\bm{K}(\bm{\theta})\bm{d}_*},
@@
 \bm{K}(\bm{\theta}) = \bm{\Sigma}(\bm{\theta})^{-1} - 
-                      \bm{\Sigma}(\bm{\theta})^{-1}\bm{P}
-         \left(\bm{P}^T\bm{\Sigma}(\bm{\theta})^{-1}\bm{P}\right)^{-1}
-         \bm{P}^T\bm{\Sigma}(\bm{\theta})^{-1}.
+                      \bm{\Sigma}(\bm{\theta})^{-1}\bm{G}
+         \left(\bm{G}^T\bm{\Sigma}(\bm{\theta})^{-1}\bm{G}\right)^{-1}
+         \bm{G}^T\bm{\Sigma}(\bm{\theta})^{-1}.
 \end{equation}
 For this section, $\bm{d}_*$ is used to denote the displacement time series for just one station.
-We are also assuming $C_u$ is zero when estimating the noise hyperparameters for this section.
+We are also assuming $C_u(\zeta,\zeta')=0$ when estimating the noise hyperparameters for this section.
 \citet{Harville1974} showed that choosing the hyperparameters which maximize eq.
@@
 We therefore do not include common mode error in our noise model.
-We then make the simplifying assumption that $v = 0$ for the easting and northing component of GNSS data.
+We then make the simplifying assumption that $\eta(\zeta) = 0$ for the easting and northing component of GNSS data.
 
@@
 We next establish our prior model for transient displacements.
-Specifically, we discuss our choice for the covariance functions $X(x,x')$ and $T(t,t')$.
+Specifically, we discuss our choice for the covariance functions $X(\vec{x},\vec{x}')$ and $T(t,t')$.
 For $X$, we use the squared exponential (SE) covariance function,
 \begin{equation}\label{eq:SE}
-X(x,x') = \exp\left(\frac{-||x - x'||_2^2}{2 \lambda^2}\right).
+X(\vec{x},\vec{x}') = \exp\left(\frac{-||\vec{x} - \vec{x}'||_2^2}{2 \ell^2}\right).
 \end{equation}
@@
 \begin{equation}\label{eq:TimeSE}
-T(t,t') = \phi^2\exp\left(\frac{-|t - t'|^2}{\theta^2}\right),
+T(t,t') = \phi^2\exp\left(\frac{-|t - t'|^2}{\tau^2}\right),
 \end{equation}
@@
 \begin{align}\label{eq:IBM}
-T(t,t') &= \int_0^t \int_0^{t'} \phi^2 \min(\tau,\tau') \,d\tau'\,d\tau \\
+T(t,t') &= \int_0^t \int_0^{t'} \phi^2 \min(s,s') \,ds'\,ds \\
         &= \frac{\phi^2}{2}\min(t,t')^2 \left(\max(t,t') - \frac{1}{3}\min(t,t')\right), \ \ \ t,t' \geq 0.
@@
 \begin{equation}\label{eq:Wendland}
-T(t,t') = \phi^2\left(1 - \frac{|t - t'|}{\theta}\right)^3_+ \left(\frac{3|t - t'|}{\theta} + 1\right), 
+T(t,t') = \phi^2\left(1 - \frac{|t - t'|}{\tau}\right)^3_+ \left(\frac{3|t - t'|}{\tau} + 1\right), 
 \end{equation}
@@
 For the outlier detection algorithm, our prior model, $u$, is chosen to have a length-scale and time-scale which is able to approximately describe SSE displacements.
-We use the SE covariance function for $X$ with $\lambda = 100$ km, and we use the Wendland covariance function for $T$, due to its computational efficiency, with $\theta = 0.1$ yr and $\phi = 1$ mm.
+We use the SE covariance function for $X$ with $\ell = 100$ km, and we use the Wendland covariance function for $T$, due to its computational efficiency, with $\tau = 0.1$ yr and $\phi = 1$ mm.
 The outlier detection algorithm is particularly effective at removing outliers for stations at high elevation, which can be adversely affected by ice or snow during the winter \citep{Lisowski2008}.
@@
 This suggests that the median of estimated hyperparameters should be an appropriate prior model for all Cascadia SSEs.
-For the IBM model, there are several anomalously large values for $\lambda$ and $\phi$, which is why they have large IQRs.
+For the IBM model, there are several anomalously large values for $\ell$ and $\phi$, which is why they have large IQRs.
 
@@
 \begin{tabular} {l l l l l l}
-$T$ & direction & $\lambda$  & $\phi$   & $\theta$  & diff. $\log$(REML) \\ \hline
+$T$ & direction & $\ell$  & $\phi$   & $\tau$  & diff. $\log$(REML) \\ \hline
 SE & east   & 92 $\pm$ 25 km  & 0.62 $\pm$ 0.11 mm  & 0.026 $\pm$ 0.011 yr  &  - \\
@@
 To further explore which covariance function for $T$ best describes SSEs, we compare the observations to the predicted displacements for each covariance function.
-let $\hat{\bm{d}} = \left(u(\bm{y}) + \bm{P}\bm{a}\right)|\bm{d}_*$ be the data prediction vector.
+let $\hat{\bm{d}} = \left(u(\bm{\xi}) + \bm{G}\bm{a}\right)|\bm{d}_*$ be the data prediction vector.
 It can be shown that $\hat{\bm{d}}$ is normally distributed with mean
@@
 \bm{\mu}_{\hat{d}} = \left[\begin{array}{cc}
-                           C_u(\bm{y},\bm{y}) & \bm{P} \\
+                           C_u(\bm{\xi},\bm{\xi}) & \bm{G} \\
                            \end{array}\right]
                      \left[\begin{array}{cc}
-                           \bm{\Sigma} & \bm{P} \\
-                           \bm{P}^T  & \bm{0} \\
+                           \bm{\Sigma} & \bm{G} \\
+                           \bm{G}^T  & \bm{0} \\
                            \end{array}\right]^{-1}
@@
 \begin{equation}\label{eq:DataPredCov}
-\bm{C}_{\hat{\bm{d}}} = C_u(\bm{y},\bm{y}) - 
+\bm{C}_{\hat{\bm{d}}} = C_u(\bm{\xi},\bm{\xi}) - 
                         \left[\begin{array}{cc}
-                              C_u(\bm{y},\bm{y}) & \bm{P} \\
+                              C_u(\bm{\xi},\bm{\xi}) & \bm{G} \\
                               \end{array}\right]
                         \left[\begin{array}{cc}
-                              \bm{\Sigma} & \bm{P} \\
-                              \bm{P}^T  & \bm{0} \\
+                              \bm{\Sigma} & \bm{G} \\
+                              \bm{G}^T  & \bm{0} \\
                               \end{array}\right]^{-1}
                         \left[\begin{array}{c}
-                              C_u(\bm{y},\bm{y}) \\
-                              \bm{P}^T \\
+                              C_u(\bm{\xi},\bm{\xi}) \\
+                              \bm{G}^T \\
                               \end{array}\right].
@@
 \begin{equation}\label{eq:Periodic}
-T(t,t') = \phi \exp\left(\frac{-\sin(\pi|t - t'|)^2}{2\theta^2}\right).
+T(t,t') = \phi \exp\left(\frac{-\sin(\pi|t - t'|)^2}{2\tau^2}\right).
 \end{equation}
-Realizations have annual periodicity and the roughness is controlled by $\theta$.
-Decreasing $\theta$ has the same effect as including higher frequency sinusoids in the seasonal model.
-The optimal value for $\theta$ can be found with the REML method as described in Section \ref{sec:NoiseModel}.
+Realizations have annual periodicity and the roughness is controlled by $\tau$.
+Decreasing $\tau$ has the same effect as including higher frequency sinusoids in the seasonal model.
+The optimal value for $\tau$ can be found with the REML method as described in Section \ref{sec:NoiseModel}.
 

### 2017-06-17 HinesHetlandGPR.tex
more

@@
 Knowing where and how quickly strain is accumulating gives insight into where we can expect stored elastic energy to be released seismically.
-Consequently, secular crustal strain rates, which can be estimated from GNSS data, have been used to constrain seismic hazard models such as UCERF3 \citep{Field2014}.
+Consequently, secular crustal strain rates estimated from GNSS data have been used to constrain seismic hazard models such as UCERF3 \citep{Field2014}.
 Dense networks of continuous GNSS stations, such as the Plate Boundary Observatory (PBO), make it feasible to estimate transient crustal strain, which we consider to be any deviation from the secular strain rate.
-Transient crustal strain can potentially illuminate geophysical signal that may not be immediately apparent from inspecting the GNSS displacement data.
+Transient crustal strain can potentially illuminate geophysical signal that may not be immediately apparent from the GNSS displacement data.
 The ability to detect transient geophysical phenomena is also relevant for seismic hazard because there are instances of major earthquakes being triggered by slow slip events \citep{Roeloffs2006} and postseismic deformation \citep{Freed2001}.
-For these reasons, it is important to develop and improve upon methods for estimating crustal strain rates from GNSS data.
 
 Most methods for estimating crustal strain rates from GNSS data assume some parametric form of the deformation signal.
-The simplest and traditional method for estimating secular crustal strain rates assumes that GNSS derived velocities can be described with a first order polynomial (i.e., having constant spatial gradients) over some subnetwork of the GNSS stations \citep[e.g][]{Frank1966,Prescott1976,Savage1986,Feigl1990,Murray2000}.
-The components of the strain rate tensor  for each subnetwork are then determined from the least squares fit to the observations.
-The assumption that deformation gradients are spatially uniform is obviously not appropriate when subnetworks span a large area.
-To help overcome this deficiency, \citet{Shen1996} and \citet{Shen2015} used an inverse distance weighting scheme where the estimated strain rate at some point is primarily controlled by observations at nearby stations.
+The simplest and traditional method for estimating secular crustal strain rates assumes that GNSS derived velocities can be described with a first order polynomial (i.e., having constant spatial gradients) over some subnetwork of the GNSS stations \citep[e.g.,][]{Frank1966,Prescott1976,Savage1986,Feigl1990,Murray2000}.
+The components of the strain rate tensor  for each subnetwork are then determined through a least squares fit to the observations.
+The assumption that deformation gradients are spatially uniform is not appropriate when subnetworks span a large area.
+To help overcome this deficiency, \citet{Shen1996,Shen2015} used an inverse distance weighting scheme, in which the estimated strain rate at some point is primarily controlled by observations at nearby stations.
 However, the methods described in \citet{Shen1996} and \citet{Shen2015} are still formulated by assuming that the deformation gradients are uniform over the entire network.
@@
 The type of basis functions and the number of degrees of freedom for a parameterization can be subjective.
-If there are too few degrees of freedom in the parameterization then estimated strain rates will be biased and the uncertainties will be underestimated.
-If there are too many degrees of freedom then there will not be any coherent features in the estimated strain rates.
+If there are too few degrees of freedom in the parameterization, then estimated strain rates will be biased and the uncertainties will be underestimated.
+On the other hand, if there are too many degrees of freedom, then there will not be any coherent features in the estimated strain rates.
 The methods described by \citet{Beavan2001}, \citet{Tape2009}, and \citet{Sandwell2016} also require the user to specify penalty parameters that control a similar trade-off between bias and variance in the solution.
 One could parameterize deformation with a physically motivated model of interseismic deformation \citep[e.g.,][]{Meade2005,McCaffrey2007}.
-In such models the lithospheric rheology and plate interfaces are assumed to be known.
+In such models the lithospheric rheology and fault geometries are assumed to be known.
 Any errors in the assumed physical model could similarly result in biased strain estimates and underestimated formal uncertainties.
@@
 Among the methods tested were the Network Strain Filter (NSF) \citep{Ohtani2010} and the Network Inversion Filter (NIF) \citep{Segall1997}.
-the NSF uses a wavelet parameterization to  describe the spatial component of geophysical signal and the NIF, which is intended for detecting slow fault slip, uses the elastic dislocation Green's functions from \citet{Okada1992}.
+the NSF uses a wavelet parameterization to  describe the spatial component of geophysical signal and the NIF, which is intended for imaging slow fault slip from geodetic data, uses the elastic dislocation Green's functions from \citet{Okada1992}.
 For the NSF and NIF, the time dependence of the geophysical signal is modeled as integrated Brownian motion.
-The method described in \citet{Holt2013} was also tested in the transient detection exercise, and they calculate strain rates using a bi-cubic spatial parameterization of displacements between time epochs.
-\citet{Holt2013} defined a detection threshold based on the strain rate magnitude, and we will demonstrate that this is indeed an effective criterion for identifying geophysical signal.
+The method described in \citet{Holt2013} was also tested in the SCEC transient detection exercise, which calculates strain rates using a bi-cubic spatial parameterization of displacements between time epochs.
+\citet{Holt2013} defined a detection threshold based on the strain rate magnitude, and below we demonstrate that this is indeed an effective criterion for identifying geophysical signal.
 For the same reasons descibed above, the transient deformation and corresponding uncertainties estimated by these methods can be biased by the chosen spatial parameterization.
@@
 
-In this paper we use Gaussian process regression (GPR) \citep{Rasmussen2006} to infer transient strain from GNSS data.
+Here we propose using Gaussian process regression (GPR) \citep{Rasmussen2006} to infer transient strain from GNSS data.
 GPR is a Bayesian, non-parametric method for inferring a continuous signal from scattered and potentially noisy data.
-GPR is ideal for synthesizing GNSS data into a spatially and temporally continuous represention of surface deformation, since GNSS stations are irregularly spaced and observation times may differ between stations.
-GPR is closely related to Kriging \citep{Cressie1992} and Least Squares Collocation \citep{Kato1998}, and it has recently gained popularity as a tool for machine learning.
+GPR is ideal for synthesizing GNSS data into a spatially and temporally continuous representation of surface deformation, since GNSS stations are irregularly spaced and observation times may differ between stations.
+GPR is closely related to kriging \citep{Cressie1992} and least squares collocation \citep{Kato1998}, and it has recently gained popularity in the machine learning community.
 GPR is Bayesian in that we describe our prior understanding of the geophysical signal with a Gaussian process.
 A Gaussian process is a normally distributed, continuous, stochastic process that is fully defined in terms of a mean function and a positive-definite covariance function.
-For example, Brownian motion is a well known Gaussian process in $\mathbb{R}^1$ which has zero mean and covariance function $\mathrm{cov}(t,t')=\min(t,t')$, where $t,t' \ge 0$.
+For example, Brownian motion, $B(t)$, is a well known Gaussian process in $\mathbb{R}^1$ which has zero mean and covariance function $\mathrm{cov}(B(t),B(t')) = \min(t,t')$, where $t,t' \ge 0$.
 If no prior information is available for the geophysical signal, then maximum likelihood methods can be used to choose a prior Gaussian process which is most consistent with the observations.
-We incorporate the GNSS observations with the prior to form a posterior estimate of transient strain.
-The posterior transient strain is also a Gaussian process and, provided that our prior is sensible, has well quantified uncertainties.
-With well quantified uncertainties we can more confidently discern geophysical signal from noise in our estimates of transient strain.
+We incorporate GNSS observations with the prior to form a posterior estimate of transient strain.
+The posterior transient strain is also a Gaussian process, and we can use its distribution to confidently discern geophysical signal from noise.
 We use GPR to infer transient strain resulting from slow slip events (SSEs) in Cascadia.
@@
 \section{Estimating Transient Strain Rates}\label{sec:Method}
-We seek a spatially and temporally dependent estimate of transient crustal strain rate, denoted as $\dot{\du{\varepsilon}}(\zeta)$.
-We use $\zeta$ to represent the ordered pair $(\vec{x},t)$, where $\vec{x}$ are spatial coordinates in $\mathbb{R}^2$, and $t$ is time.
-We consider $\dot{\du{\varepsilon}}(\zeta)$ to be spatially and temporally coherent deviations from the steady rate of strain accumulation from plate tectonics.
-We determine $\dot{\du{\varepsilon}}(\zeta)$ by spatially and temporally differentiating estimates of transient displacements, $\vec{u}(\zeta)$.
-As we discuss in Section \ref{sec:NoiseModel}, estimates of $\dot{\du{\varepsilon}}$ turn out to be more effective at illuminating geophysical signal than estimates of $\vec{u}(\zeta)$ because $\dot{\du{\varepsilon}}$ is insensitive to common mode noise in GNSS data.
+We seek a spatially and temporally dependent estimate of transient crustal strain rate, denoted as $\dot{\du{\varepsilon}}(p)$.
+We use $p$ to represent the ordered pair $(\vec{x},t)$, where $\vec{x}$ are spatial coordinates in $\mathbb{R}^2$, and $t$ is time.
+We consider $\dot{\du{\varepsilon}}(p)$ to be spatially and temporally coherent deviations from the steady rate of strain accumulation from plate tectonics.
+We determine $\dot{\du{\varepsilon}}(p)$ by spatially and temporally differentiating estimates of transient displacements, $\vec{u}(p)$.
+As we discuss in Section \ref{sec:NoiseModel}, estimates of $\dot{\du{\varepsilon}}$ turn out to be more effective at illuminating geophysical signal than estimates of $\vec{u}(p)$ because $\dot{\du{\varepsilon}}$ is insensitive to common mode noise in GNSS data.
 We make a prior assumption that each component of $\vec{u}$ is a Gaussian process,
 \begin{equation}\label{eq:TransientDeformation}
-u_i(\zeta) \sim \mathcal{N}\left(0,C_{u_i}\right),
+u_i(p) \sim \mathcal{N}\left(0,C_{u_i}\right),
 \end{equation}
-where $C_{u_i}(\zeta,\zeta')$ is a covariance function indicating how we expect $u_i(\zeta)$ to covary with $u_i(\zeta')$.
+where $C_{u_i}(p,p')$ is a covariance function indicating how we expect $u_i(p)$ to covary with $u_i(p')$.
 For simplicity, we treat each component of displacement independently and ignore any potential covariance.
-We henceforth drop the component subscripts with the understanding that the same analysis is being repeated to estimate the east, north, and vertical components of $\vec{u}$.
+We drop the component subscripts with the understanding that the same analysis is being repeated to estimate the east, north, and vertical components of $\vec{u}$.
 We assume that $C_u$ can be separated into positive definite spatial and temporal functions as
@@
 \end{align}
-where $a^{(1)}_{i}$ is an offset that is unique for each GNSS monument, $a^{(2)}_{i}$ is the steady rate of tectonic deformation at $\vec{x}_i$, and the sinusoids describe seasonal deformation (using units of years for $t$).
+where $a^{(1)}_{i}$ is an offset that is unique for each GNSS monument, $a^{(2)}_{i}$ is the secular velocity at $\vec{x}_i$, and the sinusoids describe seasonal deformation (using units of years for $t_j$).
 We use $w_{ij}$ to denote normally distributed, uncorrelated noise.
@@
 (\ref{eq:Data}) to be uncorrelated random variables distributed as $\mathcal{N}(0,\kappa^2)$ in the limit as $\kappa \to \infty$ (i.e., the coefficients have diffuse priors).
-Of course, the tectonic deformation, $a^{(2)}_{i}$, is spatially correlated and we could invoke a tectonic model to form a prior on $a^{(2)}_{i}$.
+Of course, the secular velocities, $a^{(2)}_{i}$, are spatially correlated and we could invoke a tectonic model to form a prior on $a^{(2)}_{i}$.
 However, in our application to Cascadia, we will be using displacement time series which are long enough to sufficiently constrain $a^{(2)}_{i}$ for each station, and there is no need to incorporate a prior.
@@
 We now consider the column vector of $n$ GNSS observations, $\bm{d}_*$.
-Let $\bm{\xi}$ be the set of $(\vec{x}_i, t_j)$ pairs describing where and when each of the GNSS observations have been made.
+Let $\bm{P}$ be the set of $(\vec{x}_i, t_j)$ pairs describing where and when each of the GNSS observations have been made.
 Let $\bm{a}$ be the vector of coefficients from eq.
 (\ref{eq:Data}) for each of the $m$ GNSS stations.
-We use $\bm{G}$ to represent the $n \times 6m$ matrix of corresponding basis functions evaluated at each point in $\bm{\xi}$.
+We use $\bm{G}$ to represent the $n \times 6m$ matrix of corresponding basis functions evaluated at each point in $\bm{P}$.
 We  also denote the vector of uncorrelated noise for each observation as $\bm{w}$, whose standard deviations are given by the formal data uncertainty, $\bm{\sigma}$.
@@
 \begin{equation}
-\bm{d} = u(\bm{\xi}) + \eta(\bm{\xi}) + \bm{w} + \bm{G}\bm{a},
+\bm{d} = u(\bm{P}) + \eta(\bm{P}) + \bm{w} + \bm{G}\bm{a},
 \end{equation}
@@
 \begin{equation}\label{eq:Cd}
-\bm{\Sigma} = C_u(\bm{\xi},\bm{\xi}) + C_\eta(\bm{\xi},\bm{\xi}) + 
+\bm{\Sigma} = C_u(\bm{P},\bm{P}) + C_\eta(\bm{P},\bm{P}) + 
               \mathrm{diag}\left(\bm{\sigma}^2\right).  
 \end{equation}
-It should be understood that notation such as $u(\bm{\xi})$ and $C_u(\bm{\xi},\bm{\xi})$ is shorthand for $[u(\xi_i)]_{\xi_i \in \bm{\xi}}$ and $[C_u(\xi_i,\xi_j)]_{(\xi_i,\xi_j) \in \bm{\xi} \times \bm{\xi}}$, respectively.
+It should be understood that notation such as $u(\bm{P})$ and $C_u(\bm{P},\bm{P})$ represents the column vector $[u(P_i)]_{P_i \in \bm{P}}$ and the matrix $[C_u(P_i,P_j)]_{(P_i,P_j) \in \bm{P} \times \bm{P}}$, respectively.
 
@@
 \begin{equation}\label{eq:PosteriorMean}
-\mu_{\hat{u}}(\zeta) = C_u(\zeta,\bm{\xi})\left(\bm{\Sigma} + \kappa^2\bm{G}\bm{G}^T\right)^{-1}\bm{d}_*
+\mu_{\hat{u}}(p) = C_u(p,\bm{P})\left(\bm{\Sigma} + \kappa^2\bm{G}\bm{G}^T\right)^{-1}\bm{d}_*
 \end{equation}    
@@
 \begin{equation}\label{eq:PosteriorCov}
-C_{\hat{u}}(\zeta,\zeta') = C_u(\zeta,\zeta') - C_u(\zeta,\bm{\xi})\left(\bm{\Sigma} + \kappa^2\bm{G}\bm{G}^T\right)^{-1}C_u(\bm{\xi},\zeta').
+C_{\hat{u}}(p,p') = C_u(p,p') - C_u(p,\bm{P})\left(\bm{\Sigma} + \kappa^2\bm{G}\bm{G}^T\right)^{-1}C_u(\bm{P},p').
 \end{equation}
@@
  \begin{equation}\label{eq:PosteriorMean2}
-\mu_{\hat{u}}(\zeta) = \left[\begin{array}{cc}
-                         C_u(\zeta,\bm{\xi}) & \bm{0} \\
+\mu_{\hat{u}}(p) = \left[\begin{array}{cc}
+                         C_u(p,\bm{P}) & \bm{0} \\
                          \end{array}\right]
@@
 \begin{equation}\label{eq:PosteriorCov2}
-C_{\hat{u}}(\zeta,\zeta') = C_u(\zeta,\zeta') - 
+C_{\hat{u}}(p,p') = C_u(p,p') - 
                     \left[\begin{array}{cc}
-                          C_u(\zeta,\bm{\xi}) & \bm{0} \\
+                          C_u(p,\bm{P}) & \bm{0} \\
                           \end{array}\right]
@@
                     \left[\begin{array}{c}
-                          C_u(\bm{\xi},\zeta') \\
+                          C_u(\bm{P},p') \\
                           \bm{0} \\
@@
  \begin{equation}\label{eq:PosteriorMean3}
-\mu_{\hat{u}}(\zeta) = \left[\begin{array}{cc}
-                         C_u(\zeta,\bm{\xi}) & \bm{0} \\
+\mu_{\hat{u}}(p) = \left[\begin{array}{cc}
+                         C_u(p,\bm{P}) & \bm{0} \\
                          \end{array}\right]
@@
 \begin{equation}\label{eq:PosteriorCov3}
-C_{\hat{u}}(\zeta,\zeta') = C_u(\zeta,\zeta') - 
+C_{\hat{u}}(p,p') = C_u(p,p') - 
                     \left[\begin{array}{cc}
-                          C_u(\zeta,\bm{\xi}) & \bm{0} \\
+                          C_u(p,\bm{P}) & \bm{0} \\
                           \end{array}\right]
@@
 The posterior transient displacements are differentiated temporally and spatially to form an estimate of $\dot{\du{\varepsilon}}$.
-The components of $\dot{\du{\varepsilon}}$ are given by
+Using $\hat{u}_i$ to denote the posterior transient displacements along direction $i$ and $x_i$ to represent the components of $\vec{x}$, we can write the components of $\dot{\du{\varepsilon}}$ as
 \begin{equation}\label{eq:StrainRate}
-\dot{\varepsilon}_{ij}(\zeta) = \frac{1}{2} \frac{\partial}{\partial t} \left(
-                                        \frac{\partial \hat{u}_i(\zeta)}{\partial x_j} +  
-                                        \frac{\partial \hat{u}_j(\zeta)}{\partial x_i}\right),
+\dot{\varepsilon}_{ij}(p) = \frac{1}{2} \frac{\partial}{\partial t} \left(
+                                        \frac{\partial \hat{u}_i(p)}{\partial x_j} +  
+                                        \frac{\partial \hat{u}_j(p)}{\partial x_i}\right).
 \end{equation}
-which are Gaussian processes with mean functions
+The transient strain rate components are Gaussian processes with mean functions
 \begin{equation}\label{eq:StrainMean}
-\mu_{\dot{\varepsilon}_{ij}}(\zeta) = \frac{1}{2}\frac{\partial}{\partial t}\left(
-                                  \frac{\partial \mu_{\hat{u}_i}(\zeta)}{\partial x_j} + 
-                                  \frac{\partial \mu_{\hat{u}_j}(\zeta)}{\partial x_i} \right)
+\mu_{\dot{\varepsilon}_{ij}}(p) = \frac{1}{2}\frac{\partial}{\partial t}\left(
+                                  \frac{\partial \mu_{\hat{u}_i}(p)}{\partial x_j} + 
+                                  \frac{\partial \mu_{\hat{u}_j}(p)}{\partial x_i} \right)
 \end{equation} 
@@
 \begin{equation}\label{eq:StrainCov}
-C_{\dot{\varepsilon}_{ij}}(\zeta,\zeta') = \frac{1}{4} \frac{\partial^2}{\partial t \, \partial t'}\left(
-                                   \frac{\partial^2 C_{\hat{u}_i}(\zeta,\zeta')}{\partial x_j \, \partial x'_j} + 
-                                   \frac{\partial^2 C_{\hat{u}_j}(\zeta,\zeta')}{\partial x_i \, \partial x'_i} \right).
+C_{\dot{\varepsilon}_{ij}}(p,p') = \frac{1}{4} \frac{\partial^2}{\partial t \, \partial t'}\left(
+                                   \frac{\partial^2 C_{\hat{u}_i}(p,p')}{\partial x_j \, \partial x'_j} + 
+                                   \frac{\partial^2 C_{\hat{u}_j}(p,p')}{\partial x_i \, \partial x'_i} \right).
 \end{equation}
-In the above equations we use $\hat{u}_i$ to denote posterior transient displacements along direction $i$ and $x_i$ to represent the components of $\vec{x}$.
 
@@
 We calculate the residuals between the observations and a best fitting model, and data with residuals that are anomalously large are identified as outliers.
-We treat $\bm{d}_*$ as a sample of $\bm{d}$ and assume that there is no temporally correlated noise (i.e., $\eta(\zeta) = 0$).
-The best fitting model for $\bm{d}_*$ is considered to be the expected value of the random vector $u(\bm{\xi}) + \bm{G}\bm{a}$ after conditioning it with non-outlier observations.
+We treat $\bm{d}_*$ as a sample of $\bm{d}$ and assume that there is no temporally correlated noise (i.e., $\eta(p) = 0$).
+The best fitting model for $\bm{d}_*$ is considered to be the expected value of the random vector $u(\bm{P}) + \bm{G}\bm{a}$ after conditioning it with non-outlier observations.
 We still consider $u$ to have a separable covariance function as in eq.
@@
 
-To begin the algorithm, we let $\Omega$ be the index set of non-outliers in $\bm{d}_*$ and initiate it with all $n$ indices.
+To begin the algorithm, we let $\bm{\Omega}$ be the index set of non-outliers in $\bm{d}_*$ and initiate it with all $n$ indices.
 This algorithm is iterative, and for each iteration we calculate the residual vector
 \begin{align}\label{eq:Residual}
-\bm{r} &= \frac{\bm{d}_* - \mathrm{E}\left[(u(\bm{\xi}) + \bm{G}\bm{a})|\tilde{\bm{d}}_* \right]}{\bm{\sigma}} \\
+\bm{r} &= \frac{\bm{d}_* - \mathrm{E}\left[(u(\bm{P}) + \bm{G}\bm{a})|\tilde{\bm{d}}_* \right]}{\bm{\sigma}} \\
        &= \frac{1}{\bm{\sigma}}\left(\bm{d}_*  - 
           \left[\begin{array}{cc}
-                C_u(\bm{\xi},\tilde{\bm{\xi}}) & \bm{G} \\
+                C_u(\bm{P},\tilde{\bm{P}}) & \bm{G} \\
                 \end{array}\right]
           \left[\begin{array}{cc}
-                C_u(\tilde{\bm{\xi}},\tilde{\bm{\xi}}) + \mathrm{diag}(\tilde{\bm{\sigma}}^2) & \tilde{\bm{G}} \\
+                C_u(\tilde{\bm{P}},\tilde{\bm{P}}) + \mathrm{diag}(\tilde{\bm{\sigma}}^2) & \tilde{\bm{G}} \\
                 \tilde{\bm{G}}^T  & \bm{0} \\
@@
 \end{align}
-where the tilde indicates that only elements corresponding to indices in $\Omega$ are retained (e.g., $\tilde{\bm{\xi}} = \{y_i\}_{i\in\Omega}$). We then update $\Omega$ to be
+where the tilde indicates that only elements corresponding to indices in $\bm{\Omega}$ are retained (e.g., $\tilde{\bm{P}} = \{P_i\}_{i\in\bm{\Omega}}$). We then update $\bm{\Omega}$ to be
 \begin{equation}\label{eq:Update}
-\Omega = \{i : |r_i| < \lambda \tilde{\bm{r}}_{rms}\}, \ \ \ r_i \in \bm{r}
+\bm{\Omega} = \{i : |r_i| < \lambda \tilde{\bm{r}}_{rms}\}, \ \ \ r_i \in \bm{r}
 \end{equation} 
@@
 We use $\lambda=4$ in this study, which seems to accurately identify outliers without unnecessarily decimating the data.
-Iterations continue until the new $\Omega$ is equal to the previous $\Omega$.
+Iterations continue until the new $\bm{\Omega}$ is equal to the previous $\bm{\Omega}$.
 
@@
 For this section, $\bm{d}_*$ is used to denote the displacement time series for just one station.
-We are also assuming $C_u(\zeta,\zeta')=0$ when estimating the noise hyperparameters for this section.
+We are also assuming $C_u(p,p')=0$ when estimating the noise hyperparameters for this section.
 \citet{Harville1974} showed that choosing the hyperparameters which maximize eq.
@@
 We therefore do not include common mode error in our noise model.
-We then make the simplifying assumption that $\eta(\zeta) = 0$ for the easting and northing component of GNSS data.
+We then make the simplifying assumption that $\eta(p) = 0$ for the easting and northing component of GNSS data.
 
@@
 To further explore which covariance function for $T$ best describes SSEs, we compare the observations to the predicted displacements for each covariance function.
-let $\hat{\bm{d}} = \left(u(\bm{\xi}) + \bm{G}\bm{a}\right)|\bm{d}_*$ be the data prediction vector.
+let $\hat{\bm{d}} = \left(u(\bm{P}) + \bm{G}\bm{a}\right)|\bm{d}_*$ be the data prediction vector.
 It can be shown that $\hat{\bm{d}}$ is normally distributed with mean
@@
 \bm{\mu}_{\hat{d}} = \left[\begin{array}{cc}
-                           C_u(\bm{\xi},\bm{\xi}) & \bm{G} \\
+                           C_u(\bm{P},\bm{P}) & \bm{G} \\
                            \end{array}\right]
@@
 \begin{equation}\label{eq:DataPredCov}
-\bm{C}_{\hat{\bm{d}}} = C_u(\bm{\xi},\bm{\xi}) - 
+\bm{C}_{\hat{\bm{d}}} = C_u(\bm{P},\bm{P}) - 
                         \left[\begin{array}{cc}
-                              C_u(\bm{\xi},\bm{\xi}) & \bm{G} \\
+                              C_u(\bm{P},\bm{P}) & \bm{G} \\
                               \end{array}\right]
@@
                         \left[\begin{array}{c}
-                              C_u(\bm{\xi},\bm{\xi}) \\
+                              C_u(\bm{P},\bm{P}) \\
                               \bm{G}^T \\

### 2017-06-18 HinesHetlandGPR.tex
more stuff

@@
 \section{Estimating Transient Strain Rates}\label{sec:Method}
-We seek a spatially and temporally dependent estimate of transient crustal strain rate, denoted as $\dot{\du{\varepsilon}}(p)$.
-We use $p$ to represent the ordered pair $(\vec{x},t)$, where $\vec{x}$ are spatial coordinates in $\mathbb{R}^2$, and $t$ is time.
-We consider $\dot{\du{\varepsilon}}(p)$ to be spatially and temporally coherent deviations from the steady rate of strain accumulation from plate tectonics.
+We seek a spatially and temporally dependent estimate of transient crustal strain rates, which we consider to be any deviation from the secular strain rates.
+We denote transient crustal strain rates as $\dot{\du{\varepsilon}}(p)$, where $p$ represents the ordered pair $(\vec{x},t)$, $\vec{x}$ are spatial coordinates in $\mathbb{R}^2$, and $t$ is time.
 We determine $\dot{\du{\varepsilon}}(p)$ by spatially and temporally differentiating estimates of transient displacements, $\vec{u}(p)$.
-As we discuss in Section \ref{sec:NoiseModel}, estimates of $\dot{\du{\varepsilon}}$ turn out to be more effective at illuminating geophysical signal than estimates of $\vec{u}(p)$ because $\dot{\du{\varepsilon}}$ is insensitive to common mode noise in GNSS data.
+As we discuss in Section \ref{sec:NoiseModel} and also as argued for in XXX, estimates of $\dot{\du{\varepsilon}}$ are more effective at illuminating geophysical signal than estimates of $\vec{u}(p)$.
 We make a prior assumption that each component of $\vec{u}$ is a Gaussian process,
@@
 Of course, the secular velocities, $a^{(2)}_{i}$, are spatially correlated and we could invoke a tectonic model to form a prior on $a^{(2)}_{i}$.
-However, in our application to Cascadia, we will be using displacement time series which are long enough to sufficiently constrain $a^{(2)}_{i}$ for each station, and there is no need to incorporate a prior.
+However, in our application to Cascadia, we will be using displacement time series which are long enough to sufficiently constrain $a^{(2)}_{i}$ for each station, avoiding the need to incorporate a prior on it.
 Likewise, the seasonal coefficients may be spatially correlated \citep{Langbein2008}, and it may be worth exploring and exploiting such a correlation in a future study.
@@
 The prior for each component of transient displacements is then conditioned with $\bm{d}_*$ to form a posterior estimate of transient displacements, $\hat{u} = u | \bm{d}_*$.
-We will assume that an appropriate covariance functions and corresponding hyperparameters for $X$, $T$, and $C_\eta$ have already been chosen.
-If $\kappa$ is kept finite, then we can follow \citet{Rasmussen2006} to find that $\hat{u}$ is distributed as $\mathcal{N}(\mu_{\hat{u}},C_{\hat{u}})$, where
+We will assume that an appropriate covariance function and corresponding hyperparameters for $X$, $T$, and $C_\eta$ have already been chosen.
+If $\kappa$ is kept finite, then following \citet{Rasmussen2006} we find that $\hat{u}$ is distributed as $\mathcal{N}(\mu_{\hat{u}},C_{\hat{u}})$, where
 \begin{equation}\label{eq:PosteriorMean}
@@
 
+Our motivation for estimating transient strain rates is, in part, to detect geophysical phenomena.
+In Section \ref{sec:Results}, we show that strain resulting from SSEs is immediately visible from simply inspecting $\dot{\du{\varepsilon}}$.
+However, if our goal is to automatically detect geophysical events such as SSEs, then we need to define a signal-to-noise ratio, SNR, based on $\dot{\du{\varepsilon}}$.
+We consider the Frobenius norm of $\dot{\du{\varepsilon}}$, $||\dot{\du{\varepsilon}}||_\mathrm{F}$, which is also referred to as the second invariant of strain rate in the geodetic literature.
+We use $||\dot{\du{\varepsilon}}||_\mathrm{F}$ as a metric for the strain rate ``magnitude''.
+Noting that $||\dot{\du{\varepsilon}}||_\mathrm{F}$ is a random variable, we choose SNR to be the ratio of the estimated mean and standard deviation of $||\dot{\du{\varepsilon}}||_\mathrm{F}$.
+Using nonlinear error-propagation, we find SNR to be
+\begin{equation}\label{eq:SNR}
+\mathrm{SNR}(p) = \sqrt{\frac{C_{\dot{\varepsilon}_\mathrm{nn}}(p,p)\mu_{\dot{\varepsilon}_\mathrm{nn}}(p)^2 + 
+               C_{\dot{\varepsilon}_\mathrm{ee}}(p,p)\mu_{\dot{\varepsilon}_\mathrm{ee}}(p)^2 + 
+               4C_{\dot{\varepsilon}_\mathrm{en}}(p,p)\mu_{\dot{\varepsilon}_\mathrm{en}}(p)^2}{
+               \mu_{\dot{\varepsilon}_\mathrm{nn}}(p)^2 +
+               \mu_{\dot{\varepsilon}_\mathrm{ee}}(p)^2 +
+               2\mu_{\dot{\varepsilon}_\mathrm{en}}(p)^2}},
+\end{equation}
+where the subscripts ``n'' and ``e'' denote north and east, respectively.
+
 \section{Outlier detection}\label{sec:Outlier}
 In our formulation for estimating transient strain rates, we have assumed that noise in the data vector is normally distributed.
-This is not an appropriate assumption for GNSS displacement timeseries which are prone to more outliers than would be predicted for normally distributed noise.
-It follows that proposed methods for analyzing GNSS timeseries should be robust against outliers \citep[e.g.,][]{Blewitt2016}.
+This is not an appropriate assumption for GNSS displacement time series which are prone to more outliers than would be predicted for normally distributed noise.
+It follows that proposed methods for analyzing GNSS time series should be robust against outliers \citep[e.g.,][]{Blewitt2016}.
 In order to make our estimates of transient strain more robust, we automatically identify and remove outliers in the GNSS data as a pre-processing step.
@@
 (\ref{eq:TransientCovariance}), and the choice for $X$ and $T$ does not need to be the same as that used in Section \ref{sec:Method}.
-Since outliers are determined based on how well a spatially and temporally dependent model fits the data, we are able to identify anomalous observations which may not be immediately apparent based on inspection of individual timeseries.
+Since outliers are determined based on how well a spatially and temporally dependent model fits the data, we are able to identify anomalous observations which may not be immediately apparent based on inspection of individual time series.
 
@@
 \begin{equation}\label{eq:Update}
-\bm{\Omega} = \{i : |r_i| < \lambda \tilde{\bm{r}}_{rms}\}, \ \ \ r_i \in \bm{r}
+\bm{\Omega} = \{i : |r_i| < \lambda \cdot \mathrm{RMS}\}, \ \ \ r_i \in \bm{r}
 \end{equation} 
-where $\tilde{\bm{r}}_{rms}$ is the root-mean-square of $\tilde{\bm{r}}$ and $\lambda$ is an outlier tolerance.
-We use $\lambda=4$ in this study, which seems to accurately identify outliers without unnecessarily decimating the data.
+where RMS is the root-mean-square of $\tilde{\bm{r}}$ and $\lambda$ is an outlier tolerance.
+We use $\lambda=4$ in this study, which in our experience accurately identifies outliers without unnecessarily decimating the data.
 Iterations continue until the new $\bm{\Omega}$ is equal to the previous $\bm{\Omega}$.
 
-It should be noted that this algorithm does not identify jumps in GNSS time series, which are another common defect.
-Some, but not all, jumps can be automatically removed by parsing station logs and identifying dates of equipment changes.
+It should be noted that this algorithm does not identify jumps in GNSS time series, which are another common issue.
+Some, but not all, jumps can be automatically removed by looking up the dates of equipment changes and earthquakes.
 However, it is still necessary to manually find and remove jumps of unknown origin.
@@
 We demonstrate our method for estimating transient strain rates with GNSS data from Cascadia.
-We are interested in identifying transient strain resulting from slow slip events (SSEs) \citep[e.g.,][]{Dragert2001}.
+We are interested in identifying transient strain resulting from SSEs \citep[e.g.,][]{Dragert2001}.
 Our results illuminate where elastic strain energy, which can eventually be released seismically, is accumulating during SSEs.
@@
 In Cascadia, SSEs can be detected by monitoring for associated seismic tremor \citep{Rogers2003}, which is actively being done by the Pacific Northwest Seismic Network (PNSN) \citep{Wech2010}.
-We verify that transient strain rates estimated with the method described above are indeed identifying strain from SSEs by comparing $\dot{\bm{\varepsilon}}$ to the tremor records.
-We envision that GNSS derived transient strain rates could be useful for monitoring subduction zones which have SSEs but no associated tremor \citep{Schwartz2007}.
+We verify that transient strain rates estimated with the method described above are indeed identifying strain from SSEs by comparing $\dot{\du{\varepsilon}}$ to the tremor records.
+We envision that GNSS derived transient strain rates could be useful for detecting SSEs at subduction zones where there is no associated tremor \citep{Schwartz2007}.
 
@@
 The earliest SSE considered began in August 2010, and the most recent SSE began in February 2017.
-We consider these most recent SSEs because the station coverage is sufficiently dense for us to make well constrained inferences of optimal prior models.
-The positions of GNSS stations used to constrain transient strain rates are shown in Figure \ref{fig:Context}.
+We consider these most recent SSEs because the station coverage is sufficiently dense for us to use maximum likelihood methods to constrain prior models.
+The positions of GNSS stations used to estimate transient strain rates are shown in Figure \ref{fig:Context}.
 
@@
 The red dots show the positions of GNSS stations mentioned in this paper.
-The magenta dot indicates the location of the strain rates shown in Figure \ref{fig:StrainTs} and \ref{fig:StrainMag}.
-The magenta dashed circle demarcates the spatial extent of the tremors shown in Figure \ref{fig:StrainMag}}.
+The green dot indicates the location of the strain rates shown in Figure \ref{fig:StrainTs} and \ref{fig:StrainMag}.
+The green dashed circle demarcates the spatial extent of the tremors shown in Figure \ref{fig:StrainMag}}.
 \label{fig:Context}
@@
 There have been numerous studies on temporally correlated noise in GNSS data \citep[e.g.,][]{Zhang1997,Mao1999,Williams2004,Langbein2008}.
-In these studies, temporally correlated noise was described with some combination of Brownian motion (also referred to as a Weiner process or a random walk), a first-order Gauss-Markov (FOGM) process, and/or flicker noise.
+In these studies, temporally correlated noise was described with some combination of Brownian motion, a first-order Gauss-Markov (FOGM) process, and/or flicker noise.
 There is some physical justification for using Brownian motion as a noise model because it accurately describes the power spectrum of motion resulting from instability in geodetic monuments \citep[e.g.,][]{Wyatt1982,Wyatt1989,Langbein1997}.
@@
 Our noise model which satisfies eq.
-(\ref{eq:FOGMdef}) is a Gaussian process with zero mean and covariance function
+(\ref{eq:FOGMdef}) is a Gaussian process with zero mean and the covariance function
 \begin{equation}\label{eq:FOGM}
@@
 
-We constrain the hyperparameters for $\eta$, $\alpha$ and $\beta$, with a set of 38 continuous GNSS stations in Washington and Oregon which are east of 121$^\circ$ W.
-These stations are sufficiently far from the subduction zone that they are unlikely to contain transient signal.
+We constrain the hyperparameters for $\eta$, $\alpha$ and $\beta$, with a set of 38 continuous GNSS stations in the Pacific Northwest which are east of 121$^\circ$W.
+These stations are sufficiently far from the subduction zone that they are unlikely to contain transient signal associated with the Cascadia SSEs.
 We clean the data for these stations by removing jumps at times of equipment changes, and we remove outliers that have been detected with the algorithm described in Section \ref{sec:Outlier}.
-We then find $\alpha$ and $\beta$ for each station timeseries with the Restricted Maximum Likelihood (REML) method \cite[e.g.,][]{Harville1974,Cressie1992}.
+We then find $\alpha$ and $\beta$ for each station time series with the Restricted Maximum Likelihood (REML) method \cite[e.g.,][]{Harville1974,Cressie1992,Hines2017}.
 The REML method finds the hyperparameters, which we collectively refer to as $\bm{\theta}$, that maximize the likelihood function
@@
 \end{equation}
-For this section, $\bm{d}_*$ is used to denote the displacement time series for just one station.
+We independently estimate $\bm{\theta}$ for each station, and so $\bm{d}_*$ consists of displacements for an individual station.
 We are also assuming $C_u(p,p')=0$ when estimating the noise hyperparameters for this section.
@@
 (\ref{eq:REML}) is equivalent to choosing the hyperparameters which maximize the probability of drawing $\bm{d}_*$ from $\bm{d}$.
-We use the REML method over the maximum likelihood (ML) method \citep[e.g.,][]{Langbein1997} because the REML method properly accounts for the improper prior that we assigned to $\bm{a}$.
+We use the REML method over the maximum likelihood (ML) method \citep[e.g.,][]{Langbein1997} because the REML method properly accounts for the improper prior that we assigned to $\bm{a}$ \citep{Hines2017}.
 The distribution of inferred $\alpha$ and $\beta$ are shown in Figure \ref{fig:NoiseParams}.
 The amplitude of FOGM noise, $\alpha$, for the east and north components is notable low and are clustered around 0.5 mm/yr$^{0.5}$.
-The corresponding estimates of $\beta$ tend to cluster around 0 yr$^{-1}$, suggesting that Brownian motion would have also been an appropriate noise model.
-The amplitude of FOGM noise for the vertical component is significantly larger with a median value of 13.5 mm/yr$^{0.5}$.
-The inferred values for $\beta$ are higher for the vertical component with a median value of 8.21 yr$^{-1}$.
+The corresponding estimates of $\beta$ tend to cluster around 0 yr$^{-1}$, suggesting Brownian motion.
+We also estimate hyperparameters for the vertical component of displacements, even though verticals are not used for estimating strain.
+Our hope is that the vertical deformation gradients could reveal some geophysical signal.
+The amplitude of FOGM noise for the vertical component is relatively large with a median value of 13.5 mm/yr$^{0.5}$.
+The inferred values for $\beta$ are also higher for the vertical component with a median value of 8.21 yr$^{-1}$.
 In Figure \ref{fig:NoiseSamples}, we use the median values of $\alpha$ and $\beta$ to generate two random samples of FOGM noise.
 The samples span five years and over these five years the east and north samples drift by about 1 mm.
-In the context of detecting SSEs, which produce several mm's of surface displacement on the timescale of weeks, the estimated FOGM noise for the east and north component is negligible.
+In the context of detecting SSEs, which produce several mm's of surface displacement on the time-scale of weeks, the estimated FOGM noise for the east and north component is negligible.
 In contrast, the estimated FOGM noise for the vertical component is larger than the signal we would expect from SSEs.
@@
 Based on this analysis, we henceforth ignore temporally correlated noise in the east and north component because of its low amplitude.
-We also do not estimate $u$ for vertical deformation because of the low signal to noise ratio.
+We also do not use vertical displacements because of the low signal to noise ratio.
 
@@
 \includegraphics{figures/noise/noise-params.pdf}
-\caption{Distribution of estimated FOGM hyperparameters from eq.
-(\ref{eq:FOGM}).
-Hyperparameters are independently estimated for 38 stations in Washington and Oregon that are east of $121^\circ$ W.
+\caption{Distribution of estimated FOGM hyperparameters (eq.
+\ref{eq:FOGM}).
+Hyperparameters are independently estimated for 38 stations in the Pacific Northwest that are east of $121^\circ$W.
 ``IQR'' is the inter-quartile range.}
@@
 \includegraphics{figures/noise/noise-samples.pdf}
-\caption{Two FOGM noise samples, where the hyperparameters have been set to the median values from Figure \ref{fig:NoiseParams}.}   
+\caption{Two FOGM noise samples for each component, where the hyperparameters have been set to the median values from Figure \ref{fig:NoiseParams}.}   
 \label{fig:NoiseSamples}
@@
 When not accounted for, common mode error manifests as spatially uniform undulations in $\hat{u}$.
-However, we are primarily interested in estimating strain which is insensitive to common mode error.
+However, we are primarily interested in estimating strain rates, which are insensitive to common mode error.
 We therefore do not include common mode error in our noise model.
@@
 \end{equation}
-The SE covariance function is commonly used in Kriging \citep[e.g,][]{Cressie1992} and Gaussian process regression \citep[e.g.,][]{Rasmussen2006}.
-The SE is a valid (i.e., positive definite) covariance function for any number of spatial dimensions, and it describes an isotropic Gaussian processes with realizations that are infinitely differentiable.
-In terms of geodetic applications, \citet{Kato1998} and \cite{El-Fiky1999} demonstrated that the SE covariance function accurately describes secular strain rates in Japan.
+The SE covariance function is commonly used in kriging \citep[e.g,][]{Cressie1992} and Gaussian process regression \citep[e.g.,][]{Rasmussen2006}.
+The SE is a positive definite covariance function for any number of spatial dimensions.
+A Gaussian process with an SE covariance function is isotropic and has realizations that are infinitely differentiable.
+In terms of geodetic applications, \citet{Kato1998} and \cite{El-Fiky1999} demonstrated that the SE accurately describes the covariance of secular GNSS derived velocities in Japan.
 
@@
 \begin{equation}\label{eq:TimeSE}
-T(t,t') = \phi^2\exp\left(\frac{-|t - t'|^2}{\tau^2}\right),
+T(t,t') = \phi^2\exp\left(\frac{-|t - t'|^2}{\tau^2}\right).
 \end{equation}
-because the SE covariance function is versatile and commonly used for non-parametric modeling.
 Note that $T$ includes the hyperparameter $\phi$, which serves to scale the covariance function $C_u$.
@@
 \end{align}
-IBM has been used, in the context of Kalman filtering, as a non-parametric model for the time dependence of geophysical signals \citep[e.g.,][]{Segall1997,McGuire2003,Ohtani2010,Hines2016}.
+IBM has been used in the context of Kalman filtering as a non-parametric model for the time dependence of geophysical signals \citep[e.g.,][]{Segall1997,McGuire2003,Ohtani2010,Hines2016}.
 It should be emphasized $t=0$ is a reference time at which the Gaussian process is exactly zero.
@@
 In our analysis, we exploit this sparsity with the CHOLMOD software package \citep{Chen2008}.
+We elaborate on the computational advantages of the Wendland functions in Section \ref{sec:Discussion}.
 Wendland functions are positive definite in $\mathbb{R}^d$, and they describes an isotropic Gaussian process with realizations that can be differentiated $k$ times.
 The form of the covariance function depends on the choice of $d$ and $k$.
-To describe the temporal covariance of $u$, we only require that $d=1$ and $k=1$.
+We use $d=1$ since we are describing the temporal covariance of $u$.
+We use $k=1$ because we need $u$ to be at least once differentiable in time to determine strain rates.
 The corresponding Wendland covariance function is
@@
 
-We next want to determine appropriate hyperparameters for $X$ and each of the three candidate functions for $T$.
-First, we again clean the GNSS datasets by removing offsets at times of equipment changes and removing outliers with the method describe in Section \ref{sec:Outlier}.
+We next determine appropriate hyperparameters for $X$ and each of the three candidate covariance functions for $T$.
+First, we clean the GNSS datasets by removing offsets at times of equipment changes and removing outliers with the method describe in Section \ref{sec:Outlier}.
 For the outlier detection algorithm, our prior model, $u$, is chosen to have a length-scale and time-scale which is able to approximately describe SSE displacements.
-We use the SE covariance function for $X$ with $\ell = 100$ km, and we use the Wendland covariance function for $T$, due to its computational efficiency, with $\tau = 0.1$ yr and $\phi = 1$ mm.
-The outlier detection algorithm is particularly effective at removing outliers for stations at high elevation, which can be adversely affected by ice or snow during the winter \citep{Lisowski2008}.
-As an example, Figure \ref{fig:Outliers} shows the outliers which were automatically detected for station SC03, which is located on Mount Olympus in Washington.
+We use the SE covariance function for $X$ with length-scale $\ell = 100$ km, and we use the Wendland covariance function for $T$, due to its computational efficiency, with time-scale $\tau = 0.1$ yr and variance $\phi = 1$ mm.
+The outlier detection algorithm is particularly effective at removing outliers for stations at high elevation (Figure \ref{fig:Outliers}), which can be adversely affected by ice or snow during the winter \citep{Lisowski2008}.
 After cleaning the data, we divide it into seven subsets which are four months long and each centered on the time of an SSE.
 The times of seven SSEs are determined with tremor records from \cite{Wech2010}.
-We find the optimal hyperparameters for $T$ and $X$ for each subset of data.
-The REML method is once again used to determine the hyperparameters for $T$ and $X$ which maximize the probability of drawing $\bm{d}_*$ from $\bm{d}$, where $\bm{d}_*$ now refers to a data subset.
-We chose to make each data subsets four months long because it is long enough to encompass the SSE, and it is short enough to still be computationally tractable.
+We use the REML method to find the optimal hyperparameters for $T$ and $X$ for each subset of data.
+We chose to make each data subsets four months long because it is long enough to encompass a SSE in Cascadia, while it is short enough to still be computationally tractable.
 However, four months is too short to resolve the sinusoids in $\bm{d}$, and they are omitted from $\bm{d}$ for the REML analysis.
@@
 This suggests that the median of estimated hyperparameters should be an appropriate prior model for all Cascadia SSEs.
-For the IBM model, there are several anomalously large values for $\ell$ and $\phi$, which is why they have large IQRs.
+For the IBM model, there are several anomalously large values for $\ell$ and $\phi$, which results in large IQRs.
 
@@
 The orange markers indicate outliers which were automatically detected using the algorithm from Section \ref{sec:Outlier}.
-The error bars show one standard deviation uncertainty.}
+The error bars show one standard deviation uncertainties.
+Note that outliers tend be observed in the winter, suggesting that they related to snow and ice accumulation.}
 \label{fig:Outliers}
@@
 \caption{Optimal hyperparameters for transient displacements determined with the REML method.
-The temporal covariance function is indicated by the ``$T$'' column, and the spatial covariance function, $X$, is the squared exponential in all cases.
+The temporal covariance function is indicated by the ``$T$'' column.
+The SE, IBM, and Wendland covariance functions are defined in eqs.
+(\ref{eq:TimeSE}), (\ref{eq:IBM}), and (\ref{eq:Wendland}), respectively.
+The spatial covariance function, $X$, is the squared exponential (eq.
+\ref{eq:SE}) in all cases.
 The hyperparameters are estimated for each of the seven SSEs considered in this study, and the tabulated values indicate the median and interquartile ranges of estimates.
-The ``diff $\log$(REML)'' column compares the log REML likelihood to the log REML likelihood for the SE model.
-Positive values indicate that observations are more consistent with the SE model.}
+The ``diff $\log$(REML)'' column compares the log REML likelihood to the log REML likelihood when using the SE covariance function.
+Positive values indicate that observations are more consistent with the SE covariance function.}
 \end{table}
@@
 The REML likelihoods do not definitively indicate whether the SE or Wendland covariance function is preferable.
+
 To further explore which covariance function for $T$ best describes SSEs, we compare the observations to the predicted displacements for each covariance function.
-let $\hat{\bm{d}} = \left(u(\bm{P}) + \bm{G}\bm{a}\right)|\bm{d}_*$ be the data prediction vector.
+We consider the data prediction vector to be $\hat{\bm{d}} = \left(u(\bm{P}) + \bm{G}\bm{a}\right)|\bm{d}_*$.
 It can be shown that $\hat{\bm{d}}$ is normally distributed with mean
@@
 Since the SSE signal is strongest at Station P436, it can be assumed that $T$ adequately describes the signal elsewhere.
-The prediction for the IBM model contains slightly more high frequency, and perhaps spurious, features.
+The prediction for the IBM covariance function contains slightly more high frequency, and perhaps spurious, features.
 The predictions for the Wendland and SE covariance functions are nearly indistinguishable.
@@
 \includegraphics{figures/signal_fit/signal-fit.pdf}
-\caption{Observed displacement at Station P436 and predicted displacements when using different covariance functions for $T$.
-One standard deviation uncertainty is shown for the observations and the predicted displacements when using the SE covariance function.
-For clarity, uncertainties are not shown when using the IBM and Wendland covariance functions.}
+\caption{Observed easting component of displacements at station P436 and predicted displacements when using different covariance functions for $T$.
+The one standard deviation uncertainties are shown for the observations and the predicted displacements when using the SE covariance function.
+For clarity, uncertainties are not shown for the IBM and Wendland covariance functions, but they are nearly equivalent to the uncertainties for the SE covariance function.}
 \label{fig:Fit}
@@
 
-\subsection{Transient Strain Rates} 
-Having established a noise model and a prior for transient displacements, we use the cleaned GNSS dataset described in Section \ref{sec:SignalModel} to calculate transient strain rates in the Puget Sound region.
-We calculate transient strain rate at each day from January 1, 2010 to March 15, 2017.
+\subsection{Transient Strain Rates}\label{sec:Results} 
+Having established a noise model and a prior for transient displacements, we use the cleaned GNSS dataset to calculate transient strain rates in the Puget Sound region.
+We calculate transient strain rates for each day from January 1, 2010 to March 15, 2017.
 The strain rates are estimates at a grid of points spanning the study area.
-In Figure \ref{fig:StrainMap}, we show the transient strain rates on January 1, 2016, which coincides with the height of an SSE.
+In Figure \ref{fig:StrainMap} we show the transient strain rates on January 1, 2016, which coincides with the height of an SSE.
 We have included a video showing the map view of strain rates over time as supplementary material.
@@
 For comparison, estimated secular strain rates indicate trench perpendicular compression throughout this study region \citep{Murray2000,McCaffrey2007,McCaffrey2013}.
-The SSE are thus concentrating tectonically accumulated strain energy trench-ward, and pushing the subduction zone closer to failure.
+The SSE are thus concentrating tectonically accumulated strain energy trench-ward, and presumably pushing the subduction zone closer to failure.
 Similar conclusions have been drawn based on fault slip models \citep[e.g.,][]{Dragert2001,Wech2009,Schmidt2010}, which reveal that SSEs are occurring down-dip of the seismogenic zone.
-The SSEs, which are thrust events, migrate stress up-dip and push the locked portion of the fault closer to failure.
 A key difference between the strain inferred here and strain derived from fault slip models is that our estimated strain rates are not based on an assumed physical model.
@@
 Strain glyphs show the normal strain rate along each azimuth, where orange indicates compression and blue indicates extension.
-The semi-transparent regions indicate one standard deviation uncertainties.}
+The shaded regions indicate one standard deviation uncertainties in the normal strain rate.}
 \label{fig:StrainMap}
@@
 
-In Figure \ref{fig:StrainTs} we show the time dependence of estimated transient strain rates at a position on the Olympic Peninsula, where strain rate magnitudes from SSEs are largest.
-To verify that the estimated strain rates are accurately identifying geophysical signal, we compare the strain rate magnitudes to the frequency of seismic tremor (Figure \ref{fig:StrainMag}).
-The shown strain rate magnitudes have been normalized by their uncertainties and can be considered signal-to-noise ratios.
-Estimated strain rate magnitudes greater than ${\sim}3$ can then be interpreted as a detected transient event.
-For each detected transient event there is a corresponding peak in seismic tremor.
+In Figure \ref{fig:StrainTs} we show the time dependence of estimated transient strain rates at a position on the Olympic Peninsula, where transient strain rates from SSEs are largest.
+To verify that the estimated transient strain rates are accurately identifying geophysical signal, we compare the signal-to-noise ratio from eq.
+(\ref{eq:SNR}), SNR, to the frequency of seismic tremor (Figure \ref{fig:StrainMag}).
+An SNR greater than ${\sim}3$ can be interpreted as a detected geophysical signal.
+For each detected event there is a corresponding peak in seismic tremor.
 We are also able to clearly identify transient strain associated with a more subtle inter-SSE event in August 2014.
-In between SSEs, the normalized strain rate magnitudes are consistently between 0 and 2, indicating that no anomalous non-SSE events are being detected, at least at this location.
+In between SSEs, the SNR is consistently between 0 and 2, indicating that no anomalous non-SSE events are being detected, at least at this location.
 
@@
 \includegraphics{figures/strain_ts/strain-ts.pdf}
-\caption{Three components of the transient strain rate tensor estimated at the position shown in Figure \ref{fig:Context}. The semi-transparent regions indicate one standard deviation uncertainty.}   
+\caption{Three components of the transient horizontal strain rate tensor estimated at the position shown in Figure \ref{fig:Context}.
+The shaded regions indicate one standard deviation uncertainty.}
 \label{fig:StrainTs}
@@
 \includegraphics{figures/strain_ts/mag-ts.pdf}
-\caption{(Top) Strain rate magnitudes at the position shown in Figure \ref{fig:Context} normalized by the one standard deviation uncertainties.
-(bottom) Tremor frequency in the region shown in Figure \ref{fig:Context}.}
+\caption{(Top) Signal-to-noise ratio (eq. \ref{eq:SNR}) at the position shown in Figure \ref{fig:Context}. (bottom) Frequency of tremors in the region shown in Figure \ref{fig:Context}.}   
 \label{fig:StrainMag}
@@
 \section{Discussion}\label{sec:Discussion}
-
-We have demonstrated that transient strain rates estimated with the method described in Section \ref{sec:Method} can be used to detect SSEs or other transient geophysical phenomena.
+We have demonstrated that transient strain rates estimated with the method described in Section \ref{sec:Method} can be used to detect SSEs, and will be robust to detect other transient geophysical phenomena.
 Another potential application would be to use the GNSS derived transient strain rates to develop noise models for borehole strain meters (BSMs).
@@
 The noise consists, in part, of a long-term decay resulting from the instrument equilibrating with the surrounding rock \citep{Gladwin1987}.
-Typically, this noise is dealt with in an ad-hoc manner by fitting and removing exponentials and low order polynomials.
-If we use GNSS derived strain rates as an approximation of the signal in BSM data, then we could perform a more rigorous quantification of BSM noise.
+Typically, this noise is dealt with in an ad-hoc manner by fitting and removing exponentials and low-order polynomials.
+A more rigorous quantification of BSM noise could be performed by using GNSS derived strain rates as an approximation of the signal in BSM data.
 
 We also note that there is potential for a more thorough analysis of the spatio-temporal noise in GNSS data, $v$, than what was performed in Section \ref{sec:NoiseModel}.
-In this study we have not explored the spatial covariance of $v$, since it would not have significantly affected our estimated strain rates.
+In this study, we have not explored the spatial covariance of $v$, since it would not have significantly affected our estimated strain rates.
 If, for example, we were interested in inferring fault slip from GNSS displacements then it may be necessary to account for the spatial covariance of the $v$.
@@
 The method described in this paper assumes a stochastic prior model for displacements, which can be chosen objectively with the Restricted Maximum Likelihood method.
-Because GPR is a Bayesian method, the uncertainties on our estimated transient strain rates are well quantified and we can discern geophysical signal from noise.
-This is demonstrated in our application to Cascadia, where we clearly detect transient strain resulting from slow slip events.
-GPR is not robust against outliers, which are common in GNSS data, and so we have also described an effective pre-processing method for identifying and removing outliers from GNSS data sets.
+Because GPR is a Bayesian method, the uncertainties on our estimated transient strain rates are well quantified, allowing one to discern geophysical signal from noise.
+This is demonstrated in our application to Cascadia, where we detect transient strain resulting from slow slip events.
+GPR is not robust against outliers, which are common in GNSS data, and so we also introduced an effective pre-processing method for identifying and removing outliers from GNSS data sets.
 Another complication with GPR is that it usually involves inverting a dense covariance matrix with rank equal to the number of observations.

### 2017-06-18 HinesHetlandGPR.tex
more crap

@@
 \usepackage{bm}
+\usepackage{ulem}
 \usepackage{setspace}
@@
 
-\newcommand*{\du}[1]{\underline{\underline{#1}}}
+%\newcommand*{\du}[1]{\underline{\underline{#1}}}
+\newcommand*{\du}[1]{\uuline{#1}}
 
@@
 Here we propose using Gaussian process regression (GPR) \citep{Rasmussen2006} to infer transient strain from GNSS data.
-GPR is a Bayesian, non-parametric method for inferring a continuous signal from scattered and potentially noisy data.
-GPR is ideal for synthesizing GNSS data into a spatially and temporally continuous representation of surface deformation, since GNSS stations are irregularly spaced and observation times may differ between stations.
+GPR is a Bayesian, non-parametric method for inferring a continuous signal from scattered data.
+Since GNSS stations are irregularly spaced and observation times may differ between stations, GPR is an ideal tool for synthesizing GNSS data into a spatially and temporally continuous representation of surface deformation.
 GPR is closely related to kriging \citep{Cressie1992} and least squares collocation \citep{Kato1998}, and it has recently gained popularity in the machine learning community.
@@
 The posterior transient strain is also a Gaussian process, and we can use its distribution to confidently discern geophysical signal from noise.
-We use GPR to infer transient strain resulting from slow slip events (SSEs) in Cascadia.
+Here we use GPR to infer transient strain resulting from slow slip events (SSEs) in Cascadia.
 Our results demonstrate that GPR can be an effective tool for detecting SSEs and revealing the spatio-temporal evolution of crustal strain during SSEs.
@@
 We seek a spatially and temporally dependent estimate of transient crustal strain rates, which we consider to be any deviation from the secular strain rates.
+Our attention is limited to horizontal strain rates in this study.
 We denote transient crustal strain rates as $\dot{\du{\varepsilon}}(p)$, where $p$ represents the ordered pair $(\vec{x},t)$, $\vec{x}$ are spatial coordinates in $\mathbb{R}^2$, and $t$ is time.
@@
 For simplicity, we treat each component of displacement independently and ignore any potential covariance.
-We drop the component subscripts with the understanding that the same analysis is being repeated to estimate the east, north, and vertical components of $\vec{u}$.
+We drop the component subscripts with the understanding that the same analysis is being repeated to estimate the easting and northing components of $\vec{u}$.
 We assume that $C_u$ can be separated into positive definite spatial and temporal functions as
 \begin{equation}\label{eq:TransientCovariance}
-C_{u}\left((\vec{x},t),(\vec{x}',t')\right) = X(\vec{x},\vec{x}')T(t,t'),
+C_{u}\left((\vec{x},t),(\vec{x}',t')\right) = X(\vec{x},\vec{x}')T(t,t').
 \end{equation}  
@@
 
-The prior for each component of transient displacements is then conditioned with $\bm{d}_*$ to form a posterior estimate of transient displacements, $\hat{u} = u | \bm{d}_*$.
+The prior for transient displacements is then conditioned with $\bm{d}_*$ to form a posterior estimate of transient displacements, $\hat{u} = u | \bm{d}_*$.
 We will assume that an appropriate covariance function and corresponding hyperparameters for $X$, $T$, and $C_\eta$ have already been chosen.
-If $\kappa$ is kept finite, then following \citet{Rasmussen2006} we find that $\hat{u}$ is distributed as $\mathcal{N}(\mu_{\hat{u}},C_{\hat{u}})$, where
+We discuss how the covariance functions are chosen for our application to Cascadia in Sections \ref{sec:NoiseModel} and \ref{sec:SignalModel}.
+If $\kappa$ is kept finite then, following \citet{Rasmussen2006}, we find that $\hat{u}$ is distributed as $\mathcal{N}(\mu_{\hat{u}},C_{\hat{u}})$, where
 \begin{equation}\label{eq:PosteriorMean}
@@
 (\ref{eq:PosteriorCov}) is not suitable for evaluating this limit.
-We can use the partitioned matrix inversion identity \citep[e.g.,][]{Press2007} to rewrite eq.
+We use the partitioned matrix inversion identity \citep[e.g.,][]{Press2007} to rewrite eq.
 (\ref{eq:PosteriorMean}) and eq.
@@
 We use eq.
-(\ref{eq:PosteriorMean3}) and (\ref{eq:PosteriorCov3}) to find the posterior east and north components of transient displacements.
+(\ref{eq:PosteriorMean3}) and (\ref{eq:PosteriorCov3}) to find the posterior easting and northing components of transient displacements.
 The posterior transient displacements are differentiated temporally and spatially to form an estimate of $\dot{\du{\varepsilon}}$.
@@
 Noting that $||\dot{\du{\varepsilon}}||_\mathrm{F}$ is a random variable, we choose SNR to be the ratio of the estimated mean and standard deviation of $||\dot{\du{\varepsilon}}||_\mathrm{F}$.
-Using nonlinear error-propagation, we find SNR to be
+Using nonlinear uncertainty propagation, we find SNR to be
 \begin{equation}\label{eq:SNR}
-\mathrm{SNR}(p) = \sqrt{\frac{C_{\dot{\varepsilon}_\mathrm{nn}}(p,p)\mu_{\dot{\varepsilon}_\mathrm{nn}}(p)^2 + 
+\mathrm{SNR}(p) = \frac{\mu_{\dot{\varepsilon}_\mathrm{nn}}(p)^2 +
+                        \mu_{\dot{\varepsilon}_\mathrm{ee}}(p)^2 +
+                        2\mu_{\dot{\varepsilon}_\mathrm{en}}(p)^2}
+                       {\big(C_{\dot{\varepsilon}_\mathrm{nn}}(p,p)\mu_{\dot{\varepsilon}_\mathrm{nn}}(p)^2 + 
                               C_{\dot{\varepsilon}_\mathrm{ee}}(p,p)\mu_{\dot{\varepsilon}_\mathrm{ee}}(p)^2 + 
-               4C_{\dot{\varepsilon}_\mathrm{en}}(p,p)\mu_{\dot{\varepsilon}_\mathrm{en}}(p)^2}{
-               \mu_{\dot{\varepsilon}_\mathrm{nn}}(p)^2 +
-               \mu_{\dot{\varepsilon}_\mathrm{ee}}(p)^2 +
-               2\mu_{\dot{\varepsilon}_\mathrm{en}}(p)^2}},
+                              4C_{\dot{\varepsilon}_\mathrm{en}}(p,p)\mu_{\dot{\varepsilon}_\mathrm{en}}(p)^2
+                        \big)^{\frac{1}{2}}}
+,
 \end{equation}
 where the subscripts ``n'' and ``e'' denote north and east, respectively.
+For simplicity, we have ignored covariances between the strain rate components in eq.
+(\ref{eq:SNR}), even though they are non-zero.
 
@@
 In our formulation for estimating transient strain rates, we have assumed that noise in the data vector is normally distributed.
-This is not an appropriate assumption for GNSS displacement time series which are prone to more outliers than would be predicted for normally distributed noise.
-It follows that proposed methods for analyzing GNSS time series should be robust against outliers \citep[e.g.,][]{Blewitt2016}.
+This is not an appropriate assumption for GNSS data which are prone to more outliers than would be predicted for normally distributed noise.
+It follows that proposed methods for analyzing GNSS data should be robust against outliers \citep[e.g.,][]{Blewitt2016}.
 In order to make our estimates of transient strain more robust, we automatically identify and remove outliers in the GNSS data as a pre-processing step.
@@
 We calculate the residuals between the observations and a best fitting model, and data with residuals that are anomalously large are identified as outliers.
-We treat $\bm{d}_*$ as a sample of $\bm{d}$ and assume that there is no temporally correlated noise (i.e., $\eta(p) = 0$).
+We treat $\bm{d}_*$ as a sample of $\bm{d}$ and assume that there is no correlated noise (i.e., $\eta(p) = 0$).
 The best fitting model for $\bm{d}_*$ is considered to be the expected value of the random vector $u(\bm{P}) + \bm{G}\bm{a}$ after conditioning it with non-outlier observations.
@@
 \begin{equation}\label{eq:Update}
-\bm{\Omega} = \{i : |r_i| < \lambda \cdot \mathrm{RMS}\}, \ \ \ r_i \in \bm{r}
+\bm{\Omega} = \{i : |r_i| < \lambda \cdot \mathrm{RMS}\}, \ \ \ r_i \in \bm{r},
 \end{equation} 
@@
 The error bars show one standard deviation uncertainties.
-Note that outliers tend be observed in the winter, suggesting that they related to snow and ice accumulation.}
+Note that outliers tend be observed in the winter, suggesting that they were caused by snow or ice.}
 \label{fig:Outliers}

### 2017-06-18 HinesHetlandGPR.tex
cleaned up figures

@@
 The red dots show the positions of GNSS stations mentioned in this paper.
-The green dot indicates the location of the strain rates shown in Figure \ref{fig:StrainTs} and \ref{fig:StrainMag}.
-The green dashed circle demarcates the spatial extent of the tremors shown in Figure \ref{fig:StrainMag}}.
+The blue dot indicates the location of the strain rates shown in Figure \ref{fig:StrainTs} and \ref{fig:StrainMag}.
+The blue dashed circle demarcates the spatial extent of the tremors shown in Figure \ref{fig:StrainMag}}.
 \label{fig:Context}

### 2017-06-18 figures/context_map/make-plot.py

@@ -80,20 +80,20 @@ ax.plot(x,y,'ko',ms=4,zorder=3)
 idx = np.nonzero(data['id'][...] == 'SC03')[0][0]
 stax,stay = x[idx],y[idx]
 ax.plot(stax,stay,'ro',ms=10,zorder=4)
-ax.text(stax-17000.0,stay-23000.0,'SC03',color='r')
+ax.text(stax-17000.0,stay-23000.0,'SC03',color='r',fontsize=10)
 idx = np.nonzero(data['id'][...] == 'P436')[0][0]
 stax,stay = x[idx],y[idx]
 ax.plot(stax,stay,'ro',ms=10,zorder=4)
-ax.text(stax-17000.0,stay-23000.0,'P436',color='r')
+ax.text(stax-17000.0,stay-23000.0,'P436',color='r',fontsize=10)
 # plot location of strain time series
 stax,stay = bm(-124.03,47.9)
 radius = 150000.0
-ax.plot(stax,stay,'mo',ms=10,zorder=4)
-art = Ellipse((stax,stay),2*radius,2*radius,fc='none',ls='--',ec='m',zorder=4)
+ax.plot(stax,stay,'C0o',ms=10,zorder=4)
+art = Ellipse((stax,stay),2*radius,2*radius,fc='none',ls='--',ec='C0',zorder=4)
 ax.add_artist(art)
 
 # PLOT TREMOR DENSITY
-N = 80
+N = 50
 xlim = ax.get_xlim()
 ylim = ax.get_ylim()
 xitp = np.linspace(xlim[0],xlim[1],N)
@@ -120,6 +120,7 @@ cmap = ListedColormap(colors)
 p = ax.imshow(u,cmap=cmap,zorder=2,extent=(xlim+ylim),origin='lower',
               interpolation='bicubic')
 cbar = transparent_colorbar(p,ax=ax)
+cbar.ax.tick_params(labelsize=10)
 cbar.set_label('tremors per km$\mathregular{^2}$ per year',fontsize=10)
 plt.tight_layout(pad=3.0)
 plt.savefig('context-map.pdf',fmt='pdf')

### 2017-06-18 figures/noise/make-plot.py

@@ -31,8 +31,8 @@ def get_meridians_and_parallels(bm,ticks):
 bins = 20
 files = os.listdir('results')
 
-params0 = {'east':[],'north':[],'vertical':[]}
-params1 = {'east':[],'north':[],'vertical':[]}
+params_beta = {'east':[],'north':[],'vertical':[]}
+params_alpha = {'east':[],'north':[],'vertical':[]}
 lons = []
 lats = []
 for f in files:
@@ -49,70 +49,68 @@ for f in files:
   content = content[content.find('station :'):]
   for dir in ['east','north','vertical']:
     params = _get_field('optimal %s parameters' % dir,content).strip().split(',')
-    params0[dir] += [float(params[0])]
-    params1[dir] += [float(params[1])]
+    params_beta[dir] += [float(params[0])]
+    params_alpha[dir] += [float(params[1])]
 
-east_p0 = params0['east']
-north_p0 = params0['north']
-vertical_p0 = params0['vertical']
-east_p1 = params1['east']
-north_p1 = params1['north']
-vertical_p1 = params1['vertical']
+east_beta = params_beta['east']
+north_beta = params_beta['north']
+vertical_beta = params_beta['vertical']
+east_alpha = params_alpha['east']
+north_alpha = params_alpha['north']
+vertical_alpha = params_alpha['vertical']
 
 stats_box = {'facecolor':'w','alpha':0.8,'boxstyle':'round','edgecolor':'0.8'}
 
-fig,axs = plt.subplots(3,2,figsize=(7,6))
+fig,axs = plt.subplots(3,2,figsize=(7.0,5.5))
 
-axs[0][0].hist(east_p0,np.linspace(0.0,2.0,bins),color='C0',zorder=2)
-text = 'median : %.2f\nIQR : %.2f' % (np.median(east_p0),iqr(east_p0))
-axs[0][0].text(0.55,0.68,text,transform=axs[0][0].transAxes,bbox=stats_box,fontsize=10)
-axs[0][0].grid(zorder=0)
-axs[0][0].set_ylabel('count',fontsize=10)
[…]

### 2017-06-18 figures/strain_map/make-plot.py

@@ -146,7 +146,7 @@ rect = Rectangle([0.63,0.0],0.37,0.3,transform=ax.transAxes,zorder=3,
                  joinstyle='round', 
                  capstyle='round')
 ax.add_artist(rect)
-ax.legend()
+ax.set_title('January 1, 2016',fontsize=10)
 plt.tight_layout(pad=3.0)
 plt.savefig('strain-map.pdf',fmt='pdf')
 plt.show()

### 2017-06-18 figures/strain_ts/make-plot.py

@@ -38,9 +38,6 @@ xidx = np.nonzero(fdx['id'][...] == station)[0][0]
 sta_lon = fdx['longitude'][xidx]
 sta_lat = fdx['latitude'][xidx]
 sta_x,sta_y = bm(sta_lon,sta_lat)
-print(sta_lon)
-print(sta_lat)
-
 # convert to microstrain/year
 dudx = fdx['east'][tidx,xidx] * 1.0e6 * 365.25
 dudx_std = fdx['east_std_dev'][tidx,xidx] * 1.0e6 * 365.25
@@ -63,21 +60,25 @@ mag_std = np.sqrt((exx_std*exx/mag)**2 + (eyy_std*eyy/mag)**2 + (exy_std*2*exy/m
 
 fig,axs = plt.subplots(3,1,sharex=True,figsize=(7,5))
 _setup_ts_ax(axs)
+axs[0].set_title('%.1f$^\mathregular{\circ}$W, %.1f$^\mathregular{\circ}$N' %
+                (-sta_lon,sta_lat),fontsize=10)
+            
 axs[0].plot(time,exx,'C0-',lw=1.5,zorder=2)
 axs[0].fill_between(time,exx-exx_std,exx+exx_std,edgecolor='none',facecolor='C0',alpha=0.4,zorder=2)
 axs[0].grid()
+axs[0].tick_params(labelsize=10)
 axs[0].set_ylabel('east normal\n[$\mathregular{\mu}$strain/yr]',fontsize=10)
 
 axs[1].plot(time,eyy,'C0-',lw=1.5,zorder=2)
 axs[1].fill_between(time,eyy-eyy_std,eyy+eyy_std,edgecolor='none',facecolor='C0',alpha=0.4,zorder=2)
 axs[1].grid()
+axs[1].tick_params(labelsize=10)
 axs[1].set_ylabel('north normal\n[$\mathregular{\mu}$strain/yr]',fontsize=10)
 
 axs[2].plot(time,exy,'C0-',lw=1.5,zorder=2)
 axs[2].fill_between(time,exy-exy_std,exy+exy_std,edgecolor='none',facecolor='C0',alpha=0.4,zorder=2)
 axs[2].grid()
 axs[2].set_ylabel('east-north shear\n[$\mathregular{\mu}$strain/yr]',fontsize=10)
-
 
 axs[2].set_xticks(xticks)
 axs[2].set_xticklabels(xtick_labels) 
[…]

### 2017-06-18 HinesHetlandGPR.tex
added more

@@
 However, in our application to Cascadia, we will be using displacement time series which are long enough to sufficiently constrain $a^{(2)}_{i}$ for each station, avoiding the need to incorporate a prior on it.
-Likewise, the seasonal coefficients may be spatially correlated \citep{Langbein2008}, and it may be worth exploring and exploiting such a correlation in a future study.
+Likewise, seasonal deformation is spatially correlated \citep{Dong2002,Langbein2008}, and it may be worth exploring and exploiting such a correlation in a future study.
 
@@
 \section{Discussion}\label{sec:Discussion}
-We have demonstrated that transient strain rates estimated with the method described in Section \ref{sec:Method} can be used to detect SSEs, and will be robust to detect other transient geophysical phenomena.
-Another potential application would be to use the GNSS derived transient strain rates to develop noise models for borehole strain meters (BSMs).
-The plate Boundary Observatory maintains 43 BSMs in the Pacific Northwest, and it has been demonstrated that BSMs are able to record transient geophysical events such as SSEs \citep[e.g.,][]{Dragert2011}.
-However, noise in BSM data is not well understood, which complicates the use of BSM data for identifying geophysical signal.
+We have demonstrated that transient strain rates estimated with the method described in Section \ref{sec:Method} can be used to detect SSEs and will be robust to detect other transient geophysical phenomena.
+Another potential application would be to use GNSS derived transient strain rates to develop noise models for borehole strain meters (BSMs).
+The Plate Boundary Observatory maintains 43 BSMs in the Pacific Northwest, and it has been demonstrated that BSMs are able to record transient geophysical events such as SSEs \citep[e.g.,][]{Dragert2011}.
+However, noise in BSM data is not well understood, which complicates the use of BSM data for detecting geophysical events.
 The noise consists, in part, of a long-term decay resulting from the instrument equilibrating with the surrounding rock \citep{Gladwin1987}.
 Typically, this noise is dealt with in an ad-hoc manner by fitting and removing exponentials and low-order polynomials.
-A more rigorous quantification of BSM noise could be performed by using GNSS derived strain rates as an approximation of the signal in BSM data.
+A more rigorous quantification of BSM noise could be performed by analyzing the residuals between the the BSM data and GNSS derived strain rates.
 
-We also note that there is potential for a more thorough analysis of the spatio-temporal noise in GNSS data, $v$, than what was performed in Section \ref{sec:NoiseModel}.
-In this study, we have not explored the spatial covariance of $v$, since it would not have significantly affected our estimated strain rates.
-If, for example, we were interested in inferring fault slip from GNSS displacements then it may be necessary to account for the spatial covariance of the $v$.
-
-In future studies, we may explore more sophisticated seasonal models than what is used in eq.
-(\ref{eq:Data}).
-As suggested earlier, we may want to quantify the spatial covariance of seasonal deformation.
-We may also want to explore whether the semiannual terms in eq.
-(\ref{eq:Data}) are necessary or if higher frequency terms are warranted.
+There is potential for a more thorough analysis of the spatio-temporal noise in GNSS data, $\eta$, than what was performed in Section \ref{sec:NoiseModel}.
+In this study, we did not explored the spatial covariance of $\eta$.
+Spatially correlated common mode noise \citep{Wdowinski1997}, which results from factors such as reference frame error, is a non-negligible component of GNSS data.
+We ignore common mode error in this study because transients strain rates are insensitive to it.
+For other geophysical studies based on GNSS data, such as fault slip inversions, it may be necessary to incorporate a spatially covarying noise model \citep[e.g.,][]{Miyazaki2003}.
+We can also improve upon the season model used in this study, which consists of four spatially uncorrelated sinusoids for each station.
+We did not explore the spatial covariance of seasonal deformation or the temporal ``roughness'' (i.e., the number of sinusoids needed to describe the observations).
 The periodic Gaussian process \citep{Mackay1998} is an alternative model for seasonal deformation and is well suited for exploring the roughness of seasonal deformation.

### 2017-06-19 HinesHetlandGPR.tex
more crap
added more

@@
 Transient crustal strain can potentially illuminate geophysical signal that may not be immediately apparent from the GNSS displacement data.
-The ability to detect transient geophysical phenomena is also relevant for seismic hazard because there are instances of major earthquakes being triggered by slow slip events \citep{Roeloffs2006} and postseismic deformation \citep{Freed2001}.
+The ability to detect transient geophysical phenomena is also relevant for assessing seismic hazard because there are instances of major earthquakes being triggered by slow slip events \citep{Roeloffs2006} and postseismic deformation \citep{Freed2001}.
 
 Most methods for estimating crustal strain rates from GNSS data assume some parametric form of the deformation signal.
-The simplest and traditional method for estimating secular crustal strain rates assumes that GNSS derived velocities can be described with a first order polynomial (i.e., having constant spatial gradients) over some subnetwork of the GNSS stations \citep[e.g.,][]{Frank1966,Prescott1976,Savage1986,Feigl1990,Murray2000}.
+The simplest method for estimating secular crustal strain rates assumes that GNSS derived velocities can be described with a first order polynomial (i.e., having constant deformation gradients) over some subnetwork of the GNSS stations \citep[e.g.,][]{Feigl1990,Murray2000}.
 The components of the strain rate tensor  for each subnetwork are then determined through a least squares fit to the observations.
@@
 To help overcome this deficiency, \citet{Shen1996,Shen2015} used an inverse distance weighting scheme, in which the estimated strain rate at some point is primarily controlled by observations at nearby stations.
-However, the methods described in \citet{Shen1996} and \citet{Shen2015} are still formulated by assuming that the deformation gradients are uniform over the entire network.
+However, the methods described in \citet{Shen1996,Shen2015} are still formulated by assuming that the deformation gradients are uniform over the entire network.
 The errors in this assumption manifest as implausibly low formal uncertainties for the estimated strain rates.
@@
 Among the methods tested were the Network Strain Filter (NSF) \citep{Ohtani2010} and the Network Inversion Filter (NIF) \citep{Segall1997}.
-the NSF uses a wavelet parameterization to  describe the spatial component of geophysical signal and the NIF, which is intended for imaging slow fault slip from geodetic data, uses the elastic dislocation Green's functions from \citet{Okada1992}.
+The NSF uses a wavelet parameterization to  describe the spatial component of geophysical signal.
+The NIF, which is intended for imaging slow fault slip from geodetic data, uses the elastic dislocation Green's functions from \citet{Okada1992}.
 For the NSF and NIF, the time dependence of the geophysical signal is modeled as integrated Brownian motion.
@@
 
-Here we propose using Gaussian process regression (GPR) \citep{Rasmussen2006} to infer transient strain from GNSS data.
+Here we propose using Gaussian process regression (GPR) \citep{Rasmussen2006} to estimate transient strain from GNSS data.
 GPR is a Bayesian, non-parametric method for inferring a continuous signal from scattered data.
@@
 For example, Brownian motion, $B(t)$, is a well known Gaussian process in $\mathbb{R}^1$ which has zero mean and covariance function $\mathrm{cov}(B(t),B(t')) = \min(t,t')$, where $t,t' \ge 0$.
-If no prior information is available for the geophysical signal, then maximum likelihood methods can be used to choose a prior Gaussian process which is most consistent with the observations.
+If no prior information is available for the geophysical signal, then maximum likelihood methods can be used to choose a prior Gaussian process that is most consistent with the observations.
 We incorporate GNSS observations with the prior to form a posterior estimate of transient strain.
@@
 \section{Estimating Transient Strain Rates}\label{sec:Method}
-We seek a spatially and temporally dependent estimate of transient crustal strain rates, which we consider to be any deviation from the secular strain rates.
+We seek a spatially and temporally dependent estimate of transient crustal strain rates.
 Our attention is limited to horizontal strain rates in this study.
@@
 The prior for transient displacements is then conditioned with $\bm{d}_*$ to form a posterior estimate of transient displacements, $\hat{u} = u | \bm{d}_*$.
-We will assume that an appropriate covariance function and corresponding hyperparameters for $X$, $T$, and $C_\eta$ have already been chosen.
-We discuss how the covariance functions are chosen for our application to Cascadia in Sections \ref{sec:NoiseModel} and \ref{sec:SignalModel}.
+We will assume that appropriate covariance functions and corresponding hyperparameters for $X$, $T$, and $C_\eta$ have already been chosen.
+In Section \ref{sec:NoiseModel} and \ref{sec:SignalModel}, we discuss how the covariance functions are chosen for our application to GNSS data from Cascadia.
 If $\kappa$ is kept finite then, following \citet{Rasmussen2006}, we find that $\hat{u}$ is distributed as $\mathcal{N}(\mu_{\hat{u}},C_{\hat{u}})$, where
@@
 (\ref{eq:PosteriorMean3}) and (\ref{eq:PosteriorCov3}) to find the posterior easting and northing components of transient displacements.
-The posterior transient displacements are differentiated temporally and spatially to form an estimate of $\dot{\du{\varepsilon}}$.
 Using $\hat{u}_i$ to denote the posterior transient displacements along direction $i$ and $x_i$ to represent the components of $\vec{x}$, we can write the components of $\dot{\du{\varepsilon}}$ as
@@
 Our motivation for estimating transient strain rates is, in part, to detect geophysical phenomena.
-In Section \ref{sec:Results}, we show that strain resulting from SSEs is immediately visible from simply inspecting $\dot{\du{\varepsilon}}$.
-However, if our goal is to automatically detect geophysical events such as SSEs, then we need to define a signal-to-noise ratio, SNR, based on $\dot{\du{\varepsilon}}$.
-We consider the Frobenius norm of $\dot{\du{\varepsilon}}$, $||\dot{\du{\varepsilon}}||_\mathrm{F}$, which is also referred to as the second invariant of strain rate in the geodetic literature.
-We use $||\dot{\du{\varepsilon}}||_\mathrm{F}$ as a metric for the strain rate ``magnitude''.
+We propose using a signal-to-noise ratio, SNR, based on $\dot{\du{\varepsilon}}$ as our detection criterion.
+The Frobenius norm of $\dot{\du{\varepsilon}}$, $||\dot{\du{\varepsilon}}||_\mathrm{F}$, which is also referred to as the second invariant of strain rate in the geodetic literature, is often used as a metric for the strain rate ``magnitude''.
 Noting that $||\dot{\du{\varepsilon}}||_\mathrm{F}$ is a random variable, we choose SNR to be the ratio of the estimated mean and standard deviation of $||\dot{\du{\varepsilon}}||_\mathrm{F}$.
@@
 Our method for detecting outliers is based on the data editing algorithm described in \citet{Gibbs2011}.
-We calculate the residuals between the observations and a best fitting model, and data with residuals that are anomalously large are identified as outliers.
+We calculate the residuals between the observations and a best fitting model.
+Data with residuals that are anomalously large are then identified as outliers.
 We treat $\bm{d}_*$ as a sample of $\bm{d}$ and assume that there is no correlated noise (i.e., $\eta(p) = 0$).
@@
 We use publicly available continuous GNSS data from the University Navstar Consortium (UNAVCO) \citep{Herring2016}, which can be found at www.unavco.org.
-We limit the dataset to the stations and time ranges which are pertinent to seven of the most recent SSEs in the Puget Sound region.
-The earliest SSE considered began in August 2010, and the most recent SSE began in February 2017.
-We consider these most recent SSEs because the station coverage is sufficiently dense for us to use maximum likelihood methods to constrain prior models.
+We limit the dataset to the stations and time ranges which are pertinent to the seven most recent SSEs in the Puget Sound region.
+The earliest SSE considered in this study began in August 2010, and the most recent SSE began in February 2017.
+We use these most recent SSEs because the station coverage is sufficiently dense for us to use maximum likelihood methods to constrain prior models.
 The positions of GNSS stations used to estimate transient strain rates are shown in Figure \ref{fig:Context}.
@@
 \caption{Positions of continuous GNSS stations used to estimate transient strain rates.
-The colored regions indicates the density of detected seismic tremor as determined by \citet{Wech2010}.
+The colored regions indicate the distribution of seismic tremor as determined by \citet{Wech2010}.
 The red dots show the positions of GNSS stations mentioned in this paper.
 The blue dot indicates the location of the strain rates shown in Figure \ref{fig:StrainTs} and \ref{fig:StrainMag}.
-The blue dashed circle demarcates the spatial extent of the tremors shown in Figure \ref{fig:StrainMag}}.
+The blue dashed circle demarcates the spatial extent of the tremors shown in Figure \ref{fig:StrainMag}.}
 \label{fig:Context}
@@
 \subsection{Noise model}\label{sec:NoiseModel}
-Before we determine the transient strain rates, we must establish a prior for the transient deformation, $u$, and the noise, $\eta$.
+Before we determine the transient strain rates, we must establish a prior for the transient displacements, $u$, and the noise, $\eta$.
 In this section we discuss our choice for the noise covariance function $C_\eta$.
@@
 In these studies, temporally correlated noise was described with some combination of Brownian motion, a first-order Gauss-Markov (FOGM) process, and/or flicker noise.
-There is some physical justification for using Brownian motion as a noise model because it accurately describes the power spectrum of motion resulting from instability in geodetic monuments \citep[e.g.,][]{Wyatt1982,Wyatt1989,Langbein1997}.
+There is some physical justification for using Brownian motion as a noise model because it accurately describes the power spectrum of motion resulting from instability in geodetic monuments \citep[e.g.,][]{Wyatt1982,Wyatt1989}.
 Here we describe the time dependence of $\eta$ as a FOGM process and consider $\eta$ to be spatially uncorrelated.
@@
 where $w(t)$ is white noise with unit variance.
-The FOGM process degenerates to the commonly used Brownian motion noise model under the condition that $\alpha=0$ and $\eta(0) = 0$, which is evident from the definition of Brownian motion \citep[e.g.,][]{Papoulis1991}.
-Our noise model which satisfies eq.
+The FOGM process degenerates to the commonly used Brownian motion noise model under the condition that $\alpha=0$ and $\eta(0) = 0$.
+Our noise model that satisfies eq.
 (\ref{eq:FOGMdef}) is a Gaussian process with zero mean and the covariance function
@@
 \end{equation}
-We independently estimate $\bm{\theta}$ for each station, and so $\bm{d}_*$ consists of displacements for an individual station.
-We are also assuming $C_u(p,p')=0$ when estimating the noise hyperparameters for this section.
 \citet{Harville1974} showed that choosing the hyperparameters which maximize eq.
 (\ref{eq:REML}) is equivalent to choosing the hyperparameters which maximize the probability of drawing $\bm{d}_*$ from $\bm{d}$.
-We use the REML method over the maximum likelihood (ML) method \citep[e.g.,][]{Langbein1997} because the REML method properly accounts for the improper prior that we assigned to $\bm{a}$ \citep{Hines2017}.
+We independently estimate $\bm{\theta}$ for each station, and so $\bm{d}_*$ consists of displacements for an individual station.
+We are assuming $u(p)=0$ when estimating the noise hyperparameters for this section.
+We use the REML method over the maximum likelihood (ML) method \citep[e.g.,][]{Langbein1997} because the REML method accounts for the improper prior that we assigned to $\bm{a}$ \citep{Hines2017}.
 The distribution of inferred $\alpha$ and $\beta$ are shown in Figure \ref{fig:NoiseParams}.
-The amplitude of FOGM noise, $\alpha$, for the east and north components is notable low and are clustered around 0.5 mm/yr$^{0.5}$.
-The corresponding estimates of $\beta$ tend to cluster around 0 yr$^{-1}$, suggesting Brownian motion.
+The amplitude of FOGM noise, $\beta$, for the easting and northing components is notable low and are clustered around 0.5 mm/yr$^{0.5}$.
+The corresponding estimates of $\alpha$ tend to cluster around 0 yr$^{-1}$, suggesting Brownian motion.
 We also estimate hyperparameters for the vertical component of displacements, even though verticals are not used for estimating strain.
@@
 The amplitude of FOGM noise for the vertical component is relatively large with a median value of 13.5 mm/yr$^{0.5}$.
-The inferred values for $\beta$ are also higher for the vertical component with a median value of 8.21 yr$^{-1}$.
-In Figure \ref{fig:NoiseSamples}, we use the median values of $\alpha$ and $\beta$ to generate two random samples of FOGM noise.
-The samples span five years and over these five years the east and north samples drift by about 1 mm.
-In the context of detecting SSEs, which produce several mm's of surface displacement on the time-scale of weeks, the estimated FOGM noise for the east and north component is negligible.
+The inferred values for $\alpha$ are also higher for the vertical component with a median value of 8.21 yr$^{-1}$.
+In Figure \ref{fig:NoiseSamples}, we use the median values of $\alpha$ and $\beta$ to generate two random samples of FOGM noise for each component.
+The samples span five years and over these five years the easting and northing samples drift by about 1 mm.
+In the context of detecting SSEs, which produce several mm's of surface displacement on the time-scale of weeks, the estimated FOGM noise for the easting and northing component is negligible.
 In contrast, the estimated FOGM noise for the vertical component is larger than the signal we would expect from SSEs.
-We suspect that the higher amplitude for the FOGM model in the vertical component is accommodating for deficiencies in our rather simple seasonal model.
-Based on this analysis, we henceforth ignore temporally correlated noise in the east and north component because of its low amplitude.
-We also do not use vertical displacements because of the low signal to noise ratio.
+We suspect that the higher amplitude for the FOGM noise in the vertical component is accommodating for deficiencies in our rather simple seasonal model.
+Based on this analysis, we henceforth ignore temporally correlated noise in the easting and northing component because of its low amplitude.
+We also do not use vertical displacements because of the presumably low signal-to-noise ratio.
 
@@
 \ref{eq:FOGM}).
-Hyperparameters are independently estimated for 38 stations in the Pacific Northwest that are east of $121^\circ$W.
+Hyperparameters are estimated with the REML method for 38 stations in the Pacific Northwest that are east of $121^\circ$W.
 ``IQR'' is the inter-quartile range.}
@@
 \includegraphics{figures/noise/noise-samples.pdf}
-\caption{Two FOGM noise samples for each component, where the hyperparameters have been set to the median values from Figure \ref{fig:NoiseParams}.}   
+\caption{Two FOGM noise samples for each component. The FOGM hyperparameters have been set to the median values from Figure \ref{fig:NoiseParams}.}   
 \label{fig:NoiseSamples}
@@
 For the outlier detection algorithm, our prior model, $u$, is chosen to have a length-scale and time-scale which is able to approximately describe SSE displacements.
-We use the SE covariance function for $X$ with length-scale $\ell = 100$ km, and we use the Wendland covariance function for $T$, due to its computational efficiency, with time-scale $\tau = 0.1$ yr and variance $\phi = 1$ mm.
+We use the SE covariance function for $X$ with length-scale $\ell = 100$ km, and we use the Wendland covariance function for $T$, due to its computational efficiency, with time-scale $\tau = 0.1$ yr and $\phi = 1$ mm.
 The outlier detection algorithm is particularly effective at removing outliers for stations at high elevation (Figure \ref{fig:Outliers}), which can be adversely affected by ice or snow during the winter \citep{Lisowski2008}.
-After cleaning the data, we divide it into seven subsets which are four months long and each centered on the time of an SSE.
-The times of seven SSEs are determined with tremor records from \cite{Wech2010}.
+After cleaning the dataset, we divide it into seven subsets which are four months long and each centered on the time of an SSE.
+The times of the seven SSEs are determined with tremor records from \cite{Wech2010}.
 We use the REML method to find the optimal hyperparameters for $T$ and $X$ for each subset of data.
@@
 The estimated hyperparameters for $u$ are summarized in Table 1.
-Based on the interquartile ranges (IQR), the estimated hyperparameters when using the SE and Wendland covariance functions do not vary significantly between SSEs.
+Based on the interquartile ranges (IQR), the estimated hyperparameters for the SE and Wendland covariance functions do not vary significantly between SSEs.
 This suggests that the median of estimated hyperparameters should be an appropriate prior model for all Cascadia SSEs.
@@
 \end{tabular}
-\caption{Optimal hyperparameters for transient displacements determined with the REML method.
+\caption{Optimal hyperparameters for the prior on transient displacements determined with the REML method.
 The temporal covariance function is indicated by the ``$T$'' column.
@@
 The hyperparameters are estimated for each of the seven SSEs considered in this study, and the tabulated values indicate the median and interquartile ranges of estimates.
-The ``diff $\log$(REML)'' column compares the log REML likelihood to the log REML likelihood when using the SE covariance function.
+The ``diff $\log$(REML)'' column compares the log REML likelihood to the log REML likelihood when using the SE covariance function for $T$.
 Positive values indicate that observations are more consistent with the SE covariance function.}
@@
 For other geophysical studies based on GNSS data, such as fault slip inversions, it may be necessary to incorporate a spatially covarying noise model \citep[e.g.,][]{Miyazaki2003}.
-We can also improve upon the season model used in this study, which consists of four spatially uncorrelated sinusoids for each station.
+We can also improve upon the seasonal model used in this study, which consists of four spatially uncorrelated sinusoids for each station.
 We did not explore the spatial covariance of seasonal deformation or the temporal ``roughness'' (i.e., the number of sinusoids needed to describe the observations).
@@
 
-% We could have also used a spherical spatial covariance functions
-
-%Numerical Notes/scaling. Use partitioned matrix and Cholmod to solve for strain.
+We have estimated transient strain rates constrained by about seven years of daily displacement observations from 94 GNSS stations.
+It can be computationally intensive to evaluate eqs.
+(\ref{eq:PosteriorMean3}) and (\ref{eq:PosteriorCov3}) for a dataset with this size.
+We significantly reduced the amount of memory needed to estimate transient strain rates by describing the temporal covariance of displacements with a compact Wendland covariance function.
+By using a compact covariance function for our prior, the matrix being inverted in eqs.
+(\ref{eq:PosteriorMean3}) and (\ref{eq:PosteriorCov3}) becomes sparse.
+We then evaluate eqs.
+(\ref{eq:PosteriorMean3}) and (\ref{eq:PosteriorCov3}) with CHOLMOD, which is designed for solving sparse positive definite systems of equations.
+The matrix being inverted in eqs.
+(\ref{eq:PosteriorMean3}) and (\ref{eq:PosteriorCov3}) is not positive definite; however, we can use another partitioned matrix inversion identity from \citet[e.g.,][]{Press2007} to partition it into positive definite submatrices.
+Even when using a compact covariance function, it may still be necessary to reduce the computational burden by dividing the data into subsets and evaluate transient strain rates for each subset.
 
@@
 This is demonstrated in our application to Cascadia, where we detect transient strain resulting from slow slip events.
-GPR is not robust against outliers, which are common in GNSS data, and so we also introduced an effective pre-processing method for identifying and removing outliers from GNSS data sets.
+GPR is not robust against outliers, which are common in GNSS data, and so we also introduced an effective pre-processing method for identifying and removing outliers from GNSS datasets.
 Another complication with GPR is that it usually involves inverting a dense covariance matrix with rank equal to the number of observations.

### 2017-06-20 HinesHetlandGPR.tex
added more

@@
 The components of the strain rate tensor  for each subnetwork are then determined through a least squares fit to the observations.
-The assumption that deformation gradients are spatially uniform is not appropriate when subnetworks span a large area.
+The assumption that deformation gradients are spatially uniform is not appropriate when subnetworks span too large of an area.
 To help overcome this deficiency, \citet{Shen1996,Shen2015} used an inverse distance weighting scheme, in which the estimated strain rate at some point is primarily controlled by observations at nearby stations.
@@
 We determine $\dot{\du{\varepsilon}}(p)$ by spatially and temporally differentiating estimates of transient displacements, $\vec{u}(p)$.
-As we discuss in Section \ref{sec:NoiseModel} and also as argued for in XXX, estimates of $\dot{\du{\varepsilon}}$ are more effective at illuminating geophysical signal than estimates of $\vec{u}(p)$.
 We make a prior assumption that each component of $\vec{u}$ is a Gaussian process,
@@
 Another significant source of noise in GNSS data is common mode error \citep[e.g.,][]{Wdowinski1997,Dong2006}, which is noise that is highly spatially correlated.
-When not accounted for, common mode error manifests as spatially uniform undulations in $\hat{u}$.
-However, we are primarily interested in estimating strain rates, which are insensitive to common mode error.
+When not accounted for, common mode error manifests as spatially uniform undulations in estimated transient displacements.
+However, estimated transient strain rates are insensitive to common mode error.
 We therefore do not include common mode error in our noise model.
@@
 In Table 1, we summarize how the log REML likelihoods for the Wendland and IBM covariance functions compare to the SE covariance function.
-Based on the differences in log REML likelihoods, the data is substantially more likely to come from the SE or Wendland covariance functions than from the IBM covariance function.
+Based on the differences in log REML likelihoods, the data is substantially more likely to come from a Gaussian process with a SE or Wendland covariance function than an IBM covariance function.
 The REML likelihoods do not definitively indicate whether the SE or Wendland covariance function is preferable.
@@
 We compute $\hat{\bm{d}}$ using SE, Wendland, and IBM covariance functions for $T$ and the median hyperparameters from Table 1.
-Figure \ref{fig:Fit} compares the easting component of $\bm{d}_*$ to $\hat{\bm{d}}$ for the Winter 2016 SSE at Station P436.
+Figure \ref{fig:Fit} compares the easting component of $\bm{d}_*$ to $\hat{\bm{d}}$ for the Winter 2015-2016 SSE at Station P436.
 The data prediction vector appears to accurately describe displacements throughout the SSE, regardless of the choice of $T$.
@@
 For comparison, estimated secular strain rates indicate trench perpendicular compression throughout this study region \citep{Murray2000,McCaffrey2007,McCaffrey2013}.
-The SSE are thus concentrating tectonically accumulated strain energy trench-ward, and presumably pushing the subduction zone closer to failure.
-Similar conclusions have been drawn based on fault slip models \citep[e.g.,][]{Dragert2001,Wech2009,Schmidt2010}, which reveal that SSEs are occurring down-dip of the seismogenic zone.
-A key difference between the strain inferred here and strain derived from fault slip models is that our estimated strain rates are not based on an assumed physical model.
+The SSEs are thus concentrating tectonically accumulated strain energy trench-ward, and presumably pushing the subduction zone closer to failure.
+Similar conclusions have been drawn based on fault slip models \citep[e.g.,][]{Dragert2001,Wech2009,Schmidt2010}, which reveal that SSEs are occurring down-dip of the seismogenic zone and migrating stress upward.
+A key difference between the strain inferred here and strain that can be derived from fault slip models is that our estimated strain rates are not based on an assumed physical model.
 In contrast, fault slip models can be biased by errors in the assumed fault geometry or lithospheric rheology.
@@
 \includegraphics{figures/strain_map/strain-map.pdf}
-\caption{Estimated transient strain rates during the Winter 2016 SSE.
+\caption{Estimated transient strain rates during the Winter 2015-2016 SSE.
 Strain glyphs show the normal strain rate along each azimuth, where orange indicates compression and blue indicates extension.
-The shaded regions indicate one standard deviation uncertainties in the normal strain rate.}
+The shaded regions indicate one standard deviation uncertainties in the normal strain rates.}
 \label{fig:StrainMap}
@@
 \includegraphics{figures/strain_ts/mag-ts.pdf}
-\caption{(Top) Signal-to-noise ratio (eq. \ref{eq:SNR}) at the position shown in Figure \ref{fig:Context}. (bottom) Frequency of tremors in the region shown in Figure \ref{fig:Context}.}   
+\caption{(top) Signal-to-noise ratio (eq. \ref{eq:SNR}) at the position shown in Figure \ref{fig:Context}. (bottom) Frequency of tremors in the region shown in Figure \ref{fig:Context}.}   
 \label{fig:StrainMag}
@@
 In this study, we did not explored the spatial covariance of $\eta$.
-Spatially correlated common mode noise \citep{Wdowinski1997}, which results from factors such as reference frame error, is a non-negligible component of GNSS data.
+Spatially correlated common mode noise, which results from factors such as reference frame error, is a non-negligible component of GNSS data.
 We ignore common mode error in this study because transients strain rates are insensitive to it.
@@
 The periodic Gaussian process \citep{Mackay1998} is an alternative model for seasonal deformation and is well suited for exploring the roughness of seasonal deformation.
-The periodic Gaussian process has zero mean and covariance function,
+The periodic Gaussian process has zero mean and the covariance function
 \begin{equation}\label{eq:Periodic}
@@
 
-We have estimated transient strain rates constrained by about seven years of daily displacement observations from 94 GNSS stations.
+The transient strain rates estimated in this study are constrained by about seven years of daily displacement observations from 94 GNSS stations.
 It can be computationally intensive to evaluate eqs.
 (\ref{eq:PosteriorMean3}) and (\ref{eq:PosteriorCov3}) for a dataset with this size.
-We significantly reduced the amount of memory needed to estimate transient strain rates by describing the temporal covariance of displacements with a compact Wendland covariance function.
-By using a compact covariance function for our prior, the matrix being inverted in eqs.
-(\ref{eq:PosteriorMean3}) and (\ref{eq:PosteriorCov3}) becomes sparse.
-We then evaluate eqs.
-(\ref{eq:PosteriorMean3}) and (\ref{eq:PosteriorCov3}) with CHOLMOD, which is designed for solving sparse positive definite systems of equations.
+We significantly reduce the amount of memory needed to estimate transient strain rates by describing the temporal covariance of displacements with a compact Wendland covariance function.
+Using a compact covariance function for our prior turns eqs.
+(\ref{eq:PosteriorMean3}) and (\ref{eq:PosteriorCov3}) into sparse systems of equations, which we then solve with CHOLMOD.
+CHOLMOD is designed for solving sparse, positive definite systems of equations.
 The matrix being inverted in eqs.
-(\ref{eq:PosteriorMean3}) and (\ref{eq:PosteriorCov3}) is not positive definite; however, we can use another partitioned matrix inversion identity from \citet[e.g.,][]{Press2007} to partition it into positive definite submatrices.
-Even when using a compact covariance function, it may still be necessary to reduce the computational burden by dividing the data into subsets and evaluate transient strain rates for each subset.
+(\ref{eq:PosteriorMean3}) and (\ref{eq:PosteriorCov3}) is not positive definite; however, we can use another partitioned matrix inversion identity from \citet{Press2007} to partition it into positive definite submatrices to be inverted.
+Even when using a compact covariance function, it may still be necessary to reduce the computational burden by dividing the data into subsets and evaluating transient strain rates for each subset.
 
 \section{Conclusion}\label{sec:Conclusion}
-In this paper we have discussed how Gaussian process regression (GPR) can be used to estimate transient strain rates from GNSS data.
-Most other methods for estimating strain from GNSS data assume a parametric representation of deformation, which can bias the estimated strain rates if the parameterization is not chosen carefully.
-The method described in this paper assumes a stochastic prior model for displacements, which can be chosen objectively with the Restricted Maximum Likelihood method.
+In this paper we propose using Gaussian process regression (GPR) to estimate transient strain rates from GNSS data.
+Most other methods for estimating strain rates assume a parametric representation of deformation, which can bias the results if the parameterization is not chosen carefully.
+Here we assume a stochastic, rather than parametric, prior model for displacements.
+Our prior model describes how much we expect transient displacements to covary spatially and temporally.
+If we know nothing about the underlying signal that we are trying to recover, then the prior model can be chosen objectively with maximum likelihood methods.
 Because GPR is a Bayesian method, the uncertainties on our estimated transient strain rates are well quantified, allowing one to discern geophysical signal from noise.
-This is demonstrated in our application to Cascadia, where we detect transient strain resulting from slow slip events.
-GPR is not robust against outliers, which are common in GNSS data, and so we also introduced an effective pre-processing method for identifying and removing outliers from GNSS datasets.
-Another complication with GPR is that it usually involves inverting a dense covariance matrix with rank equal to the number of observations.
-This can be prohibitive when using years of daily GNSS observations from a network of several hundred stations.
-We overcome this complication by using compact Wendland covariance function for our prior model, which allows us to utilize sparse matrix software.
-We believe that Gaussian process regression is a powerful tool that could be applied to a wider range of geophysical problems than what was considered in this paper.
+We demonstrate that GPR is an effective tool for detecting geophysical phenomena, such as slow slip events, in our application to GNSS data from Cascadia.
+One limitation with GPR is that it is not robust against outliers.
+To overcome this limitation, we have introduced an effective pre-processing method for identifying and removing outliers from GNSS datasets.
+Another complication with GPR is that it usually involves inverting a dense matrix where the number of rows and columns is equal to the number of observations.
+This is prohibitive when using several years of daily GNSS observations from a network of several hundred stations.
+We significantly reduce the computational burden of GPR by using compact Wendland covariance function to describe our prior model.
+While this paper just focuses on estimating transient strain rates, we believe that GPR is a powerful tool that can be applied to a wide range of geophysical problems.
 
-% In the previous section we demonstrate that Gaussian processes can be an effective tool for illuminating 
-% GPR can be used for more than just identifying transient strain rates and
-
-% discuss potential applications for slip models
-
-% discuss how we remove errors like reference frame and seasonals
-
-% discuss how this can be used for tikhonov regularization
-
-% we do not interpolate the strain field because it may introduce spurious artifacts Baxter2011
+\section{Acknowledgements}
+This material is based upon work supported by the National Science Foundation under grant EAR 1245263.
+This material is based on EarthScope Plate Boundary Observatory data services provided by UNAVCO through the GAGE Facility with support from the National Science Foundation (NSF) and National Aeronautics and Space Administration (NASA) under NSF Cooperative Agreement EAR-1261833.
 
@@
 
-%The material presented in this paper is primarily focused on GPS data; however, we speculate that the RBF-FD scheme can be of particular use in denoising borehole strain meter (BSM) data.
-The Plate Boundary Observatory has deployed X BSMs along the Western United States.
-BSM data contains low frequency drift resulting from relaxation of the borehole \citep{Gladwin1987}, which can obscure the geophysical signal of interest.
-We suggest that the RBF-FD scheme may be useful in denoising BSM data.
-Since the RBF-FD scheme provides a straight-forward mapping from GPS displacements to strain at any target locations, it is possible to use GPS derived strains as \textit{a priori} information for strain at BSM sites.
-GPS derived strains could then aid in discerning tectonic signal from noise in BSM data.
-
-%Additional uses for strain: transient detection, prior for strain meters,
-
-%Additional potential applications of the RBF-FD method: regularizing inverse problems,
-
-% cite eric calais for triangulation strain
-
 \bibliographystyle{apalike}

### 2017-06-20 tex/gjilguid2e.tex
added more

% gjilguid2e.tex
% V2.0 released 1998 December 18
% V2.1 released 2003 October 7 -- Gregor Hutton, updated the web address for the style files.

\documentclass{gji}
\usepackage{timet}

\title[Geophys.\ J.\ Int.: \LaTeXe\ Guide for Authors]
  {Geophysical Journal International:
   \LaTeXe\ style guide for authors}
\author[B.L.N. Kennett]
  {B.L.N. Kennett$^1$\thanks{Pacific Region Office, GJI} \\
  $^1$ Research School of Earth Sciences, Australian National
    University, Canberra ACT \emph{0200}, Australia
  }
\date{Received 1998 December 18; in original form 1998 November 22}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\volume{200}
\pubyear{1998}

%\def\LaTeX{L\kern-.36em\raise.3ex\hbox{{\small A}}\kern-.15em
%    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}
%\def\LATeX{L\kern-.36em\raise.3ex\hbox{{\Large A}}\kern-.15em
%    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}
% Authors with AMS fonts and mssymb.tex can comment out the following
% line to get the correct symbol for Geophysical Journal International.
\let\leqslant=\leq

\newtheorem{theorem}{Theorem}[section]

\begin{document}

\label{firstpage}

\maketitle

\begin{summary}
 This guide is for authors who are preparing papers for
 \textit{Geophysical Journal International} using the
\LaTeXe\ document preparation system and the GJI class file.
\end{summary}

\begin{keywords}
 \LaTeXe\ -- class files: \verb"gji.cls"\ -- sample text -- user guide.
\end{keywords}

\section{Introduction}

In addition to the standard submission of hardcopy from authors,
\textit{Geophysical Journal International} accepts machine-readable forms
of  papers in \LaTeX.
The layout design for \textit{Geophysical Journal International} has been
implemented as a \LaTeXe\ class file derived from the MN style file for
Monthly Notices of the Royal Astronomical Society. The GJI classfile is
based on the \verb"ARTICLE" style as discussed in the \LaTeX\ manual
\cite {la}. Commands which differ from the standard \LaTeX\ interface, or
which are provided in addition to the standard interface, are explained
in this guide. This guide is not a substitute for the \LaTeX\ manual
itself. Authors planning to submit their papers in \LaTeX\ are advised to
use \verb"gji.cls" as  early as possible in the creation of their
files.  This guide is modified from that produced by Woollatt et al (1994)
to describe the features of the MN style.

A very accessible guide to the features of \LaTeXe and the differences
from the earlier version is provided by Kopka \& Daly \shortcite{kd}.  This
reference provides in chapter 9 a summary of \LaTeX\ error messages and also
a full description of standard \LaTeX\ commands in Appendix F.

\subsection{The GJI document classes}

The use of \LaTeX\ document classes allows a simple change of class
(or class option) to transform the appearance of your document. The
GJI class file preserves the standard \LaTeX\ interface such that any
document which can be produced using the standard \LaTeX\ \verb"ARTICLE"
class can also be produced with the GJI class. However, the measure (or
width of text) is narrower than the default for \verb"ARTICLE", therefore
line breaks will change and long equations may need re-setting.

\subsection{General style issues}

For general style issues, authors are referred to the `Instructions for
Authors' on the inside back cover of \textit{Geophysical Journal
International}. Authors
who are interested in the details of style are referred to Butcher
\shortcite {bu} and The Chicago Manual \shortcite {cm}. The language of
the journal is British English and spelling should conform to this.

Use should be made of symbolic references (\verb"\ref") in order to
protect against late changes of order, etc.

\subsection{Submission of \LaTeX\ articles to the journal}

Papers should initially be submitted in the usual way to:
The Executive Secretary, Royal Astronomical Society, {\em or\/}
the EGS Editor,  {\em or\/} the DGG editor,  {\em or\/}
the American Editor,  {\em or\/} the Pacific Region Editor,
as set out in the Instructions for Authors on the inside back cover of
each issue of Geophysical Journal International.
Four hard copies should be supplied including figures, normally using the
\verb"[referee]" option, for papers with a  high mathematical content the
\verb"[mreferee]" option is recommended. In each case a separate page of
figure captions is preferred.
One of the copies should  be single-sided,
while the other two should  be weight-reduced, by being either
single-spaced or double-sided.   Copies of figures should
also be supplied. Authors should  ensure that their figures are suitable
(in terms of lettering size, etc.)  for the reductions they intend; they
should not attempt to include their figures inside a \TeX\ or \LaTeX\
file by using \verb"\special" or one of the style files for figure
handling.

Note that articles, or revised versions thereof, may not currently be
submitted by electronic mail. However when the article is accepted for
publication the \LaTeX\ file can be sent to the publisher by \verb"ftp"
together with appropriate forms of figures.  Instructions will be provided
following acceptance.

\section{Using the GJI class file}

If the file \verb"gji.cls" is not already in the appropriate
system directory for \LaTeX\ files, either arrange for it to be
put there, or copy it to your working directory. The class file
and related material, such as this guide, can be accessed via the
journal web-site  at
http://www.blackwellpublishing.com/journals/gji under {\em Author
Guidelines}.

The GJI document class is implemented as a complete document class, {\em
not\/} a document class option. In   order to use the GJI style, replace
\verb"article" by
\verb"gji" in the
\verb"\documentclass" command at the beginning of your document:
\begin{verbatim}
\documentclass{article}
\end{verbatim}
is replaced by
\begin{verbatim}
\documentclass{gji}
\end{verbatim}
In general, the following standard document class options should {\em
not\/} be used with the GJI style:
\begin{enumerate}
  \item \texttt{10pt}, \texttt{11pt}, \texttt{12pt} -- unavailable;
  \item \texttt{twoside} (no associated style file) --
     \texttt{twoside} is the default;
  \item \texttt{fleqn}, \texttt{leqno}, \texttt{titlepage} --
        should not be used (\verb"fleqn" is already incorporated into
        the GJI style);
  \item \texttt{twocolumn} -- is not necessary as it is the default style.
\end{enumerate}

In \LaTeX2e the use of postscript fonts and the inclusion of non-standard
options is carried out through the \verb"\usepackage" command, rather than
as options as in earlier versions.  Thus the Times font can be used for
text by including
\begin{verbatim}
\usepackage{times}
\end{verbatim}
on the line immediately after the \verb"\documentclass". If necessary,
\texttt{ifthen} and \texttt{bezier} can be included as packages.

The GJI class file has been designed to operate with the standard
version of \verb"lfonts.tex" that is distributed as part of \LaTeX
. If you have access to the source file for this guide,
\verb"gjilguid2e.tex", attempt to typeset it.  If you find font
problems you might investigate whether a non-standard version of
\verb"lfonts.tex" has been installed in your system.

\subsection{Additional document class options}\label{classoptions}

The following additional class options are available with the GJI style:
\begin{description}
  \item \texttt{onecolumn} -- to be used \textit{only} when two-column output
        is unable to accommodate long equations;
  \item \texttt{landscape} -- for producing wide figures and tables which
        need to be included in landscape format (i.e.\ sideways) rather
        than portrait (i.e.\ upright). This option is described below.
  \item \texttt{doublespacing} -- this will double-space your
        article by setting \verb"\baselinestretch" to 2.
  \item \texttt{referee} -- 12/20pt text size, single column,
        designed for submission of papers.
  \item \texttt{mreferee} -- 11/17pt text size, single column
        designed for submission of papers with mathematical content.
  \item \texttt{camera} -- designed for use with computer modern fonts to
        produce a closer representation of GJI style for camera
        ready material.
  \item \texttt{galley} -- no running heads, no attempt to align
        the bottom of columns.
\end{description}

\subsection{Landscape pages}

If a table or illustration is too wide to fit the standard measure, it
must be turned, with its caption, through 90 degrees anticlockwise.
Landscape illustrations and/or tables cannot be produced directly
using the GJI style file because \TeX\ itself cannot turn the
page, and not all device drivers provide such a facility.
The following procedure can be used to produce such pages.
\begin{enumerate}
  \item Use the \verb"table*" or \verb"figure*" environments in your
        document to create the space for your table or figure on the
        appropriate page of your document. Include an empty
        caption in this environment to ensure the correct
        numbering of subsequent tables and figures. For instance, the
        following code prints a page with the running head, a message
        half way down and the figure number towards the bottom. If you
        are including a plate, the running headline is different, and you
        need to key in the three lines which are marked with \verb"% **",
        with an appropriate headline.
\begin{verbatim}
% ** \clearpage
% ** \thispagestyle{plate}
% ** \plate{Opposite p.~812, GJI, \textbf{135}}
\begin{figure*}
  \vbox to220mm{\vfil Landscape figure to
                go here. \vfil}
  \caption{}
  \label{landfig}
\end{figure*}
\end{verbatim}
\item Create a separate document with the corresponding document style
      but also with the \verb"landscape" document style option, and
      include the \verb"\pagestyle" command, as follows:
\begin{verbatim}
\documentclass[landscape]{gji}
\pagestyle{empty}
\end{verbatim}
  \item Include your complete tables and illustrations (or space for
        these) with captions using the \verb"table*" and \verb"figure*"
        environments.
  \item Before each float environment, use the
        \verb"\setcounter" command to ensure the correct numbering of
        the caption. For example,
\begin{verbatim}
\setcounter{table}{0}
\begin{table*}
 \begin{minipage}{115mm}
 \caption{Images of global seismic tomography.}
 \label{tab1}
 \begin{tabular}{@{}llllcll}
   :
 \end{tabular}
 \end{minipage}
\end{table*}
\end{verbatim}
The corresponding example for a figure would be:
\begin{verbatim}
\clearpage
\setcounter{figure}{12}
\begin{figure*}
 \vspace{144mm}
 \caption{Travel times for regional model.}
 \label{fig13}
\end{figure*}
\end{verbatim}
\end{enumerate}

\section{Additional facilities}

In addition to all the standard \LaTeX\ design elements, the GJI style
includes the following features.
\begin{enumerate}
  \item Extended commands for specifying a short version of the title and
        author(s) for the running headlines;
  \item A \verb"summary" environment to produce a suitably indented
        Summary
  \item An \verb"abstract" environment which produces the GJI style of
        Summary
  \item A \verb"keywords" environment and a \verb"\nokeywords" command;
  \item Use of the \verb"description" environment for unnumbered lists.
  \item A starred version of the \verb"\caption" command to produce
        captions for continued figures or tables.
 \end{enumerate}
 In general, once you have used the additional \verb"gji.cls" facilities
in your document, do not process it with a standard \LaTeX\ style file.

\subsection{Titles and author's name}

In the GJI style, the title of the article and the author's name (or
authors' names) are used both at the beginning of the article for the
main title and throughout the article as running headlines at the top
of every page. The title is used on odd-numbered pages (rectos) and the
author's name appears on even-numbered pages (versos). Although the
main heading can run to several lines of text, the running headline
must be a single line ($\leqslant 45$ characters). Moreover, the main
heading can also incorporate new line commands (e.g. \verb"\\") but
these are not acceptable in a running headline. To enable you to
specify an alternative short title and an alternative short author's
name, the standard \verb"\title" and \verb"\author" commands have been
extended to take an optional argument to be used as the running
headline. The running headlines for this guide were produced using the
following code:
\begin{verbatim}
\title[Geophys.\ J.\ Int.:
       \LaTeXe\ Guide for Authors]
  {Geophysical Journal International:
   \LaTeXe\ style guide for authors}
\end{verbatim}
and
\begin{verbatim}
\author[B.L.N. Kennett]
   {B.L.N. Kennett$^1$
  \thanks{Pacific Region Office, GJI} \\
  $^{1}$Research School of Earth Sciences,
    Australian National University,
    Canberra ACT \emph{0200}, Australia
  }
\end{verbatim}
The \verb"\thanks" note produces a footnote to the title or author.

\subsection{Key words and Summary}

At the beginning of your article, the title should be generated in the
usual way using the \verb"\maketitle" command. Immediately following
the title you should include a Summary followed by a list of key
words. The summary should be enclosed within an \verb"summary"
environment, followed immediately by the key words enclosed in a
\verb"keywords" environment. For example, the titles for this guide
were produced by the following source:
\begin{verbatim}
\maketitle
\begin{summary}
 This guide is for authors who are preparing
 papers for \textit{Geophysical Journal
 International} using the \LaTeXe\ document
 preparation system and the GJI style file.
\end{summary}
\begin{keywords}
 \LaTeXe\ -- class files: \verb"gji.cls"\ --
 sample text -- user guide.
\end{keywords}

\section{Introduction}
  :
\end{verbatim}
The heading `\textbf{Key words}' is included automatically and the key
words are followed by vertical space. If, for any reason, there are no
key words, you should insert the \verb"\nokeywords" command immediately
after the end of the \verb"summary" or \verb"abstract" environment. This
ensures that the   vertical space after the abstract and/or title is
correct and that any
\verb"thanks" acknowledgments are correctly included at the bottom of
the first column. For example,
\begin{verbatim}
\maketitle
\begin{abstract}
  :
\end{abstract}
\nokeywords

\section{Introduction}
  :
\end{verbatim}

Note that the \verb"summary" and \verb"abstract" environments have the same
effect for the documentclass \verb"gji.cls"

\subsection{Lists}

The GJI style provides numbered lists using the \verb"enumerate"
environment and unnumbered lists using the \verb"description"
environment with an empty label. Bulleted lists are not part of the GJI
style and the \verb"itemize" environment should not be used.

The enumerated list numbers each list item with roman numerals:
\begin{enumerate}
  \item first item
  \item second item
  \item third item
\end{enumerate}
Alternative numbering styles can be achieved by inserting a
redefinition of the number labelling command after the
\verb"\begin{enumerate}". For example, the list
\begin{enumerate}
\renewcommand{\theenumi}{(\arabic{enumi})}
  \item first item
  \item second item
  \item etc\ldots
\end{enumerate}
was produced by:
\begin{verbatim}
\begin{enumerate}
 \renewcommand{\theenumi}{(\arabic{enumi})}
  \item first item
       :
\end{enumerate}
\end{verbatim}
Unnumbered lists are provided using the \verb"description" environment.
For example,
\begin{description}
  \item First unnumbered item which has no label and is indented from
        the left margin.
  \item Second unnumbered item.
  \item Third unnumbered item.
\end{description}
was produced by,
\begin{verbatim}
\begin{description}
 \item First unnumbered item...
 \item Second unnumbered item.
 \item Third unnumbered item.
\end{description}
\end{verbatim}

\subsection{Captions for continued figures and tables}

The \verb"\caption*" command may be used to produce a caption with the
same number as the previous caption (for the corresponding type of
float). For instance, if a very large table does not fit on one page,
it must be split into two floats; the second float should use the
\verb"caption*" command with a suitable caption:
\begin{verbatim}
\begin{table}
 \caption*{-- \textit{continued}}
  \begin{tabular}{@{}lccll}
  :
  \end{tabular}
\end{table}
\end{verbatim}

 \begin{figure}
     \vspace{5.5cm}
     \caption{An example figure in which space has been
              left for the artwork.}
     \label{sample-figure}
  \end{figure}

\section[]{Some guidelines for using\\* standard facilities}

The following notes may help you achieve the best effects with the GJI
style file.

\subsection{Sections}

\LaTeX\ provides five levels of section headings and they are all
defined in the GJI style file:
\begin{description}
  \item \verb"\section"
  \item \verb"\subsection"
  \item \verb"\subsubsection"
  \item \verb"\paragraph"
  \item \verb"\subparagraph"
\end{description}
Section numbers are given for section, subsection, subsubsection
and paragraph headings.  Section headings are automatically converted to
upper case; if you need any other style, see the example in section~\ref{headings}.

If you find your section/subsection (etc.)\ headings are wrapping round,
you must use the \verb"\\*" to end individual lines and include the
optional argument \verb"[]" in the section command. This ensures that
the turnover is flushleft.

\subsection{Illustrations (or figures)}

\begin{figure*}
 \vspace{5.5cm}
     \caption{An example figure spanning two-columns
             in which space has been left for the artwork.}
     \label{twocol-figure}
\end{figure*}

The GJI style will cope with positioning of your illustrations and
you should not use the positional qualifiers on the
\verb"figure" environment which would override these decisions. See
`Instructions for Authors' in {\em Geophysical Journal International\/}
for submission of
artwork. Figure captions should be below the figure itself, therefore
the \verb"\caption" command should appear after the figure or space
left for an illustration. For example, Fig.~\ref{sample-figure} is
produced using the following commands:
\begin{verbatim}
\begin{figure}
 \vspace{5.5cm}
 \caption{An example figure in which space has
          been left for the artwork.}
 \label{sample-figure}
\end{figure}
\end{verbatim}

Where a figure needs to span two-columns the \verb"figure*" environment
should be used as in  Fig.~\ref{twocol-figure} using the following commands
\begin{verbatim}
\begin{figure*}
 \vspace{5.5cm}
   \caption{An example figure spanning two-columns
     in which space has been left for the artwork.}
   \label{twocol-figure}
\end{figure*}
\end{verbatim}

\subsection{Tables}

The GJI style will cope with positioning of your tables and you
should not use the positional qualifiers on the
\verb"table" environment which would override these decisions. Table
captions should be at the top, therefore the \verb"\caption" command
should appear before the body of the table.

The \verb"tabular" environment can be used to produce tables with
single horizontal rules, which are allowed, if desired, at the head and
foot only. This environment has been modified for the GJI style in the
following ways:
\begin{enumerate}
  \item additional vertical space is inserted on either side of a rule;
  \item vertical lines are not produced.
\end{enumerate}
Commands to redefine quantities such as \verb"\arraystretch" should be
omitted. For example, Table~\ref{symbols} is produced using the
following commands.
\begin{table}
 \caption{Seismic velocities at major discontinuities.}
 \label{symbols}
 \begin{tabular}{@{}lcccccc}
  Class & depth & radius
        & $\alpha _{-}$ & $\alpha _{+}$
        & $\beta _{-}$ & $\beta _{+}$ \\
  ICB & 5154 & 1217 & 11.091 & 10.258
        & 3.438 &  0. \\
  CMB & 2889 & 3482 & 8.009 & 13.691
        & 0. & 7.301 \\
 \end{tabular}

 \medskip
 The ICB represents the boundary between the inner and outer cores and
the CMB the boundary between the core and the mantle.  Velocities with
subscript $-$ are evaluated just below the discontinuity and
those with subscript $+$ are evaluated just above the discontinuity.
\end{table}
\begin{verbatim}
\begin{table}
 \caption{Seismic velocities at major
          discontinuities.}
 \label{symbols}
 \begin{tabular}{@{}lcccccc}
  Class & depth & radius
        & $\alpha _{-}$ & $\alpha _{+}$
        & $\beta _{-}$ & $\beta _{+}$ \\
  ICB & 5154 & 1217 & 11.091 & 10.258
        & 3.438 &  0. \\
  CMB & 2889 & 3482 & 8.009 & 13.691
        & 0. & 7.301 \\
 \end{tabular}

 \medskip
 The ICB represents the boundary ...
... evaluated just above the discontinuity.

\end{table}
\end{verbatim}

If you have a table that is to extend over two columns, you need to use
\verb"table*" in a minipage environment, i.e., you can say
\begin{verbatim}
\begin{table*}
\begin{minipage}{80mm}
 \caption{Caption which will wrap round to the
          width of the minipage environment.}
 \begin{tabular}{%
      :
 \end{tabular}
\end{minipage}
\end{table*}
\end{verbatim}
The width of the minipage should more or less be the width of your table,
so you can only guess on a value on the first pass. The value will have to
be adjusted when your article is finally typeset, so don't worry
about making it the exact size.

\subsection{Running headlines}

As described above, the title of the article and the author's name (or
authors' names) are used as running headlines at the top of every page.
The headline on right pages can list up to three names; for more than
three use et~al. The \verb"\pagestyle" and \verb"\thispagestyle"
commands should {\em not\/} be used. Similarly, the commands
\verb"\markright" and \verb"\markboth" should not be necessary.

\subsection{Typesetting mathematics}

\subsubsection{Displayed mathematics}

The GJI style will set displayed mathematics flush with the left margin,
provided that you use the \LaTeX\ standard of open and closed square brackets
as delimiters. The equation
\[
 \sum_{i=1}^p \lambda_i =
{\mathrm{trace}}(\mathbf{S})
\]
was typeset in the GJI style using the commands
\begin{verbatim}
\[
 \sum_{i=1}^p \lambda_i =
{\mathrm{trace}}(\mathbf{S})
\]
\end{verbatim}
This correct positioning should be compared with that for
the following centred equation,
$$ \alpha_{j+1} > \bar{\alpha}+ks_{\alpha} $$
which was (wrongly) typeset using double dollars as follows:
\begin{verbatim}
$$ \alpha_{j+1} > \bar{\alpha}+ks_{\alpha} $$
\end{verbatim}
Note that \verb"\mathrm" will produce a roman character
in math mode.

For numbered equations use the \verb"equation" and \verb"eqnarray"
environments which will give the correct positioning.
If equation numbering by section is required the command
\verb"\eqsecnum" should appear after \verb"begin{document}"
at the head of the file.

\subsubsection{Bold math italic}\label{boldmathitalic}

The class file provides a font \verb"\mitbf" defined as:
\begin{verbatim}
\newcommand{\mitbf}[1]{
  \hbox{\mathversion{bold}$#1$}}
\end{verbatim}
Which can be used as follows, to typset the equation
\begin{equation}
  d(\mitbf{{s_{t_u}}}) = \langle [RM(\mitbf{{x_y}}
  + \mitbf{{s_t}}) - RM(\mitbf{{x_y}})]^2 \rangle
\end{equation}
the input should be
\begin{verbatim}
\begin{equation}
  d(\mitbf{{s_{t_u}}}) = \langle [RM(\mitbf{{x_y}}
  + \mitbf{{s_t}}) - RM(\mitbf{{x_y}})]^2 \rangle
\end{equation}
\end{verbatim}

If you are using version 1 of the New Font Selection Scheme, you may
have some messages in your log file that read something like ``Warning:
Font/shape `cmm/b/it' in size~\hbox{$< \!\! 9 \!\! >$} not available
on input line 649. Warning: Using external font `cmmi9' instead on input
line 649.'' If you have such messages, your system will have substituted
math italic characters where you wanted bold math italic ones: you
are advised to upgrade to version 2.

\subsubsection{Bold Greek}\label{boldgreek}

To get bold Greek you use the same
method as for bold math italic. Thus you can input
\begin{verbatim}
\[ \mitbf{{\alpha_{\mu}}} =
\mitbf{\Theta} \alpha. \]
\end{verbatim}
to typeset the equation
\[ \mitbf{{\alpha_{\mu}}} = \mitbf{\Theta} \alpha . \]

\subsection{Points to note in formatting text}\label{formtext}

A number of text characters require special attention
so that \LaTeX\ can properly format a file.

The following characters must be preceded by a
backslash or \LaTeX\ will interpret them as commands:
\begin{quote}
~~~~~~~~~\$~~~\&~~~\%~~~\#~~~\_~~~\{~~~and~~~\}
\end{quote}
must be typed
\begin{center}
\begin{quote}
~~~~~~\verb"\$"~~~\verb"\&"~~~\verb"\%"~~~\verb"\#"
~~~\verb"\_"~~~\verb"\{"~~~and~~~\verb"\}".
\end{quote}
\end{center}

\LaTeX\ interprets all double quotes as closing quotes.
Therefore quotation marks must be typed as pairs of
opening and closing single quotes, for example,
\texttt{ ``quoted text.''}

Note that \LaTeX\ will not recognize greater than or
less than symbols unless they are typed within math
commands (\verb"$>$" or \verb"$<$").

\subsubsection{Special symbols}

The macros for the special symbols in Tables~\ref{mathmode}
and~\ref{anymode}
have been taken from the Springer Verlag `Astronomy and Astrophysics'
design, with their permission. They are directly compatible and use the
same macro names.
These symbols will work in all text sizes, but are only guaranteed to work
in text and displaystyles. Some of the symbols will not get any smaller
when they are used in sub- or superscripts, and will therefore be
displayed at the wrong size. Don't worry about this as the typesetter
will be able to sort this out.
%
\begin{table*}
\begin{minipage}{106mm}
\caption{Special symbols which can only be used in math mode.}
\label{mathmode}
\begin{tabular}{@{}llllll}
Input & Explanation & Output & Input & Explanation & Output\\
\hline
\verb"\la"     & less or approx       & $\la$     &
  \verb"\ga"     & greater or approx    & $\ga$\\[2pt]
\verb"\getsto" & gets over to         & $\getsto$ &
  \verb"\cor"    & corresponds to       & $\cor$\\[2pt]
\verb"\lid"    & less or equal        & $\lid$    &
  \verb"\gid"    & greater or equal     & $\gid$\\[2pt]
\verb"\sol"    & similar over less    & $\sol$    &
  \verb"\sog"    & similar over greater & $\sog$\\[2pt]
\verb"\lse"    & less over simeq      & $\lse$    &
  \verb"\gse"    & greater over simeq   & $\gse$\\[2pt]
\verb"\grole"  & greater over less    & $\grole$  &
  \verb"\leogr"  & less over greater    & $\leogr$\\[2pt]
\verb"\loa"    & less over approx     & $\loa$    &
  \verb"\goa"    & greater over approx  & $\goa$\\
\hline
\end{tabular}
\end{minipage}
\end{table*}
%
\begin{table*}
\begin{minipage}{115mm}
\caption{Special symbols which don't have to be
used in math mode.}
\label{anymode}
\begin{tabular}{@{}llllll}
Input & Explanation & Output & Input & Explanation & Output\\
\hline
\verb"\sun"      & sun symbol            & $\sun$     &
  \verb"\earth"     & earth symbol         & $\earth$   \\[2pt]
\verb"\degr"     & degree                &$\degr$     &
  \verb"\micron"   & \micron               & \micron    \\[2pt]
\verb"\diameter" & diameter              & \diameter  &
  \verb"\sq"       & square                & \squareforqed\\[2pt]
\verb"\fd"       & fraction of day       & \fd        &
  \verb"\fh"       & fraction of hour      & \fh\\[2pt]
\verb"\fm"       & fraction of minute    & \fm        &
  \verb"\fs"       & fraction of second    & \fs\\[2pt]
\verb"\fdg"      & fraction of degree    & \fdg       &
  \verb"\fp"       & fraction of period    & \fp\\[2pt]
\verb"\farcs"    & fraction of arcsecond & \farcs     &
  \verb"\farcm"    & fraction of arcmin    & \farcm\\[2pt]
\verb"\arcsec"   & arcsecond             & \arcsec    &
  \verb"\arcmin"   & arcminute             & \arcmin\\

\hline
\end{tabular}
\end{minipage}
\end{table*}

The command \verb"\chemical" is provided to set chemical species with
an even level for subscripts (not produced in standard mathematics mode).
Thus \verb"\chemical{Fe_{2}^{2+}Cr_{2}O_{4}}" will produce
\chemical{Fe_{2}^{2+}Cr_{2}O_{4}}.

\subsection{Bibliography}

Two methods are provided for managing citations and
references.   The first approach uses the
\verb"\begin{thebibliography}{}"
and \verb"\end{thebibliography}{}" commands.

The second approach uses a simplified scheme using
\verb"\begin{references}" and \verb"\end{references}" commands.

References to published literature should be quoted in text by author
and date; e.g. Draine (1978) or (Begelman, Blandford \& Rees 1984).
Where more than one reference is cited having the same author(s) and date,
the letters a,b,c, \ldots\ should follow the date; e.g.\ Smith (1988a),
Smith (1988b), etc.
The first time you introduce a three-author paper, you should list all
three authors at the first citation, and thereafter, use et al.

\subsubsection{Biblography References in the text}

References in the text are given by author and date, and, whichever
method is used to produce the bibliography, the references in the text
are done in the same way. Each bibliographical entry has a key, which
is assigned by the author and used to refer to that entry in the text.
There is one form of citation -- \verb"\cite{key}" -- to produce the
author and date, and another form -- \verb"\shortcite{key}" -- which
produces the date only. Thus, Rutherford \& Hawker \shortcite{rh} is
produced by
\begin{verbatim}
Rutherford \& Hawker \shortcite{rh}
\end{verbatim}
while \cite{hi} is produced by
\begin{verbatim}
\cite{hi}
\end{verbatim}

\subsubsection{The bibliography}

The following listing shows some references prepared in the style of
the journal; the code produces the references at the end of this guide.
The following rules apply for the ordering of your references:
\begin{enumerate}
  \item if an author has written several papers, some with other authors,
        the rule is that the single-author papers precede the two-author
        papers, which, in turn, precede the multi-author papers;
  \item within the two-author paper citations, the order is determined
        by the second author's surname, regardless of date;
  \item within the multi-author paper citiations, the order is
        chronological, regardless of author's surnames.
\end{enumerate}
%
\begin{verbatim}
\begin{thebibliography}{}
  \bibitem[\protect\citename{Butcher }1992]{bu}
    Butcher J. 1992. \textit{Copy-editing: The
    Cambridge Handbook}, 3rd edn, Cambridge
    Univ. Press, Cambridge.
  \bibitem[\protect\citename{The Chicago Manual }%
    1982]{cm} \textit{The Chicago Manual of Style},
    Univ. Chicago Press, Chicago, 1982.
  \bibitem[\protect\citename{Chao }1985]{ch}
    Chao, B. F., 1985. Normal mode study of the
    Earth's rigid body motions,
    \textit{Geophys. Res. Lett.}, \textbf{12}, 526-529.
  \bibitem[\protect\citename{Hinderer }1986]{hi}
    Hinderer, J., 1986. Resonance effects of the
    earth's fluid core in earth rotation,
    in \textit{Solved and Unsolved Problems},
    pp. 277-296, ed. Cazenave A., Reidel,
    Dordrecht.
  \bibitem[\protect\citename{Lamport }1986]{la}
    Lamport L., 1986,  \LaTeX: \textit{A Document
    Preparation System}, Addison--Wesley, New York
  \bibitem[\protect\citename{Lindberg }1986]{li}
    Lindberg, C., 1986.  Multiple taper harmonic
    analysis of terrestrial free oscillations,
    \textit{PhD thesis}, University of California.
  \bibitem[\protect\citename{Maupin }1992]{ma}
    Maupin, V., 1992. Modelling of laterally
    trapped surface waves with application to
    Rayleigh waves in the Hawaiian swell,
    \textit{Geophys. J. Int.}, \textbf{110}, 553-570.
  \bibitem[\protect\citename{Rutherford
    \& Hawker }1981]{rh} Rutherford, S. R.
    \& Hawker, K. E.,  1981, Consistent coupled
    mode theory of sound propagation for a
    class of non-separable problems,
    \textit{J. acoust. Soc. Am.}, \textbf{71},
    554-564
\end{thebibliography}
\end{verbatim}
Each entry takes the form
\begin{verbatim}
\bibitem[\protect\citename{Author(s), }%
  Date]{tag} Bibliography entry
\end{verbatim}
where \verb"Author(s)" should be the author names as they are cited in
the text, \verb"Date" is the date to be cited in the text, and
\verb"tag" is the tag that is to be used as an argument for the
\verb"\cite{}" and \verb"\shortcite{}" commands. \verb"Bibliography entry"
should be the material that is to appear in the bibliography,
suitably formatted.

\subsubsection{Simplified References and Citations}

The second approach to referencing is taken with permission from
the American Geophysical Union Latex macros

The reference section is started using a
\verb"\begin{references}" command which will
automatically produce a correctly formatted
``Reference'' head.  Each reference is then
preceded by a \verb"\reference"
command.  It is the author's
responsibility to place bibliographic reference
information in the proper order with correct
punctuation.  After the last reference in your
reference section, type an \verb"\end{references}"
command.

Authors may enter properly formatted citations directly
in the manuscript text and enclose those citations in
\verb"\markcite{}" commands.  This approach
marks all citations in your manuscript, but there
is no interaction between the \verb"\markcite"
commands and the reference section.

To create in-text citations, enclose each citation
within a \verb"\markcite" command.
There are two ways to include in-text citations,
depending on the way you phrase your sentence.
You may either include an entire reference within
brackets \markcite{(Merritt et al., 1996)} or you
may mention the author as part of your sentence and
include only the year in brackets, as in \markcite{Ono (1996)}.

As an example
\begin{verbatim}
\begin{references}
\reference
Azimi, Sh.A., Kalinin, A.Y., Kalinin, V.B.,
\& Pivovarov, B.L., 1968.
Impulse and transient characteristics of media
with linear and quadratic absorption laws,
\textit{Izv. Earth Phys.} (English Transl.),
\textbf{2}, 88--93.
\reference
Dahlen, F.A., \& Smith, M.L., 1975.
The influence of rotation on the free
oscillations of the Earth,
\textit{Phil. Trans. R. Soc. London Ser. A},
\textbf{279}, 143--167.
\reference
Durek, J.J., Ritzwoller, M.H.,
\& Woodhouse, J.H., 1993.
Constraining upper mantle anelasticity
using surface wave amplitude anomalies,
\gji, \textbf{114}, 249--272.
\end{references}
\end{verbatim}
produces the reference list
\begin{references}
 \reference
Azimi, Sh.A., Kalinin, A.Y., Kalinin, V.B.,
\& Pivovarov, B.L., 1968.
Impulse and transient characteristics of media
with linear and  quadratic absorption laws,
\textit{Izv. Earth Phys.} (English Transl.),
\textbf{2}, 88--93.
\reference
Dahlen, F.A., \& Smith, M.L., 1975.
The influence of rotation on the free oscillations of the Earth,
\textit{Phil. Trans. R. Soc. London Ser. A}, \textbf{279}, 143--167.
\reference
Durek, J.J., Ritzwoller, M.H., \& Woodhouse, J.H., 1993.
Constraining upper mantle anelasticity
using surface wave amplitude anomalies,
\gji \textbf{114}, 249--272.
\end{references}

\subsubsection{Common Journals}

The following abbreviations are provided for commonly cited
journals and can be used directly in the bibliography.

In the following table the abbreviation and the form of the
associated entry are presented
\newline
\begin{tabular}{ll}
\verb"\areps" & \areps \\
\verb"\bssa"  & \bssa \\
\verb"\eos"   & \eos  \\
\verb"\eps"   & \eps \\
\verb"\epsl"  & \epsl \\
\verb"\gca"   & \gca \\
\verb"\geo"   & \geo \\
\verb"\geop"  & \geop \\
\verb"\gji"   & \gji \\
\verb"\gjras" & \gjras \\
\verb"\grl"   & \grl \\
\verb"\gsab"  & \gsab \\
\verb"\gs"    & \gs \\
\verb"\jgr"   & \jgr \\
\verb"\jseis" & \jseis \\
\verb"\mnras" & \mnras \\
\verb"\pag"   & \pag \\
\verb"\pepi"  & \pepi \\
\verb"\rg"    & \rg \\
\verb"\tecto" & \tecto \\
\end{tabular}
%
% The following two tables have been moved back in the text to
% improve page layout
%
\begin{table*}
\begin{minipage}{130mm}
\caption{Authors' notes.}
\label{authors}
\begin{tabular}{@{}ll}
\verb"\title[optional short title]{long title}"
                    & short title used in running head\\
\verb"\author[optional short author(s)]{long author(s)}"
                    & short author(s) used in running head\\
\verb"\begin{abstract}...\end{abstract}"& for summary on
titlepage\\
\verb"\begin{summary}...\end{summary}"& for abstract on
titlepage\\
\verb"\begin{keywords}...\end{keywords}"& for keywords on titlepage\\
\verb"\nokeywords"  & if there are no keywords on titlepage\\
\verb"\begin{figure*}...\end{figure*}" & for a double spanning figure in two-column mode\\
\verb"\begin{table*}...\end{table*}" & for a double spanning table in
                                       two-column mode\\
\verb"\caption*"    & for continuation figure captions\\
\verb"\resetfigno" & resets figures numbers after an appendix\\
\verb"[referee]" & documentclass option for 12/20pt, single col,
                   for manuscript submission\\
\verb"[mreferee]" & documentclass option for 11/17pt, single col,
                   for submission of papers with extensive mathematics\\
\end{tabular}
\end{minipage}
\end{table*}
%
\begin{table*}
\begin{minipage}{130mm}
\caption{Editors' notes.}
\label{editors}
\begin{tabular}{@{}lp{270pt}}
\verb"\pagerange{000--000}"& for catchline, note use of en-rule\\
\verb"\pagerange{L00--L00}"& for letters option, used in catchline\\
\verb"\volume{000}" & volume number, for catchline\\
\verb"\pubyear{0000}" & publication year, for catchline\\
\verb"\microfiche{GJI000/0}" & for articles accompanied by microfiche\\
\verb"\journal" & replace the whole catchline at one go\\
\verb"[doublespacing]" & documentclass option for doublespacing\\
\verb"[galley]" & documentclass option for running to galley\\
\verb"[landscape]" & documentclass option for landscape illustrations\\
\verb"[fasttrack]" & documentclass option, for rapid short communications
                   (adds F to folios)\\
\verb"[onecolumn]" & documentclass option for one-column \\
\verb"\bsp"& typesets the final phrase `This paper has been produced
 using the Blackwell Publishing GJI \LaTeX2e\ class file.'\\
\end{tabular}
\end{minipage}
\end{table*}

\subsection{Appendices}

The appendices in this guide were generated by typing:
\begin{verbatim}
\appendix
\section{For authors}
     :
\section{For editors}
\end{verbatim}
You only need to type \verb"\appendix" once. Thereafter, every
\verb"\section" command will generate a new appendix which will be
numbered A, B, etc.

If figure captions are to provided after an appendix the figure number can
be reset to avoid extraneous labelling using the command
\verb"\resetfigno".

\section[]{Example of section heading with\\*
  {\mdseries \textsc{S}\lowercase{\textsc{mall}}
  \textsc{C}\lowercase{\textsc{aps}}},
  \lowercase{lowercase},
  \textit{ italic}, and bold\\* Greek such as
  $\mitbf{{\mu^\kappa}}$}\label{headings}

This can be built up using text commands and the \verb"mitbf"
command introduced above

\begin{verbatim}
\section[]{Example of section heading with\\*
  {\mdseries \textsc{S}\lowercase{\textsc{mall}}
  \textsc{C}\lowercase{\textsc{aps}}},
  \lowercase{lowercase},
  \textit{ italic}, and bold\\* Greek such as
  $\mitbf{{\mu^\kappa}}$}\label{headings}
\end{verbatim}

\subsection{Acknowledgments}
Acknowledgments after the main text and before the appendices can be
included with the
\texttt{acknowledgments} environment, as
\begin{verbatim}
\begin{acknowledgments}
We wish to thank ...
\end{acknowledgments}
\end{verbatim}
There is also a corresponding \texttt{acknowledgment} environment for a
single acknowledgment.

\begin{acknowledgments}
A number of colleagues have helped with suggestions for the
improvement of this material and I would particularly like to thank
Bob Geller, University of Tokyo for his criticisms and corrections.
\end{acknowledgments}

\begin{thebibliography}{}
  \bibitem[\protect\citename{Butcher }1992]{bu}
    Butcher J. 1992. \textit{Copy-editing: The Cambridge
    Handbook}, 3rd edn, Cambridge Univ. Press, Cambridge.
  \bibitem[\protect\citename{The Chicago Manual }%
    1982]{cm} \textit{The Chicago Manual of Style}, Univ.
    Chicago Press, Chicago, 1982.
  \bibitem[\protect\citename{Chao }1985]{ch}
    Chao, B. F., 1985. Normal mode study of the Earth's rigid
     body motions, \textit{Geophys. Res. Lett.}, \textbf{12}, 526-529.
  \bibitem[\protect\citename{Hinderer }1986]{hi}
    Hinderer, J., 1986. Resonance effects of the earth's fluid
    core in earth rotation, in \textit{Solved and Unsolved
    Problems}, pp. 277-296, ed. Cazenave A., Reidel,
    Dordrecht.
  \bibitem[\protect\citename{Kopka \& Daly}1995]{kd}
    Kopka H. \& Daly P.W., 1995, \textit{A guide to} \LaTeX2e,
    Addison--Wesley, New York
  \bibitem[\protect\citename{Lamport }1986]{la}
    Lamport L., 1986,  \LaTeX: \textit{A Document
    Preparation System}, Addison--Wesley, New York
  \bibitem[\protect\citename{Lindberg }1986]{bl}
    Lindberg, C., 1986.  Multiple taper harmonic analysis of
    terrestrial free oscillations, \textit{PhD thesis},
    University of California.
  \bibitem[\protect\citename{Maupin }1992]{ma}
    Maupin, V., 1992. Modelling of laterally trapped surface
    waves with application to Rayleigh waves in the Hawaiian
    swell, \textit{Geophys. J. Int.}. \textbf{110}, 553-570.
  \bibitem[\protect\citename{Rutherford \& Hawker }%
    1981]{rh}  Rutherford, S. R. \& Hawker, K. E., 1981,
    Consistent coupled mode theory of sound propagation for a
    class of non-separable problems,
   \textit{J. acoust. Soc. Am.}, \textbf{71}, 554-564
\end{thebibliography}

\appendix
\section{For authors}

Table~\ref{authors} is a list of design macros which are unique to GJI. The
list displays each macro's name and description.

\section{For editors}

The additional features shown in Table~\ref{editors} may be used for
production purposes.

\bsp % ``This paper has been produced using the Blackwell
     %   Publishing GJI \LaTeXe\ class file.''

\label{lastpage}

\end{document}

### 2017-06-20 tex/hines_hetland_2017_gji.tex
added more

--- 2017-06-20 tex/gjilguid2e.tex
@@
 \usepackage{timet}
-
-\title[Geophys.\ J.\ Int.: \LaTeXe\ Guide for Authors]
-  {Geophysical Journal International:
-   \LaTeXe\ style guide for authors}
-\author[B.L.N. Kennett]
-  {B.L.N. Kennett$^1$\thanks{Pacific Region Office, GJI} \\
-  $^1$ Research School of Earth Sciences, Australian National
-    University, Canberra ACT \emph{0200}, Australia
-  }
-\date{Received 1998 December 18; in original form 1998 November 22}
+\usepackage{amsmath}
+\usepackage{amsfonts}
+\usepackage{amssymb}
+\usepackage{natbib}
+\usepackage{graphicx}
+\usepackage{arydshln}
+\usepackage{bm}
+\usepackage{ulem}
+
+\title[Transient strain in geodetic data]
+  {Revealing transient strain in geodetic data with Gaussian process regression}
+\author{T. T. Hines$^1$ and E. A. Hetland$^1$ \\
+        $^1$ Department of Earth and Environmental Sciences, University of Michigan, Ann Arbor, MI, USA}
+\date{Received XXX; in original form XXX}
 \pagerange{\pageref{firstpage}--\pageref{lastpage}}
-\volume{200}
-\pubyear{1998}
-
-%\def\LaTeX{L\kern-.36em\raise.3ex\hbox{{\small A}}\kern-.15em
-%    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}
-%\def\LATeX{L\kern-.36em\raise.3ex\hbox{{\Large A}}\kern-.15em
-%    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}
+\volume{XXX}
+\pubyear{XXX}
+
 % Authors with AMS fonts and mssymb.tex can comment out the following
@@
 
+\newcommand*{\du}[1]{\uuline{#1}}
+
 \newtheorem{theorem}{Theorem}[section]
@@
 \begin{summary}
- This guide is for authors who are preparing papers for
- \textit{Geophysical Journal International} using the
-\LaTeXe\ document preparation system and the GJI class file.
+TODO
 \end{summary}
@@
 \begin{keywords}
- \LaTeXe\ -- class files: \verb"gji.cls"\ -- sample text -- user guide.
+ TODO -- TODO -- TODO -- TODO.
 \end{keywords}
 
-\section{Introduction}
-
-In addition to the standard submission of hardcopy from authors,
-\textit{Geophysical Journal International} accepts machine-readable forms
-of  papers in \LaTeX.
-The layout design for \textit{Geophysical Journal International} has been
-implemented as a \LaTeXe\ class file derived from the MN style file for
-Monthly Notices of the Royal Astronomical Society. The GJI classfile is
-based on the \verb"ARTICLE" style as discussed in the \LaTeX\ manual
-\cite {la}. Commands which differ from the standard \LaTeX\ interface, or
-which are provided in addition to the standard interface, are explained
-in this guide. This guide is not a substitute for the \LaTeX\ manual
-itself. Authors planning to submit their papers in \LaTeX\ are advised to
-use \verb"gji.cls" as  early as possible in the creation of their
-files.  This guide is modified from that produced by Woollatt et al (1994)
-to describe the features of the MN style.
-
-A very accessible guide to the features of \LaTeXe and the differences
-from the earlier version is provided by Kopka \& Daly \shortcite{kd}.  This
-reference provides in chapter 9 a summary of \LaTeX\ error messages and also
-a full description of standard \LaTeX\ commands in Appendix F.
+\section{Introduction}\label{sec:Introduction}
+Crustal strain rates are fundamentally important quantities for assessing seismic hazard.
+Knowing where and how quickly strain is accumulating gives insight into where we can expect stored elastic energy to be released seismically.
+Consequently, secular crustal strain rates estimated from GNSS data have been used to constrain seismic hazard models such as UCERF3 \citep{Field2014}.
+Dense networks of continuous GNSS stations, such as the Plate Boundary Observatory (PBO), make it feasible to estimate transient crustal strain, which we consider to be any deviation from the secular strain rate.
+Transient crustal strain can potentially illuminate geophysical signal that may not be immediately apparent from the GNSS displacement data.
+The ability to detect transient geophysical phenomena is also relevant for assessing seismic hazard because there are instances of major earthquakes being triggered by slow slip events \citep{Roeloffs2006} and postseismic deformation \citep{Freed2001}.
+
+%Most methods for estimating crustal strain rates from GNSS data assume some parametric form of the deformation signal.
+The simplest method for estimating secular crustal strain rates assumes that GNSS derived velocities can be described with a first order polynomial (i.e., having constant deformation gradients) over some subnetwork of the GNSS stations \citep[e.g.,][]{Feigl1990,Murray2000}.
+The components of the strain rate tensor  for each subnetwork are then determined through a least squares fit to the observations.
+The assumption that deformation gradients are spatially uniform is not appropriate when subnetworks span too large of an area.
+To help overcome this deficiency, \citet{Shen1996,Shen2015} used an inverse distance weighting scheme, in which the estimated strain rate at some point is primarily controlled by observations at nearby stations.
+However, the methods described in \citet{Shen1996,Shen2015} are still formulated by assuming that the deformation gradients are uniform over the entire network.
+The errors in this assumption manifest as implausibly low formal uncertainties for the estimated strain rates.
+Other methods for estimating secular strain rates have parameterized GNSS derived velocities with bi-cubic splines \citep{Beavan2001}, spherical wavelets \citep{Tape2009}, and elastostatic Green's functions \citep{Sandwell2016}.
+The type of basis functions and the number of degrees of freedom for a parameterization can be subjective.
+If there are too few degrees of freedom in the parameterization, then estimated strain rates will be biased and the uncertainties will be underestimated.
+On the other hand, if there are too many degrees of freedom, then there will not be any coherent features in the estimated strain rates.
+The methods described by \citet{Beavan2001}, \citet{Tape2009}, and \citet{Sandwell2016} also require the user to specify penalty parameters that control a similar trade-off between bias and variance in the solution.
+One could parameterize deformation with a physically motivated model of interseismic deformation \citep[e.g.,][]{Meade2005,McCaffrey2007}.
+In such models the lithospheric rheology and fault geometries are assumed to be known.
+Any errors in the assumed physical model could similarly result in biased strain estimates and underestimated formal uncertainties.
+
+%The aforementioned studies are concerned with estimating secular strain rates.
+In recent years the Southern California Earthquake Center (SCEC) community has shown interest in developing methods for detecting transient strain.
+SCEC supported a transient detection exercise \citep{Lohman2013}, where several research groups tested their methods for detecting transient geophysical signals with a synthetic GNSS dataset.
+Among the methods tested were the Network Strain Filter (NSF) \citep{Ohtani2010} and the Network Inversion Filter (NIF) \citep{Segall1997}.
+The NSF uses a wavelet parameterization to  describe the spatial component of geophysical signal.
+The NIF, which is intended for imaging slow fault slip from geodetic data, uses the elastic dislocation Green's functions from \citet{Okada1992}.
+For the NSF and NIF, the time dependence of the geophysical signal is modeled as integrated Brownian motion.
+The method described in \citet{Holt2013} was also tested in the SCEC transient detection exercise, which calculates strain rates using a bi-cubic spatial parameterization of displacements between time epochs.
+\citet{Holt2013} defined a detection threshold based on the strain rate magnitude, and below we demonstrate that this is indeed an effective criterion for identifying geophysical signal.
+For the same reasons descibed above, the transient deformation and corresponding uncertainties estimated by these methods can be biased by the chosen spatial parameterization.
+It is then difficult to distinguish signal from noise with these methods, which limits their utility for transient detection.
+
+%Here we propose using Gaussian process regression (GPR) \citep{Rasmussen2006} to estimate transient strain from GNSS data.
+GPR is a Bayesian, non-parametric method for inferring a continuous signal from scattered data.
+Since GNSS stations are irregularly spaced and observation times may differ between stations, GPR is an ideal tool for synthesizing GNSS data into a spatially and temporally continuous representation of surface deformation.
+GPR is closely related to kriging \citep{Cressie1992} and least squares collocation \citep{Kato1998}, and it has recently gained popularity in the machine learning community.
+GPR is Bayesian in that we describe our prior understanding of the geophysical signal with a Gaussian process.
+A Gaussian process is a normally distributed, continuous, stochastic process that is fully defined in terms of a mean function and a positive-definite covariance function.
+For example, Brownian motion, $B(t)$, is a well known Gaussian process in $\mathbb{R}^1$ which has zero mean and covariance function $\mathrm{cov}(B(t),B(t')) = \min(t,t')$, where $t,t' \ge 0$.
+If no prior information is available for the geophysical signal, then maximum likelihood methods can be used to choose a prior Gaussian process that is most consistent with the observations.
+We incorporate GNSS observations with the prior to form a posterior estimate of transient strain.
+The posterior transient strain is also a Gaussian process, and we can use its distribution to confidently discern geophysical signal from noise.
+Here we use GPR to infer transient strain resulting from slow slip events (SSEs) in Cascadia.
+Our results demonstrate that GPR can be an effective tool for detecting SSEs and revealing the spatio-temporal evolution of crustal strain during SSEs.
 
@@
 three authors at the first citation, and thereafter, use et al.
+
+%\bibliographystyle{apalike}
+%\bibliography{mybib}
 
@@
 
-\begin{thebibliography}{}
-  \bibitem[\protect\citename{Butcher }1992]{bu}
-    Butcher J. 1992. \textit{Copy-editing: The Cambridge
-    Handbook}, 3rd edn, Cambridge Univ. Press, Cambridge.
-  \bibitem[\protect\citename{The Chicago Manual }%
-    1982]{cm} \textit{The Chicago Manual of Style}, Univ.
-    Chicago Press, Chicago, 1982.
-  \bibitem[\protect\citename{Chao }1985]{ch}
-    Chao, B. F., 1985. Normal mode study of the Earth's rigid
-     body motions, \textit{Geophys. Res. Lett.}, \textbf{12}, 526-529.
-  \bibitem[\protect\citename{Hinderer }1986]{hi}
-    Hinderer, J., 1986. Resonance effects of the earth's fluid
-    core in earth rotation, in \textit{Solved and Unsolved
-    Problems}, pp. 277-296, ed. Cazenave A., Reidel,
-    Dordrecht.
-  \bibitem[\protect\citename{Kopka \& Daly}1995]{kd}
-    Kopka H. \& Daly P.W., 1995, \textit{A guide to} \LaTeX2e,
-    Addison--Wesley, New York
-  \bibitem[\protect\citename{Lamport }1986]{la}
-    Lamport L., 1986,  \LaTeX: \textit{A Document
-    Preparation System}, Addison--Wesley, New York
-  \bibitem[\protect\citename{Lindberg }1986]{bl}
-    Lindberg, C., 1986.  Multiple taper harmonic analysis of
-    terrestrial free oscillations, \textit{PhD thesis},
-    University of California.
-  \bibitem[\protect\citename{Maupin }1992]{ma}
-    Maupin, V., 1992. Modelling of laterally trapped surface
-    waves with application to Rayleigh waves in the Hawaiian
-    swell, \textit{Geophys. J. Int.}. \textbf{110}, 553-570.
-  \bibitem[\protect\citename{Rutherford \& Hawker }%
-    1981]{rh}  Rutherford, S. R. \& Hawker, K. E., 1981,
-    Consistent coupled mode theory of sound propagation for a
-    class of non-separable problems,
-   \textit{J. acoust. Soc. Am.}, \textbf{71}, 554-564
-\end{thebibliography}
+%\begin{thebibliography}{}
+%  \bibitem[\protect\citename{Butcher }1992]{bu}
+%    Butcher J. 1992. \textit{Copy-editing: The Cambridge
+%    Handbook}, 3rd edn, Cambridge Univ. Press, Cambridge.
+%  \bibitem[\protect\citename{The Chicago Manual }%
+%    1982]{cm} \textit{The Chicago Manual of Style}, Univ.
+%    Chicago Press, Chicago, 1982.
+%  \bibitem[\protect\citename{Chao }1985]{ch}
+%    Chao, B. F., 1985. Normal mode study of the Earth's rigid
+%     body motions, \textit{Geophys. Res. Lett.}, \textbf{12}, 526-529.
+%  \bibitem[\protect\citename{Hinderer }1986]{hi}
+%    Hinderer, J., 1986. Resonance effects of the earth's fluid
+%    core in earth rotation, in \textit{Solved and Unsolved
+%    Problems}, pp. 277-296, ed. Cazenave A., Reidel,
+%    Dordrecht.
+%  \bibitem[\protect\citename{Kopka \& Daly}1995]{kd}
+%    Kopka H. \& Daly P.W., 1995, \textit{A guide to} \LaTeX2e,
+%    Addison--Wesley, New York
+%  \bibitem[\protect\citename{Lamport }1986]{la}
+%    Lamport L., 1986,  \LaTeX: \textit{A Document
+%    Preparation System}, Addison--Wesley, New York
+%  \bibitem[\protect\citename{Lindberg }1986]{bl}
+%    Lindberg, C., 1986.  Multiple taper harmonic analysis of
+%    terrestrial free oscillations, \textit{PhD thesis},
+%    University of California.
+%  \bibitem[\protect\citename{Maupin }1992]{ma}
+%    Maupin, V., 1992. Modelling of laterally trapped surface
+%    waves with application to Rayleigh waves in the Hawaiian
+%    swell, \textit{Geophys. J. Int.}. \textbf{110}, 553-570.
+%  \bibitem[\protect\citename{Rutherford \& Hawker }%
+%    1981]{rh}  Rutherford, S. R. \& Hawker, K. E., 1981,
+%    Consistent coupled mode theory of sound propagation for a
+%    class of non-separable problems,
+%   \textit{J. acoust. Soc. Am.}, \textbf{71}, 554-564
+%\end{thebibliography}
 

### 2017-06-20 tex/figures/context_map/make-plot.py

@@ -1,0 +1,128 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import pygeons
+from mpl_toolkits.basemap import Basemap
+import h5py
+from matplotlib.colors import ListedColormap
+from matplotlib.patches import Ellipse
+from matplotlib.cm import viridis
+from myplot.colorbar import pseudo_transparent_cmap,transparent_colorbar
+from sympy import sqrt,pi,exp
+import rbf
+
+r,eps = rbf.basis.get_r(),rbf.basis.get_eps()
+expr = 1.0/sqrt((2*pi)**2*eps**4)*exp(-r**2/(2*eps**2))
+normal2d = rbf.basis.RBF(expr)
+
+def get_meridians_and_parallels(bm,ticks):
+  ''' 
+  returns the meridians and parallels that should be plotted.
+  '''
+  diff_lon = (bm.urcrnrlon-bm.llcrnrlon)
+  round_digit = int(np.ceil(np.log10(ticks/diff_lon)))
+  dlon = np.round(diff_lon/ticks,round_digit)
+
+  diff_lat = (bm.urcrnrlat-bm.llcrnrlat)
+  round_digit = int(np.ceil(np.log10(ticks/diff_lat)))
+  dlat = np.round(diff_lat/ticks,round_digit)
+
+  meridians = np.arange(np.floor(bm.llcrnrlon),
+                        np.ceil(bm.urcrnrlon),dlon)
+  parallels = np.arange(np.floor(bm.llcrnrlat),
+                        np.ceil(bm.urcrnrlat),dlat)
+
+  return meridians,parallels
+
+def setup_ax(bm,ax):
+  # function which prints out the coordinates on the bottom left 
+  # corner of the figure
+  def coord_string(x,y):
+    str = 'x : %g  y : %g  ' % (x,y)
+    str += '(lon : %g E  lat : %g N)' % bm(x,y,inverse=True)
+    return str
+
+  ax.format_coord = coord_string
+  bm.drawcountries(ax=ax,zorder=1)
+  bm.drawstates(ax=ax,zorder=1)
+  bm.drawcoastlines(ax=ax,zorder=1)
+  bm.drawmapboundary(ax=ax,fill_color=(0.8,0.8,0.8),zorder=0)
[…]

### 2017-06-20 tex/figures/outliers/make_plot.py

@@ -1,0 +1,57 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import pygeons
+from matplotlib import colors
+
+def trend(u1,t1):
+  idx = np.isfinite(u1)
+  u2 = u1[idx]
+  t2 = t1[idx]
+  G = np.array([np.ones(t2.shape[0]),t2]).T
+  m = np.linalg.inv(G.T.dot(G)).dot(G.T).dot(u2)
+  return m
+  
+xtick_dates = ['2011-01-01','2012-01-01','2013-01-01','2014-01-01']
+xticks = [pygeons.mjd.mjd(i,'%Y-%m-%d') for i in xtick_dates]
+xtick_labels = ['2011-01-01','2012-01-01','2013-01-01','2014-01-01']
+
+c1 = colors.to_rgb('C0')
+c1t = tuple(0.6 + 0.4*np.array(colors.to_rgb('C0')))
+c2 = colors.to_rgb('C1')
+c2t = tuple(0.6 + 0.4*np.array(colors.to_rgb('C1')))
+
+# plot raw data
+fig,ax = plt.subplots(figsize=(7,3.0))
+pygeons.plot.plot._setup_ts_ax([ax])
+ax.set_xticks(xticks)
+ax.set_xticklabels(xtick_labels)
+ax.tick_params(labelsize=10)
+ax.set_xlim((55378.0,56839.0))
+ax.set_ylim((-10.0,15.0))
+ax.grid()
+
+data = pygeons.io.io.dict_from_hdf5('data-2017-05-17.h5')
+idx = (data['id'] == 'SC03').nonzero()[0][0]
+lon = data['longitude'][idx]
+lat = data['latitude'][idx]
+
+t = data['time']
+u = 1000*data['east'][:,idx]
+us = 1000*data['east_std_dev'][:,idx]
+m = trend(u,t)
+u = u - m[0] - m[1]*t
+ax.errorbar(t,u,us,marker='.',linestyle='None',color=c2,ecolor=c2t,ms=4.99)
+
+data = pygeons.io.io.dict_from_hdf5('data.final.h5')
+idx = (data['id'] == 'SC03').nonzero()[0][0]
+t = data['time']
+u = 1000*data['east'][:,idx]
+us = 1000*data['east_std_dev'][:,idx]
+u = u - m[0] - m[1]*t
+ax.errorbar(t,u,us,marker='.',linestyle='None',color=c1,ecolor=c1t,ms=5.0)
[…]

### 2017-06-20 InitialSubmission/hines_hetland_2017_gji.tex
reorganized

--- 2017-06-20 HinesHetlandGPR.tex
@@
-\documentclass[10pt,letter]{article}
+\documentclass[extra,mreferee]{gji}
+%\documentclass[extra,onecolumn,doublespacing]{gji}
+\usepackage{timet}
 \usepackage{amsmath}
@@
 \usepackage{amssymb}
-\usepackage{natbib}
 \usepackage{graphicx}
-\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
 \usepackage{arydshln}
@@
 \usepackage{ulem}
-\usepackage{setspace}
-\title{Revealing Transient Strain in Geodetic Data with Gaussian Process Regression}
-\author{Trever T. Hines and Eric A. Hetland}
 
-%\newcommand*{\du}[1]{\underline{\underline{#1}}}
+\title[Transient strain in geodetic data]
+      {Revealing transient strain in geodetic data with Gaussian process regression}
+
+\author[T. T. Hines and E. A. Hetland]
+       {T. T. Hines$^1$ and E. A. Hetland$^1$ \\
+        $^1$ Department of Earth and Environmental Sciences, University of Michigan, Ann Arbor, MI, USA}
+
+\date{Received XXX; in original form XXX}
+
+\pagerange{\pageref{firstpage}--\pageref{lastpage}}
+
+\volume{XXX}
+
+\pubyear{XXX}
+
+% Authors with AMS fonts and mssymb.tex can comment out the following
+% line to get the correct symbol for Geophysical Journal International.
+\let\leqslant=\leq
+
 \newcommand*{\du}[1]{\uuline{#1}}
+
+\newtheorem{theorem}{Theorem}[section]
 
@@
 
+\label{firstpage}
+
 \maketitle
 
-%\doublespacing
+\begin{summary}
+XXX
+\end{summary}
+
+\begin{keywords}
+ XXX -- XXX -- XXX -- XXX.
+\end{keywords}
 
@@
 
-We now consider the column vector of $n$ GNSS observations, $\bm{d}_*$.
-Let $\bm{P}$ be the set of $(\vec{x}_i, t_j)$ pairs describing where and when each of the GNSS observations have been made.
-Let $\bm{a}$ be the vector of coefficients from eq.
+We now consider the column vector of $n$ GNSS observations, $\mitbf{d}_*$.
+Let $\mitbf{P}$ be the set of $(\vec{x}_i, t_j)$ pairs describing where and when each of the GNSS observations have been made.
+Let $\mitbf{a}$ be the vector of coefficients from eq.
 (\ref{eq:Data}) for each of the $m$ GNSS stations.
-We use $\bm{G}$ to represent the $n \times 6m$ matrix of corresponding basis functions evaluated at each point in $\bm{P}$.
-We  also denote the vector of uncorrelated noise for each observation as $\bm{w}$, whose standard deviations are given by the formal data uncertainty, $\bm{\sigma}$.
+We use $\mitbf{G}$ to represent the $n \times 6m$ matrix of corresponding basis functions evaluated at each point in $\mitbf{P}$.
+We  also denote the vector of uncorrelated noise for each observation as $\mitbf{w}$, whose standard deviations are given by the formal data uncertainty, $\mitbf{\sigma}$.
 The observations can then be viewed as a realization of the random vector
 \begin{equation}
-\bm{d} = u(\bm{P}) + \eta(\bm{P}) + \bm{w} + \bm{G}\bm{a},
+\mitbf{d} = u(\mitbf{P}) + \eta(\mitbf{P}) + \mitbf{w} + \mitbf{G}\mitbf{a},
 \end{equation}
-which is distributed as $\mathcal{N}(\bm{0},\bm{\Sigma} + \kappa^2\bm{G}\bm{G}^T)$, where
+which is distributed as $\mathcal{N}(\mitbf{0},\mitbf{\Sigma} + \kappa^2\mitbf{G}\mitbf{G}^T)$, where
 \begin{equation}\label{eq:Cd}
-\bm{\Sigma} = C_u(\bm{P},\bm{P}) + C_\eta(\bm{P},\bm{P}) + 
-              \mathrm{diag}\left(\bm{\sigma}^2\right).  
+\mitbf{\Sigma} = C_u(\mitbf{P},\mitbf{P}) + C_\eta(\mitbf{P},\mitbf{P}) + 
+              \mathrm{diag}\left(\mitbf{\sigma}^2\right).  
 \end{equation}
-It should be understood that notation such as $u(\bm{P})$ and $C_u(\bm{P},\bm{P})$ represents the column vector $[u(P_i)]_{P_i \in \bm{P}}$ and the matrix $[C_u(P_i,P_j)]_{(P_i,P_j) \in \bm{P} \times \bm{P}}$, respectively.
+It should be understood that notation such as $u(\mitbf{P})$ and $C_u(\mitbf{P},\mitbf{P})$ represents the column vector $[u(P_i)]_{P_i \in \mitbf{P}}$ and the matrix $[C_u(P_i,P_j)]_{(P_i,P_j) \in \mitbf{P} \times \mitbf{P}}$, respectively.
 
-The prior for transient displacements is then conditioned with $\bm{d}_*$ to form a posterior estimate of transient displacements, $\hat{u} = u | \bm{d}_*$.
+The prior for transient displacements is then conditioned with $\mitbf{d}_*$ to form a posterior estimate of transient displacements, $\hat{u} = u | \mitbf{d}_*$.
 We will assume that appropriate covariance functions and corresponding hyperparameters for $X$, $T$, and $C_\eta$ have already been chosen.
@@
 \begin{equation}\label{eq:PosteriorMean}
-\mu_{\hat{u}}(p) = C_u(p,\bm{P})\left(\bm{\Sigma} + \kappa^2\bm{G}\bm{G}^T\right)^{-1}\bm{d}_*
+\mu_{\hat{u}}(p) = C_u(p,\mitbf{P})\left(\mitbf{\Sigma} + \kappa^2\mitbf{G}\mitbf{G}^T\right)^{-1}\mitbf{d}_*
 \end{equation}    
@@
 \begin{equation}\label{eq:PosteriorCov}
-C_{\hat{u}}(p,p') = C_u(p,p') - C_u(p,\bm{P})\left(\bm{\Sigma} + \kappa^2\bm{G}\bm{G}^T\right)^{-1}C_u(\bm{P},p').
+C_{\hat{u}}(p,p') = C_u(p,p') - C_u(p,\mitbf{P})\left(\mitbf{\Sigma} + \kappa^2\mitbf{G}\mitbf{G}^T\right)^{-1}C_u(\mitbf{P},p').
 \end{equation}
@@
 \mu_{\hat{u}}(p) = \left[\begin{array}{cc}
-                         C_u(p,\bm{P}) & \bm{0} \\
+                         C_u(p,\mitbf{P}) & \mitbf{0} \\
                          \end{array}\right]
                    \left[\begin{array}{cc}
-                         \bm{\Sigma} & \bm{G} \\
-                         \bm{G}^T  & -\kappa^{-2} \bm{I} \\
+                         \mitbf{\Sigma} & \mitbf{G} \\
+                         \mitbf{G}^T  & -\kappa^{-2} \mitbf{I} \\
                          \end{array}\right]^{-1}
                    \left[\begin{array}{c}
-                         \bm{d}_* \\
-                         \bm{0} \\
+                         \mitbf{d}_* \\
+                         \mitbf{0} \\
                          \end{array}\right]
@@
                     \left[\begin{array}{cc}
-                          C_u(p,\bm{P}) & \bm{0} \\
+                          C_u(p,\mitbf{P}) & \mitbf{0} \\
                           \end{array}\right]
                     \left[\begin{array}{cc}
-                          \bm{\Sigma} & \bm{G} \\
-                          \bm{G}^T  & -\kappa^{-2} \bm{I} \\
+                          \mitbf{\Sigma} & \mitbf{G} \\
+                          \mitbf{G}^T  & -\kappa^{-2} \mitbf{I} \\
                           \end{array}\right]^{-1}
                     \left[\begin{array}{c}
-                          C_u(\bm{P},p') \\
-                          \bm{0} \\
+                          C_u(\mitbf{P},p') \\
+                          \mitbf{0} \\
                           \end{array}\right].
@@
 \mu_{\hat{u}}(p) = \left[\begin{array}{cc}
-                         C_u(p,\bm{P}) & \bm{0} \\
+                         C_u(p,\mitbf{P}) & \mitbf{0} \\
                          \end{array}\right]
                    \left[\begin{array}{cc}
-                         \bm{\Sigma} & \bm{G} \\
-                         \bm{G}^T  & \bm{0} \\
+                         \mitbf{\Sigma} & \mitbf{G} \\
+                         \mitbf{G}^T  & \mitbf{0} \\
                          \end{array}\right]^{-1}
                    \left[\begin{array}{c}
-                         \bm{d}_* \\
-                         \bm{0} \\
+                         \mitbf{d}_* \\
+                         \mitbf{0} \\
                          \end{array}\right]
@@
                     \left[\begin{array}{cc}
-                          C_u(p,\bm{P}) & \bm{0} \\
+                          C_u(p,\mitbf{P}) & \mitbf{0} \\
                           \end{array}\right]
                     \left[\begin{array}{cc}
-                          \bm{\Sigma} & \bm{G} \\
-                          \bm{G}^T  & \bm{0} \\
+                          \mitbf{\Sigma} & \mitbf{G} \\
+                          \mitbf{G}^T  & \mitbf{0} \\
                           \end{array}\right]^{-1}
                     \left[\begin{array}{c}
-                          C_u(\bm{P},p') \\
-                          \bm{0} \\
+                          C_u(\mitbf{P},p') \\
+                          \mitbf{0} \\
                           \end{array}\right].
@@
 Data with residuals that are anomalously large are then identified as outliers.
-We treat $\bm{d}_*$ as a sample of $\bm{d}$ and assume that there is no correlated noise (i.e., $\eta(p) = 0$).
-The best fitting model for $\bm{d}_*$ is considered to be the expected value of the random vector $u(\bm{P}) + \bm{G}\bm{a}$ after conditioning it with non-outlier observations.
+We treat $\mitbf{d}_*$ as a sample of $\mitbf{d}$ and assume that there is no correlated noise (i.e., $\eta(p) = 0$).
+The best fitting model for $\mitbf{d}_*$ is considered to be the expected value of the random vector $u(\mitbf{P}) + \mitbf{G}\mitbf{a}$ after conditioning it with non-outlier observations.
 We still consider $u$ to have a separable covariance function as in eq.
@@
 
-To begin the algorithm, we let $\bm{\Omega}$ be the index set of non-outliers in $\bm{d}_*$ and initiate it with all $n$ indices.
+To begin the algorithm, we let $\mitbf{\Omega}$ be the index set of non-outliers in $\mitbf{d}_*$ and initiate it with all $n$ indices.
 This algorithm is iterative, and for each iteration we calculate the residual vector
 \begin{align}\label{eq:Residual}
-\bm{r} &= \frac{\bm{d}_* - \mathrm{E}\left[(u(\bm{P}) + \bm{G}\bm{a})|\tilde{\bm{d}}_* \right]}{\bm{\sigma}} \\
-       &= \frac{1}{\bm{\sigma}}\left(\bm{d}_*  - 
+\mitbf{r} &= \frac{\mitbf{d}_* - \mathrm{E}\left[(u(\mitbf{P}) + \mitbf{G}\mitbf{a})|\tilde{\mitbf{d}}_* \right]}{\mitbf{\sigma}} \\
+       &= \frac{1}{\mitbf{\sigma}}\left(\mitbf{d}_*  - 
           \left[\begin{array}{cc}
-                C_u(\bm{P},\tilde{\bm{P}}) & \bm{G} \\
+                C_u(\mitbf{P},\tilde{\mitbf{P}}) & \mitbf{G} \\
                 \end{array}\right]
           \left[\begin{array}{cc}
-                C_u(\tilde{\bm{P}},\tilde{\bm{P}}) + \mathrm{diag}(\tilde{\bm{\sigma}}^2) & \tilde{\bm{G}} \\
-                \tilde{\bm{G}}^T  & \bm{0} \\
+                C_u(\tilde{\mitbf{P}},\tilde{\mitbf{P}}) + \mathrm{diag}(\tilde{\mitbf{\sigma}}^2) & \tilde{\mitbf{G}} \\
+                \tilde{\mitbf{G}}^T  & \mitbf{0} \\
                 \end{array}\right]^{-1}
           \left[\begin{array}{c}
-                \tilde{\bm{d}}_* \\
-                \bm{0} \\
+                \tilde{\mitbf{d}}_* \\
+                \mitbf{0} \\
                 \end{array}\right]\right),
 \end{align}
-where the tilde indicates that only elements corresponding to indices in $\bm{\Omega}$ are retained (e.g., $\tilde{\bm{P}} = \{P_i\}_{i\in\bm{\Omega}}$). We then update $\bm{\Omega}$ to be
+where the tilde indicates that only elements corresponding to indices in $\mitbf{\Omega}$ are retained (e.g., $\tilde{\mitbf{P}} = \{P_i\}_{i\in\mitbf{\Omega}}$). We then update $\mitbf{\Omega}$ to be
 \begin{equation}\label{eq:Update}
-\bm{\Omega} = \{i : |r_i| < \lambda \cdot \mathrm{RMS}\}, \ \ \ r_i \in \bm{r},
+\mitbf{\Omega} = \{i : |r_i| < \lambda \cdot \mathrm{RMS}\}, \ \ \ r_i \in \mitbf{r},
 \end{equation} 
-where RMS is the root-mean-square of $\tilde{\bm{r}}$ and $\lambda$ is an outlier tolerance.
+where RMS is the root-mean-square of $\tilde{\mitbf{r}}$ and $\lambda$ is an outlier tolerance.
 We use $\lambda=4$ in this study, which in our experience accurately identifies outliers without unnecessarily decimating the data.
-Iterations continue until the new $\bm{\Omega}$ is equal to the previous $\bm{\Omega}$.
+Iterations continue until the new $\mitbf{\Omega}$ is equal to the previous $\mitbf{\Omega}$.
 
@@
 We then find $\alpha$ and $\beta$ for each station time series with the Restricted Maximum Likelihood (REML) method \cite[e.g.,][]{Harville1974,Cressie1992,Hines2017}.
-The REML method finds the hyperparameters, which we collectively refer to as $\bm{\theta}$, that maximize the likelihood function
+The REML method finds the hyperparameters, which we collectively refer to as $\mitbf{\theta}$, that maximize the likelihood function
 \begin{equation}\label{eq:REML}
-\mathcal{L}(\bm{\theta}) = \left(\frac{\left|\bm{G}^T\bm{G}\right|}
+\mathcal{L}(\mitbf{\theta}) = \left(\frac{\left|\mitbf{G}^T\mitbf{G}\right|}
                            {(2\pi)^{n-6m} 
-                           \left| \bm{\Sigma}(\bm{\theta}) \right| 
-                           \left| \bm{G}^T\bm{\Sigma}\bm{G} \right|}\right)^{\frac{1}{2}} 
-                           e^{-\tfrac{1}{2}\bm{d}_*^T\bm{K}(\bm{\theta})\bm{d}_*},
+                           \left| \mitbf{\Sigma}(\mitbf{\theta}) \right| 
+                           \left| \mitbf{G}^T\mitbf{\Sigma}\mitbf{G} \right|}\right)^{\frac{1}{2}} 
+                           e^{-\tfrac{1}{2}\mitbf{d}_*^T\mitbf{K}(\mitbf{\theta})\mitbf{d}_*},
 \end{equation}
@@
 \begin{equation}
-\bm{K}(\bm{\theta}) = \bm{\Sigma}(\bm{\theta})^{-1} - 
-                      \bm{\Sigma}(\bm{\theta})^{-1}\bm{G}
-         \left(\bm{G}^T\bm{\Sigma}(\bm{\theta})^{-1}\bm{G}\right)^{-1}
-         \bm{G}^T\bm{\Sigma}(\bm{\theta})^{-1}.
+\mitbf{K}(\mitbf{\theta}) = \mitbf{\Sigma}(\mitbf{\theta})^{-1} - 
+                      \mitbf{\Sigma}(\mitbf{\theta})^{-1}\mitbf{G}
+         \left(\mitbf{G}^T\mitbf{\Sigma}(\mitbf{\theta})^{-1}\mitbf{G}\right)^{-1}
+         \mitbf{G}^T\mitbf{\Sigma}(\mitbf{\theta})^{-1}.
 \end{equation}
 \citet{Harville1974} showed that choosing the hyperparameters which maximize eq.
-(\ref{eq:REML}) is equivalent to choosing the hyperparameters which maximize the probability of drawing $\bm{d}_*$ from $\bm{d}$.
-We independently estimate $\bm{\theta}$ for each station, and so $\bm{d}_*$ consists of displacements for an individual station.
+(\ref{eq:REML}) is equivalent to choosing the hyperparameters which maximize the probability of drawing $\mitbf{d}_*$ from $\mitbf{d}$.
+We independently estimate $\mitbf{\theta}$ for each station, and so $\mitbf{d}_*$ consists of displacements for an individual station.
 We are assuming $u(p)=0$ when estimating the noise hyperparameters for this section.
-We use the REML method over the maximum likelihood (ML) method \citep[e.g.,][]{Langbein1997} because the REML method accounts for the improper prior that we assigned to $\bm{a}$ \citep{Hines2017}.
+We use the REML method over the maximum likelihood (ML) method \citep[e.g.,][]{Langbein1997} because the REML method accounts for the improper prior that we assigned to $\mitbf{a}$ \citep{Hines2017}.
 The distribution of inferred $\alpha$ and $\beta$ are shown in Figure \ref{fig:NoiseParams}.
@@
 \end{align}
-IBM has been used in the context of Kalman filtering as a non-parametric model for the time dependence of geophysical signals \citep[e.g.,][]{Segall1997,McGuire2003,Ohtani2010,Hines2016}.
+IBM has been used in the context of Kalman filtering as a non-parametric model for the time dependence of geophysical signals \citep[e.g.,][]{Segall1997,McGuire2003,Ohtani2010,Hines2016a}.
 It should be emphasized $t=0$ is a reference time at which the Gaussian process is exactly zero.
@@
 However, if we are interesting in detecting transient events, where there is no known start time, then IBM may not be an appropriate prior and an isotropic Gaussian process should be preferred.
-In the following analysis, we make the quite arbitrary choice that $t$ is zero on the first epoch of $\bm{d}_*$.
+In the following analysis, we make the quite arbitrary choice that $t$ is zero on the first epoch of $\mitbf{d}_*$.
 Using an earlier reference time does not change the results discussed in this section.
@@
 We chose to make each data subsets four months long because it is long enough to encompass a SSE in Cascadia, while it is short enough to still be computationally tractable.
-However, four months is too short to resolve the sinusoids in $\bm{d}$, and they are omitted from $\bm{d}$ for the REML analysis.
+However, four months is too short to resolve the sinusoids in $\mitbf{d}$, and they are omitted from $\mitbf{d}$ for the REML analysis.
 The estimated hyperparameters for $u$ are summarized in Table 1.
@@
 \begin{table}\label{tab:Parameters}
+\caption{Optimal hyperparameters for the prior on transient displacements determined with the REML method.
+The temporal covariance function is indicated by the ``$T$'' column.
+The SE, IBM, and Wendland covariance functions are defined in eqs.
+(\ref{eq:TimeSE}), (\ref{eq:IBM}), and (\ref{eq:Wendland}), respectively.
+The spatial covariance function, $X$, is the squared exponential (eq.
+\ref{eq:SE}) in all cases.
+The hyperparameters are estimated for each of the seven SSEs considered in this study, and the tabulated values indicate the median and interquartile ranges of estimates.
+The ``diff $\log$(REML)'' column compares the log REML likelihood to the log REML likelihood when using the SE covariance function for $T$.
+Positive values indicate that observations are more consistent with the SE covariance function.}
 \begin{tabular} {l l l l l l}
@@
 \end{tabular}
-\caption{Optimal hyperparameters for the prior on transient displacements determined with the REML method.
-The temporal covariance function is indicated by the ``$T$'' column.
-The SE, IBM, and Wendland covariance functions are defined in eqs.
-(\ref{eq:TimeSE}), (\ref{eq:IBM}), and (\ref{eq:Wendland}), respectively.
-The spatial covariance function, $X$, is the squared exponential (eq.
-\ref{eq:SE}) in all cases.
-The hyperparameters are estimated for each of the seven SSEs considered in this study, and the tabulated values indicate the median and interquartile ranges of estimates.
-The ``diff $\log$(REML)'' column compares the log REML likelihood to the log REML likelihood when using the SE covariance function for $T$.
-Positive values indicate that observations are more consistent with the SE covariance function.}
 \end{table}
@@
 To further explore which covariance function for $T$ best describes SSEs, we compare the observations to the predicted displacements for each covariance function.
-We consider the data prediction vector to be $\hat{\bm{d}} = \left(u(\bm{P}) + \bm{G}\bm{a}\right)|\bm{d}_*$.
-It can be shown that $\hat{\bm{d}}$ is normally distributed with mean
+We consider the data prediction vector to be $\hat{\mitbf{d}} = \left(u(\mitbf{P}) + \mitbf{G}\mitbf{a}\right)|\mitbf{d}_*$.
+It can be shown that $\hat{\mitbf{d}}$ is normally distributed with mean
 \begin{equation}\label{eq:DataPredMean}
-\bm{\mu}_{\hat{d}} = \left[\begin{array}{cc}
-                           C_u(\bm{P},\bm{P}) & \bm{G} \\
+\mitbf{\mu}_{\hat{d}} = \left[\begin{array}{cc}
+                           C_u(\mitbf{P},\mitbf{P}) & \mitbf{G} \\
                            \end{array}\right]
                      \left[\begin{array}{cc}
-                           \bm{\Sigma} & \bm{G} \\
-                           \bm{G}^T  & \bm{0} \\
+                           \mitbf{\Sigma} & \mitbf{G} \\
+                           \mitbf{G}^T  & \mitbf{0} \\
                            \end{array}\right]^{-1}
                      \left[\begin{array}{c}
-                           \bm{d}_* \\
-                           \bm{0} \\
+                           \mitbf{d}_* \\
+                           \mitbf{0} \\
                            \end{array}\right]
@@
 \begin{equation}\label{eq:DataPredCov}
-\bm{C}_{\hat{\bm{d}}} = C_u(\bm{P},\bm{P}) - 
+\mitbf{C}_{\hat{\mitbf{d}}} = C_u(\mitbf{P},\mitbf{P}) - 
                         \left[\begin{array}{cc}
-                              C_u(\bm{P},\bm{P}) & \bm{G} \\
+                              C_u(\mitbf{P},\mitbf{P}) & \mitbf{G} \\
                               \end{array}\right]
                         \left[\begin{array}{cc}
-                              \bm{\Sigma} & \bm{G} \\
-                              \bm{G}^T  & \bm{0} \\
+                              \mitbf{\Sigma} & \mitbf{G} \\
+                              \mitbf{G}^T  & \mitbf{0} \\
                               \end{array}\right]^{-1}
                         \left[\begin{array}{c}
-                              C_u(\bm{P},\bm{P}) \\
-                              \bm{G}^T \\
+                              C_u(\mitbf{P},\mitbf{P}) \\
+                              \mitbf{G}^T \\
                               \end{array}\right].
 \end{equation}
-We compute $\hat{\bm{d}}$ using SE, Wendland, and IBM covariance functions for $T$ and the median hyperparameters from Table 1.
-Figure \ref{fig:Fit} compares the easting component of $\bm{d}_*$ to $\hat{\bm{d}}$ for the Winter 2015-2016 SSE at Station P436.
+We compute $\hat{\mitbf{d}}$ using SE, Wendland, and IBM covariance functions for $T$ and the median hyperparameters from Table 1.
+Figure \ref{fig:Fit} compares the easting component of $\mitbf{d}_*$ to $\hat{\mitbf{d}}$ for the Winter 2015-2016 SSE at Station P436.
 The data prediction vector appears to accurately describe displacements throughout the SSE, regardless of the choice of $T$.
@@
 
-% Discuss similarities with PCA used in Dong2006 and common mode error by Wdowinski1997
-% note that these errors are due to long wavelength features which get removed in strain calculations!!!
-% see strain method by almendinger2007
-
-% mention the scec geodetic transient exercise Lohman2013 and references therein
-
-% Note that segalls 2016 paper mentions the gap between interseismic and sse
-
-% mention howell2016, who said some crap about verticals, mention Hammond2016 so said more sensible crap about verticals
-
-\bibliographystyle{apalike}
+\bibliographystyle{gji}
 \bibliography{mybib}

### 2017-06-20 InitialSubmission/hines_hetland_2017_gji.tex
more crap

@@
 \begin{summary}
-XXX
+Transient strain rates derived from GNSS data can be used to detect and understand geophysical phenomena such as slow slip events and postseismic deformation.
+Here, we propose using Gaussian process regression (GPR) as a tool for estimating transient strain rates from GNSS data.
+GPR is a non-parametric, Bayesian method for interpolating scattered data.
+Transient strain rates estimated with GPR have meaningful uncertainties, which allow us to easily discern geophysical signal from noise.
+In our approach, we assume a stochastic prior model for transient displacements.
+The prior describes how much we expect transient displacements to covary spatially and temporally.
+A posterior estimate of transient strain rates is obtained by differentiating the prior after conditioning it with GNSS data.
+We apply our method to GNSS data from Cascadia, where our intention is to detect transient strain resulting from slow slip events.
+We use maximum likelihood methods to constrain a prior model describing surface displacements resulting from slow slip events.
+The temporal covariance of our prior model is described by a compact Wendland covariance function, which significantly reduces the computational burden that is often associated with GPR.
+Our results reveal the spatial and temporal evolution of strain from slow slip events.
+We verify that the transient strain estimated with GPR is in fact geophysical signal by comparing it to the seismic tremor that is associated with slow slip events in Cascadia.
 \end{summary}
@@
 Consequently, secular crustal strain rates estimated from GNSS data have been used to constrain seismic hazard models such as UCERF3 \citep{Field2014}.
-Dense networks of continuous GNSS stations, such as the Plate Boundary Observatory (PBO), make it feasible to estimate transient crustal strain, which we consider to be any deviation from the secular strain rate.
-Transient crustal strain can potentially illuminate geophysical signal that may not be immediately apparent from the GNSS displacement data.
-The ability to detect transient geophysical phenomena is also relevant for assessing seismic hazard because there are instances of major earthquakes being triggered by slow slip events \citep{Roeloffs2006} and postseismic deformation \citep{Freed2001}.
+Transient crustal strain, which we consider to be any deviation from secular strain rates, is also relevant for assessing seismic hazard.
+Transient strain resulting from geophysical phenomena such as slow slip events or postseismic deformation is itself innocuous and can only be detected with high precision geodetic methods.
+However, there are several instances of major earthquakes being triggered by these transient strain events \citep{Roeloffs2006,Freed2001}.
+Dense networks of continuous GNSS stations, such as the Plate Boundary Observatory (PBO), make it possible to identify transient strain with high fidelity.
+An active area of research is on developing and improving upon methods for deriving strain and, specifically transient strain, from GNSS displacement observations.
 
-Most methods for estimating crustal strain rates from GNSS data assume some parametric form of the deformation signal.
-The simplest method for estimating secular crustal strain rates assumes that GNSS derived velocities can be described with a first order polynomial (i.e., having constant deformation gradients) over some subnetwork of the GNSS stations \citep[e.g.,][]{Feigl1990,Murray2000}.
+Most methods for estimating strain rates from GNSS data assume some parametric form of the deformation signal.
+The simplest method for estimating secular strain rates assumes that GNSS derived velocities can be described with a first order polynomial (i.e., having constant deformation gradients) over some subnetwork of the GNSS stations \citep[e.g.,][]{Feigl1990,Murray2000}.
 The components of the strain rate tensor  for each subnetwork are then determined through a least squares fit to the observations.
@@
 Here we use GPR to infer transient strain resulting from slow slip events (SSEs) in Cascadia.
-Our results demonstrate that GPR can be an effective tool for detecting SSEs and revealing the spatio-temporal evolution of crustal strain during SSEs.
+Our results demonstrate that GPR can be an effective tool for detecting SSEs and revealing the spatio-temporal evolution of strain during SSEs.
 
@@
 Our motivation for estimating transient strain rates is, in part, to detect geophysical phenomena.
-We propose using a signal-to-noise ratio, SNR, based on $\dot{\du{\varepsilon}}$ as our detection criterion.
+We propose using a signal-to-noise ratio, SNR, based on $\dot{\du{\varepsilon}}$ for our detection criterion.
 The Frobenius norm of $\dot{\du{\varepsilon}}$, $||\dot{\du{\varepsilon}}||_\mathrm{F}$, which is also referred to as the second invariant of strain rate in the geodetic literature, is often used as a metric for the strain rate ``magnitude''.

### 2017-06-21 InitialSubmission/hines_hetland_2017_gji.tex
more crap
added more

@@
 
-\newcommand*{\du}[1]{\uuline{#1}}
-
 \newtheorem{theorem}{Theorem}[section]
@@
 Transient strain rates derived from GNSS data can be used to detect and understand geophysical phenomena such as slow slip events and postseismic deformation.
-Here, we propose using Gaussian process regression (GPR) as a tool for estimating transient strain rates from GNSS data.
+Here we propose using Gaussian process regression (GPR) as a tool for estimating transient strain rates from GNSS data.
 GPR is a non-parametric, Bayesian method for interpolating scattered data.
-Transient strain rates estimated with GPR have meaningful uncertainties, which allow us to easily discern geophysical signal from noise.
+Transient strain rates estimated with GPR have meaningful uncertainties, allowing geophysical signal to be easily discerned from noise.
 In our approach, we assume a stochastic prior model for transient displacements.
-The prior describes how much we expect transient displacements to covary spatially and temporally.
-A posterior estimate of transient strain rates is obtained by differentiating the prior after conditioning it with GNSS data.
-We apply our method to GNSS data from Cascadia, where our intention is to detect transient strain resulting from slow slip events.
-We use maximum likelihood methods to constrain a prior model describing surface displacements resulting from slow slip events.
-The temporal covariance of our prior model is described by a compact Wendland covariance function, which significantly reduces the computational burden that is often associated with GPR.
+The prior describes how much one expects transient displacements to covary spatially and temporally.
+A posterior estimate of transient strain rates is obtained by differentiating the posterior displacements.
+One limitation with GPR is that it is not robust against outliers, and so we introduce a pre-processing method for detecting and removing outliers.
+As a demonstration, we use GPR to detect transient strain resulting from slow slip events in Cascadia.
+Maximum likelihood methods are used to constrain a prior model for transient surface displacements in this region.
+The temporal covariance of our prior model is described by a compact Wendland covariance function, which significantly reduces the computational burden that can be associated with GPR.
 Our results reveal the spatial and temporal evolution of strain from slow slip events.
-We verify that the transient strain estimated with GPR is in fact geophysical signal by comparing it to the seismic tremor that is associated with slow slip events in Cascadia.
+We verify that the transient strain estimated with GPR is in fact geophysical signal by comparing it to the seismic tremor that is associated with Cascadia slow slip events.
 \end{summary}
@@
 Consequently, secular crustal strain rates estimated from GNSS data have been used to constrain seismic hazard models such as UCERF3 \citep{Field2014}.
-Transient crustal strain, which we consider to be any deviation from secular strain rates, is also relevant for assessing seismic hazard.
-Transient strain resulting from geophysical phenomena such as slow slip events or postseismic deformation is itself innocuous and can only be detected with high precision geodetic methods.
-However, there are several instances of major earthquakes being triggered by these transient strain events \citep{Roeloffs2006,Freed2001}.
+Transient crustal strain, which is caused by geophysical phenomena such as slow slip events or postseismic deformation, is also relevant for assessing seismic hazard.
+While transient strain itself is not damaging, there is a risk that it can trigger a major earthquake \citep{Roeloffs2006,Freed2001}.
 Dense networks of continuous GNSS stations, such as the Plate Boundary Observatory (PBO), make it possible to identify transient strain with high fidelity.
-An active area of research is on developing and improving upon methods for deriving strain and, specifically transient strain, from GNSS displacement observations.
+An active area of research is developing and improving upon methods for deriving strain and, specifically transient strain, from GNSS displacement observations.
 
@@
 On the other hand, if there are too many degrees of freedom, then there will not be any coherent features in the estimated strain rates.
-The methods described by \citet{Beavan2001}, \citet{Tape2009}, and \citet{Sandwell2016} also require the user to specify penalty parameters that control a similar trade-off between bias and variance in the solution.
+The methods described by \citet{Beavan2001}, \citet{Tape2009}, and \citet{Sandwell2016} also require the user to specify penalty parameters that control a trade-off between bias and variance in the solution.
 One could parameterize deformation with a physically motivated model of interseismic deformation \citep[e.g.,][]{Meade2005,McCaffrey2007}.
@@
 We seek a spatially and temporally dependent estimate of transient crustal strain rates.
-Our attention is limited to horizontal strain rates in this study.
-We denote transient crustal strain rates as $\dot{\du{\varepsilon}}(p)$, where $p$ represents the ordered pair $(\vec{x},t)$, $\vec{x}$ are spatial coordinates in $\mathbb{R}^2$, and $t$ is time.
-We determine $\dot{\du{\varepsilon}}(p)$ by spatially and temporally differentiating estimates of transient displacements, $\vec{u}(p)$.
+We consider transient strain rates to be any deviation from secular strain rates, and our attention is limited to horizontal strain rates in this study.
+We denote transient crustal strain rates as $\dot\varepsilon(p)$, where $p$ represents the ordered pair $(\vec{x},t)$, $\vec{x}$ are spatial coordinates in $\mathbb{R}^2$, and $t$ is time.
+We determine $\dot\varepsilon(p)$ by spatially and temporally differentiating estimates of transient displacements, $\vec{u}(p)$.
 We make a prior assumption that each component of $\vec{u}$ is a Gaussian process,
@@
 For simplicity, we treat each component of displacement independently and ignore any potential covariance.
-We drop the component subscripts with the understanding that the same analysis is being repeated to estimate the easting and northing components of $\vec{u}$.
+Hence, we drop the component subscripts with the understanding that the same analysis is being repeated to estimate the easting and northing components of $\vec{u}$.
 We assume that $C_u$ can be separated into positive definite spatial and temporal functions as
@@
 Of course, the secular velocities, $a^{(2)}_{i}$, are spatially correlated and we could invoke a tectonic model to form a prior on $a^{(2)}_{i}$.
-However, in our application to Cascadia, we will be using displacement time series which are long enough to sufficiently constrain $a^{(2)}_{i}$ for each station, avoiding the need to incorporate a prior on it.
+However, in our application to Cascadia, we will be using displacement time series which are long enough to sufficiently constrain $a^{(2)}_{i}$ for each station, avoiding the need to incorporate a prior.
 Likewise, seasonal deformation is spatially correlated \citep{Dong2002,Langbein2008}, and it may be worth exploring and exploiting such a correlation in a future study.
 
-We now consider the column vector of $n$ GNSS observations, $\mitbf{d}_*$.
+We now consider the column vector of $n$ GNSS observations make at $m$ stations, $\mitbf{d}_*$.
 Let $\mitbf{P}$ be the set of $(\vec{x}_i, t_j)$ pairs describing where and when each of the GNSS observations have been made.
@@
 The prior for transient displacements is then conditioned with $\mitbf{d}_*$ to form a posterior estimate of transient displacements, $\hat{u} = u | \mitbf{d}_*$.
-We will assume that appropriate covariance functions and corresponding hyperparameters for $X$, $T$, and $C_\eta$ have already been chosen.
+For now, we will assume that appropriate covariance functions and corresponding hyperparameters for $X$, $T$, and $C_\eta$ have already been chosen.
 In Section \ref{sec:NoiseModel} and \ref{sec:SignalModel}, we discuss how the covariance functions are chosen for our application to GNSS data from Cascadia.
@@
 (\ref{eq:PosteriorMean3}) and (\ref{eq:PosteriorCov3}) to find the posterior easting and northing components of transient displacements.
-Using $\hat{u}_i$ to denote the posterior transient displacements along direction $i$ and $x_i$ to represent the components of $\vec{x}$, we can write the components of $\dot{\du{\varepsilon}}$ as
+Using $\hat{u}_i$ to denote the posterior transient displacements along direction $i$ and $x_i$ to represent the components of $\vec{x}$, we can write the components of $\dot\varepsilon$ as
 \begin{equation}\label{eq:StrainRate}
-\dot{\varepsilon}_{ij}(p) = \frac{1}{2} \frac{\partial}{\partial t} \left(
+\dot\varepsilon_{ij}(p) = \frac{1}{2} \frac{\partial}{\partial t} \left(
                                         \frac{\partial \hat{u}_i(p)}{\partial x_j} +  
@@
 \begin{equation}\label{eq:StrainMean}
-\mu_{\dot{\varepsilon}_{ij}}(p) = \frac{1}{2}\frac{\partial}{\partial t}\left(
+\mu_{\dot\varepsilon_{ij}}(p) = \frac{1}{2}\frac{\partial}{\partial t}\left(
                                   \frac{\partial \mu_{\hat{u}_i}(p)}{\partial x_j} + 
@@
 \begin{equation}\label{eq:StrainCov}
-C_{\dot{\varepsilon}_{ij}}(p,p') = \frac{1}{4} \frac{\partial^2}{\partial t \, \partial t'}\left(
+C_{\dot\varepsilon_{ij}}(p,p') = \frac{1}{4} \frac{\partial^2}{\partial t \, \partial t'}\left(
                                    \frac{\partial^2 C_{\hat{u}_i}(p,p')}{\partial x_j \, \partial x'_j} + 
@@
 Our motivation for estimating transient strain rates is, in part, to detect geophysical phenomena.
-We propose using a signal-to-noise ratio, SNR, based on $\dot{\du{\varepsilon}}$ for our detection criterion.
-The Frobenius norm of $\dot{\du{\varepsilon}}$, $||\dot{\du{\varepsilon}}||_\mathrm{F}$, which is also referred to as the second invariant of strain rate in the geodetic literature, is often used as a metric for the strain rate ``magnitude''.
-Noting that $||\dot{\du{\varepsilon}}||_\mathrm{F}$ is a random variable, we choose SNR to be the ratio of the estimated mean and standard deviation of $||\dot{\du{\varepsilon}}||_\mathrm{F}$.
+We propose using a signal-to-noise ratio, SNR, based on $\dot\varepsilon$ for our detection criterion.
+The Frobenius norm of $\dot\varepsilon$, $||\dot\varepsilon||_\mathrm{F}$, which is sometimes referred to as the second invariant of strain rate in the geodetic literature, is often used as a metric for the strain rate ``magnitude''.
+Noting that $||\dot\varepsilon||_\mathrm{F}$ is a random variable, SNR can be taken as the ratio of the estimated mean and standard deviation of $||\dot\varepsilon||_\mathrm{F}$.
 Using nonlinear uncertainty propagation, we find SNR to be
 \begin{equation}\label{eq:SNR}
-\mathrm{SNR}(p) = \frac{\mu_{\dot{\varepsilon}_\mathrm{nn}}(p)^2 +
-                        \mu_{\dot{\varepsilon}_\mathrm{ee}}(p)^2 +
-                        2\mu_{\dot{\varepsilon}_\mathrm{en}}(p)^2}
-                       {\big(C_{\dot{\varepsilon}_\mathrm{nn}}(p,p)\mu_{\dot{\varepsilon}_\mathrm{nn}}(p)^2 + 
-                              C_{\dot{\varepsilon}_\mathrm{ee}}(p,p)\mu_{\dot{\varepsilon}_\mathrm{ee}}(p)^2 + 
-                              4C_{\dot{\varepsilon}_\mathrm{en}}(p,p)\mu_{\dot{\varepsilon}_\mathrm{en}}(p)^2
+\mathrm{SNR}(p) = \frac{\mu_{\dot\varepsilon_\mathrm{nn}}(p)^2 +
+                        \mu_{\dot\varepsilon_\mathrm{ee}}(p)^2 +
+                        2\mu_{\dot\varepsilon_\mathrm{en}}(p)^2}
+                       {\big(C_{\dot\varepsilon_\mathrm{nn}}(p,p)\mu_{\dot\varepsilon_\mathrm{nn}}(p)^2 + 
+                              C_{\dot\varepsilon_\mathrm{ee}}(p,p)\mu_{\dot\varepsilon_\mathrm{ee}}(p)^2 + 
+                              4C_{\dot\varepsilon_\mathrm{en}}(p,p)\mu_{\dot\varepsilon_\mathrm{en}}(p)^2
                         \big)^{\frac{1}{2}}}
@@
 In Cascadia, SSEs can be detected by monitoring for associated seismic tremor \citep{Rogers2003}, which is actively being done by the Pacific Northwest Seismic Network (PNSN) \citep{Wech2010}.
-We verify that transient strain rates estimated with the method described above are indeed identifying strain from SSEs by comparing $\dot{\du{\varepsilon}}$ to the tremor records.
+We verify that transient strain rates estimated with the method described above are indeed identifying strain from SSEs by comparing $\dot\varepsilon$ to the tremor records.
 We envision that GNSS derived transient strain rates could be useful for detecting SSEs at subduction zones where there is no associated tremor \citep{Schwartz2007}.
@@
 
-\subsection{Transient displacement model}\label{sec:SignalModel}
+\subsection{Prior model}\label{sec:SignalModel}
 We next establish our prior model for transient displacements.

### 2017-06-22 InitialSubmission/hines_hetland_2017_gji.tex
added animation

@@
 A posterior estimate of transient strain rates is obtained by differentiating the posterior displacements.
-One limitation with GPR is that it is not robust against outliers, and so we introduce a pre-processing method for detecting and removing outliers.
+One limitation with GPR is that it is not robust against outliers, so we introduce a pre-processing method for detecting and removing outliers in GNSS data.
 As a demonstration, we use GPR to detect transient strain resulting from slow slip events in Cascadia.
@@
 
-We now consider the column vector of $n$ GNSS observations make at $m$ stations, $\mitbf{d}_*$.
+We now consider the column vector of $n$ GNSS observations made at $m$ stations, $\mitbf{d}_*$.
 Let $\mitbf{P}$ be the set of $(\vec{x}_i, t_j)$ pairs describing where and when each of the GNSS observations have been made.
@@
 We are interested in identifying transient strain resulting from SSEs \citep[e.g.,][]{Dragert2001}.
-Our results illuminate where elastic strain energy, which can eventually be released seismically, is accumulating during SSEs.
+Our results illuminate where elastic strain energy is accumulating during SSEs, which can eventually be released seismically.
 We demonstrate that transient strain rates estimated with the method described above are an effective tool for detecting SSEs and potentially other geophysical signal in GNSS data.
@@
 
-We use publicly available continuous GNSS data from the University Navstar Consortium (UNAVCO) \citep{Herring2016}, which can be found at www.unavco.org.
+We use publicly available continuous GNSS data from UNAVCO \citep{Herring2016}, which can be found at www.unavco.org.
 We limit the dataset to the stations and time ranges which are pertinent to the seven most recent SSEs in the Puget Sound region.
@@
 The corresponding estimates of $\alpha$ tend to cluster around 0 yr$^{-1}$, suggesting Brownian motion.
-We also estimate hyperparameters for the vertical component of displacements, even though verticals are not used for estimating strain.
-Our hope is that the vertical deformation gradients could reveal some geophysical signal.
+We also estimate hyperparameters for the vertical component of displacements, under the hope that vertical deformation gradients could reveal some geophysical signal.
 The amplitude of FOGM noise for the vertical component is relatively large with a median value of 13.5 mm/yr$^{0.5}$.
@@
 We chose to make each data subsets four months long because it is long enough to encompass a SSE in Cascadia, while it is short enough to still be computationally tractable.
-However, four months is too short to resolve the sinusoids in $\mitbf{d}$, and they are omitted from $\mitbf{d}$ for the REML analysis.
+However, four months is too short to resolve the sinusoids in $\mitbf{d}$, and they are omitted from $\mitbf{d}$ in this REML analysis for Cascadia SSEs.
 The estimated hyperparameters for $u$ are summarized in Table 1.
@@
 Having established a noise model and a prior for transient displacements, we use the cleaned GNSS dataset to calculate transient strain rates in the Puget Sound region.
-We calculate transient strain rates for each day from January 1, 2010 to March 15, 2017.
+We calculate transient strain rates for each day from January 1, 2010 to May 15, 2017.
 The strain rates are estimates at a grid of points spanning the study area.
 In Figure \ref{fig:StrainMap} we show the transient strain rates on January 1, 2016, which coincides with the height of an SSE.
-We have included a video showing the map view of strain rates over time as supplementary material.
+We have included a video showing the map view of strain rates through time as supplementary material.
 The strain rates shown in Figure \ref{fig:StrainMap} are generally similar to the strain rates for the other six SSEs considered in this study.
@@
 There is potential for a more thorough analysis of the spatio-temporal noise in GNSS data, $\eta$, than what was performed in Section \ref{sec:NoiseModel}.
-In this study, we did not explored the spatial covariance of $\eta$.
+We did not explore the spatial covariance of $\eta$.
 Spatially correlated common mode noise, which results from factors such as reference frame error, is a non-negligible component of GNSS data.
@@
 This material is based upon work supported by the National Science Foundation under grant EAR 1245263.
-This material is based on EarthScope Plate Boundary Observatory data services provided by UNAVCO through the GAGE Facility with support from the National Science Foundation (NSF) and National Aeronautics and Space Administration (NASA) under NSF Cooperative Agreement EAR-1261833.
+The EarthScope Plate Boundary Observatory data is provided by UNAVCO through the GAGE Facility with support from the National Science Foundation (NSF) and National Aeronautics and Space Administration (NASA) under NSF Cooperative Agreement EAR-1261833.
+An implementation of the method described in this paper is named Python-based Geodetic Network Strain software (PyGeoNS).
+PyGeoNS is distributed under the MIT License and can be found at www.github.com/treverhines/PyGeoNS.
 

### 2017-06-26 InitialSubmission/hines_hetland_2017_gji.tex
did more

@@
 As a demonstration, we use GPR to detect transient strain resulting from slow slip events in Cascadia.
-Maximum likelihood methods are used to constrain a prior model for transient surface displacements in this region.
+Maximum likelihood methods are used to constrain a prior model for transient displacements in this region.
 The temporal covariance of our prior model is described by a compact Wendland covariance function, which significantly reduces the computational burden that can be associated with GPR.
@@
 Consequently, secular crustal strain rates estimated from GNSS data have been used to constrain seismic hazard models such as UCERF3 \citep{Field2014}.
-Transient crustal strain, which is caused by geophysical phenomena such as slow slip events or postseismic deformation, is also relevant for assessing seismic hazard.
+Transient crustal strain, which is caused by geophysical phenomena such as slow slip events (SSEs) or postseismic deformation, is also relevant for assessing seismic hazard.
 While transient strain itself is not damaging, there is a risk that it can trigger a major earthquake \citep{Roeloffs2006,Freed2001}.
 Dense networks of continuous GNSS stations, such as the Plate Boundary Observatory (PBO), make it possible to identify transient strain with high fidelity.
-An active area of research is developing and improving upon methods for deriving strain and, specifically transient strain, from GNSS displacement observations.
+An active area of research is developing and improving upon methods for deriving transient strain from GNSS data.
 
 Most methods for estimating strain rates from GNSS data assume some parametric form of the deformation signal.
-The simplest method for estimating secular strain rates assumes that GNSS derived velocities can be described with a first order polynomial (i.e., having constant deformation gradients) over some subnetwork of the GNSS stations \citep[e.g.,][]{Feigl1990,Murray2000}.
+The simplest method for estimating secular strain rates assumes that GNSS derived velocities can be described with a first-order polynomial (i.e., having constant deformation gradients) over some subnetwork of the GNSS stations \citep[e.g.,][]{Feigl1990,Murray2000}.
 The components of the strain rate tensor  for each subnetwork are then determined through a least squares fit to the observations.
@@
 On the other hand, if there are too many degrees of freedom, then there will not be any coherent features in the estimated strain rates.
-The methods described by \citet{Beavan2001}, \citet{Tape2009}, and \citet{Sandwell2016} also require the user to specify penalty parameters that control a trade-off between bias and variance in the solution.
+The methods described by \citet{Beavan2001} and \citet{Tape2009} also require the user to specify penalty parameters that control a similar trade-off between bias and variance in the solution.
 One could parameterize deformation with a physically motivated model of interseismic deformation \citep[e.g.,][]{Meade2005,McCaffrey2007}.
 In such models the lithospheric rheology and fault geometries are assumed to be known.
-Any errors in the assumed physical model could similarly result in biased strain estimates and underestimated formal uncertainties.
+Any errors in the assumed physical model could result in biased strain estimates and underestimated formal uncertainties.
 
@@
 In recent years the Southern California Earthquake Center (SCEC) community has shown interest in developing methods for detecting transient strain.
-SCEC supported a transient detection exercise \citep{Lohman2013}, where several research groups tested their methods for detecting transient geophysical signals with a synthetic GNSS dataset.
+SCEC supported a transient detection exercise \citep{Lohman2013}, where several research groups tested their methods for detecting transient geophysical signal with a synthetic GNSS dataset.
 Among the methods tested were the Network Strain Filter (NSF) \citep{Ohtani2010} and the Network Inversion Filter (NIF) \citep{Segall1997}.
@@
 Since GNSS stations are irregularly spaced and observation times may differ between stations, GPR is an ideal tool for synthesizing GNSS data into a spatially and temporally continuous representation of surface deformation.
-GPR is closely related to kriging \citep{Cressie1992} and least squares collocation \citep{Kato1998}, and it has recently gained popularity in the machine learning community.
+GPR is closely related to kriging \citep{Cressie1992} and least squares collocation \citep{Moritz1978}.
+The latter has been used by \citet{Kato1998} and \citet{El-Fiky1999} to estimate secular strain rates from GNSS data.
 GPR is Bayesian in that we describe our prior understanding of the geophysical signal with a Gaussian process.
-A Gaussian process is a normally distributed, continuous, stochastic process that is fully defined in terms of a mean function and a positive-definite covariance function.
+A Gaussian process is a normally distributed stochastic process that is fully defined in terms of a mean function and a positive-definite covariance function.
 For example, Brownian motion, $B(t)$, is a well known Gaussian process in $\mathbb{R}^1$ which has zero mean and covariance function $\mathrm{cov}(B(t),B(t')) = \min(t,t')$, where $t,t' \ge 0$.
@@
 The posterior transient strain is also a Gaussian process, and we can use its distribution to confidently discern geophysical signal from noise.
-Here we use GPR to infer transient strain resulting from slow slip events (SSEs) in Cascadia.
-Our results demonstrate that GPR can be an effective tool for detecting SSEs and revealing the spatio-temporal evolution of strain during SSEs.
+We use GPR to infer transient strain resulting from SSEs in Cascadia.
+Our results demonstrate that GPR is an effective tool for detecting SSEs and revealing how strain evolves during SSEs.
 
@@
 where $C_{u_i}(p,p')$ is a covariance function indicating how we expect $u_i(p)$ to covary with $u_i(p')$.
-For simplicity, we treat each component of displacement independently and ignore any potential covariance.
+For simplicity, we treat each component of displacements independently and ignore any potential covariance.
 Hence, we drop the component subscripts with the understanding that the same analysis is being repeated to estimate the easting and northing components of $\vec{u}$.
@@
 \end{equation}  
-The appropriate choice for $X$ and $T$ may vary depending on the geophysical signal we are trying to describe (e.g.
-postseismic deformation or deformation from slow slip events), and we discuss this matter in Section \ref{sec:SignalModel}.
+The appropriate choice for $X$ and $T$ may vary depending on the geophysical signal we are trying to describe, and we discuss this matter in Section \ref{sec:SignalModel}.
 
-We constrain $u$ with GNSS data, which records $u$ as well as other physical and non-physical processes which we are not interested in.
+We constrain $u$ with GNSS data, which records $u$ as well as other physical and non-physical processes that we are not interested in.
 We describe GNSS observations at position $\vec{x}_i$ and time $t_j$ as a realization of the random variable
@@
 where the subscripts ``n'' and ``e'' denote north and east, respectively.
-For simplicity, we have ignored covariances between the strain rate components in eq.
+For simplicity, we have ignored covariances between the transient strain rate components in eq.
 (\ref{eq:SNR}), even though they are non-zero.
@@
 In our formulation for estimating transient strain rates, we have assumed that noise in the data vector is normally distributed.
-This is not an appropriate assumption for GNSS data which are prone to more outliers than would be predicted for normally distributed noise.
+This is not an appropriate assumption for GNSS data, which often have more outliers than would be predicted for normally distributed noise.
 It follows that proposed methods for analyzing GNSS data should be robust against outliers \citep[e.g.,][]{Blewitt2016}.
@@
 (\ref{eq:TransientCovariance}), and the choice for $X$ and $T$ does not need to be the same as that used in Section \ref{sec:Method}.
-Since outliers are determined based on how well a spatially and temporally dependent model fits the data, we are able to identify anomalous observations which may not be immediately apparent based on inspection of individual time series.
+Since outliers are determined based on how well a spatially and temporally dependent model fits the data, we are able to identify anomalous observations which may not be immediately apparent from inspecting individual time series.
 
@@
 \section{Application to Cascadia Slow Slip Events}\label{sec:Cascadia}
-We demonstrate our method for estimating transient strain rates with GNSS data from Cascadia.
-We are interested in identifying transient strain resulting from SSEs \citep[e.g.,][]{Dragert2001}.
-Our results illuminate where elastic strain energy is accumulating during SSEs, which can eventually be released seismically.
-We demonstrate that transient strain rates estimated with the method described above are an effective tool for detecting SSEs and potentially other geophysical signal in GNSS data.
+We use our method to estimate transient strain rates in Cascadia, and we are specifically interested in identifying strain resulting from SSEs \citep[e.g.,][]{Dragert2001}.
 In Cascadia, SSEs can be detected by monitoring for associated seismic tremor \citep{Rogers2003}, which is actively being done by the Pacific Northwest Seismic Network (PNSN) \citep{Wech2010}.
-We verify that transient strain rates estimated with the method described above are indeed identifying strain from SSEs by comparing $\dot\varepsilon$ to the tremor records.
-We envision that GNSS derived transient strain rates could be useful for detecting SSEs at subduction zones where there is no associated tremor \citep{Schwartz2007}.
+We can thus compare the tremor records to the transient strain rates estimated with GPR to verify that we are indeed identifying strain from SSEs.
 
@@
 The red dots show the positions of GNSS stations mentioned in this paper.
-The blue dot indicates the location of the strain rates shown in Figure \ref{fig:StrainTs} and \ref{fig:StrainMag}.
+The blue dot indicates the location of the transient strain rates shown in Figure \ref{fig:StrainTs} and the signal-to-noise ratio shown in Figure \ref{fig:StrainMag}.
 The blue dashed circle demarcates the spatial extent of the tremors shown in Figure \ref{fig:StrainMag}.}
@@
 
-We constrain the hyperparameters for $\eta$, $\alpha$ and $\beta$, with a set of 38 continuous GNSS stations in the Pacific Northwest which are east of 121$^\circ$W.
-These stations are sufficiently far from the subduction zone that they are unlikely to contain transient signal associated with the Cascadia SSEs.
+We constrain the hyperparameters for $\eta$, $\alpha$ and $\beta$, with a set of 38 continuous GNSS stations in Cascadia that are east of 121$^\circ$W.
+These stations are sufficiently far from the subduction zone that they are unlikely to contain transient signal associated with SSEs.
 We clean the data for these stations by removing jumps at times of equipment changes, and we remove outliers that have been detected with the algorithm described in Section \ref{sec:Outlier}.
@@
                            \left| \mitbf{\Sigma}(\mitbf{\theta}) \right| 
-                           \left| \mitbf{G}^T\mitbf{\Sigma}\mitbf{G} \right|}\right)^{\frac{1}{2}} 
+                           \left| \mitbf{G}^T\mitbf{\Sigma}(\mitbf{\theta})^{-1}\mitbf{G} \right|}\right)^{\frac{1}{2}} 
                            e^{-\tfrac{1}{2}\mitbf{d}_*^T\mitbf{K}(\mitbf{\theta})\mitbf{d}_*},
@@
 The amplitude of FOGM noise, $\beta$, for the easting and northing components is notable low and are clustered around 0.5 mm/yr$^{0.5}$.
-The corresponding estimates of $\alpha$ tend to cluster around 0 yr$^{-1}$, suggesting Brownian motion.
+The corresponding estimates of $\alpha$ tend to cluster around 0 yr$^{-1}$, suggesting that noise can be described with Brownian motion.
 We also estimate hyperparameters for the vertical component of displacements, under the hope that vertical deformation gradients could reveal some geophysical signal.
@@
 \ref{eq:FOGM}).
-Hyperparameters are estimated with the REML method for 38 stations in the Pacific Northwest that are east of $121^\circ$W.
-``IQR'' is the inter-quartile range.}
+Hyperparameters are estimated with the REML method for 38 stations in Cascadia that are east of $121^\circ$W.
+``IQR'' is the interquartile range.}
 \label{fig:NoiseParams}
@@
 \begin{equation}\label{eq:TimeSE}
-T(t,t') = \phi^2\exp\left(\frac{-|t - t'|^2}{\tau^2}\right).
+T(t,t') = \phi^2\exp\left(\frac{-|t - t'|^2}{2\tau^2}\right).
 \end{equation}
@@
 For example, if we are trying to identify postseismic deformation then $t$ should be zero at the time of the earthquake.
-However, if we are interesting in detecting transient events, where there is no known start time, then IBM may not be an appropriate prior and an isotropic Gaussian process should be preferred.
+However, if we are interesting in detecting transient events, where there is no known start time, then IBM may not be an appropriate prior, and an isotropic Gaussian process should be preferred.
 In the following analysis, we make the quite arbitrary choice that $t$ is zero on the first epoch of $\mitbf{d}_*$.
@@
 In our analysis, we exploit this sparsity with the CHOLMOD software package \citep{Chen2008}.
-We elaborate on the computational advantages of the Wendland functions in Section \ref{sec:Discussion}.
 Wendland functions are positive definite in $\mathbb{R}^d$, and they describes an isotropic Gaussian process with realizations that can be differentiated $k$ times.
@@
 We use $d=1$ since we are describing the temporal covariance of $u$.
-We use $k=1$ because we need $u$ to be at least once differentiable in time to determine strain rates.
+We use $k=2$, giving samples of $u$ continuous velocities and accelerations.
 The corresponding Wendland covariance function is
 \begin{equation}\label{eq:Wendland}
-T(t,t') = \phi^2\left(1 - \frac{|t - t'|}{\tau}\right)^3_+ \left(\frac{3|t - t'|}{\tau} + 1\right), 
+T(t,t') = \phi^2\left(1 - \frac{|t - t'|}{\tau}\right)^5_+ \left(\frac{8|t - t'|^2}{\tau^2} + \frac{5|t - t'|}{\tau} + 1\right), 
 \end{equation}
@@
 The outlier detection algorithm is particularly effective at removing outliers for stations at high elevation (Figure \ref{fig:Outliers}), which can be adversely affected by ice or snow during the winter \citep{Lisowski2008}.
-After cleaning the dataset, we divide it into seven subsets which are four months long and each centered on the time of an SSE.
+After cleaning the dataset, we divide it into seven subsets which are four months long and each centered on the time of a SSE.
 The times of the seven SSEs are determined with tremor records from \cite{Wech2010}.
 We use the REML method to find the optimal hyperparameters for $T$ and $X$ for each subset of data.
-We chose to make each data subsets four months long because it is long enough to encompass a SSE in Cascadia, while it is short enough to still be computationally tractable.
+We choose to make each data subsets four months long because it is long enough to encompass a SSE in Cascadia, while it is short enough to still be computationally tractable.
 However, four months is too short to resolve the sinusoids in $\mitbf{d}$, and they are omitted from $\mitbf{d}$ in this REML analysis for Cascadia SSEs.
 The estimated hyperparameters for $u$ are summarized in Table 1.
-Based on the interquartile ranges (IQR), the estimated hyperparameters for the SE and Wendland covariance functions do not vary significantly between SSEs.
+Based on the interquartile ranges, the estimated hyperparameters for the SE and Wendland covariance functions do not vary significantly between SSEs.
 This suggests that the median of estimated hyperparameters should be an appropriate prior model for all Cascadia SSEs.
-For the IBM model, there are several anomalously large values for $\ell$ and $\phi$, which results in large IQRs.
+For the IBM model, there are several anomalously large values for $\ell$ and $\phi$, which results in large interquartile ranges.
 
@@
 \includegraphics{figures/outliers/outliers.pdf}
-\caption{Detrended easting component of GNSS displacement data for station SC03, which is located on Mount Olympus in Washington.
-The orange markers indicate outliers which were automatically detected using the algorithm from Section \ref{sec:Outlier}.
+\caption{Detrended easting component of displacements at station SC03, which is located on Mount Olympus in Washington.
+The orange markers indicate outliers that were automatically detected using the algorithm from Section \ref{sec:Outlier}.
 The error bars show one standard deviation uncertainties.
@@
 The ``diff $\log$(REML)'' column compares the log REML likelihood to the log REML likelihood when using the SE covariance function for $T$.
-Positive values indicate that observations are more consistent with the SE covariance function.}
+Negative values indicate that observations are more consistent with the SE covariance function.}
 \begin{tabular} {l l l l l l}
@@
 Next we identify which covariance function for $T$ best describes the SSEs.
-One approach is to compare the REML likelihood for each covariance function, similar to the analysis in \citet{Langbein2004}.
+One approach is to compare the REML likelihoods for each covariance function, similar to the analysis in \citet{Langbein2004}.
 In Table 1, we summarize how the log REML likelihoods for the Wendland and IBM covariance functions compare to the SE covariance function.
@@
 To verify that the estimated transient strain rates are accurately identifying geophysical signal, we compare the signal-to-noise ratio from eq.
-(\ref{eq:SNR}), SNR, to the frequency of seismic tremor (Figure \ref{fig:StrainMag}).
-An SNR greater than ${\sim}3$ can be interpreted as a detected geophysical signal.
+(\ref{eq:SNR}) to the frequency of seismic tremor (Figure \ref{fig:StrainMag}).
+A signal-to-noise ratio greater than ${\sim}3$ can be interpreted as a detected geophysical signal.
 For each detected event there is a corresponding peak in seismic tremor.
-We are also able to clearly identify transient strain associated with a more subtle inter-SSE event in August 2014.
-In between SSEs, the SNR is consistently between 0 and 2, indicating that no anomalous non-SSE events are being detected, at least at this location.
+We are also able to clearly identify transient strain associated with a more subtle SSE in August 2014.
+In between SSEs, the signal-to-noise ratio is consistently between 0 and 2, indicating that no anomalous non-SSE events are being detected, at least at this location.
 
@@
 Another potential application would be to use GNSS derived transient strain rates to develop noise models for borehole strain meters (BSMs).
-The Plate Boundary Observatory maintains 43 BSMs in the Pacific Northwest, and it has been demonstrated that BSMs are able to record transient geophysical events such as SSEs \citep[e.g.,][]{Dragert2011}.
+The Plate Boundary Observatory maintains 43 BSMs in Cascadia, and it has been demonstrated that BSMs are able to record transient geophysical events such as SSEs \citep[e.g.,][]{Dragert2011}.
 However, noise in BSM data is not well understood, which complicates the use of BSM data for detecting geophysical events.
@@
 We can also improve upon the seasonal model used in this study, which consists of four spatially uncorrelated sinusoids for each station.
-We did not explore the spatial covariance of seasonal deformation or the temporal ``roughness'' (i.e., the number of sinusoids needed to describe the observations).
+We did not explore the spatial covariance of seasonal deformation or the temporal roughness (i.e., the number of sinusoids needed to describe the observations).
 The periodic Gaussian process \citep{Mackay1998} is an alternative model for seasonal deformation and is well suited for exploring the roughness of seasonal deformation.
@@
 \begin{equation}\label{eq:Periodic}
-T(t,t') = \phi \exp\left(\frac{-\sin(\pi|t - t'|)^2}{2\tau^2}\right).
+T(t,t') = \phi^2 \exp\left(\frac{-\sin(\pi|t - t'|)^2}{2\tau^2}\right).
 \end{equation}

### 2017-06-26 InitialSubmission/hines_hetland_2017_gji.tex
added more

@@
 Transient crustal strain, which is caused by geophysical phenomena such as slow slip events (SSEs) or postseismic deformation, is also relevant for assessing seismic hazard.
-While transient strain itself is not damaging, there is a risk that it can trigger a major earthquake \citep{Roeloffs2006,Freed2001}.
+While transient strain itself is not damaging, there is a risk that it can trigger major earthquakes \citep{Roeloffs2006,Freed2001}.
 Dense networks of continuous GNSS stations, such as the Plate Boundary Observatory (PBO), make it possible to identify transient strain with high fidelity.
-An active area of research is developing and improving upon methods for deriving transient strain from GNSS data.
+Developing and improving upon methods for deriving secular and transient strain from GNSS data is an active area of research.
 

### 2017-07-14 RevisedSubmission/gji_extraguide.tex
started making edits

%%gji_extra_guide.tex

\documentclass[extra,mreferee]{gji}
%\usepackage{timet}

\title[\texttt{gji\_extra.sty} ]  
  {\texttt{gji\_extra} - additional facilities for use with GJI
manuscripts}

\author[]  {}

\newcommand{\btx}{\textsc{BibTeX}}
\begin{document}

\maketitle

\section{Introduction}
The following set of features are derived from other packages with
slight modifications to make them compatible with the \verb"gji"
documentclass. They can be accessed by including the \verb"extra"
option to the \verb"gji" document class e.g.
\begin{verbatim}
  \documentclass[extra,mreferee]{gji}
\end{verbatim}

\section{Reserved Space and Boxed Figures }

It is possible to reserve space for a figure (or to draw a box of
a specified size) using the \verb!\figbox! command.

The \verb!\figbox! command takes 3 arguments: the
horizontal and vertical sizes of the space that is to be left blank, as well
as the contents of the box.

For example,
\begin{verbatim}
\begin{figure}
\figbox{6cm}{5cm}{Paste orbits here}
\caption{The orbits of some planets}
\end{figure}
\end{verbatim}
\begin{figure*}
\figbox{6cm}{5cm}{Paste orbits here}
\caption{Illustrating the use of the figbox command for a
display of the orbits of some planets}
\end{figure*}
makes a box 4~cm wide and 6~cm high with the caption text below it.
The text `Paste orbits here' is simply a note to the author that is
printed centered in the framed box. This third argument could be left empty,
or could contain commands for importing a computer plot.

There is also a starred version \verb!\figbox*! that behaves exactly the same
as \verb!\figbox! except that no frame is drawn around the figure. This is
most useful for real figures, whereas the unstarred command is more
appropriate for reserved space for glued-in figures.

It is possible to have \verb!\figbox! and \verb!\figbox*! scale automatically
to the size of its contents, something that is also appropriate when the
contents are an imported graphics file. In this case, leave both the
dimensions empty. As an example, if the orbit plot exists as an encapsulated
PostScript file named \texttt{orbit.eps}, then it could be included directly
with the commands
\begin{verbatim}
\begin{figure}
\figbox*{}{}{%
 \includegraphics[width=4cm]{orbit.eps}}
\caption{The orbits of some planets}
\end{figure}
\end{verbatim}
For this to work, you must have loaded the \texttt{graphicx} package with
\verb!\usepackage! at the beginning of the file, and you must have a
PostScript driver for the output. (There are other packages with
different syntaxes for importing graphics; use the one that you are
most familiar with.)

\section{Alternative texts for one or two columns}
Mathematical formulas often have to be fiddled to fit them into the
narrow confines of a single column in two-column format, whereas they
will fit with no problem in the wide columns of the manuscript mode.
This often results in the author having to massage his formulas when he
changes between manuscript and camera-ready options, and then back again
when he want to print the manuscript once more.

The special command \verb!\iftwocol! allows both versions of the text to
be included in the one document, for automatic selection depending on
whether two-column mode is active or not. Its syntax is\label{iftwocol}
\begin{quote}
\normalsize
\verb!\iftwocol{!{\em yes\/}\verb!}{!{\em no\/}\verb!}!
\end{quote}
where {\em yes\/} is the text that is inserted if two-columns are in
effect, and {\em no\/} the text that is otherwise taken.

This command may be used in other situations, but the main application 
is in the case of mathematics.

\section{Literature citations}
\textit{Geophysical Journal International} uses the
author-year system of literature citation, something that is not
supported by standard \LaTeX. The \verb"gji" documentclass
provides partial support but more comprehensive features are 
available using features from the \verb"egs" package and the
\verb"natbib" module developed by P.W.Daly.

Since there are two ways of making a citation in the author-year system,
either as Jones et al.\ (1990) or as (Jones et al., 1990), there
are two variations of the original \verb!\cite! command.
Suppose the key for the above reference is \texttt{jones90}, then use
\begin{flushleft}
\verb!\citet{jones90}! for Jones et al.\ (1990)\\
\verb!\citep{jones90}! for (Jones et al., 1990)\\
\verb!\citep[p.~32]{jones90}! for (Jones et al., 1990, p.~32)\\
\verb!\citep[e.g.,][]{jones90}! for (e.g., Jones et al., 1990)\\
\verb!\citep[e.g.,][p.~32]{jones90}! for (e.g., Jones et al., 1990, p.~32).

\end{flushleft}
Note that the use of the optional arguments to add notes within the
brackets of the citation:
a single note behaves as in standard \LaTeX, as a note
\emph{after} the citation; however, with two notes (non-standard), the
first goes \emph{before}, the second \emph{after} it.

Two other citation commands are available:
\begin{flushleft}
\verb!\citeauthor{jones90}! prints Jones et al.\\
\verb!\citeyear{jones90}! prints 1990.
\end{flushleft}

For the above examples to function properly, either the \verb"gji"
bibliography style must be used with \btx, or the
\texttt{thebibliography} environment must be formatted accordingly.

\begin{flushleft}
With \btx\\[1ex]
\verb!\bibliographystyle{!gji\verb!}!\\
\verb!\bibliography{!{\em bib\_file\_name\/}\verb!}!\\[1ex]
Without \btx\\[0.5ex]
\begin{verbatim}
\begin{thebibliography}{}
\bibitem[Jones et al.(1990)]{jones90}
  Jones, J. K., Thomas, P. R. 
  \& Peters, R. F., 1990.
  The best results of fitting curves. 
  \textit{J. Math. Dev.}, \textbf{12}, 1245--1261.
\end{thebibliography}
\end{verbatim}
\end{flushleft}

As an example the References at the end of these notes are
the same set as used in \verb"gjilguid2e.tex"
but now cast in the appropriate bibliography environment.

\section{Balancing columns}
It is possible to make the two columns on the last page to be nearly the
same height. It is only necessary to give the command \verb!\balance!
somewhere within the text of the first column on the last page. Issuing
\verb!\balance! too soon can yield strange results on earlier pages;
calling it too late has no effect.

It is best to print out the last page without \verb!\balance! when the
work is nearly completed. If this page has more than one column of text,
then find some appropriate place within the {\em first\/} column,
preferably between paragraphs, and add \verb!\balance! to the text there.
If there is less than one column of text on the last page, then
\verb!\balance! may be inserted anywhere within the text of that column.

\section{planotables}
The {\tt planotable} environment that facilitates the 
formatting of lengthy tables. Tables longer than 
one page will require the use of {\tt planotable}.

It is possible to create fairly complex tables with 
arbitrary spacing, straddle heads, rules, etc. in \LaTeX.
Authors who need to create complicated tables should 
consult the \LaTeX\ manual [{\it Lamport}, 1985] for 
details.  Most of \LaTeX's {\tt tabular} capabilities 
are applicable to {\tt planotables} as well.

Planotables automatically produce table lines and
vertical table spacing.
In addition, planotables have several capabilities that 
facilitate the formatting of tables.  For instance, 
it is possible to break long planotables across pages, 
and, where tabular tables print at an automatic width,
authors may choose a specific planotable width.

Type \verb"\begin{planotable}{COLS}", 
where {\tt COLS}
sets the justification for each column.
Choose one letter (``l,'' ``c,'' or ``r'') for each 
column, indicating left, center, or right justification.

To set a planotable to a specific width, type 
\verb"\tablewidth{}", with the desired table width between 
the curly braces, after the \verb"\begin{planotable}" 
command.  For instance, to create a single-column table 
type \verb"\tablewidth{20pc}", where ``20pc'' represents 20 picas.

Type planotable titles within the curly braces of
\verb"\tablecaption{}" commands.  Capitalize the first
letter of each word (except for prepositions, conjunctions,
and articles that are three letters or shorter).  Do not
allow table caption lines to hyphenate; use a
\verb"\protect\\" command to break lines where necessary.
The \verb"\tablecaption" command automatically generates
the ``{\bf Table \#.}'' information.

Type a \verb"\startdata" command, and then type your table
data.  The startdata command formats column headings, 
engages the tabular formatting, and produces the table 
caption.  Data within a table row are separated by {\tt \&} 
(ampersand) characters.  The end of each row is indicated 
with a \verb"\nl" command.  Extra vertical space can be 
inserted between rows with a \verb"\vspace{}" command 
(type the desired amount of space between the curly
braces).

If a planotable is longer than one page, \LaTeX\ will 
automatically break it across pages.  To force a page 
break in a particular place, type a \verb"\tablebreak" 
command.  This command affects the following line of 
table data (not the line it appears with) as shown 
in the sample.tex document.

If a cell contains no data, type a \verb"\nodata" 
command to create a ``no data'' symbol.

\subsection{Table footnotes and table comments}

If your table contains material requiring footnotes, use a
\verb"\tablemark{TAG}" command to indicate the footnote,
then type the associated information (with a corresponding
{\tt TAG}) within a \verb"\tablenotetext{TAG}{}" command.

Short table comments can be created using a\\
\verb"\tablenotetext{\null}{TEXT}"\\
command, with a 
\verb"\null" argument as the tag.  Longer comments may
be placed within a \verb"\tablecomment{TEXT}" command.
Lists of table references should use the following
format: \verb"\tablecomment{References:  Names of your references.}".
Only one paragraph of material is permitted  at the end of a table,
(excluding superscripted footnotes), so if both references and notes
exist, they should be run in together.

Both \verb"\tablenotetext" and \verb"\tablecomment" 
commands must appear after the column heading information
and before the table data; otherwise, they will affect
indentation of the last table cell.  These commands 
work with planotables and tabular tables.

\begin{acknowledgments}
Much of this material has been derived from the work of Patrick W. Daly
for the European Geophysical Society and his efforts in providing
robust and easy to use features are to be commended.  The American
Geophysical Union is also to be thanked for permission to use the
\verb"planotable"  material
\end{acknowledgments}

\begin{thebibliography}{}  
  \bibitem[Butcher(1992)]{bu}  
    Butcher J. 1992. {\it Copy-editing: The Cambridge  
    Handbook}, 3rd edn, Cambridge Univ. Press, Cambridge.
  \bibitem[The Chicago Manual(1982)]{cm}  
    {\it The Chicago Manual of Style}, Univ.   
    Chicago Press, Chicago, 1982.
  \bibitem[Chao(1985)]{ch} 
    Chao, B. F., 1985. Normal mode study of the Earth's rigid 
     body motions, {\it Geophys. Res. Lett.}, {\bf 12}, 526-529.
  \bibitem[Hinderer(1986)]{hi}  
    Hinderer, J., 1986. Resonance effects of the earth's fluid
    core in earth rotation, in {\it Solved and Unsolved 
    Problems}, pp. 277-296, ed. Cazenave A., Reidel, 
    Dordrecht.
  \bibitem[Kopka \& Daly(1995)]{kd}  
    Kopka H. \& Daly P.W., 1995, \textit{A guide to} \LaTeX2e,
    Addison--Wesley, New York  
  \bibitem[Lamport(1986)]{la}  
    Lamport L., 1986,  \LaTeX: {\it A Document   
    Preparation System}, Addison--Wesley, New York  
  \bibitem[Lindberg (1986)]{bl}  
    Lindberg, C., 1986.  Multiple taper harmonic analysis of 
    terrestrial free oscillations, {\it PhD thesis}, 
    University of California.
  \bibitem[Maupin(1992)]{ma} 
    Maupin, V., 1992. Modelling of laterally trapped surface 
    waves with application to Rayleigh waves in the Hawaiian 
    swell, {\it Geophys. J. Int.}. {\bf 110}, 553-570.     	
  \bibitem[Rutherford \& Hawker(1981)]{rh}
      Rutherford, S. R. \& Hawker, K. E., 1981, 
    Consistent coupled mode theory of sound propagation for a 
    class of non-separable problems, 
   {\it J. acoust. Soc. Am.}, {\bf 71}, 554-564
\end{thebibliography}

\end{document}

### 2017-07-14 RevisedSubmission/hines_hetland_2017_gji.tex
started making edits

--- 2017-06-26 InitialSubmission/hines_hetland_2017_gji.tex
@@
 GPR is Bayesian in that we describe our prior understanding of the geophysical signal with a Gaussian process.
-A Gaussian process is a normally distributed stochastic process that is fully defined in terms of a mean function and a positive-definite covariance function.
+A Gaussian process is a normally distributed stochastic process that is fully defined in terms of a mean function and a positive definite covariance function.
 For example, Brownian motion, $B(t)$, is a well known Gaussian process in $\mathbb{R}^1$ which has zero mean and covariance function $\mathrm{cov}(B(t),B(t')) = \min(t,t')$, where $t,t' \ge 0$.
@@
 where $C_{u_i}(p,p')$ is a covariance function indicating how we expect $u_i(p)$ to covary with $u_i(p')$.
-For simplicity, we treat each component of displacements independently and ignore any potential covariance.
+For simplicity, we treat each component of displacements independently and ignore any potential covariance between them.
 Hence, we drop the component subscripts with the understanding that the same analysis is being repeated to estimate the easting and northing components of $\vec{u}$.
-We assume that $C_u$ can be separated into positive definite spatial and temporal functions as
+We assume that $C_u$ can be separated into spatial and temporal functions as
 \begin{equation}\label{eq:TransientCovariance}
@@
 \end{equation}  
+As long as the functions $X$ and $T$ are positive definite, $C_u$ is guaranteed to also be positive definite and thus a valid covariance function \citep[sec.
+4.2.4]{Rasmussen2006}.
 The appropriate choice for $X$ and $T$ may vary depending on the geophysical signal we are trying to describe, and we discuss this matter in Section \ref{sec:SignalModel}.
@@
 The prior for transient displacements is then conditioned with $\mitbf{d}_*$ to form a posterior estimate of transient displacements, $\hat{u} = u | \mitbf{d}_*$.
-For now, we will assume that appropriate covariance functions and corresponding hyperparameters for $X$, $T$, and $C_\eta$ have already been chosen.
+For now, we will assume that appropriate covariance functions and corresponding hyperparameters for $C_u$ and $C_\eta$ have already been chosen.
 In Section \ref{sec:NoiseModel} and \ref{sec:SignalModel}, we discuss how the covariance functions are chosen for our application to GNSS data from Cascadia.
-If $\kappa$ is kept finite then, following \citet{Rasmussen2006}, we find that $\hat{u}$ is distributed as $\mathcal{N}(\mu_{\hat{u}},C_{\hat{u}})$, where
+If $\kappa$ is kept finite then, from \citet[sec.
+2.2]{Rasmussen2006}, we find that $\hat{u}$ is distributed as $\mathcal{N}(\mu_{\hat{u}},C_{\hat{u}})$, where
 \begin{equation}\label{eq:PosteriorMean}
@@
 (\ref{eq:PosteriorCov}) is not suitable for evaluating this limit.
-We use the partitioned matrix inversion identity \citep[e.g.,][]{Press2007} to rewrite eq.
+We use a partitioned matrix inversion identity \citep[sec.
+2.7.4]{Press2007} to rewrite eq.
 (\ref{eq:PosteriorMean}) and eq.
@@
 \end{equation}
-The transient strain rate components are Gaussian processes with mean functions
+The transient strain rate components are stochastic processes whose distributions can be interpreted as the distributions of samples of $\hat{u}$ that have been differentiated by the operator from eq.
+(\ref{eq:StrainRate}).
+Since differentiation is a linear operation, the transient strain rate components are Gaussian processes.
+From \citet[sec.
+10.2]{Papoulis1991}, we find that the components of $\dot{\varepsilon}$ have mean functions
 \begin{equation}\label{eq:StrainMean}

### 2017-07-14 RevisedSubmission/figures/signal_fit/make-plot.py

@@ -1,0 +1,71 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import pygeons
+from matplotlib import colors
+
+xtick_dates = ['2015-11-01','2015-12-01','2016-01-01','2016-02-01','2016-03-01']
+xticks = [pygeons.mjd.mjd(i,'%Y-%m-%d') for i in xtick_dates]
+xtick_labels = ['2015-11-01','2015-12-01','2016-01-01','2016-02-01','2016-03-01']
+
+c0 = colors.to_rgb('C0')
+c0t = tuple(0.6 + 0.4*np.array(colors.to_rgb('C0')))
+
+fig,ax = plt.subplots(figsize=(7,3.0))
+pygeons.plot.plot._setup_ts_ax([ax])
+ax.set_xticks(xticks)
+ax.set_xticklabels(xtick_labels)
+ax.tick_params(labelsize=10)
+ax.set_xlim((57327.0,57448.0))
+ax.set_ylabel('displacement [mm]')
+#ax.set_ylim((-10.0,15.0))
+ax.grid()
+
+# OBSERVED
+data = pygeons.io.io.dict_from_hdf5('data/data.h5')
+idx = (data['id'] == 'P436').nonzero()[0][0]
+lon = data['longitude'][idx]
+lat = data['latitude'][idx]
+t = data['time']
+u = 1000*data['east'][:,idx]
+us = 1000*data['east_std_dev'][:,idx]
+
+ax.set_title('station P436 (%.1f$^\mathregular{\circ}$W, %.1f$^\mathregular{\circ}$N)' %
+            (-lon,lat),fontsize=10)
+
+ax.errorbar(t,u,us,marker='.',linestyle='None',color=c0,ecolor=c0t,ms=5.0,zorder=1,label='observed')
+
+# SE
+data = pygeons.io.io.dict_from_hdf5('data/se.per.fit.h5')
+idx = (data['id'] == 'P436').nonzero()[0][0]
+lon = data['longitude'][idx]
+lat = data['latitude'][idx]
+t = data['time']
+u = 1000*data['east'][:,idx]
+us = 1000*data['east_std_dev'][:,idx]
+ax.plot(t,u,'-',color='C1',zorder=2,lw=1.5,label='SE')
+ax.fill_between(t,u-us,u+us,edgecolor='none',facecolor='C1',zorder=2,alpha=0.4)
+
+# WEN
[…]

### 2017-07-14 RevisedSubmission/figures/signal_reml/make_plot.py

@@ -1,0 +1,163 @@
+import numpy as np
+from pygeons.io.parser import _get_field
+import os
+import matplotlib.pyplot as plt
+
+def iqr(x):
+  return np.subtract(*np.percentile(x,[75.0,25.0],interpolation='linear'))
+
+msg = 'cov. function & direction & $\\lambda$  & $\\phi$   & $\\theta$  & diff. REML \\\\ \\hline\n'
+#####################################################################
+files = os.listdir('se-se-results')
+sese_like = {'east':[],'north':[],'vertical':[]}
+params0 = {'east':[],'north':[],'vertical':[]}
+params1 = {'east':[],'north':[],'vertical':[]}
+params2 = {'east':[],'north':[],'vertical':[]}
+for f in files:
+  buff = open('se-se-results/%s' % f,'r')
+  content = buff.read()
+  buff.close()
+  content = content[content.find('REML RESULTS'):]
+  content = content[content.find('network :'):]
+  units = _get_field('parameter units',content).strip().split(',')
+  for dir in ['east','north','vertical']:
+    params = _get_field('optimal %s parameters' % dir,content).strip().split(',')
+    if np.isfinite(float(params[0])):
+      params0[dir] += [float(params[0])]
+
+    if np.isfinite(float(params[1])):
+      params1[dir] += [float(params[1])]
+
+    if np.isfinite(float(params[2])):
+      params2[dir] += [float(params[2])]
+
+    likelihood_content = content[content.find('likelihood :'):]
+    sese_like[dir] += [float(_get_field(dir,likelihood_content))]
+    
+east_p0 = params0['east']
+north_p0 = params0['north']
+vertical_p0 = params0['vertical']
+
+east_p1 = params1['east']
+north_p1 = params1['north']
+vertical_p1 = params1['vertical']
+
+east_p2 = params2['east']
+north_p2 = params2['north']
+vertical_p2 = params2['vertical']
+
+msg += 'SE-SE & east   & %.5g $\pm$ %.5g km  & %.5g $\pm$ %.5g mm  & %.5g $\pm$ %.5g yr & 0.0 \\\\\n' % (
+                                                                                      np.median(east_p2),iqr(east_p2),
+                                                                                      np.median(east_p0),iqr(east_p0),
+                                                                                      np.median(east_p1),iqr(east_p1))
[…]

### 2017-07-15 RevisedSubmission/hines_hetland_2017_gji.tex
added more

@@
 For example, Brownian motion, $B(t)$, is a well known Gaussian process in $\mathbb{R}^1$ which has zero mean and covariance function $\mathrm{cov}(B(t),B(t')) = \min(t,t')$, where $t,t' \ge 0$.
-If no prior information is available for the geophysical signal, then maximum likelihood methods can be used to choose a prior Gaussian process that is most consistent with the observations.
+If no prior information is available for the geophysical signal, then maximum likelihood methods can be used to objectively choose a prior Gaussian process that is most consistent with the observations.
 We incorporate GNSS observations with the prior to form a posterior estimate of transient strain.
 The posterior transient strain is also a Gaussian process, and we can use its distribution to confidently discern geophysical signal from noise.
-We use GPR to infer transient strain resulting from SSEs in Cascadia.
-Our results demonstrate that GPR is an effective tool for detecting SSEs and revealing how strain evolves during SSEs.
+We use GPR to infer strain resulting from SSEs in Cascadia, demonstrating that GPR is an effective tool for detecting transient geophysical processes.
 
@@
 We use our method to estimate transient strain rates in Cascadia, and we are specifically interested in identifying strain resulting from SSEs \citep[e.g.,][]{Dragert2001}.
-In Cascadia, SSEs can be detected by monitoring for associated seismic tremor \citep{Rogers2003}, which is actively being done by the Pacific Northwest Seismic Network (PNSN) \citep{Wech2010}.
+In Cascadia, SSEs can be detected by monitoring for associated seismic tremor \citep{Rogers2003}, which is actively being done by the Pacific Northwest Seismic Network \citep{Wech2010}.
 We can thus compare the tremor records to the transient strain rates estimated with GPR to verify that we are indeed identifying strain from SSEs.
 
-We use publicly available continuous GNSS data from UNAVCO \citep{Herring2016}, which can be found at www.unavco.org.
+We use the daily displacement solutions generated by the Geodesy Advancing Geosciences and EarthScope (GAGE) facility for continuous GNSS stations \citep{Herring2016}.
+This data is publicly available and can be found at www.unavco.org.
 We limit the dataset to the stations and time ranges which are pertinent to the seven most recent SSEs in the Puget Sound region.
@@
 (\ref{eq:REML}) is equivalent to choosing the hyperparameters which maximize the probability of drawing $\mitbf{d}_*$ from $\mitbf{d}$.
+We use the REML method over the maximum likelihood method \citep[e.g.,][]{Langbein1997} because the REML method accounts for the improper prior that we assigned to $\mitbf{a}$ \citep{Hines2017}.
 We independently estimate $\mitbf{\theta}$ for each station, and so $\mitbf{d}_*$ consists of displacements for an individual station.
 We are assuming $u(p)=0$ when estimating the noise hyperparameters for this section.
-We use the REML method over the maximum likelihood (ML) method \citep[e.g.,][]{Langbein1997} because the REML method accounts for the improper prior that we assigned to $\mitbf{a}$ \citep{Hines2017}.
 The distribution of inferred $\alpha$ and $\beta$ are shown in Figure \ref{fig:NoiseParams}.
@@
 Our third option for $T$ is the Wendland class of covariance functions \citep{Wendland2005}.
-Wendland covariance functions have compact support and hence their corresponding covariance matrices will be sparse.
+Wendland covariance functions have compact support and hence their corresponding covariance matrices are sparse.
 In our analysis, we exploit this sparsity with the CHOLMOD software package \citep{Chen2008}.
@@
 We consider the data prediction vector to be $\hat{\mitbf{d}} = \left(u(\mitbf{P}) + \mitbf{G}\mitbf{a}\right)|\mitbf{d}_*$.
-It can be shown that $\hat{\mitbf{d}}$ is normally distributed with mean
+Following a similar procedure as in Section \ref{sec:Method}, it can be shown that $\hat{\mitbf{d}}$ is normally distributed with mean
 \begin{equation}\label{eq:DataPredMean}
@@
 We compute $\hat{\mitbf{d}}$ using SE, Wendland, and IBM covariance functions for $T$ and the median hyperparameters from Table 1.
-Figure \ref{fig:Fit} compares the easting component of $\mitbf{d}_*$ to $\hat{\mitbf{d}}$ for the Winter 2015-2016 SSE at Station P436.
-The data prediction vector appears to accurately describe displacements throughout the SSE, regardless of the choice of $T$.
-Since the SSE signal is strongest at Station P436, it can be assumed that $T$ adequately describes the signal elsewhere.
+Figure \ref{fig:Fit} compares the easting component of $\mitbf{d}_*$ to $\hat{\mitbf{d}}$ for the winter 2015-2016 SSE at station P436, which is among the stations that record the strongest signal.
+The data prediction vector reasonable fits the displacements throughout the SSE, regardless of the choice of $T$.
 The prediction for the IBM covariance function contains slightly more high frequency, and perhaps spurious, features.
 The predictions for the Wendland and SE covariance functions are nearly indistinguishable.
+Overall, the predicted displacements are not strongly sensitive to the choice of temporal covariance function.
 In our estimates of transient strain discussed in the next section, we ultimately settle on the Wendland covariance function for $T$ and use the median values from Table 1 for the hyperparameters.
@@
 In Figure \ref{fig:StrainMap} we show the transient strain rates on January 1, 2016, which coincides with the height of an SSE.
-We have included a video showing the map view of strain rates through time as supplementary material.
+We have included an animation showing the map view of strain rates through time as supplementary material.
 The strain rates shown in Figure \ref{fig:StrainMap} are generally similar to the strain rates for the other six SSEs considered in this study.
@@
 We are also able to clearly identify transient strain associated with a more subtle SSE in August 2014.
-In between SSEs, the signal-to-noise ratio is consistently between 0 and 2, indicating that no anomalous non-SSE events are being detected, at least at this location.
+In between peaks in seismic tremor, the signal-to-noise ratio is consistently between 0 and 2, suggesting that all the transient strain detected at this location is associated with SSEs.
 
@@
 
+The results we have presented thus far indicate that we are identifying the strain that we should expect to see.
+There are, however, features in our estimated transient strain rates which we were not expecting.
+For example, there is a brief period of east-west extension on the Olympic peninsula several days prior to some of the SSEs.
+This feature can be seen before the summer 2012 and winter 2015-2016 SSEs in Figure \ref{fig:StrainTs} as well as in the supplementary animation.
+A discussion on the mechanisms causing this deformation is outside the scope of this study.
+
 \section{Discussion}\label{sec:Discussion}
 We have demonstrated that transient strain rates estimated with the method described in Section \ref{sec:Method} can be used to detect SSEs and will be robust to detect other transient geophysical phenomena.
-Another potential application would be to use GNSS derived transient strain rates to develop noise models for borehole strain meters (BSMs).
+Another potential application would be to use GNSS derived transient strain rates to better understand the data from borehole strain meters (BSMs).
 The Plate Boundary Observatory maintains 43 BSMs in Cascadia, and it has been demonstrated that BSMs are able to record transient geophysical events such as SSEs \citep[e.g.,][]{Dragert2011}.
-However, noise in BSM data is not well understood, which complicates the use of BSM data for detecting geophysical events.
+However, there are complications that prevent BSM data from being used quantitatively in geophysical studies.
+One difficulty is that BSM data should be calibrated with a well known strain source, such as diurnal and semidiurnal tides \citep{Hart1996,Roeloffs2010,Hodgkinson2013}.
+Unfortunately, the tidal forces at BSMs which record SSEs are strongly influenced by local bodies of water such as the Straight of Juan de Fuca, making it difficult to form a theoretical prediction of tidal strains \citep{Roeloffs2010}.
+Another complication is that noise in BSM data is not well understood.
 The noise consists, in part, of a long-term decay resulting from the instrument equilibrating with the surrounding rock \citep{Gladwin1987}.
 Typically, this noise is dealt with in an ad-hoc manner by fitting and removing exponentials and low-order polynomials.
-A more rigorous quantification of BSM noise could be performed by analyzing the residuals between the the BSM data and GNSS derived strain rates.
+We envision that the GNSS derived strain rates from this paper can be used as a reference strain for calibrating BSM data and quantify its noise.
 

### 2017-07-17 RevisedSubmission/hines_hetland_2017_gji.tex
added more

@@
 We have included an animation showing the map view of strain rates through time as supplementary material.
-The strain rates shown in Figure \ref{fig:StrainMap} are generally similar to the strain rates for the other six SSEs considered in this study.
-The SSEs cause compression in the Olympic Peninsula and extension east of Puget Sound.
-For comparison, estimated secular strain rates indicate trench perpendicular compression throughout this study region \citep{Murray2000,McCaffrey2007,McCaffrey2013}.
-The SSEs are thus concentrating tectonically accumulated strain energy trench-ward, and presumably pushing the subduction zone closer to failure.
-Similar conclusions have been drawn based on fault slip models \citep[e.g.,][]{Dragert2001,Wech2009,Schmidt2010}, which reveal that SSEs are occurring down-dip of the seismogenic zone and migrating stress upward.
-A key difference between the strain inferred here and strain that can be derived from fault slip models is that our estimated strain rates are not based on an assumed physical model.
+The strain rates shown in Figure \ref{fig:StrainMap} are generally similar to the strain rates during the other six SSEs considered in this study.
+The SSEs cause trench perpendicular compression in the Olympic Peninsula and extension east of Puget Sound.
+The strain transitions from compressional to extensional strain around the southern tip of Vancouver Island, which coincides with the location of where thrust slip tends to be inferred for SSEs in the Puget Sound region \citep[e.g.,][]{Dragert2001,Wech2009,Schmidt2010}.
+Thus, this pattern of strain is to be expected.
+During the period in between SSEs, secular strain rates indicate trench perpendicular compression throughout this study region \citep{Murray2000,McCaffrey2007,McCaffrey2013}.
+When comparing inferred strain rates from SSEs to the secular strain rates, we see that SSEs are concentrating tectonically accumulated strain energy towards the trench, and presumably pushing the subduction zone closer to failure.
+
+%A key difference between the strain inferred here and strain that can be derived from fault slip models is that our estimated strain rates are not based on an assumed physical model.
 In contrast, fault slip models can be biased by errors in the assumed fault geometry or lithospheric rheology.
@@
 The results we have presented thus far indicate that we are identifying the strain that we should expect to see.
-There are, however, features in our estimated transient strain rates which we were not expecting.
-For example, there is a brief period of east-west extension on the Olympic peninsula several days prior to some of the SSEs.
+There are, however, subtle features in our estimated transient strain rates which we were not expecting.
+For example, there is a brief period of east-west extension on the Olympic Peninsula several days prior to some of the SSEs.
 This feature can be seen before the summer 2012 and winter 2015-2016 SSEs in Figure \ref{fig:StrainTs} as well as in the supplementary animation.
-A discussion on the mechanisms causing this deformation is outside the scope of this study.
+While this deformation is noteworthy, a discussion on the mechanisms causing it is outside the scope of this study.
 
 \section{Discussion}\label{sec:Discussion}
-We have demonstrated that transient strain rates estimated with the method described in Section \ref{sec:Method} can be used to detect SSEs and will be robust to detect other transient geophysical phenomena.
-Another potential application would be to use GNSS derived transient strain rates to better understand the data from borehole strain meters (BSMs).
+Our results demonstrate that GPR is an effective tool for estimating transient strain from GNSS data, which can then be used to detect geophysical processes.
+Alternatively, one could detect geophysical signal by merely inspecting raw GNSS displacement time series.
+Indeed, the SSEs identified in Figure \ref{fig:StrainMag} do produce visually discernible displacements in the GNSS time series.
+However, the GNSS data also contains outliers and non-tectonic deformation that is localized to individual stations.
+In contrast, our estimates of transient strain only identify features that are sufficiently spatially and temporally coherent, based on our chosen prior model.
+Furthermore, our estimates of transient strain are insensitive to common mode noise, which is highly spatially correlated noise resulting from factors such as reference frame error.
+Common mode noise can obscure geophysical signal in the GNSS data, but it gets canceled out when computing the spatial gradients to obtain transient strain.
+Lastly, if we want to detect transient geophysical signal for the purpose of assessing seismic hazard, then strain is perhaps a more meaningful metric to monitor than displacements.
+
+In addition to detecting geophysical processes, the GNSS derived transient strain rates can be used to better understand the data from borehole strain meters (BSMs).
 The Plate Boundary Observatory maintains 43 BSMs in Cascadia, and it has been demonstrated that BSMs are able to record transient geophysical events such as SSEs \citep[e.g.,][]{Dragert2011}.
@@
 There is potential for a more thorough analysis of the spatio-temporal noise in GNSS data, $\eta$, than what was performed in Section \ref{sec:NoiseModel}.
-We did not explore the spatial covariance of $\eta$.
-Spatially correlated common mode noise, which results from factors such as reference frame error, is a non-negligible component of GNSS data.
-We ignore common mode error in this study because transients strain rates are insensitive to it.
-For other geophysical studies based on GNSS data, such as fault slip inversions, it may be necessary to incorporate a spatially covarying noise model \citep[e.g.,][]{Miyazaki2003}.
+We did not explore the spatial covariance of $\eta$, which would describe common mode noise.
+We are able to ignore common mode error in this study; however, for other geophysical studies based on GNSS data, such as fault slip inversions, it may be necessary to incorporate a spatially covarying noise model \citep[e.g.,][]{Miyazaki2003}.
 We can also improve upon the seasonal model used in this study, which consists of four spatially uncorrelated sinusoids for each station.

### 2017-07-18 RevisedSubmission/hines_hetland_2017_gji.tex
added more

@@
 Our results demonstrate that GPR is an effective tool for estimating transient strain from GNSS data, which can then be used to detect geophysical processes.
-Alternatively, one could detect geophysical signal by merely inspecting raw GNSS displacement time series.
-Indeed, the SSEs identified in Figure \ref{fig:StrainMag} do produce visually discernible displacements in the GNSS time series.
+One may argue that geophysical signal can also be detected by merely inspecting the GNSS displacement time series.
+Indeed, the SSEs identified in Figure \ref{fig:StrainMag} do produce visible displacements in the GNSS data.
 However, the GNSS data also contains outliers and non-tectonic deformation that is localized to individual stations.
@@
 Furthermore, our estimates of transient strain are insensitive to common mode noise, which is highly spatially correlated noise resulting from factors such as reference frame error.
-Common mode noise can obscure geophysical signal in the GNSS data, but it gets canceled out when computing the spatial gradients to obtain transient strain.
-Lastly, if we want to detect transient geophysical signal for the purpose of assessing seismic hazard, then strain is perhaps a more meaningful metric to monitor than displacements.
+Common mode noise can obscure geophysical signal in the GNSS data, but it gets canceled out when computing the transient strain.
+Lastly, our estimates of transient strain rates are a spatial and temporal derivative of displacements, and thus any geophysical signal in the transient strain rates tends to be more pronounced than in the GNSS data.
+For these reasons, we argue that transient strain rates estimated with the method described in Section \ref{sec:Method} can illuminate geophysical signal that may not be discernible from the noise in the GNSS displacement data.
 

### 2017-08-10 RevisedSubmission/hines_hetland_2017_gji.tex
showing fit for more stations

@@
 Our motivation for estimating transient strain rates is, in part, to detect geophysical phenomena.
+As we will see, geophysical signal can be easily identified by visually inspecting the solution for transient strain rates.
+However, if we want to detect geophysical phenomena automatically, then we need to define a detection criterion.
 We propose using a signal-to-noise ratio, SNR, based on $\dot\varepsilon$ for our detection criterion.
@@
 We compute $\hat{\mitbf{d}}$ using SE, Wendland, and IBM covariance functions for $T$ and the median hyperparameters from Table 1.
-Figure \ref{fig:Fit} compares the easting component of $\mitbf{d}_*$ to $\hat{\mitbf{d}}$ for the winter 2015-2016 SSE at station P436, which is among the stations that record the strongest signal.
-The data prediction vector reasonable fits the displacements throughout the SSE, regardless of the choice of $T$.
-The prediction for the IBM covariance function contains slightly more high frequency, and perhaps spurious, features.
-The predictions for the Wendland and SE covariance functions are nearly indistinguishable.
-Overall, the predicted displacements are not strongly sensitive to the choice of temporal covariance function.
-In our estimates of transient strain discussed in the next section, we ultimately settle on the Wendland covariance function for $T$ and use the median values from Table 1 for the hyperparameters.
+Figure \ref{fig:Fit} compares the easting component of $\mitbf{d}_*$ to $\hat{\mitbf{d}}$ during the winter 2015-2016 SSE at the three stations that record the strongest SSE signal, ALBH, P436, and SC02.
+Based on Figure \ref{fig:Fit}, $\hat{\mitbf{d}}$ is insensitive to the choice of covariance function for $T$.
+The predicted displacements for the IBM covariance function contain slightly more high frequency, and perhaps spurious, features.
+Regardless of the choice of covariance function, $\hat{\mitbf{d}}$ at stations ALBH and SC02 appears slightly too smooth when compared to $\mitbf{d}_*$.
+This could indicate that the chosen timescale hyperparameter, $\tau$, is too large.
+The predicted displacements at the remaining stations, which record more subtle SSE deformation, seem to faithfully describe $\mitbf{d}_*$.
+
+For our estimates of transient strain discussed in the next section, we ultimately settle on the Wendland covariance function for $T$ and use the median values from Table 1 for the hyperparameters.
 We choose the Wendland covariance function over the SE covariance function because of its computational advantages.
@@
 \includegraphics{figures/signal_fit/signal-fit.pdf}
-\caption{Observed easting component of displacements at station P436 and predicted displacements when using different covariance functions for $T$.
-The one standard deviation uncertainties are shown for the observations and the predicted displacements when using the SE covariance function.
-For clarity, uncertainties are not shown for the IBM and Wendland covariance functions, but they are nearly equivalent to the uncertainties for the SE covariance function.}
+\caption{Easting component of the observed displacements, $\mitbf{d}_*$, and predicted displacements, $\hat{\mitbf{d}}$, during an SSE at three stations where the SSE signal is strongest.
+The predicted displacements are shown for when $T$ is a squared exponential (SE), Wendland, and integrated Brownian motion (IBM) covariance function.
+The one standard deviation uncertainties are shown for the observations and the SE predictions.
+For clarity, uncertainties are not shown for the IBM and Wendland predictions, but they are nearly equivalent to the uncertainties for the SE predictions.
+The SE and Wendland predictions are practically indistinguishable.}
 \label{fig:Fit}
@@
 In addition to detecting geophysical processes, the GNSS derived transient strain rates can be used to better understand the data from borehole strain meters (BSMs).
-The Plate Boundary Observatory maintains 43 BSMs in Cascadia, and it has been demonstrated that BSMs are able to record transient geophysical events such as SSEs \citep[e.g.,][]{Dragert2011}.
+The Plate Boundary Observatory maintains about forty BSMs in the Pacific Northwest, and it has been demonstrated that BSMs are able to record transient geophysical events such as SSEs \citep[e.g.,][]{Dragert2011}.
 However, there are complications that prevent BSM data from being used quantitatively in geophysical studies.

### 2017-08-10 RevisedSubmission/figures/context_map/make-plot.py

@@ -76,15 +76,28 @@ lon = data['longitude'][...]
 lat = data['latitude'][...]
 x,y = bm(lon,lat)
 ax.plot(x,y,'ko',ms=4,zorder=3)
-# Show locations for SC03 and P436
+# Show locations for SC03, SC02, P436, and ALBH
 idx = np.nonzero(data['id'][...] == 'SC03')[0][0]
 stax,stay = x[idx],y[idx]
 ax.plot(stax,stay,'ro',ms=10,zorder=4)
 ax.text(stax-17000.0,stay-23000.0,'SC03',color='r',fontsize=10)
+
+idx = np.nonzero(data['id'][...] == 'SC02')[0][0]
+stax,stay = x[idx],y[idx]
+ax.plot(stax,stay,'ro',ms=10,zorder=4)
+ax.text(stax-17000.0,stay-23000.0,'SC02',color='r',fontsize=10)
+
 idx = np.nonzero(data['id'][...] == 'P436')[0][0]
 stax,stay = x[idx],y[idx]
 ax.plot(stax,stay,'ro',ms=10,zorder=4)
 ax.text(stax-17000.0,stay-23000.0,'P436',color='r',fontsize=10)
+
+idx = np.nonzero(data['id'][...] == 'ALBH')[0][0]
+stax,stay = x[idx],y[idx]
+ax.plot(stax,stay,'ro',ms=10,zorder=4)
+ax.text(stax-17000.0,stay-23000.0,'ALBH',color='r',fontsize=10)
+
+
 # plot location of strain time series
 stax,stay = bm(-124.03,47.9)
 radius = 150000.0

### 2017-08-10 RevisedSubmission/figures/signal_fit/make-new-plot.py

@@ -1,0 +1,75 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import pygeons
+from matplotlib import colors
+
+xtick_dates = ['2015-11-01','2015-12-01','2016-01-01','2016-02-01','2016-03-01']
+xticks = [pygeons.mjd.mjd(i,'%Y-%m-%d') for i in xtick_dates]
+xtick_labels = ['2015-11-01','2015-12-01','2016-01-01','2016-02-01','2016-03-01']
+
+c0 = colors.to_rgb('C0')
+c0t = tuple(0.6 + 0.4*np.array(colors.to_rgb('C0')))
+
+fig,axs = plt.subplots(3,1,figsize=(7,6),sharex=True)
+pygeons.plot.plot._setup_ts_ax(axs)
+for i,name in enumerate(['ALBH','P436','SC02']):
+  axs[i].set_xticks(xticks)
+  axs[i].set_xticklabels(xtick_labels)
+  axs[i].tick_params(labelsize=10)
+  axs[i].set_xlim((57327.0,57448.0))
+  axs[i].set_ylabel('easting [mm]')
+  #axs[i].set_ylim((-10.0,15.0))
+  axs[i].grid()
+  
+  # OBSERVED
+  data = pygeons.io.io.dict_from_hdf5('data/data.h5')
+  idx = (data['id'] == name).nonzero()[0][0]
+  lon = data['longitude'][idx]
+  lat = data['latitude'][idx]
+  t = data['time']
+  u = 1000*data['east'][:,idx]
+  us = 1000*data['east_std_dev'][:,idx]
+  
+  axs[i].set_title('station %s (%.1f$^\mathregular{\circ}$W, %.1f$^\mathregular{\circ}$N)' %
+              (name,-lon,lat),fontsize=10)
+  
+  axs[i].errorbar(t,u,us,marker='.',linestyle='None',color=c0,ecolor=c0t,ms=5.0,zorder=1,label='observed')
+  
+  # SE
+  data = pygeons.io.io.dict_from_hdf5('data/se.per.fit.h5')
+  idx = (data['id'] == name).nonzero()[0][0]
+  lon = data['longitude'][idx]
+  lat = data['latitude'][idx]
+  t = data['time']
+  u = 1000*data['east'][:,idx]
+  us = 1000*data['east_std_dev'][:,idx]
[…]

### 2017-08-10 RevisedSubmission/hines_hetland_2017_gji.tex
reworded thing

@@
 The predicted displacements for the IBM covariance function contain slightly more high frequency, and perhaps spurious, features.
-Regardless of the choice of covariance function, $\hat{\mitbf{d}}$ at stations ALBH and SC02 appears slightly too smooth when compared to $\mitbf{d}_*$.
-This could indicate that the chosen timescale hyperparameter, $\tau$, is too large.
+When comparing $\hat{\mitbf{d}}$ to $\mitbf{d}_*$ at station ALBH and SC02, it appears that $\hat{\mitbf{d}}$ slightly underestimates the rate of deformation during the SSE, regardless of the chosen covariance function.
+This over-smoothing could indicate that the chosen time-scale hyperparameter, $\tau$, is too large.
 The predicted displacements at the remaining stations, which record more subtle SSE deformation, seem to faithfully describe $\mitbf{d}_*$.

### 2017-08-13 RevisedSubmission/hines_hetland_2017_gji.tex
added more

@@
 The SSEs cause trench perpendicular compression in the Olympic Peninsula and extension east of Puget Sound.
-The strain transitions from compressional to extensional strain around the southern tip of Vancouver Island, which coincides with the location of where thrust slip tends to be inferred for SSEs in the Puget Sound region \citep[e.g.,][]{Dragert2001,Wech2009,Schmidt2010}.
+The strain transitions from compression to extension around the southern tip of Vancouver Island, which coincides with the location of thrust slip for SSEs in the Puget Sound region \citep[e.g.,][]{Dragert2001,Wech2009,Schmidt2010}.
 Thus, this pattern of strain is to be expected.
@@
 A signal-to-noise ratio greater than ${\sim}3$ can be interpreted as a detected geophysical signal.
-For each detected event there is a corresponding peak in seismic tremor.
-We are also able to clearly identify transient strain associated with a more subtle SSE in August 2014.
-In between peaks in seismic tremor, the signal-to-noise ratio is consistently between 0 and 2, suggesting that all the transient strain detected at this location is associated with SSEs.
+We detect nine distinct events, which each correspond to peaks in seismic tremor.
+The smaller events detected in August 2014 and February 2017 can be considered inter-SSE events.
+They were not among the SSEs used to constrain the prior covariance function.
+In between peaks in seismic tremor, the signal-to-noise ratio is consistently between 0 and 2, suggesting that all the transient strain detected at this location is associated with SSEs and inter-SSE events.
 
@@
 There are, however, subtle features in our estimated transient strain rates which we were not expecting.
-For example, there is a brief period of east-west extension on the Olympic Peninsula several days prior to some of the SSEs.
-This feature can be seen before the summer 2012 and winter 2015-2016 SSEs in Figure \ref{fig:StrainTs} as well as in the supplementary animation.
+For example, some SSEs are preceded by a brief period of east-west extension on the Olympic Peninsula.
+This feature can be seen several days before the summer 2012 and winter 2015-2016 SSEs in Figure \ref{fig:StrainTs} as well as in the supplementary animation.
 While this deformation is noteworthy, a discussion on the mechanisms causing it is outside the scope of this study.

### 2017-08-25 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 One limitation with GPR is that it is not robust against outliers, so we introduce a pre-processing method for detecting and removing outliers in GNSS data.
-As a demonstration, we use GPR to detect transient strain resulting from slow slip events in Cascadia.
+As a demonstration, we use GPR to detect transient strain resulting from slow slip events in the Pacific Northwest.
 Maximum likelihood methods are used to constrain a prior model for transient displacements in this region.
@@
 Our results reveal the spatial and temporal evolution of strain from slow slip events.
-We verify that the transient strain estimated with GPR is in fact geophysical signal by comparing it to the seismic tremor that is associated with Cascadia slow slip events.
+We verify that the transient strain estimated with GPR is in fact geophysical signal by comparing it to the seismic tremor that is associated with Pacific Northwest slow slip events.
 \end{summary}
@@
 The posterior transient strain is also a Gaussian process, and we can use its distribution to confidently discern geophysical signal from noise.
-We use GPR to infer strain resulting from SSEs in Cascadia, demonstrating that GPR is an effective tool for detecting transient geophysical processes.
+We use GPR to infer strain resulting from SSEs in the Pacific Northwest, demonstrating that GPR is an effective tool for detecting transient geophysical processes.
 
@@
 Of course, the secular velocities, $a^{(2)}_{i}$, are spatially correlated and we could invoke a tectonic model to form a prior on $a^{(2)}_{i}$.
-However, in our application to Cascadia, we will be using displacement time series which are long enough to sufficiently constrain $a^{(2)}_{i}$ for each station, avoiding the need to incorporate a prior.
+However, in our application to the Pacific Northwest, we will be using displacement time series which are long enough to sufficiently constrain $a^{(2)}_{i}$ for each station, avoiding the need to incorporate a prior.
 Likewise, seasonal deformation is spatially correlated \citep{Dong2002,Langbein2008}, and it may be worth exploring and exploiting such a correlation in a future study.
@@
 For now, we will assume that appropriate covariance functions and corresponding hyperparameters for $C_u$ and $C_\eta$ have already been chosen.
-In Section \ref{sec:NoiseModel} and \ref{sec:SignalModel}, we discuss how the covariance functions are chosen for our application to GNSS data from Cascadia.
+In Section \ref{sec:NoiseModel} and \ref{sec:SignalModel}, we discuss how the covariance functions are chosen for our application to GNSS data from the Pacific Northwest.
 If $\kappa$ is kept finite then, from \citet[sec.
@@
 
-Our method for detecting outliers is based on the data editing algorithm described in \citet{Gibbs2011}.
-We calculate the residuals between the observations and a best fitting model.
-Data with residuals that are anomalously large are then identified as outliers.
-We treat $\mitbf{d}_*$ as a sample of $\mitbf{d}$ and assume that there is no correlated noise (i.e., $\eta(p) = 0$).
-The best fitting model for $\mitbf{d}_*$ is considered to be the expected value of the random vector $u(\mitbf{P}) + \mitbf{G}\mitbf{a}$ after conditioning it with non-outlier observations.
+Our method for detecting outliers is based on the data editing algorithm described in \citet{Acheson1975}.
+We identify outliers by using the residuals between the observations, $\mitbf{d}_*$, and a best fitting model.
+Data with residuals that are anomalously large are flagged as outliers.
+We treat $\mitbf{d}_*$ as a sample of $\mitbf{d}$, but we ignore any correlated noise (i.e., $\eta(p) = 0$).
+The best fitting model for $\mitbf{d}_*$ is then considered to be the expected value of the random vector $u(\mitbf{P}) + \mitbf{G}\mitbf{a}$ after conditioning it with non-outlier observations.
 We still consider $u$ to have a separable covariance function as in eq.
 (\ref{eq:TransientCovariance}), and the choice for $X$ and $T$ does not need to be the same as that used in Section \ref{sec:Method}.
+We only desire that $u$ is sufficiently flexible to describe any signal that cannot be described by the linear trend or seasonal sinusoids in $\mitbf{G}$.
 Since outliers are determined based on how well a spatially and temporally dependent model fits the data, we are able to identify anomalous observations which may not be immediately apparent from inspecting individual time series.
@@
 
-\section{Application to Cascadia Slow Slip Events}\label{sec:Cascadia}
-We use our method to estimate transient strain rates in Cascadia, and we are specifically interested in identifying strain resulting from SSEs \citep[e.g.,][]{Dragert2001}.
-In Cascadia, SSEs can be detected by monitoring for associated seismic tremor \citep{Rogers2003}, which is actively being done by the Pacific Northwest Seismic Network \citep{Wech2010}.
+\section{Application to Pacific Northwest Slow Slip Events}\label{sec:Cascadia}
+We use our method to estimate transient strain rates in the Pacific Northwest, and we are specifically interested in identifying strain resulting from SSEs \citep[e.g.,][]{Dragert2001}.
+In the Pacific Northwest, SSEs can be detected by monitoring for associated seismic tremor \citep{Rogers2003}, which is actively being done by the Pacific Northwest Seismic Network \citep{Wech2010}.
 We can thus compare the tremor records to the transient strain rates estimated with GPR to verify that we are indeed identifying strain from SSEs.
@@
 The FOGM process degenerates to the commonly used Brownian motion noise model under the condition that $\alpha=0$ and $\eta(0) = 0$.
-Our noise model that satisfies eq.
+Our spatially uncorrelated noise model that satisfies eq.
 (\ref{eq:FOGMdef}) is a Gaussian process with zero mean and the covariance function
 \begin{equation}\label{eq:FOGM}
-C_\eta\left((\vec{x},t),(\vec{x}',t')\right) = \frac{\beta^2}{2\alpha}\exp\left(-\alpha|t - t'|\right) \delta(||\vec{x} - \vec{x}'||_2). 
+C_\eta\left((\vec{x},t),(\vec{x}',t')\right) = 
+\begin{cases}
+\frac{\beta^2}{2\alpha}\exp\left(-\alpha|t - t'|\right), &\vec{x} = \vec{x}' \\
+0, &\vec{x} \neq \vec{x}'
+\end{cases}. 
 \end{equation}
 
-We constrain the hyperparameters for $\eta$, $\alpha$ and $\beta$, with a set of 38 continuous GNSS stations in Cascadia that are east of 121$^\circ$W.
+We constrain the hyperparameters for $\eta$, $\alpha$ and $\beta$, with a set of 38 continuous GNSS stations in the Pacific Northwest that are east of 121$^\circ$W.
 These stations are sufficiently far from the subduction zone that they are unlikely to contain transient signal associated with SSEs.
@@
 The distribution of inferred $\alpha$ and $\beta$ are shown in Figure \ref{fig:NoiseParams}.
-The amplitude of FOGM noise, $\beta$, for the easting and northing components is notable low and are clustered around 0.5 mm/yr$^{0.5}$.
+The amplitude of FOGM noise, $\beta$, for the easting and northing components is notably low and are clustered around 0.5 mm/yr$^{0.5}$.
 The corresponding estimates of $\alpha$ tend to cluster around 0 yr$^{-1}$, suggesting that noise can be described with Brownian motion.
@@
 \ref{eq:FOGM}).
-Hyperparameters are estimated with the REML method for 38 stations in Cascadia that are east of $121^\circ$W.
+Hyperparameters are estimated with the REML method for 38 stations in the Pacific Northwest that are east of $121^\circ$W.
 ``IQR'' is the interquartile range.}
@@
 The SE covariance function is commonly used in kriging \citep[e.g,][]{Cressie1992} and Gaussian process regression \citep[e.g.,][]{Rasmussen2006}.
+In terms of geodetic applications, \citet{Kato1998} and \cite{El-Fiky1999} demonstrated that the SE accurately describes the covariance of secular GNSS derived velocities in Japan.
 The SE is a positive definite covariance function for any number of spatial dimensions.
 A Gaussian process with an SE covariance function is isotropic and has realizations that are infinitely differentiable.
-In terms of geodetic applications, \citet{Kato1998} and \cite{El-Fiky1999} demonstrated that the SE accurately describes the covariance of secular GNSS derived velocities in Japan.
 
@@
 First, we clean the GNSS datasets by removing offsets at times of equipment changes and removing outliers with the method describe in Section \ref{sec:Outlier}.
-For the outlier detection algorithm, our prior model, $u$, is chosen to have a length-scale and time-scale which is able to approximately describe SSE displacements.
+For the outlier detection algorithm, our prior model, $u$, is chosen to have a length-scale and time-scale that is able to approximately describe SSE displacements.
 We use the SE covariance function for $X$ with length-scale $\ell = 100$ km, and we use the Wendland covariance function for $T$, due to its computational efficiency, with time-scale $\tau = 0.1$ yr and $\phi = 1$ mm.
@@
 We use the REML method to find the optimal hyperparameters for $T$ and $X$ for each subset of data.
-We choose to make each data subsets four months long because it is long enough to encompass a SSE in Cascadia, while it is short enough to still be computationally tractable.
-However, four months is too short to resolve the sinusoids in $\mitbf{d}$, and they are omitted from $\mitbf{d}$ in this REML analysis for Cascadia SSEs.
+We choose to make each data subsets four months long because it is long enough to encompass a SSE in the Pacific Northwest, while it is short enough to still be computationally tractable.
+However, four months is too short to resolve the sinusoids in $\mitbf{d}$, and they are omitted from $\mitbf{d}$ in this REML analysis for Pacific Northwest SSEs.
 The estimated hyperparameters for $u$ are summarized in Table 1.
 Based on the interquartile ranges, the estimated hyperparameters for the SE and Wendland covariance functions do not vary significantly between SSEs.
-This suggests that the median of estimated hyperparameters should be an appropriate prior model for all Cascadia SSEs.
+This suggests that the median of estimated hyperparameters should be an appropriate prior model for all Pacific Northwest SSEs.
 For the IBM model, there are several anomalously large values for $\ell$ and $\phi$, which results in large interquartile ranges.
@@
 The strain rates are estimates at a grid of points spanning the study area.
-In Figure \ref{fig:StrainMap} we show the transient strain rates on January 1, 2016, which coincides with the height of an SSE.
+In Figure \ref{fig:StrainMap} we show the transient strain rates on January 1, 2016, which coincides with the height of the winter 2015-2016 SSE.
 We have included an animation showing the map view of strain rates through time as supplementary material.
@@
 Because GPR is a Bayesian method, the uncertainties on our estimated transient strain rates are well quantified, allowing one to discern geophysical signal from noise.
-We demonstrate that GPR is an effective tool for detecting geophysical phenomena, such as slow slip events, in our application to GNSS data from Cascadia.
+We demonstrate that GPR is an effective tool for detecting geophysical phenomena, such as slow slip events, in our application to GNSS data from the Pacific Northwest.
 One limitation with GPR is that it is not robust against outliers.

### 2017-09-14 RevisedSubmission/hines_hetland_2017_gji.tex
did more shit

@@
 where $C_{u_i}(p,p')$ is a covariance function indicating how we expect $u_i(p)$ to covary with $u_i(p')$.
-For simplicity, we treat each component of displacements independently and ignore any potential covariance between them.
-Hence, we drop the component subscripts with the understanding that the same analysis is being repeated to estimate the easting and northing components of $\vec{u}$.
-We assume that $C_u$ can be separated into spatial and temporal functions as
+The function $C_{u_i}$ must be positive definite, which means that the matrix $[C_{u_i}(p,p')]_{(p,p') \in \mitbf{P} \times \mitbf{P}}$ is positive definite for any set of ordered pairs $\mitbf{P}$ \citep[sec.
+2.5]{Cressie1992}.
+We assume that $C_{u_i}$ can be separated into spatial and temporal functions as
 \begin{equation}\label{eq:TransientCovariance}
-C_{u}\left((\vec{x},t),(\vec{x}',t')\right) = X(\vec{x},\vec{x}')T(t,t').
+C_{u_i}\left((\vec{x},t),(\vec{x}',t')\right) = X_i(\vec{x},\vec{x}')T_i(t,t').
 \end{equation}  
-As long as the functions $X$ and $T$ are positive definite, $C_u$ is guaranteed to also be positive definite and thus a valid covariance function \citep[sec.
+As long as the functions $X_i$ and $T_i$ are positive definite, $C_{u_i}$ is guaranteed to also be positive definite and thus a valid covariance function \citep[sec.
 4.2.4]{Rasmussen2006}.
-The appropriate choice for $X$ and $T$ may vary depending on the geophysical signal we are trying to describe, and we discuss this matter in Section \ref{sec:SignalModel}.
+The appropriate choice for $X_i$ and $T_i$ may vary depending on the geophysical signal we are trying to describe, and we discuss this matter in Section \ref{sec:SignalModel}.
 
-We constrain $u$ with GNSS data, which records $u$ as well as other physical and non-physical processes that we are not interested in.
+A posterior estimate of transient displacement is formed by conditioning $\vec{u}$ with GNSS observations.
+GNSS data records transient displacements as well as other physical and non-physical processes that we are not interested in.
 We describe GNSS observations at position $\vec{x}_i$ and time $t_j$ as a realization of the random variable
@@
 \begin{split}
-d_{ij} = &u(\vec{x}_i,t_j) + \eta(\vec{x}_i,t_j) + w_{ij} + a^{(1)}_i + a^{(2)}_it_j + \\
-         &a^{(3)}_i\sin(2 \pi t_j) + a^{(4)}_i\cos(2 \pi t_j) + a^{(5)}_i\sin(4 \pi t_j) + a^{(6)}_i\cos(4 \pi t_j), 
+\vec{d}_{ij} = &\vec{u}(\vec{x}_i,t_j) + \vec{\eta}(\vec{x}_i,t_j) + \vec{w}_{ij} + \vec{a}^{(1)}_i + \vec{a}^{(2)}_it_j + \\
+               &\vec{a}^{(3)}_i\sin(2 \pi t_j) + \vec{a}^{(4)}_i\cos(2 \pi t_j) + \vec{a}^{(5)}_i\sin(4 \pi t_j) + \vec{a}^{(6)}_i\cos(4 \pi t_j), 
 \end{split}
 \end{align}
-where $a^{(1)}_{i}$ is an offset that is unique for each GNSS monument, $a^{(2)}_{i}$ is the secular velocity at $\vec{x}_i$, and the sinusoids describe seasonal deformation (using units of years for $t_j$).
-We use $w_{ij}$ to denote normally distributed, uncorrelated noise.
-Correlated noise which does not have a parametric representation is denoted by $\eta$.
-For example, $\eta$ can consist of temporally correlated noise describing benchmark wobble \citep[e.g.,][]{Wyatt1982,Wyatt1989}, and/or spatially correlated noise describing common mode error \citep[e.g.,][]{Wdowinski1997}.
-For now, we will only assume that $\eta \sim \mathcal{N}(0,C_\eta)$.
-We consider the six coefficients in eq.
+where $\vec{a}^{(1)}_{i}$ is an offset that is unique for each GNSS station, $\vec{a}^{(2)}_{i}$ is the secular velocity at $\vec{x}_i$, and the sinusoids describe seasonal deformation (using units of years for $t_j$).
+We use $\vec{w}_{ij}$ to denote normally distributed, uncorrelated noise.
+Correlated noise which does not have a parametric representation is denoted by $\vec{\eta}$.
+For example, $\vec{\eta}$ can consist of temporally correlated noise describing benchmark wobble \citep[e.g.,][]{Wyatt1982,Wyatt1989}, and/or spatially correlated noise describing common mode error \citep[e.g.,][]{Wdowinski1997}.
+For now, we will assume that each component of $\vec{\eta}$ is a Gaussian process distributed as $\eta_i \sim \mathcal{N}(0,C_{\eta_i})$, and $C_{\eta_i}$ is a spatially and temporally dependent covariance function.
+We consider the coefficients in eq.
 (\ref{eq:Data}) to be uncorrelated random variables distributed as $\mathcal{N}(0,\kappa^2)$ in the limit as $\kappa \to \infty$ (i.e., the coefficients have diffuse priors).
-Of course, the secular velocities, $a^{(2)}_{i}$, are spatially correlated and we could invoke a tectonic model to form a prior on $a^{(2)}_{i}$.
-However, in our application to the Pacific Northwest, we will be using displacement time series which are long enough to sufficiently constrain $a^{(2)}_{i}$ for each station, avoiding the need to incorporate a prior.
+Of course, the secular velocities, $\vec{a}^{(2)}_{i}$, are spatially correlated and we could invoke a tectonic model to form a prior on $\vec{a}^{(2)}_{i}$.
+However, in our application to the Pacific Northwest, we will be using displacement time series which are long enough to sufficiently constrain $\vec{a}^{(2)}_{i}$ for each station, avoiding the need to incorporate a prior.
 Likewise, seasonal deformation is spatially correlated \citep{Dong2002,Langbein2008}, and it may be worth exploring and exploiting such a correlation in a future study.
+We ignore any covariances between the easting and northing components of $\vec{d}_{ij}$.
+This simplifies our analysis by allowing us to constrain each component of transient displacements independently.
 
-We now consider the column vector of $n$ GNSS observations made at $m$ stations, $\mitbf{d}_*$.
-Let $\mitbf{P}$ be the set of $(\vec{x}_i, t_j)$ pairs describing where and when each of the GNSS observations have been made.
-Let $\mitbf{a}$ be the vector of coefficients from eq.
-(\ref{eq:Data}) for each of the $m$ GNSS stations.
+Each component of the posterior transient displacements, denoted as $\hat{u}_i$, is constrained by conditioning $u_i$ with GNSS observations of displacements along direction $i$, which we write as $\mitbf{d}^*_i$.
+The data vector consists of $n$ observations, which have been made at $m$ stations.
+We let $\mitbf{P}$ be the set of $(\vec{x}_i, t_j)$ pairs describing where and when each of the observations have been made.
+We also let $\mitbf{a}_i$ be the length $6m$ vector of coefficients from eq.
+(\ref{eq:Data}) for each station and for direction $i$.
 We use $\mitbf{G}$ to represent the $n \times 6m$ matrix of corresponding basis functions evaluated at each point in $\mitbf{P}$.
-We  also denote the vector of uncorrelated noise for each observation as $\mitbf{w}$, whose standard deviations are given by the formal data uncertainty, $\mitbf{\sigma}$.
-The observations can then be viewed as a realization of the random vector
+We can then view $\mitbf{d}^*_i$ as a realization of the random vector
 \begin{equation}
-\mitbf{d} = u(\mitbf{P}) + \eta(\mitbf{P}) + \mitbf{w} + \mitbf{G}\mitbf{a},
+\mitbf{d}_i = u_i(\mitbf{P}) + \eta_i(\mitbf{P}) + \mitbf{G}\mitbf{a}_i,
 \end{equation}

### 2017-09-16 RevisedSubmission/hines_hetland_2017_gji.tex

@@
 \section{Estimating Transient Strain Rates}\label{sec:Method}
-We seek a spatially and temporally dependent estimate of transient crustal strain rates.
-We consider transient strain rates to be any deviation from secular strain rates, and our attention is limited to horizontal strain rates in this study.
-We denote transient crustal strain rates as $\dot\varepsilon(p)$, where $p$ represents the ordered pair $(\vec{x},t)$, $\vec{x}$ are spatial coordinates in $\mathbb{R}^2$, and $t$ is time.
-We determine $\dot\varepsilon(p)$ by spatially and temporally differentiating estimates of transient displacements, $\vec{u}(p)$.
-We make a prior assumption that each component of $\vec{u}$ is a Gaussian process,
+We seek a spatially and temporally dependent estimate of transient strain rates.
+We consider transient strain rates to be any deviation from secular strain rates, and our attention is limited to horizontal strain rates.
+We denote transient strain rates as $\dot\varepsilon(p)$, where $p$ represents the ordered pair $(\vec{x},t)$, $\vec{x}$ are spatial coordinates in a 2-D map projection aligned with the cardinal directions (i.e, $\vec{x} = (x_\mathrm{e},x_\mathrm{n})$), and $t$ is time.
+We determine $\dot\varepsilon$ by spatially and temporally differentiating estimates of transient displacements, $\vec{u}(p)$.
+Transient displacements have an easting and northing component (i.e., $\vec{u} = (u_\mathrm{e}, u_\mathrm{n})$), and we make a prior assumption that each component of $\vec{u}$ is a Gaussian process,
 \begin{equation}\label{eq:TransientDeformation}
-u_i(p) \sim \mathcal{N}\left(0,C_{u_i}\right),
+u_i \sim \mathcal{N}\left(0,C_{u_i}\right),
 \end{equation}
@@
 
-A posterior estimate of transient displacement is formed by conditioning $\vec{u}$ with GNSS observations.
 GNSS data records transient displacements as well as other physical and non-physical processes that we are not interested in.
-We describe GNSS observations at position $\vec{x}_i$ and time $t_j$ as a realization of the random variable
+We describe GNSS displacement observations at position $\vec{x}_j$ and time $t_k$ as a realization of the random variable
 \begin{align}\label{eq:Data}
 \begin{split}
-\vec{d}_{ij} = &\vec{u}(\vec{x}_i,t_j) + \vec{\eta}(\vec{x}_i,t_j) + \vec{w}_{ij} + \vec{a}^{(1)}_i + \vec{a}^{(2)}_it_j + \\
-               &\vec{a}^{(3)}_i\sin(2 \pi t_j) + \vec{a}^{(4)}_i\cos(2 \pi t_j) + \vec{a}^{(5)}_i\sin(4 \pi t_j) + \vec{a}^{(6)}_i\cos(4 \pi t_j), 
+\vec{d}_{jk} = &\vec{u}(\vec{x}_j,t_k) + \vec{\eta}(\vec{x}_j,t_k) + \vec{a}^{(1)}_j + \vec{a}^{(2)}_jt_k + \vec{a}^{(3)}_j\sin(2 \pi t_k) + \\ 
+               &\vec{a}^{(4)}_j\cos(2 \pi t_k) + \vec{a}^{(5)}_j\sin(4 \pi t_k) + \vec{a}^{(6)}_j\cos(4 \pi t_k), 
 \end{split}
 \end{align}
-where $\vec{a}^{(1)}_{i}$ is an offset that is unique for each GNSS station, $\vec{a}^{(2)}_{i}$ is the secular velocity at $\vec{x}_i$, and the sinusoids describe seasonal deformation (using units of years for $t_j$).
-We use $\vec{w}_{ij}$ to denote normally distributed, uncorrelated noise.
-Correlated noise which does not have a parametric representation is denoted by $\vec{\eta}$.
-For example, $\vec{\eta}$ can consist of temporally correlated noise describing benchmark wobble \citep[e.g.,][]{Wyatt1982,Wyatt1989}, and/or spatially correlated noise describing common mode error \citep[e.g.,][]{Wdowinski1997}.
-For now, we will assume that each component of $\vec{\eta}$ is a Gaussian process distributed as $\eta_i \sim \mathcal{N}(0,C_{\eta_i})$, and $C_{\eta_i}$ is a spatially and temporally dependent covariance function.
+where $\vec{a}^{(1)}_j$ is an offset that is unique for each GNSS station, $\vec{a}^{(2)}_j$ is the secular velocity at $\vec{x}_j$, and the sinusoids describe seasonal deformation (using units of years for $t_k$).
+We use $\vec{\eta}$ to represent all the stochastic noise processes.
+For example, $\vec{\eta}$ can consist of uncorrelated white noise, temporally correlated noise describing benchmark wobble \citep[e.g.,][]{Wyatt1982,Wyatt1989}, and/or spatially correlated noise describing common mode error \citep[e.g.,][]{Wdowinski1997}.
+For now, we only make the assumption that each component of $\vec{\eta}$ is a Gaussian process distributed as $\eta_i \sim \mathcal{N}(0,C_{\eta_i})$, and $C_{\eta_i}$ is a spatially and temporally dependent covariance function.
 We consider the coefficients in eq.
 (\ref{eq:Data}) to be uncorrelated random variables distributed as $\mathcal{N}(0,\kappa^2)$ in the limit as $\kappa \to \infty$ (i.e., the coefficients have diffuse priors).
-Of course, the secular velocities, $\vec{a}^{(2)}_{i}$, are spatially correlated and we could invoke a tectonic model to form a prior on $\vec{a}^{(2)}_{i}$.
-However, in our application to the Pacific Northwest, we will be using displacement time series which are long enough to sufficiently constrain $\vec{a}^{(2)}_{i}$ for each station, avoiding the need to incorporate a prior.
+Of course, the secular velocities, $\vec{a}^{(2)}_j$, are spatially correlated and we could invoke a tectonic model to form a prior on $\vec{a}^{(2)}_j$.
+However, in our application to the Pacific Northwest, we will be using displacement time series which are long enough to sufficiently constrain $\vec{a}^{(2)}_j$ for each station, avoiding the need to incorporate a prior.
 Likewise, seasonal deformation is spatially correlated \citep{Dong2002,Langbein2008}, and it may be worth exploring and exploiting such a correlation in a future study.
-We ignore any covariances between the easting and northing components of $\vec{d}_{ij}$.
-This simplifies our analysis by allowing us to constrain each component of transient displacements independently.
 
-Each component of the posterior transient displacements, denoted as $\hat{u}_i$, is constrained by conditioning $u_i$ with GNSS observations of displacements along direction $i$, which we write as $\mitbf{d}^*_i$.
-The data vector consists of $n$ observations, which have been made at $m$ stations.
-We let $\mitbf{P}$ be the set of $(\vec{x}_i, t_j)$ pairs describing where and when each of the observations have been made.
-We also let $\mitbf{a}_i$ be the length $6m$ vector of coefficients from eq.
-(\ref{eq:Data}) for each station and for direction $i$.
-We use $\mitbf{G}$ to represent the $n \times 6m$ matrix of corresponding basis functions evaluated at each point in $\mitbf{P}$.
+To reduce the computational complexity of our analysis, we assume the easting and northing components of $\vec{u}$ are uncorrelated.
+Likewise, we ignore any covariances between the components in $\vec{\eta}$.
+By ignoring covariances between the easting and northing components, we are inherently assuming that the principle components of $\vec{u}$ and $\vec{\eta}$ are aligned with the cardinal directions, which is admittedly an arbitrary assumption.
+
+We combine the prior on transient displacements with GNSS displacement data to form a posterior estimate of transient displacements.
+Each component of the posterior, denoted as $\hat{u}_i$, is constrained by conditioning $u_i$ with GNSS data for the corresponding direction.
+We denote the data vector consisting of $n$ GNSS displacement observations as $\mitbf{d}^*_i$.
+We let $\mitbf{P}$ be the set of $(\vec{x}_j, t_k)$ pairs describing where and when each of the observations have been made.
+We collect the coefficients from eq.
+(\ref{eq:Data}) into the vector $\mitbf{a}_i$, which contains 6 elements for each GNSS station.
+We use $\mitbf{G}$ to represent the matrix of corresponding basis functions evaluated at each point in $\mitbf{P}$.
 We can then view $\mitbf{d}^*_i$ as a realization of the random vector

### 2017-09-16 RevisedSubmission/hines_hetland_2017_gji.tex
did more shit

@@
 GNSS data records transient displacements as well as other physical and non-physical processes that we are not interested in.
-We describe GNSS displacement observations at position $\vec{x}_j$ and time $t_k$ as a realization of the random variable
+We describe component $i$ of GNSS displacement observations at station location $\vec{x}^{(j)}$ and time $t$ as a realization of the random variable
 \begin{align}\label{eq:Data}
 \begin{split}
-\vec{d}_{jk} = &\vec{u}(\vec{x}_j,t_k) + \vec{\eta}(\vec{x}_j,t_k) + \vec{a}^{(1)}_j + \vec{a}^{(2)}_jt_k + \vec{a}^{(3)}_j\sin(2 \pi t_k) + \\ 
-               &\vec{a}^{(4)}_j\cos(2 \pi t_k) + \vec{a}^{(5)}_j\sin(4 \pi t_k) + \vec{a}^{(6)}_j\cos(4 \pi t_k), 
+d^{(j)}_i(t) = &u_i(\vec{x}^{(j)},t) + \eta_i(\vec{x}^{(j)},t) + a_i^{(j)} + b_i^{(j)}t + c_i^{(j)}\sin(2 \pi t) + \\ 
+               &e_i^{(j)}\cos(2 \pi t) + f_i^{(j)}\sin(4 \pi t)  + g_i^{(j)}\cos(4 \pi t), 
 \end{split}
 \end{align}
-where $\vec{a}^{(1)}_j$ is an offset that is unique for each GNSS station, $\vec{a}^{(2)}_j$ is the secular velocity at $\vec{x}_j$, and the sinusoids describe seasonal deformation (using units of years for $t_k$).
-We use $\vec{\eta}$ to represent all the stochastic noise processes.
-For example, $\vec{\eta}$ can consist of uncorrelated white noise, temporally correlated noise describing benchmark wobble \citep[e.g.,][]{Wyatt1982,Wyatt1989}, and/or spatially correlated noise describing common mode error \citep[e.g.,][]{Wdowinski1997}.
-For now, we only make the assumption that each component of $\vec{\eta}$ is a Gaussian process distributed as $\eta_i \sim \mathcal{N}(0,C_{\eta_i})$, and $C_{\eta_i}$ is a spatially and temporally dependent covariance function.
+where $a_i^{(j)}$ is an offset that is unique for each GNSS station, $b_i^{(j)}$ is the secular velocity at $\vec{x}^{(j)}$, and the sinusoids describe seasonal deformation (using units of years for $t$).
+We use $\eta_i$ to represent all the stochastic noise processes.
+For example, $\eta_i$ can consist of uncorrelated white noise, temporally correlated noise describing benchmark wobble \citep[e.g.,][]{Wyatt1982,Wyatt1989}, and/or spatially correlated noise describing common mode error \citep[e.g.,][]{Wdowinski1997}.
+For now, we only make the assumption that each component of $\eta_i$ is a Gaussian process distributed as $\eta_i \sim \mathcal{N}(0,C_{\eta_i})$, and $C_{\eta_i}$ is a spatially and temporally dependent covariance function.
 We consider the coefficients in eq.
 (\ref{eq:Data}) to be uncorrelated random variables distributed as $\mathcal{N}(0,\kappa^2)$ in the limit as $\kappa \to \infty$ (i.e., the coefficients have diffuse priors).
-Of course, the secular velocities, $\vec{a}^{(2)}_j$, are spatially correlated and we could invoke a tectonic model to form a prior on $\vec{a}^{(2)}_j$.
-However, in our application to the Pacific Northwest, we will be using displacement time series which are long enough to sufficiently constrain $\vec{a}^{(2)}_j$ for each station, avoiding the need to incorporate a prior.
+Of course, the secular velocities, $b_i^{(j)}$, are spatially correlated and we could invoke a tectonic model to form a prior on $b_i^{(j)}$.
+However, in our application to the Pacific Northwest, we will be using displacement time series which are long enough to sufficiently constrain $b_i^{(j)}$ for each station, avoiding the need to incorporate a prior.
 Likewise, seasonal deformation is spatially correlated \citep{Dong2002,Langbein2008}, and it may be worth exploring and exploiting such a correlation in a future study.
 
-To reduce the computational complexity of our analysis, we assume the easting and northing components of $\vec{u}$ are uncorrelated.
-Likewise, we ignore any covariances between the components in $\vec{\eta}$.
-By ignoring covariances between the easting and northing components, we are inherently assuming that the principle components of $\vec{u}$ and $\vec{\eta}$ are aligned with the cardinal directions, which is admittedly an arbitrary assumption.
-
-We combine the prior on transient displacements with GNSS displacement data to form a posterior estimate of transient displacements.
-Each component of the posterior, denoted as $\hat{u}_i$, is constrained by conditioning $u_i$ with GNSS data for the corresponding direction.
-We denote the data vector consisting of $n$ GNSS displacement observations as $\mitbf{d}^*_i$.
-We let $\mitbf{P}$ be the set of $(\vec{x}_j, t_k)$ pairs describing where and when each of the observations have been made.
+We now consider component $i$ of $n$ GNSS displacement observations, denoted as the column vector $\mitbf{d}^*_i$.
+We let $\mitbf{P}$ be the set of $(\vec{x}^{(j)}, t)$ pairs describing where and when each of the observations has been made.
 We collect the coefficients from eq.
-(\ref{eq:Data}) into the vector $\mitbf{a}_i$, which contains 6 elements for each GNSS station.
-We use $\mitbf{G}$ to represent the matrix of corresponding basis functions evaluated at each point in $\mitbf{P}$.
+(\ref{eq:Data}) into the vector $\mitbf{m}_i$, which contains 6 elements for each of the $m$ GNSS station.
+We use $\mitbf{G}$ to represent the $n \times 6m$ matrix of corresponding basis functions evaluated at each point in $\mitbf{P}$.
 We can then view $\mitbf{d}^*_i$ as a realization of the random vector
 \begin{equation}
-\mitbf{d}_i = u_i(\mitbf{P}) + \eta_i(\mitbf{P}) + \mitbf{G}\mitbf{a}_i,
+\mitbf{d}_i = u_i(\mitbf{P}) + \eta_i(\mitbf{P}) + \mitbf{G}\mitbf{m}_i,
 \end{equation}
@@
 \begin{equation}\label{eq:Cd}
-\mitbf{\Sigma} = C_u(\mitbf{P},\mitbf{P}) + C_\eta(\mitbf{P},\mitbf{P}) + 
-              \mathrm{diag}\left(\mitbf{\sigma}^2\right).  
+\mitbf{\Sigma} = C_{u_i}(\mitbf{P},\mitbf{P}) + C_{\eta_i}(\mitbf{P},\mitbf{P}).  
 \end{equation}
-It should be understood that notation such as $u(\mitbf{P})$ and $C_u(\mitbf{P},\mitbf{P})$ represents the column vector $[u(P_i)]_{P_i \in \mitbf{P}}$ and the matrix $[C_u(P_i,P_j)]_{(P_i,P_j) \in \mitbf{P} \times \mitbf{P}}$, respectively.
+It should be understood that notation such as $u_i(\mitbf{P})$ and $C_{u_i}(\mitbf{P},\mitbf{P})$ represents the column vector $[u_i(p)]_{p \in \mitbf{P}}$ and the matrix $[C_{u_i}(p,p')]_{(p,p') \in \mitbf{P} \times \mitbf{P}}$, respectively.
 
-The prior for transient displacements is then conditioned with $\mitbf{d}_*$ to form a posterior estimate of transient displacements, $\hat{u} = u | \mitbf{d}_*$.
-For now, we will assume that appropriate covariance functions and corresponding hyperparameters for $C_u$ and $C_\eta$ have already been chosen.
-In Section \ref{sec:NoiseModel} and \ref{sec:SignalModel}, we discuss how the covariance functions are chosen for our application to GNSS data from the Pacific Northwest.
+We combine the prior on transient displacements with GNSS displacement data to form a posterior estimate of transient displacements.
+We ignore any covariances between the easting and northing components, which reduces the computational complexity of our analysis because we can determine each component of the posterior independently.
+However, we are inherently assuming that the principle components of transient displacements and the noise processes are aligned with the cardinal directions, which is admittedly an arbitrary assumption.
+We form each component of the posterior, denoted as $\hat{u}_i$, by conditioning $u_i$ with $\mitbf{d}^*_i$.
+That is to say $\hat{u}_i = u_i | \mitbf{d}^*_i$.
 If $\kappa$ is kept finite then, from \citet[sec.

### 2017-09-23 RevisedSubmission/hines_hetland_2017_gji.tex
did more shit

@@
 We consider transient strain rates to be any deviation from secular strain rates, and our attention is limited to horizontal strain rates.
-We denote transient strain rates as $\dot\varepsilon(p)$, where $p$ represents the ordered pair $(\vec{x},t)$, $\vec{x}$ are spatial coordinates in a 2-D map projection aligned with the cardinal directions (i.e, $\vec{x} = (x_\mathrm{e},x_\mathrm{n})$), and $t$ is time.
+We denote transient strain rates as $\dot\varepsilon(p)$, where $p$ represents the ordered pair $(\vec{x},t)$, $\vec{x}$ are spatial coordinates, and $t$ is time.
+For simplicity, we assume that the study region is sufficiently small that $\vec{x}$ can be considered a point in a 2-D map projection that is aligned with the cardinal directions, $\vec{x} = (x_\mathrm{e},x_\mathrm{n})$.
 We determine $\dot\varepsilon$ by spatially and temporally differentiating estimates of transient displacements, $\vec{u}(p)$.
-Transient displacements have an easting and northing component (i.e., $\vec{u} = (u_\mathrm{e}, u_\mathrm{n})$), and we make a prior assumption that each component of $\vec{u}$ is a Gaussian process,
-\begin{equation}\label{eq:TransientDeformation}
-u_i \sim \mathcal{N}\left(0,C_{u_i}\right),
-\end{equation}
-where $C_{u_i}(p,p')$ is a covariance function indicating how we expect $u_i(p)$ to covary with $u_i(p')$.
-The function $C_{u_i}$ must be positive definite, which means that the matrix $[C_{u_i}(p,p')]_{(p,p') \in \mitbf{P} \times \mitbf{P}}$ is positive definite for any set of ordered pairs $\mitbf{P}$ \citep[sec.
+Transient displacements have an easting and northing component, $\vec{u} = (u_\mathrm{e}, u_\mathrm{n})$, and we make a prior assumption that each component of $\vec{u}$ is a Gaussian process with zero mean and a generic covariance function $\mathrm{cov}\left(u_i(p),u_i(p')\right) = C_{u_i}(p,p')$.
+Using a more concise notation, we describe each component of $\vec{u}$ as $u_i \sim \mathcal{N}\left(0,C_{u_i}\right)$.
+The function $C_{u_i}$ must be positive definite, which means that the matrix $[C_{u_i}(p,p')]_{(p,p') \in \mitbf{P} \times \mitbf{P}}$ is positive definite for any set of points $\mitbf{P}$ \citep[sec.
 2.5]{Cressie1992}.
@@
 GNSS data records transient displacements as well as other physical and non-physical processes that we are not interested in.
-We describe component $i$ of GNSS displacement observations at station location $\vec{x}^{(j)}$ and time $t$ as a realization of the random variable
+We first consider component $i$ of a single GNSS displacement observation made at station $j$, which is located at $\vec{x}^{(j)}$, and time $t$.
+We describe this observation as a realization of the random variable
 \begin{align}\label{eq:Data}
@@
 \end{align}
-where $a_i^{(j)}$ is an offset that is unique for each GNSS station, $b_i^{(j)}$ is the secular velocity at $\vec{x}^{(j)}$, and the sinusoids describe seasonal deformation (using units of years for $t$).
+where $a_i^{(j)}$ is an offset that is unique for each station, $b_i^{(j)}$ is the secular velocity at $\vec{x}^{(j)}$, and the sinusoids describe seasonal deformation (using units of years for $t$).
 We use $\eta_i$ to represent all the stochastic noise processes.
@@
 
-We now consider component $i$ of $n$ GNSS displacement observations, denoted as the column vector $\mitbf{d}^*_i$.
-We let $\mitbf{P}$ be the set of $(\vec{x}^{(j)}, t)$ pairs describing where and when each of the observations has been made.
-We collect the coefficients from eq.
+We move on to consider the column vector of $n$ GNSS displacement observations, $\mitbf{d}^*_i$, where the subscript indicates that we are still only considering component $i$ of displacements.
+The times and positions for each observation are described by the set $\mitbf{P}$.
+$\mitbf{d}^*_i$ can be considered a realization of the random vector $\mitbf{d}_i$, which is formed by evaluating eq.
+(\ref{eq:Data}) at each point in $\mitbf{P}$.
+To write out $\mitbf{d}_i$ explicitly, we collect the coefficients from eq.
 (\ref{eq:Data}) into the vector $\mitbf{m}_i$, which contains 6 elements for each of the $m$ GNSS station.
 We use $\mitbf{G}$ to represent the $n \times 6m$ matrix of corresponding basis functions evaluated at each point in $\mitbf{P}$.
-We can then view $\mitbf{d}^*_i$ as a realization of the random vector
+We can write $\mitbf{d}_i$ as
 \begin{equation}
@@
 However, we are inherently assuming that the principle components of transient displacements and the noise processes are aligned with the cardinal directions, which is admittedly an arbitrary assumption.
-We form each component of the posterior, denoted as $\hat{u}_i$, by conditioning $u_i$ with $\mitbf{d}^*_i$.
+We then form each component of the posterior, denoted as $\hat{u}_i$, by conditioning $u_i$ with $\mitbf{d}^*_i$.
 That is to say $\hat{u}_i = u_i | \mitbf{d}^*_i$.
 If $\kappa$ is kept finite then, from \citet[sec.
-2.2]{Rasmussen2006}, we find that $\hat{u}$ is distributed as $\mathcal{N}(\mu_{\hat{u}},C_{\hat{u}})$, where
+2.2]{Rasmussen2006}, we find that $\hat{u}_i$ is distributed as $\mathcal{N}(\mu_{\hat{u}_i},C_{\hat{u}_i})$, where
 \begin{equation}\label{eq:PosteriorMean}
-\mu_{\hat{u}}(p) = C_u(p,\mitbf{P})\left(\mitbf{\Sigma} + \kappa^2\mitbf{G}\mitbf{G}^T\right)^{-1}\mitbf{d}_*
+\mu_{\hat{u}_i}(p) = C_{u_i}(p,\mitbf{P})\left(\mitbf{\Sigma} + \kappa^2\mitbf{G}\mitbf{G}^T\right)^{-1}\mitbf{d}_*
 \end{equation}    
@@
 \begin{equation}\label{eq:PosteriorCov}
-C_{\hat{u}}(p,p') = C_u(p,p') - C_u(p,\mitbf{P})\left(\mitbf{\Sigma} + \kappa^2\mitbf{G}\mitbf{G}^T\right)^{-1}C_u(\mitbf{P},p').
+C_{\hat{u}_i}(p,p') = C_{u_i}(p,p') - C_{u_i}(p,\mitbf{P})\left(\mitbf{\Sigma} + \kappa^2\mitbf{G}\mitbf{G}^T\right)^{-1}C_{u_i}(\mitbf{P},p').
 \end{equation}
@@
  \begin{equation}\label{eq:PosteriorMean2}
-\mu_{\hat{u}}(p) = \left[\begin{array}{cc}
-                         C_u(p,\mitbf{P}) & \mitbf{0} \\
+\mu_{\hat{u}_i}(p) = \left[\begin{array}{cc}
+                         C_{u_i}(p,\mitbf{P}) & \mitbf{0} \\
                          \end{array}\right]
@@
 \begin{equation}\label{eq:PosteriorCov2}
-C_{\hat{u}}(p,p') = C_u(p,p') - 
+C_{\hat{u}_i}(p,p') = C_{u_i}(p,p') - 
                     \left[\begin{array}{cc}
-                          C_u(p,\mitbf{P}) & \mitbf{0} \\
+                          C_{u_i}(p,\mitbf{P}) & \mitbf{0} \\
                           \end{array}\right]
@@
                     \left[\begin{array}{c}
-                          C_u(\mitbf{P},p') \\
+                          C_{u_i}(\mitbf{P},p') \\
                           \mitbf{0} \\
@@
 \end{equation}
-Taking the limit as $\kappa \to \infty$, we get the solution for the mean and covariance of $\hat{u}$,
+Taking the limit as $\kappa \to \infty$, we get the solution for the mean and covariance of $\hat{u}_i$,
  \begin{equation}\label{eq:PosteriorMean3}
-\mu_{\hat{u}}(p) = \left[\begin{array}{cc}
-                         C_u(p,\mitbf{P}) & \mitbf{0} \\
+\mu_{\hat{u}_i}(p) = \left[\begin{array}{cc}
+                         C_{u_i}(p,\mitbf{P}) & \mitbf{0} \\
                          \end{array}\right]
@@
 \begin{equation}\label{eq:PosteriorCov3}
-C_{\hat{u}}(p,p') = C_u(p,p') - 
+C_{\hat{u}_i}(p,p') = C_{u_i}(p,p') - 
                     \left[\begin{array}{cc}
-                          C_u(p,\mitbf{P}) & \mitbf{0} \\
+                          C_{u_i}(p,\mitbf{P}) & \mitbf{0} \\
                           \end{array}\right]
@@
 
-We use eq.
-(\ref{eq:PosteriorMean3}) and (\ref{eq:PosteriorCov3}) to find the posterior easting and northing components of transient displacements.
-Using $\hat{u}_i$ to denote the posterior transient displacements along direction $i$ and $x_i$ to represent the components of $\vec{x}$, we can write the components of $\dot\varepsilon$ as
+Now that we have a posterior estimate of transient displacements, we can write the components of $\dot\varepsilon$ as 
 \begin{equation}\label{eq:StrainRate}
@@
 We propose using a signal-to-noise ratio, SNR, based on $\dot\varepsilon$ for our detection criterion.
-The Frobenius norm of $\dot\varepsilon$, $||\dot\varepsilon||_\mathrm{F}$, which is sometimes referred to as the second invariant of strain rate in the geodetic literature, is often used as a metric for the strain rate ``magnitude''.
+The Frobenius norm of $\dot\varepsilon$, $||\dot\varepsilon||_\mathrm{F} = \left(\dot\varepsilon_{\mathrm{ee}}^2 + \dot\varepsilon_{\mathrm{nn}}^2 + 2\dot\varepsilon_{\mathrm{en}}^2\right)^{\frac{1}{2}}$, which is sometimes referred to as the second invariant of strain rate in the geodetic literature, is often used as a metric for the strain rate ``magnitude''.
 Noting that $||\dot\varepsilon||_\mathrm{F}$ is a random variable, SNR can be taken as the ratio of the estimated mean and standard deviation of $||\dot\varepsilon||_\mathrm{F}$.
@@
 \end{equation}
-where the subscripts ``n'' and ``e'' denote north and east, respectively.
-For simplicity, we have ignored covariances between the transient strain rate components in eq.
-(\ref{eq:SNR}), even though they are non-zero.
+For simplicity, we have ignored covariances between the transient strain rate components in eq. (\ref{eq:SNR}), even though they are non-zero.
 

### 2017-09-24 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 \begin{equation}\label{eq:PosteriorMean}
-\mu_{\hat{u}_i}(p) = C_{u_i}(p,\mitbf{P})\left(\mitbf{\Sigma} + \kappa^2\mitbf{G}\mitbf{G}^T\right)^{-1}\mitbf{d}_*
+\mu_{\hat{u}_i}(p) = C_{u_i}(p,\mitbf{P})\left(\mitbf{\Sigma} + \kappa^2\mitbf{G}\mitbf{G}^T\right)^{-1}\mitbf{d}^*_i
 \end{equation}    
@@
                    \left[\begin{array}{c}
-                         \mitbf{d}_* \\
+                         \mitbf{d}^*_i \\
                          \mitbf{0} \\
@@
 \end{equation}
+To ensure that the inverse matrices in eqs.
+(\ref{eq:PosteriorMean3}) and (\ref{eq:PosteriorCov3}) exist, each column in $\mitbf{G}$ must be linearly independent and there must be at least as many observations as basis functions ($n \geq 6m$).
 
@@
 \end{equation}
-The transient strain rate components are stochastic processes whose distributions can be interpreted as the distributions of samples of $\hat{u}$ that have been differentiated by the operator from eq.
+The components of $\dot\varepsilon$ are stochastic processes that are obtained from differentiating, adding, and scaling the Gaussian processes $\hat{u}_\mathrm{e}$ and $\hat{u}_\mathrm{n}$.
+To give a more intuitive interpretation of $\dot\varepsilon_{ij}$, we can consider realizations of $\dot{\varepsilon}_{ij}$ to be realizations of $\hat{u}_i$ and $\hat{u}_j$ that have undergone the operations from eq.
 (\ref{eq:StrainRate}).
-Since differentiation is a linear operation, the transient strain rate components are Gaussian processes.
+Since the operations from eq.
+(\ref{eq:StrainRate}) are linear, we know that $\dot\varepsilon_{ij}$ is also a Gaussian process.
 From \citet[sec.
-10.2]{Papoulis1991}, we find that the components of $\dot{\varepsilon}$ have mean functions
+10.2]{Papoulis1991}, we find that $\dot{\varepsilon}_{ij}$ has a mean function
 \begin{equation}\label{eq:StrainMean}
@@
 \end{equation} 
-and covariance functions
+and covariance function
 \begin{equation}\label{eq:StrainCov}

### 2017-09-25 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 However, we are inherently assuming that the principle components of transient displacements and the noise processes are aligned with the cardinal directions, which is admittedly an arbitrary assumption.
-We then form each component of the posterior, denoted as $\hat{u}_i$, by conditioning $u_i$ with $\mitbf{d}^*_i$.
-That is to say $\hat{u}_i = u_i | \mitbf{d}^*_i$.
+We then form each component of the posterior, denoted as $\hat{u}_i$, by updating $u_i$ with the fact that $\mitbf{d}^*_i$ has been drawn from the random vector $\mitbf{d}_i$.
+That is to say $\hat{u}_i = u_i |\mitbf{d}^*_i$.
 If $\kappa$ is kept finite then, from \citet[sec.
@@
 To ensure that the inverse matrices in eqs.
-(\ref{eq:PosteriorMean3}) and (\ref{eq:PosteriorCov3}) exist, each column in $\mitbf{G}$ must be linearly independent and there must be at least as many observations as basis functions ($n \geq 6m$).
+(\ref{eq:PosteriorMean3}) and (\ref{eq:PosteriorCov3}) exist, each column in $\mitbf{G}$ must be linearly independent.
+This condition tends to be violated when there are too few observations at a station.
+In that case, a singular value decomposition can be used to remove linearly dependent components from $\mitbf{G}$.
 
@@
 \end{equation}
-The components of $\dot\varepsilon$ are stochastic processes that are obtained from differentiating, adding, and scaling the Gaussian processes $\hat{u}_\mathrm{e}$ and $\hat{u}_\mathrm{n}$.
-To give a more intuitive interpretation of $\dot\varepsilon_{ij}$, we can consider realizations of $\dot{\varepsilon}_{ij}$ to be realizations of $\hat{u}_i$ and $\hat{u}_j$ that have undergone the operations from eq.
+The components of $\dot\varepsilon$ are stochastic processes, and we can consider realizations of $\dot{\varepsilon}_{ij}$ to be realizations of $\hat{u}_i$ and $\hat{u}_j$ that have undergone the operations from eq.
 (\ref{eq:StrainRate}).

### 2017-09-25 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 4.2.4]{Rasmussen2006}.
+In order to compute transient strain rates from transient displacements, it is also necessary that the derivatives $\partial^2 X_i(\vec{x},\vec{x}')/ \partial x_j \partial x_j'$ and $\partial^2 T_i(t,t') / \partial t \partial t'$ exist everywhere.
 The appropriate choice for $X_i$ and $T_i$ may vary depending on the geophysical signal we are trying to describe, and we discuss this matter in Section \ref{sec:SignalModel}.
@@
 
-Now that we have a posterior estimate of transient displacements, we can write the components of $\dot\varepsilon$ as 
+We now have a posterior estimate of transient displacements that can be evaluated at any $p$. We can write the components of $\dot\varepsilon$ as 
 \begin{equation}\label{eq:StrainRate}

### 2017-09-26 RevisedSubmission/hines_hetland_2017_gji.tex
did more

--- 2017-09-25 RevisedSubmission/hines_hetland_2017_gji.tex
@@
 4.2.4]{Rasmussen2006}.
+We also require $u_i$ to be spatially and temporally differentiable (See \citet[sec.
+2.2]{Adler1981} for a formal definition of differentiation for Gaussian processes).
+This requirement is satisfied if the derivatives $\partial^2 X_i(\vec{x},\vec{x}')/ \partial x_j \partial x_j'$ and $\partial^2 T_i(t,t') / \partial t \partial t'$ exist everywhere.
 The appropriate choice for $X_i$ and $T_i$ may vary depending on the geophysical signal we are trying to describe, and we discuss this matter in Section \ref{sec:SignalModel}.
@@
 
-Now that we have a posterior estimate of transient displacements, we can write the components of $\dot\varepsilon$ as 
+The posterior transient displacements can be evaluated at any $p$.
+Furthermore, $\hat{u}_i$ is spatially and temporally differentiable, allowing us to formulate $\dot\varepsilon$, which can be evaluated at any $p$.
+The components of $\dot\varepsilon$ can be written as
 \begin{equation}\label{eq:StrainRate}
@@
 \end{equation}
-The components of $\dot\varepsilon$ are stochastic processes, and we can consider realizations of $\dot{\varepsilon}_{ij}$ to be realizations of $\hat{u}_i$ and $\hat{u}_j$ that have undergone the operations from eq.
-(\ref{eq:StrainRate}).
-Since the operations from eq.
-(\ref{eq:StrainRate}) are linear, we know that $\dot\varepsilon_{ij}$ is also a Gaussian process.
-From \citet[sec.
-10.2]{Papoulis1991}, we find that $\dot{\varepsilon}_{ij}$ has a mean function
+Since eq.
+(\ref{eq:StrainRate}) is a linear operation on the Gaussian processes $u_i$ and $u_j$, we know that $\dot\varepsilon_{ij}$ is also a Gaussian process.
+The mean and covariance functions for $\dot{\varepsilon}_{ij}$ are given by
 \begin{equation}\label{eq:StrainMean}
@@
 \end{equation} 
-and covariance function
+and
 \begin{equation}\label{eq:StrainCov}
@@
                                    \frac{\partial^2 C_{\hat{u}_i}(p,p')}{\partial x_j \, \partial x'_j} + 
-                                   \frac{\partial^2 C_{\hat{u}_j}(p,p')}{\partial x_i \, \partial x'_i} \right).
+                                   \frac{\partial^2 C_{\hat{u}_j}(p,p')}{\partial x_i \, \partial x'_i} \right),
 \end{equation} 
+respectively.
 

### 2017-09-26 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 4.2.4]{Rasmussen2006}.
-We also require $u_i$ to be spatially and temporally differentiable (See \citet[sec.
+We also require that the derivatives $\partial^2 X_i(\vec{x},\vec{x}')/ \partial x_j \partial x_j'$ and $\partial^2 T_i(t,t') / \partial t \partial t'$ exist everywhere, which ensures that $u_i$ is spatially and temporally differentiable (See \citet[sec.
 2.2]{Adler1981} for a formal definition of differentiation for Gaussian processes).
-This requirement is satisfied if the derivatives $\partial^2 X_i(\vec{x},\vec{x}')/ \partial x_j \partial x_j'$ and $\partial^2 T_i(t,t') / \partial t \partial t'$ exist everywhere.
-The appropriate choice for $X_i$ and $T_i$ may vary depending on the geophysical signal we are trying to describe, and we discuss this matter in Section \ref{sec:SignalModel}.
+The appropriate choice for $X_i$ and $T_i$ may vary depending on the geophysical signal we are trying to describe.
+To keep this section sufficiently general, we hold off on specifying $X_i$ and $T_i$ until Section \ref{sec:SignalModel}, where we estimate transient strain from slow slip events in the Pacific Northwest.
 
@@
 However, we are inherently assuming that the principle components of transient displacements and the noise processes are aligned with the cardinal directions, which is admittedly an arbitrary assumption.
-We then form each component of the posterior, denoted as $\hat{u}_i$, by updating $u_i$ with the fact that $\mitbf{d}^*_i$ has been drawn from the random vector $\mitbf{d}_i$.
+We then form each component of the posterior, denoted as $\hat{u}_i$, by updating $u_i$ with the fact that $\mitbf{d}^*_i$ has been sampled from the random vector $\mitbf{d}_i$.
 That is to say $\hat{u}_i = u_i |\mitbf{d}^*_i$.
@@
 
-The posterior transient displacements can be evaluated at any $p$.
-Furthermore, $\hat{u}_i$ is spatially and temporally differentiable, allowing us to formulate $\dot\varepsilon$, which can be evaluated at any $p$.
+We can use eqs.
+(\ref{eq:PosteriorMean3}) and (\ref{eq:PosteriorCov3}) to evaluate $\hat{u}_i$ at any $p$.
+Furthermore, since $\hat{u}_i$ is spatially and temporally differentiable, we can use $\hat{u}_i$ to formulate $\dot\varepsilon$, which can also be evaluated at any $p$.
 The components of $\dot\varepsilon$ can be written as

### 2017-09-28 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 GPR is a Bayesian, non-parametric method for inferring a continuous signal from scattered data.
-Since GNSS stations are irregularly spaced and observation times may differ between stations, GPR is an ideal tool for synthesizing GNSS data into a spatially and temporally continuous representation of surface deformation.
+Since GNSS stations are irregularly spaced and observation times may differ between stations, GPR is an ideal tool for synthesizing discrete GNSS data into a spatially and temporally continuous representation of surface deformation.
 GPR is closely related to kriging \citep{Cressie1992} and least squares collocation \citep{Moritz1978}.
@@
 We determine $\dot\varepsilon$ by spatially and temporally differentiating estimates of transient displacements, $\vec{u}(p)$.
-Transient displacements have an easting and northing component, $\vec{u} = (u_\mathrm{e}, u_\mathrm{n})$, and we make a prior assumption that each component of $\vec{u}$ is a Gaussian process with zero mean and a generic covariance function $\mathrm{cov}\left(u_i(p),u_i(p')\right) = C_{u_i}(p,p')$.
+Transient displacements have an easting and northing component, $\vec{u} = (u_\mathrm{e}, u_\mathrm{n})$, and we make a prior assumption that each component of $\vec{u}$ is a Gaussian process with zero mean and a generic covariance function $\mathrm{Cov}\left(u_i(p),u_i(p')\right) = C_{u_i}(p,p')$.
 Using a more concise notation, we describe each component of $\vec{u}$ as $u_i \sim \mathcal{N}\left(0,C_{u_i}\right)$.
@@
 For example, $\eta_i$ can consist of uncorrelated white noise, temporally correlated noise describing benchmark wobble \citep[e.g.,][]{Wyatt1982,Wyatt1989}, and/or spatially correlated noise describing common mode error \citep[e.g.,][]{Wdowinski1997}.
-For now, we only make the assumption that each component of $\eta_i$ is a Gaussian process distributed as $\eta_i \sim \mathcal{N}(0,C_{\eta_i})$, and $C_{\eta_i}$ is a spatially and temporally dependent covariance function.
+For now, we only make the assumption that $\eta_i$ is a Gaussian process distributed as $\eta_i \sim \mathcal{N}(0,C_{\eta_i})$, and $C_{\eta_i}$ is a spatially and temporally dependent covariance function.
 We consider the coefficients in eq.
@@
 
-We can use eqs.
+The posterior transient displacements are spatially and temporally continuous, and we can use eqs.
 (\ref{eq:PosteriorMean3}) and (\ref{eq:PosteriorCov3}) to evaluate $\hat{u}_i$ at any $p$.
-Furthermore, since $\hat{u}_i$ is spatially and temporally differentiable, we can use $\hat{u}_i$ to formulate $\dot\varepsilon$, which can also be evaluated at any $p$.
+Furthermore, $\hat{u}_i$ is spatially and temporally differentiable, allowing us to formulate $\dot\varepsilon$ at any $p$ that we may be interested in.
 The components of $\dot\varepsilon$ can be written as
@@
 However, if we want to detect geophysical phenomena automatically, then we need to define a detection criterion.
-We propose using a signal-to-noise ratio, SNR, based on $\dot\varepsilon$ for our detection criterion.
-The Frobenius norm of $\dot\varepsilon$, $||\dot\varepsilon||_\mathrm{F} = \left(\dot\varepsilon_{\mathrm{ee}}^2 + \dot\varepsilon_{\mathrm{nn}}^2 + 2\dot\varepsilon_{\mathrm{en}}^2\right)^{\frac{1}{2}}$, which is sometimes referred to as the second invariant of strain rate in the geodetic literature, is often used as a metric for the strain rate ``magnitude''.
-Noting that $||\dot\varepsilon||_\mathrm{F}$ is a random variable, SNR can be taken as the ratio of the estimated mean and standard deviation of $||\dot\varepsilon||_\mathrm{F}$.
+We use a signal-to-noise ratio, SNR, that is based on the Frobenius norm of $\dot\varepsilon$, $||\dot\varepsilon||_\mathrm{F} = \left(\dot\varepsilon_{\mathrm{ee}}^2 + \dot\varepsilon_{\mathrm{nn}}^2 + 2\dot\varepsilon_{\mathrm{en}}^2\right)^{\frac{1}{2}}$, for our detection criterion.
+In the geodetic literature, $||\dot{\varepsilon}||_\mathrm{F}$ is often used as a metric for the strain rate ``magnitude", and it is sometimes referred to as the second invariant of strain rate.
+Noting that $||\dot\varepsilon||_\mathrm{F}$ is a random variable, we take SNR to be the ratio of the estimated mean and standard deviation of $||\dot\varepsilon||_\mathrm{F}$.
+An estimate of the mean is found by evaluating $||\dot\varepsilon||_\mathrm{F}$ at the mean of $\dot\varepsilon$,
+\begin{align}
+\mu_{||\dot\varepsilon||_\mathrm{F}} &\approx \left. ||\dot\varepsilon||_F \right|_{\dot\varepsilon = \mu_{\dot\varepsilon}} \\
+                                     &= \left(\mu_{\dot\varepsilon_\mathrm{nn}}^2 + 
+                                              \mu_{\dot\varepsilon_\mathrm{ee}}^2 + 
+                                              2\mu_{\dot\varepsilon_\mathrm{en}}^2
+                                     \right)^\frac{1}{2},
+\end{align}
+and we use nonlinear uncertainty propagation to estimate the standard deviation of $||\dot\varepsilon||_\mathrm{F}$,
+\begin{equation}
+\sigma_{||\dot\varepsilon||_\mathrm{F}} \approx
+\left(\left(\left. \frac{\partial ||\dot\varepsilon||_\mathrm{F}}{\partial \dot\varepsilon_\mathrm{nn}}\right|_{\dot\varepsilon = \mu_{\dot\varepsilon}} \right)^2 \sigma^2_{\varepsilon_\mathrm{nn}} +
+\left(\left. \frac{\partial ||\dot\varepsilon||_\mathrm{F}}{\partial \dot\varepsilon_\mathrm{ee}}\right|_{\dot\varepsilon = \mu_{\dot\varepsilon}} \right)^2 \sigma^2_{\varepsilon_\mathrm{ee}} +
+\left(\left. \frac{\partial ||\dot\varepsilon||_\mathrm{F}}{\partial \dot\varepsilon_\mathrm{en}}\right|_{\dot\varepsilon = \mu_{\dot\varepsilon}} \right)^2 \sigma^2_{\varepsilon_\mathrm{en}}
+\right)^\frac{1}{2},
+\end{equation}
+where $\sigma^2_{\varepsilon_{ij}}(p) = C_{\varepsilon_{ij}}(p,p)$. 
+ by evaluating it at $\dot\varepsilon_{ij} = \mu_{\dot\varepsilon_{ij}}$.
+point $p$ is approximately $(\mu_{\dot\varepsilon_\mathrm{nn}}(p) + \mu_{\dot\varepsilon_\mathrm{ee}}(p) + 2\mu_{\dot\varepsilon_\mathrm{en}}(p))$.
+Using nonlinear uncertainty propagation, the standard deviation of $||\dot\varepsilon||$ is approximated as
+
+can be found with nonlinear uncertainty propagation,
+
  Using nonlinear uncertainty propagation, we find SNR to be  

### 2017-09-28 RevisedSubmission/hines_hetland_2017_gji.tex
did more shit

@@
 Since eq.
-(\ref{eq:StrainRate}) is a linear operation on the Gaussian processes $u_i$ and $u_j$, we know that $\dot\varepsilon_{ij}$ is also a Gaussian process.
+(\ref{eq:StrainRate}) is a linear operation on the Gaussian processes $\hat{u}_i$ and $\hat{u}_j$, we know that $\dot\varepsilon_{ij}$ is also a Gaussian process.
 The mean and covariance functions for $\dot{\varepsilon}_{ij}$ are given by
@@
 respectively.
-
+\subsection{Transient detection criterion}\label{sec:TransientDetection}
 Our motivation for estimating transient strain rates is, in part, to detect geophysical phenomena.
-As we will see, geophysical signal can be easily identified by visually inspecting the solution for transient strain rates.
+As we will see, geophysical signal can be easily identified by visually inspecting the solution for $\dot\varepsilon$ from eqs.
+(\ref{eq:StrainMean}) and (\ref{eq:StrainCov}).
 However, if we want to detect geophysical phenomena automatically, then we need to define a detection criterion.
@@
 \begin{align}
-\mu_{||\dot\varepsilon||_\mathrm{F}} &\approx \left. ||\dot\varepsilon||_F \right|_{\dot\varepsilon = \mu_{\dot\varepsilon}} \\
-                                     &= \left(\mu_{\dot\varepsilon_\mathrm{nn}}^2 + 
-                                              \mu_{\dot\varepsilon_\mathrm{ee}}^2 + 
+\mu_{||\dot\varepsilon||_\mathrm{F}} &\approx ||\dot\varepsilon||_F \big|_{\dot\varepsilon = \mu_{\dot\varepsilon}} \\
+                                     &= \left(\mu_{\dot\varepsilon_\mathrm{ee}}^2 + 
+                                              \mu_{\dot\varepsilon_\mathrm{nn}}^2 + 
                                               2\mu_{\dot\varepsilon_\mathrm{en}}^2
@@
 \end{align}
-and we use nonlinear uncertainty propagation to estimate the standard deviation of $||\dot\varepsilon||_\mathrm{F}$,
+and we use nonlinear uncertainty propagation to estimate the standard deviation,
 \begin{equation}
 \sigma_{||\dot\varepsilon||_\mathrm{F}} \approx
-\left(\left(\left. \frac{\partial ||\dot\varepsilon||_\mathrm{F}}{\partial \dot\varepsilon_\mathrm{nn}}\right|_{\dot\varepsilon = \mu_{\dot\varepsilon}} \right)^2 \sigma^2_{\varepsilon_\mathrm{nn}} +
-\left(\left. \frac{\partial ||\dot\varepsilon||_\mathrm{F}}{\partial \dot\varepsilon_\mathrm{ee}}\right|_{\dot\varepsilon = \mu_{\dot\varepsilon}} \right)^2 \sigma^2_{\varepsilon_\mathrm{ee}} +
-\left(\left. \frac{\partial ||\dot\varepsilon||_\mathrm{F}}{\partial \dot\varepsilon_\mathrm{en}}\right|_{\dot\varepsilon = \mu_{\dot\varepsilon}} \right)^2 \sigma^2_{\varepsilon_\mathrm{en}}
+\left(\left(\left.
+\frac{\partial ||\dot\varepsilon||_\mathrm{F}}{\partial \dot\varepsilon_\mathrm{ee}}\right|_{\dot\varepsilon = \mu_{\dot\varepsilon}} \right)^2 \sigma^2_{\dot\varepsilon_\mathrm{ee}} +
+\left(\left. \frac{\partial ||\dot\varepsilon||_\mathrm{F}}{\partial \dot\varepsilon_\mathrm{nn}}\right|_{\dot\varepsilon = \mu_{\dot\varepsilon}} \right)^2 \sigma^2_{\dot\varepsilon_\mathrm{nn}} +
+\left(\left. \frac{\partial ||\dot\varepsilon||_\mathrm{F}}{\partial \dot\varepsilon_\mathrm{en}}\right|_{\dot\varepsilon = \mu_{\dot\varepsilon}} \right)^2 \sigma^2_{\dot\varepsilon_\mathrm{en}}
 \right)^\frac{1}{2},
 \end{equation}
-where $\sigma^2_{\varepsilon_{ij}}(p) = C_{\varepsilon_{ij}}(p,p)$. 
- by evaluating it at $\dot\varepsilon_{ij} = \mu_{\dot\varepsilon_{ij}}$.
-point $p$ is approximately $(\mu_{\dot\varepsilon_\mathrm{nn}}(p) + \mu_{\dot\varepsilon_\mathrm{ee}}(p) + 2\mu_{\dot\varepsilon_\mathrm{en}}(p))$.
-Using nonlinear uncertainty propagation, the standard deviation of $||\dot\varepsilon||$ is approximated as
+where $\sigma^2_{\dot\varepsilon_{ij}}(p) = C_{\dot\varepsilon_{ij}}(p,p)$. After some calculations, we find SNR to be
+\begin{align}\label{eq:SNR}
+\mathrm{SNR}(p) &= \frac{\mu_{||\dot\varepsilon||_\mathrm{F}}(p)}{\sigma_{||\dot\varepsilon||_\mathrm{F}}(p)} \\
+                &= \frac{\mu_{\dot\varepsilon_\mathrm{ee}}(p)^2 +
+                        \mu_{\dot\varepsilon_\mathrm{nn}}(p)^2 +
+                        2\mu_{\dot\varepsilon_\mathrm{en}}(p)^2}
+                       {\big(\sigma^2_{\dot\varepsilon_\mathrm{ee}}(p)\mu_{\dot\varepsilon_\mathrm{ee}}(p)^2 + 
+                             \sigma^2_{\dot\varepsilon_\mathrm{nn}}(p)\mu_{\dot\varepsilon_\mathrm{nn}}(p)^2 + 
+                              4\sigma^2_{\dot\varepsilon_\mathrm{en}}(p)\mu_{\dot\varepsilon_\mathrm{en}}(p)^2
+                        \big)^{\frac{1}{2}}}
+.
+\end{align}
+We explicitly show that SNR is a function of $p$ to emphasize that it identifies the position and time of anomalous deformation.
+We can reasonably suspect that some transient geophysical phenomena is occurring wherever and whenever SNR is larger than ${\sim}3$.
 
-can be found with nonlinear uncertainty propagation,
-
- Using nonlinear uncertainty propagation, we find SNR to be  
-\begin{equation}\label{eq:SNR}
-\mathrm{SNR}(p) = \frac{\mu_{\dot\varepsilon_\mathrm{nn}}(p)^2 +
-                        \mu_{\dot\varepsilon_\mathrm{ee}}(p)^2 +
-                        2\mu_{\dot\varepsilon_\mathrm{en}}(p)^2}
-                       {\big(C_{\dot\varepsilon_\mathrm{nn}}(p,p)\mu_{\dot\varepsilon_\mathrm{nn}}(p)^2 + 
-                              C_{\dot\varepsilon_\mathrm{ee}}(p,p)\mu_{\dot\varepsilon_\mathrm{ee}}(p)^2 + 
-                              4C_{\dot\varepsilon_\mathrm{en}}(p,p)\mu_{\dot\varepsilon_\mathrm{en}}(p)^2
-                        \big)^{\frac{1}{2}}}
-,
-\end{equation}
-For simplicity, we have ignored covariances between the transient strain rate components in eq. (\ref{eq:SNR}), even though they are non-zero.
-
-\section{Outlier detection}\label{sec:Outlier}
+\subsection{Outlier detection}\label{sec:Outlier}
 In our formulation for estimating transient strain rates, we have assumed that noise in the data vector is normally distributed.

### 2017-09-30 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 where $a_i^{(j)}$ is an offset that is unique for each station, $b_i^{(j)}$ is the secular velocity at $\vec{x}^{(j)}$, and the sinusoids describe seasonal deformation (using units of years for $t$).
-We use $\eta_i$ to represent all the stochastic noise processes.
+We use $\eta_i$ to represent all the non-parametric noise processes.
 For example, $\eta_i$ can consist of uncorrelated white noise, temporally correlated noise describing benchmark wobble \citep[e.g.,][]{Wyatt1982,Wyatt1989}, and/or spatially correlated noise describing common mode error \citep[e.g.,][]{Wdowinski1997}.
@@
 We can write $\mitbf{d}_i$ as
-\begin{equation}
+\begin{equation}\label{eq:DataVec}
 \mitbf{d}_i = u_i(\mitbf{P}) + \eta_i(\mitbf{P}) + \mitbf{G}\mitbf{m}_i,
@@
 However, we are inherently assuming that the principle components of transient displacements and the noise processes are aligned with the cardinal directions, which is admittedly an arbitrary assumption.
-We then form each component of the posterior, denoted as $\hat{u}_i$, by updating $u_i$ with the fact that $\mitbf{d}^*_i$ has been sampled from the random vector $\mitbf{d}_i$.
-That is to say $\hat{u}_i = u_i |\mitbf{d}^*_i$.
+We then form each component of the posterior, denoted as $\hat{u}_i$, by conditioning $u_i$ with the fact that $\mitbf{d}^*_i$ has been realized from the random vector $\mitbf{d}_i$.
+That is to say $\hat{u}_i = u_i |(\mitbf{d}_i = \mitbf{d}^*_i)$.
 If $\kappa$ is kept finite then, from \citet[sec.
@@
 \subsection{Outlier detection}\label{sec:Outlier}
-In our formulation for estimating transient strain rates, we have assumed that noise in the data vector is normally distributed.
+In deriving our formulation for transient strain rates, we have assumed that noise in the data vector is normally distributed.
 This is not an appropriate assumption for GNSS data, which often have more outliers than would be predicted for normally distributed noise.
+Methods for analyzing GNSS data should either be robust against outliers or should involve a preprocessing step in which outliers are detected and removed.
+Examples of the former include the MIDAS method for estimating secular velocities \citep{Blewitt2016} and the GPS Imaging method for detecting spatially coherent features \citep{Hammond2016}.
+We adopt the latter approach and remove outliers in the GNSS data prior to estimating $\dot\varepsilon$.
+We identify outliers based on the residuals between the observations and a best fitting model.
+Observations with residuals that exceed some threshold are removed.
+This strategy is commonly used to detect outliers in GNSS data, where the residuals are typically computed for each station by estimating and removing a temporally dependent model consisting of a linear trend and seasonal terms \citep[e.g.,][]{Johansson2002,Dong2006,Bos2013}.
+For the outlier detection method used in this study, the residuals are determined by removing a best fitting spatially and temporally dependent model from the GNSS data.
+That model that we want to fit to the $n$ GNSS observations, $\mitbf{d}^*$, consists of the linear trend and sinusoids from eq.
+(\ref{eq:DataVec}), $\mitbf{G}\mitbf{m}$, and a Gaussian process, $v(\mitbf{P})$, that is used to describe any spatially and temporally correlated features in the GNSS data.
+We now consider $\mitbf{d}^*$ to be a realization of the random vector
+\begin{equation}
+\mitbf{d} = \mitbf{G}\mitbf{m} + v(\mitbf{P}) + \mitbf{w},
+\end{equation}   
+where $\mitbf{w}$ is a vector of uncorrelated Gaussian noise distributed as $\mitbf{w} = \mathcal{N}\left(0,\mathrm{diag}(\mitbf{\sigma}^2)\right)$, and $\mitbf{\sigma} = [\sigma_1, \sigma_2, ..., \sigma_n]$ are the data uncertainties.
+The residual vector, $\mitbf{r} = [r_1,r_2,...,r_n]$, is the difference between $\mitbf{d}^*$, and the mean of the $\mitbf{G}\mitbf{m} + v(\mitbf{P})$ conditioned on $\mitbf{d} = \mitbf{d}^*$.
+
+This best fitting model is found using Gaussian process regression, where the model consists of a parametric and
+
+We again use Gaussian process regression to find the best fitting model, which consists of parametric and non-parametric
+
+We now describe column vector of $n$ GNSS observations $\mitbf{d}*$ to be a
+
+With this approach, outliers are considered to by any observation that is spatially or temporally anomalous.
+
+approach for identifying outliers differs
+
+is involving fitting a model to , in principle, similar to
+
+Methods for analyzing GNSS data should either be robust against outliers or a outliers , such as the MIDAS method for estimating secular velocities \citep{Blewitt2016}, or
+
 It follows that proposed methods for analyzing GNSS data should be robust against outliers \citep[e.g.,][]{Blewitt2016}.

### 2017-10-02 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 A Gaussian process is a normally distributed stochastic process that is fully defined in terms of a mean function and a positive definite covariance function.
-For example, Brownian motion, $B(t)$, is a well known Gaussian process in $\mathbb{R}^1$ which has zero mean and covariance function $\mathrm{cov}(B(t),B(t')) = \min(t,t')$, where $t,t' \ge 0$.
+For example, Brownian motion, $B(t)$, is a well known Gaussian process in $\mathbb{R}^1$ which has zero mean and covariance function $\mathrm{Cov}\left[B(t),B(t')\right] = \min(t,t')$, where $t,t' \ge 0$.
 If no prior information is available for the geophysical signal, then maximum likelihood methods can be used to objectively choose a prior Gaussian process that is most consistent with the observations.
@@
 We determine $\dot\varepsilon$ by spatially and temporally differentiating estimates of transient displacements, $\vec{u}(p)$.
-Transient displacements have an easting and northing component, $\vec{u} = (u_\mathrm{e}, u_\mathrm{n})$, and we make a prior assumption that each component of $\vec{u}$ is a Gaussian process with zero mean and a generic covariance function $\mathrm{Cov}\left(u_i(p),u_i(p')\right) = C_{u_i}(p,p')$.
+Transient displacements have an easting and northing component, $\vec{u} = (u_\mathrm{e}, u_\mathrm{n})$, and we make a prior assumption that each component of $\vec{u}$ is a Gaussian process with zero mean and a generic covariance function $\mathrm{Cov}\left[u_i(p),u_i(p')\right] = C_{u_i}(p,p')$.
 Using a more concise notation, we describe each component of $\vec{u}$ as $u_i \sim \mathcal{N}\left(0,C_{u_i}\right)$.
@@
 That is to say $\hat{u}_i = u_i |(\mitbf{d}_i = \mitbf{d}^*_i)$.
-If $\kappa$ is kept finite then, from \citet[sec.
-2.2]{Rasmussen2006}, we find that $\hat{u}_i$ is distributed as $\mathcal{N}(\mu_{\hat{u}_i},C_{\hat{u}_i})$, where
-\begin{equation}\label{eq:PosteriorMean}
-\mu_{\hat{u}_i}(p) = C_{u_i}(p,\mitbf{P})\left(\mitbf{\Sigma} + \kappa^2\mitbf{G}\mitbf{G}^T\right)^{-1}\mitbf{d}^*_i
-\end{equation}    
-and
-\begin{equation}\label{eq:PosteriorCov}
-C_{\hat{u}_i}(p,p') = C_{u_i}(p,p') - C_{u_i}(p,\mitbf{P})\left(\mitbf{\Sigma} + \kappa^2\mitbf{G}\mitbf{G}^T\right)^{-1}C_{u_i}(\mitbf{P},p').
-\end{equation}
+A general solution for $\hat{u}_i$ is derived in \citet[sec.
+8.9]{VonMises1964}, where we find that $\hat{u}_i$ is distributed as $\mathcal{N}(\mu_{\hat{u}_i},C_{\hat{u}_i})$ with mean function
+\begin{align}\label{eq:PosteriorMean}
+\mu_{\hat{u}_i}(p) &= \mathrm{E}\left[u_i(p)\right] + 
+                      \mathrm{Cov}\left[u_i(p),\mitbf{d}_i\right] 
+                      \mathrm{Cov}\left[\mitbf{d}_i\right]^{-1}
+                      \left(\mitbf{d}^*_i - \mathrm{E}\left[\mitbf{d}_i\right]\right)\nonumber \\
+                   &= C_{u_i}(p,\mitbf{P})\left(\mitbf{\Sigma} + \kappa^2\mitbf{G}\mitbf{G}^T\right)^{-1}\mitbf{d}^*_i
+\end{align}    
+and covariance function
+\begin{align}\label{eq:PosteriorCov}
+C_{\hat{u}_i}(p,p') &= \mathrm{Cov}\left[u_i(p),u_i(p')\right] - 
+                       \mathrm{Cov}\left[u_i(p),\mitbf{d}_i\right] 
+                       \mathrm{Cov}\left[\mitbf{d}_i\right]^{-1}
+                       \mathrm{Cov}\left[\mitbf{d}_i,u_i(p')\right]\nonumber \\
+                    &= C_{u_i}(p,p') - C_{u_i}(p,\mitbf{P})\left(\mitbf{\Sigma} + \kappa^2\mitbf{G}\mitbf{G}^T\right)^{-1}C_{u_i}(\mitbf{P},p').
+\end{align}
 However, we are interested in the limit as $\kappa \to \infty$, and the form for eq.
@@
 \begin{align}
-\mu_{||\dot\varepsilon||_\mathrm{F}} &\approx ||\dot\varepsilon||_F \big|_{\dot\varepsilon = \mu_{\dot\varepsilon}} \\
+\mu_{||\dot\varepsilon||_\mathrm{F}} &\approx ||\dot\varepsilon||_F \big|_{\dot\varepsilon = \mu_{\dot\varepsilon}}\nonumber \\
                                      &= \left(\mu_{\dot\varepsilon_\mathrm{ee}}^2 + 
@@
 Examples of the former include the MIDAS method for estimating secular velocities \citep{Blewitt2016} and the GPS Imaging method for detecting spatially coherent features \citep{Hammond2016}.
-We adopt the latter approach and remove outliers in the GNSS data prior to estimating $\dot\varepsilon$.
-We identify outliers based on the residuals between the observations and a best fitting model.
+In this study, we identify and remove outliers as a preprocessing step before estimating $\dot\varepsilon$.
+Outliers are identified based on the residuals for a model that best fits the observed data.
 Observations with residuals that exceed some threshold are removed.
-This strategy is commonly used to detect outliers in GNSS data, where the residuals are typically computed for each station by estimating and removing a temporally dependent model consisting of a linear trend and seasonal terms \citep[e.g.,][]{Johansson2002,Dong2006,Bos2013}.
+This strategy is commonly used to detect outliers in GNSS data, where the residuals are typically computed by estimating and removing a linear trend and seasonal terms for each time series \citep[e.g.,][]{Johansson2002,Dong2006,Bos2013}.
+The outlier detection method used here is different from this more standard approach because the model we fit to the data additionally consists of a spatially and temporally correlated Gaussian process.
+With this more complex model, we are able to detect outliers that are spatially or temporally anomalous.
+
 For the outlier detection method used in this study, the residuals are determined by removing a best fitting spatially and temporally dependent model from the GNSS data.

### 2017-10-03 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 \end{align}
-However, we are interested in the limit as $\kappa \to \infty$, and the form for eq.
-(\ref{eq:PosteriorMean}) and eq.
-(\ref{eq:PosteriorCov}) is not suitable for evaluating this limit.
+However, we are interested in the limit as $\kappa \to \infty$, and the form for eqs.
+(\ref{eq:PosteriorMean}) and (\ref{eq:PosteriorCov}) is not suitable for evaluating this limit.
 We use a partitioned matrix inversion identity \citep[sec.
-2.7.4]{Press2007} to rewrite eq.
-(\ref{eq:PosteriorMean}) and eq.
-(\ref{eq:PosteriorCov}) as
+2.7.4]{Press2007} to rewrite eqs.
+(\ref{eq:PosteriorMean}) and (\ref{eq:PosteriorCov}) as
  \begin{equation}\label{eq:PosteriorMean2}
@@
 This strategy is commonly used to detect outliers in GNSS data, where the residuals are typically computed by estimating and removing a linear trend and seasonal terms for each time series \citep[e.g.,][]{Johansson2002,Dong2006,Bos2013}.
-The outlier detection method used here is different from this more standard approach because the model we fit to the data additionally consists of a spatially and temporally correlated Gaussian process.
+The outlier detection method used here is different from this more standard approach because the model we fit to the data additionally consists of a spatially and temporally correlated Gaussian process, $v \sim \mathcal{N}(0,C_v)$.
+We use $v$ to explain any coherent features in the data that cannot be explained by the linear trend and sinusoids.
 With this more complex model, we are able to detect outliers that are spatially or temporally anomalous.
 
-For the outlier detection method used in this study, the residuals are determined by removing a best fitting spatially and temporally dependent model from the GNSS data.
-That model that we want to fit to the $n$ GNSS observations, $\mitbf{d}^*$, consists of the linear trend and sinusoids from eq.
-(\ref{eq:DataVec}), $\mitbf{G}\mitbf{m}$, and a Gaussian process, $v(\mitbf{P})$, that is used to describe any spatially and temporally correlated features in the GNSS data.
+To describe our outlier detection algorithm in detail, we let $\mitbf{d}^*$ denote all $n$ GNSS displacement observations for a single component (note that we dropped the component subscripts since outliers are detected for the easting and northing components independently).
+Our goal is to find the index set of non-outliers in $\mitbf{d}^*$, which we denote as $\mitbf{\Omega}$.
+For notational ease, we use a tilde to indicate that an array only contains elements corresponding to $\mitbf{\Omega}$ (e.g., the vector of non-outlier observations is denoted $\tilde{\mitbf{d}}^* = [d_i^*]_{i \in \mitbf{\Omega}}$).
 We now consider $\mitbf{d}^*$ to be a realization of the random vector
@@
 \end{equation}
-where $\mitbf{w}$ is a vector of uncorrelated Gaussian noise distributed as $\mitbf{w} = \mathcal{N}\left(0,\mathrm{diag}(\mitbf{\sigma}^2)\right)$, and $\mitbf{\sigma} = [\sigma_1, \sigma_2, ..., \sigma_n]$ are the data uncertainties.
-The residual vector, $\mitbf{r} = [r_1,r_2,...,r_n]$, is the difference between $\mitbf{d}^*$, and the mean of the $\mitbf{G}\mitbf{m} + v(\mitbf{P})$ conditioned on $\mitbf{d} = \mitbf{d}^*$.
-
-This best fitting model is found using Gaussian process regression, where the model consists of a parametric and
-
-We again use Gaussian process regression to find the best fitting model, which consists of parametric and non-parametric
-
-We now describe column vector of $n$ GNSS observations $\mitbf{d}*$ to be a
-
-With this approach, outliers are considered to by any observation that is spatially or temporally anomalous.
-
-approach for identifying outliers differs
-
-is involving fitting a model to , in principle, similar to
-
-Methods for analyzing GNSS data should either be robust against outliers or a outliers , such as the MIDAS method for estimating secular velocities \citep{Blewitt2016}, or
-
-It follows that proposed methods for analyzing GNSS data should be robust against outliers \citep[e.g.,][]{Blewitt2016}.
-In order to make our estimates of transient strain more robust, we automatically identify and remove outliers in the GNSS data as a pre-processing step.
-
-Our method for detecting outliers is based on the data editing algorithm described in \citet{Acheson1975}.
-We identify outliers by using the residuals between the observations, $\mitbf{d}_*$, and a best fitting model.
-Data with residuals that are anomalously large are flagged as outliers.
-We treat $\mitbf{d}_*$ as a sample of $\mitbf{d}$, but we ignore any correlated noise (i.e., $\eta(p) = 0$).
-The best fitting model for $\mitbf{d}_*$ is then considered to be the expected value of the random vector $u(\mitbf{P}) + \mitbf{G}\mitbf{a}$ after conditioning it with non-outlier observations.
-We still consider $u$ to have a separable covariance function as in eq.
-(\ref{eq:TransientCovariance}), and the choice for $X$ and $T$ does not need to be the same as that used in Section \ref{sec:Method}.
-We only desire that $u$ is sufficiently flexible to describe any signal that cannot be described by the linear trend or seasonal sinusoids in $\mitbf{G}$.
-Since outliers are determined based on how well a spatially and temporally dependent model fits the data, we are able to identify anomalous observations which may not be immediately apparent from inspecting individual time series.
-
-To begin the algorithm, we let $\mitbf{\Omega}$ be the index set of non-outliers in $\mitbf{d}_*$ and initiate it with all $n$ indices.
-This algorithm is iterative, and for each iteration we calculate the residual vector
-\begin{align}\label{eq:Residual}
-\mitbf{r} &= \frac{\mitbf{d}_* - \mathrm{E}\left[(u(\mitbf{P}) + \mitbf{G}\mitbf{a})|\tilde{\mitbf{d}}_* \right]}{\mitbf{\sigma}} \\
-       &= \frac{1}{\mitbf{\sigma}}\left(\mitbf{d}_*  - 
+where $\mitbf{G}$ and $\mitbf{m}$ are the same as in eq.
+(\ref{eq:DataVec}), and $\mitbf{w}$ is a vector of uncorrelated noise with assumed known standard deviations $\mitbf{\sigma} = [\sigma_1, \sigma_2, ..., \sigma_n]$.
+Our best fitting model for $\mitbf{d}^*$ is taken to be the mean of $\mitbf{G}\mitbf{m} + v(\mitbf{P})$ after it is conditioned with non-outlier observations.
+The residual vector is then
+\begin{equation}\label{eq:Residual1}
+\mitbf{r} = \mitbf{d}^* - \mathrm{E}\left[\Big(\mitbf{G}\mitbf{m} + v(\mitbf{P})\Big) \Big| \Big(\tilde{\mitbf{d}} = \tilde{\mitbf{d}}^*\Big)\right].
+\end{equation}
+Following a similar procedure as above, we can write eq. (\ref{eq:Residual1}) as
+\begin{equation}\label{eq:Residual}
+\mitbf{r} = \mitbf{d}^*  - 
             \left[\begin{array}{cc}
-                C_u(\mitbf{P},\tilde{\mitbf{P}}) & \mitbf{G} \\
+                  C_v(\mitbf{P},\tilde{\mitbf{P}}) & \mitbf{G} \\
                   \end{array}\right]
             \left[\begin{array}{cc}
-                C_u(\tilde{\mitbf{P}},\tilde{\mitbf{P}}) + \mathrm{diag}(\tilde{\mitbf{\sigma}}^2) & \tilde{\mitbf{G}} \\
+                  C_v(\tilde{\mitbf{P}},\tilde{\mitbf{P}}) + \mathrm{diag}(\tilde{\mitbf{\sigma}}^2) & \tilde{\mitbf{G}} \\
                   \tilde{\mitbf{G}}^T  & \mitbf{0} \\
@@
             \left[\begin{array}{c}
-                \tilde{\mitbf{d}}_* \\
+                  \tilde{\mitbf{d}}^* \\
                   \mitbf{0} \\
-                \end{array}\right]\right),
-\end{align}
-where the tilde indicates that only elements corresponding to indices in $\mitbf{\Omega}$ are retained (e.g., $\tilde{\mitbf{P}} = \{P_i\}_{i\in\mitbf{\Omega}}$). We then update $\mitbf{\Omega}$ to be
+                  \end{array}\right].
+\end{equation}
+We start the algorithm by letting $\mitbf{\Omega}$ contain all $n$ indices, and we refine $\mitbf{\Omega}$ iteratively.
+For each iteration, we compute $\mitbf{r}$ with eq.
+(\ref{eq:Residual}) and then update $\mitbf{\Omega}$ to be
 \begin{equation}\label{eq:Update}
-\mitbf{\Omega} = \{i : |r_i| < \lambda \cdot \mathrm{RMS}\}, \ \ \ r_i \in \mitbf{r},
+\mitbf{\Omega} = \left\{i : \left|\frac{r_i}{\sigma_i}\right| < \lambda \cdot \mathrm{RMS}\right\},
 \end{equation} 
-where RMS is the root-mean-square of $\tilde{\mitbf{r}}$ and $\lambda$ is an outlier tolerance.
+where RMS is the root-mean-square of $\tilde{\mitbf{r}}$ and $\lambda$ is an outlier tolerance. Iterations continue until the new $\mitbf{\Omega}$ is the same as the previous $\mitbf{\Omega}$.
+
 We use $\lambda=4$ in this study, which in our experience accurately identifies outliers without unnecessarily decimating the data.
@@
 
-It should be noted that this algorithm does not identify jumps in GNSS time series, which are another common issue.
+It should be noted that this algorithm does not identify jumps in GNSS time series (cite Gazeaux 2013), which are another common issue.
 Some, but not all, jumps can be automatically removed by looking up the dates of equipment changes and earthquakes.

### 2017-10-07 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 \begin{equation}\label{eq:TransientCovariance}
-C_{u_i}\left((\vec{x},t),(\vec{x}',t')\right) = X_i(\vec{x},\vec{x}')T_i(t,t').
+C_{u_i}(p,p') = C_{u_i}\left((\vec{x},t),(\vec{x}',t')\right) = X_i(\vec{x},\vec{x}')T_i(t,t').
 \end{equation}  
@@
 2.2]{Adler1981} for a formal definition of differentiation for Gaussian processes).
+For example, the squared exponential function is a commonly used covariance function that is sufficiently differentiable for our purposes.
+We can make a potential covariance function for $u_i$ by using a squared exponential for $X$ and $T$,
+\begin{equation}\label{eq:ExampleCov}
+C_{u_i}\left(p,p'\right) = \phi^2\exp\left(\frac{-||\vec{x} - \vec{x}'||_2^2}{2\ell^2}\right)
+                           \exp\left(\frac{-|t - t'|^2}{2\tau^2}\right),
+\end{equation}
+where $\ell$ and $\tau$ describe a characteristic length-scale and time-scale for realizations of $u_i$.
+The parameter $\phi$ controls the amplitude of realizations of $u_i$.
+We refer to the parameters of a covariance function, such as $\ell$, $\tau$, and $\phi$, as hyperparameters.
+These hyperparameters can be chosen subjectively or they can be chosen objectively with maximum likelihood techniques.
+
+commonly used covariance function that is sufficiently differentiable for our purposes.
+
 The appropriate choice for $X_i$ and $T_i$ may vary depending on the geophysical signal we are trying to describe.
@@
 GNSS data records transient displacements as well as other physical and non-physical processes that we are not interested in.
-We first consider component $i$ of a single GNSS displacement observation made at station $j$, which is located at $\vec{x}^{(j)}$, and time $t$.
+We first consider component $i$ of a single GNSS displacement observation made at station $j$, which is located at $\vec{x}^{(j)}$, and time $t^{(k)}$.
 We describe this observation as a realization of the random variable
@@
 \begin{split}
-d^{(j)}_i(t) = &u_i(\vec{x}^{(j)},t) + \eta_i(\vec{x}^{(j)},t) + a_i^{(j)} + b_i^{(j)}t + c_i^{(j)}\sin(2 \pi t) + \\ 
-               &e_i^{(j)}\cos(2 \pi t) + f_i^{(j)}\sin(4 \pi t)  + g_i^{(j)}\cos(4 \pi t), 
+d_i^{(jk)} = &u_i\left(\vec{x}^{(j)},t^{(k)}\right) + 
+              \eta_i\left(\vec{x}^{(j)},t^{(k)}\right) + 
+              a_i^{(j)} + b_i^{(j)}t^{(k)} + \\
+             &c_i^{(j)}\sin\left(2 \pi t^{(k)}\right) +  
+              e_i^{(j)}\cos\left(2 \pi t^{(k)}\right) + \\ 
+             &f_i^{(j)}\sin\left(4 \pi t^{(k)}\right)  + 
+              g_i^{(j)}\cos\left(4 \pi t^{(k)}\right), 
 \end{split}
 \end{align}
-where $a_i^{(j)}$ is an offset that is unique for each station, $b_i^{(j)}$ is the secular velocity at $\vec{x}^{(j)}$, and the sinusoids describe seasonal deformation (using units of years for $t$).
+where $a_i^{(j)}$ is an offset that is unique for each station, $b_i^{(j)}$ is the secular velocity at $\vec{x}^{(j)}$, and the sinusoids describe seasonal deformation (using units of years for $t^{(k)}$).
 We use $\eta_i$ to represent all the non-parametric noise processes.
@@
 We move on to consider the column vector of $n$ GNSS displacement observations, $\mitbf{d}^*_i$, where the subscript indicates that we are still only considering component $i$ of displacements.
-The times and positions for each observation are described by the set $\mitbf{P}$.
+The observations are made at $m$ stations, and the times and positions for each observation are described by the set $\mitbf{P}$.
 $\mitbf{d}^*_i$ can be considered a realization of the random vector $\mitbf{d}_i$, which is formed by evaluating eq.
 (\ref{eq:Data}) at each point in $\mitbf{P}$.
-To write out $\mitbf{d}_i$ explicitly, we collect the coefficients from eq.
-(\ref{eq:Data}) into the vector $\mitbf{m}_i$, which contains 6 elements for each of the $m$ GNSS station.
-We use $\mitbf{G}$ to represent the $n \times 6m$ matrix of corresponding basis functions evaluated at each point in $\mitbf{P}$.
-We can write $\mitbf{d}_i$ as
+To write out $\mitbf{d}_i$ explicitly, we let $\mitbf{G}$ be an $n \times 6m$ matrix consisting of the basis functions from eq.
+(\ref{eq:Data}) (i.e., the linear trends and sinusoids for each station) evaluated at each point in $\mitbf{P}$.
+The coefficients corresponding to each basis function are collected into the column vector $\mitbf{m}_i$.
+We can then write $\mitbf{d}_i$ as
 \begin{equation}\label{eq:DataVec}
@@
 For each iteration, we compute $\mitbf{r}$ with eq.
-(\ref{eq:Residual}) and then update $\mitbf{\Omega}$ to be
+(\ref{eq:Residual}) and then update $\mitbf{\Omega}$ so that it contains the indices of $\mitbf{r}$ whose weighted values are less than $\lambda$ times the weighted root mean square of $\tilde{\mitbf{r}}$,
 \begin{equation}\label{eq:Update}
-\mitbf{\Omega} = \left\{i : \left|\frac{r_i}{\sigma_i}\right| < \lambda \cdot \mathrm{RMS}\right\},
+\mitbf{\Omega} \leftarrow \left\{i : \left|\frac{r_i}{\sigma_i}\right| < \lambda \cdot \sqrt{\frac{1}{|\mitbf{\Omega}|} \sum_{j \in \mitbf{\Omega}} \frac{r_j^2}{\sigma_j^2}} \right\}.
 \end{equation} 
-where RMS is the root-mean-square of $\tilde{\mitbf{r}}$ and $\lambda$ is an outlier tolerance. Iterations continue until the new $\mitbf{\Omega}$ is the same as the previous $\mitbf{\Omega}$.
+Iterations continue until the new $\mitbf{\Omega}$ is the same as the previous $\mitbf{\Omega}$.
 
@@
 That being said, this outlier detection algorithm significantly reduces the effort needed to manually clean GNSS data.
+
+\begin{figure*}
+\includegraphics{figures/outliers_new/outliers.pdf}
+\caption{Detrended easting component of displacements at station SC03, which is located on Mount Olympus in Washington.
+The orange markers indicate outliers that were automatically detected using the algorithm from Section \ref{sec:Outlier}.
+The error bars show one standard deviation uncertainties.
+Note that outliers tend be observed in the winter, suggesting that they were caused by snow or ice.}
+\label{fig:Outliers}
+\end{figure*}
 

### 2017-10-07 RevisedSubmission/figures/outliers_new/make_plot.py

@@ -1,0 +1,125 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import pygeons
+from matplotlib import colors
+import rbf
+import logging
+logging.basicConfig(level=logging.DEBUG)
+
+def trend(u1,t1):
+  idx = np.isfinite(u1)
+  u2 = u1[idx]
+  t2 = t1[idx]
+  G = np.array([np.ones(t2.shape[0]),t2]).T
+  m = np.linalg.inv(G.T.dot(G)).dot(G.T).dot(u2)
+  return m
+  
+xtick_dates = ['2011-01-01','2012-01-01','2013-01-01','2014-01-01']
+xticks = [pygeons.mjd.mjd(i,'%Y-%m-%d') for i in xtick_dates]
+xtick_labels = ['2011-01-01','2012-01-01','2013-01-01','2014-01-01']
+# start and end time in MJD
+xmin,xmax = 55378.0,56839.0
+
+c1 = colors.to_rgb('C0')
+c1t = tuple(0.6 + 0.4*np.array(colors.to_rgb('C0')))
+c2 = colors.to_rgb('C1')
+c2t = tuple(0.6 + 0.4*np.array(colors.to_rgb('C1')))
+
+# plot raw data
+fig,axs = plt.subplots(2,1,figsize=(7,4.5),sharex=True)
+pygeons.plot.plot._setup_ts_ax(axs)
+
+axs[0].set_xticks(xticks)
+axs[0].set_xticklabels(xtick_labels)
+axs[0].tick_params(labelsize=10)
+axs[0].set_xlim((xmin,xmax))
+axs[0].set_ylim((-10.0,15.0))
+axs[0].grid()
+
+axs[1].set_xticks(xticks)
+axs[1].set_xticklabels(xtick_labels)
+axs[1].tick_params(labelsize=10)
+axs[1].set_xlim((xmin,xmax))
+axs[1].set_ylim((-10.0,15.0))
+axs[1].grid()
+
+data = pygeons.io.io.dict_from_hdf5('data-2017-05-17.h5')
+idx = (data['id'] == 'SC03').nonzero()[0][0]
+lon = data['longitude'][idx]
+lat = data['latitude'][idx]
+
+t = data['time']
+u = 1000*data['east'][:,idx]
+us = 1000*data['east_std_dev'][:,idx]
+
+# toss out times that are out of range
+tidx = (t >= xmin) & (t <= xmax)
[…]

### 2017-10-09 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 Here we propose using Gaussian process regression (GPR) \citep{Rasmussen2006} to estimate transient strain from GNSS data.
+GPR is closely related to kriging \citep{Cressie1992} and least squares collocation \citep{Moritz1978}.
+The latter has been used by \citet{Kato1998} and \citet{El-Fiky1999} to estimate secular strain rates from GNSS data.
 GPR is a Bayesian, non-parametric method for inferring a continuous signal from scattered data.
 Since GNSS stations are irregularly spaced and observation times may differ between stations, GPR is an ideal tool for synthesizing discrete GNSS data into a spatially and temporally continuous representation of surface deformation.
-GPR is closely related to kriging \citep{Cressie1992} and least squares collocation \citep{Moritz1978}.
-The latter has been used by \citet{Kato1998} and \citet{El-Fiky1999} to estimate secular strain rates from GNSS data.
 GPR is Bayesian in that we describe our prior understanding of the geophysical signal with a Gaussian process.
-A Gaussian process is a normally distributed stochastic process that is fully defined in terms of a mean function and a positive definite covariance function.
-For example, Brownian motion, $B(t)$, is a well known Gaussian process in $\mathbb{R}^1$ which has zero mean and covariance function $\mathrm{Cov}\left[B(t),B(t')\right] = \min(t,t')$, where $t,t' \ge 0$.
-If no prior information is available for the geophysical signal, then maximum likelihood methods can be used to objectively choose a prior Gaussian process that is most consistent with the observations.
-We incorporate GNSS observations with the prior to form a posterior estimate of transient strain.
+A Gaussian process is a continuous stochastic process where any collection of points in the process can be described as a multivariate Gaussian random variable.
+Just as multivariate Gaussian random variables are fully determined by a mean vector and a covariance matrix, a Gaussian process is fully determined by a mean function and a covariance function.
+For example, Brownian motion, $B(t)$, is a well known Gaussian process in $\mathbb{R}^1$ which has a mean function $\mathrm{E}[B(t)] = 0$ and a covariance function $\mathrm{Cov}\left[B(t),B(t')\right] = \min(t,t')$, where $t,t' \ge 0$.
+If there is no information available to help choose an appropriate prior Gaussian process, then maximum likelihood methods can be used to objectively choose one that is most consistent with the data.
+Once a prior is selected, we incorporate GNSS observations with the prior to form a posterior estimate of transient strain.
 The posterior transient strain is also a Gaussian process, and we can use its distribution to confidently discern geophysical signal from noise.
@@
 Transient displacements have an easting and northing component, $\vec{u} = (u_\mathrm{e}, u_\mathrm{n})$, and we make a prior assumption that each component of $\vec{u}$ is a Gaussian process with zero mean and a generic covariance function $\mathrm{Cov}\left[u_i(p),u_i(p')\right] = C_{u_i}(p,p')$.
-Using a more concise notation, we describe each component of $\vec{u}$ as $u_i \sim \mathcal{N}\left(0,C_{u_i}\right)$.
-The function $C_{u_i}$ must be positive definite, which means that the matrix $[C_{u_i}(p,p')]_{(p,p') \in \mitbf{P} \times \mitbf{P}}$ is positive definite for any set of points $\mitbf{P}$ \citep[sec.
-2.5]{Cressie1992}.
+Using a more concise notation, we describe each component of $\vec{u}$ as $u_i \sim \mathcal{GP}\left(0,C_{u_i}\right)$.
+The function $C_{u_i}$ must be positive definite, which means that the matrix $[C_{u_i}(p,p')]_{(p,p') \in \mitbf{P} \times \mitbf{P}}$ is positive definite for any set of points $\mitbf{P}$.
 We assume that $C_{u_i}$ can be separated into spatial and temporal functions as
 \begin{equation}\label{eq:TransientCovariance}
-C_{u_i}(p,p') = C_{u_i}\left((\vec{x},t),(\vec{x}',t')\right) = X_i(\vec{x},\vec{x}')T_i(t,t').
+C_{u_i}(p,p') = C_{u_i}\left((\vec{x},t),(\vec{x}',t')\right) = \phi^2 X_i(\vec{x},\vec{x}')T_i(t,t'),
 \end{equation}   
+where the parameter $\phi$ controls the amplitude of realizations of $u_i$.
 As long as the functions $X_i$ and $T_i$ are positive definite, $C_{u_i}$ is guaranteed to also be positive definite and thus a valid covariance function \citep[sec.
@@
 2.2]{Adler1981} for a formal definition of differentiation for Gaussian processes).
-For example, the squared exponential function is a commonly used covariance function that is sufficiently differentiable for our purposes.
-We can make a potential covariance function for $u_i$ by using a squared exponential for $X$ and $T$,
+As an example, we can satisfy our requirements for positive definiteness and differentiability by using a squared exponential function for $X_i$ and $T_i$,
 \begin{equation}\label{eq:ExampleCov}
-C_{u_i}\left(p,p'\right) = \phi^2\exp\left(\frac{-||\vec{x} - \vec{x}'||_2^2}{2\ell^2}\right)
-                           \exp\left(\frac{-|t - t'|^2}{2\tau^2}\right),
+X_i(\vec{x},\vec{x}') = \exp\left(\frac{-||\vec{x} - \vec{x}'||_2^2}{2\ell^2}\right), \ \ \ \ T_i(t,t') = \exp\left(\frac{-|t - t'|^2}{2\tau^2}\right).
 \end{equation}
-where $\ell$ and $\tau$ describe a characteristic length-scale and time-scale for realizations of $u_i$.
-The parameter $\phi$ controls the amplitude of realizations of $u_i$.
-We refer to the parameters of a covariance function, such as $\ell$, $\tau$, and $\phi$, as hyperparameters.
-These hyperparameters can be chosen subjectively or they can be chosen objectively with maximum likelihood techniques.
-
-commonly used covariance function that is sufficiently differentiable for our purposes.
-
+The parameters $\ell$ and $\tau$ control the length-scale and time-scale, respectively, of realizations of $u_i$.
+While the squared exponential function is a commonly used covariance function for GPR, it is not appropriate for every application.
 The appropriate choice for $X_i$ and $T_i$ may vary depending on the geophysical signal we are trying to describe.
@@
 For example, $\eta_i$ can consist of uncorrelated white noise, temporally correlated noise describing benchmark wobble \citep[e.g.,][]{Wyatt1982,Wyatt1989}, and/or spatially correlated noise describing common mode error \citep[e.g.,][]{Wdowinski1997}.
-For now, we only make the assumption that $\eta_i$ is a Gaussian process distributed as $\eta_i \sim \mathcal{N}(0,C_{\eta_i})$, and $C_{\eta_i}$ is a spatially and temporally dependent covariance function.
+For now, we only make the assumption that $\eta_i$ is a Gaussian process distributed as $\eta_i \sim \mathcal{GP}(0,C_{\eta_i})$, and $C_{\eta_i}$ is a spatially and temporally dependent covariance function.
 We consider the coefficients in eq.
@@
 A general solution for $\hat{u}_i$ is derived in \citet[sec.
-8.9]{VonMises1964}, where we find that $\hat{u}_i$ is distributed as $\mathcal{N}(\mu_{\hat{u}_i},C_{\hat{u}_i})$ with mean function
+8.9]{VonMises1964}, where we find that $\hat{u}_i$ is distributed as $\mathcal{GP}(\mu_{\hat{u}_i},C_{\hat{u}_i})$ with mean function
 \begin{align}\label{eq:PosteriorMean}

### 2017-10-10 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 GPR is Bayesian in that we describe our prior understanding of the geophysical signal with a Gaussian process.
-A Gaussian process is a continuous stochastic process where any collection of points in the process can be described as a multivariate Gaussian random variable.
-Just as multivariate Gaussian random variables are fully determined by a mean vector and a covariance matrix, a Gaussian process is fully determined by a mean function and a covariance function.
+A Gaussian process is a continuous stochastic process whose value at any collection of points can be described as a multivariate Gaussian random variable.
+Just as a multivariate Gaussian random variable is fully determined by a mean vector and a covariance matrix, a Gaussian process is fully determined by a mean function and a covariance function.
 For example, Brownian motion, $B(t)$, is a well known Gaussian process in $\mathbb{R}^1$ which has a mean function $\mathrm{E}[B(t)] = 0$ and a covariance function $\mathrm{Cov}\left[B(t),B(t')\right] = \min(t,t')$, where $t,t' \ge 0$.
@@
 In deriving our formulation for transient strain rates, we have assumed that noise in the data vector is normally distributed.
-This is not an appropriate assumption for GNSS data, which often have more outliers than would be predicted for normally distributed noise.
+This is not an appropriate assumption for GNSS data, which often have more outliers than would be expected for normally distributed noise.
 Methods for analyzing GNSS data should either be robust against outliers or should involve a preprocessing step in which outliers are detected and removed.
@@
 Observations with residuals that exceed some threshold are removed.
-This strategy is commonly used to detect outliers in GNSS data, where the residuals are typically computed by estimating and removing a linear trend and seasonal terms for each time series \citep[e.g.,][]{Johansson2002,Dong2006,Bos2013}.
-The outlier detection method used here is different from this more standard approach because the model we fit to the data additionally consists of a spatially and temporally correlated Gaussian process, $v \sim \mathcal{N}(0,C_v)$.
-We use $v$ to explain any coherent features in the data that cannot be explained by the linear trend and sinusoids.
-With this more complex model, we are able to detect outliers that are spatially or temporally anomalous.
+This strategy for detecting outliers is commonly used for GNSS data, where the model being fit to the data typically consists of a linear trend and seasonal terms for each GNSS station \citep[e.g.,][]{Johansson2002,Dong2006,Bos2013}.
+To prevent transient geophysical signal from being erroneously identified as outliers, the model used in our outlier detection algorithm additionally consists of a temporally correlated Gaussian process.
+The details of our algorithm are given in Appendix A.
 
-To describe our outlier detection algorithm in detail, we let $\mitbf{d}^*$ denote all $n$ GNSS displacement observations for a single component (note that we dropped the component subscripts since outliers are detected for the easting and northing components independently).
-Our goal is to find the index set of non-outliers in $\mitbf{d}^*$, which we denote as $\mitbf{\Omega}$.
-For notational ease, we use a tilde to indicate that an array only contains elements corresponding to $\mitbf{\Omega}$ (e.g., the vector of non-outlier observations is denoted $\tilde{\mitbf{d}}^* = [d_i^*]_{i \in \mitbf{\Omega}}$).
-We now consider $\mitbf{d}^*$ to be a realization of the random vector
-\begin{equation}
-\mitbf{d} = \mitbf{G}\mitbf{m} + v(\mitbf{P}) + \mitbf{w},
-\end{equation}
-where $\mitbf{G}$ and $\mitbf{m}$ are the same as in eq.
-(\ref{eq:DataVec}), and $\mitbf{w}$ is a vector of uncorrelated noise with assumed known standard deviations $\mitbf{\sigma} = [\sigma_1, \sigma_2, ..., \sigma_n]$.
-Our best fitting model for $\mitbf{d}^*$ is taken to be the mean of $\mitbf{G}\mitbf{m} + v(\mitbf{P})$ after it is conditioned with non-outlier observations.
-The residual vector is then
-\begin{equation}\label{eq:Residual1}
-\mitbf{r} = \mitbf{d}^* - \mathrm{E}\left[\Big(\mitbf{G}\mitbf{m} + v(\mitbf{P})\Big) \Big| \Big(\tilde{\mitbf{d}} = \tilde{\mitbf{d}}^*\Big)\right].
-\end{equation}
-Following a similar procedure as above, we can write eq. (\ref{eq:Residual1}) as
-\begin{equation}\label{eq:Residual}
-\mitbf{r} = \mitbf{d}^*  - 
-            \left[\begin{array}{cc}
-                  C_v(\mitbf{P},\tilde{\mitbf{P}}) & \mitbf{G} \\
-                  \end{array}\right]
-            \left[\begin{array}{cc}
-                  C_v(\tilde{\mitbf{P}},\tilde{\mitbf{P}}) + \mathrm{diag}(\tilde{\mitbf{\sigma}}^2) & \tilde{\mitbf{G}} \\
-                  \tilde{\mitbf{G}}^T  & \mitbf{0} \\
-                  \end{array}\right]^{-1}
-            \left[\begin{array}{c}
-                  \tilde{\mitbf{d}}^* \\
-                  \mitbf{0} \\
-                  \end{array}\right].
-\end{equation}
-We start the algorithm by letting $\mitbf{\Omega}$ contain all $n$ indices, and we refine $\mitbf{\Omega}$ iteratively.
-For each iteration, we compute $\mitbf{r}$ with eq.
-(\ref{eq:Residual}) and then update $\mitbf{\Omega}$ so that it contains the indices of $\mitbf{r}$ whose weighted values are less than $\lambda$ times the weighted root mean square of $\tilde{\mitbf{r}}$,
-\begin{equation}\label{eq:Update}
-\mitbf{\Omega} \leftarrow \left\{i : \left|\frac{r_i}{\sigma_i}\right| < \lambda \cdot \sqrt{\frac{1}{|\mitbf{\Omega}|} \sum_{j \in \mitbf{\Omega}} \frac{r_j^2}{\sigma_j^2}} \right\}.
-\end{equation} 
-Iterations continue until the new $\mitbf{\Omega}$ is the same as the previous $\mitbf{\Omega}$.
-
-We use $\lambda=4$ in this study, which in our experience accurately identifies outliers without unnecessarily decimating the data.
-Iterations continue until the new $\mitbf{\Omega}$ is equal to the previous $\mitbf{\Omega}$.
-
-It should be noted that this algorithm does not identify jumps in GNSS time series (cite Gazeaux 2013), which are another common issue.
-Some, but not all, jumps can be automatically removed by looking up the dates of equipment changes and earthquakes.
+It should be noted that our algorithm does not identify jumps in GNSS time series, which are another common issue.
+Some, but not all, jumps can be automatically removed by looking up the dates of equipment changes and earthquakes \citep{Gazeaux2013}.
 However, it is still necessary to manually find and remove jumps of unknown origin.
-That being said, this outlier detection algorithm significantly reduces the effort needed to manually clean GNSS data.
-
-\begin{figure*}
-\includegraphics{figures/outliers_new/outliers.pdf}
-\caption{Detrended easting component of displacements at station SC03, which is located on Mount Olympus in Washington.
-The orange markers indicate outliers that were automatically detected using the algorithm from Section \ref{sec:Outlier}.
-The error bars show one standard deviation uncertainties.
-Note that outliers tend be observed in the winter, suggesting that they were caused by snow or ice.}
-\label{fig:Outliers}
-\end{figure*}
 
@@
 
+\appendix
+
+\section{Outlier detection algorithm}
+Our outlier detection algorithm is loosely based on the data editing algorithm from \citet{Acheson1975}.
+Let $\mitbf{d}^*$ denote all $n$ GNSS displacement observations for a single directional component, which have been made at positions and times $\mitbf{P}$.
+We model $\mitbf{d}^*$ as a realization of the random vector
+\begin{equation}\label{eq:OutlierDataVec}
+\mitbf{d} = \mitbf{G}\mitbf{m} + v(\mitbf{P}) + \mitbf{w},
+\end{equation}
+where $\mitbf{G}$ and $\mitbf{m}$ are the same as in eq.
+(\ref{eq:DataVec}), $v$ is a Gaussian process distributed as $\mathcal{GP}(0,C_v)$, and $\mitbf{w}$ is a vector of uncorrelated noise with known standard deviations $\mitbf{\sigma} = [\sigma_1, \sigma_2, ..., \sigma_n]$.
+The Gaussian process $v$ is intended to describe transient features in the data that cannot be explained by the linear trend or seasonal terms in $\mitbf{G}$.
+We let the temporal covariance of $v$ be a squared exponential, and we let $v$ be spatially uncorrelated so that,
+\begin{equation}
+C_v(p,p') = \phi^2\exp\left(\frac{-|t - t'|^2}{2\tau^2}\right) C_w(\vec{x},\vec{x}'),
+\end{equation}
+where 
+\begin{equation}\label{eq:WhiteCov}
+C_w(\vec{x},\vec{x}') = 
+\begin{cases}
+1, &\vec{x} = \vec{x}' \\
+0, &\vec{x} \neq \vec{x}'
+\end{cases}.
+\end{equation}
+The spatial covariance of $v$ has little effect on the detected outliers, and so we have assumed that $v$ is spatially uncorrelated for simplicity.
+Based on our experience, $v$ can reasonably describe most transient features in the data when we set $\phi = 1$ mm and $\tau = 10$ days.
+
+Our goal is to find the index set of non-outliers in $\mitbf{d}^*$, which we denote as $\mitbf{\Omega}$.
+We use a tilde to indicate that an array only contains elements corresponding to $\mitbf{\Omega}$ (e.g., the vector of non-outlier observations is denoted $\tilde{\mitbf{d}}^* = [d_i^*]_{i \in \mitbf{\Omega}}$).
+The outliers are identified iteratively, and we initiate $\mitbf{\Omega}$ with all $n$ indices.
+We consider outliers to be data that are inconsistent with eq.
+(\ref{eq:OutlierDataVec}), which we determine with the residual vector
+\begin{align}\label{eq:Residual}
+\mitbf{r} &= \mitbf{d}^* - \mathrm{E}\left[\Big(\mitbf{G}\mitbf{m} + v(\mitbf{P})\Big) \Big| \Big(\tilde{\mitbf{d}} = \tilde{\mitbf{d}}^*\Big)\right] \\
+          &= \mitbf{d}^*  - 
+            \left[\begin{array}{cc}
+                  C_v(\mitbf{P},\tilde{\mitbf{P}}) & \mitbf{G} \nonumber \\
+                  \end{array}\right]
+            \left[\begin{array}{cc}
+                  C_v(\tilde{\mitbf{P}},\tilde{\mitbf{P}}) + \mathrm{diag}(\tilde{\mitbf{\sigma}}^2) & \tilde{\mitbf{G}} \\
+                  \tilde{\mitbf{G}}^T  & \mitbf{0} \\
+                  \end{array}\right]^{-1}
+            \left[\begin{array}{c}
+                  \tilde{\mitbf{d}}^* \\
+                  \mitbf{0} \\
+                  \end{array}\right].
+\end{align}
+Data with abnormally large residuals are identified as outliers.
+For each iteration, we compute $\mitbf{r}$ and then update $\mitbf{\Omega}$ so that it contains the indices of $\mitbf{r}$ whose weighted values are less than $\lambda$ times the weighted root mean square of $\tilde{\mitbf{r}}$,
+\begin{equation}\label{eq:Update}
+\mitbf{\Omega} \leftarrow \left\{i : \left|\frac{r_i}{\sigma_i}\right| < \lambda \cdot \sqrt{\frac{1}{|\mitbf{\Omega}|} \sum_{j \in \mitbf{\Omega}} \frac{r_j^2}{\sigma_j^2}} \right\}.
+\end{equation} 
+Iterations continue until the new $\mitbf{\Omega}$ is the same as the previous $\mitbf{\Omega}$.
+We use $\lambda=4$ in this study, which in our experience accurately identifies outliers without unnecessarily decimating the data.
+
+To demostrate this algorithm ...
+
+\begin{figure*}
+\includegraphics{figures/outliers_new/outliers.pdf}
+\caption{Detrended easting component of displacements at station SC03, which is located on Mount Olympus in Washington.
+The orange markers indicate outliers that were automatically detected using the algorithm from Section \ref{sec:Outlier}.
+The error bars show one standard deviation uncertainties.
+Note that outliers tend be observed in the winter, suggesting that they were caused by snow or ice.}
+\label{fig:Outliers}
+\end{figure*}
+
 \bibliographystyle{gji}

### 2017-10-10 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 where $\mitbf{G}$ and $\mitbf{m}$ are the same as in eq.
-(\ref{eq:DataVec}), $v$ is a Gaussian process distributed as $\mathcal{GP}(0,C_v)$, and $\mitbf{w}$ is a vector of uncorrelated noise with known standard deviations $\mitbf{\sigma} = [\sigma_1, \sigma_2, ..., \sigma_n]$.
+(\ref{eq:DataVec}), $v$ is a Gaussian process distributed as $\mathcal{GP}(0,C_v)$, and $\mitbf{w}$ is a vector of uncorrelated Gaussian noise with known standard deviations $\mitbf{\sigma} = [\sigma_1, \sigma_2, ..., \sigma_n]$.
 The Gaussian process $v$ is intended to describe transient features in the data that cannot be explained by the linear trend or seasonal terms in $\mitbf{G}$.

### 2017-10-10 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 Let $\mitbf{d}^*$ denote all $n$ GNSS displacement observations for a single directional component, which have been made at positions and times $\mitbf{P}$.
-We model $\mitbf{d}^*$ as a realization of the random vector
+We describe $\mitbf{d}^*$ as a realization of the random vector
 \begin{equation}\label{eq:OutlierDataVec}
@@
 The outliers are identified iteratively, and we initiate $\mitbf{\Omega}$ with all $n$ indices.
-We consider outliers to be data that are inconsistent with eq.
-(\ref{eq:OutlierDataVec}), which we determine with the residual vector
+We consider outliers to be data that are poorly explained by the model $\mitbf{G}\mitbf{m} + v(\mitbf{P})$, which is determine by the residual vector
 \begin{align}\label{eq:Residual}

### 2017-10-11 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 
-Most methods for estimating strain rates from GNSS data assume some parametric form of the deformation signal.
-The simplest method for estimating secular strain rates assumes that GNSS derived velocities can be described with a first-order polynomial (i.e., having constant deformation gradients) over some subnetwork of the GNSS stations \citep[e.g.,][]{Feigl1990,Murray2000}.
-The components of the strain rate tensor  for each subnetwork are then determined through a least squares fit to the observations.
+Most methods for estimating strain rates from GNSS data assume some parametric form for the deformation signal.
+The simplest method for estimating secular strain rates assumes that GNSS derived velocities can be described with a first-order polynomial (i.e, the deformation gradients are constant) over some subnetwork of the GNSS stations \citep[e.g.,][]{Feigl1990,Murray2000}.
+The components of the strain rate tensor for each subnetwork are then determined from the least squares fit to the observations.
 The assumption that deformation gradients are spatially uniform is not appropriate when subnetworks span too large of an area.
-To help overcome this deficiency, \citet{Shen1996,Shen2015} used an inverse distance weighting scheme, in which the estimated strain rate at some point is primarily controlled by observations at nearby stations.
-However, the methods described in \citet{Shen1996,Shen2015} are still formulated by assuming that the deformation gradients are uniform over the entire network.
-The errors in this assumption manifest as implausibly low formal uncertainties for the estimated strain rates.
+To help overcome this deficiency, \citet{Shen1996,Shen2015} used an inverse distance weighting scheme, in which the estimated strain rate at a point is primarily controlled by observations at nearby stations.
+The method  of \citet{Shen1996,Shen2015} can be viewed as a form of local least squares regression with a first-order polynomial \citep[e.g.,][sec.
+6]{Hastie2009}.
 Other methods for estimating secular strain rates have parameterized GNSS derived velocities with bi-cubic splines \citep{Beavan2001}, spherical wavelets \citep{Tape2009}, and elastostatic Green's functions \citep{Sandwell2016}.
-The type of basis functions and the number of degrees of freedom for a parameterization can be subjective.
+The type of basis functions and the number of degrees of freedom for a parameterization, which are often chosen subjectively, have a strong influence on the strain solution.
 If there are too few degrees of freedom in the parameterization, then estimated strain rates will be biased and the uncertainties will be underestimated.

### 2017-10-12 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 
-To demostrate this algorithm ...
+The outlier detection algorithm is demonstrated in Figure \ref{fig:Outliers}.
+In the demonstration we use the easting component of displacements at a single station, SC03, which is located on Mt.
+Olympus in Washington state.
+Station SC03 records anomalous observations during the winter, presumably because of snow and ice accumulation, and we want to remove these observations.
+The station also records periodic westward motion from slow slip events, and we want to keep this deformation intact.
+The detected outliers are shown in (Panel B).
+For comparison, we also show the detected outliers when we do not include the Gaussian process $v$ in our model for the data (Panel A).
+When $v$ is not included, real transient deformation resulting from slow slip events is erroneously identified as outliers.
+When $v$ is included, the identified outliers only consist of the anomalous deformation that lacks temporal continuity.
 
@@
 \includegraphics{figures/outliers_new/outliers.pdf}
-\caption{Detrended easting component of displacements at station SC03, which is located on Mount Olympus in Washington.
-The orange markers indicate outliers that were automatically detected using the algorithm from Section \ref{sec:Outlier}.
-The error bars show one standard deviation uncertainties.
-Note that outliers tend be observed in the winter, suggesting that they were caused by snow or ice.}
+\caption{Outliers detected in the easting component of displacements at station SC03.
+The orange markers indicate detected outliers.
+The orange line is the best fit model to the data, which is used to compute the residual vector $\mitbf{r}$.
+The model being fit to the data in Panel A is $\mitbf{G}\mitbf{m}$, and the model in Panel B is $\mitbf{G}\mitbf{m} + v(\mitbf{P})$.}
 \label{fig:Outliers}

### 2017-10-12 RevisedSubmission/figures/outliers_new/make_plot.py

@@ -11,7 +11,7 @@ def trend(u1,t1):
   u2 = u1[idx]
   t2 = t1[idx]
   G = np.array([np.ones(t2.shape[0]),t2]).T
-  m = np.linalg.inv(G.T.dot(G)).dot(G.T).dot(u2)
+  m = np.linalg.pinv(G.T.dot(G)).dot(G.T).dot(u2)
   return m
   
 xtick_dates = ['2011-01-01','2012-01-01','2013-01-01','2014-01-01']
@@ -26,6 +26,21 @@ c2 = colors.to_rgb('C1')
 c2t = tuple(0.6 + 0.4*np.array(colors.to_rgb('C1')))
 
 # plot raw data
+
+data = pygeons.io.io.dict_from_hdf5('data-2017-05-17.h5')
+
+name = 'SC03'
+idx = (data['id'] == name).nonzero()[0][0]
+lon = data['longitude'][idx]
+lat = data['latitude'][idx]
+
+t = data['time']
+u = 1000*data['east'][:,idx]
+us = 1000*data['east_std_dev'][:,idx]
+
+# toss out times that are out of range or have missing data
+tidx = (t >= xmin) & (t <= xmax) & np.isfinite(u)
+
 fig,axs = plt.subplots(2,1,figsize=(7,4.5),sharex=True)
 pygeons.plot.plot._setup_ts_ax(axs)
 
@@ -43,22 +58,9 @@ axs[1].set_xlim((xmin,xmax))
 axs[1].set_ylim((-10.0,15.0))
 axs[1].grid()
 
-data = pygeons.io.io.dict_from_hdf5('data-2017-05-17.h5')
-idx = (data['id'] == 'SC03').nonzero()[0][0]
-lon = data['longitude'][idx]
-lat = data['latitude'][idx]
-
-t = data['time']
-u = 1000*data['east'][:,idx]
-us = 1000*data['east_std_dev'][:,idx]
-
-# toss out times that are out of range
-tidx = (t >= xmin) & (t <= xmax)
 t = t[tidx]
 u = u[tidx]
 us = us[tidx]
-
-print(t.shape)
 
 # detrend for viewing ease
 m = trend(u,t)
@@ -74,13 +76,14 @@ def basis(t):
                    np.cos(4*np.pi*t[:,0]/365.25)]).T
                    
 # parameters
-tol = 3.0
+tol = 2.5
 
 # using just basis functions
 gp = rbf.gauss.gpbfci(basis)
 
 # find outliers
[…]

### 2017-10-12 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 Iterations continue until the new $\mitbf{\Omega}$ is the same as the previous $\mitbf{\Omega}$.
-We use $\lambda=4$ in this study, which in our experience accurately identifies outliers without unnecessarily decimating the data.
 
 The outlier detection algorithm is demonstrated in Figure \ref{fig:Outliers}.
-In the demonstration we use the easting component of displacements at a single station, SC03, which is located on Mt.
+For the demonstration, we use the easting component of displacements at a single station, SC03, which is located on Mt.
 Olympus in Washington state.
@@
 The station also records periodic westward motion from slow slip events, and we want to keep this deformation intact.
-The detected outliers are shown in (Panel B).
+The detected outliers are shown in Panel B.
 For comparison, we also show the detected outliers when we do not include the Gaussian process $v$ in our model for the data (Panel A).
@@
 When $v$ is included, the identified outliers only consist of the anomalous deformation that lacks temporal continuity.
+It should be noted that we use $\lambda = 2.5$ for this demonstration, which causes the outlier detection algorithm to be particularly aggressive.
+In Section \ref{sec:Cascadia}, we clean the data using a more tolerant $\lambda = 4.0$.
 

### 2017-10-13 RevisedSubmission/hines_hetland_2017_gji.tex
did more shit

@@
 GPR is Bayesian in that we describe our prior understanding of the geophysical signal with a Gaussian process.
-A Gaussian process is a continuous stochastic process whose value at any collection of points can be described as a multivariate Gaussian random variable.
-Just as a multivariate Gaussian random variable is fully determined by a mean vector and a covariance matrix, a Gaussian process is fully determined by a mean function and a covariance function.
+A Gaussian process is a continuous stochastic process whose value at any collection of points can be described as a Gaussian random vector.
+Just as a Gaussian random vector is fully determined by a mean vector and a covariance matrix, a Gaussian process is fully determined by a mean function and a covariance function.
 For example, Brownian motion, $B(t)$, is a well known Gaussian process in $\mathbb{R}^1$ which has a mean function $\mathrm{E}[B(t)] = 0$ and a covariance function $\mathrm{Cov}\left[B(t),B(t')\right] = \min(t,t')$, where $t,t' \ge 0$.
@@
 We consider transient strain rates to be any deviation from secular strain rates, and our attention is limited to horizontal strain rates.
-We denote transient strain rates as $\dot\varepsilon(p)$, where $p$ represents the ordered pair $(\vec{x},t)$, $\vec{x}$ are spatial coordinates, and $t$ is time.
-For simplicity, we assume that the study region is sufficiently small that $\vec{x}$ can be considered a point in a 2-D map projection that is aligned with the cardinal directions, $\vec{x} = (x_\mathrm{e},x_\mathrm{n})$.
-We determine $\dot\varepsilon$ by spatially and temporally differentiating estimates of transient displacements, $\vec{u}(p)$.
-Transient displacements have an easting and northing component, $\vec{u} = (u_\mathrm{e}, u_\mathrm{n})$, and we make a prior assumption that each component of $\vec{u}$ is a Gaussian process with zero mean and a generic covariance function $\mathrm{Cov}\left[u_i(p),u_i(p')\right] = C_{u_i}(p,p')$.
+We denote transient strain rates as $\dot\varepsilon(p)$, where $p$ represents the ordered pair $(\vec{x},t)$, $\vec{x} = (x_\mathrm{e},x_\mathrm{n})$ are spatial coordinates, and $t$ is time.
+For simplicity, we assume that the study region is sufficiently small that $\vec{x}$ can be considered a point in a 2-D map projection that is aligned with the cardinal directions.
+We determine $\dot\varepsilon$ by spatially and temporally differentiating estimates of transient displacements, $\vec{u}(p) = (u_\mathrm{e},u_\mathrm{n})$.
+
+We make a prior assumption that each component of $\vec{u}$ is a Gaussian process with zero mean and a generic covariance function $\mathrm{Cov}\left[u_i(p),u_i(p')\right] = C_{u_i}(p,p')$.
 Using a more concise notation, we describe each component of $\vec{u}$ as $u_i \sim \mathcal{GP}\left(0,C_{u_i}\right)$.
@@
 We first consider component $i$ of a single GNSS displacement observation made at station $j$, which is located at $\vec{x}^{(j)}$, and time $t^{(k)}$.
-We describe this observation as a realization of the random variable
+We describe this observation, $d_i^{*(jk)}$, as a realization of the random variable
 \begin{align}\label{eq:Data}
@@
 d_i^{(jk)} = &u_i\left(\vec{x}^{(j)},t^{(k)}\right) + 
-              \eta_i\left(\vec{x}^{(j)},t^{(k)}\right) + 
+              \eta_i^{(jk)} + 
               a_i^{(j)} + b_i^{(j)}t^{(k)} + \\
@@
 \end{align}
-where $a_i^{(j)}$ is an offset that is unique for each station, $b_i^{(j)}$ is the secular velocity at $\vec{x}^{(j)}$, and the sinusoids describe seasonal deformation (using units of years for $t^{(k)}$).
-We use $\eta_i$ to represent all the non-parametric noise processes.
-For example, $\eta_i$ can consist of uncorrelated white noise, temporally correlated noise describing benchmark wobble \citep[e.g.,][]{Wyatt1982,Wyatt1989}, and/or spatially correlated noise describing common mode error \citep[e.g.,][]{Wdowinski1997}.
-For now, we only make the assumption that $\eta_i$ is a Gaussian process distributed as $\eta_i \sim \mathcal{GP}(0,C_{\eta_i})$, and $C_{\eta_i}$ is a spatially and temporally dependent covariance function.
-We consider the coefficients in eq.
-(\ref{eq:Data}) to be uncorrelated random variables distributed as $\mathcal{N}(0,\kappa^2)$ in the limit as $\kappa \to \infty$ (i.e., the coefficients have diffuse priors).
-Of course, the secular velocities, $b_i^{(j)}$, are spatially correlated and we could invoke a tectonic model to form a prior on $b_i^{(j)}$.
-However, in our application to the Pacific Northwest, we will be using displacement time series which are long enough to sufficiently constrain $b_i^{(j)}$ for each station, avoiding the need to incorporate a prior.
-Likewise, seasonal deformation is spatially correlated \citep{Dong2002,Langbein2008}, and it may be worth exploring and exploiting such a correlation in a future study.
-
-We move on to consider the column vector of $n$ GNSS displacement observations, $\mitbf{d}^*_i$, where the subscript indicates that we are still only considering component $i$ of displacements.
+where $\eta_i^{(jk)}$ describes noise, $a_i^{(j)}$ is an offset that is unique for each station, $b_i^{(j)}$ is the secular velocity at $\vec{x}^{(j)}$, and the sinusoids describe seasonal deformation (using units of years for $t^{(k)}$).
+We then consider the column vector of $n$ GNSS displacement observations, $\mitbf{d}^*_i$, where the subscript indicates that we are still only considering component $i$ of displacements.
 The observations are made at $m$ stations, and the times and positions for each observation are described by the set $\mitbf{P}$.
@@
 The coefficients corresponding to each basis function are collected into the column vector $\mitbf{m}_i$.
+The noise for all the observations are described by the column vector $\mitbf{\eta}_i$.
 We can then write $\mitbf{d}_i$ as
 \begin{equation}\label{eq:DataVec}
-\mitbf{d}_i = u_i(\mitbf{P}) + \eta_i(\mitbf{P}) + \mitbf{G}\mitbf{m}_i,
+\mitbf{d}_i = u_i(\mitbf{P}) + \mitbf{\eta}_i + \mitbf{G}\mitbf{m}_i,
 \end{equation}
-which is distributed as $\mathcal{N}(\mitbf{0},\mitbf{\Sigma} + \kappa^2\mitbf{G}\mitbf{G}^T)$, where
-\begin{equation}\label{eq:Cd}
-\mitbf{\Sigma} = C_{u_i}(\mitbf{P},\mitbf{P}) + C_{\eta_i}(\mitbf{P},\mitbf{P}).  
+where the notation $u_i(\mitbf{P})$ represents the column vector $[u_i(p)]_{p \in \mitbf{P}}$.
+
+We assume a diffuse prior for the components of $\mitbf{m}_i$, that is to say $\mitbf{m}_i \sim \mathcal{N}(\mitbf{0},\kappa^2\mitbf{I})$ in the limit as $\kappa \to \infty$.
+Of course, the secular velocities, $b_i^{(j)}$, are spatially correlated and we could invoke a tectonic model to form a prior on $b_i^{(j)}$.
+However, in our application to the Pacific Northwest, we will be using displacement time series which are long enough to sufficiently constrain $b_i^{(j)}$ for each station, avoiding the need to incorporate a prior.
+Likewise, seasonal deformation is spatially correlated \citep{Dong2002,Langbein2008}, and it may be worth exploring and exploiting such a correlation in a future study.
+We assume that $\mitbf{\eta}_i$ is a spatially and/or temporally corelated random vector distributed as $\mathcal{N}(\mitbf{0},\mitbf{C}_{\eta_i})$.
+For example, $\mitbf{\eta}_i$ can be uncorrelated white noise, temporally correlated noise describing benchmark wobble \citep[e.g.,][]{Wyatt1982,Wyatt1989}, and/or spatially correlated noise describing common mode error \citep[e.g.,][]{Wdowinski1997}.
+The appropriate noise model may vary depending on the application, and we hold off on specifying the covariance matrix, $\mitbf{C}_{\eta_i}$, until Section \ref{sec:NoiseModel}.
+We are now able to write the distributed of $\mitbf{d}_i$ as
+\begin{equation}\label{eq:DataDist}
+\mitbf{d}_i \sim \mathcal{N}\left(\mitbf{0}, C_{u_i}(\mitbf{P},\mitbf{P}) + \mitbf{C}_{\eta_i}  + \kappa^2\mitbf{G}\mitbf{G}^T \right),
 \end{equation}
-It should be understood that notation such as $u_i(\mitbf{P})$ and $C_{u_i}(\mitbf{P},\mitbf{P})$ represents the column vector $[u_i(p)]_{p \in \mitbf{P}}$ and the matrix $[C_{u_i}(p,p')]_{(p,p') \in \mitbf{P} \times \mitbf{P}}$, respectively.
+where $C_{u_i}(\mitbf{P},\mitbf{P})$ represents the matrix $[C_{u_i}(p,p')]_{(p,p') \in \mitbf{P} \times \mitbf{P}}$.
 
-We combine the prior on transient displacements with GNSS displacement data to form a posterior estimate of transient displacements.
-We ignore any covariances between the easting and northing components, which reduces the computational complexity of our analysis because we can determine each component of the posterior independently.
-However, we are inherently assuming that the principle components of transient displacements and the noise processes are aligned with the cardinal directions, which is admittedly an arbitrary assumption.
-We then form each component of the posterior, denoted as $\hat{u}_i$, by conditioning $u_i$ with the fact that $\mitbf{d}^*_i$ has been realized from the random vector $\mitbf{d}_i$.
+We form a posterior estimate of transient displacements, denoted as $\hat{u}_i$, by updating $u_i$ with the fact that $\mitbf{d}_i^*$ was realized from the random vector $\mitbf{d}_i$.
 That is to say $\hat{u}_i = u_i |(\mitbf{d}_i = \mitbf{d}^*_i)$.
@@
                       \left(\mitbf{d}^*_i - \mathrm{E}\left[\mitbf{d}_i\right]\right)\nonumber \\
-                   &= C_{u_i}(p,\mitbf{P})\left(\mitbf{\Sigma} + \kappa^2\mitbf{G}\mitbf{G}^T\right)^{-1}\mitbf{d}^*_i
+                   &= C_{u_i}(p,\mitbf{P})\left( C_{u_i}(\mitbf{P},\mitbf{P}) + \mitbf{C}_{\eta_i} + \kappa^2\mitbf{G}\mitbf{G}^T\right)^{-1}\mitbf{d}^*_i
 \end{align}    
@@
                        \mathrm{Cov}\left[\mitbf{d}_i,u_i(p')\right]\nonumber \\
-                    &= C_{u_i}(p,p') - C_{u_i}(p,\mitbf{P})\left(\mitbf{\Sigma} + \kappa^2\mitbf{G}\mitbf{G}^T\right)^{-1}C_{u_i}(\mitbf{P},p').
+                    &= C_{u_i}(p,p') - C_{u_i}(p,\mitbf{P})\left( C_{u_i}(\mitbf{P},\mitbf{P}) + \mitbf{C}_{\eta_i} + \kappa^2\mitbf{G}\mitbf{G}^T \right)^{-1}C_{u_i}(\mitbf{P},p').
 \end{align}
@@
                    \left[\begin{array}{cc}
-                         \mitbf{\Sigma} & \mitbf{G} \\
+                         C_{u_i}(\mitbf{P},\mitbf{P}) + \mitbf{C}_{\eta_i} & \mitbf{G} \\
                          \mitbf{G}^T  & -\kappa^{-2} \mitbf{I} \\
@@
                     \left[\begin{array}{cc}
-                          \mitbf{\Sigma} & \mitbf{G} \\
+                          C_{u_i}(\mitbf{P},\mitbf{P}) + \mitbf{C}_{\eta_i} & \mitbf{G} \\
                           \mitbf{G}^T  & -\kappa^{-2} \mitbf{I} \\
@@
                    \left[\begin{array}{cc}
-                         \mitbf{\Sigma} & \mitbf{G} \\
+                         C_{u_i}(\mitbf{P},\mitbf{P}) + \mitbf{C}_{\eta_i} & \mitbf{G} \\
                          \mitbf{G}^T  & \mitbf{0} \\
@@
                     \left[\begin{array}{cc}
-                          \mitbf{\Sigma} & \mitbf{G} \\
+                          C_{u_i}(\mitbf{P},\mitbf{P}) + \mitbf{C}_{\eta_i} & \mitbf{G} \\
                           \mitbf{G}^T  & \mitbf{0} \\
@@
 In that case, a singular value decomposition can be used to remove linearly dependent components from $\mitbf{G}$.
+
+It should be noted that we ignore any covariances between the easting and northing components, which reduces the computational complexity of our analysis because we can determine each component of the posterior independently.
+However, we are inherently assuming that the principle components of transient displacements and the noise processes are aligned with the cardinal directions, which is admittedly an arbitrary assumption.
 
@@
 \begin{equation}
-C_v(p,p') = \phi^2\exp\left(\frac{-|t - t'|^2}{2\tau^2}\right) C_w(\vec{x},\vec{x}'),
+C_v(p,p') = \phi^2\exp\left(\frac{-|t - t'|^2}{2\tau^2}\right) \delta_{\vec{x},\vec{x}'},
 \end{equation}
-where 
-\begin{equation}\label{eq:WhiteCov}
-C_w(\vec{x},\vec{x}') = 
-\begin{cases}
-1, &\vec{x} = \vec{x}' \\
-0, &\vec{x} \neq \vec{x}'
-\end{cases}.
-\end{equation}
+where $\delta_{\vec{x},\vec{x}'} = 1$ if $\vec{x} = \vec{x}'$ and $0$ otherwise.
 The spatial covariance of $v$ has little effect on the detected outliers, and so we have assumed that $v$ is spatially uncorrelated for simplicity.

### 2017-10-14 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 Likewise, seasonal deformation is spatially correlated \citep{Dong2002,Langbein2008}, and it may be worth exploring and exploiting such a correlation in a future study.
-We assume that $\mitbf{\eta}_i$ is a spatially and/or temporally corelated random vector distributed as $\mathcal{N}(\mitbf{0},\mitbf{C}_{\eta_i})$.
+We assume that $\mitbf{\eta}_i$ is a spatially and/or temporally correlated random vector distributed as $\mathcal{N}(\mitbf{0},\mitbf{C}_{\eta_i})$.
 For example, $\mitbf{\eta}_i$ can be uncorrelated white noise, temporally correlated noise describing benchmark wobble \citep[e.g.,][]{Wyatt1982,Wyatt1989}, and/or spatially correlated noise describing common mode error \citep[e.g.,][]{Wdowinski1997}.
 The appropriate noise model may vary depending on the application, and we hold off on specifying the covariance matrix, $\mitbf{C}_{\eta_i}$, until Section \ref{sec:NoiseModel}.
-We are now able to write the distributed of $\mitbf{d}_i$ as
+We are now able to write the distribution of $\mitbf{d}_i$ as
 \begin{equation}\label{eq:DataDist}
@@
 
-We form a posterior estimate of transient displacements, denoted as $\hat{u}_i$, by updating $u_i$ with the fact that $\mitbf{d}_i^*$ was realized from the random vector $\mitbf{d}_i$.
-That is to say $\hat{u}_i = u_i |(\mitbf{d}_i = \mitbf{d}^*_i)$.
+We form a posterior estimate of transient displacements, denoted as $\hat{u}_i$, by updating $u_i$ with the fact that $\mitbf{d}_i^*$ was realized from the random vector $\mitbf{d}_i$, that is to say $\hat{u}_i = u_i |(\mitbf{d}_i = \mitbf{d}^*_i)$.
 A general solution for $\hat{u}_i$ is derived in \citet[sec.
@@
 and
-\begin{equation}\label{eq:PosteriorCov2}
-C_{\hat{u}_i}(p,p') = C_{u_i}(p,p') - 
+\begin{multline}\label{eq:PosteriorCov2}
+C_{\hat{u}_i}(p,p') = C_{u_i}(p,p') - \\ 
                       \left[\begin{array}{cc}
@@
                           \end{array}\right].
-\end{equation}
+\end{multline}
 Taking the limit as $\kappa \to \infty$, we get the solution for the mean and covariance of $\hat{u}_i$,
@@
 
-It should be noted that we ignore any covariances between the easting and northing components, which reduces the computational complexity of our analysis because we can determine each component of the posterior independently.
-However, we are inherently assuming that the principle components of transient displacements and the noise processes are aligned with the cardinal directions, which is admittedly an arbitrary assumption.
+It should be noted that we have ignored any covariances between the easting and northing components of $\vec{u}$ and $\vec{\mitbf{d}}$.
+This simplification reduces the computational complexity of evaluating the posterior transient displacements because each component can be evaluated independently.
+However, we are inherently assuming that the principle axes describing the distribution of $\vec{u}(p)$ and $\vec{d}^{(jk)}$ are aligned with the cardinal directions, which is admittedly an arbitrary assumption.
 

### 2017-10-19 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 We make a prior assumption that each component of $\vec{u}$ is a Gaussian process with zero mean and a generic covariance function $\mathrm{Cov}\left[u_i(p),u_i(p')\right] = C_{u_i}(p,p')$.
-Using a more concise notation, we describe each component of $\vec{u}$ as $u_i \sim \mathcal{GP}\left(0,C_{u_i}\right)$.
+Using a more concise notation, we can equivalently state our prior on each component of $\vec{u}$ as $u_i \sim \mathcal{GP}\left(0,C_{u_i}\right)$.
 The function $C_{u_i}$ must be positive definite, which means that the matrix $[C_{u_i}(p,p')]_{(p,p') \in \mitbf{P} \times \mitbf{P}}$ is positive definite for any set of points $\mitbf{P}$.
@@
 4.2.4]{Rasmussen2006}.
-We also require that the derivatives $\partial^2 X_i(\vec{x},\vec{x}')/ \partial x_j \partial x_j'$ and $\partial^2 T_i(t,t') / \partial t \partial t'$ exist everywhere, which ensures that $u_i$ is spatially and temporally differentiable (See \citet[sec.
-2.2]{Adler1981} for a formal definition of differentiation for Gaussian processes).
-As an example, we can satisfy our requirements for positive definiteness and differentiability by using a squared exponential function for $X_i$ and $T_i$,
+We also require that the derivatives $\partial^2 X_i(\vec{x},\vec{x}')/ \partial x_j \partial x_j'$ and $\partial^2 T_i(t,t') / \partial t \partial t'$ exist.
+This ensures that $u_i$ is spatially and temporally differentiable, allowing us to compute $\dot\varepsilon$ (See \citet[sec.
+2.2]{Adler1981} or \citet[sec.
+10A]{Papoulis1991} for a definition of stochastic differentiation and the conditions for differentiability).
+
+We now provide an example to give our prior on $\vec{u}$ a more tangible meaning.
+We can satisfy our requirements for positive definiteness and differentiability by using a squared exponential function for $X_i$ and $T_i$,
 \begin{equation}\label{eq:ExampleCov}
@@
 \end{equation}
-The parameters $\ell$ and $\tau$ control the length-scale and time-scale, respectively, of realizations of $u_i$.
+The parameters $\ell$ and $\tau$ control the length-scale and time-scale, respectively, of realizations of $u_i$. An example realization is shown in Figure ...
+
 While the squared exponential function is a commonly used covariance function for GPR, it is not appropriate for every application.
@@
 \end{equation} 
-respectively.
+respectively \citep[sec. 10]{Papoulis1991}.
 \subsection{Transient detection criterion}\label{sec:TransientDetection}

### 2017-10-20 RevisedSubmission/hines_hetland_2017_gji.tex
add shit

@@
 Just as a Gaussian random vector is fully determined by a mean vector and a covariance matrix, a Gaussian process is fully determined by a mean function and a covariance function.
-For example, Brownian motion, $B(t)$, is a well known Gaussian process in $\mathbb{R}^1$ which has a mean function $\mathrm{E}[B(t)] = 0$ and a covariance function $\mathrm{Cov}\left[B(t),B(t')\right] = \min(t,t')$, where $t,t' \ge 0$.
+For example, Brownian motion, $B(t)$, is a well known Gaussian process in $\mathbb{R}^1$ which has the mean function $\mathrm{E}[B(t)] = 0$ and the covariance function $\mathrm{Cov}\left[B(t),B(t')\right] = \min(t,t')$, where $t,t' \ge 0$.
 If there is no information available to help choose an appropriate prior Gaussian process, then maximum likelihood methods can be used to objectively choose one that is most consistent with the data.
@@
 We seek a spatially and temporally dependent estimate of transient strain rates.
-We consider transient strain rates to be any deviation from secular strain rates, and our attention is limited to horizontal strain rates.
+We consider transient strain rates to be any deviation from secular strain rates.
 We denote transient strain rates as $\dot\varepsilon(p)$, where $p$ represents the ordered pair $(\vec{x},t)$, $\vec{x} = (x_\mathrm{e},x_\mathrm{n})$ are spatial coordinates, and $t$ is time.
 For simplicity, we assume that the study region is sufficiently small that $\vec{x}$ can be considered a point in a 2-D map projection that is aligned with the cardinal directions.
-We determine $\dot\varepsilon$ by spatially and temporally differentiating estimates of transient displacements, $\vec{u}(p) = (u_\mathrm{e},u_\mathrm{n})$.
+We determine $\dot\varepsilon$ by spatially and temporally differentiating estimates of transient displacements, which are constrained by GNSS data.
 
-We make a prior assumption that each component of $\vec{u}$ is a Gaussian process with zero mean and a generic covariance function $\mathrm{Cov}\left[u_i(p),u_i(p')\right] = C_{u_i}(p,p')$.
-Using a more concise notation, we can equivalently state our prior on each component of $\vec{u}$ as $u_i \sim \mathcal{GP}\left(0,C_{u_i}\right)$.
-The function $C_{u_i}$ must be positive definite, which means that the matrix $[C_{u_i}(p,p')]_{(p,p') \in \mitbf{P} \times \mitbf{P}}$ is positive definite for any set of points $\mitbf{P}$.
+We let $\vec{u}(p) = (u_\mathrm{e}(p),u_\mathrm{n}(p))$ be our prior understanding of transient displacements.
+Since transient displacements are not precisely known, we do not consider $\vec{u}$ to be a deterministic function.
+Instead, $\vec{u}$ is a stochastic process containing a distribution of functions that could potentially describe transient displacements.
+We let each component of $\vec{u}$ be a Gaussian process with zero mean and a generic covariance function $\mathrm{Cov}\left[u_i(p),u_i(p')\right] = C_{u_i}(p,p')$.
+Using a more concise notation, we write our prior on each component of $\vec{u}$ as $u_i \sim \mathcal{GP}\left(0,C_{u_i}\right)$.
+The function $C_{u_i}$ must be positive definite in order to be a valid covariance function.
+By definition, $C_{u_i}$ is a positive definite function if the matrix $[C_{u_i}(p,p')]_{(p,p') \in \mitbf{P} \times \mitbf{P}}$ is positive definite for any set of points $\mitbf{P}$ \citep[sec.
+2.5]{Cressie1992}.
 We assume that $C_{u_i}$ can be separated into spatial and temporal functions as
 \begin{equation}\label{eq:TransientCovariance}
-C_{u_i}(p,p') = C_{u_i}\left((\vec{x},t),(\vec{x}',t')\right) = \phi^2 X_i(\vec{x},\vec{x}')T_i(t,t'),
+C_{u_i}(p,p') = C_{u_i}\left((\vec{x},t),(\vec{x}',t')\right) = X_i(\vec{x},\vec{x}')T_i(t,t').
 \end{equation}   
-where the parameter $\phi$ controls the amplitude of realizations of $u_i$.
-As long as the functions $X_i$ and $T_i$ are positive definite, $C_{u_i}$ is guaranteed to also be positive definite and thus a valid covariance function \citep[sec.
+As long as the functions $X_i$ and $T_i$ are positive definite, $C_{u_i}$ is guaranteed to also be positive definite \citep[sec.
 4.2.4]{Rasmussen2006}.
 We also require that the derivatives $\partial^2 X_i(\vec{x},\vec{x}')/ \partial x_j \partial x_j'$ and $\partial^2 T_i(t,t') / \partial t \partial t'$ exist.
-This ensures that $u_i$ is spatially and temporally differentiable, allowing us to compute $\dot\varepsilon$ (See \citet[sec.
+This ensures that $\vec{u}$ is spatially and temporally differentiable, allowing us to compute $\dot\varepsilon$ (See \citet[sec.
 2.2]{Adler1981} or \citet[sec.
@@
 
-We now provide an example to give our prior on $\vec{u}$ a more tangible meaning.
-We can satisfy our requirements for positive definiteness and differentiability by using a squared exponential function for $X_i$ and $T_i$,
+To give an example of $u_i$, we can use a squared exponential function for $X_i$ and $T_i$,
 \begin{equation}\label{eq:ExampleCov}
-X_i(\vec{x},\vec{x}') = \exp\left(\frac{-||\vec{x} - \vec{x}'||_2^2}{2\ell^2}\right), \ \ \ \ T_i(t,t') = \exp\left(\frac{-|t - t'|^2}{2\tau^2}\right).
+X_i(\vec{x},\vec{x}') = \exp\left(\frac{-||\vec{x} - \vec{x}'||_2^2}{2\ell^2}\right), \ \ \ \ T_i(t,t') = \phi^2 \exp\left(\frac{-|t - t'|^2}{2\tau^2}\right),
 \end{equation}
-The parameters $\ell$ and $\tau$ control the length-scale and time-scale, respectively, of realizations of $u_i$. An example realization is shown in Figure ...
-
+which satisfies our requirements for positive definiteness and differentiability.
+The parameters $\ell$ and $\tau$ control the length-scale and time-scale, respectively, of realizations (i.e., sample functions) of $u_i$.
+The parameter $\phi$, which we have arbitrarily chosen to incorporate into $T_i$ rather than $X_i$, controls the amplitude of realizations of $u_i$.
+Ideally, we want realizations of $u_i$ to have a length-scale, time-scale, and amplitude that resembles what we expect from true signal.
+A few realizations for $u_i$ are shown in Figure XXX.
 While the squared exponential function is a commonly used covariance function for GPR, it is not appropriate for every application.
@@
                                         \frac{\partial \hat{u}_i(p)}{\partial x_j} +  
-                                        \frac{\partial \hat{u}_j(p)}{\partial x_i}\right).
+                                        \frac{\partial \hat{u}_j(p)}{\partial x_i} \right).
 \end{equation}
@@
 (\ref{eq:StrainRate}) is a linear operation on the Gaussian processes $\hat{u}_i$ and $\hat{u}_j$, we know that $\dot\varepsilon_{ij}$ is also a Gaussian process.
-The mean and covariance functions for $\dot{\varepsilon}_{ij}$ are given by
-\begin{equation}\label{eq:StrainMean}
-\mu_{\dot\varepsilon_{ij}}(p) = \frac{1}{2}\frac{\partial}{\partial t}\left(
-                                  \frac{\partial \mu_{\hat{u}_i}(p)}{\partial x_j} + 
-                                  \frac{\partial \mu_{\hat{u}_j}(p)}{\partial x_i} \right)
-\end{equation} 
+From \citet[sec.
+10]{Papoulis1991}, we find that the mean and covariance functions for $\dot{\varepsilon}_{ij}$ are
+\begin{align}\label{eq:StrainMean}
+\mu_{\dot\varepsilon_\mathrm{ee}}(p) &= \frac{\partial^2 \mu_{\hat{u}_\mathrm{e}}(p)}{\partial t \, \partial x_\mathrm{e}} \\
+\mu_{\dot\varepsilon_\mathrm{nn}}(p) &= \frac{\partial^2 \mu_{\hat{u}_\mathrm{n}}(p)}{\partial t \, \partial x_\mathrm{n}} \\
+\mu_{\dot\varepsilon_\mathrm{en}}(p) &= \frac{1}{2}\frac{\partial}{\partial t}\left(
+                                        \frac{\partial \mu_{\hat{u}_\mathrm{e}}(p)}{\partial x_\mathrm{n}} + 
+                                        \frac{\partial \mu_{\hat{u}_\mathrm{n}}(p)}{\partial x_\mathrm{e}} \right)                                  
+\end{align} 
 and
-\begin{equation}\label{eq:StrainCov}
-C_{\dot\varepsilon_{ij}}(p,p') = \frac{1}{4} \frac{\partial^2}{\partial t \, \partial t'}\left(
-                                   \frac{\partial^2 C_{\hat{u}_i}(p,p')}{\partial x_j \, \partial x'_j} + 
-                                   \frac{\partial^2 C_{\hat{u}_j}(p,p')}{\partial x_i \, \partial x'_i} \right),
-\end{equation} 
-respectively \citep[sec. 10]{Papoulis1991}.
+\begin{align}\label{eq:StrainCov}
+C_{\dot\varepsilon_\mathrm{ee}}(p,p') &= \frac{\partial^4 C_{\hat{u}_\mathrm{e}}(p,p')}{\partial t \, \partial t' \, \partial x_\mathrm{e} \, \partial x'_\mathrm{e}} \\
+C_{\dot\varepsilon_\mathrm{nn}}(p,p') &= \frac{\partial^4 C_{\hat{u}_\mathrm{n}}(p,p')}{\partial t \, \partial t' \, \partial x_\mathrm{n} \, \partial x'_\mathrm{n}} \\
+C_{\dot\varepsilon_\mathrm{en}}(p,p') &= \frac{1}{4} \frac{\partial^2}{\partial t \, \partial t'}\left(
+                                         \frac{\partial^2 C_{\hat{u}_\mathrm{e}}(p,p')}{\partial x_\mathrm{n} \, \partial x'_\mathrm{n}} + 
+                                         \frac{\partial^2 C_{\hat{u}_\mathrm{n}}(p,p')}{\partial x_\mathrm{e} \, \partial x'_\mathrm{e}} \right),
+\end{align} 
+respectively.
 \subsection{Transient detection criterion}\label{sec:TransientDetection}

### 2017-10-21 RevisedSubmission/compile.sh
started making macros

@@ -1,0 +1,14 @@
+#!/bin/bash
+#pdflatex -synctex=1 -interaction=nonstopmode dissertation.tex
+#bibtex ch1/ch1.aux
+#bibtex ch2/ch2.aux
+#bibtex ch3/ch3.aux
+#bibtex ch4/ch4.aux
+#bibtex ch5/ch5.aux
+#bibtex ch6/ch6.aux
+#pdflatex -synctex=1 -interaction=nonstopmode dissertation.tex
+
+pdflatex -synctex=1 -interaction=nonstopmode hines_hetland_2017_gji.tex
+bibtex hines_hetland_2017_gji.aux
+pdflatex -synctex=1 -interaction=nonstopmode hines_hetland_2017_gji.tex
+pdflatex -synctex=1 -interaction=nonstopmode hines_hetland_2017_gji.tex

### 2017-10-21 RevisedSubmission/hines_hetland_2017_gji.tex
justified text file

@@
 
+%% DEFINE MATH MODE MACROS
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newcommand{\pos}{\vec{x}} % position vector
+\newcommand{\data}{\mitbf{d}} % data vector
+\newcommand{\post}{\hat{u}} % posterior
+\newcommand{\points}{\mitbf{P}} % observation points
+\newcommand{\strain}{\dot\varepsilon} % strain rate
+\newcommand{\E}[1]{\e\left[ #1 \right]} % Expected value
+\newcommand{\Cov}[1]{\mathrm{Cov}\left[ #1 \right]} % Expected value
+\newcommand{\e}{\mathrm{e}} % east subscript
+\newcommand{\n}{\mathrm{n}} % north subscript
+\newcommand{\G}{\mitbf{G}} % basis functions
+\newcommand{\zeros}{\mitbf{0}} % zeros matrix
+\newcommand{\eye}{\mitbf{I}} % identity matrix
+
+%% SUMMARY
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{summary}
-Transient strain rates derived from GNSS data can be used to detect and understand geophysical phenomena such as slow slip events and postseismic deformation.
-Here we propose using Gaussian process regression (GPR) as a tool for estimating transient strain rates from GNSS data.
-GPR is a non-parametric, Bayesian method for interpolating scattered data.
-Transient strain rates estimated with GPR have meaningful uncertainties, allowing geophysical signal to be easily discerned from noise.
-In our approach, we assume a stochastic prior model for transient displacements.
-The prior describes how much one expects transient displacements to covary spatially and temporally.
-A posterior estimate of transient strain rates is obtained by differentiating the posterior displacements.
-One limitation with GPR is that it is not robust against outliers, so we introduce a pre-processing method for detecting and removing outliers in GNSS data.
-As a demonstration, we use GPR to detect transient strain resulting from slow slip events in the Pacific Northwest.
-Maximum likelihood methods are used to constrain a prior model for transient displacements in this region.
-The temporal covariance of our prior model is described by a compact Wendland covariance function, which significantly reduces the computational burden that can be associated with GPR.
-Our results reveal the spatial and temporal evolution of strain from slow slip events.
-We verify that the transient strain estimated with GPR is in fact geophysical signal by comparing it to the seismic tremor that is associated with Pacific Northwest slow slip events.
+
+Transient strain rates derived from GNSS data can be used to detect
+and understand geophysical phenomena such as slow slip events and
+postseismic deformation. Here we propose using Gaussian process
+regression (GPR) as a tool for estimating transient strain rates from
+GNSS data. GPR is a non-parametric, Bayesian method for interpolating
+scattered data. Transient strain rates estimated with GPR have
+meaningful uncertainties, allowing geophysical signal to be easily
+discerned from noise. In our approach, we assume a stochastic prior
+model for transient displacements. The prior describes how much one
+expects transient displacements to covary spatially and temporally. A
+posterior estimate of transient strain rates is obtained by
+differentiating the posterior displacements. One limitation with GPR
+is that it is not robust against outliers, so we introduce a
+pre-processing method for detecting and removing outliers in GNSS
+data. As a demonstration, we use GPR to detect transient strain
+resulting from slow slip events in the Pacific Northwest. Maximum
+likelihood methods are used to constrain a prior model for transient
+displacements in this region. The temporal covariance of our prior
+model is described by a compact Wendland covariance function, which
+significantly reduces the computational burden that can be associated
+with GPR. Our results reveal the spatial and temporal evolution of
+strain from slow slip events. We verify that the transient strain
+estimated with GPR is in fact geophysical signal by comparing it to
+the seismic tremor that is associated with Pacific Northwest slow slip
+events.
+
 \end{summary}
@@
 
+%% INTRODUCTION
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}\label{sec:Introduction}
-Crustal strain rates are fundamentally important quantities for assessing seismic hazard.
-Knowing where and how quickly strain is accumulating gives insight into where we can expect stored elastic energy to be released seismically.
-Consequently, secular crustal strain rates estimated from GNSS data have been used to constrain seismic hazard models such as UCERF3 \citep{Field2014}.
-Transient crustal strain, which is caused by geophysical phenomena such as slow slip events (SSEs) or postseismic deformation, is also relevant for assessing seismic hazard.
-While transient strain itself is not damaging, there is a risk that it can trigger major earthquakes \citep{Roeloffs2006,Freed2001}.
-Dense networks of continuous GNSS stations, such as the Plate Boundary Observatory (PBO), make it possible to identify transient strain with high fidelity.
-Developing and improving upon methods for deriving secular and transient strain from GNSS data is an active area of research.
-
-Most methods for estimating strain rates from GNSS data assume some parametric form for the deformation signal.
-The simplest method for estimating secular strain rates assumes that GNSS derived velocities can be described with a first-order polynomial (i.e, the deformation gradients are constant) over some subnetwork of the GNSS stations \citep[e.g.,][]{Feigl1990,Murray2000}.
-The components of the strain rate tensor for each subnetwork are then determined from the least squares fit to the observations.
-The assumption that deformation gradients are spatially uniform is not appropriate when subnetworks span too large of an area.
-To help overcome this deficiency, \citet{Shen1996,Shen2015} used an inverse distance weighting scheme, in which the estimated strain rate at a point is primarily controlled by observations at nearby stations.
-The method  of \citet{Shen1996,Shen2015} can be viewed as a form of local least squares regression with a first-order polynomial \citep[e.g.,][sec.
-6]{Hastie2009}.
-Other methods for estimating secular strain rates have parameterized GNSS derived velocities with bi-cubic splines \citep{Beavan2001}, spherical wavelets \citep{Tape2009}, and elastostatic Green's functions \citep{Sandwell2016}.
-The type of basis functions and the number of degrees of freedom for a parameterization, which are often chosen subjectively, have a strong influence on the strain solution.
-If there are too few degrees of freedom in the parameterization, then estimated strain rates will be biased and the uncertainties will be underestimated.
-On the other hand, if there are too many degrees of freedom, then there will not be any coherent features in the estimated strain rates.
-The methods described by \citet{Beavan2001} and \citet{Tape2009} also require the user to specify penalty parameters that control a similar trade-off between bias and variance in the solution.
-One could parameterize deformation with a physically motivated model of interseismic deformation \citep[e.g.,][]{Meade2005,McCaffrey2007}.
-In such models the lithospheric rheology and fault geometries are assumed to be known.
-Any errors in the assumed physical model could result in biased strain estimates and underestimated formal uncertainties.
-
-The aforementioned studies are concerned with estimating secular strain rates.
-In recent years the Southern California Earthquake Center (SCEC) community has shown interest in developing methods for detecting transient strain.
-SCEC supported a transient detection exercise \citep{Lohman2013}, where several research groups tested their methods for detecting transient geophysical signal with a synthetic GNSS dataset.
-Among the methods tested were the Network Strain Filter (NSF) \citep{Ohtani2010} and the Network Inversion Filter (NIF) \citep{Segall1997}.
-The NSF uses a wavelet parameterization to  describe the spatial component of geophysical signal.
-The NIF, which is intended for imaging slow fault slip from geodetic data, uses the elastic dislocation Green's functions from \citet{Okada1992}.
-For the NSF and NIF, the time dependence of the geophysical signal is modeled as integrated Brownian motion.
-The method described in \citet{Holt2013} was also tested in the SCEC transient detection exercise, which calculates strain rates using a bi-cubic spatial parameterization of displacements between time epochs.
-\citet{Holt2013} defined a detection threshold based on the strain rate magnitude, and below we demonstrate that this is indeed an effective criterion for identifying geophysical signal.
-For the same reasons descibed above, the transient deformation and corresponding uncertainties estimated by these methods can be biased by the chosen spatial parameterization.
-It is then difficult to distinguish signal from noise with these methods, which limits their utility for transient detection.
-
-Here we propose using Gaussian process regression (GPR) \citep{Rasmussen2006} to estimate transient strain from GNSS data.
-GPR is closely related to kriging \citep{Cressie1992} and least squares collocation \citep{Moritz1978}.
-The latter has been used by \citet{Kato1998} and \citet{El-Fiky1999} to estimate secular strain rates from GNSS data.
-GPR is a Bayesian, non-parametric method for inferring a continuous signal from scattered data.
-Since GNSS stations are irregularly spaced and observation times may differ between stations, GPR is an ideal tool for synthesizing discrete GNSS data into a spatially and temporally continuous representation of surface deformation.
-GPR is Bayesian in that we describe our prior understanding of the geophysical signal with a Gaussian process.
-A Gaussian process is a continuous stochastic process whose value at any collection of points can be described as a Gaussian random vector.
-Just as a Gaussian random vector is fully determined by a mean vector and a covariance matrix, a Gaussian process is fully determined by a mean function and a covariance function.
-For example, Brownian motion, $B(t)$, is a well known Gaussian process in $\mathbb{R}^1$ which has the mean function $\mathrm{E}[B(t)] = 0$ and the covariance function $\mathrm{Cov}\left[B(t),B(t')\right] = \min(t,t')$, where $t,t' \ge 0$.
-If there is no information available to help choose an appropriate prior Gaussian process, then maximum likelihood methods can be used to objectively choose one that is most consistent with the data.
-Once a prior is selected, we incorporate GNSS observations with the prior to form a posterior estimate of transient strain.
-The posterior transient strain is also a Gaussian process, and we can use its distribution to confidently discern geophysical signal from noise.
-We use GPR to infer strain resulting from SSEs in the Pacific Northwest, demonstrating that GPR is an effective tool for detecting transient geophysical processes.
-
+
+Crustal strain rates are fundamentally important quantities for
+assessing seismic hazard. Knowing where and how quickly strain is
+accumulating gives insight into where we can expect stored elastic
+energy to be released seismically. Consequently, secular crustal
+strain rates estimated from GNSS data have been used to constrain
+seismic hazard models such as UCERF3 \citep{Field2014}. Transient
+crustal strain, which is caused by geophysical phenomena such as slow
+slip events (SSEs) or postseismic deformation, is also relevant for
+assessing seismic hazard. While transient strain itself is not
+damaging, there is a risk that it can trigger major earthquakes
+\citep{Roeloffs2006,Freed2001}. Dense networks of continuous GNSS
+stations, such as the Plate Boundary Observatory (PBO), make it
+possible to identify transient strain with high fidelity. Developing
+and improving upon methods for deriving secular and transient strain
+from GNSS data is an active area of research.
+
+Most methods for estimating strain rates from GNSS data assume some
+parametric form for the deformation signal. The simplest method for
+estimating secular strain rates assumes that GNSS derived velocities
+can be described with a first-order polynomial (i.e, the deformation
+gradients are constant) over some subnetwork of the GNSS stations
+\citep[e.g.,][]{Feigl1990,Murray2000}. The components of the strain
+rate tensor for each subnetwork are then determined from the least
+squares fit to the observations. The assumption that deformation
+gradients are spatially uniform is not appropriate when subnetworks
+span too large of an area. To help overcome this deficiency,
+\citet{Shen1996,Shen2015} used an inverse distance weighting scheme,
+in which the estimated strain rate at a point is primarily controlled
+by observations at nearby stations. The method of
+\citet{Shen1996,Shen2015} can be viewed as a form of local least
+squares regression with a first-order polynomial \citep[e.g.,][sec.
+6]{Hastie2009}. Other methods for estimating secular strain rates have
+parameterized GNSS derived velocities with bi-cubic splines
+\citep{Beavan2001}, spherical wavelets \citep{Tape2009}, and
+elastostatic Green's functions \citep{Sandwell2016}. The type of basis
+functions and the number of degrees of freedom for a parameterization,
+which are often chosen subjectively, have a strong influence on the
+strain solution. If there are too few degrees of freedom in the
+parameterization, then estimated strain rates will be biased and the
+uncertainties will be underestimated. On the other hand, if there are
+too many degrees of freedom, then there will not be any coherent
+features in the estimated strain rates. The methods described by
+\citet{Beavan2001} and \citet{Tape2009} also require the user to
+specify penalty parameters that control a similar trade-off between
+bias and variance in the solution. One could parameterize deformation
+with a physically motivated model of interseismic deformation
+\citep[e.g.,][]{Meade2005,McCaffrey2007}. In such models the
+lithospheric rheology and fault geometries are assumed to be known.
+Any errors in the assumed physical model could result in biased strain
+estimates and underestimated formal uncertainties.
+
+The aforementioned studies are concerned with estimating secular
+strain rates. In recent years the Southern California Earthquake
+Center (SCEC) community has shown interest in developing methods for
+detecting transient strain. SCEC supported a transient detection
+exercise \citep{Lohman2013}, where several research groups tested
+their methods for detecting transient geophysical signal with a
+synthetic GNSS dataset. Among the methods tested were the Network
+Strain Filter (NSF) \citep{Ohtani2010} and the Network Inversion
+Filter (NIF) \citep{Segall1997}. The NSF uses a wavelet
+parameterization to describe the spatial component of geophysical
+signal. The NIF, which is intended for imaging slow fault slip from
+geodetic data, uses the elastic dislocation Green's functions from
+\citet{Okada1992}. For the NSF and NIF, the time dependence of the
+geophysical signal is modeled as integrated Brownian motion. The
+method described in \citet{Holt2013} was also tested in the SCEC
+transient detection exercise, which calculates strain rates using a
+bi-cubic spatial parameterization of displacements between time
+epochs. \citet{Holt2013} defined a detection threshold based on the
+strain rate magnitude, and below we demonstrate that this is indeed an
+effective criterion for identifying geophysical signal. For the same
+reasons descibed above, the transient deformation and corresponding
+uncertainties estimated by these methods can be biased by the chosen
+spatial parameterization. It is then difficult to distinguish signal
+from noise with these methods, which limits their utility for
+transient detection.
+
+Here we propose using Gaussian process regression (GPR)
+\citep{Rasmussen2006} to estimate transient strain from GNSS data. GPR
+is closely related to kriging \citep{Cressie1992} and least squares
+collocation \citep{Moritz1978}. The latter has been used by
+\citet{Kato1998} and \citet{El-Fiky1999} to estimate secular strain
+rates from GNSS data. GPR is a Bayesian, non-parametric method for
+inferring a continuous signal from scattered data. Since GNSS stations
+are irregularly spaced and observation times may differ between
+stations, GPR is an ideal tool for synthesizing discrete GNSS data
+into a spatially and temporally continuous representation of surface
+deformation. GPR is Bayesian in that we describe our prior
+understanding of the geophysical signal with a Gaussian process. A
+Gaussian process is a continuous stochastic process whose value at any
+collection of points can be described as a Gaussian random vector.
+Just as a Gaussian random vector is fully determined by a mean vector
+and a covariance matrix, a Gaussian process is fully determined by a
+mean function and a covariance function. For example, Brownian motion,
+$B(t)$, is a well known Gaussian process in $\mathbb{R}^1$ which has
+the mean function $\E{B(t)} = 0$ and the covariance function
+$\Cov{B(t),B(t')} = \min(t,t')$, where $t,t' \ge 0$. If there is no
+information available to help choose an appropriate prior Gaussian
+process, then maximum likelihood methods can be used to objectively
+choose one that is most consistent with the data. Once a prior is
+selected, we incorporate GNSS observations with the prior to form a
+posterior estimate of transient strain. The posterior transient strain
+is also a Gaussian process, and we can use its distribution to
+confidently discern geophysical signal from noise. We use GPR to infer
+strain resulting from SSEs in the Pacific Northwest, demonstrating
+that GPR is an effective tool for detecting transient geophysical
+processes.
+
+%% ESTIMATING TRANSIENT STRAIN RATES
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Estimating Transient Strain Rates}\label{sec:Method}
-We seek a spatially and temporally dependent estimate of transient strain rates.
-We consider transient strain rates to be any deviation from secular strain rates.
-We denote transient strain rates as $\dot\varepsilon(p)$, where $p$ represents the ordered pair $(\vec{x},t)$, $\vec{x} = (x_\mathrm{e},x_\mathrm{n})$ are spatial coordinates, and $t$ is time.
-For simplicity, we assume that the study region is sufficiently small that $\vec{x}$ can be considered a point in a 2-D map projection that is aligned with the cardinal directions.
-We determine $\dot\varepsilon$ by spatially and temporally differentiating estimates of transient displacements, which are constrained by GNSS data.
-
-We let $\vec{u}(p) = (u_\mathrm{e}(p),u_\mathrm{n}(p))$ be our prior understanding of transient displacements.
-Since transient displacements are not precisely known, we do not consider $\vec{u}$ to be a deterministic function.
-Instead, $\vec{u}$ is a stochastic process containing a distribution of functions that could potentially describe transient displacements.
-We let each component of $\vec{u}$ be a Gaussian process with zero mean and a generic covariance function $\mathrm{Cov}\left[u_i(p),u_i(p')\right] = C_{u_i}(p,p')$.
-Using a more concise notation, we write our prior on each component of $\vec{u}$ as $u_i \sim \mathcal{GP}\left(0,C_{u_i}\right)$.
-The function $C_{u_i}$ must be positive definite in order to be a valid covariance function.
-By definition, $C_{u_i}$ is a positive definite function if the matrix $[C_{u_i}(p,p')]_{(p,p') \in \mitbf{P} \times \mitbf{P}}$ is positive definite for any set of points $\mitbf{P}$ \citep[sec.
-2.5]{Cressie1992}.
-We assume that $C_{u_i}$ can be separated into spatial and temporal functions as
+
+We seek a spatially and temporally dependent estimate of transient
+strain rates. We consider transient strain rates to be any deviation
+from secular strain rates. We denote transient strain rates as
+$\strain(p)$, where $p$ represents the ordered pair $(\pos,t)$, $\pos
+= (x_\e,x_\n)$ are spatial coordinates, and $t$ is time. For
+simplicity, we assume that the study region is sufficiently small that
+$\pos$ can be considered a point in a 2-D map projection that is
+aligned with the cardinal directions. We determine $\strain$ by
+spatially and temporally differentiating estimates of transient
+displacements, which are constrained by GNSS data.
+
+We let $\vec{u}(p) = (u_\e(p),u_\n(p))$ be our prior understanding of
+transient displacements. Since transient displacements are not
+precisely known, we do not consider $\vec{u}$ to be a deterministic
+function. Instead, $\vec{u}$ is a stochastic process containing a
+distribution of functions that could potentially describe transient
+displacements. We let each component of $\vec{u}$ be a Gaussian
+process with zero mean and a generic covariance function
+$\Cov{u_i(p),u_i(p')} = C_{u_i}(p,p')$. Using a more concise notation,
+we write our prior on each component of $\vec{u}$ as $u_i \sim
+\mathcal{GP}\left(0,C_{u_i}\right)$. The function $C_{u_i}$ must be
+positive definite in order to be a valid covariance function. By
+definition, $C_{u_i}$ is a positive definite function if the matrix
+$[C_{u_i}(p,p')]_{(p,p') \in \points \times \points}$ is positive
+definite for any set of points $\points$ \citep[sec.
+2.5]{Cressie1992}. We assume that $C_{u_i}$ can be separated into
+spatial and temporal functions as
 \begin{equation}\label{eq:TransientCovariance}
-C_{u_i}(p,p') = C_{u_i}\left((\vec{x},t),(\vec{x}',t')\right) = X_i(\vec{x},\vec{x}')T_i(t,t').
-\end{equation}   
-As long as the functions $X_i$ and $T_i$ are positive definite, $C_{u_i}$ is guaranteed to also be positive definite \citep[sec.
-4.2.4]{Rasmussen2006}.
-We also require that the derivatives $\partial^2 X_i(\vec{x},\vec{x}')/ \partial x_j \partial x_j'$ and $\partial^2 T_i(t,t') / \partial t \partial t'$ exist.
-This ensures that $\vec{u}$ is spatially and temporally differentiable, allowing us to compute $\dot\varepsilon$ (See \citet[sec.
-2.2]{Adler1981} or \citet[sec.
-10A]{Papoulis1991} for a definition of stochastic differentiation and the conditions for differentiability).
-
-To give an example of $u_i$, we can use a squared exponential function for $X_i$ and $T_i$,
+C_{u_i}(p,p') = C_{u_i}\left((\pos,t),(\pos',t')\right) 
+              = X_i(\pos,\pos')T_i(t,t').
+\end{equation}   
+As long as the functions $X_i$ and $T_i$ are positive definite,
+$C_{u_i}$ is guaranteed to also be positive definite \citep[sec.
+4.2.4]{Rasmussen2006}. We also require that the derivatives
+$\partial^2 X_i(\pos,\pos')/ \partial x_j \partial x_j'$ and
+$\partial^2 T_i(t,t') / \partial t \partial t'$ exist. This ensures
+that $\vec{u}$ is spatially and temporally differentiable, allowing us
+to compute $\strain$ (See \citet[sec. 2.2]{Adler1981} or \citet[sec.
+10A]{Papoulis1991} for a definition of stochastic differentiation and
+the conditions for differentiability).
+
+To give an example of $u_i$, we can use a squared exponential function
+for $X_i$ and $T_i$,
 \begin{equation}\label{eq:ExampleCov}
-X_i(\vec{x},\vec{x}') = \exp\left(\frac{-||\vec{x} - \vec{x}'||_2^2}{2\ell^2}\right), \ \ \ \ T_i(t,t') = \phi^2 \exp\left(\frac{-|t - t'|^2}{2\tau^2}\right),
-\end{equation}
-which satisfies our requirements for positive definiteness and differentiability.
-The parameters $\ell$ and $\tau$ control the length-scale and time-scale, respectively, of realizations (i.e., sample functions) of $u_i$.
-The parameter $\phi$, which we have arbitrarily chosen to incorporate into $T_i$ rather than $X_i$, controls the amplitude of realizations of $u_i$.
-Ideally, we want realizations of $u_i$ to have a length-scale, time-scale, and amplitude that resembles what we expect from true signal.
-A few realizations for $u_i$ are shown in Figure XXX.
-While the squared exponential function is a commonly used covariance function for GPR, it is not appropriate for every application.
-The appropriate choice for $X_i$ and $T_i$ may vary depending on the geophysical signal we are trying to describe.
-To keep this section sufficiently general, we hold off on specifying $X_i$ and $T_i$ until Section \ref{sec:SignalModel}, where we estimate transient strain from slow slip events in the Pacific Northwest.
-
-GNSS data records transient displacements as well as other physical and non-physical processes that we are not interested in.
-We first consider component $i$ of a single GNSS displacement observation made at station $j$, which is located at $\vec{x}^{(j)}$, and time $t^{(k)}$.
-We describe this observation, $d_i^{*(jk)}$, as a realization of the random variable
+X_i(\pos,\pos') = \exp\left(\frac{-||\pos - \pos'||_2^2}{2\ell^2}\right), \ \ \ \ 
+T_i(t,t') = \phi^2 \exp\left(\frac{-|t - t'|^2}{2\tau^2}\right),
+\end{equation}
+which satisfies our requirements for positive definiteness and
+differentiability. The parameters $\ell$ and $\tau$ control the
+length-scale and time-scale, respectively, of realizations (i.e.,
+sample functions) of $u_i$. The parameter $\phi$, which we have
+arbitrarily chosen to incorporate into $T_i$ rather than $X_i$,
+controls the amplitude of realizations of $u_i$. Ideally, we want
+realizations of $u_i$ to have a length-scale, time-scale, and
+amplitude that resembles what we expect from true signal. A few
+realizations for $u_i$ are shown in Figure XXX. While the squared
+exponential function is a commonly used covariance function for GPR,
+it is not appropriate for every application. The appropriate choice
+for $X_i$ and $T_i$ may vary depending on the geophysical signal we
+are trying to describe. To keep this section sufficiently general, we
+hold off on specifying $X_i$ and $T_i$ until Section
+\ref{sec:SignalModel}, where we estimate transient strain from slow
+slip events in the Pacific Northwest.
+
+GNSS data records transient displacements as well as other physical
+and non-physical processes that we are not interested in. We first
+consider component $i$ of a single GNSS displacement observation made
+at station $j$, which is located at $\pos^{(j)}$, and time $t^{(k)}$.
+We describe this observation, $d_i^{*(jk)}$, as a realization of the
+random variable
 \begin{align}\label{eq:Data}
 \begin{split}
-d_i^{(jk)} = &u_i\left(\vec{x}^{(j)},t^{(k)}\right) + 
+d_i^{(jk)} = &u_i\left(\pos^{(j)},t^{(k)}\right) + 
               \eta_i^{(jk)} + 
@@
 \end{align}
-where $\eta_i^{(jk)}$ describes noise, $a_i^{(j)}$ is an offset that is unique for each station, $b_i^{(j)}$ is the secular velocity at $\vec{x}^{(j)}$, and the sinusoids describe seasonal deformation (using units of years for $t^{(k)}$).
-We then consider the column vector of $n$ GNSS displacement observations, $\mitbf{d}^*_i$, where the subscript indicates that we are still only considering component $i$ of displacements.
-The observations are made at $m$ stations, and the times and positions for each observation are described by the set $\mitbf{P}$.
-$\mitbf{d}^*_i$ can be considered a realization of the random vector $\mitbf{d}_i$, which is formed by evaluating eq.
-(\ref{eq:Data}) at each point in $\mitbf{P}$.
-To write out $\mitbf{d}_i$ explicitly, we let $\mitbf{G}$ be an $n \times 6m$ matrix consisting of the basis functions from eq.
-(\ref{eq:Data}) (i.e., the linear trends and sinusoids for each station) evaluated at each point in $\mitbf{P}$.
-The coefficients corresponding to each basis function are collected into the column vector $\mitbf{m}_i$.
-The noise for all the observations are described by the column vector $\mitbf{\eta}_i$.
-We can then write $\mitbf{d}_i$ as
+where $\eta_i^{(jk)}$ describes noise, $a_i^{(j)}$ is an offset that
+is unique for each station, $b_i^{(j)}$ is the secular velocity at
+$\pos^{(j)}$, and the sinusoids describe seasonal deformation (using
+units of years for $t^{(k)}$). We then consider the column vector of
+$n$ GNSS displacement observations, $\data^*_i$, where the subscript
+indicates that we are still only considering component $i$ of
+displacements. The observations are made at $m$ stations, and the
+times and positions for each observation are described by the set
+$\points$. $\data^*_i$ can be considered a realization of the random
+vector $\data_i$, which is formed by evaluating eq. (\ref{eq:Data}) at
+each point in $\points$. To write out $\data_i$ explicitly, we let
+$\G$ be an $n \times 6m$ matrix consisting of the basis functions from
+eq. (\ref{eq:Data}) (i.e., the linear trends and sinusoids for each
+station) evaluated at each point in $\points$. The coefficients
+corresponding to each basis function are collected into the column
+vector $\mitbf{m}_i$. The noise for all the observations are described
+by the column vector $\mitbf{\eta}_i$. We can then write $\data_i$ as
 \begin{equation}\label{eq:DataVec}
-\mitbf{d}_i = u_i(\mitbf{P}) + \mitbf{\eta}_i + \mitbf{G}\mitbf{m}_i,
-\end{equation}
-where the notation $u_i(\mitbf{P})$ represents the column vector $[u_i(p)]_{p \in \mitbf{P}}$.
-
-We assume a diffuse prior for the components of $\mitbf{m}_i$, that is to say $\mitbf{m}_i \sim \mathcal{N}(\mitbf{0},\kappa^2\mitbf{I})$ in the limit as $\kappa \to \infty$.
-Of course, the secular velocities, $b_i^{(j)}$, are spatially correlated and we could invoke a tectonic model to form a prior on $b_i^{(j)}$.
-However, in our application to the Pacific Northwest, we will be using displacement time series which are long enough to sufficiently constrain $b_i^{(j)}$ for each station, avoiding the need to incorporate a prior.
-Likewise, seasonal deformation is spatially correlated \citep{Dong2002,Langbein2008}, and it may be worth exploring and exploiting such a correlation in a future study.
-We assume that $\mitbf{\eta}_i$ is a spatially and/or temporally correlated random vector distributed as $\mathcal{N}(\mitbf{0},\mitbf{C}_{\eta_i})$.
-For example, $\mitbf{\eta}_i$ can be uncorrelated white noise, temporally correlated noise describing benchmark wobble \citep[e.g.,][]{Wyatt1982,Wyatt1989}, and/or spatially correlated noise describing common mode error \citep[e.g.,][]{Wdowinski1997}.
-The appropriate noise model may vary depending on the application, and we hold off on specifying the covariance matrix, $\mitbf{C}_{\eta_i}$, until Section \ref{sec:NoiseModel}.
-We are now able to write the distribution of $\mitbf{d}_i$ as
+\data_i = u_i(\points) + \mitbf{\eta}_i + \G\mitbf{m}_i,
+\end{equation}
+where the notation $u_i(\points)$ represents the column vector
+$[u_i(p)]_{p \in \points}$.
+
+We assume a diffuse prior for the components of $\mitbf{m}_i$, that is
+to say $\mitbf{m}_i \sim \mathcal{N}(\zeros,\kappa^2\eye)$ in the
+limit as $\kappa \to \infty$. Of course, the secular velocities,
+$b_i^{(j)}$, are spatially correlated and we could invoke a tectonic
+model to form a prior on $b_i^{(j)}$. However, in our application to
+the Pacific Northwest, we will be using displacement time series which
+are long enough to sufficiently constrain $b_i^{(j)}$ for each
+station, avoiding the need to incorporate a prior. Likewise, seasonal
+deformation is spatially correlated \citep{Dong2002,Langbein2008}, and
+it may be worth exploring and exploiting such a correlation in a
+future study. We assume that $\mitbf{\eta}_i$ is a spatially and/or
+temporally correlated random vector distributed as
+$\mathcal{N}(\zeros,\mitbf{C}_{\eta_i})$. For example,
+$\mitbf{\eta}_i$ can be uncorrelated white noise, temporally
+correlated noise describing benchmark wobble
+\citep[e.g.,][]{Wyatt1982,Wyatt1989}, and/or spatially correlated
+noise describing common mode error \citep[e.g.,][]{Wdowinski1997}. The
+appropriate noise model may vary depending on the application, and we
+hold off on specifying the covariance matrix, $\mitbf{C}_{\eta_i}$,
+until Section \ref{sec:NoiseModel}. We are now able to write the
+distribution of $\data_i$ as
 \begin{equation}\label{eq:DataDist}
-\mitbf{d}_i \sim \mathcal{N}\left(\mitbf{0}, C_{u_i}(\mitbf{P},\mitbf{P}) + \mitbf{C}_{\eta_i}  + \kappa^2\mitbf{G}\mitbf{G}^T \right),
-\end{equation}
-where $C_{u_i}(\mitbf{P},\mitbf{P})$ represents the matrix $[C_{u_i}(p,p')]_{(p,p') \in \mitbf{P} \times \mitbf{P}}$.
-
-We form a posterior estimate of transient displacements, denoted as $\hat{u}_i$, by updating $u_i$ with the fact that $\mitbf{d}_i^*$ was realized from the random vector $\mitbf{d}_i$, that is to say $\hat{u}_i = u_i |(\mitbf{d}_i = \mitbf{d}^*_i)$.
-A general solution for $\hat{u}_i$ is derived in \citet[sec.
-8.9]{VonMises1964}, where we find that $\hat{u}_i$ is distributed as $\mathcal{GP}(\mu_{\hat{u}_i},C_{\hat{u}_i})$ with mean function
+\data_i \sim \mathcal{N}\left(\zeros, C_{u_i}(\points,\points) + 
+                                      \mitbf{C}_{\eta_i}  + 
+                                      \kappa^2\G\G^T \right),
+\end{equation}
+where $C_{u_i}(\points,\points)$ represents the matrix
+$[C_{u_i}(p,p')]_{(p,p') \in \points \times \points}$.
+
+We form a posterior estimate of transient displacements, denoted as
+$\post_i$, by updating $u_i$ with the fact that $\data_i^*$ was
+realized from the random vector $\data_i$, that is to say $\post_i =
+u_i |(\data_i = \data^*_i)$. A general solution for $\post_i$ is
+derived in \citet[sec. 8.9]{VonMises1964}, where we find that
+$\post_i$ is distributed as $\mathcal{GP}(\mu_{\post_i},C_{\post_i})$
+with mean function
 \begin{align}\label{eq:PosteriorMean}
-\mu_{\hat{u}_i}(p) &= \mathrm{E}\left[u_i(p)\right] + 
-                      \mathrm{Cov}\left[u_i(p),\mitbf{d}_i\right] 
-                      \mathrm{Cov}\left[\mitbf{d}_i\right]^{-1}
-                      \left(\mitbf{d}^*_i - \mathrm{E}\left[\mitbf{d}_i\right]\right)\nonumber \\
-                   &= C_{u_i}(p,\mitbf{P})\left( C_{u_i}(\mitbf{P},\mitbf{P}) + \mitbf{C}_{\eta_i} + \kappa^2\mitbf{G}\mitbf{G}^T\right)^{-1}\mitbf{d}^*_i
+\mu_{\post_i}(p) &= \E{u_i(p)} + 
+                    \Cov{u_i(p),\data_i} 
+                    \Cov{\data_i}^{-1}
+                    \left(\data^*_i - \E{\data_i}\right)\nonumber \\
+                 &= C_{u_i}(p,\points)\left( C_{u_i}(\points,\points) + 
+                                             \mitbf{C}_{\eta_i} + 
+                                             \kappa^2\G\G^T\right)^{-1}
+                    \data^*_i
 \end{align}    
@@
 \begin{align}\label{eq:PosteriorCov}
-C_{\hat{u}_i}(p,p') &= \mathrm{Cov}\left[u_i(p),u_i(p')\right] - 
-                       \mathrm{Cov}\left[u_i(p),\mitbf{d}_i\right] 
-                       \mathrm{Cov}\left[\mitbf{d}_i\right]^{-1}
-                       \mathrm{Cov}\left[\mitbf{d}_i,u_i(p')\right]\nonumber \\
-                    &= C_{u_i}(p,p') - C_{u_i}(p,\mitbf{P})\left( C_{u_i}(\mitbf{P},\mitbf{P}) + \mitbf{C}_{\eta_i} + \kappa^2\mitbf{G}\mitbf{G}^T \right)^{-1}C_{u_i}(\mitbf{P},p').
+C_{\post_i}(p,p') &= \Cov{u_i(p),u_i(p')} - 
+                     \Cov{u_i(p),\data_i} 
+                     \Cov{\data_i}^{-1}
+                     \Cov{\data_i,u_i(p')} \nonumber \\
+                  &= C_{u_i}(p,p') - 
+                     C_{u_i}(p,\points)\left( C_{u_i}(\points,\points) + 
+                                              \mitbf{C}_{\eta_i} + 
+                                              \kappa^2\G\G^T \right)^{-1}
+                     C_{u_i}(\points,p').
 \end{align}
-However, we are interested in the limit as $\kappa \to \infty$, and the form for eqs.
-(\ref{eq:PosteriorMean}) and (\ref{eq:PosteriorCov}) is not suitable for evaluating this limit.
-We use a partitioned matrix inversion identity \citep[sec.
-2.7.4]{Press2007} to rewrite eqs.
+However, we are interested in the limit as $\kappa \to \infty$, and
+the form for eqs. (\ref{eq:PosteriorMean}) and (\ref{eq:PosteriorCov})
+is not suitable for evaluating this limit. We use a partitioned matrix
+inversion identity \citep[sec. 2.7.4]{Press2007} to rewrite eqs.
 (\ref{eq:PosteriorMean}) and (\ref{eq:PosteriorCov}) as
 \begin{equation}\label{eq:PosteriorMean2}
-\mu_{\hat{u}_i}(p) = \left[\begin{array}{cc}
-                         C_{u_i}(p,\mitbf{P}) & \mitbf{0} \\
+\mu_{\post_i}(p) = \left[\begin{array}{cc} 
+                         C_{u_i}(p,\points) & \zeros \\
                          \end{array}\right]
                    \left[\begin{array}{cc}
-                         C_{u_i}(\mitbf{P},\mitbf{P}) + \mitbf{C}_{\eta_i} & \mitbf{G} \\
-                         \mitbf{G}^T  & -\kappa^{-2} \mitbf{I} \\
+                         C_{u_i}(\points,\points) + \mitbf{C}_{\eta_i} & \G \\
+                         \G^T  & -\kappa^{-2} \eye \\
                          \end{array}\right]^{-1}
                    \left[\begin{array}{c}
-                         \mitbf{d}^*_i \\
-                         \mitbf{0} \\
+                         \data^*_i \\
+                         \zeros \\
                          \end{array}\right]
@@
 \begin{multline}\label{eq:PosteriorCov2}
-C_{\hat{u}_i}(p,p') = C_{u_i}(p,p') - \\ 
+C_{\post_i}(p,p') = C_{u_i}(p,p') - \\ 
                     \left[\begin{array}{cc}
-                          C_{u_i}(p,\mitbf{P}) & \mitbf{0} \\
+                          C_{u_i}(p,\points) & \zeros \\
                           \end{array}\right]
                     \left[\begin{array}{cc}
-                          C_{u_i}(\mitbf{P},\mitbf{P}) + \mitbf{C}_{\eta_i} & \mitbf{G} \\
-                          \mitbf{G}^T  & -\kappa^{-2} \mitbf{I} \\
+                          C_{u_i}(\points,\points) + \mitbf{C}_{\eta_i} & \G \\
+                          \G^T  & -\kappa^{-2} \eye \\
                           \end{array}\right]^{-1}
                     \left[\begin{array}{c}
-                          C_{u_i}(\mitbf{P},p') \\
-                          \mitbf{0} \\
+                          C_{u_i}(\points,p') \\
+                          \zeros \\
                           \end{array}\right].
 \end{multline}
-Taking the limit as $\kappa \to \infty$, we get the solution for the mean and covariance of $\hat{u}_i$,
+Taking the limit as $\kappa \to \infty$, we get the solution for the
+mean and covariance of $\post_i$,
 \begin{equation}\label{eq:PosteriorMean3}
-\mu_{\hat{u}_i}(p) = \left[\begin{array}{cc}
-                         C_{u_i}(p,\mitbf{P}) & \mitbf{0} \\
+\mu_{\post_i}(p) = \left[\begin{array}{cc}
+                         C_{u_i}(p,\points) & \zeros \\
                          \end{array}\right]
                    \left[\begin{array}{cc}
-                         C_{u_i}(\mitbf{P},\mitbf{P}) + \mitbf{C}_{\eta_i} & \mitbf{G} \\
-                         \mitbf{G}^T  & \mitbf{0} \\
+                         C_{u_i}(\points,\points) + \mitbf{C}_{\eta_i} & \G \\
+                         \G^T  & \zeros \\
                          \end{array}\right]^{-1}
                    \left[\begin{array}{c}
-                         \mitbf{d}^*_i \\
-                         \mitbf{0} \\
+                         \data^*_i \\
+                         \zeros \\
                          \end{array}\right]
@@
 \begin{equation}\label{eq:PosteriorCov3}
-C_{\hat{u}_i}(p,p') = C_{u_i}(p,p') - 
+C_{\post_i}(p,p') = C_{u_i}(p,p') - 
                     \left[\begin{array}{cc}
-                          C_{u_i}(p,\mitbf{P}) & \mitbf{0} \\
+                          C_{u_i}(p,\points) & \zeros \\
                           \end{array}\right]
                     \left[\begin{array}{cc}
-                          C_{u_i}(\mitbf{P},\mitbf{P}) + \mitbf{C}_{\eta_i} & \mitbf{G} \\
-                          \mitbf{G}^T  & \mitbf{0} \\
+                          C_{u_i}(\points,\points) + \mitbf{C}_{\eta_i} & \G \\
+                          \G^T  & \zeros \\
                           \end{array}\right]^{-1}
                     \left[\begin{array}{c}
-                          C_{u_i}(\mitbf{P},p') \\
-                          \mitbf{0} \\
+                          C_{u_i}(\points,p') \\
+                          \zeros \\
                           \end{array}\right].
 \end{equation}
-To ensure that the inverse matrices in eqs.
-(\ref{eq:PosteriorMean3}) and (\ref{eq:PosteriorCov3}) exist, each column in $\mitbf{G}$ must be linearly independent.
-This condition tends to be violated when there are too few observations at a station.
-In that case, a singular value decomposition can be used to remove linearly dependent components from $\mitbf{G}$.
-
-It should be noted that we have ignored any covariances between the easting and northing components of $\vec{u}$ and $\vec{\mitbf{d}}$.
-This simplification reduces the computational complexity of evaluating the posterior transient displacements because each component can be evaluated independently.
-However, we are inherently assuming that the principle axes describing the distribution of $\vec{u}(p)$ and $\vec{d}^{(jk)}$ are aligned with the cardinal directions, which is admittedly an arbitrary assumption.
-
-The posterior transient displacements are spatially and temporally continuous, and we can use eqs.
-(\ref{eq:PosteriorMean3}) and (\ref{eq:PosteriorCov3}) to evaluate $\hat{u}_i$ at any $p$.
-Furthermore, $\hat{u}_i$ is spatially and temporally differentiable, allowing us to formulate $\dot\varepsilon$ at any $p$ that we may be interested in.
-The components of $\dot\varepsilon$ can be written as
+To ensure that the inverse matrices in eqs. (\ref{eq:PosteriorMean3})
+and (\ref{eq:PosteriorCov3}) exist, each column in $\G$ must be
+linearly independent. This condition tends to be violated when there
+are too few observations at a station. In that case, a singular value
+decomposition can be used to remove linearly dependent components from
+$\G$.
+
+It should be noted that we have ignored any covariances between the
+easting and northing components of $\vec{u}$ and $\vec{\data}$.
+This simplification reduces the computational complexity of evaluating
+the posterior transient displacements because each component can be
+evaluated independently. However, we are inherently assuming that the
+principle axes describing the distribution of $\vec{u}(p)$ and
+$\vec{d}^{(jk)}$ are aligned with the cardinal directions, which is
+admittedly an arbitrary assumption.
+
+The posterior transient displacements are spatially and temporally
+continuous, and we can use eqs. (\ref{eq:PosteriorMean3}) and
+(\ref{eq:PosteriorCov3}) to evaluate $\post_i$ at any $p$.
+Furthermore, $\post_i$ is spatially and temporally differentiable,
+allowing us to formulate $\strain$ at any $p$ that we may be
+interested in. The components of $\strain$ can be written as
 \begin{equation}\label{eq:StrainRate}
-\dot\varepsilon_{ij}(p) = \frac{1}{2} \frac{\partial}{\partial t} \left(
-                                        \frac{\partial \hat{u}_i(p)}{\partial x_j} +  
-                                        \frac{\partial \hat{u}_j(p)}{\partial x_i} \right).
-\end{equation}
-Since eq.
-(\ref{eq:StrainRate}) is a linear operation on the Gaussian processes $\hat{u}_i$ and $\hat{u}_j$, we know that $\dot\varepsilon_{ij}$ is also a Gaussian process.
-From \citet[sec.
-10]{Papoulis1991}, we find that the mean and covariance functions for $\dot{\varepsilon}_{ij}$ are
+\strain_{ij}(p) = \frac{1}{2} \frac{\partial}{\partial t} 
+                  \left(\frac{\partial \post_i(p)}{\partial x_j} +  
+                        \frac{\partial \post_j(p)}{\partial x_i} \right).
+\end{equation}
+Since eq. (\ref{eq:StrainRate}) is a linear operation on the Gaussian
+processes $\post_i$ and $\post_j$, we know that $\strain_{ij}$ is also
+a Gaussian process. From \citet[sec. 10]{Papoulis1991}, we find that
+the mean and covariance functions for $\dot{\varepsilon}_{ij}$ are
 \begin{align}\label{eq:StrainMean}
-\mu_{\dot\varepsilon_\mathrm{ee}}(p) &= \frac{\partial^2 \mu_{\hat{u}_\mathrm{e}}(p)}{\partial t \, \partial x_\mathrm{e}} \\
-\mu_{\dot\varepsilon_\mathrm{nn}}(p) &= \frac{\partial^2 \mu_{\hat{u}_\mathrm{n}}(p)}{\partial t \, \partial x_\mathrm{n}} \\
-\mu_{\dot\varepsilon_\mathrm{en}}(p) &= \frac{1}{2}\frac{\partial}{\partial t}\left(
-                                        \frac{\partial \mu_{\hat{u}_\mathrm{e}}(p)}{\partial x_\mathrm{n}} + 
-                                        \frac{\partial \mu_{\hat{u}_\mathrm{n}}(p)}{\partial x_\mathrm{e}} \right)                                  
+\mu_{\strain_\e\e}(p) &= \frac{\partial^2 \mu_{\post_\e}(p)}{\partial t \, \partial x_\e} \\
+\mu_{\strain_\n\n}(p) &= \frac{\partial^2 \mu_{\post_\n}(p)}{\partial t \, \partial x_\n} \\
+\mu_{\strain_\e\n}(p) &= \frac{1}{2}\frac{\partial}{\partial t}\left(
+                                    \frac{\partial \mu_{\post_\e}(p)}{\partial x_\n} + 
+                                    \frac{\partial \mu_{\post_\n}(p)}{\partial x_\e} \right)
 \end{align} 
@@
 \begin{align}\label{eq:StrainCov}
-C_{\dot\varepsilon_\mathrm{ee}}(p,p') &= \frac{\partial^4 C_{\hat{u}_\mathrm{e}}(p,p')}{\partial t \, \partial t' \, \partial x_\mathrm{e} \, \partial x'_\mathrm{e}} \\
-C_{\dot\varepsilon_\mathrm{nn}}(p,p') &= \frac{\partial^4 C_{\hat{u}_\mathrm{n}}(p,p')}{\partial t \, \partial t' \, \partial x_\mathrm{n} \, \partial x'_\mathrm{n}} \\
-C_{\dot\varepsilon_\mathrm{en}}(p,p') &= \frac{1}{4} \frac{\partial^2}{\partial t \, \partial t'}\left(
-                                         \frac{\partial^2 C_{\hat{u}_\mathrm{e}}(p,p')}{\partial x_\mathrm{n} \, \partial x'_\mathrm{n}} + 
-                                         \frac{\partial^2 C_{\hat{u}_\mathrm{n}}(p,p')}{\partial x_\mathrm{e} \, \partial x'_\mathrm{e}} \right),
+C_{\strain_\e\e}(p,p') &= \frac{\partial^4 C_{\post_\e}(p,p')}{\partial t \, \partial t' \, \partial x_\e \, \partial x'_\e} \\
+C_{\strain_\n\n}(p,p') &= \frac{\partial^4 C_{\post_\n}(p,p')}{\partial t \, \partial t' \, \partial x_\n \, \partial x'_\n} \\
+C_{\strain_\e\n}(p,p') &= \frac{1}{4} \frac{\partial^2}{\partial t \, \partial t'}\left(
+                                      \frac{\partial^2 C_{\post_\e}(p,p')}{\partial x_\n \, \partial x'_\n} + 
+                                      \frac{\partial^2 C_{\post_\n}(p,p')}{\partial x_\e \, \partial x'_\e} \right),
 \end{align} 
 respectively.
+
+%% TRANSIENT DETECTION CRITERION
 \subsection{Transient detection criterion}\label{sec:TransientDetection}
-Our motivation for estimating transient strain rates is, in part, to detect geophysical phenomena.
-As we will see, geophysical signal can be easily identified by visually inspecting the solution for $\dot\varepsilon$ from eqs.
-(\ref{eq:StrainMean}) and (\ref{eq:StrainCov}).
-However, if we want to detect geophysical phenomena automatically, then we need to define a detection criterion.
-We use a signal-to-noise ratio, SNR, that is based on the Frobenius norm of $\dot\varepsilon$, $||\dot\varepsilon||_\mathrm{F} = \left(\dot\varepsilon_{\mathrm{ee}}^2 + \dot\varepsilon_{\mathrm{nn}}^2 + 2\dot\varepsilon_{\mathrm{en}}^2\right)^{\frac{1}{2}}$, for our detection criterion.
-In the geodetic literature, $||\dot{\varepsilon}||_\mathrm{F}$ is often used as a metric for the strain rate ``magnitude", and it is sometimes referred to as the second invariant of strain rate.
-Noting that $||\dot\varepsilon||_\mathrm{F}$ is a random variable, we take SNR to be the ratio of the estimated mean and standard deviation of $||\dot\varepsilon||_\mathrm{F}$.
-An estimate of the mean is found by evaluating $||\dot\varepsilon||_\mathrm{F}$ at the mean of $\dot\varepsilon$,
+
+Our motivation for estimating transient strain rates is, in part, to
+detect geophysical phenomena. As we will see, geophysical signal can
+be easily identified by visually inspecting the solution for $\strain$
+from eqs. (\ref{eq:StrainMean}) and (\ref{eq:StrainCov}). However, if
+we want to detect geophysical phenomena automatically, then we need to
+define a detection criterion. We use a signal-to-noise ratio, SNR,
+that is based on the Frobenius norm of $\strain$,
+$||\strain||_\mathrm{F} = \left(\strain_{\e\e}^2 + \strain_{\n\n}^2 +
+2\strain_{\e\n}^2\right)^{\frac{1}{2}}$, for our detection criterion.
+In the geodetic literature, $||\dot{\varepsilon}||_\mathrm{F}$ is
+often used as a metric for the strain rate ``magnitude", and it is
+sometimes referred to as the second invariant of strain rate. Noting
+that $||\strain||_\mathrm{F}$ is a random variable, we take SNR to be
+the ratio of the estimated mean and standard deviation of
+$||\strain||_\mathrm{F}$. An estimate of the mean is found by
+evaluating $||\strain||_\mathrm{F}$ at the mean of $\strain$,
 \begin{align}
-\mu_{||\dot\varepsilon||_\mathrm{F}} &\approx ||\dot\varepsilon||_F \big|_{\dot\varepsilon = \mu_{\dot\varepsilon}}\nonumber \\
-                                     &= \left(\mu_{\dot\varepsilon_\mathrm{ee}}^2 + 
-                                              \mu_{\dot\varepsilon_\mathrm{nn}}^2 + 
-                                              2\mu_{\dot\varepsilon_\mathrm{en}}^2
-                                     \right)^\frac{1}{2},
+\mu_{||\strain||_\mathrm{F}} &\approx ||\strain||_F \big|_{\strain = \mu_{\strain}}\nonumber \\
+                             &= \left(\mu_{\strain_\e\e}^2 + 
+                                      \mu_{\strain_\n\n}^2 + 
+                                      2\mu_{\strain_\e\n}^2 \right)^\frac{1}{2},
 \end{align}
-and we use nonlinear uncertainty propagation to estimate the standard deviation,
+and we use nonlinear uncertainty propagation to estimate the standard
+deviation,
 \begin{equation}
-\sigma_{||\dot\varepsilon||_\mathrm{F}} \approx
-\left(\left(\left.
-\frac{\partial ||\dot\varepsilon||_\mathrm{F}}{\partial \dot\varepsilon_\mathrm{ee}}\right|_{\dot\varepsilon = \mu_{\dot\varepsilon}} \right)^2 \sigma^2_{\dot\varepsilon_\mathrm{ee}} +
-\left(\left. \frac{\partial ||\dot\varepsilon||_\mathrm{F}}{\partial \dot\varepsilon_\mathrm{nn}}\right|_{\dot\varepsilon = \mu_{\dot\varepsilon}} \right)^2 \sigma^2_{\dot\varepsilon_\mathrm{nn}} +
-\left(\left. \frac{\partial ||\dot\varepsilon||_\mathrm{F}}{\partial \dot\varepsilon_\mathrm{en}}\right|_{\dot\varepsilon = \mu_{\dot\varepsilon}} \right)^2 \sigma^2_{\dot\varepsilon_\mathrm{en}}
-\right)^\frac{1}{2},
-\end{equation}
-where $\sigma^2_{\dot\varepsilon_{ij}}(p) = C_{\dot\varepsilon_{ij}}(p,p)$. After some calculations, we find SNR to be
+\sigma_{||\strain||_\mathrm{F}} \approx
+\left( \left(\left. \frac{\partial ||\strain||_\mathrm{F}}{\partial \strain_\e\e}\right|_{\strain = \mu_{\strain}} \right)^2 
+       \sigma^2_{\strain_\e\e} +
+       \left(\left. \frac{\partial ||\strain||_\mathrm{F}}{\partial \strain_\n\n}\right|_{\strain = \mu_{\strain}} \right)^2
+       \sigma^2_{\strain_\n\n} +
+       \left(\left. \frac{\partial ||\strain||_\mathrm{F}}{\partial \strain_\e\n}\right|_{\strain = \mu_{\strain}} \right)^2 
+       \sigma^2_{\strain_\e\n} \right)^\frac{1}{2},
+\end{equation}
+where $\sigma^2_{\strain_{ij}}(p) = C_{\strain_{ij}}(p,p)$. After some
+calculations, we find SNR to be
 \begin{align}\label{eq:SNR}
-\mathrm{SNR}(p) &= \frac{\mu_{||\dot\varepsilon||_\mathrm{F}}(p)}{\sigma_{||\dot\varepsilon||_\mathrm{F}}(p)} \\
-                &= \frac{\mu_{\dot\varepsilon_\mathrm{ee}}(p)^2 +
-                        \mu_{\dot\varepsilon_\mathrm{nn}}(p)^2 +
-                        2\mu_{\dot\varepsilon_\mathrm{en}}(p)^2}
-                       {\big(\sigma^2_{\dot\varepsilon_\mathrm{ee}}(p)\mu_{\dot\varepsilon_\mathrm{ee}}(p)^2 + 
-                             \sigma^2_{\dot\varepsilon_\mathrm{nn}}(p)\mu_{\dot\varepsilon_\mathrm{nn}}(p)^2 + 
-                              4\sigma^2_{\dot\varepsilon_\mathrm{en}}(p)\mu_{\dot\varepsilon_\mathrm{en}}(p)^2
-                        \big)^{\frac{1}{2}}}
-.
+\mathrm{SNR}(p) &= \frac{\mu_{||\strain||_\mathrm{F}}(p)}{\sigma_{||\strain||_\mathrm{F}}(p)} \\
+                &= \frac{\mu_{\strain_\e\e}(p)^2 +
+                         \mu_{\strain_\n\n}(p)^2 +
+                         2\mu_{\strain_\e\n}(p)^2}
+                        {\big(\sigma^2_{\strain_\e\e}(p)\mu_{\strain_\e\e}(p)^2 + 
+                              \sigma^2_{\strain_\n\n}(p)\mu_{\strain_\n\n}(p)^2 + 
+                              4\sigma^2_{\strain_\e\n}(p)\mu_{\strain_\e\n}(p)^2
+                         \big)^{\frac{1}{2}}}.
 \end{align}
-We explicitly show that SNR is a function of $p$ to emphasize that it identifies the position and time of anomalous deformation.
-We can reasonably suspect that some transient geophysical phenomena is occurring wherever and whenever SNR is larger than ${\sim}3$.
-
+We explicitly show that SNR is a function of $p$ to emphasize that it
+identifies the position and time of anomalous deformation. We can
+reasonably suspect that some transient geophysical phenomena is
+occurring wherever and whenever SNR is larger than ${\sim}3$.
+
+%% OUTLIER DETECTION
 \subsection{Outlier detection}\label{sec:Outlier}
-In deriving our formulation for transient strain rates, we have assumed that noise in the data vector is normally distributed.
-This is not an appropriate assumption for GNSS data, which often have more outliers than would be expected for normally distributed noise.
-Methods for analyzing GNSS data should either be robust against outliers or should involve a preprocessing step in which outliers are detected and removed.
-Examples of the former include the MIDAS method for estimating secular velocities \citep{Blewitt2016} and the GPS Imaging method for detecting spatially coherent features \citep{Hammond2016}.
-In this study, we identify and remove outliers as a preprocessing step before estimating $\dot\varepsilon$.
-Outliers are identified based on the residuals for a model that best fits the observed data.
-Observations with residuals that exceed some threshold are removed.
-This strategy for detecting outliers is commonly used for GNSS data, where the model being fit to the data typically consists of a linear trend and seasonal terms for each GNSS station \citep[e.g.,][]{Johansson2002,Dong2006,Bos2013}.
-To prevent transient geophysical signal from being erroneously identified as outliers, the model used in our outlier detection algorithm additionally consists of a temporally correlated Gaussian process.
-The details of our algorithm are given in Appendix A.
-
-It should be noted that our algorithm does not identify jumps in GNSS time series, which are another common issue.
-Some, but not all, jumps can be automatically removed by looking up the dates of equipment changes and earthquakes \citep{Gazeaux2013}.
-However, it is still necessary to manually find and remove jumps of unknown origin.
-
+
+In deriving our formulation for transient strain rates, we have
+assumed that noise in the data vector is normally distributed. This is
+not an appropriate assumption for GNSS data, which often have more
+outliers than would be expected for normally distributed noise.
+Methods for analyzing GNSS data should either be robust against
+outliers or should involve a preprocessing step in which outliers are
+detected and removed. Examples of the former include the MIDAS method
+for estimating secular velocities \citep{Blewitt2016} and the GPS
+Imaging method for detecting spatially coherent features
+\citep{Hammond2016}. In this study, we identify and remove outliers as
+a preprocessing step before estimating $\strain$. Outliers are
+identified based on the residuals for a model that best fits the
+observed data. Observations with residuals that exceed some threshold
+are removed. This strategy for detecting outliers is commonly used for
+GNSS data, where the model being fit to the data typically consists of
+a linear trend and seasonal terms for each GNSS station
+\citep[e.g.,][]{Johansson2002,Dong2006,Bos2013}. To prevent transient
+geophysical signal from being erroneously identified as outliers, the
+model used in our outlier detection algorithm additionally consists of
+a temporally correlated Gaussian process. The details of our algorithm
+are given in Appendix A.
+
+It should be noted that our algorithm does not identify jumps in GNSS
+time series, which are another common issue. Some, but not all, jumps
+can be automatically removed by looking up the dates of equipment
+changes and earthquakes \citep{Gazeaux2013}. However, it is still
+necessary to manually find and remove jumps of unknown origin.
+
+%% Application to Pacific Northwest Slow Slip Events
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Application to Pacific Northwest Slow Slip Events}\label{sec:Cascadia}
-We use our method to estimate transient strain rates in the Pacific Northwest, and we are specifically interested in identifying strain resulting from SSEs \citep[e.g.,][]{Dragert2001}.
-In the Pacific Northwest, SSEs can be detected by monitoring for associated seismic tremor \citep{Rogers2003}, which is actively being done by the Pacific Northwest Seismic Network \citep{Wech2010}.
-We can thus compare the tremor records to the transient strain rates estimated with GPR to verify that we are indeed identifying strain from SSEs.
-
-We use the daily displacement solutions generated by the Geodesy Advancing Geosciences and EarthScope (GAGE) facility for continuous GNSS stations \citep{Herring2016}.
-This data is publicly available and can be found at www.unavco.org.
-We limit the dataset to the stations and time ranges which are pertinent to the seven most recent SSEs in the Puget Sound region.
-The earliest SSE considered in this study began in August 2010, and the most recent SSE began in February 2017.
-We use these most recent SSEs because the station coverage is sufficiently dense for us to use maximum likelihood methods to constrain prior models.
-The positions of GNSS stations used to estimate transient strain rates are shown in Figure \ref{fig:Context}.
+
+We use our method to estimate transient strain rates in the Pacific
+Northwest, and we are specifically interested in identifying strain
+resulting from SSEs \citep[e.g.,][]{Dragert2001}. In the Pacific
+Northwest, SSEs can be detected by monitoring for associated seismic
+tremor \citep{Rogers2003}, which is actively being done by the Pacific
+Northwest Seismic Network \citep{Wech2010}. We can thus compare the
+tremor records to the transient strain rates estimated with GPR to
+verify that we are indeed identifying strain from SSEs.
+
+We use the daily displacement solutions generated by the Geodesy
+Advancing Geosciences and EarthScope (GAGE) facility for continuous
+GNSS stations \citep{Herring2016}. This data is publicly available and
+can be found at www.unavco.org. We limit the dataset to the stations
+and time ranges which are pertinent to the seven most recent SSEs in
+the Puget Sound region. The earliest SSE considered in this study
+began in August 2010, and the most recent SSE began in February 2017.
+We use these most recent SSEs because the station coverage is
+sufficiently dense for us to use maximum likelihood methods to
+constrain prior models.  The positions of GNSS stations used to
+estimate transient strain rates are shown in Figure \ref{fig:Context}.
 
@@
 \includegraphics{figures/context_map/context-map.pdf}
-\caption{Positions of continuous GNSS stations used to estimate transient strain rates.
-The colored regions indicate the distribution of seismic tremor as determined by \citet{Wech2010}.
-The red dots show the positions of GNSS stations mentioned in this paper.
-The blue dot indicates the location of the transient strain rates shown in Figure \ref{fig:StrainTs} and the signal-to-noise ratio shown in Figure \ref{fig:StrainMag}.
-The blue dashed circle demarcates the spatial extent of the tremors shown in Figure \ref{fig:StrainMag}.}
+\caption{
+Positions of continuous GNSS stations used to estimate transient
+strain rates. The colored regions indicate the distribution of seismic
+tremor as determined by \citet{Wech2010}. The red dots show the
+positions of GNSS stations mentioned in this paper. The blue dot
+indicates the location of the transient strain rates shown in Figure
+\ref{fig:StrainTs} and the signal-to-noise ratio shown in Figure
+\ref{fig:StrainMag}. The blue dashed circle demarcates the spatial
+extent of the tremors shown in Figure \ref{fig:StrainMag}.
+}    
 \label{fig:Context}
@@
 
+%% NOISE MODEL
 \subsection{Noise model}\label{sec:NoiseModel}
-Before we determine the transient strain rates, we must establish a prior for the transient displacements, $u$, and the noise, $\eta$.
-In this section we discuss our choice for the noise covariance function $C_\eta$.
-There have been numerous studies on temporally correlated noise in GNSS data \citep[e.g.,][]{Zhang1997,Mao1999,Williams2004,Langbein2008}.
-In these studies, temporally correlated noise was described with some combination of Brownian motion, a first-order Gauss-Markov (FOGM) process, and/or flicker noise.
-There is some physical justification for using Brownian motion as a noise model because it accurately describes the power spectrum of motion resulting from instability in geodetic monuments \citep[e.g.,][]{Wyatt1982,Wyatt1989}.
-Here we describe the time dependence of $\eta$ as a FOGM process and consider $\eta$ to be spatially uncorrelated.
-A FOGM process is a solution to the stochastic differential equation
+
+Before we determine the transient strain rates, we must establish a
+prior for the transient displacements, $u$, and the noise, $\eta$. In
+this section we discuss our choice for the noise covariance function
+$C_\eta$. There have been numerous studies on temporally correlated
+noise in GNSS data \citep[e.g.,][]{Zhang1997, Mao1999, Williams2004,
+Langbein2008}. In these studies, temporally correlated noise was
+described with some combination of Brownian motion, a first-order
+Gauss-Markov (FOGM) process, and/or flicker noise. There is some
+physical justification for using Brownian motion as a noise model
+because it accurately describes the power spectrum of motion resulting
+from instability in geodetic monuments \citep[e.g.,][]{Wyatt1982,
+Wyatt1989}. Here we describe the time dependence of $\eta$ as a FOGM
+process and consider $\eta$ to be spatially uncorrelated. A FOGM
+process is a solution to the stochastic differential equation
 \begin{equation}\label{eq:FOGMdef}
@@
 \end{equation}
-where $w(t)$ is white noise with unit variance.
-The FOGM process degenerates to the commonly used Brownian motion noise model under the condition that $\alpha=0$ and $\eta(0) = 0$.
-Our spatially uncorrelated noise model that satisfies eq.
-(\ref{eq:FOGMdef}) is a Gaussian process with zero mean and the covariance function
+where $w(t)$ is white noise with unit variance. The FOGM process
+degenerates to the commonly used Brownian motion noise model under the
+condition that $\alpha=0$ and $\eta(0) = 0$. Our spatially
+uncorrelated noise model that satisfies eq. (\ref{eq:FOGMdef}) is a
+Gaussian process with zero mean and the covariance function
 \begin{equation}\label{eq:FOGM}
-C_\eta\left((\vec{x},t),(\vec{x}',t')\right) = 
+C_\eta\left((\pos,t),(\pos',t')\right) = 
 \begin{cases}
-\frac{\beta^2}{2\alpha}\exp\left(-\alpha|t - t'|\right), &\vec{x} = \vec{x}' \\
-0, &\vec{x} \neq \vec{x}'
+\frac{\beta^2}{2\alpha}\exp\left(-\alpha|t - t'|\right), &\pos = \pos' \\
+0, &\pos \neq \pos'
 \end{cases}. 
@@
 
-We constrain the hyperparameters for $\eta$, $\alpha$ and $\beta$, with a set of 38 continuous GNSS stations in the Pacific Northwest that are east of 121$^\circ$W.
-These stations are sufficiently far from the subduction zone that they are unlikely to contain transient signal associated with SSEs.
-We clean the data for these stations by removing jumps at times of equipment changes, and we remove outliers that have been detected with the algorithm described in Section \ref{sec:Outlier}.
-We then find $\alpha$ and $\beta$ for each station time series with the Restricted Maximum Likelihood (REML) method \cite[e.g.,][]{Harville1974,Cressie1992,Hines2017}.
-The REML method finds the hyperparameters, which we collectively refer to as $\mitbf{\theta}$, that maximize the likelihood function
+We constrain the hyperparameters for $\eta$, $\alpha$ and $\beta$,
+with a set of 38 continuous GNSS stations in the Pacific Northwest
+that are east of 121$^\circ$W.  These stations are sufficiently far
+from the subduction zone that they are unlikely to contain transient
+signal associated with SSEs.  We clean the data for these stations by
+removing jumps at times of equipment changes, and we remove outliers
+that have been detected with the algorithm described in Section
+\ref{sec:Outlier}. We then find $\alpha$ and $\beta$ for each station
+time series with the Restricted Maximum Likelihood (REML) method
+\cite[e.g.,][]{Harville1974,Cressie1992,Hines2017}. The REML method
+finds the hyperparameters, which we collectively refer to as
+$\mitbf{\theta}$, that maximize the likelihood function
 \begin{equation}\label{eq:REML}
-\mathcal{L}(\mitbf{\theta}) = \left(\frac{\left|\mitbf{G}^T\mitbf{G}\right|}
-                           {(2\pi)^{n-6m} 
+\mathcal{L}(\mitbf{\theta}) = \left(\frac{\left|\G^T\G\right|}{(2\pi)^{n-6m} 
                                     \left| \mitbf{\Sigma}(\mitbf{\theta}) \right| 
-                           \left| \mitbf{G}^T\mitbf{\Sigma}(\mitbf{\theta})^{-1}\mitbf{G} \right|}\right)^{\frac{1}{2}} 
-                           e^{-\tfrac{1}{2}\mitbf{d}_*^T\mitbf{K}(\mitbf{\theta})\mitbf{d}_*},
+                                    \left| \G^T\mitbf{\Sigma}(\mitbf{\theta})^{-1}\G \right|}\right)^{\frac{1}{2}} 
+                              e^{-\tfrac{1}{2}\data_*^T\mitbf{K}(\mitbf{\theta})\data_*},
 \end{equation}
@@
 \mitbf{K}(\mitbf{\theta}) = \mitbf{\Sigma}(\mitbf{\theta})^{-1} - 
-                      \mitbf{\Sigma}(\mitbf{\theta})^{-1}\mitbf{G}
-         \left(\mitbf{G}^T\mitbf{\Sigma}(\mitbf{\theta})^{-1}\mitbf{G}\right)^{-1}
-         \mitbf{G}^T\mitbf{\Sigma}(\mitbf{\theta})^{-1}.
-\end{equation}
-\citet{Harville1974} showed that choosing the hyperparameters which maximize eq.
-(\ref{eq:REML}) is equivalent to choosing the hyperparameters which maximize the probability of drawing $\mitbf{d}_*$ from $\mitbf{d}$.
-We use the REML method over the maximum likelihood method \citep[e.g.,][]{Langbein1997} because the REML method accounts for the improper prior that we assigned to $\mitbf{a}$ \citep{Hines2017}.
-We independently estimate $\mitbf{\theta}$ for each station, and so $\mitbf{d}_*$ consists of displacements for an individual station.
-We are assuming $u(p)=0$ when estimating the noise hyperparameters for this section.
-The distribution of inferred $\alpha$ and $\beta$ are shown in Figure \ref{fig:NoiseParams}.
-The amplitude of FOGM noise, $\beta$, for the easting and northing components is notably low and are clustered around 0.5 mm/yr$^{0.5}$.
-The corresponding estimates of $\alpha$ tend to cluster around 0 yr$^{-1}$, suggesting that noise can be described with Brownian motion.
-We also estimate hyperparameters for the vertical component of displacements, under the hope that vertical deformation gradients could reveal some geophysical signal.
-The amplitude of FOGM noise for the vertical component is relatively large with a median value of 13.5 mm/yr$^{0.5}$.
-The inferred values for $\alpha$ are also higher for the vertical component with a median value of 8.21 yr$^{-1}$.
-In Figure \ref{fig:NoiseSamples}, we use the median values of $\alpha$ and $\beta$ to generate two random samples of FOGM noise for each component.
-The samples span five years and over these five years the easting and northing samples drift by about 1 mm.
-In the context of detecting SSEs, which produce several mm's of surface displacement on the time-scale of weeks, the estimated FOGM noise for the easting and northing component is negligible.
-In contrast, the estimated FOGM noise for the vertical component is larger than the signal we would expect from SSEs.
-We suspect that the higher amplitude for the FOGM noise in the vertical component is accommodating for deficiencies in our rather simple seasonal model.
-Based on this analysis, we henceforth ignore temporally correlated noise in the easting and northing component because of its low amplitude.
-We also do not use vertical displacements because of the presumably low signal-to-noise ratio.
+                            \mitbf{\Sigma}(\mitbf{\theta})^{-1}\G
+                            \left(\G^T\mitbf{\Sigma}(\mitbf{\theta})^{-1}\G\right)^{-1}
+                            \G^T\mitbf{\Sigma}(\mitbf{\theta})^{-1}.
+\end{equation}
+\citet{Harville1974} showed that choosing the hyperparameters which
+maximize eq. (\ref{eq:REML}) is equivalent to choosing the
+hyperparameters which maximize the probability of drawing $\data_*$
+from $\data$. We use the REML method over the maximum likelihood
+method \citep[e.g.,][]{Langbein1997} because the REML method accounts
+for the improper prior that we assigned to $\mitbf{a}$
+\citep{Hines2017}. We independently estimate $\mitbf{\theta}$ for each
+station, and so $\data_*$ consists of displacements for an individual
+station. We are assuming $u(p)=0$ when estimating the noise
+hyperparameters for this section. The distribution of inferred
+$\alpha$ and $\beta$ are shown in Figure \ref{fig:NoiseParams}. The
+amplitude of FOGM noise, $\beta$, for the easting and northing
+components is notably low and are clustered around 0.5 mm/yr$^{0.5}$.
+The corresponding estimates of $\alpha$ tend to cluster around 0
+yr$^{-1}$, suggesting that noise can be described with Brownian
+motion. We also estimate hyperparameters for the vertical component of
+displacements, under the hope that vertical deformation gradients
+could reveal some geophysical signal. The amplitude of FOGM noise for
+the vertical component is relatively large with a median value of 13.5
+mm/yr$^{0.5}$.  The inferred values for $\alpha$ are also higher for
+the vertical component with a median value of 8.21 yr$^{-1}$. In
+Figure \ref{fig:NoiseSamples}, we use the median values of $\alpha$
+and $\beta$ to generate two random samples of FOGM noise for each
+component. The samples span five years and over these five years the
+easting and northing samples drift by about 1 mm. In the context of
+detecting SSEs, which produce several mm's of surface displacement on
+the time-scale of weeks, the estimated FOGM noise for the easting and
+northing component is negligible. In contrast, the estimated FOGM
+noise for the vertical component is larger than the signal we would
+expect from SSEs. We suspect that the higher amplitude for the FOGM
+noise in the vertical component is accommodating for deficiencies in
+our rather simple seasonal model. Based on this analysis, we
+henceforth ignore temporally correlated noise in the easting and
+northing component because of its low amplitude. We also do not use
+vertical displacements because of the presumably low signal-to-noise
+ratio.
 
@@
 \includegraphics{figures/noise/noise-params.pdf}
-\caption{Distribution of estimated FOGM hyperparameters (eq.
-\ref{eq:FOGM}).
-Hyperparameters are estimated with the REML method for 38 stations in the Pacific Northwest that are east of $121^\circ$W.
-``IQR'' is the interquartile range.}
+\caption{
+Distribution of estimated FOGM hyperparameters (eq. \ref{eq:FOGM}).
+Hyperparameters are estimated with the REML method for 38 stations in
+the Pacific Northwest that are east of $121^\circ$W. ``IQR'' is the
+interquartile range.
+}   
 \label{fig:NoiseParams}
@@
 \includegraphics{figures/noise/noise-samples.pdf}
-\caption{Two FOGM noise samples for each component. The FOGM hyperparameters have been set to the median values from Figure \ref{fig:NoiseParams}.}   
+\caption{
+Two FOGM noise samples for each component. The FOGM hyperparameters
+have been set to the median values from Figure \ref{fig:NoiseParams}.
+}   
 \label{fig:NoiseSamples}
@@
 
-Another significant source of noise in GNSS data is common mode error \citep[e.g.,][]{Wdowinski1997,Dong2006}, which is noise that is highly spatially correlated.
-When not accounted for, common mode error manifests as spatially uniform undulations in estimated transient displacements.
-However, estimated transient strain rates are insensitive to common mode error.
-We therefore do not include common mode error in our noise model.
-We then make the simplifying assumption that $\eta(p) = 0$ for the easting and northing component of GNSS data.
-
+Another significant source of noise in GNSS data is common mode error
+\citep[e.g.,][]{Wdowinski1997,Dong2006}, which is noise that is highly
+spatially correlated. When not accounted for, common mode error
+manifests as spatially uniform undulations in estimated transient
+displacements. However, estimated transient strain rates are
+insensitive to common mode error. We therefore do not include common
+mode error in our noise model. We then make the simplifying assumption
+that $\eta(p) = 0$ for the easting and northing component of GNSS
+data.
+
+%% PRIOR MODEL
 \subsection{Prior model}\label{sec:SignalModel}
+
 We next establish our prior model for transient displacements.
-Specifically, we discuss our choice for the covariance functions $X(\vec{x},\vec{x}')$ and $T(t,t')$.
-For $X$, we use the squared exponential (SE) covariance function,
+Specifically, we discuss our choice for the covariance functions
+$X(\pos,\pos')$ and $T(t,t')$. For $X$, we use the squared exponential
+(SE) covariance function,
 \begin{equation}\label{eq:SE}
-X(\vec{x},\vec{x}') = \exp\left(\frac{-||\vec{x} - \vec{x}'||_2^2}{2 \ell^2}\right).
-\end{equation}
-The SE covariance function is commonly used in kriging \citep[e.g,][]{Cressie1992} and Gaussian process regression \citep[e.g.,][]{Rasmussen2006}.
-In terms of geodetic applications, \citet{Kato1998} and \cite{El-Fiky1999} demonstrated that the SE accurately describes the covariance of secular GNSS derived velocities in Japan.
-The SE is a positive definite covariance function for any number of spatial dimensions.
-A Gaussian process with an SE covariance function is isotropic and has realizations that are infinitely differentiable.
-
-We consider three potential models for the temporal covariance of $u$. First, we consider the one-dimensional SE covariance function, 
+X(\pos,\pos') = \exp\left(\frac{-||\pos - \pos'||_2^2}{2 \ell^2}\right).
+\end{equation}
+The SE covariance function is commonly used in kriging
+\citep[e.g,][]{Cressie1992} and Gaussian process regression
+\citep[e.g.,][]{Rasmussen2006}. In terms of geodetic applications,
+\citet{Kato1998} and \cite{El-Fiky1999} demonstrated that the SE
+accurately describes the covariance of secular GNSS derived velocities
+in Japan. The SE is a positive definite covariance function for any
+number of spatial dimensions. A Gaussian process with an SE covariance
+function is isotropic and has realizations that are infinitely
+differentiable.
+
+We consider three potential models for the temporal covariance of $u$.
+First, we consider the one-dimensional SE covariance function,
 \begin{equation}\label{eq:TimeSE}
@@
 \end{equation}
-Note that $T$ includes the hyperparameter $\phi$, which serves to scale the covariance function $C_u$.
-Second, we consider integrated Brownian motion (IBM).
-IBM has zero mean and its covariance function can be found by integrating the covariance function for Brownian motion as
+Note that $T$ includes the hyperparameter $\phi$, which serves to
+scale the covariance function $C_u$. Second, we consider integrated
+Brownian motion (IBM). IBM has zero mean and its covariance function
+can be found by integrating the covariance function for Brownian
+motion as
 \begin{align}\label{eq:IBM}
@@
 \end{align}
-IBM has been used in the context of Kalman filtering as a non-parametric model for the time dependence of geophysical signals \citep[e.g.,][]{Segall1997,McGuire2003,Ohtani2010,Hines2016a}.
-It should be emphasized $t=0$ is a reference time at which the Gaussian process is exactly zero.
-For some geophysical signals, it is appropriate to have this reference time.
-For example, if we are trying to identify postseismic deformation then $t$ should be zero at the time of the earthquake.
-However, if we are interesting in detecting transient events, where there is no known start time, then IBM may not be an appropriate prior, and an isotropic Gaussian process should be preferred.
-In the following analysis, we make the quite arbitrary choice that $t$ is zero on the first epoch of $\mitbf{d}_*$.
-Using an earlier reference time does not change the results discussed in this section.
-Our third option for $T$ is the Wendland class of covariance functions \citep{Wendland2005}.
-Wendland covariance functions have compact support and hence their corresponding covariance matrices are sparse.
-In our analysis, we exploit this sparsity with the CHOLMOD software package \citep{Chen2008}.
-Wendland functions are positive definite in $\mathbb{R}^d$, and they describes an isotropic Gaussian process with realizations that can be differentiated $k$ times.
-The form of the covariance function depends on the choice of $d$ and $k$.
+IBM has been used in the context of Kalman filtering as a
+non-parametric model for the time dependence of geophysical signals
+\citep[e.g.,][]{Segall1997, McGuire2003, Ohtani2010, Hines2016a}. It
+should be emphasized $t=0$ is a reference time at which the Gaussian
+process is exactly zero. For some geophysical signals, it is
+appropriate to have this reference time. For example, if we are trying
+to identify postseismic deformation then $t$ should be zero at the
+time of the earthquake.  However, if we are interesting in detecting
+transient events, where there is no known start time, then IBM may not
+be an appropriate prior, and an isotropic Gaussian process should be
+preferred. In the following analysis, we make the quite arbitrary
+choice that $t$ is zero on the first epoch of $\data_*$. Using an
+earlier reference time does not change the results discussed in this
+section. Our third option for $T$ is the Wendland class of covariance
+functions \citep{Wendland2005}. Wendland covariance functions have
+compact support and hence their corresponding covariance matrices are
+sparse. In our analysis, we exploit this sparsity with the CHOLMOD
+software package \citep{Chen2008}. Wendland functions are positive
+definite in $\mathbb{R}^d$, and they describes an isotropic Gaussian
+process with realizations that can be differentiated $k$ times. The
+form of the covariance function depends on the choice of $d$ and $k$.
 We use $d=1$ since we are describing the temporal covariance of $u$.
-We use $k=2$, giving samples of $u$ continuous velocities and accelerations.
-The corresponding Wendland covariance function is
+We use $k=2$, giving samples of $u$ continuous velocities and
+accelerations. The corresponding Wendland covariance function is
 \begin{equation}\label{eq:Wendland}
@@
 
-We next determine appropriate hyperparameters for $X$ and each of the three candidate covariance functions for $T$.
-First, we clean the GNSS datasets by removing offsets at times of equipment changes and removing outliers with the method describe in Section \ref{sec:Outlier}.
-For the outlier detection algorithm, our prior model, $u$, is chosen to have a length-scale and time-scale that is able to approximately describe SSE displacements.
-We use the SE covariance function for $X$ with length-scale $\ell = 100$ km, and we use the Wendland covariance function for $T$, due to its computational efficiency, with time-scale $\tau = 0.1$ yr and $\phi = 1$ mm.
-The outlier detection algorithm is particularly effective at removing outliers for stations at high elevation (Figure \ref{fig:Outliers}), which can be adversely affected by ice or snow during the winter \citep{Lisowski2008}.
-After cleaning the dataset, we divide it into seven subsets which are four months long and each centered on the time of a SSE.
-The times of the seven SSEs are determined with tremor records from \cite{Wech2010}.
-We use the REML method to find the optimal hyperparameters for $T$ and $X$ for each subset of data.
-We choose to make each data subsets four months long because it is long enough to encompass a SSE in the Pacific Northwest, while it is short enough to still be computationally tractable.
-However, four months is too short to resolve the sinusoids in $\mitbf{d}$, and they are omitted from $\mitbf{d}$ in this REML analysis for Pacific Northwest SSEs.
-The estimated hyperparameters for $u$ are summarized in Table 1.
-Based on the interquartile ranges, the estimated hyperparameters for the SE and Wendland covariance functions do not vary significantly between SSEs.
-This suggests that the median of estimated hyperparameters should be an appropriate prior model for all Pacific Northwest SSEs.
-For the IBM model, there are several anomalously large values for $\ell$ and $\phi$, which results in large interquartile ranges.
+We next determine appropriate hyperparameters for $X$ and each of the
+three candidate covariance functions for $T$. First, we clean the GNSS
+datasets by removing offsets at times of equipment changes and
+removing outliers with the method describe in Section
+\ref{sec:Outlier}. For the outlier detection algorithm, our prior
+model, $u$, is chosen to have a length-scale and time-scale that is
+able to approximately describe SSE displacements. We use the SE
+covariance function for $X$ with length-scale $\ell = 100$ km, and we
+use the Wendland covariance function for $T$, due to its computational
+efficiency, with time-scale $\tau = 0.1$ yr and $\phi = 1$ mm.  The
+outlier detection algorithm is particularly effective at removing
+outliers for stations at high elevation (Figure \ref{fig:Outliers}),
+which can be adversely affected by ice or snow during the winter
+\citep{Lisowski2008}. After cleaning the dataset, we divide it into
+seven subsets which are four months long and each centered on the time
+of a SSE. The times of the seven SSEs are determined with tremor
+records from \cite{Wech2010}. We use the REML method to find the
+optimal hyperparameters for $T$ and $X$ for each subset of data. We
+choose to make each data subsets four months long because it is long
+enough to encompass a SSE in the Pacific Northwest, while it is short
+enough to still be computationally tractable. However, four months is
+too short to resolve the sinusoids in $\data$, and they are omitted
+from $\data$ in this REML analysis for Pacific Northwest SSEs. The
+estimated hyperparameters for $u$ are summarized in Table 1. Based on
+the interquartile ranges, the estimated hyperparameters for the SE and
+Wendland covariance functions do not vary significantly between SSEs.
+This suggests that the median of estimated hyperparameters should be
+an appropriate prior model for all Pacific Northwest SSEs. For the IBM
+model, there are several anomalously large values for $\ell$ and
+$\phi$, which results in large interquartile ranges.
 
@@
 \includegraphics{figures/outliers/outliers.pdf}
-\caption{Detrended easting component of displacements at station SC03, which is located on Mount Olympus in Washington.
-The orange markers indicate outliers that were automatically detected using the algorithm from Section \ref{sec:Outlier}.
-The error bars show one standard deviation uncertainties.
-Note that outliers tend be observed in the winter, suggesting that they were caused by snow or ice.}
+\caption{
+Detrended easting component of displacements at station SC03, which is
+located on Mount Olympus in Washington. The orange markers indicate
+outliers that were automatically detected using the algorithm from
+Section \ref{sec:Outlier}. The error bars show one standard deviation
+uncertainties. Note that outliers tend be observed in the winter,
+suggesting that they were caused by snow or ice.
+}   
 \label{fig:Outliers}
@@
 \begin{table}\label{tab:Parameters}
-\caption{Optimal hyperparameters for the prior on transient displacements determined with the REML method.
-The temporal covariance function is indicated by the ``$T$'' column.
-The SE, IBM, and Wendland covariance functions are defined in eqs.
-(\ref{eq:TimeSE}), (\ref{eq:IBM}), and (\ref{eq:Wendland}), respectively.
-The spatial covariance function, $X$, is the squared exponential (eq.
-\ref{eq:SE}) in all cases.
-The hyperparameters are estimated for each of the seven SSEs considered in this study, and the tabulated values indicate the median and interquartile ranges of estimates.
-The ``diff $\log$(REML)'' column compares the log REML likelihood to the log REML likelihood when using the SE covariance function for $T$.
-Negative values indicate that observations are more consistent with the SE covariance function.}
+\caption{
+Optimal hyperparameters for the prior on transient displacements
+determined with the REML method. The temporal covariance function is
+indicated by the ``$T$'' column. The SE, IBM, and Wendland covariance
+functions are defined in eqs. (\ref{eq:TimeSE}), (\ref{eq:IBM}), and
+(\ref{eq:Wendland}), respectively. The spatial covariance function,
+$X$, is the squared exponential (eq. \ref{eq:SE}) in all cases. The
+hyperparameters are estimated for each of the seven SSEs considered in
+this study, and the tabulated values indicate the median and
+interquartile ranges of estimates. The ``diff $\log$(REML)'' column
+compares the log REML likelihood to the log REML likelihood when using
+the SE covariance function for $T$. Negative values indicate that
+observations are more consistent with the SE covariance function.
+} 
 \begin{tabular} {l l l l l l}
@@
 
-Next we identify which covariance function for $T$ best describes the SSEs.
-One approach is to compare the REML likelihoods for each covariance function, similar to the analysis in \citet{Langbein2004}.
-In Table 1, we summarize how the log REML likelihoods for the Wendland and IBM covariance functions compare to the SE covariance function.
-Based on the differences in log REML likelihoods, the data is substantially more likely to come from a Gaussian process with a SE or Wendland covariance function than an IBM covariance function.
-The REML likelihoods do not definitively indicate whether the SE or Wendland covariance function is preferable.
-
-To further explore which covariance function for $T$ best describes SSEs, we compare the observations to the predicted displacements for each covariance function.
-We consider the data prediction vector to be $\hat{\mitbf{d}} = \left(u(\mitbf{P}) + \mitbf{G}\mitbf{a}\right)|\mitbf{d}_*$.
-Following a similar procedure as in Section \ref{sec:Method}, it can be shown that $\hat{\mitbf{d}}$ is normally distributed with mean
+Next we identify which covariance function for $T$ best describes the
+SSEs. One approach is to compare the REML likelihoods for each
+covariance function, similar to the analysis in \citet{Langbein2004}.
+In Table 1, we summarize how the log REML likelihoods for the Wendland
+and IBM covariance functions compare to the SE covariance function.
+Based on the differences in log REML likelihoods, the data is
+substantially more likely to come from a Gaussian process with a SE or
+Wendland covariance function than an IBM covariance function. The REML
+likelihoods do not definitively indicate whether the SE or Wendland
+covariance function is preferable.
+
+To further explore which covariance function for $T$ best describes
+SSEs, we compare the observations to the predicted displacements for
+each covariance function. We consider the data prediction vector to be
+$\hat{\data} = \left(u(\points) + \G\mitbf{a}\right)|\data_*$.
+Following a similar procedure as in Section \ref{sec:Method}, it can
+be shown that $\hat{\data}$ is normally distributed with mean
 \begin{equation}\label{eq:DataPredMean}
 \mitbf{\mu}_{\hat{d}} = \left[\begin{array}{cc}
-                           C_u(\mitbf{P},\mitbf{P}) & \mitbf{G} \\
+                           C_u(\points,\points) & \G \\
                            \end{array}\right]
                      \left[\begin{array}{cc}
-                           \mitbf{\Sigma} & \mitbf{G} \\
-                           \mitbf{G}^T  & \mitbf{0} \\
+                           \mitbf{\Sigma} & \G \\
+                           \G^T  & \zeros \\
                            \end{array}\right]^{-1}
                      \left[\begin{array}{c}
-                           \mitbf{d}_* \\
-                           \mitbf{0} \\
+                           \data_* \\
+                           \zeros \\
                            \end{array}\right]
@@
 \begin{equation}\label{eq:DataPredCov}
-\mitbf{C}_{\hat{\mitbf{d}}} = C_u(\mitbf{P},\mitbf{P}) - 
+\mitbf{C}_{\hat{\data}} = C_u(\points,\points) - 
                         \left[\begin{array}{cc}
-                              C_u(\mitbf{P},\mitbf{P}) & \mitbf{G} \\
+                              C_u(\points,\points) & \G \\
                               \end{array}\right]
                         \left[\begin{array}{cc}
-                              \mitbf{\Sigma} & \mitbf{G} \\
-                              \mitbf{G}^T  & \mitbf{0} \\
+                              \mitbf{\Sigma} & \G \\
+                              \G^T  & \zeros \\
                               \end{array}\right]^{-1}
                         \left[\begin{array}{c}
-                              C_u(\mitbf{P},\mitbf{P}) \\
-                              \mitbf{G}^T \\
+                              C_u(\points,\points) \\
+                              \G^T \\
                               \end{array}\right].
 \end{equation}
-We compute $\hat{\mitbf{d}}$ using SE, Wendland, and IBM covariance functions for $T$ and the median hyperparameters from Table 1.
-Figure \ref{fig:Fit} compares the easting component of $\mitbf{d}_*$ to $\hat{\mitbf{d}}$ during the winter 2015-2016 SSE at the three stations that record the strongest SSE signal, ALBH, P436, and SC02.
-Based on Figure \ref{fig:Fit}, $\hat{\mitbf{d}}$ is insensitive to the choice of covariance function for $T$.
-The predicted displacements for the IBM covariance function contain slightly more high frequency, and perhaps spurious, features.
-When comparing $\hat{\mitbf{d}}$ to $\mitbf{d}_*$ at station ALBH and SC02, it appears that $\hat{\mitbf{d}}$ slightly underestimates the rate of deformation during the SSE, regardless of the chosen covariance function.
-This over-smoothing could indicate that the chosen time-scale hyperparameter, $\tau$, is too large.
-The predicted displacements at the remaining stations, which record more subtle SSE deformation, seem to faithfully describe $\mitbf{d}_*$.
-
-For our estimates of transient strain discussed in the next section, we ultimately settle on the Wendland covariance function for $T$ and use the median values from Table 1 for the hyperparameters.
-We choose the Wendland covariance function over the SE covariance function because of its computational advantages.
+We compute $\hat{\data}$ using SE, Wendland, and IBM covariance
+functions for $T$ and the median hyperparameters from Table 1. Figure
+\ref{fig:Fit} compares the easting component of $\data_*$ to
+$\hat{\data}$ during the winter 2015-2016 SSE at the three stations
+that record the strongest SSE signal, ALBH, P436, and SC02. Based on
+Figure \ref{fig:Fit}, $\hat{\data}$ is insensitive to the choice of
+covariance function for $T$. The predicted displacements for the IBM
+covariance function contain slightly more high frequency, and perhaps
+spurious, features. When comparing $\hat{\data}$ to $\data_*$ at
+station ALBH and SC02, it appears that $\hat{\data}$ slightly
+underestimates the rate of deformation during the SSE, regardless of
+the chosen covariance function. This over-smoothing could indicate
+that the chosen time-scale hyperparameter, $\tau$, is too large. The
+predicted displacements at the remaining stations, which record more
+subtle SSE deformation, seem to faithfully describe $\data_*$.
+
+For our estimates of transient strain discussed in the next section,
+we ultimately settle on the Wendland covariance function for $T$ and
+use the median values from Table 1 for the hyperparameters. We choose
+the Wendland covariance function over the SE covariance function
+because of its computational advantages.
 
@@
 \includegraphics{figures/signal_fit/signal-fit.pdf}
-\caption{Easting component of the observed displacements, $\mitbf{d}_*$, and predicted displacements, $\hat{\mitbf{d}}$, during an SSE at three stations where the SSE signal is strongest.
-The predicted displacements are shown for when $T$ is a squared exponential (SE), Wendland, and integrated Brownian motion (IBM) covariance function.
-The one standard deviation uncertainties are shown for the observations and the SE predictions.
-For clarity, uncertainties are not shown for the IBM and Wendland predictions, but they are nearly equivalent to the uncertainties for the SE predictions.
-The SE and Wendland predictions are practically indistinguishable.}
+\caption{
+Easting component of the observed displacements, $\data_*$, and
+predicted displacements, $\hat{\data}$, during an SSE at three
+stations where the SSE signal is strongest. The predicted
+displacements are shown for when $T$ is a squared exponential (SE),
+Wendland, and integrated Brownian motion (IBM) covariance function.
+The one standard deviation uncertainties are shown for the
+observations and the SE predictions. For clarity, uncertainties are
+not shown for the IBM and Wendland predictions, but they are nearly
+equivalent to the uncertainties for the SE predictions. The SE and
+Wendland predictions are practically indistinguishable.
+}   
 \label{fig:Fit}
@@
 
+%% TRANSIENT STRAIN RATES
 \subsection{Transient Strain Rates}\label{sec:Results}
-Having established a noise model and a prior for transient displacements, we use the cleaned GNSS dataset to calculate transient strain rates in the Puget Sound region.
-We calculate transient strain rates for each day from January 1, 2010 to May 15, 2017.
-The strain rates are estimates at a grid of points spanning the study area.
-In Figure \ref{fig:StrainMap} we show the transient strain rates on January 1, 2016, which coincides with the height of the winter 2015-2016 SSE.
-We have included an animation showing the map view of strain rates through time as supplementary material.
-The strain rates shown in Figure \ref{fig:StrainMap} are generally similar to the strain rates during the other six SSEs considered in this study.
-The SSEs cause trench perpendicular compression in the Olympic Peninsula and extension east of Puget Sound.
-The strain transitions from compression to extension around the southern tip of Vancouver Island, which coincides with the location of thrust slip for SSEs in the Puget Sound region \citep[e.g.,][]{Dragert2001,Wech2009,Schmidt2010}.
-Thus, this pattern of strain is to be expected.
-During the period in between SSEs, secular strain rates indicate trench perpendicular compression throughout this study region \citep{Murray2000,McCaffrey2007,McCaffrey2013}.
-When comparing inferred strain rates from SSEs to the secular strain rates, we see that SSEs are concentrating tectonically accumulated strain energy towards the trench, and presumably pushing the subduction zone closer to failure.
-
-%A key difference between the strain inferred here and strain that can be derived from fault slip models is that our estimated strain rates are not based on an assumed physical model.
-In contrast, fault slip models can be biased by errors in the assumed fault geometry or lithospheric rheology.
-Moreover, the degrees of freedom in fault slip models usually cannot be constrained by GNSS data alone, and it is necessary to impose regularization which further biases the results.
-Since our estimated strain rates lack such systematic errors, we can be more confident that our solution is unbiased and has meaningful uncertainties.
+
+Having established a noise model and a prior for transient
+displacements, we use the cleaned GNSS dataset to calculate transient
+strain rates in the Puget Sound region.  We calculate transient strain
+rates for each day from January 1, 2010 to May 15, 2017. The strain
+rates are estimates at a grid of points spanning the study area. In
+Figure \ref{fig:StrainMap} we show the transient strain rates on
+January 1, 2016, which coincides with the height of the winter
+2015-2016 SSE. We have included an animation showing the map view of
+strain rates through time as supplementary material. The strain rates
+shown in Figure \ref{fig:StrainMap} are generally similar to the
+strain rates during the other six SSEs considered in this study. The
+SSEs cause trench perpendicular compression in the Olympic Peninsula
+and extension east of Puget Sound. The strain transitions from
+compression to extension around the southern tip of Vancouver Island,
+which coincides with the location of thrust slip for SSEs in the Puget
+Sound region \citep[e.g.,][]{Dragert2001,Wech2009,Schmidt2010}. Thus,
+this pattern of strain is to be expected. During the period in between
+SSEs, secular strain rates indicate trench perpendicular compression
+throughout this study region \citep{Murray2000, McCaffrey2007,
+McCaffrey2013}. When comparing inferred strain rates from SSEs to the
+secular strain rates, we see that SSEs are concentrating tectonically
+accumulated strain energy towards the trench, and presumably pushing
+the subduction zone closer to failure.
+
+%A key difference between the strain inferred here and strain that can
+%be derived from fault slip models is that our estimated strain rates
+%are not based on an assumed physical model. In contrast, fault slip
+%models can be biased by errors in the assumed fault geometry or
+%lithospheric rheology. Moreover, the degrees of freedom in fault slip
+%models usually cannot be constrained by GNSS data alone, and it is
+%necessary to impose regularization which further biases the results.
+%Since our estimated strain rates lack such systematic errors, we can
+%be more confident that our solution is unbiased and has meaningful
+%uncertainties.
 
@@
 \includegraphics{figures/strain_map/strain-map.pdf}
-\caption{Estimated transient strain rates during the Winter 2015-2016 SSE.
-Strain glyphs show the normal strain rate along each azimuth, where orange indicates compression and blue indicates extension.
-The shaded regions indicate one standard deviation uncertainties in the normal strain rates.}
+\caption{
+Estimated transient strain rates during the Winter 2015-2016 SSE.
+Strain glyphs show the normal strain rate along each azimuth, where
+orange indicates compression and blue indicates extension. The shaded
+regions indicate one standard deviation uncertainties in the normal
+strain rates.
+}   
 \label{fig:StrainMap}
@@
 
-In Figure \ref{fig:StrainTs} we show the time dependence of estimated transient strain rates at a position on the Olympic Peninsula, where transient strain rates from SSEs are largest.
-To verify that the estimated transient strain rates are accurately identifying geophysical signal, we compare the signal-to-noise ratio from eq.
-(\ref{eq:SNR}) to the frequency of seismic tremor (Figure \ref{fig:StrainMag}).
-A signal-to-noise ratio greater than ${\sim}3$ can be interpreted as a detected geophysical signal.
-We detect nine distinct events, which each correspond to peaks in seismic tremor.
-The smaller events detected in August 2014 and February 2017 can be considered inter-SSE events.
-They were not among the SSEs used to constrain the prior covariance function.
-In between peaks in seismic tremor, the signal-to-noise ratio is consistently between 0 and 2, suggesting that all the transient strain detected at this location is associated with SSEs and inter-SSE events.
+In Figure \ref{fig:StrainTs} we show the time dependence of estimated
+transient strain rates at a position on the Olympic Peninsula, where
+transient strain rates from SSEs are largest. To verify that the
+estimated transient strain rates are accurately identifying
+geophysical signal, we compare the signal-to-noise ratio from eq.
+(\ref{eq:SNR}) to the frequency of seismic tremor (Figure
+\ref{fig:StrainMag}). A signal-to-noise ratio greater than ${\sim}3$
+can be interpreted as a detected geophysical signal. We detect nine
+distinct events, which each correspond to peaks in seismic tremor. The
+smaller events detected in August 2014 and February 2017 can be
+considered inter-SSE events. They were not among the SSEs used to
+constrain the prior covariance function. In between peaks in seismic
+tremor, the signal-to-noise ratio is consistently between 0 and 2,
+suggesting that all the transient strain detected at this location is
+associated with SSEs and inter-SSE events.
 
@@
 \includegraphics{figures/strain_ts/strain-ts.pdf}
-\caption{Three components of the transient horizontal strain rate tensor estimated at the position shown in Figure \ref{fig:Context}.
-The shaded regions indicate one standard deviation uncertainty.}
+\caption{
+Three components of the transient horizontal strain rate tensor
+estimated at the position shown in Figure \ref{fig:Context}. The
+shaded regions indicate one standard deviation uncertainty.
+}   
 \label{fig:StrainTs}
@@
 \includegraphics{figures/strain_ts/mag-ts.pdf}
-\caption{(top) Signal-to-noise ratio (eq. \ref{eq:SNR}) at the position shown in Figure \ref{fig:Context}. (bottom) Frequency of tremors in the region shown in Figure \ref{fig:Context}.}   
+\caption{
+(top) Signal-to-noise ratio (eq. \ref{eq:SNR}) at the position shown
+in Figure \ref{fig:Context}. (bottom) Frequency of tremors in the
+region shown in Figure \ref{fig:Context}.
+}   
 \label{fig:StrainMag}
@@
 
-The results we have presented thus far indicate that we are identifying the strain that we should expect to see.
-There are, however, subtle features in our estimated transient strain rates which we were not expecting.
-For example, some SSEs are preceded by a brief period of east-west extension on the Olympic Peninsula.
-This feature can be seen several days before the summer 2012 and winter 2015-2016 SSEs in Figure \ref{fig:StrainTs} as well as in the supplementary animation.
-While this deformation is noteworthy, a discussion on the mechanisms causing it is outside the scope of this study.
-
+The results we have presented thus far indicate that we are
+identifying the strain that we should expect to see. There are,
+however, subtle features in our estimated transient strain rates which
+we were not expecting. For example, some SSEs are preceded by a brief
+period of east-west extension on the Olympic Peninsula. This feature
+can be seen several days before the summer 2012 and winter 2015-2016
+SSEs in Figure \ref{fig:StrainTs} as well as in the supplementary
+animation. While this deformation is noteworthy, a discussion on the
+mechanisms causing it is outside the scope of this study.
+
+%% DISCUSSION
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Discussion}\label{sec:Discussion}
-Our results demonstrate that GPR is an effective tool for estimating transient strain from GNSS data, which can then be used to detect geophysical processes.
-One may argue that geophysical signal can also be detected by merely inspecting the GNSS displacement time series.
-Indeed, the SSEs identified in Figure \ref{fig:StrainMag} do produce visible displacements in the GNSS data.
-However, the GNSS data also contains outliers and non-tectonic deformation that is localized to individual stations.
-In contrast, our estimates of transient strain only identify features that are sufficiently spatially and temporally coherent, based on our chosen prior model.
-Furthermore, our estimates of transient strain are insensitive to common mode noise, which is highly spatially correlated noise resulting from factors such as reference frame error.
-Common mode noise can obscure geophysical signal in the GNSS data, but it gets canceled out when computing the transient strain.
-Lastly, our estimates of transient strain rates are a spatial and temporal derivative of displacements, and thus any geophysical signal in the transient strain rates tends to be more pronounced than in the GNSS data.
-For these reasons, we argue that transient strain rates estimated with the method described in Section \ref{sec:Method} can illuminate geophysical signal that may not be discernible from the noise in the GNSS displacement data.
-
-In addition to detecting geophysical processes, the GNSS derived transient strain rates can be used to better understand the data from borehole strain meters (BSMs).
-The Plate Boundary Observatory maintains about forty BSMs in the Pacific Northwest, and it has been demonstrated that BSMs are able to record transient geophysical events such as SSEs \citep[e.g.,][]{Dragert2011}.
-However, there are complications that prevent BSM data from being used quantitatively in geophysical studies.
-One difficulty is that BSM data should be calibrated with a well known strain source, such as diurnal and semidiurnal tides \citep{Hart1996,Roeloffs2010,Hodgkinson2013}.
-Unfortunately, the tidal forces at BSMs which record SSEs are strongly influenced by local bodies of water such as the Straight of Juan de Fuca, making it difficult to form a theoretical prediction of tidal strains \citep{Roeloffs2010}.
-Another complication is that noise in BSM data is not well understood.
-The noise consists, in part, of a long-term decay resulting from the instrument equilibrating with the surrounding rock \citep{Gladwin1987}.
-Typically, this noise is dealt with in an ad-hoc manner by fitting and removing exponentials and low-order polynomials.
-We envision that the GNSS derived strain rates from this paper can be used as a reference strain for calibrating BSM data and quantify its noise.
-
-There is potential for a more thorough analysis of the spatio-temporal noise in GNSS data, $\eta$, than what was performed in Section \ref{sec:NoiseModel}.
-We did not explore the spatial covariance of $\eta$, which would describe common mode noise.
-We are able to ignore common mode error in this study; however, for other geophysical studies based on GNSS data, such as fault slip inversions, it may be necessary to incorporate a spatially covarying noise model \citep[e.g.,][]{Miyazaki2003}.
-We can also improve upon the seasonal model used in this study, which consists of four spatially uncorrelated sinusoids for each station.
-We did not explore the spatial covariance of seasonal deformation or the temporal roughness (i.e., the number of sinusoids needed to describe the observations).
-The periodic Gaussian process \citep{Mackay1998} is an alternative model for seasonal deformation and is well suited for exploring the roughness of seasonal deformation.
-The periodic Gaussian process has zero mean and the covariance function
+Our results demonstrate that GPR is an effective tool for estimating
+transient strain from GNSS data, which can then be used to detect
+geophysical processes. One may argue that geophysical signal can also
+be detected by merely inspecting the GNSS displacement time series.
+Indeed, the SSEs identified in Figure \ref{fig:StrainMag} do produce
+visible displacements in the GNSS data. However, the GNSS data also
+contains outliers and non-tectonic deformation that is localized to
+individual stations. In contrast, our estimates of transient strain
+only identify features that are sufficiently spatially and temporally
+coherent, based on our chosen prior model. Furthermore, our estimates
+of transient strain are insensitive to common mode noise, which is
+highly spatially correlated noise resulting from factors such as
+reference frame error. Common mode noise can obscure geophysical
+signal in the GNSS data, but it gets canceled out when computing the
+transient strain. Lastly, our estimates of transient strain rates are
+a spatial and temporal derivative of displacements, and thus any
+geophysical signal in the transient strain rates tends to be more
+pronounced than in the GNSS data. For these reasons, we argue that
+transient strain rates estimated with the method described in Section
+\ref{sec:Method} can illuminate geophysical signal that may not be
+discernible from the noise in the GNSS displacement data.
+
+In addition to detecting geophysical processes, the GNSS derived
+transient strain rates can be used to better understand the data from
+borehole strain meters (BSMs). The Plate Boundary Observatory
+maintains about forty BSMs in the Pacific Northwest, and it has been
+demonstrated that BSMs are able to record transient geophysical events
+such as SSEs \citep[e.g.,][]{Dragert2011}. However, there are
+complications that prevent BSM data from being used quantitatively in
+geophysical studies. One difficulty is that BSM data should be
+calibrated with a well known strain source, such as diurnal and
+semidiurnal tides \citep{Hart1996,Roeloffs2010,Hodgkinson2013}.
+Unfortunately, the tidal forces at BSMs which record SSEs are strongly
+influenced by local bodies of water such as the Straight of Juan de
+Fuca, making it difficult to form a theoretical prediction of tidal
+strains \citep{Roeloffs2010}. Another complication is that noise in
+BSM data is not well understood. The noise consists, in part, of a
+long-term decay resulting from the instrument equilibrating with the
+surrounding rock \citep{Gladwin1987}. Typically, this noise is dealt
+with in an ad-hoc manner by fitting and removing exponentials and
+low-order polynomials. We envision that the GNSS derived strain rates
+from this paper can be used as a reference strain for calibrating BSM
+data and quantify its noise.
+
+There is potential for a more thorough analysis of the spatio-temporal
+noise in GNSS data, $\eta$, than what was performed in Section
+\ref{sec:NoiseModel}. We did not explore the spatial covariance of
+$\eta$, which would describe common mode noise. We are able to ignore
+common mode error in this study; however, for other geophysical
+studies based on GNSS data, such as fault slip inversions, it may be
+necessary to incorporate a spatially covarying noise model
+\citep[e.g.,][]{Miyazaki2003}. We can also improve upon the seasonal
+model used in this study, which consists of four spatially
+uncorrelated sinusoids for each station. We did not explore the
+spatial covariance of seasonal deformation or the temporal roughness
+(i.e., the number of sinusoids needed to describe the observations).
+The periodic Gaussian process \citep{Mackay1998} is an alternative
+model for seasonal deformation and is well suited for exploring the
+roughness of seasonal deformation.  The periodic Gaussian process has
+zero mean and the covariance function
 \begin{equation}\label{eq:Periodic}
@@
 \end{equation}
-Realizations have annual periodicity and the roughness is controlled by $\tau$.
-Decreasing $\tau$ has the same effect as including higher frequency sinusoids in the seasonal model.
-The optimal value for $\tau$ can be found with the REML method as described in Section \ref{sec:NoiseModel}.
-
-The transient strain rates estimated in this study are constrained by about seven years of daily displacement observations from 94 GNSS stations.
-It can be computationally intensive to evaluate eqs.
-(\ref{eq:PosteriorMean3}) and (\ref{eq:PosteriorCov3}) for a dataset with this size.
-We significantly reduce the amount of memory needed to estimate transient strain rates by describing the temporal covariance of displacements with a compact Wendland covariance function.
-Using a compact covariance function for our prior turns eqs.
-(\ref{eq:PosteriorMean3}) and (\ref{eq:PosteriorCov3}) into sparse systems of equations, which we then solve with CHOLMOD.
-CHOLMOD is designed for solving sparse, positive definite systems of equations.
-The matrix being inverted in eqs.
-(\ref{eq:PosteriorMean3}) and (\ref{eq:PosteriorCov3}) is not positive definite; however, we can use another partitioned matrix inversion identity from \citet{Press2007} to partition it into positive definite submatrices to be inverted.
-Even when using a compact covariance function, it may still be necessary to reduce the computational burden by dividing the data into subsets and evaluating transient strain rates for each subset.
-
+Realizations have annual periodicity and the roughness is controlled
+by $\tau$. Decreasing $\tau$ has the same effect as including higher
+frequency sinusoids in the seasonal model. The optimal value for
+$\tau$ can be found with the REML method as described in Section
+\ref{sec:NoiseModel}.
+
+The transient strain rates estimated in this study are constrained by
+about seven years of daily displacement observations from 94 GNSS
+stations. It can be computationally intensive to evaluate eqs.
+(\ref{eq:PosteriorMean3}) and (\ref{eq:PosteriorCov3}) for a dataset
+with this size. We significantly reduce the amount of memory needed to
+estimate transient strain rates by describing the temporal covariance
+of displacements with a compact Wendland covariance function. Using a
+compact covariance function for our prior turns eqs.
+(\ref{eq:PosteriorMean3}) and (\ref{eq:PosteriorCov3}) into sparse
+systems of equations, which we then solve with CHOLMOD. CHOLMOD is
+designed for solving sparse, positive definite systems of equations.
+The matrix being inverted in eqs. (\ref{eq:PosteriorMean3}) and
+(\ref{eq:PosteriorCov3}) is not positive definite; however, we can use
+another partitioned matrix inversion identity from \citet{Press2007}
+to partition it into positive definite submatrices to be inverted.
+Even when using a compact covariance function, it may still be
+necessary to reduce the computational burden by dividing the data into
+subsets and evaluating transient strain rates for each subset.
+
+%% CONCLUSION
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Conclusion}\label{sec:Conclusion}
-In this paper we propose using Gaussian process regression (GPR) to estimate transient strain rates from GNSS data.
-Most other methods for estimating strain rates assume a parametric representation of deformation, which can bias the results if the parameterization is not chosen carefully.
-Here we assume a stochastic, rather than parametric, prior model for displacements.
-Our prior model describes how much we expect transient displacements to covary spatially and temporally.
-If we know nothing about the underlying signal that we are trying to recover, then the prior model can be chosen objectively with maximum likelihood methods.
-Because GPR is a Bayesian method, the uncertainties on our estimated transient strain rates are well quantified, allowing one to discern geophysical signal from noise.
-We demonstrate that GPR is an effective tool for detecting geophysical phenomena, such as slow slip events, in our application to GNSS data from the Pacific Northwest.
-One limitation with GPR is that it is not robust against outliers.
-To overcome this limitation, we have introduced an effective pre-processing method for identifying and removing outliers from GNSS datasets.
-Another complication with GPR is that it usually involves inverting a dense matrix where the number of rows and columns is equal to the number of observations.
-This is prohibitive when using several years of daily GNSS observations from a network of several hundred stations.
-We significantly reduce the computational burden of GPR by using compact Wendland covariance function to describe our prior model.
-While this paper just focuses on estimating transient strain rates, we believe that GPR is a powerful tool that can be applied to a wide range of geophysical problems.
-
+In this paper we propose using Gaussian process regression (GPR) to
+estimate transient strain rates from GNSS data. Most other methods for
+estimating strain rates assume a parametric representation of
+deformation, which can bias the results if the parameterization is not
+chosen carefully. Here we assume a stochastic, rather than parametric,
+prior model for displacements. Our prior model describes how much we
+expect transient displacements to covary spatially and temporally. If
+we know nothing about the underlying signal that we are trying to
+recover, then the prior model can be chosen objectively with maximum
+likelihood methods. Because GPR is a Bayesian method, the
+uncertainties on our estimated transient strain rates are well
+quantified, allowing one to discern geophysical signal from noise. We
+demonstrate that GPR is an effective tool for detecting geophysical
+phenomena, such as slow slip events, in our application to GNSS data
+from the Pacific Northwest. One limitation with GPR is that it is not
+robust against outliers. To overcome this limitation, we have
+introduced an effective pre-processing method for identifying and
+removing outliers from GNSS datasets. Another complication with GPR is
+that it usually involves inverting a dense matrix where the number of
+rows and columns is equal to the number of observations. This is
+prohibitive when using several years of daily GNSS observations from a
+network of several hundred stations. We significantly reduce the
+computational burden of GPR by using compact Wendland covariance
+function to describe our prior model. While this paper just focuses on
+estimating transient strain rates, we believe that GPR is a powerful
+tool that can be applied to a wide range of geophysical problems.
+
+%% ACKNOWLEDGEMENTS
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Acknowledgements}
-This material is based upon work supported by the National Science Foundation under grant EAR 1245263.
-The EarthScope Plate Boundary Observatory data is provided by UNAVCO through the GAGE Facility with support from the National Science Foundation (NSF) and National Aeronautics and Space Administration (NASA) under NSF Cooperative Agreement EAR-1261833.
-An implementation of the method described in this paper is named Python-based Geodetic Network Strain software (PyGeoNS).
-PyGeoNS is distributed under the MIT License and can be found at www.github.com/treverhines/PyGeoNS.
-
+This material is based upon work supported by the National Science
+Foundation under grant EAR 1245263. The EarthScope Plate Boundary
+Observatory data is provided by UNAVCO through the GAGE Facility with
+support from the National Science Foundation (NSF) and National
+Aeronautics and Space Administration (NASA) under NSF Cooperative
+Agreement EAR-1261833. An implementation of the method described in
+this paper is named Python-based Geodetic Network Strain software
+(PyGeoNS). PyGeoNS is distributed under the MIT License and can be
+found at www.github.com/treverhines/PyGeoNS.
+
+%% APPENDIX
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \appendix
-
 \section{Outlier detection algorithm}
-Our outlier detection algorithm is loosely based on the data editing algorithm from \citet{Acheson1975}.
-Let $\mitbf{d}^*$ denote all $n$ GNSS displacement observations for a single directional component, which have been made at positions and times $\mitbf{P}$.
-We describe $\mitbf{d}^*$ as a realization of the random vector
+
+Our outlier detection algorithm is loosely based on the data editing
+algorithm from \citet{Acheson1975}. Let $\data^*$ denote all $n$ GNSS
+displacement observations for a single directional component, which
+have been made at positions and times $\points$.  We describe
+$\data^*$ as a realization of the random vector
 \begin{equation}\label{eq:OutlierDataVec}
-\mitbf{d} = \mitbf{G}\mitbf{m} + v(\mitbf{P}) + \mitbf{w},
-\end{equation}
-where $\mitbf{G}$ and $\mitbf{m}$ are the same as in eq.
-(\ref{eq:DataVec}), $v$ is a Gaussian process distributed as $\mathcal{GP}(0,C_v)$, and $\mitbf{w}$ is a vector of uncorrelated Gaussian noise with known standard deviations $\mitbf{\sigma} = [\sigma_1, \sigma_2, ..., \sigma_n]$.
-The Gaussian process $v$ is intended to describe transient features in the data that cannot be explained by the linear trend or seasonal terms in $\mitbf{G}$.
-We let the temporal covariance of $v$ be a squared exponential, and we let $v$ be spatially uncorrelated so that,
+\data = \G\mitbf{m} + v(\points) + \mitbf{w},
+\end{equation}
+where $\G$ and $\mitbf{m}$ are the same as in eq. (\ref{eq:DataVec}),
+$v$ is a Gaussian process distributed as $\mathcal{GP}(0,C_v)$, and
+$\mitbf{w}$ is a vector of uncorrelated Gaussian noise with known
+standard deviations $\mitbf{\sigma} = [\sigma_1, \sigma_2, ...,
+\sigma_n]$. The Gaussian process $v$ is intended to describe transient
+features in the data that cannot be explained by the linear trend or
+seasonal terms in $\G$. We let the temporal covariance of $v$ be a
+squared exponential, and we let $v$ be spatially uncorrelated so that,
 \begin{equation}
-C_v(p,p') = \phi^2\exp\left(\frac{-|t - t'|^2}{2\tau^2}\right) \delta_{\vec{x},\vec{x}'},
-\end{equation}
-where $\delta_{\vec{x},\vec{x}'} = 1$ if $\vec{x} = \vec{x}'$ and $0$ otherwise.
-The spatial covariance of $v$ has little effect on the detected outliers, and so we have assumed that $v$ is spatially uncorrelated for simplicity.
-Based on our experience, $v$ can reasonably describe most transient features in the data when we set $\phi = 1$ mm and $\tau = 10$ days.
-
-Our goal is to find the index set of non-outliers in $\mitbf{d}^*$, which we denote as $\mitbf{\Omega}$.
-We use a tilde to indicate that an array only contains elements corresponding to $\mitbf{\Omega}$ (e.g., the vector of non-outlier observations is denoted $\tilde{\mitbf{d}}^* = [d_i^*]_{i \in \mitbf{\Omega}}$).
-The outliers are identified iteratively, and we initiate $\mitbf{\Omega}$ with all $n$ indices.
-We consider outliers to be data that are poorly explained by the model $\mitbf{G}\mitbf{m} + v(\mitbf{P})$, which is determine by the residual vector
+C_v(p,p') = \phi^2\exp\left(\frac{-|t - t'|^2}{2\tau^2}\right) \delta_{\pos,\pos'},
+\end{equation}
+where $\delta_{\pos,\pos'} = 1$ if $\pos = \pos'$ and $0$ otherwise.
+The spatial covariance of $v$ has little effect on the detected
+outliers, and so we have assumed that $v$ is spatially uncorrelated
+for simplicity. Based on our experience, $v$ can reasonably describe
+most transient features in the data when we set $\phi = 1$ mm and
+$\tau = 10$ days.
+
+Our goal is to find the index set of non-outliers in $\data^*$, which
+we denote as $\mitbf{\Omega}$. We use a tilde to indicate that an
+array only contains elements corresponding to $\mitbf{\Omega}$ (e.g.,
+the vector of non-outlier observations is denoted $\tilde{\data}^* =
+[d_i^*]_{i \in \mitbf{\Omega}}$). The outliers are identified
+iteratively, and we initiate $\mitbf{\Omega}$ with all $n$ indices. We
+consider outliers to be data that are poorly explained by the model
+$\G\mitbf{m} + v(\points)$, which is determine by the residual vector
 \begin{align}\label{eq:Residual} 
-\mitbf{r} &= \mitbf{d}^* - \mathrm{E}\left[\Big(\mitbf{G}\mitbf{m} + v(\mitbf{P})\Big) \Big| \Big(\tilde{\mitbf{d}} = \tilde{\mitbf{d}}^*\Big)\right] \\
-          &= \mitbf{d}^*  - 
+\mitbf{r} &= \data^* - \E{\Big(\G\mitbf{m} + v(\points)\Big) \Big| \Big(\tilde{\data} =
+                                                                        \tilde{\data}^*\Big)} \\
+          &= \data^*  - 
             \left[\begin{array}{cc}
-                  C_v(\mitbf{P},\tilde{\mitbf{P}}) & \mitbf{G} \nonumber \\
+                  C_v(\points,\tilde{\points}) & \G \nonumber \\
                   \end{array}\right]
             \left[\begin{array}{cc}
-                  C_v(\tilde{\mitbf{P}},\tilde{\mitbf{P}}) + \mathrm{diag}(\tilde{\mitbf{\sigma}}^2) & \tilde{\mitbf{G}} \\
-                  \tilde{\mitbf{G}}^T  & \mitbf{0} \\
+                  C_v(\tilde{\points},\tilde{\points}) + \mathrm{diag}(\tilde{\mitbf{\sigma}}^2) & \tilde{\G} \\
+                  \tilde{\G}^T  & \zeros \\
                   \end{array}\right]^{-1}
             \left[\begin{array}{c}
-                  \tilde{\mitbf{d}}^* \\
-                  \mitbf{0} \\
+                  \tilde{\data}^* \\
+                  \zeros \\
                   \end{array}\right].
 \end{align}
-Data with abnormally large residuals are identified as outliers.
-For each iteration, we compute $\mitbf{r}$ and then update $\mitbf{\Omega}$ so that it contains the indices of $\mitbf{r}$ whose weighted values are less than $\lambda$ times the weighted root mean square of $\tilde{\mitbf{r}}$,
+Data with abnormally large residuals are identified as outliers. For
+each iteration, we compute $\mitbf{r}$ and then update
+$\mitbf{\Omega}$ so that it contains the indices of $\mitbf{r}$ whose
+weighted values are less than $\lambda$ times the weighted root mean
+square of $\tilde{\mitbf{r}}$,
 \begin{equation}\label{eq:Update}
@@
 \end{equation} 
-Iterations continue until the new $\mitbf{\Omega}$ is the same as the previous $\mitbf{\Omega}$.
-
-The outlier detection algorithm is demonstrated in Figure \ref{fig:Outliers}.
-For the demonstration, we use the easting component of displacements at a single station, SC03, which is located on Mt.
-Olympus in Washington state.
-Station SC03 records anomalous observations during the winter, presumably because of snow and ice accumulation, and we want to remove these observations.
-The station also records periodic westward motion from slow slip events, and we want to keep this deformation intact.
-The detected outliers are shown in Panel B.
-For comparison, we also show the detected outliers when we do not include the Gaussian process $v$ in our model for the data (Panel A).
-When $v$ is not included, real transient deformation resulting from slow slip events is erroneously identified as outliers.
-When $v$ is included, the identified outliers only consist of the anomalous deformation that lacks temporal continuity.
-It should be noted that we use $\lambda = 2.5$ for this demonstration, which causes the outlier detection algorithm to be particularly aggressive.
-In Section \ref{sec:Cascadia}, we clean the data using a more tolerant $\lambda = 4.0$.
+Iterations continue until the new $\mitbf{\Omega}$ is the same as the
+previous $\mitbf{\Omega}$.
+
+The outlier detection algorithm is demonstrated in Figure
+\ref{fig:Outliers}. For the demonstration, we use the easting
+component of displacements at a single station, SC03, which is located
+on Mt. Olympus in Washington state. Station SC03 records anomalous
+observations during the winter, presumably because of snow and ice
+accumulation, and we want to remove these observations. The station
+also records periodic westward motion from slow slip events, and we
+want to keep this deformation intact. The detected outliers are shown
+in Panel B. For comparison, we also show the detected outliers when we
+do not include the Gaussian process $v$ in our model for the data
+(Panel A). When $v$ is not included, real transient deformation
+resulting from slow slip events is erroneously identified as outliers.
+When $v$ is included, the identified outliers only consist of the
+anomalous deformation that lacks temporal continuity.  It should be
+noted that we use $\lambda = 2.5$ for this demonstration, which causes
+the outlier detection algorithm to be particularly aggressive. In
+Section \ref{sec:Cascadia}, we clean the data using a more tolerant
+$\lambda = 4.0$.
 
@@
 \includegraphics{figures/outliers_new/outliers.pdf}
-\caption{Outliers detected in the easting component of displacements at station SC03.
-The orange markers indicate detected outliers.
-The orange line is the best fit model to the data, which is used to compute the residual vector $\mitbf{r}$.
-The model being fit to the data in Panel A is $\mitbf{G}\mitbf{m}$, and the model in Panel B is $\mitbf{G}\mitbf{m} + v(\mitbf{P})$.}
+\caption{
+Outliers detected in the easting component of displacements at station
+SC03. The orange markers indicate detected outliers. The orange line
+is the best fit model to the data, which is used to compute the
+residual vector $\mitbf{r}$. The model being fit to the data in Panel
+A is $\G\mitbf{m}$, and the model in Panel B is $\G\mitbf{m} +
+v(\points)$.
+}
 \label{fig:Outliers}

### 2017-10-22 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 \newcommand{\strain}{\dot\varepsilon} % strain rate
-\newcommand{\E}[1]{\e\left[ #1 \right]} % Expected value
+\newcommand{\E}[1]{\mathrm{E}\left[ #1 \right]} % Expected value
 \newcommand{\Cov}[1]{\mathrm{Cov}\left[ #1 \right]} % Expected value
@@
 
+% Why we care about determining strain
+
 Crustal strain rates are fundamentally important quantities for
@@
 from GNSS data is an active area of research.
+
+% Existing methods for determining secular strain from GPS
 
@@
 
+% Existing methods for determining transient strain from GPS
+
 The aforementioned studies are concerned with estimating secular
@@
 
+% What we are doing in this paper and why it improves upon existing
+% methods
+
 Here we propose using Gaussian process regression (GPR)
@@
 stations, GPR is an ideal tool for synthesizing discrete GNSS data
-into a spatially and temporally continuous representation of surface
-deformation. GPR is Bayesian in that we describe our prior
-understanding of the geophysical signal with a Gaussian process. A
-Gaussian process is a continuous stochastic process whose value at any
-collection of points can be described as a Gaussian random vector.
-Just as a Gaussian random vector is fully determined by a mean vector
-and a covariance matrix, a Gaussian process is fully determined by a
-mean function and a covariance function. For example, Brownian motion,
-$B(t)$, is a well known Gaussian process in $\mathbb{R}^1$ which has
-the mean function $\E{B(t)} = 0$ and the covariance function
-$\Cov{B(t),B(t')} = \min(t,t')$, where $t,t' \ge 0$. If there is no
-information available to help choose an appropriate prior Gaussian
-process, then maximum likelihood methods can be used to objectively
-choose one that is most consistent with the data. Once a prior is
-selected, we incorporate GNSS observations with the prior to form a
-posterior estimate of transient strain. The posterior transient strain
-is also a Gaussian process, and we can use its distribution to
-confidently discern geophysical signal from noise. We use GPR to infer
-strain resulting from SSEs in the Pacific Northwest, demonstrating
-that GPR is an effective tool for detecting transient geophysical
-processes.
+into a spatially and temporally continuous representation of transient
+strain. GPR is Bayesian in that we use a prior model to control the
+spatial and temporal roughness of the inferred transient strain. The
+prior is specified as a stochastic process, namely a Gaussian process.
+If there is no information available to help choose an appropriate
+prior Gaussian process, then maximum likelihood methods can be used to
+objectively choose one that is most consistent with the data. We use
+GPR to infer transient strain resulting from SSEs in the Pacific
+Northwest, demonstrating that GPR is an effective tool for detecting
+transient geophysical processes.
 
@@
 \section{Estimating Transient Strain Rates}\label{sec:Method}
+
+% Overview of what we are trying to do in this section and how
 
@@
 simplicity, we assume that the study region is sufficiently small that
-$\pos$ can be considered a point in a 2-D map projection that is
-aligned with the cardinal directions. We determine $\strain$ by
-spatially and temporally differentiating estimates of transient
-displacements, which are constrained by GNSS data.
+$\pos$ can be considered a point in a 2-D Cartesian map projection
+that is aligned with the cardinal directions. We determine $\strain$
+by spatially and temporally differentiating estimates of transient
+displacements (i.e., displacements that deviate from the secular
+trend), which are constrained by GNSS data.
+
+% Defining u as a prior Gaussian process
 
@@
 transient displacements. Since transient displacements are not
-precisely known, we do not consider $\vec{u}$ to be a deterministic
-function. Instead, $\vec{u}$ is a stochastic process containing a
-distribution of functions that could potentially describe transient
-displacements. We let each component of $\vec{u}$ be a Gaussian
-process with zero mean and a generic covariance function
-$\Cov{u_i(p),u_i(p')} = C_{u_i}(p,p')$. Using a more concise notation,
-we write our prior on each component of $\vec{u}$ as $u_i \sim
-\mathcal{GP}\left(0,C_{u_i}\right)$. The function $C_{u_i}$ must be
-positive definite in order to be a valid covariance function. By
-definition, $C_{u_i}$ is a positive definite function if the matrix
-$[C_{u_i}(p,p')]_{(p,p') \in \points \times \points}$ is positive
-definite for any set of points $\points$ \citep[sec.
-2.5]{Cressie1992}. We assume that $C_{u_i}$ can be separated into
-spatial and temporal functions as
+precisely known, we cannot consider $\vec{u}$ to be a deterministic
+function. Instead, $\vec{u}$ is considered to be a stochastic process
+containing a distribution of functions that could potentially describe
+transient displacements. Specifically, we let each component of
+$\vec{u}$ be a Gaussian process. A Gaussian process is a stochastic
+process whose value at any collection of points can be described as a
+Gaussian random vector. Just as a Gaussian random vector is fully
+determined by a mean vector and a covariance matrix, a Gaussian
+process is fully determined by a mean function and a covariance
+function. For example, Brownian motion, $B(t)$, is a well known
+Gaussian process in $\mathbb{R}^1$ which has the mean function
+$\E{B(t)} = 0$ and the covariance function $\Cov{B(t),B(t')} =
+\min(t,t')$, where $t,t' \ge 0$. We let each component of $\vec{u}$
+have zero mean and a generic covariance function $\Cov{u_i(p),u_i(p')}
+= C_{u_i}(p,p')$. Using a more concise notation, we write our prior on
+each component of $\vec{u}$ as $u_i \sim
+\mathcal{GP}\left(0,C_{u_i}\right)$.
+
+% Constraints on the Covariance function
+
+The function $C_{u_i}$ must be positive definite in order to be a
+valid covariance function. By definition, $C_{u_i}$ is a positive
+definite function if the matrix $[C_{u_i}(p,p')]_{(p,p') \in \points
+\times \points}$ is positive definite for any set of points $\points$
+\citep[sec. 2.5]{Cressie1992}. We assume that $C_{u_i}$ can be
+separated into spatial and temporal functions as
 \begin{equation}\label{eq:TransientCovariance}
@@
 
-To give an example of $u_i$, we can use a squared exponential function
-for $X_i$ and $T_i$,
+We provide an example to give our prior on $\vec{u}$ a more tangible
+meaning. we can use a squared exponential function for $X_i$ and
+$T_i$,
 \begin{equation}\label{eq:ExampleCov}

### 2017-10-23 RevisedSubmission/hines_hetland_2017_gji.tex
added prior demo

@@
 strain rates. We consider transient strain rates to be any deviation
-from secular strain rates. We denote transient strain rates as
+from the secular strain rates. We denote transient strain rates as
 $\strain(p)$, where $p$ represents the ordered pair $(\pos,t)$, $\pos
@@
 by spatially and temporally differentiating estimates of transient
-displacements (i.e., displacements that deviate from the secular
-trend), which are constrained by GNSS data.
+displacements, which are constrained by GNSS data.
 
@@
 
+% Example of the prior
+
 We provide an example to give our prior on $\vec{u}$ a more tangible
@@
 realizations of $u_i$ to have a length-scale, time-scale, and
-amplitude that resembles what we expect from true signal. A few
-realizations for $u_i$ are shown in Figure XXX. While the squared
-exponential function is a commonly used covariance function for GPR,
-it is not appropriate for every application. The appropriate choice
-for $X_i$ and $T_i$ may vary depending on the geophysical signal we
-are trying to describe. To keep this section sufficiently general, we
-hold off on specifying $X_i$ and $T_i$ until Section
-\ref{sec:SignalModel}, where we estimate transient strain from slow
-slip events in the Pacific Northwest.
+amplitude that resembles what we expect for the true transient
+displacements. A few realizations for $u_i$ are shown in Figure XXX.
+While the squared exponential function is a commonly used covariance
+function for GPR, it is not appropriate for every application. The
+appropriate choice for $X_i$ and $T_i$ may vary depending on the
+geophysical signal we are trying to describe. To keep this section
+sufficiently general, we hold off on specifying $X_i$ and $T_i$ until
+Section \ref{sec:SignalModel}, where we estimate transient strain from
+slow slip events in the Pacific Northwest.
 
@@
 \begin{align}\label{eq:StrainMean}
-\mu_{\strain_\e\e}(p) &= \frac{\partial^2 \mu_{\post_\e}(p)}{\partial t \, \partial x_\e} \\
-\mu_{\strain_\n\n}(p) &= \frac{\partial^2 \mu_{\post_\n}(p)}{\partial t \, \partial x_\n} \\
-\mu_{\strain_\e\n}(p) &= \frac{1}{2}\frac{\partial}{\partial t}\left(
+\mu_{\strain_{\e\e}}(p) &= \frac{\partial^2 \mu_{\post_\e}(p)}{\partial t \, \partial x_\e} \\
+\mu_{\strain_{\n\n}}(p) &= \frac{\partial^2 \mu_{\post_\n}(p)}{\partial t \, \partial x_\n} \\
+\mu_{\strain_{\e\n}}(p) &= \frac{1}{2}\frac{\partial}{\partial t}\left(
                                     \frac{\partial \mu_{\post_\e}(p)}{\partial x_\n} + 
@@
 \begin{align}\label{eq:StrainCov}
-C_{\strain_\e\e}(p,p') &= \frac{\partial^4 C_{\post_\e}(p,p')}{\partial t \, \partial t' \, \partial x_\e \, \partial x'_\e} \\
-C_{\strain_\n\n}(p,p') &= \frac{\partial^4 C_{\post_\n}(p,p')}{\partial t \, \partial t' \, \partial x_\n \, \partial x'_\n} \\
-C_{\strain_\e\n}(p,p') &= \frac{1}{4} \frac{\partial^2}{\partial t \, \partial t'}\left(
+C_{\strain_{\e\e}}(p,p') &= \frac{\partial^4 C_{\post_\e}(p,p')}{\partial t \, \partial t' \, \partial x_\e \, \partial x'_\e} \\
+C_{\strain_{\n\n}}(p,p') &= \frac{\partial^4 C_{\post_\n}(p,p')}{\partial t \, \partial t' \, \partial x_\n \, \partial x'_\n} \\
+C_{\strain_{\e\n}}(p,p') &= \frac{1}{4} \frac{\partial^2}{\partial t \, \partial t'}\left(
                                       \frac{\partial^2 C_{\post_\e}(p,p')}{\partial x_\n \, \partial x'_\n} + 
@@
 \mu_{||\strain||_\mathrm{F}} &\approx ||\strain||_F \big|_{\strain = \mu_{\strain}}\nonumber \\
-                             &= \left(\mu_{\strain_\e\e}^2 + 
-                                      \mu_{\strain_\n\n}^2 + 
-                                      2\mu_{\strain_\e\n}^2 \right)^\frac{1}{2},
+                             &= \left(\mu_{\strain_{\e\e}}^2 + 
+                                      \mu_{\strain_{\n\n}}^2 + 
+                                      2\mu_{\strain_{\e\n}}^2 \right)^\frac{1}{2},
 \end{align}
@@
 \sigma_{||\strain||_\mathrm{F}} \approx
-\left( \left(\left. \frac{\partial ||\strain||_\mathrm{F}}{\partial \strain_\e\e}\right|_{\strain = \mu_{\strain}} \right)^2 
-       \sigma^2_{\strain_\e\e} +
-       \left(\left. \frac{\partial ||\strain||_\mathrm{F}}{\partial \strain_\n\n}\right|_{\strain = \mu_{\strain}} \right)^2
-       \sigma^2_{\strain_\n\n} +
-       \left(\left. \frac{\partial ||\strain||_\mathrm{F}}{\partial \strain_\e\n}\right|_{\strain = \mu_{\strain}} \right)^2 
-       \sigma^2_{\strain_\e\n} \right)^\frac{1}{2},
+\left( \left(\left. \frac{\partial ||\strain||_\mathrm{F}}{\partial \strain_{\e\e}}\right|_{\strain = \mu_{\strain}} \right)^2 
+       \sigma^2_{\strain_{\e\e}} +
+       \left(\left. \frac{\partial ||\strain||_\mathrm{F}}{\partial \strain_{\n\n}}\right|_{\strain = \mu_{\strain}} \right)^2
+       \sigma^2_{\strain_{\n\n}} +
+       \left(\left. \frac{\partial ||\strain||_\mathrm{F}}{\partial \strain_{\e\n}}\right|_{\strain = \mu_{\strain}} \right)^2 
+       \sigma^2_{\strain_{\e\n}} \right)^\frac{1}{2},
 \end{equation}
@@
 \mathrm{SNR}(p) &= \frac{\mu_{||\strain||_\mathrm{F}}(p)}{\sigma_{||\strain||_\mathrm{F}}(p)} \\
-                &= \frac{\mu_{\strain_\e\e}(p)^2 +
-                         \mu_{\strain_\n\n}(p)^2 +
-                         2\mu_{\strain_\e\n}(p)^2}
-                        {\big(\sigma^2_{\strain_\e\e}(p)\mu_{\strain_\e\e}(p)^2 + 
-                              \sigma^2_{\strain_\n\n}(p)\mu_{\strain_\n\n}(p)^2 + 
-                              4\sigma^2_{\strain_\e\n}(p)\mu_{\strain_\e\n}(p)^2
+                &= \frac{\mu_{\strain_{\e\e}}(p)^2 +
+                         \mu_{\strain_{\n\n}}(p)^2 +
+                         2\mu_{\strain_{\e\n}}(p)^2}
+                        {\big(\sigma^2_{\strain_{\e\e}}(p)\mu_{\strain_{\e\e}}(p)^2 + 
+                              \sigma^2_{\strain_{\n\n}}(p)\mu_{\strain_{\n\n}}(p)^2 + 
+                              4\sigma^2_{\strain_{\e\n}}(p)\mu_{\strain_{\e\n}}(p)^2
                          \big)^{\frac{1}{2}}}.

### 2017-10-23 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 $\partial^2 T_i(t,t') / \partial t \partial t'$ exist. This ensures
-that $\vec{u}$ is spatially and temporally differentiable, allowing us
-to compute $\strain$ (See \citet[sec. 2.2]{Adler1981} or \citet[sec.
-10A]{Papoulis1991} for a definition of stochastic differentiation and
-the conditions for differentiability).
+that $u_i$ is spatially and temporally differentiable, allowing us to
+compute transient strain rates (See \citet[sec. 2.2]{Adler1981} or
+\citet[sec. 10A]{Papoulis1991} for a definition of stochastic
+differentiation and the conditions for differentiability).
 
@@
 
-We provide an example to give our prior on $\vec{u}$ a more tangible
-meaning. we can use a squared exponential function for $X_i$ and
-$T_i$,
+We provide an example to give a more tangible meaning to the
+covariance function for $u_i$. we can use a squared exponential
+function for $X_i$ and $T_i$,
 \begin{equation}\label{eq:ExampleCov}
@@
 \end{equation}
+
 which satisfies our requirements for positive definiteness and
@@
 realizations of $u_i$ to have a length-scale, time-scale, and
-amplitude that resembles what we expect for the true transient
-displacements. A few realizations for $u_i$ are shown in Figure XXX.
-While the squared exponential function is a commonly used covariance
-function for GPR, it is not appropriate for every application. The
-appropriate choice for $X_i$ and $T_i$ may vary depending on the
-geophysical signal we are trying to describe. To keep this section
-sufficiently general, we hold off on specifying $X_i$ and $T_i$ until
-Section \ref{sec:SignalModel}, where we estimate transient strain from
-slow slip events in the Pacific Northwest.
+amplitude that resemble what we expect for the true transient
+displacements. The covariance function for $u_i$ and one realizations
+for $u_i$ are shown in Figure \ref{fig:PriorDemo} using the parameters
+$\ell = 100$ km, $\tau = 10$ days, and $\phi = 1$ mm. While the
+squared exponential function is a commonly used covariance function
+for GPR, it is not appropriate for every application. The appropriate
+choice for $X_i$ and $T_i$ may vary depending on the geophysical
+signal we are trying to describe. To keep this section sufficiently
+general, we hold off on specifying $X_i$ and $T_i$ until Section
+\ref{sec:SignalModel}, where we estimate transient strain from slow
+slip events in the Pacific Northwest.
+
+\begin{figure*}
+\includegraphics{figures/prior_demo/prior-demo.pdf}
+\caption{ 
+Panel A) The covariance function $C_{u_i}$ using a squared exponential
+to describe the spatial and temporal covariance (eq.
+\ref{eq:ExampleCov}). Panel B) A single realization of $u_i$ using the
+covariance function from Panel A. The realization is really a function
+of three variables, $t$, $x_\mathrm{e}$, and $x_\mathrm{n}$, although
+we only show its dependence on $t$ and one of the spatial dimensions.
+}    
+\label{fig:PriorDemo}
+\end{figure*}
 

### 2017-10-23 RevisedSubmission/figures/outliers_new/make_plot.py

@@ -41,7 +41,7 @@ us = 1000*data['east_std_dev'][:,idx]
 # toss out times that are out of range or have missing data
 tidx = (t >= xmin) & (t <= xmax) & np.isfinite(u)
 
-fig,axs = plt.subplots(2,1,figsize=(7,4.5),sharex=True)
+fig,axs = plt.subplots(2,1,figsize=(7.0,4.5),sharex=True)
 pygeons.plot.plot._setup_ts_ax(axs)
 
 axs[0].set_xticks(xticks)

### 2017-10-23 RevisedSubmission/figures/prior_demo/make-plot.py

@@ -20,7 +20,7 @@ gp = rbf.gauss.GaussianProcess(
 #####################################################################
 Nx = 20
 Nt = 20
-figsize = (7.0,3.25)
+figsize = (7.0,3.0)
 
 x = np.linspace(0.0,300.0,Nx)
 t = np.linspace(0.0,30.0,Nt)
@@ -69,7 +69,7 @@ p = ax.plot_surface(xg,tg,sample,cmap='viridis',edgecolor='k',linewidth=0.5)
 # set labels
 ax.set_ylabel(r'$\mathregular{t}$ [days]',fontsize=10)
 ax.set_xlabel(r'$\mathregular{x_j}$ [km]',fontsize=10)
-ax.set_zlabel('realization [mm]',fontsize=10)
+ax.set_zlabel('displacement [mm]',fontsize=10)
 ax.tick_params(pad=1.0)
 ax.tick_params(labelsize=10)
 ax.text2D(-0.05, 0.9, "B", fontsize=16, transform=ax.transAxes)

### 2017-10-23 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 
-We provide an example to give a more tangible meaning to the
-covariance function for $u_i$. we can use a squared exponential
-function for $X_i$ and $T_i$,
+We provide an example to give the covariance function for $u_i$ a more
+tangible meaning. we can use a squared exponential function for $X_i$
+and $T_i$,
+
 \begin{equation}\label{eq:ExampleCov}
@@
 amplitude that resemble what we expect for the true transient
-displacements. The covariance function for $u_i$ and one realizations
+displacements. In Figure \ref{fig:PriorDemo}A, we show $C_{u_i}$ using
+the squared exponential function $X_i$ and $T_i$ and the parameters
+$\ell = 100$ km, $\tau = 10$ days, and $\phi = 1$ mm. A single
+realization of $u_i$ corresponsing to that choice of covariance
+function is shown in Figure \ref{fig:PriorDemo}B (See \citet[sec.
+A3]{Rasmussen2006} for details on drawing realization from Gaussian
+processes).
+
+The covariance function for $u_i$ and one realizations
 for $u_i$ are shown in Figure \ref{fig:PriorDemo} using the parameters
-$\ell = 100$ km, $\tau = 10$ days, and $\phi = 1$ mm. While the
+While the
 squared exponential function is a commonly used covariance function
@@
 Panel A) The covariance function $C_{u_i}$ using a squared exponential
-to describe the spatial and temporal covariance (eq.
+to describe the spatial and temporal covariance of $u_i$ (eq.
 \ref{eq:ExampleCov}). Panel B) A single realization of $u_i$ using the

### 2017-10-23 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 
-We provide an example to give the covariance function for $u_i$ a more
-tangible meaning. we can use a squared exponential function for $X_i$
-and $T_i$,
+We provide an example to give the prior on transient displacements a
+more tangible meaning. we can use a squared exponential function for
+$X_i$ and $T_i$,
 
@@
 \caption{ 
-Panel A) The covariance function $C_{u_i}$ using a squared exponential
-to describe the spatial and temporal covariance of $u_i$ (eq.
-\ref{eq:ExampleCov}). Panel B) A single realization of $u_i$ using the
-covariance function from Panel A. The realization is really a function
-of three variables, $t$, $x_\mathrm{e}$, and $x_\mathrm{n}$, although
-we only show its dependence on $t$ and one of the spatial dimensions.
+Panel A) The covariance between $u_i(p)$ and $u_i(p')$,
+$C_{u_i}(p,p')$, as a function of the spatial and temporal distance
+between $p$ and $p'$. The covariance function $C_{u_i}$ is composed of
+squared exponentials to describe the spatial and temporal covariance
+of $u_i$ (eq. \ref{eq:ExampleCov}). Panel B) A single realization of
+$u_i$ using the covariance function from Panel A. The realization is
+really a function of three variables, $t$, $x_\mathrm{e}$, and
+$x_\mathrm{n}$, although we only show its dependence on $t$ and one of
+the spatial dimensions.
 }    

### 2017-10-23 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 $X_i$ and $T_i$,
-
 \begin{equation}\label{eq:ExampleCov}
@@
 \end{equation}
-
 which satisfies our requirements for positive definiteness and
@@
 displacements. In Figure \ref{fig:PriorDemo}A, we show $C_{u_i}$ using
-the squared exponential function $X_i$ and $T_i$ and the parameters
-$\ell = 100$ km, $\tau = 10$ days, and $\phi = 1$ mm. A single
-realization of $u_i$ corresponsing to that choice of covariance
+the squared exponential function for $X_i$ and $T_i$ and the
+parameters $\ell = 100$ km, $\tau = 10$ days, and $\phi = 1$ mm. A
+single realization of $u_i$ corresponsing to that choice of covariance
 function is shown in Figure \ref{fig:PriorDemo}B (See \citet[sec.
 A3]{Rasmussen2006} for details on drawing realization from Gaussian
-processes).
-
-The covariance function for $u_i$ and one realizations
-for $u_i$ are shown in Figure \ref{fig:PriorDemo} using the parameters
-While the
-squared exponential function is a commonly used covariance function
-for GPR, it is not appropriate for every application. The appropriate
-choice for $X_i$ and $T_i$ may vary depending on the geophysical
-signal we are trying to describe. To keep this section sufficiently
-general, we hold off on specifying $X_i$ and $T_i$ until Section
-\ref{sec:SignalModel}, where we estimate transient strain from slow
-slip events in the Pacific Northwest.
+processes). While the squared exponential function is a commonly used
+covariance function for GPR, it is not appropriate for every
+application. The appropriate choice for $X_i$ and $T_i$ may vary
+depending on the geophysical signal we are trying to describe. To keep
+this section sufficiently general, we hold off on specifying $X_i$ and
+$T_i$ until Section \ref{sec:SignalModel}, where we estimate transient
+strain from slow slip events in the Pacific Northwest.
 

### 2017-10-24 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 by spatially and temporally differentiating estimates of transient
-displacements, which are constrained by GNSS data.
+displacements, which we constrain with GNSS data.
 
@@
 of $u_i$ (eq. \ref{eq:ExampleCov}). Panel B) A single realization of
-$u_i$ using the covariance function from Panel A. The realization is
-really a function of three variables, $t$, $x_\mathrm{e}$, and
+$u_i$ corresponding to the covariance function from Panel A. The
+realization is a function of three variables, $t$, $x_\mathrm{e}$, and
 $x_\mathrm{n}$, although we only show its dependence on $t$ and one of
@@
 \end{figure*}
+
+% define the components of a datum
 
@@
 $\pos^{(j)}$, and the sinusoids describe seasonal deformation (using
-units of years for $t^{(k)}$). We then consider the column vector of
-$n$ GNSS displacement observations, $\data^*_i$, where the subscript
-indicates that we are still only considering component $i$ of
-displacements. The observations are made at $m$ stations, and the
-times and positions for each observation are described by the set
-$\points$. $\data^*_i$ can be considered a realization of the random
-vector $\data_i$, which is formed by evaluating eq. (\ref{eq:Data}) at
-each point in $\points$. To write out $\data_i$ explicitly, we let
-$\G$ be an $n \times 6m$ matrix consisting of the basis functions from
-eq. (\ref{eq:Data}) (i.e., the linear trends and sinusoids for each
-station) evaluated at each point in $\points$. The coefficients
-corresponding to each basis function are collected into the column
-vector $\mitbf{m}_i$. The noise for all the observations are described
-by the column vector $\mitbf{\eta}_i$. We can then write $\data_i$ as
+units of years for $t^{(k)}$).
+
+% define the data vector
+
+We then consider the column vector of $n$ GNSS displacement
+observations, $\data^*_i$, where the subscript indicates that we are
+still only considering component $i$ of displacements. The
+observations are made at $m$ stations, and the times and positions for
+each observation are described by the set $\points$. $\data^*_i$ can
+be considered a realization of the random vector $\data_i$, which is
+formed by evaluating eq. (\ref{eq:Data}) at each point in $\points$.
+To write out $\data_i$ explicitly, we let $\G$ be an $n \times 6m$
+matrix consisting of the basis functions from eq. (\ref{eq:Data})
+(i.e., the linear trends and sinusoids for each station) evaluated at
+each point in $\points$. The coefficients corresponding to each basis
+function are collected into the column vector $\mitbf{m}_i$. The noise
+for all the observations are described by the column vector
+$\mitbf{\eta}_i$. We can then write $\data_i$ as
 \begin{equation}\label{eq:DataVec}
@@
 $[u_i(p)]_{p \in \points}$.
+
+% establishing the distribution of the parameters in d and d itself
 
@@
 $[C_{u_i}(p,p')]_{(p,p') \in \points \times \points}$.
+
+% Combining the data and prior to form posterior transient
+% displacements
 
@@
 
+% Noting that we ignore covariance between components
+
 It should be noted that we have ignored any covariances between the
@@
 admittedly an arbitrary assumption.
+
+% converting posterior transient displacements to posterior transient
+% strain
 
@@
 
+% Non-technical description of outlier detection methods and our
+% outlier detection method
+
 In deriving our formulation for transient strain rates, we have
@@
 
+% Acknowledging that jumps are not detected
+
 It should be noted that our algorithm does not identify jumps in GNSS
@@
 
-We use our method to estimate transient strain rates in the Pacific
-Northwest, and we are specifically interested in identifying strain
-resulting from SSEs \citep[e.g.,][]{Dragert2001}. In the Pacific
-Northwest, SSEs can be detected by monitoring for associated seismic
-tremor \citep{Rogers2003}, which is actively being done by the Pacific
-Northwest Seismic Network \citep{Wech2010}. We can thus compare the
-tremor records to the transient strain rates estimated with GPR to
-verify that we are indeed identifying strain from SSEs.
-
-We use the daily displacement solutions generated by the Geodesy
-Advancing Geosciences and EarthScope (GAGE) facility for continuous
-GNSS stations \citep{Herring2016}. This data is publicly available and
-can be found at www.unavco.org. We limit the dataset to the stations
-and time ranges which are pertinent to the seven most recent SSEs in
-the Puget Sound region. The earliest SSE considered in this study
-began in August 2010, and the most recent SSE began in February 2017.
-We use these most recent SSEs because the station coverage is
-sufficiently dense for us to use maximum likelihood methods to
-constrain prior models.  The positions of GNSS stations used to
-estimate transient strain rates are shown in Figure \ref{fig:Context}.
+% Overview of what we are doing in this section
+
+In this section we estimate transient strain rates in the Pacific
+Northwest, and we are specifically interested in identifying transient
+strain resulting from SSEs \citep[e.g.,][]{Dragert2001}. Before
+estimating transient strain rates, we establish a noise model for GNSS
+stations in this region, and we establish a prior Gaussian process to
+describe displacements from SSEs. SSEs in the Pacific Northwest can be
+detected by monitoring for associated seismic tremor
+\citep{Rogers2003}, which is actively being done by the Pacific
+Northwest Seismic Network \citep{Wech2010}. We can then compare the
+tremor records to our estimated transient strain rates to verify that
+we are indeed identifying strain from SSEs.
+
+% Dataset used here
+
+We use the daily displacement solutions for continuous GNSS stations
+generated by the Geodesy Advancing Geosciences and EarthScope (GAGE)
+Facility \citep{Herring2016}. We limit the dataset to the stations and
+times that are pertinent to the seven most recent SSEs in the Puget
+Sound region. The earliest SSE considered in this study began in
+August 2010, and the most recent SSE began in February 2017. The
+positions of GNSS stations used to estimate transient strain rates are
+shown in Figure \ref{fig:Context}.
 
@@
 
-Before we determine the transient strain rates, we must establish a
-prior for the transient displacements, $u$, and the noise, $\eta$. In
-this section we discuss our choice for the noise covariance function
-$C_\eta$. There have been numerous studies on temporally correlated
-noise in GNSS data \citep[e.g.,][]{Zhang1997, Mao1999, Williams2004,
-Langbein2008}. In these studies, temporally correlated noise was
-described with some combination of Brownian motion, a first-order
-Gauss-Markov (FOGM) process, and/or flicker noise. There is some
-physical justification for using Brownian motion as a noise model
-because it accurately describes the power spectrum of motion resulting
-from instability in geodetic monuments \citep[e.g.,][]{Wyatt1982,
-Wyatt1989}. Here we describe the time dependence of $\eta$ as a FOGM
-process and consider $\eta$ to be spatially uncorrelated. A FOGM
-process is a solution to the stochastic differential equation
+% landing on the FOGM noise model
+
+In this section we establish a noise model for the GNSS data. The
+noise in GNSS data, specifically the temporally correlated noise, has
+been thoroughly studied over the past two decades
+\citep[e.g.,][]{Zhang1997, Mao1999, Williams2004, Langbein2008}. In
+these studies, temporally correlated noise is described with some
+combination of Brownian motion, a first-order Gauss-Markov (FOGM)
+process, and/or flicker noise. There is some physical justification
+for using Brownian motion as a noise model because it accurately
+describes the power spectrum of motion resulting from instability in
+geodetic monuments \citep[e.g.,][]{Wyatt1982, Wyatt1989}. In
+particular, the power spectrums for Brownian motion and monument
+motion are both proportional to $f^{-2}$, where $f$ is frequency.
+Brownian motion (which is also known as a Weiner process), by
+definition, contains a reference time at which the process begins.
+Since there is no notion of when noise ``begins" in GNSS data, we do
+not find Brownian motion to be an appropriate noise model. On the
+other hand, a FOGM process also has a power spectrum that is
+proportional to $f^{-2}$, above some corner frequency, and it has no
+reference time (i.e., a FOGM process is stationary). In this study, we
+describe the temporal covariance of GNSS noise with a FOGM process. We
+do not explore spatial covariance in GNSS noise and simply assume that
+it is spatially uncorrelated.
+
+Here we
+describe the time dependence of $\eta$ as a FOGM process and consider
+$\eta$ to be spatially uncorrelated. A FOGM process is a solution to
+the stochastic differential equation
+
 \begin{equation}\label{eq:FOGMdef}

### 2017-10-24 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 $\strain(p)$, where $p$ represents the ordered pair $(\pos,t)$, $\pos
-= (x_\e,x_\n)$ are spatial coordinates, and $t$ is time. For
-simplicity, we assume that the study region is sufficiently small that
-$\pos$ can be considered a point in a 2-D Cartesian map projection
-that is aligned with the cardinal directions. We determine $\strain$
-by spatially and temporally differentiating estimates of transient
-displacements, which we constrain with GNSS data.
+= (x_\e,x_\n)$ are spatial coordinates, and $t$ is time. The
+subscripts ``e" and ``n" indicate the east and north component, and we
+assume that the study region is sufficiently small that $\pos$ can be
+considered a point in a 2-D Cartesian map projection that is aligned
+with the cardinal directions. We determine $\strain$ by spatially and
+temporally differentiating estimates of transient displacements, which
+we constrain with GNSS data.
 
@@
 $\vec{u}$ be a Gaussian process. A Gaussian process is a stochastic
-process whose value at any collection of points can be described as a
-Gaussian random vector. Just as a Gaussian random vector is fully
-determined by a mean vector and a covariance matrix, a Gaussian
-process is fully determined by a mean function and a covariance
-function. For example, Brownian motion, $B(t)$, is a well known
-Gaussian process in $\mathbb{R}^1$ which has the mean function
-$\E{B(t)} = 0$ and the covariance function $\Cov{B(t),B(t')} =
-\min(t,t')$, where $t,t' \ge 0$. We let each component of $\vec{u}$
-have zero mean and a generic covariance function $\Cov{u_i(p),u_i(p')}
-= C_{u_i}(p,p')$. Using a more concise notation, we write our prior on
-each component of $\vec{u}$ as $u_i \sim
-\mathcal{GP}\left(0,C_{u_i}\right)$.
+process whose value at any collection of points can be described with
+a multivariate normal distribution. That is to say, the random vector
+$[u_i(p)]_{p \in \points}$ has a multivariate normal distribution for
+any collection of points $\points$. A realization of the random vector
+$[u_i(p)]_{p \in \points}$ can be interpretted as a realization of
+$u_i$ (i.e., a sample function) that is evaluated at $\points$. Just
+as a multivariate normal distribution is fully determined by a mean
+vector and a covariance matrix, a Gaussian process is fully determined
+by a mean function and a covariance function. For example, Brownian
+motion, $B(t)$, is a well known Gaussian process in $\mathbb{R}^1$
+which has the mean function $\E{B(t)} = 0$ and the covariance function
+$\Cov{B(t),B(t')} = \min(t,t')$, where $t,t' \ge 0$. We let each
+component of $\vec{u}$ have zero mean and a generic covariance
+function $\Cov{u_i(p),u_i(p')} = C_{u_i}(p,p')$. Using a more concise
+notation, we write our prior on each component of $\vec{u}$ as $u_i
+\sim \mathcal{GP}\left(0,C_{u_i}\right)$.
 
@@
 differentiability. The parameters $\ell$ and $\tau$ control the
-length-scale and time-scale, respectively, of realizations (i.e.,
-sample functions) of $u_i$. The parameter $\phi$, which we have
-arbitrarily chosen to incorporate into $T_i$ rather than $X_i$,
-controls the amplitude of realizations of $u_i$. Ideally, we want
-realizations of $u_i$ to have a length-scale, time-scale, and
-amplitude that resemble what we expect for the true transient
-displacements. In Figure \ref{fig:PriorDemo}A, we show $C_{u_i}$ using
-the squared exponential function for $X_i$ and $T_i$ and the
-parameters $\ell = 100$ km, $\tau = 10$ days, and $\phi = 1$ mm. A
-single realization of $u_i$ corresponsing to that choice of covariance
-function is shown in Figure \ref{fig:PriorDemo}B (See \citet[sec.
-A3]{Rasmussen2006} for details on drawing realization from Gaussian
-processes). While the squared exponential function is a commonly used
-covariance function for GPR, it is not appropriate for every
-application. The appropriate choice for $X_i$ and $T_i$ may vary
+length-scale and time-scale, respectively, of realizations of $u_i$.
+The parameter $\phi$, which we have arbitrarily chosen to incorporate
+into $T_i$ rather than $X_i$, controls the amplitude of realizations
+of $u_i$. Ideally, we want realizations of $u_i$ to have a
+length-scale, time-scale, and amplitude that resemble what we expect
+for the true transient displacements. In Figure \ref{fig:PriorDemo}A,
+we show $C_{u_i}$ using the squared exponential function for $X_i$ and
+$T_i$ and the parameters $\ell = 100$ km, $\tau = 10$ days, and $\phi
+= 1$ mm. A single realization of $u_i$ corresponsing to that choice of
+covariance function is shown in Figure \ref{fig:PriorDemo}B (See
+\citet[sec. A3]{Rasmussen2006} for details on drawing realization from
+Gaussian processes). While the squared exponential function is a
+commonly used covariance function for GPR, it is not appropriate for
+every application. The appropriate choice for $X_i$ and $T_i$ may vary
 depending on the geophysical signal we are trying to describe. To keep

### 2017-10-25 RevisedSubmission/hines_hetland_2017_gji.tex
added more

@@
 
-% landing on the FOGM noise model
-
-In this section we establish a noise model for the GNSS data. The
-noise in GNSS data, specifically the temporally correlated noise, has
-been thoroughly studied over the past two decades
-\citep[e.g.,][]{Zhang1997, Mao1999, Williams2004, Langbein2008}. In
-these studies, temporally correlated noise is described with some
-combination of Brownian motion, a first-order Gauss-Markov (FOGM)
-process, and/or flicker noise. There is some physical justification
-for using Brownian motion as a noise model because it accurately
-describes the power spectrum of motion resulting from instability in
-geodetic monuments \citep[e.g.,][]{Wyatt1982, Wyatt1989}. In
-particular, the power spectrums for Brownian motion and monument
-motion are both proportional to $f^{-2}$, where $f$ is frequency.
-Brownian motion (which is also known as a Weiner process), by
-definition, contains a reference time at which the process begins.
-Since there is no notion of when noise ``begins" in GNSS data, we do
-not find Brownian motion to be an appropriate noise model. On the
-other hand, a FOGM process also has a power spectrum that is
-proportional to $f^{-2}$, above some corner frequency, and it has no
-reference time (i.e., a FOGM process is stationary). In this study, we
-describe the temporal covariance of GNSS noise with a FOGM process. We
-do not explore spatial covariance in GNSS noise and simply assume that
-it is spatially uncorrelated.
-
-Here we
-describe the time dependence of $\eta$ as a FOGM process and consider
-$\eta$ to be spatially uncorrelated. A FOGM process is a solution to
-the stochastic differential equation
-
-\begin{equation}\label{eq:FOGMdef}
-\dot{\eta}(t) + \alpha \eta(t) = \beta w(t),
-\end{equation}
-where $w(t)$ is white noise with unit variance. The FOGM process
-degenerates to the commonly used Brownian motion noise model under the
-condition that $\alpha=0$ and $\eta(0) = 0$. Our spatially
-uncorrelated noise model that satisfies eq. (\ref{eq:FOGMdef}) is a
-Gaussian process with zero mean and the covariance function
+% Defining the components of the noise vector
+
+We consider the noise vector, $\mitbf{\eta}_i$, to be composed of a
+temporally correlated Gaussian process, $\mathcal{L}_i \sim
+\mathcal{GP}(0,C_{\mathcal{L}_i})$, and a vector of uncorrelated
+Gaussian noise, $\mitbf{w}_i$, so that
+\begin{equation}
+\mitbf{\eta}_i = \mathcal{L}_i(\points) + \mitbf{w}_i.
+\end{equation}
+The standard deviations for $\mitbf{w}_i$ are given by the
+uncertainties derived for the GNSS displacement solutions,
+$\mitbf{\sigma}_i$. The noise vector then has zero mean and the
+covariance matrix
+\begin{equation}
+\mitbf{C}_{\eta_i} = C_{\mathcal{L}_i}(\points,\points) +
+                     \mathrm{diag}(\mitbf{\sigma}_i^2).
+\end{equation}
+
+The temporally correlated noise in GNSS data has been thoroughly
+studied over the past two decades \citep[e.g.,][]{Zhang1997, Mao1999,
+Williams2004, Langbein2008}. In these studies, temporally correlated
+noise tends to be described with some combination of Brownian motion
+(also known as random walk noise or a Weiner process), a first-order
+Gauss-Markov (FOGM) process, and/or flicker noise. There is some
+physical justification for using Brownian motion as a noise model
+because it accurately describes the power spectrum of motion resulting
+from instability in geodetic monuments \citep[e.g.,][]{Wyatt1982,
+Wyatt1989}. In particular, the power spectrums for Brownian motion and
+monument motion are both proportional to $f^{-2}$, where $f$ is
+frequency. Brownian motion, by definition, contains a reference time
+at which the process begins. Since there is no notion of when noise
+``begins" in GNSS data, we do not find Brownian motion to be an
+appropriate noise model. On the other hand, a FOGM process also has a
+power spectrum that is proportional to $f^{-2}$, above some corner
+frequency, and it has no reference time (i.e., a FOGM process is
+stationary) \citep{Langbein2004}. We then let $\mathcal{L}_i$ be a
+spatially uncorrelated FOGM process, which has the covariance function
 \begin{equation}\label{eq:FOGM}
-C_\eta\left((\pos,t),(\pos',t')\right) = 
-\begin{cases}
-\frac{\beta^2}{2\alpha}\exp\left(-\alpha|t - t'|\right), &\pos = \pos' \\
-0, &\pos \neq \pos'
-\end{cases}. 
-\end{equation}
-
-We constrain the hyperparameters for $\eta$, $\alpha$ and $\beta$,
-with a set of 38 continuous GNSS stations in the Pacific Northwest
-that are east of 121$^\circ$W.  These stations are sufficiently far
-from the subduction zone that they are unlikely to contain transient
-signal associated with SSEs.  We clean the data for these stations by
-removing jumps at times of equipment changes, and we remove outliers
-that have been detected with the algorithm described in Section
-\ref{sec:Outlier}. We then find $\alpha$ and $\beta$ for each station
-time series with the Restricted Maximum Likelihood (REML) method
-\cite[e.g.,][]{Harville1974,Cressie1992,Hines2017}. The REML method
-finds the hyperparameters, which we collectively refer to as
-$\mitbf{\theta}$, that maximize the likelihood function
+C_{\mathcal{L}_i}\left((\pos,t),(\pos',t')\right) = 
+\frac{\beta^2}{2\alpha}
+\exp\left(-\alpha|t - t'|\right)\delta_{\pos,\pos'}
+\end{equation}
+\citep[sec. B2.1]{Rasmussen2006}, where $\delta_{\pos,\pos'}$ is 1 if
+$\pos = \pos'$ and 0 otherwise.
+
+% How we constrain the parameters
+
+We constrain the parameters for $\mitbf{\eta}_i$, $\alpha$ and
+$\beta$, with a set of 38 continuous GNSS stations in the Pacific
+Northwest that are east of 121$^\circ$W.  These stations are
+sufficiently far from the subduction zone that they are unlikely to
+contain transient signal associated with SSEs. We find $\alpha$ and
+$\beta$ for each station time series with the Restricted Maximum
+Likelihood (REML) method \cite[e.g.,][]{Harville1974, Cressie1992}.
+The REML method finds the hyperparameters, which we collectively refer
+to as $\mitbf{\theta}$, that maximize the likelihood function
 \begin{equation}\label{eq:REML}
@@
 station. We are assuming $u(p)=0$ when estimating the noise
-hyperparameters for this section. The distribution of inferred
-$\alpha$ and $\beta$ are shown in Figure \ref{fig:NoiseParams}. The
-amplitude of FOGM noise, $\beta$, for the easting and northing
-components is notably low and are clustered around 0.5 mm/yr$^{0.5}$.
-The corresponding estimates of $\alpha$ tend to cluster around 0
-yr$^{-1}$, suggesting that noise can be described with Brownian
-motion. We also estimate hyperparameters for the vertical component of
-displacements, under the hope that vertical deformation gradients
-could reveal some geophysical signal. The amplitude of FOGM noise for
-the vertical component is relatively large with a median value of 13.5
-mm/yr$^{0.5}$.  The inferred values for $\alpha$ are also higher for
-the vertical component with a median value of 8.21 yr$^{-1}$. In
-Figure \ref{fig:NoiseSamples}, we use the median values of $\alpha$
-and $\beta$ to generate two random samples of FOGM noise for each
-component. The samples span five years and over these five years the
-easting and northing samples drift by about 1 mm. In the context of
-detecting SSEs, which produce several mm's of surface displacement on
-the time-scale of weeks, the estimated FOGM noise for the easting and
-northing component is negligible. In contrast, the estimated FOGM
-noise for the vertical component is larger than the signal we would
-expect from SSEs. We suspect that the higher amplitude for the FOGM
-noise in the vertical component is accommodating for deficiencies in
-our rather simple seasonal model. Based on this analysis, we
-henceforth ignore temporally correlated noise in the easting and
-northing component because of its low amplitude. We also do not use
-vertical displacements because of the presumably low signal-to-noise
-ratio.
+hyperparameters for this section.
+
+% results of estimated parameters
+
+The distribution of inferred $\alpha$ and $\beta$ are shown in Figure
+\ref{fig:NoiseParams}. The amplitude of FOGM noise, $\beta$, for the
+easting and northing components is notably low and are clustered
+around 0.5 mm/yr$^{0.5}$. The corresponding estimates of $\alpha$ tend
+to cluster around 0 yr$^{-1}$, suggesting that noise can be described
+with Brownian motion. We also estimate hyperparameters for the
+vertical component of displacements, under the hope that vertical
+deformation gradients could reveal some geophysical signal. The
+amplitude of FOGM noise for the vertical component is relatively large
+with a median value of 13.5 mm/yr$^{0.5}$.  The inferred values for
+$\alpha$ are also higher for the vertical component with a median
+value of 8.21 yr$^{-1}$. In Figure \ref{fig:NoiseSamples}, we use the
+median values of $\alpha$ and $\beta$ to generate two random samples
+of FOGM noise for each component. The samples span five years and over
+these five years the easting and northing samples drift by about 1 mm.
+In the context of detecting SSEs, which produce several mm's of
+surface displacement on the time-scale of weeks, the estimated FOGM
+noise for the easting and northing component is negligible. In
+contrast, the estimated FOGM noise for the vertical component is
+larger than the signal we would expect from SSEs. We suspect that the
+higher amplitude for the FOGM noise in the vertical component is
+accommodating for deficiencies in our rather simple seasonal model.
+Based on this analysis, we henceforth ignore temporally correlated
+noise in the easting and northing component because of its low
+amplitude. We also do not use vertical displacements because of the
+presumably low signal-to-noise ratio.
 
@@
 \end{figure*}
+
+% Note spatially correlated noise
 
@@
 We next determine appropriate hyperparameters for $X$ and each of the
-three candidate covariance functions for $T$. First, we clean the GNSS
-datasets by removing offsets at times of equipment changes and
-removing outliers with the method describe in Section
-\ref{sec:Outlier}. For the outlier detection algorithm, our prior
-model, $u$, is chosen to have a length-scale and time-scale that is
-able to approximately describe SSE displacements. We use the SE
-covariance function for $X$ with length-scale $\ell = 100$ km, and we
-use the Wendland covariance function for $T$, due to its computational
-efficiency, with time-scale $\tau = 0.1$ yr and $\phi = 1$ mm.  The
-outlier detection algorithm is particularly effective at removing
-outliers for stations at high elevation (Figure \ref{fig:Outliers}),
-which can be adversely affected by ice or snow during the winter
-\citep{Lisowski2008}. After cleaning the dataset, we divide it into
-seven subsets which are four months long and each centered on the time
-of a SSE. The times of the seven SSEs are determined with tremor
+three candidate covariance functions for $T$. we divide the dataset
+into seven subsets which are four months long and each centered on the
+time of a SSE. The times of the seven SSEs are determined with tremor
 records from \cite{Wech2010}. We use the REML method to find the
@@
 $\phi$, which results in large interquartile ranges.
-
-\begin{figure*}
-\includegraphics{figures/outliers/outliers.pdf}
-\caption{
-Detrended easting component of displacements at station SC03, which is
-located on Mount Olympus in Washington. The orange markers indicate
-outliers that were automatically detected using the algorithm from
-Section \ref{sec:Outlier}. The error bars show one standard deviation
-uncertainties. Note that outliers tend be observed in the winter,
-suggesting that they were caused by snow or ice.
-}   
-\label{fig:Outliers}
-\end{figure*}
 
@@
 Having established a noise model and a prior for transient
-displacements, we use the cleaned GNSS dataset to calculate transient
-strain rates in the Puget Sound region.  We calculate transient strain
-rates for each day from January 1, 2010 to May 15, 2017. The strain
-rates are estimates at a grid of points spanning the study area. In
-Figure \ref{fig:StrainMap} we show the transient strain rates on
-January 1, 2016, which coincides with the height of the winter
-2015-2016 SSE. We have included an animation showing the map view of
-strain rates through time as supplementary material. The strain rates
-shown in Figure \ref{fig:StrainMap} are generally similar to the
-strain rates during the other six SSEs considered in this study. The
-SSEs cause trench perpendicular compression in the Olympic Peninsula
-and extension east of Puget Sound. The strain transitions from
-compression to extension around the southern tip of Vancouver Island,
-which coincides with the location of thrust slip for SSEs in the Puget
-Sound region \citep[e.g.,][]{Dragert2001,Wech2009,Schmidt2010}. Thus,
-this pattern of strain is to be expected. During the period in between
+displacements, we use the GNSS dataset to calculate transient strain
+rates in the Puget Sound region. We calculate transient strain rates
+for each day from January 1, 2010 to May 15, 2017. The strain rates
+are estimates at a grid of points spanning the study area. In Figure
+\ref{fig:StrainMap} we show the transient strain rates on January 1,
+2016, which coincides with the height of the winter 2015-2016 SSE. We
+have included an animation showing the map view of strain rates
+through time as supplementary material. The strain rates shown in
+Figure \ref{fig:StrainMap} are generally similar to the strain rates
+during the other six SSEs considered in this study. The SSEs cause
+trench perpendicular compression in the Olympic Peninsula and
+extension east of Puget Sound. The strain transitions from compression
+to extension around the southern tip of Vancouver Island, which
+coincides with the location of thrust slip for SSEs in the Puget Sound
+region \citep[e.g.,][]{Dragert2001,Wech2009,Schmidt2010}. Thus, this
+pattern of strain is to be expected. During the period in between
 SSEs, secular strain rates indicate trench perpendicular compression

### 2017-10-26 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 \end{equation}
-The standard deviations for $\mitbf{w}_i$ are given by the
+The standard deviations for $\mitbf{w}_i$ are taken to be the
 uncertainties derived for the GNSS displacement solutions,
@@
 \end{equation}
+
+% Justifying our temporal noise model
 
@@
 frequency, and it has no reference time (i.e., a FOGM process is
-stationary) \citep{Langbein2004}. We then let $\mathcal{L}_i$ be a
-spatially uncorrelated FOGM process, which has the covariance function
+stationary). We then let $\mathcal{L}_i$ be a spatially uncorrelated
+FOGM process, which has the covariance function
 \begin{equation}\label{eq:FOGM}
@@
 
-We constrain the parameters for $\mitbf{\eta}_i$, $\alpha$ and
-$\beta$, with a set of 38 continuous GNSS stations in the Pacific
+By choosing a FOGM process for $\mathcal{L}_i$, we have introduced two
+parameters that need to be constrained, $\alpha$ and $\beta$. We
+constrain these parameters, which we collectively refer to as
+$\mitbf{\theta}$, with the Restricted Maximum Likelihood (REML) method
+\citep{Harville1974}. The REML method is conceptually similar to the
+Maximum Likelihood Estimation (MLE) method from \citet{Langbein1997},
+where we pick $\mitbf{\theta}$ to maximize the likelihood of drawing
+the observed data, $\mitbf{d}_i^*$, from the random vector
+$\mitbf{d}_i$. However, unlike the MLE method, the REML method
+produces unbiased estimates of $\mitbf{\theta}$ \citep[sec.
+2.6]{Cressie1992}. We use the REML method to independently estimate
+$\mitbf{\theta}$ at 38 continuous GNSS stations in the Pacific
 Northwest that are east of 121$^\circ$W. These stations are
 sufficiently far from the subduction zone that they are unlikely to
-contain transient signal associated with SSEs. We find $\alpha$ and
-$\beta$ for each station time series with the Restricted Maximum
-Likelihood (REML) method \cite[e.g.,][]{Harville1974, Cressie1992}.
-The REML method finds the hyperparameters, which we collectively refer
-to as $\mitbf{\theta}$, that maximize the likelihood function
+contain transient signal associated with SSEs. We assume that the
+noise at these inland stations is representative of the noise at all
+the stations considered in this study, which is probably a poor
+assumption since the inland stations are subject to distinctly
+different climatic conditions. Nonetheless, we find the
+$\mitbf{\theta}$ for each inland station and for each displacement
+component that maximizes the REML likelihood function
 \begin{equation}\label{eq:REML}
 \mathcal{L}(\mitbf{\theta}) = \left(\frac{\left|\G^T\G\right|}{(2\pi)^{n-6m} 
-                                    \left| \mitbf{\Sigma}(\mitbf{\theta}) \right| 
-                                    \left| \G^T\mitbf{\Sigma}(\mitbf{\theta})^{-1}\G \right|}\right)^{\frac{1}{2}} 
-                              e^{-\tfrac{1}{2}\data_*^T\mitbf{K}(\mitbf{\theta})\data_*},
-\end{equation}
-where
+                                    \left| \mitbf{C}_{\eta_i}(\mitbf{\theta}) \right| 
+                                    \left| \G^T\mitbf{C}_{\eta_i}(\mitbf{\theta})^{-1}\G \right|}\right)^{\frac{1}{2}} 
+                              e^{-\tfrac{1}{2}\data_i^{*T}\mitbf{K}(\mitbf{\theta})\data_i^*},
+\end{equation}
+with respect to $\mitbf{\theta}$, where
 \begin{equation}
-\mitbf{K}(\mitbf{\theta}) = \mitbf{\Sigma}(\mitbf{\theta})^{-1} - 
-                            \mitbf{\Sigma}(\mitbf{\theta})^{-1}\G
-                            \left(\G^T\mitbf{\Sigma}(\mitbf{\theta})^{-1}\G\right)^{-1}
-                            \G^T\mitbf{\Sigma}(\mitbf{\theta})^{-1}.
-\end{equation}
-\citet{Harville1974} showed that choosing the hyperparameters which
-maximize eq. (\ref{eq:REML}) is equivalent to choosing the
-hyperparameters which maximize the probability of drawing $\data_*$
-from $\data$. We use the REML method over the maximum likelihood
-method \citep[e.g.,][]{Langbein1997} because the REML method accounts
-for the improper prior that we assigned to $\mitbf{a}$
-\citep{Hines2017}. We independently estimate $\mitbf{\theta}$ for each
-station, and so $\data_*$ consists of displacements for an individual
-station. We are assuming $u(p)=0$ when estimating the noise
-hyperparameters for this section.
+\mitbf{K}(\mitbf{\theta}) = \mitbf{C}_{\eta_i}(\mitbf{\theta})^{-1} - 
+                            \mitbf{C}_{\eta_i}(\mitbf{\theta})^{-1}\G
+                            \left(\G^T\mitbf{C}_{\eta_i}(\mitbf{\theta})^{-1}\G\right)^{-1}
+                            \G^T\mitbf{C}_{\eta_i}(\mitbf{\theta})^{-1}.
+\end{equation}
 
@@
 The distribution of inferred $\alpha$ and $\beta$ are shown in Figure
-\ref{fig:NoiseParams}. The amplitude of FOGM noise, $\beta$, for the
-easting and northing components is notably low and are clustered
-around 0.5 mm/yr$^{0.5}$. The corresponding estimates of $\alpha$ tend
-to cluster around 0 yr$^{-1}$, suggesting that noise can be described
-with Brownian motion. We also estimate hyperparameters for the
-vertical component of displacements, under the hope that vertical
-deformation gradients could reveal some geophysical signal. The
-amplitude of FOGM noise for the vertical component is relatively large
-with a median value of 13.5 mm/yr$^{0.5}$.  The inferred values for
-$\alpha$ are also higher for the vertical component with a median
-value of 8.21 yr$^{-1}$. In Figure \ref{fig:NoiseSamples}, we use the
-median values of $\alpha$ and $\beta$ to generate two random samples
-of FOGM noise for each component. The samples span five years and over
-these five years the easting and northing samples drift by about 1 mm.
-In the context of detecting SSEs, which produce several mm's of
-surface displacement on the time-scale of weeks, the estimated FOGM
-noise for the easting and northing component is negligible. In
-contrast, the estimated FOGM noise for the vertical component is
-larger than the signal we would expect from SSEs. We suspect that the
-higher amplitude for the FOGM noise in the vertical component is
-accommodating for deficiencies in our rather simple seasonal model.
-Based on this analysis, we henceforth ignore temporally correlated
-noise in the easting and northing component because of its low
-amplitude. We also do not use vertical displacements because of the
-presumably low signal-to-noise ratio.
+\ref{fig:NoiseParams}. The estimates of $\beta$ for the easting and
+northing components are notably low and are clustered around 0.5
+mm/yr$^{0.5}$. The corresponding estimates of $\alpha$ tend to cluster
+around 0 yr$^{-1}$. We also estimate $\alpha$ and $\beta$ for the vertical
+component of displacements, with the hope that vertical deformation
+gradients could reveal some geophysical signal. For the vertical
+component, $\beta$ is relatively large with a median value of 13.5
+mm/yr$^{0.5}$. The inferred values for $\alpha$ are also higher for
+the vertical component with a median value of 8.21 yr$^{-1}$. In
+Figure \ref{fig:NoiseSamples}, we use the median values of $\alpha$
+and $\beta$ to generate two realizations of FOGM noise for each
+component. The realizations span five years, and the easting and
+northing realizations drift by about 1 mm over these five years. In
+the context of detecting SSEs, which produce several mm's of surface
+displacement on the time-scale of weeks, the estimated FOGM noise for
+the easting and northing component is negligible. In contrast, the
+estimated FOGM noise for the vertical component is larger than the
+signal we would expect from SSEs. We suspect that the higher amplitude
+for the FOGM noise in the vertical component is accommodating for
+deficiencies in our rather simple seasonal model. Based on this
+analysis, we henceforth ignore temporally correlated noise in the
+easting and northing component because of its low amplitude. We also
+do not use vertical displacements because of the presumably low
+signal-to-noise ratio.
 
@@
 \caption{
-Distribution of estimated FOGM hyperparameters (eq. \ref{eq:FOGM}).
-Hyperparameters are estimated with the REML method for 38 stations in
+Distribution of estimated FOGM parameters (eq. \ref{eq:FOGM}).
+parameters are estimated with the REML method for 38 stations in
 the Pacific Northwest that are east of $121^\circ$W. ``IQR'' is the
@@
 \caption{
-Two FOGM noise samples for each component. The FOGM hyperparameters
+Two FOGM noise samples for each component. The FOGM parameters
 have been set to the median values from Figure \ref{fig:NoiseParams}.
@@
 displacements. However, estimated transient strain rates are
-insensitive to common mode error. We therefore do not include common
-mode error in our noise model. We then make the simplifying assumption
-that $\eta(p) = 0$ for the easting and northing component of GNSS
-data.
+insensitive to common mode error, and so we do not need to include
+common mode error in our noise model.
 

### 2017-10-26 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 We consider the noise vector, $\mitbf{\eta}_i$, to be composed of a
-temporally correlated Gaussian process, $\mathcal{L}_i \sim
-\mathcal{GP}(0,C_{\mathcal{L}_i})$, and a vector of uncorrelated
+temporally correlated Gaussian process, $z_i \sim
+\mathcal{GP}(0,C_{z_i})$, and a vector of uncorrelated
 Gaussian noise, $\mitbf{w}_i$, so that
 \begin{equation}
-\mitbf{\eta}_i = \mathcal{L}_i(\points) + \mitbf{w}_i.
+\mitbf{\eta}_i = z_i(\points) + \mitbf{w}_i.
 \end{equation}
@@
 \begin{equation}
-\mitbf{C}_{\eta_i} = C_{\mathcal{L}_i}(\points,\points) +
+\mitbf{C}_{\eta_i} = C_{z_i}(\points,\points) +
                      \mathrm{diag}(\mitbf{\sigma}_i^2).
@@
 because it accurately describes the power spectrum of motion resulting
-from instability in geodetic monuments \citep[e.g.,][]{Wyatt1982,
+from instability in a geodetic monument \citep[e.g.,][]{Wyatt1982,
 Wyatt1989}. In particular, the power spectrums for Brownian motion and
-monument motion are both proportional to $f^{-2}$, where $f$ is
-frequency. Brownian motion, by definition, contains a reference time
-at which the process begins. Since there is no notion of when noise
-``begins" in GNSS data, we do not find Brownian motion to be an
-appropriate noise model. On the other hand, a FOGM process also has a
-power spectrum that is proportional to $f^{-2}$, above some corner
-frequency, and it has no reference time (i.e., a FOGM process is
-stationary). We then let $\mathcal{L}_i$ be a spatially uncorrelated
-FOGM process, which has the covariance function
+monument motion both decay proportionally to $f^{-2}$, where $f$ is
+frequency. However, Brownian motion necessarily contains a reference
+time at which the process begins. Since there is no notion of when
+noise ``begins" in GNSS data, we do not find Brownian motion to be an
+appropriate noise model. On the other hand, a FOGM process has no
+reference time (i.e., it is stationary), and its power spectrum,
+\begin{equation}\label{eq:FOGMPS}
+S(f) = \frac{\beta^2}{(2 \pi f)^2 + \alpha^2},
+\end{equation}
+decays proportionally to $f^{-2}$ above the cutoff frequency $\alpha /
+(2 \pi)$. We then choose $z_i$ to be a spatially
+uncorrelated FOGM process, which has the covariance function
 \begin{equation}\label{eq:FOGM}
-C_{\mathcal{L}_i}\left((\pos,t),(\pos',t')\right) = 
+C_{z_i}\left((\pos,t),(\pos',t')\right) = 
 \frac{\beta^2}{2\alpha}
-\exp\left(-\alpha|t - t'|\right)\delta_{\pos,\pos'}
-\end{equation}
-\citep[sec. B2.1]{Rasmussen2006}, where $\delta_{\pos,\pos'}$ is 1 if
-$\pos = \pos'$ and 0 otherwise.
+\exp\left(-\alpha|t - t'|\right)\delta_{\pos,\pos'},
+\end{equation}
+where $\delta_{\pos,\pos'}$ is 1 if $\pos = \pos'$ and 0 otherwise.
+See \citet[sec. B2]{Rasmussen2006} for more information on the FOGM
+process.
 
@@
 
-By choosing a FOGM process for $\mathcal{L}_i$, we have introduced two
+By choosing a FOGM process for $z_i$, we have introduced two
 parameters that need to be constrained, $\alpha$ and $\beta$. We
@@
 Maximum Likelihood Estimation (MLE) method from \citet{Langbein1997},
-where we pick $\mitbf{\theta}$ to maximize the likelihood of drawing
-the observed data, $\mitbf{d}_i^*$, from the random vector
-$\mitbf{d}_i$. However, unlike the MLE method, the REML method
-produces unbiased estimates of $\mitbf{\theta}$ \citep[sec.
-2.6]{Cressie1992}. We use the REML method to independently estimate
-$\mitbf{\theta}$ at 38 continuous GNSS stations in the Pacific
-Northwest that are east of 121$^\circ$W. These stations are
-sufficiently far from the subduction zone that they are unlikely to
-contain transient signal associated with SSEs. We assume that the
-noise at these inland stations is representative of the noise at all
-the stations considered in this study, which is probably a poor
-assumption since the inland stations are subject to distinctly
-different climatic conditions. Nonetheless, we find the
-$\mitbf{\theta}$ for each inland station and for each displacement
-component that maximizes the REML likelihood function
+in that we attempt to pick $\mitbf{\theta}$ to maximize the
+probability of drawing the observed data, $\mitbf{d}_i^*$, from the
+random vector $\mitbf{d}_i$. However, unlike the MLE method, the REML
+method produces unbiased estimates of $\mitbf{\theta}$ \citep[sec.
+2.6]{Cressie1992}. We use the REML method to estimate $\mitbf{\theta}$
+at 38 continuous GNSS stations in the Pacific Northwest that are east
+of 121$^\circ$W. These stations are sufficiently far from the
+subduction zone that they are unlikely to contain transient signal
+associated with SSEs, allowing us to ignore displacements from $u_i$
+in $\mitbf{d}_i$. We assume the noise at these inland stations is
+representative of the noise at all the stations considered in this
+study, which is probably a poor assumption since the inland stations
+are subject to distinctly different climatic conditions. Nonetheless,
+we find $\mitbf{\theta}$ for each of the inland stations and for each
+displacement component by maximizing the REML likelihood function
+
 \begin{equation}\label{eq:REML}
@@
                                     \left| \G^T\mitbf{C}_{\eta_i}(\mitbf{\theta})^{-1}\G \right|}\right)^{\frac{1}{2}} 
-                              e^{-\tfrac{1}{2}\data_i^{*T}\mitbf{K}(\mitbf{\theta})\data_i^*},
+                              e^{-\tfrac{1}{2}\data_i^{*T}\mitbf{K}(\mitbf{\theta})\data_i^*}
 \end{equation}
@@
 \ref{fig:NoiseParams}. The estimates of $\beta$ for the easting and
-northing components are notably low and are clustered around 0.5
-mm/yr$^{0.5}$. The corresponding estimates of $\alpha$ tend to cluster
-around 0 yr$^{-1}$. We also estimate $\alpha$ and $\beta$ for the vertical
-component of displacements, with the hope that vertical deformation
-gradients could reveal some geophysical signal. For the vertical
-component, $\beta$ is relatively large with a median value of 13.5
-mm/yr$^{0.5}$. The inferred values for $\alpha$ are also higher for
-the vertical component with a median value of 8.21 yr$^{-1}$. In
-Figure \ref{fig:NoiseSamples}, we use the median values of $\alpha$
-and $\beta$ to generate two realizations of FOGM noise for each
-component. The realizations span five years, and the easting and
-northing realizations drift by about 1 mm over these five years. In
-the context of detecting SSEs, which produce several mm's of surface
-displacement on the time-scale of weeks, the estimated FOGM noise for
-the easting and northing component is negligible. In contrast, the
-estimated FOGM noise for the vertical component is larger than the
-signal we would expect from SSEs. We suspect that the higher amplitude
-for the FOGM noise in the vertical component is accommodating for
-deficiencies in our rather simple seasonal model. Based on this
-analysis, we henceforth ignore temporally correlated noise in the
-easting and northing component because of its low amplitude. We also
-do not use vertical displacements because of the presumably low
-signal-to-noise ratio.
+northing components are clustered around 0.5 mm/yr$^{0.5}$. The
+corresponding estimates of $\alpha$ tend to cluster around 0
+yr$^{-1}$, indicating that the power spectrum of noise indeed decays
+proportionally to $f^{-2}$ over a wide band of frequencies. We also
+estimate $\alpha$ and $\beta$ for the vertical component of
+displacements, with the hope that vertical deformation gradients could
+reveal some geophysical signal. For the vertical component, $\beta$ is
+relatively large with a median value of 13.5 mm/yr$^{0.5}$. The
+inferred values for $\alpha$ are also higher for the vertical
+component with a median value of 8.21 yr$^{-1}$. In Figure
+\ref{fig:NoiseSamples}, we use the median values of $\alpha$ and
+$\beta$ to generate two realizations of FOGM noise for each component.
+The realizations span five years, and the easting and northing
+realizations drift by about 1 mm over these five years. In the context
+of detecting SSEs, which produce several mm's of surface displacement
+on the time-scale of weeks, the estimated FOGM noise for the easting
+and northing component is negligible. In contrast, the estimated FOGM
+noise for the vertical component is larger than the signal we would
+expect from SSEs. We suspect that the higher amplitude for the FOGM
+noise in the vertical component is accommodating for deficiencies in
+our rather simple seasonal model. Based on this analysis, we
+henceforth ignore temporally correlated noise in the easting and
+northing component because of its low amplitude. We also do not use
+vertical displacements because of the presumably low signal-to-noise
+ratio.
 

### 2017-10-26 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 \end{align}
-respectively.
+
+respectively. The above derivatives of $\mu_{\post_i}$ and
+$C_{\post_i}$ can be computed analytically by replacing the terms
+$C_{u_i}(p,\points)$, $C_{u_i}(\points,p')$, and $C_{u_i}(p,p')$ in
+eqs. (\ref{eq:PosteriorMean3}) and (\ref{eq:PosteriorCov3}) with their
+appropriate derivatives. For example, the mean and covariance
+functions for $\strain_{\e\e}$ are
+\begin{equation}\label{eq:StrainMeanEE}
+\mu_{\strain_{\e\e}}(p) = \left[\begin{array}{cc}
+                          \frac{\partial^2 C_{u_\e}\left( p,\points \right)}{\partial t \, \partial x_\e} & 
+                          \zeros \\
+                          \end{array}\right]
+                          \left[\begin{array}{cc}
+                          C_{u_\e}(\points,\points) + \mitbf{C}_{\eta_\e} & \G \\
+                          \G^T  & \zeros \\
+                          \end{array}\right]^{-1}
+                          \left[\begin{array}{c}
+                          \data^*_\e \\
+                          \zeros \\
+                          \end{array}\right]
+\end{equation}    
+and
+\begin{multline}\label{eq:StrainCovEE}
+C_{\strain_{\e\e}}(p,p') = \frac{\partial^4 C_{u_\e}(p,p')}{\partial t \, \partial t' \, \partial x_\e \, \partial x'_\e} - \\
+                           \left[\begin{array}{cc}
+                           \frac{\partial^2 C_{u_\e}\left( p,\points \right)}{\partial t \, \partial x_\e} & \zeros \\
+                           \end{array}\right]
+                           \left[\begin{array}{cc}
+                           C_{u_\e}(\points,\points) + \mitbf{C}_{\eta_\e} & \G \\
+                           \G^T  & \zeros \\
+                           \end{array}\right]^{-1}
+                           \left[\begin{array}{c}
+                           \frac{\partial^2 C_{u_\e}\left(\points, p' \right)}{\partial t' \, \partial x'_\e} \\
+                           \zeros \\
+                           \end{array}\right].
+\end{multline}
 

### 2017-10-26 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 \end{align} 
-
 respectively. The above derivatives of $\mu_{\post_i}$ and
-$C_{\post_i}$ can be computed analytically by replacing the terms
+$C_{\post_i}$ are computed analytically by replacing the terms
 $C_{u_i}(p,\points)$, $C_{u_i}(\points,p')$, and $C_{u_i}(p,p')$ in
@@
 appropriate derivatives. For example, the mean and covariance
-functions for $\strain_{\e\e}$ are
+functions for $\strain_{\e\e}$ can be written more verbosely as
 \begin{equation}\label{eq:StrainMeanEE}

### 2017-10-27 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 \mu_{\strain_{\e\e}}(p) = \left[\begin{array}{cc}
-                          \frac{\partial^2 C_{u_\e}\left( p,\points \right)}{\partial t \, \partial x_\e} & 
-                          \zeros \\
+                          \frac{\partial^2 C_{u_\e}\left( p,\points \right)}{\partial t \, \partial x_\e} & \zeros \\
                           \end{array}\right]
@@
 
+% State our choice for X_i
+
 We next establish our prior model for transient displacements.
-Specifically, we discuss our choice for the covariance functions
-$X(\pos,\pos')$ and $T(t,t')$. For $X$, we use the squared exponential
-(SE) covariance function,
+Specifically, we discuss our choice for the functions $X_i$ and $T_i$.
+For $X_i$, we use the squared exponential (SE) function,
 \begin{equation}\label{eq:SE}
-X(\pos,\pos') = \exp\left(\frac{-||\pos - \pos'||_2^2}{2 \ell^2}\right).
-\end{equation}
-The SE covariance function is commonly used in kriging
+X_i(\pos,\pos') = \exp\left(\frac{-||\pos - \pos'||_2^2}{2 \ell^2}\right).
+\end{equation}
+As we have mentioned, the SE function is commonly used for kriging
 \citep[e.g,][]{Cressie1992} and Gaussian process regression
@@
 \citet{Kato1998} and \cite{El-Fiky1999} demonstrated that the SE
-accurately describes the covariance of secular GNSS derived velocities
-in Japan. The SE is a positive definite covariance function for any
-number of spatial dimensions. A Gaussian process with an SE covariance
-function is isotropic and has realizations that are infinitely
-differentiable.
-
-We consider three potential models for the temporal covariance of $u$.
-First, we consider the one-dimensional SE covariance function,
+function accurately describes the spatial covariance of secular GNSS
+derived velocities in Japan.
+
+% List our choices for T_i
+
+We consider three potential choices for $T_i$. The first option is the
+one-dimensional SE function,
 \begin{equation}\label{eq:TimeSE}
-T(t,t') = \phi^2\exp\left(\frac{-|t - t'|^2}{2\tau^2}\right).
-\end{equation}
-Note that $T$ includes the hyperparameter $\phi$, which serves to
-scale the covariance function $C_u$. Second, we consider integrated
-Brownian motion (IBM). IBM has zero mean and its covariance function
-can be found by integrating the covariance function for Brownian
-motion as
+T_i(t,t') = \phi^2\exp\left(\frac{-|t - t'|^2}{2\tau^2}\right).
+\end{equation}
+Note that $T_i$ includes the parameter $\phi$, which serves to scale
+the covariance function $C_{u_i}$. The second option is the Wendland
+class of covariance functions \citep{Wendland2005}. Wendland
+covariance functions have compact support and hence their
+corresponding covariance matrices are sparse. In our analysis, we
+exploit this sparsity with the CHOLMOD software package
+\citep{Chen2008}. Wendland functions are positive definite in
+$\mathbb{R}^d$, and they can be differentiated $k$ times. The form of
+the covariance function depends on the choice of $d$ and $k$. We use
+$d=1$ since we are describing the temporal covariance of $u_i$. We use
+$k=2$, giving samples of $u_i$ continuous velocities and
+accelerations. The corresponding Wendland covariance function is
+\begin{equation}\label{eq:Wendland} 
+T_i(t,t') = \phi^2\left(1 - \frac{|t - t'|}{\tau}\right)^5_+ 
+            \left(\frac{8|t - t'|^2}{\tau^2} + \frac{5|t - t'|}{\tau} + 1\right), 
+\end{equation} 
+where $(t)_+ = \mathrm{min}(0,t)$.
+The third option for $T_i$ is the covariance function for integrated
+Brownian motion (IBM). IBM is a Gaussian process with zero mean and
+its covariance function can be found by integrating the covariance
+function for Brownian motion,
 \begin{align}\label{eq:IBM}
-T(t,t') &= \int_0^t \int_0^{t'} \phi^2 \min(s,s') \,ds'\,ds \\
+T_i(t,t') &= \int_0^t \int_0^{t'} \phi^2 \min(s,s') \,ds'\,ds \\
           &= \frac{\phi^2}{2}\min(t,t')^2 \left(\max(t,t') - \frac{1}{3}\min(t,t')\right), \ \ \ t,t' \geq 0.
@@
 non-parametric model for the time dependence of geophysical signals
-\citep[e.g.,][]{Segall1997, McGuire2003, Ohtani2010, Hines2016a}. It
-should be emphasized $t=0$ is a reference time at which the Gaussian
-process is exactly zero. For some geophysical signals, it is
-appropriate to have this reference time. For example, if we are trying
-to identify postseismic deformation then $t$ should be zero at the
-time of the earthquake.  However, if we are interesting in detecting
-transient events, where there is no known start time, then IBM may not
-be an appropriate prior, and an isotropic Gaussian process should be
-preferred. In the following analysis, we make the quite arbitrary
-choice that $t$ is zero on the first epoch of $\data_*$. Using an
-earlier reference time does not change the results discussed in this
-section. Our third option for $T$ is the Wendland class of covariance
-functions \citep{Wendland2005}. Wendland covariance functions have
-compact support and hence their corresponding covariance matrices are
-sparse. In our analysis, we exploit this sparsity with the CHOLMOD
-software package \citep{Chen2008}. Wendland functions are positive
-definite in $\mathbb{R}^d$, and they describes an isotropic Gaussian
-process with realizations that can be differentiated $k$ times. The
-form of the covariance function depends on the choice of $d$ and $k$.
-We use $d=1$ since we are describing the temporal covariance of $u$.
-We use $k=2$, giving samples of $u$ continuous velocities and
-accelerations. The corresponding Wendland covariance function is
-\begin{equation}\label{eq:Wendland}
-T(t,t') = \phi^2\left(1 - \frac{|t - t'|}{\tau}\right)^5_+ \left(\frac{8|t - t'|^2}{\tau^2} + \frac{5|t - t'|}{\tau} + 1\right), 
-\end{equation}
-where
-\begin{equation}
-(t)_+ = 
-\begin{cases}
-t, \ \ \ t > 0 \\
-0, \ \ \ \mathrm{otherwise}.
-\end{cases}
-\end{equation}
-
-We next determine appropriate hyperparameters for $X$ and each of the
-three candidate covariance functions for $T$. we divide the dataset
+\citep[e.g.,][]{Segall1997, McGuire2003, Ohtani2010, Hines2016a}.
+Similar to Brownian motion, IBM has a reference time, $t=0$, at which
+the process begins. For some geophysical signals, it is appropriate to
+have this reference time. For example, if we are trying to identify
+postseismic deformation then $t$ should be zero at the time of the
+earthquake.  However, if we are interesting in detecting transient
+events, where there is no known start time, then IBM may not be an
+appropriate prior. Instead, one should prefer an isotropic choice for
+$T_i$, meaning that it is invariant to time shifts and can be written
+in terms of $|t - t'|$. In the following analysis, we make the choice
+that $t$ is zero on the first epoch of $\data_i^*$. Using an earlier
+reference time does not change the results discussed in this section.
+
+% Choosing the parameters
+
+% SHOW REML EQUATION IN HERE
+
+We next use the REML method to determine appropriate parameters for
+$X_i$ and each of the three choices for $T_i$. we divide the dataset
 into seven subsets which are four months long and each centered on the
 time of a SSE. The times of the seven SSEs are determined with tremor
-records from \cite{Wech2010}. We use the REML method to find the
-optimal hyperparameters for $T$ and $X$ for each subset of data. We
-choose to make each data subsets four months long because it is long
-enough to encompass a SSE in the Pacific Northwest, while it is short
-enough to still be computationally tractable. However, four months is
-too short to resolve the sinusoids in $\data$, and they are omitted
-from $\data$ in this REML analysis for Pacific Northwest SSEs. The
-estimated hyperparameters for $u$ are summarized in Table 1. Based on
-the interquartile ranges, the estimated hyperparameters for the SE and
-Wendland covariance functions do not vary significantly between SSEs.
-This suggests that the median of estimated hyperparameters should be
-an appropriate prior model for all Pacific Northwest SSEs. For the IBM
-model, there are several anomalously large values for $\ell$ and
-$\phi$, which results in large interquartile ranges.
+records from \cite{Wech2010}. We find the optimal parameters for $T_i$
+and $X_i$ for each subset of data. We choose to make each data subsets
+four months long because it is long enough to encompass a SSE in the
+Pacific Northwest, while it is short enough to still be
+computationally tractable. However, four months is too short to
+resolve the sinusoids in $\data_i$, and they are omitted from
+$\data_i$ in this REML analysis for Pacific Northwest SSEs.
+
+% Results for our parameters
+
+The estimated hyperparameters for $u$ are summarized in Table 1. Based
+on the interquartile ranges, the estimated hyperparameters for the SE
+and Wendland covariance functions do not vary significantly between
+SSEs. This suggests that the median of estimated hyperparameters
+should be an appropriate prior model for all Pacific Northwest SSEs.
+For the IBM model, there are several anomalously large values for
+$\ell$ and $\phi$, which results in large interquartile ranges.
 

### 2017-10-29 RevisedSubmission/hines_hetland_2017_gji.tex
did more shit

@@
 Maximum Likelihood Estimation (MLE) method from \citet{Langbein1997},
-in that we attempt to pick $\mitbf{\theta}$ to maximize the
-probability of drawing the observed data, $\mitbf{d}_i^*$, from the
-random vector $\mitbf{d}_i$. However, unlike the MLE method, the REML
-method produces unbiased estimates of $\mitbf{\theta}$ \citep[sec.
+where we pick $\mitbf{\theta}$ to maximize the probability of drawing
+the observed data, $\mitbf{d}_i^*$, from the random vector
+$\mitbf{d}_i$. However, unlike the MLE method, the REML method
+produces unbiased estimates of $\mitbf{\theta}$ \citep[sec.
 2.6]{Cressie1992}. We use the REML method to estimate $\mitbf{\theta}$
@@
 displacement component by maximizing the REML likelihood function
-
-\begin{equation}\label{eq:REML}
+\begin{equation}\label{eq:REMLNoise}
 \mathcal{L}(\mitbf{\theta}) = \left(\frac{\left|\G^T\G\right|}{(2\pi)^{n-6m} 
@@
 \end{equation}
+In the above equation, $\data^*_i$ consists of the data for a single station.
 
@@
 Specifically, we discuss our choice for the functions $X_i$ and $T_i$.
-For $X_i$, we use the squared exponential (SE) function,
+For $X_i$, we use the squared exponential (SE) covariance function,
 \begin{equation}\label{eq:SE}
@@
 \end{equation}
-As we have mentioned, the SE function is commonly used for kriging
+The SE covariance function is commonly used for kriging
 \citep[e.g,][]{Cressie1992} and Gaussian process regression
@@
 \citet{Kato1998} and \cite{El-Fiky1999} demonstrated that the SE
-function accurately describes the spatial covariance of secular GNSS
-derived velocities in Japan.
+covariance function accurately describes the spatial covariance of
+secular GNSS derived velocities in Japan.
 
@@
 We consider three potential choices for $T_i$. The first option is the
-one-dimensional SE function,
+one-dimensional SE covariance function,
 \begin{equation}\label{eq:TimeSE}
@@
 Note that $T_i$ includes the parameter $\phi$, which serves to scale
-the covariance function $C_{u_i}$. The second option is the Wendland
-class of covariance functions \citep{Wendland2005}. Wendland
-covariance functions have compact support and hence their
+the covariance function $C_{u_i}$. The second option is a member of
+the Wendland class of covariance functions \citep{Wendland2005}.
+Wendland covariance functions have compact support and hence their
 corresponding covariance matrices are sparse. In our analysis, we
 exploit this sparsity with the CHOLMOD software package
-\citep{Chen2008}. Wendland functions are positive definite in
-$\mathbb{R}^d$, and they can be differentiated $k$ times. The form of
-the covariance function depends on the choice of $d$ and $k$. We use
-$d=1$ since we are describing the temporal covariance of $u_i$. We use
-$k=2$, giving samples of $u_i$ continuous velocities and
+\citep{Chen2008}. Wendland covariance functions can be constructed
+such that they are positive definite in $\mathbb{R}^d$ and their
+corresponding Gaussian process can be differentiated $k$ times. We use
+$d=1$, because we are describing the temporal covariance of $u_i$, and
+we use $k=2$, giving samples of $u_i$ continuous velocities and
 accelerations. The corresponding Wendland covariance function is
@@
 \end{equation} 
-where $(t)_+ = \mathrm{min}(0,t)$.
-The third option for $T_i$ is the covariance function for integrated
-Brownian motion (IBM). IBM is a Gaussian process with zero mean and
-its covariance function can be found by integrating the covariance
-function for Brownian motion,
+where $(t)_+ = \mathrm{max}(0,t)$. The third option for $T_i$ is the
+covariance function for integrated Brownian motion (IBM). IBM is a
+Gaussian process with zero mean and its covariance function can be
+found by integrating the covariance function for Brownian motion,
 \begin{align}\label{eq:IBM}
@@
 events, where there is no known start time, then IBM may not be an
-appropriate prior. Instead, one should prefer an isotropic choice for
-$T_i$, meaning that it is invariant to time shifts and can be written
-in terms of $|t - t'|$. In the following analysis, we make the choice
-that $t$ is zero on the first epoch of $\data_i^*$. Using an earlier
-reference time does not change the results discussed in this section.
+appropriate prior model. Instead, one may prefer to use the SE or
+Wendland covariance functions because they are stationary. In the
+following analysis, we make the choice that $t$ is zero on the first
+epoch of $\data_i^*$. Using an earlier reference time does not change
+the results discussed in this section.
 
@@
 
-% SHOW REML EQUATION IN HERE
-
-We next use the REML method to determine appropriate parameters for
-$X_i$ and each of the three choices for $T_i$. we divide the dataset
-into seven subsets which are four months long and each centered on the
-time of a SSE. The times of the seven SSEs are determined with tremor
-records from \cite{Wech2010}. We find the optimal parameters for $T_i$
-and $X_i$ for each subset of data. We choose to make each data subsets
-four months long because it is long enough to encompass a SSE in the
-Pacific Northwest, while it is short enough to still be
-computationally tractable. However, four months is too short to
-resolve the sinusoids in $\data_i$, and they are omitted from
-$\data_i$ in this REML analysis for Pacific Northwest SSEs.
-
-% Results for our parameters
-
-The estimated hyperparameters for $u$ are summarized in Table 1. Based
-on the interquartile ranges, the estimated hyperparameters for the SE
-and Wendland covariance functions do not vary significantly between
-SSEs. This suggests that the median of estimated hyperparameters
-should be an appropriate prior model for all Pacific Northwest SSEs.
-For the IBM model, there are several anomalously large values for
-$\ell$ and $\phi$, which results in large interquartile ranges.
+We use the REML method to determine the appropriate values for the
+parameters in $X_i$ and $T_i$, $\mitbf{\theta} = \{\ell, \phi, \tau
+\}$. we divide the GNSS data into seven subsets that are four months
+long and each centered on the time of a SSE. The times of the seven
+SSEs are determined with tremor records from \cite{Wech2010}. We find
+the optimal $\mitbf{\theta}$ for each subset of data, for each
+displacement component, and for each choice of $T_i$. We choose to
+make each data subsets four months long because it is long enough to
+encompass a SSE in the Pacific Northwest, while it is still
+computationally tractable to estimate $\mitbf{\theta}$. However, four
+months is too short to resolve the sinusoids in $\data_i$, and they
+are omitted from $\data_i$ in this REML analysis. The REML likelihood
+function that we are maximizing with respect to $\mitbf{\theta}$ is
+now 
+\begin{equation}\label{eq:REMLPrior} 
+\mathcal{L}(\mitbf{\theta}) =
+\left(\frac{\left| \G^T \G \right|}{(2\pi)^{n-6m}
+                                    \left| \mitbf{\Sigma}(\mitbf{\theta}) \right| 
+                                    \left| \G^T \mitbf{\Sigma}(\mitbf{\theta})^{-1} \G \right|}\right)^{\frac{1}{2}} 
+                              e^{-\tfrac{1}{2} \data_i^{*T} \mitbf{K}(\mitbf{\theta}) \data_i^*},
+\end{equation}
+where
+\begin{equation}
+\mitbf{K}(\mitbf{\theta}) = \mitbf{\Sigma}(\mitbf{\theta})^{-1} - 
+                            \mitbf{\Sigma}(\mitbf{\theta})^{-1} \G
+                            \left( \G^T \mitbf{\Sigma}(\mitbf{\theta})^{-1} \G \right)^{-1}
+                            \G^T \mitbf{\Sigma}(\mitbf{\theta})^{-1}
+\end{equation}
+and $\mitbf{\Sigma}(\mitbf{\theta}) = \mitbf{C}_{\eta_i} +
+C_{u_i}(\points,\points;\mitbf{\theta})$. In the above equation,
+$\data^*_i$ consists of a single subset of data. The estimated
+parameters are summarized in Table 1. Based on the interquartile
+ranges, the estimated parameters for the SE and Wendland covariance
+functions do not vary significantly between SSEs. This suggests that
+the median of estimated parameters should be an appropriate prior
+model for all Pacific Northwest SSEs. For the IBM model, there are
+several anomalously large values for $\ell$ and $\phi$, which is why
+the interquartile ranges are large.
 
@@
 \caption{
-Optimal hyperparameters for the prior on transient displacements
-determined with the REML method. The temporal covariance function is
-indicated by the ``$T$'' column. The SE, IBM, and Wendland covariance
-functions are defined in eqs. (\ref{eq:TimeSE}), (\ref{eq:IBM}), and
+Optimal parameters for the prior on transient displacements determined
+with the REML method. The temporal covariance function is indicated by
+the ``$T$'' column. The SE, IBM, and Wendland covariance functions are
+defined in eqs. (\ref{eq:TimeSE}), (\ref{eq:IBM}), and
 (\ref{eq:Wendland}), respectively. The spatial covariance function,
 $X$, is the squared exponential (eq. \ref{eq:SE}) in all cases. The
-hyperparameters are estimated for each of the seven SSEs considered in
+parameters are estimated for each of the seven SSEs considered in
 this study, and the tabulated values indicate the median and
@@
 We compute $\hat{\data}$ using SE, Wendland, and IBM covariance
-functions for $T$ and the median hyperparameters from Table 1. Figure
+functions for $T$ and the median parameters from Table 1. Figure
 \ref{fig:Fit} compares the easting component of $\data_*$ to
@@
 the chosen covariance function. This over-smoothing could indicate
-that the chosen time-scale hyperparameter, $\tau$, is too large. The
+that the chosen time-scale parameter, $\tau$, is too large. The
 predicted displacements at the remaining stations, which record more
@@
 we ultimately settle on the Wendland covariance function for $T$ and
-use the median values from Table 1 for the hyperparameters. We choose
+use the median values from Table 1 for the parameters. We choose
 the Wendland covariance function over the SE covariance function

### 2017-10-30 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 postseismic deformation then $t$ should be zero at the time of the
-earthquake.  However, if we are interesting in detecting transient
+earthquake.  However, if we are interested in detecting transient
 events, where there is no known start time, then IBM may not be an
@@
 parameters in $X_i$ and $T_i$, $\mitbf{\theta} = \{\ell, \phi, \tau
-\}$. we divide the GNSS data into seven subsets that are four months
+\}$. We divide the GNSS data into seven subsets that are four months
 long and each centered on the time of a SSE. The times of the seven
-SSEs are determined with tremor records from \cite{Wech2010}. We find
-the optimal $\mitbf{\theta}$ for each subset of data, for each
-displacement component, and for each choice of $T_i$. We choose to
-make each data subsets four months long because it is long enough to
-encompass a SSE in the Pacific Northwest, while it is still
+SSEs are determined with tremor records from \cite{Wech2010}. We
+choose to make each data subsets four months long because it is long
+enough to encompass a SSE in the Pacific Northwest, while it is still
 computationally tractable to estimate $\mitbf{\theta}$. However, four
 months is too short to resolve the sinusoids in $\data_i$, and they
-are omitted from $\data_i$ in this REML analysis. The REML likelihood
-function that we are maximizing with respect to $\mitbf{\theta}$ is
-now 
+are omitted from $\data_i$ in this REML analysis. We find the optimal
+$\mitbf{\theta}$ for each subset of data, for each displacement
+component, and for each choice of $T_i$. The REML likelihood function
+that we are maximizing with respect to $\mitbf{\theta}$ is now
 \begin{equation}\label{eq:REMLPrior} 
@@
 functions do not vary significantly between SSEs. This suggests that
-the median of estimated parameters should be an appropriate prior
-model for all Pacific Northwest SSEs. For the IBM model, there are
-several anomalously large values for $\ell$ and $\phi$, which is why
-the interquartile ranges are large.
+the median of the estimated parameters should be appropriate for all
+the Pacific Northwest SSEs. For the IBM model, there are several
+anomalously large values for $\ell$ and $\phi$, which is why the
+interquartile ranges are large.
 
@@
 \caption{
-Optimal parameters for the prior on transient displacements determined
-with the REML method. The temporal covariance function is indicated by
-the ``$T$'' column. The SE, IBM, and Wendland covariance functions are
-defined in eqs. (\ref{eq:TimeSE}), (\ref{eq:IBM}), and
-(\ref{eq:Wendland}), respectively. The spatial covariance function,
-$X$, is the squared exponential (eq. \ref{eq:SE}) in all cases. The
-parameters are estimated for each of the seven SSEs considered in
+Optimal values for the parameters in $X_i$ and $T_i$ determined with
+the REML method. The covariance function for $T_i$ is indicated by the
+``$T_i$'' column. The SE, Wendland, and IBM covariance functions are
+defined in eqs. (\ref{eq:TimeSE}), (\ref{eq:Wendland}), and
+(\ref{eq:IBM}), respectively. We use eq. (\ref{eq:SE}) for $X_i$.
+The parameters are estimated for each of the seven SSEs considered in
 this study, and the tabulated values indicate the median and
 interquartile ranges of estimates. The ``diff $\log$(REML)'' column
-compares the log REML likelihood to the log REML likelihood when using
-the SE covariance function for $T$. Negative values indicate that
-observations are more consistent with the SE covariance function.
+compares the optimal log REML likelihoods to those for when $T_i$ is
+the SE covariance function. Negative values indicate that observations
+are more consistent with the SE covariance function.
 } 
 \begin{tabular} {l l l l l l}
-$T$ & direction & $\ell$  & $\phi$   & $\tau$  & diff. $\log$(REML) \\ \hline
+$T_i$ & component & $\ell$  & $\phi$   & $\tau$  & diff. $\log$(REML) \\ \hline
 SE & east   & 92 $\pm$ 25 km  & 0.62 $\pm$ 0.11 mm  & 0.026 $\pm$ 0.011 yr  &  - \\
@@
 
-Next we identify which covariance function for $T$ best describes the
-SSEs. One approach is to compare the REML likelihoods for each
-covariance function, similar to the analysis in \citet{Langbein2004}.
-In Table 1, we summarize how the log REML likelihoods for the Wendland
-and IBM covariance functions compare to the SE covariance function.
-Based on the differences in log REML likelihoods, the data is
-substantially more likely to come from a Gaussian process with a SE or
-Wendland covariance function than an IBM covariance function. The REML
-likelihoods do not definitively indicate whether the SE or Wendland
-covariance function is preferable.
-
-To further explore which covariance function for $T$ best describes
-SSEs, we compare the observations to the predicted displacements for
-each covariance function. We consider the data prediction vector to be
-$\hat{\data} = \left(u(\points) + \G\mitbf{a}\right)|\data_*$.
-Following a similar procedure as in Section \ref{sec:Method}, it can
-be shown that $\hat{\data}$ is normally distributed with mean
+% Choosing the covariance function for T_i with REML
+
+Next, we identify which covariance function for $T_i$ best describes
+the time dependence of deformation from SSEs. One approach is to
+choose the covariance function that produces the largest optimal REML
+likelihoods, similar to the analysis in \citet{Langbein2004}. In Table
+1, we summarize how the optimal REML likelihoods for the Wendland and
+IBM covariance functions compare to those for the SE covariance
+function. Based on the differences in optimal REML likelihoods, the
+data is substantially more likely to come from a Gaussian process with
+a SE or Wendland covariance function than an IBM covariance function.
+The REML likelihoods do not definitively indicate whether the SE or
+Wendland covariance function is preferable.
+
+% Choosing the covariance function for T_i with posterior fit
+
+A more intuitive way of deciding which function to use for $T_i$ is to
+compare the posterior displacements and the observed displacements.
+The posterior displacements consist of our estimate of transient
+displacements, secular trends, and seasonal deformation. Specifically,
+the posterior displacement vector is $\hat{\data}_i = \left(
+u_i(\points) + \G\mitbf{m}_i \right) | \left( \data_i = \data_i^*
+\right)$. Following a similar procedure as in Section
+\ref{sec:Method}, it can be shown that $\hat{\data}_i$ has a
+multivariate normal distribution with mean vector
 \begin{equation}\label{eq:DataPredMean}
-\mitbf{\mu}_{\hat{d}} = \left[\begin{array}{cc}
-                           C_u(\points,\points) & \G \\
+\mitbf{\mu}_{\hat{d}_i} = \left[\begin{array}{cc}
+                          C_{u_i}(\points,\points) & \G \\
                           \end{array}\right]
                           \left[\begin{array}{cc}
-                           \mitbf{\Sigma} & \G \\
+                          \mitbf{C}_{\eta_i} + C_{u_i}(\points,\points) & \G \\
                           \G^T  & \zeros \\
@@
                           \left[\begin{array}{c}
-                           \data_* \\
+                          \data_i^* \\
                           \zeros \\
@@
 \end{equation}  
-and covariance
+and covariance matrix
 \begin{equation}\label{eq:DataPredCov}
-\mitbf{C}_{\hat{\data}} = C_u(\points,\points) - 
+\mitbf{C}_{\hat{d}_i} = C_{u_i}(\points,\points) - 
                         \left[\begin{array}{cc}
-                              C_u(\points,\points) & \G \\
+                              C_{u_i}(\points,\points) & \G \\
                               \end{array}\right]
                         \left[\begin{array}{cc}
-                              \mitbf{\Sigma} & \G \\
+                              \mitbf{C}_{\eta_i} + C_{u_i}(\points,\points) & \G \\
                               \G^T  & \zeros \\
@@
                         \left[\begin{array}{c}
-                              C_u(\points,\points) \\
+                              C_{u_i}(\points,\points) \\
                               \G^T \\
@@
 \end{equation}
-We compute $\hat{\data}$ using SE, Wendland, and IBM covariance
-functions for $T$ and the median parameters from Table 1. Figure
-\ref{fig:Fit} compares the easting component of $\data_*$ to
-$\hat{\data}$ during the winter 2015-2016 SSE at the three stations
-that record the strongest SSE signal, ALBH, P436, and SC02. Based on
-Figure \ref{fig:Fit}, $\hat{\data}$ is insensitive to the choice of
-covariance function for $T$. The predicted displacements for the IBM
-covariance function contain slightly more high frequency, and perhaps
-spurious, features. When comparing $\hat{\data}$ to $\data_*$ at
-station ALBH and SC02, it appears that $\hat{\data}$ slightly
-underestimates the rate of deformation during the SSE, regardless of
-the chosen covariance function. This over-smoothing could indicate
-that the chosen time-scale parameter, $\tau$, is too large. The
-predicted displacements at the remaining stations, which record more
-subtle SSE deformation, seem to faithfully describe $\data_*$.
+We compute $\hat{\data}_i$ using the SE, Wendland, and IBM covariance
+functions for $T_i$, and we use the median values from Table 1 for
+$\mitbf{\theta}$. Figure \ref{fig:Fit} compares $\data_i^*$ to
+$\hat{\data}_i$ for the winter 2015-2016 SSE at stations ALBH, P436,
+and SC02. Based on Figure \ref{fig:Fit}, the chosen covariance
+function for $T_i$ has almost no effect on $\hat{\data}_i$. The
+posterior displacements for the IBM covariance function contain
+slightly more high frequency, and perhaps spurious, features.
+Regardless of the chosen covariance function, $\hat{\data}_i$ appears
+to underestimate the rate of deformation during the SSE at stations
+ALBH and SC02. The deformation at these two stations is particularly
+rapid compared to the surrounding stations, and the misfit between
+$\data_i^*$ and $\hat{\data}_i$ is likely due to over-smoothing. This
+over-smoothing could indicate that the chosen values for $\tau$ or
+$\ell$ are too large. However, $\hat{\data}_i$ does faithfully fit
+$\data_i^*$ at the remaining stations, and so we do not attempt to
+further refine the parameters.
+
+% The choice
 
 For our estimates of transient strain discussed in the next section,
-we ultimately settle on the Wendland covariance function for $T$ and
-use the median values from Table 1 for the parameters. We choose
-the Wendland covariance function over the SE covariance function
-because of its computational advantages.
+we ultimately settle on the Wendland covariance function for $T_i$,
+and we use the median values from Table 1 for the parameters. We
+choose the Wendland covariance function over the SE covariance
+function because of its computational advantages.
 
@@
 \caption{
-Easting component of the observed displacements, $\data_*$, and
-predicted displacements, $\hat{\data}$, during an SSE at three
-stations where the SSE signal is strongest. The predicted
-displacements are shown for when $T$ is a squared exponential (SE),
-Wendland, and integrated Brownian motion (IBM) covariance function.
-The one standard deviation uncertainties are shown for the
-observations and the SE predictions. For clarity, uncertainties are
-not shown for the IBM and Wendland predictions, but they are nearly
-equivalent to the uncertainties for the SE predictions. The SE and
-Wendland predictions are practically indistinguishable.
+Easting component of the observed displacements, $\data_\e^*$, and
+posterior displacements, $\hat{\data}_\e$, during the winter 2015-2016
+SSE. We show $\hat{\data}_\e$ for when $T_i$ is a SE, Wendland, and
+IBM covariance function. The one standard deviation uncertainties are
+shown for $\data_\e^*$ and the $\hat{\data}_\e$ that has a SE
+covariance function for $T_i$. For clarity, uncertainties are not
+shown for $\hat{\data}_\e$ when $T_i$ is an IBM or Wendland covariance
+function, but the uncertainties are nearly equivalent to the
+uncertainties for when $T_i$ is the SE covariance function. The
+posterior displacements for the different covariance functions are all
+practically indistinguishable.
 }   

### 2017-10-30 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 We next establish our prior model for transient displacements.
-Specifically, we discuss our choice for the functions $X_i$ and $T_i$.
-For $X_i$, we use the squared exponential (SE) covariance function,
+Specifically, we discuss our choice for the spatial and temporal
+functions making up $C_{u_i}$, $X_i$ and $T_i$. For $X_i$, we use the
+squared exponential (SE) covariance function,
 \begin{equation}\label{eq:SE}
@@
 We use the REML method to determine the appropriate values for the
-parameters in $X_i$ and $T_i$, $\mitbf{\theta} = \{\ell, \phi, \tau
-\}$. We divide the GNSS data into seven subsets that are four months
-long and each centered on the time of a SSE. The times of the seven
-SSEs are determined with tremor records from \cite{Wech2010}. We
-choose to make each data subsets four months long because it is long
-enough to encompass a SSE in the Pacific Northwest, while it is still
-computationally tractable to estimate $\mitbf{\theta}$. However, four
-months is too short to resolve the sinusoids in $\data_i$, and they
-are omitted from $\data_i$ in this REML analysis. We find the optimal
-$\mitbf{\theta}$ for each subset of data, for each displacement
-component, and for each choice of $T_i$. The REML likelihood function
-that we are maximizing with respect to $\mitbf{\theta}$ is now
+parameters in $C_{u_i}$, $\mitbf{\theta} = \{\ell, \phi, \tau \}$.
+Just as in Section \ref{sec:NoiseModel}, we pick $\mitbf{\theta}$ to
+maximize the probability of drawing $\data_i^*$ from $\data_i$, but
+now $\data_i$ includes $u_i(\points)$, and $\mitbf{\theta}$ refers to
+the parameters for $C_{u_i}(\points,\points)$ rather than
+$\mitbf{C}_{\eta_i}$. We divide the GNSS data into seven subsets that
+are four months long and each centered on the time of a SSE. The times
+of the seven SSEs are determined with tremor records from
+\cite{Wech2010}. We find the optimal $\mitbf{\theta}$ for each subset
+of data, for each displacement component, and for each choice of
+$T_i$. We choose to make each data subsets four months long because it
+is long enough to encompass a SSE in the Pacific Northwest, while it
+is still computationally tractable to estimate $\mitbf{\theta}$. The
+REML likelihood function that we are maximizing with respect to
+$\mitbf{\theta}$ is now
 \begin{equation}\label{eq:REMLPrior} 

### 2017-11-01 RevisedSubmission/hines_hetland_2017_gji.tex
did more

--- 2017-10-30 RevisedSubmission/hines_hetland_2017_gji.tex
@@
 of 121$^\circ$W. These stations are sufficiently far from the
-subduction zone that they are unlikely to contain transient signal
-associated with SSEs, allowing us to ignore displacements from $u_i$
-in $\mitbf{d}_i$. We assume the noise at these inland stations is
+subduction zone that they are unlikely to record transient deformation
+from SSEs, allowing us to ingore the term $u_i(\points)$ in
+$\mitbf{d}_i$. We assume the noise at these inland stations is
 representative of the noise at all the stations considered in this
@@
 We use the REML method to determine the appropriate values for the
-parameters in $X_i$ and $T_i$, $\mitbf{\theta} = \{\ell, \phi, \tau
-\}$. We divide the GNSS data into seven subsets that are four months
-long and each centered on the time of a SSE. The times of the seven
-SSEs are determined with tremor records from \cite{Wech2010}. We
-choose to make each data subsets four months long because it is long
-enough to encompass a SSE in the Pacific Northwest, while it is still
-computationally tractable to estimate $\mitbf{\theta}$. However, four
-months is too short to resolve the sinusoids in $\data_i$, and they
-are omitted from $\data_i$ in this REML analysis. We find the optimal
-$\mitbf{\theta}$ for each subset of data, for each displacement
-component, and for each choice of $T_i$. The REML likelihood function
-that we are maximizing with respect to $\mitbf{\theta}$ is now
+parameters in $C_{u_i}$, which are $\ell$, $\phi$, and $\tau$. We
+refer to these parameters collectively as $\mitbf{\theta}$. Just as in
+Section \ref{sec:NoiseModel}, we pick $\mitbf{\theta}$ to maximize the
+probability of drawing $\data_i^*$ from $\data_i$, but now we are
+including $u_i(\points)$ in $\data_i$. and we are optimizing the
+parameters for $C_{u_i}$ rather than $\mitbf{C}_{\eta_i}$. We divide
+the GNSS data into seven subsets that are four months long and each
+centered on the time of a SSE. The times of the seven SSEs are
+determined with tremor records from \cite{Wech2010}. We find the
+optimal $\mitbf{\theta}$ for each subset of data, for each
+displacement component, and for each choice of $T_i$. The REML
+likelihood function that we are maximizing with respect to
+$\mitbf{\theta}$ is now
 \begin{equation}\label{eq:REMLPrior} 
@@
 $\mitbf{\theta}$. Figure \ref{fig:Fit} compares $\data_i^*$ to
-$\hat{\data}_i$ for the winter 2015-2016 SSE at stations ALBH, P436,
-and SC02. Based on Figure \ref{fig:Fit}, the chosen covariance
-function for $T_i$ has almost no effect on $\hat{\data}_i$. The
-posterior displacements for the IBM covariance function contain
+$\hat{\data}_i$ during the winter 2015-2016 SSE at stations ALBH,
+P436, and SC02, which are the stations that recorded the strongest
+signal from this SSE. Based on Figure \ref{fig:Fit}, the chosen
+covariance function for $T_i$ has almost no effect on $\hat{\data}_i$.
+The posterior displacements for the IBM covariance function contain
 slightly more high frequency, and perhaps spurious, features.
@@
 covariance function for $T_i$. For clarity, uncertainties are not
-shown for $\hat{\data}_\e$ when $T_i$ is an IBM or Wendland covariance
-function, but the uncertainties are nearly equivalent to the
-uncertainties for when $T_i$ is the SE covariance function. The
-posterior displacements for the different covariance functions are all
-practically indistinguishable.
+shown for the $\hat{\data}_\e$ that have an IBM or Wendland covariance
+function for $T_i$. The posterior displacements for the different
+covariance functions are all practically indistinguishable.
 }   

### 2017-11-02 RevisedSubmission/hines_hetland_2017_gji.tex
did more fucking shit

@@
 
+% Stating how we calculate strain and refering to the figures
+
 Having established a noise model and a prior for transient
-displacements, we use the GNSS dataset to calculate transient strain
-rates in the Puget Sound region. We calculate transient strain rates
-for each day from January 1, 2010 to May 15, 2017. The strain rates
-are estimates at a grid of points spanning the study area. In Figure
-\ref{fig:StrainMap} we show the transient strain rates on January 1,
-2016, which coincides with the height of the winter 2015-2016 SSE. We
-have included an animation showing the map view of strain rates
-through time as supplementary material. The strain rates shown in
-Figure \ref{fig:StrainMap} are generally similar to the strain rates
-during the other six SSEs considered in this study. The SSEs cause
-trench perpendicular compression in the Olympic Peninsula and
-extension east of Puget Sound. The strain transitions from compression
-to extension around the southern tip of Vancouver Island, which
-coincides with the location of thrust slip for SSEs in the Puget Sound
-region \citep[e.g.,][]{Dragert2001,Wech2009,Schmidt2010}. Thus, this
-pattern of strain is to be expected. During the period in between
-SSEs, secular strain rates indicate trench perpendicular compression
-throughout this study region \citep{Murray2000, McCaffrey2007,
-McCaffrey2013}. When comparing inferred strain rates from SSEs to the
-secular strain rates, we see that SSEs are concentrating tectonically
-accumulated strain energy towards the trench, and presumably pushing
-the subduction zone closer to failure.
+displacements, we can now use the GNSS data to calculate transient
+strain rates, $\strain$, in the Puget Sound region. We evaluate
+$\strain$ at a grid of points spanning the study area for each day
+from January 1, 2010 to May 15, 2017. In Figure \ref{fig:StrainMap},
+we show a map view of $\strain$ on January 1, 2016, which coincides
+with the height of the winter 2015-2016 SSE. In Figure
+\ref{fig:StrainTs}, we show a time series of $\strain$ at a position
+on the Olympic Peninsula, which tends to be where $\strain$ is the
+largest during SSEs. We also include a supplementary animation showing
+a map view of $\strain$ over time.
+
+% Describe the strain during SSEs and tectonic implications
+
+The strain rates shown in Figure \ref{fig:StrainMap} are generally
+similar to the strain rates during the other six SSEs considered in
+this study. The SSEs cause trench perpendicular compression in the
+Olympic Peninsula and extension east of Puget Sound. The strain
+transitions from compression to extension around the southern tip of
+Vancouver Island, which coincides with the location of thrust slip for
+SSEs in the Puget Sound region \citep[e.g.,][]{Dragert2001, Wech2009,
+Schmidt2010}. Thus, this pattern of strain is to be expected. During
+the period in between SSEs, secular strain rates indicate trench
+perpendicular compression throughout this study region
+\citep{Murray2000, McCaffrey2007, McCaffrey2013}. When comparing
+inferred strain rates from SSEs to the secular strain rates, we see
+that SSEs are concentrating tectonically accumulated strain energy
+towards the trench, and presumably pushing the subduction zone closer
+to failure.
 
@@
 \caption{
-Estimated transient strain rates during the Winter 2015-2016 SSE.
-Strain glyphs show the normal strain rate along each azimuth, where
-orange indicates compression and blue indicates extension. The shaded
-regions indicate one standard deviation uncertainties in the normal
-strain rates.
+Transient strain rates during the Winter 2015-2016 SSE. The glyphs
+show the normal strain rates as a function of azimuth, where orange
+indicates compression and blue indicates extension. The shaded regions
+indicate the one standard deviation uncertainties for the normal
+strain rates. It is common to depict strain rates by showing just the
+principal normal strain rates; however, we deviate from this
+convention because there is no straight-forward way of showing the
+uncertainties for the principal normal strain rates.
 }   
@@
 
-In Figure \ref{fig:StrainTs} we show the time dependence of estimated
-transient strain rates at a position on the Olympic Peninsula, where
-transient strain rates from SSEs are largest. To verify that the
-estimated transient strain rates are accurately identifying
-geophysical signal, we compare the signal-to-noise ratio from eq.
-(\ref{eq:SNR}) to the frequency of seismic tremor (Figure
+% compare seismic tremor to SNR
+
+To verify that the estimated transient strain rates are accurately
+identifying geophysical signal, we compare the signal-to-noise ratio
+from eq. (\ref{eq:SNR}) to the frequency of seismic tremor (Figure
 \ref{fig:StrainMag}). A signal-to-noise ratio greater than ${\sim}3$

### 2017-11-02 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 Our motivation for estimating transient strain rates is, in part, to
-detect geophysical phenomena. As we will see, geophysical signal can
-be easily identified by visually inspecting the solution for $\strain$
-from eqs. (\ref{eq:StrainMean}) and (\ref{eq:StrainCov}). However, if
-we want to detect geophysical phenomena automatically, then we need to
-define a detection criterion. We use a signal-to-noise ratio, SNR,
-that is based on the Frobenius norm of $\strain$,
-$||\strain||_\mathrm{F} = \left(\strain_{\e\e}^2 + \strain_{\n\n}^2 +
-2\strain_{\e\n}^2\right)^{\frac{1}{2}}$, for our detection criterion.
-In the geodetic literature, $||\dot{\varepsilon}||_\mathrm{F}$ is
-often used as a metric for the strain rate ``magnitude", and it is
-sometimes referred to as the second invariant of strain rate. Noting
-that $||\strain||_\mathrm{F}$ is a random variable, we take SNR to be
-the ratio of the estimated mean and standard deviation of
-$||\strain||_\mathrm{F}$. An estimate of the mean is found by
-evaluating $||\strain||_\mathrm{F}$ at the mean of $\strain$,
+detect transient geophysical processes. As we will see, geophysical
+signal can be easily identified by visually inspecting the solution
+for $\strain$ from eqs. (\ref{eq:StrainMean}) and
+(\ref{eq:StrainCov}). However, if we want to detect geophysical signal
+automatically, then we need to define a detection criterion. We use a
+signal-to-noise ratio, SNR, that is based on the Frobenius norm of
+$\strain$, $||\strain||_\mathrm{F} = \left(\strain_{\e\e}^2 +
+\strain_{\n\n}^2 + 2\strain_{\e\n}^2\right)^{\frac{1}{2}}$, for our
+detection criterion. In the geodetic literature,
+$||\dot{\varepsilon}||_\mathrm{F}$ is often used as a metric for the
+strain rate ``magnitude", and it is sometimes referred to as the
+second invariant of strain rate. Noting that $||\strain||_\mathrm{F}$
+is a random variable, we take SNR to be the ratio of the estimated
+mean and standard deviation of $||\strain||_\mathrm{F}$. An estimate
+of the mean is found by evaluating $||\strain||_\mathrm{F}$ at the
+mean of $\strain$,
 \begin{align}
@@
 
-We next establish our prior model for transient displacements.
+We next establish a prior model for transient displacements.
 Specifically, we discuss our choice for the spatial and temporal
@@
 Having established a noise model and a prior for transient
-displacements, we can now use the GNSS data to calculate transient
-strain rates, $\strain$, in the Puget Sound region. We evaluate
-$\strain$ at a grid of points spanning the study area for each day
-from January 1, 2010 to May 15, 2017. In Figure \ref{fig:StrainMap},
-we show a map view of $\strain$ on January 1, 2016, which coincides
-with the height of the winter 2015-2016 SSE. In Figure
-\ref{fig:StrainTs}, we show a time series of $\strain$ at a position
-on the Olympic Peninsula, which tends to be where $\strain$ is the
-largest during SSEs. We also include a supplementary animation showing
-a map view of $\strain$ over time.
+displacements, we can now calculate transient strain rates, $\strain$,
+in the Puget Sound region from the GNSS data. We evaluate $\strain$ at
+a grid of points spanning the study area for each day from January 1,
+2010 to May 15, 2017. In Figure \ref{fig:StrainMap}, we show a map
+view of $\strain$ on January 1, 2016, which coincides with the height
+of the winter 2015-2016 SSE. In Figure \ref{fig:StrainTs}, we show a
+time series of $\strain$ at a position on the Olympic Peninsula, which
+is where $\strain$ tends to be the largest during SSEs. We also
+include a supplementary animation showing a map view of $\strain$ over
+time. In each figure, we show the mean and standard deviation of
+$\strain$, making it easy to identify which features are statistically
+significant.
 
@@
 
-The strain rates shown in Figure \ref{fig:StrainMap} are generally
-similar to the strain rates during the other six SSEs considered in
-this study. The SSEs cause trench perpendicular compression in the
-Olympic Peninsula and extension east of Puget Sound. The strain
+The transient strain rates shown for the winter 2015-2016 SSE in
+Figure \ref{fig:StrainMap} are characteristic of most SSEs in the
+Puget Sound region. The SSEs cause trench perpendicular compression in
+the Olympic Peninsula and extension east of Puget Sound. The strain
 transitions from compression to extension around the southern tip of
-Vancouver Island, which coincides with the location of thrust slip for
-SSEs in the Puget Sound region \citep[e.g.,][]{Dragert2001, Wech2009,
-Schmidt2010}. Thus, this pattern of strain is to be expected. During
-the period in between SSEs, secular strain rates indicate trench
-perpendicular compression throughout this study region
-\citep{Murray2000, McCaffrey2007, McCaffrey2013}. When comparing
-inferred strain rates from SSEs to the secular strain rates, we see
-that SSEs are concentrating tectonically accumulated strain energy
-towards the trench, and presumably pushing the subduction zone closer
-to failure.
-
-%A key difference between the strain inferred here and strain that can
-%be derived from fault slip models is that our estimated strain rates
-%are not based on an assumed physical model. In contrast, fault slip
-%models can be biased by errors in the assumed fault geometry or
-%lithospheric rheology. Moreover, the degrees of freedom in fault slip
-%models usually cannot be constrained by GNSS data alone, and it is
-%necessary to impose regularization which further biases the results.
-%Since our estimated strain rates lack such systematic errors, we can
-%be more confident that our solution is unbiased and has meaningful
-%uncertainties.
+Vancouver Island, coinciding with where fault slip tends to be
+inferred \citep[e.g.,][]{Dragert2001, Wech2009, Schmidt2010}. Thus,
+this pattern of strain is to be expected. Over decadal timescales,
+there is trench perpendicular compression throughout this study region
+caused by steady tectonic plate motion \citep{Murray2000,
+McCaffrey2007, McCaffrey2013}. When comparing the transient strain
+rates caused by SSEs to the secular strain rates, we see that SSEs are
+concentrating tectonically accumulated strain energy towards the
+trench, and presumably pushing the subduction zone closer to failure.
 
@@
 identifying geophysical signal, we compare the signal-to-noise ratio
-from eq. (\ref{eq:SNR}) to the frequency of seismic tremor (Figure
-\ref{fig:StrainMag}). A signal-to-noise ratio greater than ${\sim}3$
-can be interpreted as a detected geophysical signal. We detect nine
-distinct events, which each correspond to peaks in seismic tremor. The
-smaller events detected in August 2014 and February 2017 can be
-considered inter-SSE events. They were not among the SSEs used to
-constrain the prior covariance function. In between peaks in seismic
-tremor, the signal-to-noise ratio is consistently between 0 and 2,
-suggesting that all the transient strain detected at this location is
-associated with SSEs and inter-SSE events.
+from eq. (\ref{eq:SNR}) at a point on the Olympic Peninsula to the
+frequency of seismic tremor (Figure \ref{fig:StrainMag}). A
+signal-to-noise ratio greater than ${\sim}3$ can be interpreted as a
+detected geophysical signal. We detect nine distinct events, which
+each correspond to peaks in seismic tremor. The smaller events
+detected in August 2014 and February 2017 can be considered inter-SSE
+events. They were not among the SSEs used to constrain the prior
+covariance function. In between peaks in seismic tremor, the
+signal-to-noise ratio is consistently between 0 and 2, suggesting that
+all the transient strain detected at this point on the Olympic
+Peninsula is associated with SSEs and inter-SSE events.
 
@@
 
-The results we have presented thus far indicate that we are
-identifying the strain that we should expect to see. There are,
-however, subtle features in our estimated transient strain rates which
-we were not expecting. For example, some SSEs are preceded by a brief
-period of east-west extension on the Olympic Peninsula. This feature
-can be seen several days before the summer 2012 and winter 2015-2016
-SSEs in Figure \ref{fig:StrainTs} as well as in the supplementary
-animation. While this deformation is noteworthy, a discussion on the
-mechanisms causing it is outside the scope of this study.
+% note that we are only discussing strain with known origin
+
+The results we have presented indicate that we are identifying the
+transient strain that we should expect to see. This is not to say that
+there are no unexpected features in $\strain$ that are worth further
+exploration. However, a discussion on features in $\strain$ that have
+unknown origin would be outside the scope of this paper.
 

### 2017-11-02 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 \caption{ 
-Panel A) The covariance between $u_i(p)$ and $u_i(p')$,
-$C_{u_i}(p,p')$, as a function of the spatial and temporal distance
-between $p$ and $p'$. The covariance function $C_{u_i}$ is composed of
-squared exponentials to describe the spatial and temporal covariance
-of $u_i$ (eq. \ref{eq:ExampleCov}). Panel B) A single realization of
-$u_i$ corresponding to the covariance function from Panel A. The
-realization is a function of three variables, $t$, $x_\mathrm{e}$, and
-$x_\mathrm{n}$, although we only show its dependence on $t$ and one of
-the spatial dimensions.
+(A) An example covariance function for transient displacements,
+$C_{u_i}(p,p')$, shown as a function of the spatial and temporal
+distance between $p$ and $p'$. (B) A single realization of transient
+displacements, $u_i(p)$, corresponding to the covariance function from
+Panel A. The realization is a function of three variables, $t$,
+$x_\mathrm{e}$, and $x_\mathrm{n}$, although we only show its
+dependence on $t$ and one of the spatial dimensions.
 }    
@@
 observations are made at $m$ stations, and the times and positions for
-each observation are described by the set $\points$. $\data^*_i$ can
-be considered a realization of the random vector $\data_i$, which is
-formed by evaluating eq. (\ref{eq:Data}) at each point in $\points$.
-To write out $\data_i$ explicitly, we let $\G$ be an $n \times 6m$
-matrix consisting of the basis functions from eq. (\ref{eq:Data})
-(i.e., the linear trends and sinusoids for each station) evaluated at
-each point in $\points$. The coefficients corresponding to each basis
-function are collected into the column vector $\mitbf{m}_i$. The noise
-for all the observations are described by the column vector
-$\mitbf{\eta}_i$. We can then write $\data_i$ as
+each observation are described by the set $\points$. The data vector
+$\data^*_i$ can be considered a realization of the random vector
+$\data_i$, which is formed by evaluating eq. (\ref{eq:Data}) at each
+point in $\points$. To write out $\data_i$ explicitly, we let $\G$ be
+an $n \times 6m$ matrix consisting of the basis functions from eq.
+(\ref{eq:Data}) (i.e., the linear trends and sinusoids for each
+station) evaluated at each point in $\points$. The coefficients
+corresponding to each basis function are collected into the column
+vector $\mitbf{m}_i$. The noise for all the observations are described
+by the column vector $\mitbf{\eta}_i$. We can then write $\data_i$ as
 \begin{equation}\label{eq:DataVec}
@@
 \caption{
-Positions of continuous GNSS stations used to estimate transient
-strain rates. The colored regions indicate the distribution of seismic
-tremor as determined by \citet{Wech2010}. The red dots show the
-positions of GNSS stations mentioned in this paper. The blue dot
-indicates the location of the transient strain rates shown in Figure
+Positions of the GNSS stations used to estimate transient strain
+rates. The colored regions indicate the distribution of seismic tremor
+as determined by \citet{Wech2010}. The red dots show the positions of
+GNSS stations mentioned in this paper. The blue dot indicates the
+location of the transient strain rates shown in Figure
 \ref{fig:StrainTs} and the signal-to-noise ratio shown in Figure
@@
 
-The distribution of inferred $\alpha$ and $\beta$ are shown in Figure
-\ref{fig:NoiseParams}. The estimates of $\beta$ for the easting and
-northing components are clustered around 0.5 mm/yr$^{0.5}$. The
-corresponding estimates of $\alpha$ tend to cluster around 0
-yr$^{-1}$, indicating that the power spectrum of noise indeed decays
-proportionally to $f^{-2}$ over a wide band of frequencies. We also
-estimate $\alpha$ and $\beta$ for the vertical component of
+The distribution of inferred values for $\alpha$ and $\beta$ are shown
+in Figure \ref{fig:NoiseParams}. The estimates of $\beta$ for the
+easting and northing components are clustered around 0.5
+mm/yr$^{0.5}$. The corresponding estimates of $\alpha$ tend to cluster
+around 0 yr$^{-1}$, indicating that the power spectrum of noise indeed
+decays proportionally to $f^{-2}$ over a wide band of frequencies. We
+also estimate $\alpha$ and $\beta$ for the vertical component of
 displacements, with the hope that vertical deformation gradients could
@@
 \caption{
-Distribution of estimated FOGM parameters (eq. \ref{eq:FOGM}).
-parameters are estimated with the REML method for 38 stations in
-the Pacific Northwest that are east of $121^\circ$W. ``IQR'' is the
-interquartile range.
+Distribution of inferred values for the parameters in the FOGM noise
+model (eq. \ref{eq:FOGM}). ``IQR'' is the interquartile range of
+inferred values.
 }   
@@
 \caption{
-Two FOGM noise samples for each component. The FOGM parameters
-have been set to the median values from Figure \ref{fig:NoiseParams}.
+Two samples of FOGM noise for each displacement component. The
+parameters for the FOGM noise model have been set to the median values
+from Figure \ref{fig:NoiseParams}.
 }   
@@
 this study, and the tabulated values indicate the median and
-interquartile ranges of estimates. The ``diff $\log$(REML)'' column
+interquartile ranges of estimated values. The ``diff $\log$(REML)'' column
 compares the optimal log REML likelihoods to those for when $T_i$ is
@@
 indicate the one standard deviation uncertainties for the normal
-strain rates. It is common to depict strain rates by showing just the
-principal normal strain rates; however, we deviate from this
-convention because there is no straight-forward way of showing the
-uncertainties for the principal normal strain rates.
+strain rates.
 }   
@@
 \caption{
-Three components of the transient horizontal strain rate tensor
-estimated at the position shown in Figure \ref{fig:Context}. The
-shaded regions indicate one standard deviation uncertainty.
+Components of the transient strain rates at the position shown in
+Figure \ref{fig:Context}. The shaded regions indicate the one standard
+deviation uncertainties.
 }   
@@
 (top) Signal-to-noise ratio (eq. \ref{eq:SNR}) at the position shown
-in Figure \ref{fig:Context}. (bottom) Frequency of tremors in the
-region shown in Figure \ref{fig:Context}.
+in Figure \ref{fig:Context}. (bottom) Frequency of seismic tremors in
+the region shown in Figure \ref{fig:Context}.
 }   

### 2017-11-08 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 \begin{summary}
+
+% focus on subjective vs object rather than "we have uncertainties"
 
@@
 epochs. \citet{Holt2013} defined a detection threshold based on the
-strain rate magnitude, and below we demonstrate that this is indeed an
+strain rate magnitude, and we demonstrate that this is indeed an
 effective criterion for identifying geophysical signal. For the same
@@
 spatially correlated. When not accounted for, common mode error
-manifests as spatially uniform undulations in estimated transient
-displacements. However, estimated transient strain rates are
-insensitive to common mode error, and so we do not need to include
-common mode error in our noise model.
+manifests as spatially uniform undulations in our estimated transient
+displacements. These undulations have no effect on the transient
+strain rates, and so we do not need to include common mode error in
+our noise model.
 
@@
 \section{Discussion}\label{sec:Discussion}
+
+% issues and potential improvements on noise
+
 Our results demonstrate that GPR is an effective tool for estimating
-transient strain from GNSS data, which can then be used to detect
-geophysical processes. One may argue that geophysical signal can also
-be detected by merely inspecting the GNSS displacement time series.
-Indeed, the SSEs identified in Figure \ref{fig:StrainMag} do produce
-visible displacements in the GNSS data. However, the GNSS data also
-contains outliers and non-tectonic deformation that is localized to
-individual stations. In contrast, our estimates of transient strain
-only identify features that are sufficiently spatially and temporally
-coherent, based on our chosen prior model. Furthermore, our estimates
-of transient strain are insensitive to common mode noise, which is
-highly spatially correlated noise resulting from factors such as
-reference frame error. Common mode noise can obscure geophysical
-signal in the GNSS data, but it gets canceled out when computing the
-transient strain. Lastly, our estimates of transient strain rates are
-a spatial and temporal derivative of displacements, and thus any
-geophysical signal in the transient strain rates tends to be more
-pronounced than in the GNSS data. For these reasons, we argue that
-transient strain rates estimated with the method described in Section
-\ref{sec:Method} can illuminate geophysical signal that may not be
-discernible from the noise in the GNSS displacement data.
+transient strain rates and detecting geophysical processes from GNSS
+data. However, there are some aspects of our method that warrant
+further research. For example, there is potential for a more thorough
+analysis of the spatio-temporal noise in the GNSS data than what was
+performed in Section \ref{sec:NoiseModel}. Specifically, we did not
+explore spatially correlated noise, such as from common mode error. We
+have assumed that any spatially correlated noise has a sufficiently
+long wavelength that it has a negligible effect on transient strain
+rates. We can also improve upon the seasonal noise model used in this
+study, which consists of four spatially uncorrelated sinusoids for
+each station. We did not explore the spatial covariance of seasonal
+deformation or the temporal roughness (i.e., the number of sinusoids
+needed to describe the observations). The periodic Gaussian process
+\citep{Mackay1998} is an alternative model for seasonal deformation
+and is well suited for exploring the roughness of seasonal
+deformation.  The periodic Gaussian process has zero mean and the
+covariance function
+
+\begin{equation}\label{eq:Periodic}
+  T(t,t') = \phi^2 \exp\left(\frac{-\sin(\pi|t - t'|)^2}{2\tau^2}\right). 
+\end{equation}
+
+Realizations have annual periodicity and the roughness is controlled
+by $\tau$. Decreasing $\tau$ has the same effect as including higher
+frequency sinusoids in the seasonal model. The optimal value for
+$\tau$ can be found with the REML method as described in Section
+\ref{sec:NoiseModel}.
+
+% issues and potential improvements on numerical consideratiosn
+
+% Numerical considerations
+
+% note that the compact covariance function is not a perfect solution
+% because it is only compact if the signal has a short timescale. Long
+% wavelength features will still have a dense covariance matrix. This is
+% part of the reason why we omitted the temporal noise model.
+
+There are numerical considerations as well ...
+The transient strain rates estimated in this study are constrained by
+about seven years of daily displacement observations from 94 GNSS
+stations. It can be computationally intensive to evaluate eqs.
+(\ref{eq:PosteriorMean3}) and (\ref{eq:PosteriorCov3}) for a dataset
+with this size. We significantly reduce the amount of memory needed to
+estimate transient strain rates by describing the temporal covariance
+of displacements with a compact Wendland covariance function. Using a
+compact covariance function for our prior turns eqs.
+(\ref{eq:PosteriorMean3}) and (\ref{eq:PosteriorCov3}) into sparse
+systems of equations, which we then solve with CHOLMOD. CHOLMOD is
+designed for solving sparse, positive definite systems of equations.
+The matrix being inverted in eqs. (\ref{eq:PosteriorMean3}) and
+(\ref{eq:PosteriorCov3}) is not positive definite; however, we can use
+another partitioned matrix inversion identity from \citet{Press2007}
+to partition it into positive definite submatrices to be inverted.
+Even when using a compact covariance function, it may still be
+necessary to reduce the computational burden by dividing the data into
+subsets and evaluating transient strain rates for each subset.
+
+% additional applications
 
@@
 data and quantify its noise.
-
-There is potential for a more thorough analysis of the spatio-temporal
-noise in GNSS data, $\eta$, than what was performed in Section
-\ref{sec:NoiseModel}. We did not explore the spatial covariance of
-$\eta$, which would describe common mode noise. We are able to ignore
-common mode error in this study; however, for other geophysical
-studies based on GNSS data, such as fault slip inversions, it may be
-necessary to incorporate a spatially covarying noise model
-\citep[e.g.,][]{Miyazaki2003}. We can also improve upon the seasonal
-model used in this study, which consists of four spatially
-uncorrelated sinusoids for each station. We did not explore the
-spatial covariance of seasonal deformation or the temporal roughness
-(i.e., the number of sinusoids needed to describe the observations).
-The periodic Gaussian process \citep{Mackay1998} is an alternative
-model for seasonal deformation and is well suited for exploring the
-roughness of seasonal deformation.  The periodic Gaussian process has
-zero mean and the covariance function
-\begin{equation}\label{eq:Periodic}
-T(t,t') = \phi^2 \exp\left(\frac{-\sin(\pi|t - t'|)^2}{2\tau^2}\right).
-\end{equation}
-Realizations have annual periodicity and the roughness is controlled
-by $\tau$. Decreasing $\tau$ has the same effect as including higher
-frequency sinusoids in the seasonal model. The optimal value for
-$\tau$ can be found with the REML method as described in Section
-\ref{sec:NoiseModel}.
-
-The transient strain rates estimated in this study are constrained by
-about seven years of daily displacement observations from 94 GNSS
-stations. It can be computationally intensive to evaluate eqs.
-(\ref{eq:PosteriorMean3}) and (\ref{eq:PosteriorCov3}) for a dataset
-with this size. We significantly reduce the amount of memory needed to
-estimate transient strain rates by describing the temporal covariance
-of displacements with a compact Wendland covariance function. Using a
-compact covariance function for our prior turns eqs.
-(\ref{eq:PosteriorMean3}) and (\ref{eq:PosteriorCov3}) into sparse
-systems of equations, which we then solve with CHOLMOD. CHOLMOD is
-designed for solving sparse, positive definite systems of equations.
-The matrix being inverted in eqs. (\ref{eq:PosteriorMean3}) and
-(\ref{eq:PosteriorCov3}) is not positive definite; however, we can use
-another partitioned matrix inversion identity from \citet{Press2007}
-to partition it into positive definite submatrices to be inverted.
-Even when using a compact covariance function, it may still be
-necessary to reduce the computational burden by dividing the data into
-subsets and evaluating transient strain rates for each subset.
 

### 2017-11-08 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 
-% issues and potential improvements on noise
+%% Areas for improvement and further study
+% noise model
 
@@
 performed in Section \ref{sec:NoiseModel}. Specifically, we did not
-explore spatially correlated noise, such as from common mode error. We
+explore spatially correlated noise, such as common mode error. We
 have assumed that any spatially correlated noise has a sufficiently
@@
 rates. We can also improve upon the seasonal noise model used in this
-study, which consists of four spatially uncorrelated sinusoids for
-each station. We did not explore the spatial covariance of seasonal
-deformation or the temporal roughness (i.e., the number of sinusoids
-needed to describe the observations). The periodic Gaussian process
-\citep{Mackay1998} is an alternative model for seasonal deformation
-and is well suited for exploring the roughness of seasonal
+study, which consists of four sinusoids for each station. We did not
+explore the roughness of the seasonal deformation (i.e., the number of
+sinusoids needed to describe the deformation). The periodic Gaussian
+process \citep{Mackay1998} is an alternative model for seasonal
+deformation and is well suited for exploring the roughness of seasonal
 deformation.  The periodic Gaussian process has zero mean and the
 covariance function
-
 \begin{equation}\label{eq:Periodic}
@@
 \end{equation}
-
 Realizations have annual periodicity and the roughness is controlled
@@
 frequency sinusoids in the seasonal model. The optimal value for
-$\tau$ can be found with the REML method as described in Section
-\ref{sec:NoiseModel}.
+$\tau$ can be found with the REML method.
 
@@
 
-There are numerical considerations as well ...
-The transient strain rates estimated in this study are constrained by
-about seven years of daily displacement observations from 94 GNSS
-stations. It can be computationally intensive to evaluate eqs.
-(\ref{eq:PosteriorMean3}) and (\ref{eq:PosteriorCov3}) for a dataset
-with this size. We significantly reduce the amount of memory needed to
-estimate transient strain rates by describing the temporal covariance
-of displacements with a compact Wendland covariance function. Using a
-compact covariance function for our prior turns eqs.
-(\ref{eq:PosteriorMean3}) and (\ref{eq:PosteriorCov3}) into sparse
-systems of equations, which we then solve with CHOLMOD. CHOLMOD is
-designed for solving sparse, positive definite systems of equations.
-The matrix being inverted in eqs. (\ref{eq:PosteriorMean3}) and
-(\ref{eq:PosteriorCov3}) is not positive definite; however, we can use
-another partitioned matrix inversion identity from \citet{Press2007}
-to partition it into positive definite submatrices to be inverted.
-Even when using a compact covariance function, it may still be
-necessary to reduce the computational burden by dividing the data into
-subsets and evaluating transient strain rates for each subset.
+% numerical issues
+
+Another potential research direction would be to reduce the
+computational cost of our method for estimating transient strain
+rates. GPR is generally computationally expensive when there are many
+observations. The transient strain rates estimated in this study are
+constrained by seven years of daily displacement observations from 94
+GNSS stations, which amounts to about 240,000 observations for each
+displacement component. For a dataset with this size, it is difficult
+to evaluate the matrix inverses in eqs. (\ref{eq:PosteriorMean3}) and
+(\ref{eq:PosteriorCov3}). We alleviate this computational cost by
+using a compact Wendland covariance function for our prior. By using a
+compact covariance function for our prior, eqs.
+(\ref{eq:PosteriorMean3}) and (\ref{eq:PosteriorCov3}) become sparse
+systems of equations that are more tractable to evaluate. However, the
+sparsity decreases when the Wendland covariance function has a larger
+time-scale parameter. The Wendland covariance then offers less of a
+computational advantage when we were interested in geophysical
+processes that occur over longer times-scales. An alternative approach
+to dealing with large datasets would be to explore approximation
+methods for GPR \citep[sec. 8]{Rasmussen2006}.
 

### 2017-11-09 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 Transient strain rates derived from GNSS data can be used to detect
-and understand geophysical phenomena such as slow slip events and
+and understand geophysical processes such as slow slip events and
 postseismic deformation. Here we propose using Gaussian process
@@
 GNSS data. GPR is a non-parametric, Bayesian method for interpolating
-scattered data. Transient strain rates estimated with GPR have
-meaningful uncertainties, allowing geophysical signal to be easily
-discerned from noise. In our approach, we assume a stochastic prior
-model for transient displacements. The prior describes how much one
-expects transient displacements to covary spatially and temporally. A
+scattered data. In our approach, we assume a stochastic prior model
+for transient displacements. The prior describes how much we expect
+transient displacements to covary spatially and temporally. A
 posterior estimate of transient strain rates is obtained by
-differentiating the posterior displacements. One limitation with GPR
-is that it is not robust against outliers, so we introduce a
-pre-processing method for detecting and removing outliers in GNSS
-data. As a demonstration, we use GPR to detect transient strain
-resulting from slow slip events in the Pacific Northwest. Maximum
-likelihood methods are used to constrain a prior model for transient
-displacements in this region. The temporal covariance of our prior
-model is described by a compact Wendland covariance function, which
-significantly reduces the computational burden that can be associated
-with GPR. Our results reveal the spatial and temporal evolution of
-strain from slow slip events. We verify that the transient strain
-estimated with GPR is in fact geophysical signal by comparing it to
-the seismic tremor that is associated with Pacific Northwest slow slip
-events.
+differentiating the posterior displacements, which are formed by
+conditioning the prior with the GNSS data. As a demonstration, we use
+GPR to detect transient strain resulting from slow slip events in the
+Pacific Northwest. Maximum likelihood methods are used to constrain a
+prior model for transient displacements in this region. The temporal
+covariance of our prior model is described by a compact Wendland
+covariance function, which significantly reduces the computational
+burden that can be associated with GPR. Our results reveal the spatial
+and temporal evolution of strain from slow slip events. We verify that
+the transient strain estimated with GPR is in fact geophysical signal
+by comparing it to the seismic tremor that is associated with Pacific
+Northwest slow slip events.
 
@@
 transient strain rates can be used to better understand the data from
-borehole strain meters (BSMs). The Plate Boundary Observatory
-maintains about forty BSMs in the Pacific Northwest, and it has been
+borehole strain meters (BSMs). The Plate Boundary Observatory contains
+about forty BSMs in the Pacific Northwest, and it has been
 demonstrated that BSMs are able to record transient geophysical events
@@
 \section{Conclusion}\label{sec:Conclusion}
-In this paper we propose using Gaussian process regression (GPR) to
-estimate transient strain rates from GNSS data. Most other methods for
+
+We propose using Gaussian process regression (GPR) to estimate
+transient strain rates from GNSS data. Most other methods for
 estimating strain rates assume a parametric representation of
@@
 recover, then the prior model can be chosen objectively with maximum
-likelihood methods. Because GPR is a Bayesian method, the
-uncertainties on our estimated transient strain rates are well
-quantified, allowing one to discern geophysical signal from noise. We
-demonstrate that GPR is an effective tool for detecting geophysical
-phenomena, such as slow slip events, in our application to GNSS data
-from the Pacific Northwest. One limitation with GPR is that it is not
-robust against outliers. To overcome this limitation, we have
-introduced an effective pre-processing method for identifying and
-removing outliers from GNSS datasets. Another complication with GPR is
-that it usually involves inverting a dense matrix where the number of
-rows and columns is equal to the number of observations. This is
-prohibitive when using several years of daily GNSS observations from a
-network of several hundred stations. We significantly reduce the
-computational burden of GPR by using compact Wendland covariance
-function to describe our prior model. While this paper just focuses on
-estimating transient strain rates, we believe that GPR is a powerful
-tool that can be applied to a wide range of geophysical problems.
+likelihood methods. We demonstrate that GPR is an effective tool for
+detecting geophysical processes, such as slow slip events, in our
+application to the Pacific Northwest. While this paper just focuses on
+using GPR to estimate transient strain rates and detect geophysical
+processes, we believe that GPR is a powerful tool that can be applied
+to a wide range of geophysical problems.
 

### 2017-11-09 ResponseToReviewers/ResponseToReviewers.tex
began response to reviewer

\documentclass[10pt,a4paper]{letter}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}

\begin{document}

\signature{Trever T. Hines}

\begin{letter}{}
\opening{Dear Editor,}

We were pleased with the positive comments from you and the reviewers, and we feel that the comments brought up valid concerns. We hope that we have addressed these concerns sufficiently in our revision. In order to fully address the comments, we have included a supplementary material document.  We have also made some changes to the manuscript which were not specifically requested by the reviewers.  Most of these changes were stylistic and do not change the content of the manuscript.  One of the more substantial changes was that we merged Figures 7 and 14, which are both showing observed and predicted radial components of postseismic displacements, facilitating direct comparison of the figures.

We added supporting figures S3 and S4, which break down the predicted displacements of our preferred model into an elastic and viscoelastic component.  We feel that these figures help support our discussion on lines 509 to 523, where we attribute inferences of fault slip and viscoelastic relaxation to different aspects of the observed postseismic displacements.

On line 500 to 502 of our previous version of this manuscript we make the claim that “After one year, afterslip is inferred to be deeper down on the Sierra Cucapah segment, which is describing much of the sustained near-field postseismic deformation.”  This sentence is not accurate of the preferred model we present in the manuscript. As can be clearly seen in the new figures S3-S4, afterslip is describing some of the later near-field deformation, although most of it is being described by viscoelastic relaxation.  We have reworded this passage and updated the introduction and conclusion to accurately describe our interpretations of the mechanisms driving later near-field deformation.  This does not significantly alter the main conclusions of our manuscript, because we are very clear, both in the original and revised versions, that the deformation mechanism driving later near-field deformation cannot be definitively resolved with these data.

Each of the reviewer comments are included below followed by our response and a description of how these comments were addressed in this revised manuscript.

%% REVIEWER 1
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\textbf{Reviewer \#1:}\newline

\textbf{Review of the paper : Revealing transient strain in geodetic data
with Gaussian process regression}

\textbf{Overview of the work: The paper focuses on a new non-parametric
Bayesian method to estimate transient strain rate from GNSS data. It
is based on Gaussian Process regression. The method also includes an
outlier removal processing step. This technique is then applied to the
detection of slow slip events in Cascadia. The validity of the
detected transient signals is accomplished via comparison with the
seismic tremor recorded along the Cascadia range.}

\textbf{The article is well-written and your algorithm is very interesting. I
think the use of Gaussian Process Regression is novel and open new
roads in GNSS data analysis. I believe that you spend a lot of efforts
in trying to justify most of the assumptions behind your model,
together with the effects of using different kind of prior. Perhaps,
that overshadows slightly your results. You could consider to
reorganize the article to make it more readable for geophysicists.}

\textbf{I also require that a comparison between your outlier removal model
and one other algorithm could help to support your claims about the
efficiency of your method.}

\textbf{Otherwise, I have underlined a few clarifications needed about the
stochastic model and some paragraphs within the text.}

\textbf{General Comments:}

\textbf{About the method:}

\textbf{The method is well established. The assumptions to justify the various
hypothesis on the deterministic models of the GNSS time series are
generally well supported (seasonal signal, tectonic rate, offsets).}

\textbf{1- Now, the correlation of the stochastic processes is not completely
clear. When the authors justify the model behind equation 3, wij is a
normally distributed, uncorrelated noise. In statistic, it is equal to
a WGN process. Then, the authors introduce a separate parameter (eta)
which models the temporally correlated noise. This parameter is also
following a Gaussian distribution, with a zero-mean and a covariance
matrix $C_{\eta}$ (i.e. line 10 p5). This approach is not common in
modelling GNSS time series. The standard is based on Williams 2003,
where the author established a stochastic model with a covariance
matrix as sum of identity matrix for the white noise and another
matrix which represents the coloured noise. The coloured noise
covariance matrix is defined differently (see the literature – Bock and
Melgar Reviews of Geophysics 2016, He et al., Journal of geodynamics
2017).}

\textbf{2- First, the author should give a summary of stochastic noise
modelling in GNSS. I also think that a discussion is needed to relate
to previous models. Why using separated stochastic models? Also I
would comment on the possibility that the (low) spatial correlation
between the parameters in your deterministic model (i.e. secular
velocities) could be absorbed in the estimation of the covariance
matrix $C_{\eta}$.}

\textbf{Now, I am confused When looking at p.8 Section 4.1, the description of
the noise models should be given as a subsection of Section 2 p.4. I
would also improve the literature review. Several other models have
also been discussed (i.e. Montillet et al., 2014 uses a fractional
Brownian motion model; ARMA, ARFIMA, GGM or Band pass noise – see Het
et al., 2017).}

\textbf{Outlier detection:}

\textbf{1- P7 : I would give a longer summary in the introduction about all
the efforts in outlier detection in GNSS time series. The first
paragraph p.7 is not enough.}

\textbf{2- Also, the paper would be improved if the proposed outlier detection
method is compared with existing ones such as the Hector software
package (Bos et al., 2013). For example, the comparison between this
method and another one could be done when applying the algorithm in p.
13 (last paragraph below equation 18).}

\textbf{Results:}

\textbf{The results are generally well explained. However, the use of
different priors and the comparison in the text is sometimes
confusing. Perhaps, a table could summary the different priors and the
main results. It would ease the reading of Section 4.2, 4.3 and then
the discussion in Section 5.}

\textbf{Minor issues:}

\textbf{P2 line 30 : “fidelity” replace with “reliability”.}

\textbf{P2 line 30-31 “ Developing and improving upon methods for deriving
secular...area of research.” Need references.}

\textbf{p.2 line 43 :”too large of an area” ... perhaps “ a very large area”}

\textbf{P5 line 32 : “formal data uncertainty”: can you define it?}

\textbf{P 6 : Need reference to show Equation 15.}

\textbf{P 7 line 54-60: This paragraph can be further enhanced with references
(Bock and Melgar, 2016, Gazeaux et al., 2013).}

\textbf{P 8 : first paragraph. add references on Cascadia range (Aguiar,
Melbourne, Scriver 2009), (Melbourne and Szeliga 2005), (Melbourne and
Webb, 2003, 2002).}

\textbf{P.8 line 20 delete www.unavco.org (only in Acknowledgment)}

\textbf{p. 11 line 45 “common mode error” you need to add some references and
define it properly.}

\textbf{p.13 : “Wendland functions have compact support and hence their
corresponding covariance.. sparse”. I would just mention that the
associated covariance matrices are sparse.}

\textbf{p.13 line 22 what is an “isotropic Gaussian process”? please add a
definition.}

\textbf{P.20 : “ Using a compact covariance function ..” Again I would just
say that the covariance matrix is sparse and that s why you have the
mathematical simplification.}

\textbf{p.20 line 33 “ computational burden” I think that refers to p.20 line
60 “prohibitive when using sveral years of daily GNSS”. It would
interesting to have some number. How do you quantify the computation
time to process longer and longer time series?}

%% REVIEWER 2
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\textbf{Reviewer \#2:}\newline

\textbf{Comments to the Author(s)}
\textbf{Hines/Hetland GJI Review 08/2017}

\textbf{This is a well written paper and I have just a few minor
clarifications and suggestions below.  One point that might be of
interest to address is would it be possible to “approximate” this
approach so that it can run much faster?  I gather from the text that
the method is intensive for CPU and memory use.}

\textbf{Page 3: Last line: min(t,t’) needs as scaling factor to make units for
covariance e.g., m$^2$/yr}

\textbf{Page 4: Stochastic model: It would seem that the assumption of no
covariance between the north and east components is not a very good
one since there is a tectonic framework in which the transients occur
and its unlikely this framework aligns along the Cardinal directions.
Maybe add a statement of impact of the neglect and possibly the idea
of reorienting the “axes” to align with the tectonic framework e.g.,
perpendicular to the subduction zone interface in this case?}

\textbf{Page 4/5 Equation (3): Is it worth discussing at this point whether
the estimates of u and n (eta) can be separated without explicit prior
knowledge (which we may not have).  For example, orbit modeling errors
on one satellite (e.g., due to an unknown yaw problem in the
spacecraft) wilt generate a spatially and temporally correlated error
in $d_{ij}$.  How does this (stochastically non-modeled) error not project
in the u estimates?}

\textbf{Page 6: Equation 10: Maybe it is discussed later in the paper but it
is probably worthwhile stating at this point how the inverse is
performed with a block 0 in the matrix?}

\textbf{Page 6: Equation 12: Maybe some additional explanation is needed here.
How are these partial derivatives formed.  My understanding the u
estimates at this point correspond one-to-one with the positions and
times of the GNSS position determinations.  Did I miss something?}

\textbf{Page 8: Partly addresses issue raised about eqn 3.  It probably should
be noted that wether conditions east of -121 deg longitude are
different to the coast so there could be possible problems with this
noise model.}

%% Editor Review
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\textbf{Editor:}\newline

\textbf{Editor’s Review of GJI MS 17.0593 ‘‘Revealing transient strain...’’ by
Hines and Hetland This is an novel algorithm applied to a problem of
great current interest. Based on the reviewer’s comments and my own
reading, I am recommending moderate revision. The authors’ response
should include a description of their changes, and also a version of
the paper on which these can be seen (just as highlights in the text).}

\textbf{Many of the reviewers’ comments address places where the paper could
be more clearly written, and my overall comment would be the same. The
current version perhaps spends a little too much time on the
background and not enough on explaining the method and its derivation.
As shown by the references below (these include some the authors
cite), and the Dermanis paper, which I have also attached, strain
estimation from scattered data is an area in which there has been a
lot more prior work than the authors mention. I am including these
references not because the authors need to cite many or even most of
them, but only to make the point that the problem of estimating
strain, and of looking for transients, is a familiar one: for many
readers, myself among them, much more familiar than the statistical
methods applied here. So a slightly fuller discussion of these methods
(perhaps a little more background and some reminders of, eg, what a
hyperparameter is) would be welcome to many readers; and increased
clarity of presentation usu- ally means that the paper will be used
and cited more.}

\textbf{I feel the same way about the section on detecting outliers: again, a
well-worked area. Iter- ative fitting and outlier detection, as an
overall strategy, hav e been used for a long time. I appreci- ate that
by imposing spatial and temporal conditions you can identify outliers
not otherwise obvi- ous, but it would help to show an example of this,
rather than the current Figure 4, in which most of the identified
cases are ones that any method could easily find.}

\textbf{A few specifics (numbers are page+line):}

\textbf{2+26: possibility not risk: the latter is a technical term in seismic
hazard.}

\textbf{2+47-48: I do not think this description of Shen’s method (which I’ve
used and programmed) is right, since it allows the deformation
gradients to vary in space. I’ve always thought of it as a kind of
adaptive kernel smoother, somewhat like applying 2-D loess; so it is
only depen- dent on nearby points, and has no assumption of uniformity
anywhere. (Actually this raises a more general question: is the method
given here use local or global support?)}

\textbf{3+7: could also}

\textbf{3+32: redo the sentence; as it stands, you are saying that the SCEC
exercise calculates strain rates.}

\textbf{3+37: a geophysical signal}

\textbf{3+47: function not signal}

\textbf{3+56: by a Gaussian, not with}

\textbf{4+37: displacement (not plural)}

\textbf{6+56: I suspect that equation (15) took some effort to derive; could
some details go into an Appendix?}

\textbf{Figure 6: I like that the strain uncertainties can be shown, but is
there a reason not to plot principle strains using arrows, showing
errors by an ellipse around the tips?}

\closing{Thank you for your continued consideration,}

\end{letter}
\end{document}

### 2017-11-09 ResponseToReviewers/compile.sh

@@ -1,0 +1,2 @@
+#!/bin/bash
+pdflatex -synctex=1 -interaction=nonstopmode ResponseToReviewers.tex

### 2017-11-09 ResponseToReviewers/ResponseToReviewers.tex
did more

@@
 
-We were pleased with the positive comments from you and the reviewers, and we feel that the comments brought up valid concerns.
-We hope that we have addressed these concerns sufficiently in our revision.
-In order to fully address the comments, we have included a supplementary material document.
-We have also made some changes to the manuscript which were not specifically requested by the reviewers.
-Most of these changes were stylistic and do not change the content of the manuscript.
-One of the more substantial changes was that we merged Figures 7 and 14, which are both showing observed and predicted radial components of postseismic displacements, facilitating direct comparison of the figures.
+We were pleased with the positive comments from you and the reviewers,
+and we feel that the comments brought up valid concerns. We hope that
+we have addressed these concerns sufficiently in our revision. In
+order to fully address the comments, we have included a supplementary
+material document.  We have also made some changes to the manuscript
+which were not specifically requested by the reviewers.  Most of these
+changes were stylistic and do not change the content of the
+manuscript.  One of the more substantial changes was that we merged
+Figures 7 and 14, which are both showing observed and predicted radial
+components of postseismic displacements, facilitating direct
+comparison of the figures.
 
-We added supporting figures S3 and S4, which break down the predicted displacements of our preferred model into an elastic and viscoelastic component.
-We feel that these figures help support our discussion on lines 509 to 523, where we attribute inferences of fault slip and viscoelastic relaxation to different aspects of the observed postseismic displacements.
+We added supporting figures S3 and S4, which break down the predicted
+displacements of our preferred model into an elastic and viscoelastic
+component.  We feel that these figures help support our discussion on
+lines 509 to 523, where we attribute inferences of fault slip and
+viscoelastic relaxation to different aspects of the observed
+postseismic displacements.
 
-On line 500 to 502 of our previous version of this manuscript we make the claim that “After one year, afterslip is inferred to be deeper down on the Sierra Cucapah segment, which is describing much of the sustained near-field postseismic deformation.”  This sentence is not accurate of the preferred model we present in the manuscript.
-As can be clearly seen in the new figures S3-S4, afterslip is describing some of the later near-field deformation, although most of it is being described by viscoelastic relaxation.
-We have reworded this passage and updated the introduction and conclusion to accurately describe our interpretations of the mechanisms driving later near-field deformation.
-This does not significantly alter the main conclusions of our manuscript, because we are very clear, both in the original and revised versions, that the deformation mechanism driving later near-field deformation cannot be definitively resolved with these data.
+On line 500 to 502 of our previous version of this manuscript we make
+the claim that “After one year, afterslip is inferred to be deeper
+down on the Sierra Cucapah segment, which is describing much of the
+sustained near-field postseismic deformation.” This sentence is not
+accurate of the preferred model we present in the manuscript. As can
+be clearly seen in the new figures S3-S4, afterslip is describing some
+of the later near-field deformation, although most of it is being
+described by viscoelastic relaxation.  We have reworded this passage
+and updated the introduction and conclusion to accurately describe our
+interpretations of the mechanisms driving later near-field
+deformation.  This does not significantly alter the main conclusions
+of our manuscript, because we are very clear, both in the original and
+revised versions, that the deformation mechanism driving later
+near-field deformation cannot be definitively resolved with these
+data.
 
-Each of the reviewer comments are included below followed by our response and a description of how these comments were addressed in this revised manuscript.
+Each of the reviewer comments are included below followed by our
+response and a description of how these comments were addressed in
+this revised manuscript.
 
@@
 
+I have rewritten the methods section to contain more examples and I
+elaborated on concepts that may not be familiar to a reader that does
+not have a strong background in stochastic processes. I sincerely hope
+that my revised draft is more readable to geophysicists.
+
 \textbf{I also require that a comparison between your outlier removal model
@@
 efficiency of your method.}
+
+(See my response to the comment ``Also, the paper would be improved if
+the proposed outlier detection ...")
 
@@
 
+I have changed my description of the data so that it includes a single
+term, $\eta$, that encompasses correlated and uncorrelated noise. I
+then break $\eta$ down into correlated and uncorrelated components in
+Section 3.1 (eq. 25), where I go into detail about the noise model for
+the Pacific Northwest.
+
 \textbf{2- First, the author should give a summary of stochastic noise
@@
 
-\textbf{Now, I am confused When looking at p.8 Section 4.1, the description of
-the noise models should be given as a subsection of Section 2 p.4. I
-would also improve the literature review. Several other models have
-also been discussed (i.e. Montillet et al., 2014 uses a fractional
-Brownian motion model; ARMA, ARFIMA, GGM or Band pass noise – see Het
-et al., 2017).}
+In Section 3.1, I acknowledge that noise models for GNSS data have
+been thoroughly researched and my FOGM noise model is not a new
+concept. I fear that a full summary of stochastic noise modelling
+would be too much of a distraction to the reader. However, I have
+referred the reader to He et al 2017 for such a review TODO.
+
+We agree that the spatial correlation between paramaters in the
+deterministic model can be incorporated into our noise model. We
+acknowledge this possibility in Section when we say ``the secular
+velocities, $b_i$, are spatially correlated and we could invoke a
+tectonic model to form a prior on $b_i$. However, in our application
+to the Pacific Northwest, we will be using displacement time series
+which are long enough to sufficiently constrain $b_i$ for each
+station, avoiding the need to incorporate a prior. Likewise, seasonal
+deformation is spatially correlated (Dong et al. 2002; Langbein 2008),
+and it may be worth exploring and exploiting such a correlation in a
+future study."
+
+\textbf{Now, I am confused When looking at p.8 Section 4.1, the
+description of the noise models should be given as a subsection of
+Section 2 p.4. I would also improve the literature review. Several
+other models have also been discussed (i.e. Montillet et al., 2014
+uses a fractional Brownian motion model; ARMA, ARFIMA, GGM or Band
+pass noise – see Het et al., 2017).}
+
+The paper is laid out so that Section 2 contains a general description
+of the method, and Section 3 contains details that are specific to
+each application. The choice of noise model is specific to the
+application, and so we feel that it should remain in Section 3. To
+help clarify any confusion caused by this layout, we say in Section 2,
+``The appropriate noise model may vary depending on the application,
+and we hold off on specifying the covariance matrix, $C_{\eta_i}$,
+until Section 3.1".
+
+We feel that a full review on stochastic noise models would be a
+distraction to the reader, but we refer the reader to He et al 2017
+for such as review.
 
@@
 
+In Section 2.2, we have expanded our discussion of previous efforts to
+detect outliers in GNSS data.
+
 \textbf{2- Also, the paper would be improved if the proposed outlier detection
@@
 13 (last paragraph below equation 18).}
+
+My outlier detection algorithm is quite similar to what is used in Bos
+et al 2013 (and many other papers), in that the outliers are
+determined based on the residuals for a best fitting model. The only
+difference is that I am fitting the data with a stochastic model
+rather than a parametric model consisting of a linear trend and
+sinusoids. When using a parametric model, I found that deformation
+from slow slip events would sometimes be erroneously identified as an
+outlier. The stochastic model used in my outlier detection method is
+more flexible and able to describe this transient deformation, and so
+only high frequency anomalous observations get identified as outliers.
+I have included Figure A1 to demonstrate this.
+
+Of course, a parametric model that includes, for example, B-splines
+would be just as effective at detecting outliers without removing
+geophysical signal. So I am not suggesting that my outlier detection
+method is a significant advancement or even a novel method. Instead, I
+am only describing my outlier detection method for the sake of
+completeness. The previous version of my paper seems to have put too
+much emphasis on the outlier detection method, when the main focus
+should remain on the method for estimating transient strain rates. To
+help keep my paper focused, I placed the description of the outlier
+detection method in the appendix.
 
@@
 
+In the previous draft, I was discussing two different priors: a prior
+for the outlier detection algorithm, and a prior used in computing
+transient strain rates. I acknowledge that this was convoluted. In the
+revised draft, all the details of the outlier detection algorithm are
+confined to the appendix. I hope this clarifies any confusion.
+
 \textbf{Minor issues:}
@@
 \textbf{P2 line 30 : “fidelity” replace with “reliability”.}
+
+Done
 
@@
 
+Done
+
 \textbf{P5 line 32 : “formal data uncertainty”: can you define it?}
 
+By "formal data uncertainties" I meant "the derived uncertainties that
+accompany the GNSS displacement solutions". The text has been
+clarified.
+
 \textbf{P 6 : Need reference to show Equation 15.}
+
+I have shown a few more steps in the derivation of eq. 15.
 
@@
 \textbf{P.8 line 20 delete www.unavco.org (only in Acknowledgment)}
+
+Done
 

### 2017-11-09 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 stations, such as the Plate Boundary Observatory (PBO), make it
-possible to identify transient strain with high fidelity. Developing
+possible to identify transient strain with high reliability. Developing
 and improving upon methods for deriving secular and transient strain
@@
 gradients are spatially uniform is not appropriate when subnetworks
-span too large of an area. To help overcome this deficiency,
+span a very large area. To help overcome this deficiency,
 \citet{Shen1996,Shen2015} used an inverse distance weighting scheme,
@@
 Any errors in the assumed physical model could result in biased strain
-estimates and underestimated formal uncertainties.
+estimates and underestimated uncertainties.
 

### 2017-11-09 AnnotatedRevision/compile.sh

@@ -1,0 +1,14 @@
+#!/bin/bash
+#pdflatex -synctex=1 -interaction=nonstopmode dissertation.tex
+#bibtex ch1/ch1.aux
+#bibtex ch2/ch2.aux
+#bibtex ch3/ch3.aux
+#bibtex ch4/ch4.aux
+#bibtex ch5/ch5.aux
+#bibtex ch6/ch6.aux
+#pdflatex -synctex=1 -interaction=nonstopmode dissertation.tex
+
+pdflatex -synctex=1 -interaction=nonstopmode diff.tex
+bibtex diff.aux
+pdflatex -synctex=1 -interaction=nonstopmode diff.tex
+pdflatex -synctex=1 -interaction=nonstopmode diff.tex

### 2017-11-09 AnnotatedRevision/figures/prior_demo/make-plot.py

@@ -1,0 +1,79 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import rbf
+from mpl_toolkits.mplot3d import Axes3D
+np.random.seed(7)
+
+
+gp_x = rbf.gauss.gpse((0.0,1.0,100.0))
+gp_t = rbf.gauss.gpse((0.0,1.0,10.0))
+
+# combined
+gp = rbf.gauss.GaussianProcess(
+       lambda x: np.zeros(len(x)), 
+       lambda x1,x2: gp_x.covariance(x1[:,[0]],x2[:,[0]])*
+                     gp_t.covariance(x1[:,[1]],x2[:,[1]]),
+       dim=2)
+
+
+###### PLOT COVARIANCE
+#####################################################################
+Nx = 20
+Nt = 20
+figsize = (7.0,3.0)
+
+x = np.linspace(0.0,300.0,Nx)
+t = np.linspace(0.0,30.0,Nt)
+xg,tg = np.meshgrid(x,t)
+p = np.array([xg.flatten(),tg.flatten()]).T
+
+# make cov array
+cov = gp.covariance([[0.0,0.0]],p)
+cov = cov.reshape((Nx,Nt))
+
+fig = plt.figure(figsize=figsize)
+ax = fig.add_subplot(1,2,1,projection='3d')
+ax.view_init(35.0,35.0)
+
+p = ax.plot_surface(xg,tg,cov,cmap='viridis',edgecolor='k',linewidth=0.5)
+
+# set labels
+ax.set_ylabel(r'$\mathregular{|t - t\prime|}$ [days]',fontsize=10)
+ax.set_xlabel(r'$\mathregular{||\vec{x} - \vec{x}\prime||_2}$ [km]',fontsize=10)
+ax.set_zlabel('covariance [mm$^2$]',fontsize=10)
+ax.set_xticks([0,75,150,225,300])
+ax.tick_params(pad=0)
+ax.tick_params(labelsize=10)
+ax.text2D(-0.05, 0.9, "A", fontsize=16, transform=ax.transAxes)
+
+###### PLOT SAMPLE
+#####################################################################
+Nx = 30
+Nt = 30
+figsize = (5,4)
+
+x = np.linspace(0.0,500.0,Nx)
+t = np.linspace(0.0,50.0,Nt)
+xg,tg = np.meshgrid(x,t)
[…]

### 2017-11-09 AnnotatedRevision/figures/strain_map/make-plot.py

@@ -1,0 +1,152 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import pygeons
+from mpl_toolkits.basemap import Basemap
+import h5py
+from pygeons.plot.strain_glyph import strain_glyph
+from matplotlib.text import Text
+from matplotlib.patches import Rectangle
+
+# PARAMETERS
+strain_scale = 25000.0
+plot_date = '2016-01-01'
+
+def get_meridians_and_parallels(bm,ticks):
+  ''' 
+  returns the meridians and parallels that should be plotted.
+  '''
+  diff_lon = (bm.urcrnrlon-bm.llcrnrlon)
+  round_digit = int(np.ceil(np.log10(ticks/diff_lon)))
+  dlon = np.round(diff_lon/ticks,round_digit)
+
+  diff_lat = (bm.urcrnrlat-bm.llcrnrlat)
+  round_digit = int(np.ceil(np.log10(ticks/diff_lat)))
+  dlat = np.round(diff_lat/ticks,round_digit)
+
+  meridians = np.arange(np.floor(bm.llcrnrlon),
+                        np.ceil(bm.urcrnrlon),dlon)
+  parallels = np.arange(np.floor(bm.llcrnrlat),
+                        np.ceil(bm.urcrnrlat),dlat)
+
+  return meridians,parallels
+
+def setup_ax(bm,ax):
+  # function which prints out the coordinates on the bottom left 
+  # corner of the figure
+  def coord_string(x,y):
+    str = 'x : %g  y : %g  ' % (x,y)
+    str += '(lon : %g E  lat : %g N)' % bm(x,y,inverse=True)
+    return str
+
+  ax.format_coord = coord_string
+  bm.drawcountries(ax=ax,zorder=1)
+  bm.drawstates(ax=ax,zorder=1)
+  bm.drawcoastlines(ax=ax,zorder=1)
+  bm.drawmapboundary(ax=ax,fill_color=(0.8,0.8,0.8),zorder=0)
+  bm.fillcontinents(ax=ax,color=(1.0,1.0,1.0),lake_color=(0.8,0.8,0.8),zorder=0)
+  mer,par = get_meridians_and_parallels(bm,3)
+  bm.drawmeridians(mer,
+                   labels=[0,0,0,1],dashes=[2,2],fontsize=10,
+                   ax=ax,zorder=1,color=(0.3,0.3,0.3,1.0))
+  bm.drawparallels(par,
[…]

### 2017-11-10 ResponseToReviewers/ResponseToReviewers.tex
did more

--- 2017-11-09 ResponseToReviewers/ResponseToReviewers.tex
@@
 
+I have taken several steps to make my revised draft more accessible to
+someone who does not have a strong background in stochastic processes:
+I rewritten the methods section to be clearer, contain more examples,
+and elaborate on potentially unfamiliar concepts. I have also added
+more references to the methods section to make it easier for the
+reader to find additional information. I sincerely hope that my
+revised draft is more readable to geophysicists.
+
 \textbf{I also require that a comparison between your outlier removal model
@@
 
+I have changed my description of the data (eq. 3 in the revised draft)
+so that it includes a single term, $\eta$, that encompasses correlated
+and uncorrelated noise. I then break $\eta$ down into correlated and
+uncorrelated components in Section 3.1 (eq. 25). The covariance matrix
+for $\eta$ in eq. 26 is more consistent with the literature.
+
 \textbf{2- First, the author should give a summary of stochastic noise
@@
 
-\textbf{Now, I am confused When looking at p.8 Section 4.1, the description of
-the noise models should be given as a subsection of Section 2 p.4. I
-would also improve the literature review. Several other models have
-also been discussed (i.e. Montillet et al., 2014 uses a fractional
-Brownian motion model; ARMA, ARFIMA, GGM or Band pass noise – see Het
-et al., 2017).}
+I believe the reviewer is refering to page 4-5 of the original draft,
+where I first introduce $\eta$. I do not think this is an appropriate
+place for a discussion on the noise model, or for a summary of past
+studies. The paper is laid out so that Section 2 contains a general
+description of the method for estimating transient strain rates, and
+Section 3 contains details that are specific to the particular
+application. There is no universally appropriate noise model so I
+decided to save a discussion on the noise model for Section 3. This is
+clarified in Section 2 of the revised draft when I say "The
+appropriate noise model may vary depending on the application, and we
+hold off on specifying the covariance matrix, $\mitbf{C}_{\eta_i}$,
+until Section \ref{sec:NoiseModel}."
+
+\textbf{Now, I am confused When looking at p.8 Section 4.1, the
+description of the noise models should be given as a subsection of
+Section 2 p.4. I would also improve the literature review. Several
+other models have also been discussed (i.e. Montillet et al., 2014
+uses a fractional Brownian motion model; ARMA, ARFIMA, GGM or Band
+pass noise – see Het et al., 2017).}
+
+Again, we think that the description of the noise model should remain
+in Section 3 because it is specific to the Pacific Northwest
+application.
+
+We have added a few more references in Section 3.1 of the revised
+draft TODO to emphasize that GNSS noise modeling is a well-worked
+area. We refer the reader to Bock and Melgar 2016 and He et al 2017
+for a more thorough review.
 
@@
 
+In Section 2.2 of the revised draft, I have expanded the discussion of
+previous efforts to detect outliers in GNSS data.
+
 \textbf{2- Also, the paper would be improved if the proposed outlier detection
@@
 
+Done
+
 \textbf{P2 line 30-31 “ Developing and improving upon methods for deriving
@@
 
+TODO
+
 \textbf{p.2 line 43 :”too large of an area” ... perhaps “ a very large area”}
+
+Done
 
@@
 
+By "formal data uncertainties" I meant "the derived uncertainties that
+accompany the GNSS displacement solutions". The revised text has been
+clarified.
+
 \textbf{P 6 : Need reference to show Equation 15.}
+
+I have added a few more steps in the derivation of eq. 15.
 
@@
 (Bock and Melgar, 2016, Gazeaux et al., 2013).}
+
+I have added a reference to Gazeaux et al. 2013.
 
@@
 
+We agree that we need to acknowledge the extensive amount of research
+on Cascadia SSEs. In the revised draft, we refer the reader to the
+review by Schartz and Rokosky 2007.
+
 \textbf{P.8 line 20 delete www.unavco.org (only in Acknowledgment)}
+
+Done
 
@@
 define it properly.}
+
+In Section 3.1 of the revised draft, where we elaborate on the noise
+model, we describe common mode error by saying "Another significant
+source of noise in GNSS data is common mode error (e.g., Wdowinski et
+al. 1997; Dong et al. 2006), which is noise that is highly spatially
+correlated." We believe this definition and these references are
+sufficient.
 
@@
 
+I recognize that it may be a bit redundant to say that the covariance
+function is compact and thus the covariance matrices are sparse.
+However, Wendland covariance functions are typically described as
+"compact" in the literature, and I think it is a key adjective that
+should not be dropped out. I have left the wording as is.
+
 \textbf{p.13 line 22 what is an “isotropic Gaussian process”? please add a
 definition.}
+
+An Gaussian process is isotropic if its covariance function,
+$C(x,x')$, can be written as a function of $||x - x'||_2^2$. An
+isotropic Gaussian process is also stationary, meaning that its
+statistical properties are invariant to translations. We have replace
+"isotropic" with "stationary" in the text because "stationary" is a
+more appropriate, and perhaps more familiar, term.
 
@@
 
+My personal preference is to leave the wording as is.
+
 \textbf{p.20 line 33 “ computational burden” I think that refers to p.20 line
@@
 time to process longer and longer time series?}
+
+I agree that numerical values would be interesting; however, I am
+reluctant to quantify the computational cost of evaluation transient
+strain rates. There are many factors that influence the computational
+cost (whether sparse matrices are used, the type of sparse matrices
+that are used, the linear solver algorithm, etc.) and I am not
+convinced that my implementation is the most efficient. So I do not
+give details on the computation time, but I note in the discussion
+that this is an area for further research.
 

### 2017-11-10 ResponseToReviewers/ResponseToReviewers.tex
did more

@@
 appropriate noise model may vary depending on the application, and we
-hold off on specifying the covariance matrix, $\mitbf{C}_{\eta_i}$,
-until Section \ref{sec:NoiseModel}."
+hold off on specifying the covariance matrix, $C_{\eta_i}$,
+until Section 3.1."
 

### 2017-11-10 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 Northwest, and we are specifically interested in identifying transient
-strain resulting from SSEs \citep[e.g.,][]{Dragert2001}. Before
+strain resulting from SSEs. SSEs in the Pacific Northwest were first
+discovered by \citep[e.g.,][]{Dragert2001} and we refer the reader to
+\citet{Schwartz2007} for a review of the subsequent research. Before
 estimating transient strain rates, we establish a noise model for GNSS
@@
 studied over the past two decades \citep[e.g.,][]{Zhang1997, Mao1999,
-Williams2004, Langbein2008}. In these studies, temporally correlated
-noise tends to be described with some combination of Brownian motion
-(also known as random walk noise or a Weiner process), a first-order
-Gauss-Markov (FOGM) process, and/or flicker noise. There is some
-physical justification for using Brownian motion as a noise model
-because it accurately describes the power spectrum of motion resulting
-from instability in a geodetic monument \citep[e.g.,][]{Wyatt1982,
-Wyatt1989}. In particular, the power spectrums for Brownian motion and
-monument motion both decay proportionally to $f^{-2}$, where $f$ is
-frequency. However, Brownian motion necessarily contains a reference
-time at which the process begins. Since there is no notion of when
-noise ``begins" in GNSS data, we do not find Brownian motion to be an
-appropriate noise model. On the other hand, a FOGM process has no
-reference time (i.e., it is stationary), and its power spectrum,
-\begin{equation}\label{eq:FOGMPS}
-S(f) = \frac{\beta^2}{(2 \pi f)^2 + \alpha^2},
-\end{equation}
-decays proportionally to $f^{-2}$ above the cutoff frequency $\alpha /
-(2 \pi)$. We then choose $z_i$ to be a spatially
-uncorrelated FOGM process, which has the covariance function
+Williams2004, Langbein2008} (See \citet{Bock2016} or \citet{He2017}
+for a review of GNSS noise modelling). In these studies, temporally
+correlated noise tends to be described with some combination of
+Brownian motion (also known as random walk noise or a Weiner process),
+a first-order Gauss-Markov (FOGM) process, and/or flicker noise. There
+is some physical justification for using Brownian motion as a noise
+model because it accurately describes the power spectrum of motion
+resulting from instability in a geodetic monument
+\citep[e.g.,][]{Wyatt1982, Wyatt1989}. In particular, the power
+spectrums for Brownian motion and monument motion both decay
+proportionally to $f^{-2}$, where $f$ is frequency. However, Brownian
+motion necessarily contains a reference time at which the process
+begins. Since there is no notion of when noise ``begins" in GNSS data,
+we do not find Brownian motion to be an appropriate noise model. On
+the other hand, a FOGM process has no reference time (i.e., it is
+stationary), and its power spectrum, \begin{equation}\label{eq:FOGMPS}
+S(f) = \frac{\beta^2}{(2 \pi f)^2 + \alpha^2}, \end{equation} decays
+proportionally to $f^{-2}$ above the cutoff frequency $\alpha / (2
+\pi)$. We then choose $z_i$ to be a spatially uncorrelated FOGM
+process, which has the covariance function
 \begin{equation}\label{eq:FOGM}
-C_{z_i}\left((\pos,t),(\pos',t')\right) = 
-\frac{\beta^2}{2\alpha}
-\exp\left(-\alpha|t - t'|\right)\delta_{\pos,\pos'},
-\end{equation}
+C_{z_i}\left((\pos,t),(\pos',t')\right) = \frac{\beta^2}{2\alpha}
+\exp\left(-\alpha|t - t'|\right)\delta_{\pos,\pos'}, \end{equation}
 where $\delta_{\pos,\pos'}$ is 1 if $\pos = \pos'$ and 0 otherwise.
@@
 \end{equation}
+
 Note that $T_i$ includes the parameter $\phi$, which serves to scale
@@
 Wendland covariance functions have compact support and hence their
-corresponding covariance matrices are sparse. In our analysis, we
-exploit this sparsity with the CHOLMOD software package
-\citep{Chen2008}. Wendland covariance functions can be constructed
-such that they are positive definite in $\mathbb{R}^d$ and their
-corresponding Gaussian process can be differentiated $k$ times. We use
-$d=1$, because we are describing the temporal covariance of $u_i$, and
-we use $k=2$, giving samples of $u_i$ continuous velocities and
-accelerations. The corresponding Wendland covariance function is
+corresponding covariance matrices are sparse. We exploit this sparsity
+with the CHOLMOD software package \citep{Chen2008}. Wendland
+covariance functions can be constructed such that they are positive
+definite in $\mathbb{R}^d$ and their corresponding Gaussian process
+can be differentiated $k$ times. We use $d=1$, because we are
+describing the temporal covariance of $u_i$, and we use $k=2$, giving
+samples of $u_i$ continuous velocities and accelerations. The
+corresponding Wendland covariance function is
+
 \begin{equation}\label{eq:Wendland} 
@@
 probability of drawing $\data_i^*$ from $\data_i$, but now we are
-including $u_i(\points)$ in $\data_i$. and we are optimizing the
+including $u_i(\points)$ in $\data_i$, and we are optimizing the
 parameters for $C_{u_i}$ rather than $\mitbf{C}_{\eta_i}$. We divide

### 2017-11-10 ResponseToReviewers/ResponseToReviewers.tex
did more

@@
 I believe the reviewer is refering to page 4-5 of the original draft,
-where I first introduce $\eta$. I do not think this is an appropriate
-place for a discussion on the noise model, or for a summary of past
-studies. The paper is laid out so that Section 2 contains a general
-description of the method for estimating transient strain rates, and
-Section 3 contains details that are specific to the particular
-application. There is no universally appropriate noise model so I
-decided to save a discussion on the noise model for Section 3. This is
-clarified in Section 2 of the revised draft when I say "The
-appropriate noise model may vary depending on the application, and we
-hold off on specifying the covariance matrix, $C_{\eta_i}$,
-until Section 3.1."
+where I first introduce $\eta$. I agree that a more thorough
+discussion on the noise model is necessary, but I do not think this is
+an appropriate place in the text for it. The paper is laid out so that
+Section 2 contains a general description of the method for estimating
+transient strain rates, and Section 3 contains details that are
+specific to the particular application. There is no universally
+appropriate noise model, so I decided to save a discussion on the
+noise model for Section 3. This is clarified in Section 2 of the
+revised draft when I say "The appropriate noise model may vary
+depending on the application, and we hold off on specifying the
+covariance matrix, $C_{\eta_i}$, until Section 3.1."
 
@@
 area. We refer the reader to Bock and Melgar 2016 and He et al 2017
-for a more thorough review.
+for a more complete review.
 
@@
 
-In Section 2.2 of the revised draft, I have expanded the discussion of
-previous efforts to detect outliers in GNSS data.
+In Section 2.2 of the revised draft, I have expanded my discussion on
+how previous studies have treated outliers in GNSS data.
 
@@
 
+Also, I have rewritten most of the results section to be more
+structured and hopefully have a clearer focus.
+
 \textbf{Minor issues:}
@@
 
-TODO
+The introduction has been revised and this sentence has been removed.
 
@@
 
-I agree that numerical values would be interesting; however, I am
-reluctant to quantify the computational cost of evaluation transient
-strain rates. There are many factors that influence the computational
-cost (whether sparse matrices are used, the type of sparse matrices
-that are used, the linear solver algorithm, etc.) and I am not
-convinced that my implementation is the most efficient. So I do not
-give details on the computation time, but I note in the discussion
-that this is an area for further research.
+I agree that numerical values would be useful; however, I am reluctant
+to quantify the computational cost of evaluation transient strain
+rates. There are many factors that influence the computational cost
+(whether sparse matrices are used, the type of sparse matrices that
+are used, the linear solver algorithm, etc.) and I am not convinced
+that my implementation is the most efficient. So I do not give details
+on the computation time, but I note in the discussion that this is an
+area for further research.
 

### 2017-11-10 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 
-% Why we care about determining strain
+% Why we care about determining secular and transient strain
 
@@
 accumulating gives insight into where we can expect stored elastic
-energy to be released seismically. Consequently, secular crustal
-strain rates estimated from GNSS data have been used to constrain
-seismic hazard models such as UCERF3 \citep{Field2014}. Transient
-crustal strain, which is caused by geophysical phenomena such as slow
-slip events (SSEs) or postseismic deformation, is also relevant for
-assessing seismic hazard. While transient strain itself is not
-damaging, there is a risk that it can trigger major earthquakes
-\citep{Roeloffs2006,Freed2001}. Dense networks of continuous GNSS
-stations, such as the Plate Boundary Observatory (PBO), make it
-possible to identify transient strain with high reliability. Developing
-and improving upon methods for deriving secular and transient strain
-from GNSS data is an active area of research.
-
-% Existing methods for determining secular strain from GPS
-
-Most methods for estimating strain rates from GNSS data assume some
-parametric form for the deformation signal. The simplest method for
-estimating secular strain rates assumes that GNSS derived velocities
-can be described with a first-order polynomial (i.e, the deformation
-gradients are constant) over some subnetwork of the GNSS stations
-\citep[e.g.,][]{Feigl1990,Murray2000}. The components of the strain
-rate tensor for each subnetwork are then determined from the least
-squares fit to the observations. The assumption that deformation
-gradients are spatially uniform is not appropriate when subnetworks
-span a very large area. To help overcome this deficiency,
-\citet{Shen1996,Shen2015} used an inverse distance weighting scheme,
-in which the estimated strain rate at a point is primarily controlled
-by observations at nearby stations. The method of
-\citet{Shen1996,Shen2015} can be viewed as a form of local least
-squares regression with a first-order polynomial \citep[e.g.,][sec.
-6]{Hastie2009}. Other methods for estimating secular strain rates have
-parameterized GNSS derived velocities with bi-cubic splines
-\citep{Beavan2001}, spherical wavelets \citep{Tape2009}, and
-elastostatic Green's functions \citep{Sandwell2016}. The type of basis
-functions and the number of degrees of freedom for a parameterization,
-which are often chosen subjectively, have a strong influence on the
-strain solution. If there are too few degrees of freedom in the
-parameterization, then estimated strain rates will be biased and the
-uncertainties will be underestimated. On the other hand, if there are
-too many degrees of freedom, then there will not be any coherent
-features in the estimated strain rates. The methods described by
-\citet{Beavan2001} and \citet{Tape2009} also require the user to
-specify penalty parameters that control a similar trade-off between
-bias and variance in the solution. One could parameterize deformation
-with a physically motivated model of interseismic deformation
-\citep[e.g.,][]{Meade2005,McCaffrey2007}. In such models the
-lithospheric rheology and fault geometries are assumed to be known.
-Any errors in the assumed physical model could result in biased strain
-estimates and underestimated uncertainties.
+energy to be released seismically. Estimates of secular, or long-term,
+strain rates have been used to constrain seismic hazard models such as
+UCERF3 \citep{Field2014}. Transient crustal strain, which is caused by
+geophysical processes such as slow slip events (SSEs) or postseismic
+deformation, is also relevant for assessing seismic hazard. While
+transient strain itself is not damaging, there is a risk that it can
+trigger major earthquakes \citep{Roeloffs2006,Freed2001}.
+
+% How secular strain rates are estimated from GNSS
+
+Displacement data from dense GNSS networks, such as the Plate Boundary
+Observatory (PBO), can be used to infer secular and transient strain.
+Numerous methods for converting GNSS displacement data to strain have
+been proposed, and most involve assuming some parametric form for the
+deformation signal. The simplest method for estimating secular strain
+rates assumes that GNSS derived velocities can be described with a
+first-order polynomial (i.e, the deformation gradients are constant)
+over some subnetwork of the GNSS stations \citep[e.g.,][]{Feigl1990,
+Murray2000}. The components of the strain rate tensor for each
+subnetwork are then determined from the least squares fit to the
+observations. The assumption that deformation gradients are spatially
+uniform is not appropriate when subnetworks span a very large area. To
+help overcome this deficiency, \citet{Shen1996,Shen2015} used an
+inverse distance weighting scheme, in which the estimated strain rate
+at a point is primarily controlled by observations at nearby stations.
+The method of \citet{Shen1996,Shen2015} can be viewed as a form of
+local least squares regression with a first-order polynomial. Other
+methods for estimating secular strain rates have parameterized GNSS
+derived velocities with bi-cubic splines \citep{Beavan2001}, spherical
+wavelets \citep{Tape2009}, and elastostatic Green's functions
+\citep{Sandwell2016}. The type of basis functions and the number of
+degrees of freedom for a parameterization, which are often chosen
+subjectively, have a strong influence on the strain solution. If there
+are too few degrees of freedom in the parameterization, then estimated
+strain rates will be biased and the uncertainties will be
+underestimated. On the other hand, if there are too many degrees of
+freedom, then there will not be any coherent features in the estimated
+strain rates. The methods described by \citet{Beavan2001} and
+\citet{Tape2009} also require the user to specify penalty parameters
+that control a similar trade-off between bias and variance in the
+solution. One could parameterize deformation with a physically
+motivated model of interseismic deformation \citep[e.g.,][]{Meade2005,
+McCaffrey2007}. In such models the lithospheric rheology and fault
+geometries are assumed to be known. Any errors in the assumed physical
+model could result in biased strain estimates and underestimated
+uncertainties.
 
@@
 The temporally correlated noise in GNSS data has been thoroughly
-studied over the past two decades \citep[e.g.,][]{Zhang1997, Mao1999,
-Williams2004, Langbein2008} (See \citet{Bock2016} or \citet{He2017}
-for a review of GNSS noise modelling). In these studies, temporally
-correlated noise tends to be described with some combination of
-Brownian motion (also known as random walk noise or a Weiner process),
-a first-order Gauss-Markov (FOGM) process, and/or flicker noise. There
-is some physical justification for using Brownian motion as a noise
-model because it accurately describes the power spectrum of motion
-resulting from instability in a geodetic monument
-\citep[e.g.,][]{Wyatt1982, Wyatt1989}. In particular, the power
-spectrums for Brownian motion and monument motion both decay
-proportionally to $f^{-2}$, where $f$ is frequency. However, Brownian
-motion necessarily contains a reference time at which the process
-begins. Since there is no notion of when noise ``begins" in GNSS data,
-we do not find Brownian motion to be an appropriate noise model. On
-the other hand, a FOGM process has no reference time (i.e., it is
-stationary), and its power spectrum, \begin{equation}\label{eq:FOGMPS}
-S(f) = \frac{\beta^2}{(2 \pi f)^2 + \alpha^2}, \end{equation} decays
-proportionally to $f^{-2}$ above the cutoff frequency $\alpha / (2
-\pi)$. We then choose $z_i$ to be a spatially uncorrelated FOGM
-process, which has the covariance function
+studied over the past two decades (See \citet{Bock2016} or
+\citet{He2017} for a review). Most studies found that the temporally
+correlated noise can be described with Brownian motion (also known as
+random walk noise or a Weiner process) or flicker noise
+\citep[e.g.,][]{Zhang1997, Mao1999, Williams2004, Langbein2008,
+Murray2017}. Less frequently used noise models include a band-pass
+process \citep[e.g.,][]{Langbein2004} and a first-order Gauss-Markov
+(FOGM) process \citep[e.g.,][]{Langbein2004,Dmitrieva2015}. There is
+some physical justification for using Brownian motion as a noise model
+because it accurately describes the power spectrum of motion resulting
+from instability in a geodetic monument \citep[e.g.,][]{Wyatt1982,
+Wyatt1989}. In particular, the power spectrums for Brownian motion and
+monument motion both decay proportionally to $f^{-2}$, where $f$ is
+frequency. However, Brownian motion necessarily contains a reference
+time at which the process begins. Since there is no notion of when
+noise ``begins" in GNSS data, we do not find Brownian motion to be an
+appropriate noise model. On the other hand, a FOGM process has no
+reference time (i.e., it is stationary), and its power spectrum,
+\begin{equation}\label{eq:FOGMPS} 
+S(f) = \frac{\beta^2}{(2 \pi f)^2 + \alpha^2}, 
+\end{equation} 
+decays proportionally to $f^{-2}$ above the cutoff frequency $\alpha /
+(2 \pi)$. We choose $z_i$ to be a spatially uncorrelated FOGM process.
+The covariance function for $z_i$ is then
 \begin{equation}\label{eq:FOGM}
 C_{z_i}\left((\pos,t),(\pos',t')\right) = \frac{\beta^2}{2\alpha} 
-\exp\left(-\alpha|t - t'|\right)\delta_{\pos,\pos'}, \end{equation}
-where $\delta_{\pos,\pos'}$ is 1 if $\pos = \pos'$ and 0 otherwise.
-See \citet[sec. B2]{Rasmussen2006} for more information on the FOGM
-process.
+                                          \exp\left(-\alpha|t - t'|\right)\delta_{\pos,\pos'}, 
+\end{equation}
+where $\delta_{\pos,\pos'}$ is 1 if $\pos = \pos'$ and 0 otherwise
+\citep[sec. B2]{Rasmussen2006}.
 

### 2017-11-10 ResponseToReviewers/ResponseToReviewers.tex
did more

@@
 \opening{Dear Editor,}
+
+% NOTE THE FOLLOWING ERRORS
+
+% STRAIN COVARIANCE WAS INCORRECT
+
+% THE NOISE MODEL WAS INCORRECT BECAUSE I USED A DIRAC RATHER THAN A
+% KRONECKER DELTA.
 
@@
 We have added a few more references in Section 3.1 of the revised
-draft TODO to emphasize that GNSS noise modeling is a well-worked
-area. We refer the reader to Bock and Melgar 2016 and He et al 2017
-for a more complete review.
+draft to emphasize that GNSS noise modeling is a well-worked area. We
+refer the reader to Bock and Melgar 2016 and He et al 2017 for a more
+complete review.
 
@@
 
-\textbf{Page 3: Last line: min(t,t’) needs as scaling factor to make units for
-covariance e.g., m$^2$/yr}
+Our approach is indeed computationally intensive and approximation
+methods for GPR do exist. In the discussion section for the revised
+draft, we suggest exploring these approximation methods in future
+research.
+
+\textbf{Page 3: Last line: min(t,t’) needs as scaling factor to make
+units for covariance e.g., m$^2$/yr}
+
+We added the scale factor $\phi$ to the covariance function for
+Brownian motion.
 
@@
 
+This is a good point that we neglected to mention in the previous
+draft. We have added a paragraph to Section 2 that explains the
+reasoning and implications for this assumption. The paragraph starts
+with "It should be noted that we have ignored any covariances ...".
+
 \textbf{Page 4/5 Equation (3): Is it worth discussing at this point whether
@@
 
+TODO
+
 \textbf{Page 6: Equation 10: Maybe it is discussed later in the paper but it
@@
 performed with a block 0 in the matrix?}
+
+The matrix being inverted in eq. 10 and 11 of the revised draft is
+still invertible with the block 0, provided that the columns of
+$\mitbf{G}$ are linearly dependent. This is explained in the revised
+draft.
 
@@
 
-\textbf{Page 8: Partly addresses issue raised about eqn 3.  It probably should
-be noted that wether conditions east of -121 deg longitude are
-different to the coast so there could be possible problems with this
-noise model.}
+The posterior displacements, $\hat{u}$, are actually a spatially and
+temporally continuous stochastic process. We can evaluated $\hat{u}$
+at any position and time that we may be interested in. We are also
+able to compute the spatial and temporal derivatives analytically. In
+Section 2 of the revised draft, we explain how the derivatives are
+computed in greater detail. We hope this clarifies any confusion.
+
+\textbf{Page 8: Partly addresses issue raised about eqn 3.  It
+probably should be noted that wether conditions east of -121 deg
+longitude are different to the coast so there could be possible
+problems with this noise model.}
+
+This is also a good point that we neglected to mention in the first
+draft. In Section 3.1 of the revised draft, we acknowledge the
+potentially dubious assumption that noise east of -121 is
+representative of the entire region. We added the disclaimer saying
+"We assume the noise at these inland stations is representative of the
+noise at all the stations considered in this study, which is probably
+a poor assumption since the inland stations are subject to distincly
+different climatic conditions".
 

### 2017-11-10 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 which has the mean function $\E{B(t)} = 0$ and the covariance function
-$\Cov{B(t),B(t')} = \min(t,t')$, where $t,t' \ge 0$. We let each
-component of $\vec{u}$ have zero mean and a generic covariance
-function $\Cov{u_i(p),u_i(p')} = C_{u_i}(p,p')$. Using a more concise
-notation, we write our prior on each component of $\vec{u}$ as $u_i
-\sim \mathcal{GP}\left(0,C_{u_i}\right)$.
+$\Cov{B(t),B(t')} = \phi^2 \min(t,t')$, where $t,t' \ge 0$ and $\phi$
+is a scale factor. We let each component of $\vec{u}$ have zero
+mean and a generic covariance function $\Cov{u_i(p),u_i(p')} =
+C_{u_i}(p,p')$. Using a more concise notation, we write our prior on
+each component of $\vec{u}$ as $u_i \sim
+\mathcal{GP}\left(0,C_{u_i}\right)$.
 
@@
 inferred \citep[e.g.,][]{Dragert2001, Wech2009, Schmidt2010}. Thus,
-this pattern of strain is to be expected. Over decadal timescales,
+this pattern of strain is to be expected. Over decadal time-scales,
 there is trench perpendicular compression throughout this study region

### 2017-11-10 ResponseToReviewers/ResponseToReviewers.tex
did more shit

@@
 % KRONECKER DELTA.
+
+% WE ADDED MORE POSTERIOR DISPLACEMENT FIGURES
 
@@
 someone who does not have a strong background in stochastic processes:
-I rewritten the methods section to be clearer, contain more examples,
-and elaborate on potentially unfamiliar concepts. I have also added
-more references to the methods section to make it easier for the
-reader to find additional information. I sincerely hope that my
-revised draft is more readable to geophysicists.
+I rewritten the methods section to be clearer, avoid esoteric jargon,
+contain more examples, and elaborate on potentially unfamiliar
+concepts. I have also added more references to the methods section to
+make it easier for the reader to find additional information. I
+sincerely hope that my revised draft is more readable to
+geophysicists.
 
@@
 
+If we have absolutely no prior knowledge about u or eta then it would
+not be possible to distinguish transient displacements from noise. I
+think this should go without saying.
+
+In this paper, our "prior knowledge" is that inland stations are
+unlikely to record SSE displacements, which allows us to isolate and
+characterize the statistical properties of eta. In Section 3.1, we
+only explored the temporal covariance of eta, but we could have also
+explored the spatial covariance of eta and potentially identified
+noise from orbit modeling errors. If we did find spatially correlated
+noise in eta, then our estimates of u would not be invalidated;
+rather, the estimates of u would just have larger uncertainties.
+However, we did not look for spatially correlated noise in this paper,
+and it is certainly possible that an unknown noise process could have
+propagated into our estimates of u.
+
+While the reviewer makes a valid point, I think it should go without
+saying that the an inappropriate model for $u$ and $\eta$ would
+invalidate our results for transient strain rates.
+
+it is unclear how we should
+update the text. In section 2 of the text, the statistical properties
+of $u$ and $\eta$ are assumed known, although not yet specified.
+
+It
+should go without saying that an inappropriate model for $u$ or $\eta$
+would invalidate our estimates of transient strain rates.
+
+ We make it clear in the text that we do that
+
+If there are unknown orbit modeling errors, for example, then that noise could conceivably be
+
+Once eta is
+characterized, we are able to characterize u.
+
+ We then characterize the eta
+
+If a spatially and temporally correlated noise exists and we have incorporated it into
+
 TODO

### 2017-11-11 ResponseToReviewers/ResponseToReviewers.tex
did more

--- 2017-11-10 ResponseToReviewers/ResponseToReviewers.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \textbf{Reviewer \#2:}\newline
@@
 
-TODO
+At this point in the text, it is assumed that the statistical
+properties of u and eta are known, although not yet specified. So I do
+not think this would be an appropriate place to discuss a scenario in
+which we do not know u or eta. I also think that it should go with
+saying that the validity of our results are contingent on having a
+good model for u and eta.
+
+In Section 3.1 and 3.2, we discuss how we choose a model for u and
+eta. The key piece of "prior knowledge" that allows us to discern u
+from eta is that u is negligible for inland stations. This allowed us
+to islolate and constrain eta first, and then we constrained u. If we
+allowed eta to be spatially correlated (which we did not), then we
+would have conceivably been able to describe the spatially and
+temporally correlated noise from, for example, orbit modeling errors.
 
@@
 still invertible with the block 0, provided that the columns of
-$\mitbf{G}$ are linearly dependent. This is explained in the revised
+$G$ are linearly dependent. This is explained in the revised
 draft.
 
-\textbf{Page 6: Equation 12: Maybe some additional explanation is needed here.
-How are these partial derivatives formed.  My understanding the u
-estimates at this point correspond one-to-one with the positions and
-times of the GNSS position determinations.  Did I miss something?}
+\textbf{Page 6: Equation 12: Maybe some additional explanation is
+needed here. How are these partial derivatives formed.  My
+understanding the u estimates at this point correspond one-to-one with
+the positions and times of the GNSS position determinations.  Did I
+miss something?}
 
@@
 temporally continuous stochastic process. We can evaluated $\hat{u}$
-at any position and time that we may be interested in. We are also
-able to compute the spatial and temporal derivatives analytically. In
-Section 2 of the revised draft, we explain how the derivatives are
-computed in greater detail. We hope this clarifies any confusion.
+at any position and time that we may be interested in (not just the
+position and times of the data). We are also able to compute the
+spatial and temporal derivatives analytically. In Section 2 of the
+revised draft, we explain how the derivatives are computed in greater
+detail. We hope this clarifies any confusion.
 
@@
 
-\textbf{Editor’s Review of GJI MS 17.0593 ‘‘Revealing transient strain...’’ by
-Hines and Hetland This is an novel algorithm applied to a problem of
-great current interest. Based on the reviewer’s comments and my own
-reading, I am recommending moderate revision. The authors’ response
-should include a description of their changes, and also a version of
-the paper on which these can be seen (just as highlights in the text).}
+\textbf{Editor’s Review of GJI MS 17.0593 ‘‘Revealing transient
+strain...’’ by Hines and Hetland This is an novel algorithm applied to
+a problem of great current interest. Based on the reviewer’s comments
+and my own reading, I am recommending moderate revision. The authors’
+response should include a description of their changes, and also a
+version of the paper on which these can be seen (just as highlights in
+the text).}
 
-\textbf{Many of the reviewers’ comments address places where the paper could
-be more clearly written, and my overall comment would be the same. The
-current version perhaps spends a little too much time on the
+\textbf{Many of the reviewers’ comments address places where the paper
+could be more clearly written, and my overall comment would be the
+same. The current version perhaps spends a little too much time on the
 background and not enough on explaining the method and its derivation.
@@
 hyperparameter is) would be welcome to many readers; and increased
-clarity of presentation usu- ally means that the paper will be used
+clarity of presentation usually means that the paper will be used
 and cited more.}
 
+I appreciate this comment, and I have extensively revised the paper in
+order to make it more readable. To list some of the major changes:
+
+  1) I properly define a Gaussian process in Section 2. I elaborate on
+     its definition to give it a more tangible meaning, and I use
+     Brownian motion as an example of a Gaussian process.
+
+  2) I give more references, including section numbers for books, to
+     make it easier for the reader to get additional information.
+
+  3) In section 2 of the revised draft, I give an example of a prior
+     for transient displacements. I include a figure showing what the
+     covariance function looks like and what a corresponding
+     realization looks like.
+
+  4) I provide a clearer derivation of the posterior transient
+     displacements (eq. 6 and 7 of the revised draft).
+
+  5) In the revised draft, eq. 19 and 20 were added to help the reader
+     understand how the transient strain rates are computed.
+
+  6) I provide a clearer derivation of the signal-to-noise ratio.
+
+  7) I moved details of the outlier detection method to the appendix
+     because it is not essential to the text.
+
+  8) I expanded the explanation of the REML method, which is used to
+     constrain u and eta in Section 3.1 and 3.2.
+
+  9) Most of the paper was reworded / restructed to improve
+     clarity. Despite the fact that very little of the original text
+     is in the revised daft, there are almost no differences in the
+     content being presented.
+
 \textbf{I feel the same way about the section on detecting outliers: again, a
-well-worked area. Iter- ative fitting and outlier detection, as an
-overall strategy, hav e been used for a long time. I appreci- ate that
+well-worked area. Iterative fitting and outlier detection, as an
+overall strategy, hav e been used for a long time. I appreciate that
 by imposing spatial and temporal conditions you can identify outliers

### 2017-11-11 ResponseToReviewers/ResponseToReviewers.tex
did more

@@
 
+I recognize that outlier detection is a well-worked area. I included
+the description of the outlier detection method for completeness, and
+not because I thought it was an innovative or superior method. In
+order to keep the focus of my paper on my method for computing
+transient strain, I move the outlier detection method to the appendix.
+
+In my experimentation with the outlier detection algorithm, it turns
+out that the prior spatial covariance has virtually no effect on the
+detected outliers. However, the temporal covariance is important. The
+prior stochastic model needs to be flexible enough to describe
+deformation from, for example, SSEs. This is illustrated in Figure A1
+of the revised draft.
+
 \textbf{A few specifics (numbers are page+line):}
@@
 hazard.}
+
+Done
 
@@
 
+My previous description of Shen's method was poor and misleading. I
+have my description in the revised draft.
+
 \textbf{3+7: could also}
+
+Done
 
@@
 
+TODO
+
 \textbf{3+37: a geophysical signal}
+
+Done
 
@@
 
+Done
+
 \textbf{3+56: by a Gaussian, not with}
+
+This sentence has been removed in the revised draft
 
@@
 
+This sentence has been removed in the revised draft
+
 \textbf{6+56: I suspect that equation (15) took some effort to derive; could
 some details go into an Appendix?}
+
+I added a clearer derivation in the revised draft.
 
@@
 
+I have put a lot of thought into representing strain, and I recognize
+that strain is conventionally represented by showing the principle
+strains axes. However, there is no good way of showing the
+uncertainties for the principle strain axes. For example, consider a
+perfectly dilational strain that is perturbed by a small amount of
+noise. The orientation of principle strains will be highly uncertain,
+but the magnitudes of the principle strains will be well constrained.
+If we were to draw some sort of confidence interval around the tips of
+the principle strains, then they would have a crescent shape (not an
+ellipse). It is unclear how I would write a program to draw these
+confidence intervals.
+
+Instead of showing the principle strains, I am showing the normal
+strains as a function of azimuth. For a given azimuth, the uncertainty
+in the normal strain is computed by linear propagation of the
+uncertainties in the strain. In my opinion, it is much easier to show
+the uncertainties in an azimuthal representation of strain.
+
+I did not include this justification in the revised draft because I
+think it would be too much of a digression.
+
 \closing{Thank you for your continued consideration,}

### 2017-11-11 RevisedSubmission/hines_hetland_2017_gji.tex
did more

@@
 deformation, is also relevant for assessing seismic hazard. While
-transient strain itself is not damaging, there is a risk that it can
+transient strain itself is not damaging, there is a possibility that it can
 trigger major earthquakes \citep{Roeloffs2006,Freed2001}.
@@
 that control a similar trade-off between bias and variance in the
-solution. One could parameterize deformation with a physically
+solution. One could also parameterize deformation with a physically
 motivated model of interseismic deformation \citep[e.g.,][]{Meade2005,
@@
 strain rate magnitude, and we demonstrate that this is indeed an
-effective criterion for identifying geophysical signal. For the same
+effective criterion for identifying a geophysical signal. For the same
 reasons descibed above, the transient deformation and corresponding
@@
 rates from GNSS data. GPR is a Bayesian, non-parametric method for
-inferring a continuous signal from scattered data. Since GNSS stations
-are irregularly spaced and observation times may differ between
-stations, GPR is an ideal tool for synthesizing discrete GNSS data
-into a spatially and temporally continuous representation of transient
-strain. GPR is Bayesian in that we use a prior model to control the
-spatial and temporal roughness of the inferred transient strain. The
-prior is specified as a stochastic process, namely a Gaussian process.
-If there is no information available to help choose an appropriate
-prior Gaussian process, then maximum likelihood methods can be used to
-objectively choose one that is most consistent with the data. We use
-GPR to infer transient strain resulting from SSEs in the Pacific
-Northwest, demonstrating that GPR is an effective tool for detecting
-transient geophysical processes.
+inferring a continuous function from scattered data. Since GNSS
+stations are irregularly spaced and observation times may differ
+between stations, GPR is an ideal tool for synthesizing discrete GNSS
+data into a spatially and temporally continuous representation of
+transient strain. GPR is Bayesian in that we use a prior model to
+control the spatial and temporal roughness of the inferred transient
+strain. The prior is specified as a stochastic process, namely a
+Gaussian process. If there is no information available to help choose
+an appropriate prior Gaussian process, then maximum likelihood methods
+can be used to objectively choose one that is most consistent with the
+data. We use GPR to infer transient strain resulting from SSEs in the
+Pacific Northwest, demonstrating that GPR is an effective tool for
+detecting transient geophysical processes.
 

### 2017-11-11 ResponseToReviewers/ResponseToReviewers.tex
did more

@@
 
-We were pleased with the positive comments from you and the reviewers,
-and we feel that the comments brought up valid concerns. We hope that
-we have addressed these concerns sufficiently in our revision. In
-order to fully address the comments, we have included a supplementary
-material document.  We have also made some changes to the manuscript
-which were not specifically requested by the reviewers.  Most of these
-changes were stylistic and do not change the content of the
-manuscript.  One of the more substantial changes was that we merged
-Figures 7 and 14, which are both showing observed and predicted radial
-components of postseismic displacements, facilitating direct
-comparison of the figures.
+We are grateful for the comments from you and the reviewers. We are
+pleased that the comments are generally positive and that most of the
+concerns seem to be with the presentation of the material rather than
+the material itself. We agree that the previous draft required a
+substantial amount of clarification in order to be more accessible to
+geophysicists. Consequently, we have revised most of the methods
+section (Section 2) with the hopes of improving clarity. Details of
+our revisions are given below in our response to the reviewers'
+comments.
 
-We added supporting figures S3 and S4, which break down the predicted
-displacements of our preferred model into an elastic and viscoelastic
-component.  We feel that these figures help support our discussion on
-lines 509 to 523, where we attribute inferences of fault slip and
-viscoelastic relaxation to different aspects of the observed
-postseismic displacements.
+In the revised draft, we have addressed several issues that we found,
+which were not brought up by the reviewers:
 
-On line 500 to 502 of our previous version of this manuscript we make
-the claim that “After one year, afterslip is inferred to be deeper
-down on the Sierra Cucapah segment, which is describing much of the
-sustained near-field postseismic deformation.” This sentence is not
-accurate of the preferred model we present in the manuscript. As can
-be clearly seen in the new figures S3-S4, afterslip is describing some
-of the later near-field deformation, although most of it is being
-described by viscoelastic relaxation.  We have reworded this passage
-and updated the introduction and conclusion to accurately describe our
-interpretations of the mechanisms driving later near-field
-deformation.  This does not significantly alter the main conclusions
-of our manuscript, because we are very clear, both in the original and
-revised versions, that the deformation mechanism driving later
-near-field deformation cannot be definitively resolved with these
-data.
+1) The strain covariance function, eq. 14 of the previous draft, was
+incorrect. That covariance function was only correct for the shear
+strain. The normal strains differ by a factor of 2. This error was NOT
+in our software implementation of the method, and so our presented
+results have not changed.
 
-Each of the reviewer comments are included below followed by our
-response and a description of how these comments were addressed in
-this revised manuscript.
+2) There was an error in eq. 20 of the previous draft. We used a Dirac
+delta function where we should have used a Kronecker delta.
+
+3) In the previous draft, Figure 5 was the only figure that showed how
+well our model fits the GNSS data. The data consists of seven years of
+daily displacements from 94 stations, so it is not possible to show
+all the data. However, we are now showing the fit at 3 stations,
+rather than 1. These stations were chosen to illustrate a slight
+over-smoothing in our fit to the data.
+
+4) In addition to clarifying Section 2, we have significantly
+rewritten Sections 3 (application to the Pacific Northwest) and
+Section 4 (the discussion) so that they have a clearer focus and
+structure. However, the content being presented has not changed.
+
+Each of the reviewers comments are written below and followed by our
+response. We hope that we have addressed their concerns sufficiently
+in our revision.
 
@@
 
-I have taken several steps to make my revised draft more accessible to
-someone who does not have a strong background in stochastic processes:
-I rewritten the methods section to be clearer, avoid esoteric jargon,
-contain more examples, and elaborate on potentially unfamiliar
-concepts. I have also added more references to the methods section to
-make it easier for the reader to find additional information. I
-sincerely hope that my revised draft is more readable to
-geophysicists.
+I have taken several steps to make my revised draft more readable for
+geophysicists. For details, see my response to the editor's comment
+``Many of the reviewers’ comments ... ".
 
@@
 
-(See my response to the comment ``Also, the paper would be improved if
-the proposed outlier detection ...")
+See my response to the comment ``Also, the paper would be improved if
+the proposed outlier detection ..."
 
@@
 
-  9) Most of the paper was reworded / restructed to improve
+  9) I attempt to avoid esoteric jargon such as "hyperparameter". In
+     the revised draft, "hyperparameter" has been replaced with
+     "parameter", which does not degrade the precision of our
+     description of the method.
+
+  10) Most of the paper was reworded / restructed to improve
      clarity. Despite the fact that very little of the original text
@@
 
-My previous description of Shen's method was poor and misleading. I
-have my description in the revised draft.
+I have updated my description of Shen's method in the revised draft.
 
@@
 
-TODO
+The text has been clarified say "The method described in Holt 2013 was
+also tested in the SCEC transient detection exercise. Holt 2013
+detects transient signal from an estimate of transient strain rates,
+and they estimate transient strain rates using a ..."
 

### 2017-11-11 ResponseToReviewers/ResponseToReviewers.tex
handed over to eric

@@
 matrix $C_{\eta}$ (i.e. line 10 p5). This approach is not common in
-modelling GNSS time series. The standard is based on Williams 2003,
+modeling GNSS time series. The standard is based on Williams 2003,
 where the author established a stochastic model with a covariance
@@
 \textbf{2- First, the author should give a summary of stochastic noise
-modelling in GNSS. I also think that a discussion is needed to relate
+modeling in GNSS. I also think that a discussion is needed to relate
 to previous models. Why using separated stochastic models? Also I
@@
 
-I believe the reviewer is refering to page 4-5 of the original draft,
+I believe the reviewer is referring to page 4-5 of the original draft,
 where I first introduce $\eta$. I agree that a more thorough
@@
 
-Again, we think that the description of the noise model should remain
+Again, I think that the description of the noise model should remain
 in Section 3 because it is specific to the Pacific Northwest
@@
 
-We have added a few more references in Section 3.1 of the revised
-draft to emphasize that GNSS noise modeling is a well-worked area. We
-refer the reader to Bock and Melgar 2016 and He et al 2017 for a more
+I have added a few more references in Section 3.1 of the revised draft
+to emphasize that GNSS noise modeling is a well-worked area. I refer
+the reader to Bock and Melgar 2016 and He et al 2017 for a more
 complete review.
@@
 \textbf{p.20 line 33 “ computational burden” I think that refers to p.20 line
-60 “prohibitive when using sveral years of daily GNSS”. It would
+60 “prohibitive when using several years of daily GNSS”. It would
 interesting to have some number. How do you quantify the computation
@@
 from eta is that u is negligible for inland stations. This allowed us
-to islolate and constrain eta first, and then we constrained u. If we
+to isolate and constrain eta first, and then we constrained u. If we
 allowed eta to be spatially correlated (which we did not), then we
@@
 noise at all the stations considered in this study, which is probably
-a poor assumption since the inland stations are subject to distincly
+a poor assumption since the inland stations are subject to distinctly
 different climatic conditions".
@@
   1) I properly define a Gaussian process in Section 2. I elaborate on
-     its definition to give it a more tangible meaning, and I use
-     Brownian motion as an example of a Gaussian process.
+     its definition to give it a more tangible meaning.
 
@@
 
-  10) Most of the paper was reworded / restructed to improve
+  10) Most of the paper was reworded / restructured to improve
      clarity. Despite the fact that very little of the original text
@@
 well-worked area. Iterative fitting and outlier detection, as an
-overall strategy, hav e been used for a long time. I appreciate that
+overall strategy, ha e been used for a long time. I appreciate that
 by imposing spatial and temporal conditions you can identify outliers
-not otherwise obvi- ous, but it would help to show an example of this,
+not otherwise obvious, but it would help to show an example of this,
 rather than the current Figure 4, in which most of the identified
@@
 adaptive kernel smoother, somewhat like applying 2-D loess; so it is
-only depen- dent on nearby points, and has no assumption of uniformity
+only dependent on nearby points, and has no assumption of uniformity
 anywhere. (Actually this raises a more general question: is the method
@@
 
-The text has been clarified say "The method described in Holt 2013 was
-also tested in the SCEC transient detection exercise. Holt 2013
+The text has been clarified to say "The method described in Holt 2013
+was also tested in the SCEC transient detection exercise. Holt 2013
 detects transient signal from an estimate of transient strain rates,
@@
 
-\closing{Thank you for your continued consideration,}
+\closing{Thank you for your consideration,}
 

### 2017-11-11 RevisedSubmission/hines_hetland_2017_gji.tex
handed over to eric

@@
 over some subnetwork of the GNSS stations \citep[e.g.,][]{Feigl1990,
-Murray2000}. The components of the strain rate tensor for each
+Murray2000}. The components of the secular strain rate tensor for each
 subnetwork are then determined from the least squares fit to the
@@
 help overcome this deficiency, \citet{Shen1996,Shen2015} used an
-inverse distance weighting scheme, in which the estimated strain rate
-at a point is primarily controlled by observations at nearby stations.
-The method of \citet{Shen1996,Shen2015} can be viewed as a form of
-local least squares regression with a first-order polynomial. Other
-methods for estimating secular strain rates have parameterized GNSS
-derived velocities with bi-cubic splines \citep{Beavan2001}, spherical
-wavelets \citep{Tape2009}, and elastostatic Green's functions
-\citep{Sandwell2016}. The type of basis functions and the number of
-degrees of freedom for a parameterization, which are often chosen
-subjectively, have a strong influence on the strain solution. If there
-are too few degrees of freedom in the parameterization, then estimated
-strain rates will be biased and the uncertainties will be
-underestimated. On the other hand, if there are too many degrees of
-freedom, then there will not be any coherent features in the estimated
-strain rates. The methods described by \citet{Beavan2001} and
-\citet{Tape2009} also require the user to specify penalty parameters
-that control a similar trade-off between bias and variance in the
-solution. One could also parameterize deformation with a physically
-motivated model of interseismic deformation \citep[e.g.,][]{Meade2005,
-McCaffrey2007}. In such models the lithospheric rheology and fault
-geometries are assumed to be known. Any errors in the assumed physical
-model could result in biased strain estimates and underestimated
+inverse distance weighting scheme, in which the estimated secular
+strain rate at a point is primarily controlled by observations at
+nearby stations. The method of \citet{Shen1996,Shen2015} can be viewed
+as a form of local least squares regression with a first-order
+polynomial. Other methods for estimating secular strain rates have
+parameterized GNSS derived velocities with bi-cubic splines
+\citep{Beavan2001}, spherical wavelets \citep{Tape2009}, and
+elastostatic Green's functions \citep{Sandwell2016}. The type of basis
+functions and the number of degrees of freedom for a parameterization,
+which are often chosen subjectively, have a strong influence on the
+solution. If there are too few degrees of freedom in the
+parameterization, then estimated secular strain rates will be biased
+and the uncertainties will be underestimated. On the other hand, if
+there are too many degrees of freedom, then there will not be any
+coherent features in the estimated secular strain rates. The methods
+described by \citet{Beavan2001} and \citet{Tape2009} also require the
+user to specify penalty parameters that control a similar trade-off
+between bias and variance in the solution. One could also parameterize
+deformation with a physically motivated model of interseismic
+deformation \citep[e.g.,][]{Meade2005, McCaffrey2007}. In such models
+the lithospheric rheology and fault geometries are assumed to be
+known. Any errors in the assumed physical model could result in biased
+estimates of the secular strain rates and underestimated
 uncertainties.
@@
 Center (SCEC) community has shown interest in developing methods for
-detecting transient strain. SCEC supported a transient detection
+detecting transient deformation. SCEC supported a transient detection
 exercise \citep{Lohman2013}, where several research groups tested
@@
 method described in \citet{Holt2013} was also tested in the SCEC
-transient detection exercise, which calculates strain rates using a
-bi-cubic spatial parameterization of displacements between time
-epochs. \citet{Holt2013} defined a detection threshold based on the
-strain rate magnitude, and we demonstrate that this is indeed an
-effective criterion for identifying a geophysical signal. For the same
-reasons descibed above, the transient deformation and corresponding
-uncertainties estimated by these methods can be biased by the chosen
-spatial parameterization. It is then difficult to distinguish signal
-from noise with these methods, which limits their utility for
-transient detection.
+transient detection exercise. \citet{Holt2013} detect transient
+signal from an estimate of strain rates, and they estimate strain
+rates using a bi-cubic spatial parameterization of displacements
+between time epochs. \citet{Holt2013} define a detection threshold
+based on the strain rate magnitude, and we demonstrate that this is
+indeed an effective criterion for identifying a geophysical signal.
+For the same reasons described above, the transient deformation and
+corresponding uncertainties estimated by these methods can be biased
+by the chosen spatial parameterization. It is then difficult to
+distinguish signal from noise with these methods, which limits their
+utility for transient detection.
 
@@
 Here we propose using Gaussian process regression (GPR)
-\citep{Rasmussen2006} to estimate transient strain from GNSS data. GPR
-is closely related to kriging \citep{Cressie1992} and least squares
-collocation \citep{Moritz1978}. The latter has been used by
+\citep{Rasmussen2006} to estimate transient strain rates from GNSS
+data. GPR is closely related to kriging \citep{Cressie1992} and least
+squares collocation \citep{Moritz1978}. The latter has been used by
 \citet{Kato1998} and \citet{El-Fiky1999} to estimate secular strain
@@
 data into a spatially and temporally continuous representation of
-transient strain. GPR is Bayesian in that we use a prior model to
-control the spatial and temporal roughness of the inferred transient
-strain. The prior is specified as a stochastic process, namely a
-Gaussian process. If there is no information available to help choose
-an appropriate prior Gaussian process, then maximum likelihood methods
-can be used to objectively choose one that is most consistent with the
-data. We use GPR to infer transient strain resulting from SSEs in the
-Pacific Northwest, demonstrating that GPR is an effective tool for
-detecting transient geophysical processes.
+transient strain rates. GPR is Bayesian in that we use a prior model
+to control the spatial and temporal roughness of the inferred
+transient strain rates. The prior is specified as a stochastic
+process, namely a Gaussian process. If there is no information
+available to help choose an appropriate prior Gaussian process, then
+maximum likelihood methods can be used to objectively choose one that
+is most consistent with the data. We use GPR to infer transient strain
+rates resulting from SSEs in the Pacific Northwest, demonstrating that
+GPR is an effective tool for detecting transient geophysical
+processes.
 
@@
 any collection of points $\points$. A realization of the random vector
-$[u_i(p)]_{p \in \points}$ can be interpretted as a realization of
+$[u_i(p)]_{p \in \points}$ can be interpreted as a realization of
 $u_i$ (i.e., a sample function) that is evaluated at $\points$. Just
@@
 $T_i$ and the parameters $\ell = 100$ km, $\tau = 10$ days, and $\phi
-= 1$ mm. A single realization of $u_i$ corresponsing to that choice of
+= 1$ mm. A single realization of $u_i$ corresponding to that choice of
 covariance function is shown in Figure \ref{fig:PriorDemo}B (See
@@
 from instability in a geodetic monument \citep[e.g.,][]{Wyatt1982,
-Wyatt1989}. In particular, the power spectrums for Brownian motion and
+Wyatt1989}. In particular, the power spectra for Brownian motion and
 monument motion both decay proportionally to $f^{-2}$, where $f$ is
@@
 subduction zone that they are unlikely to record transient deformation
-from SSEs, allowing us to ingore the term $u_i(\points)$ in
+from SSEs, allowing us to ignore the term $u_i(\points)$ in
 $\mitbf{d}_i$. We assume the noise at these inland stations is
@@
 calibrated with a well known strain source, such as diurnal and
-semidiurnal tides \citep{Hart1996,Roeloffs2010,Hodgkinson2013}.
+semi-diurnal tides \citep{Hart1996,Roeloffs2010,Hodgkinson2013}.
 Unfortunately, the tidal forces at BSMs which record SSEs are strongly

### 2017-11-12 RevisedSubmission/hines_hetland_2017_gji.tex
added more

@@
 
-Transient strain rates derived from GNSS data can be used to detect
-and understand geophysical processes such as slow slip events and
+Transient strain derived from GNSS data can be used to detect and
+understand geophysical processes such as slow slip events and
 postseismic deformation. Here we propose using Gaussian process
-regression (GPR) as a tool for estimating transient strain rates from
-GNSS data. GPR is a non-parametric, Bayesian method for interpolating
+regression (GPR) as a tool for estimating transient strain from GNSS
+data. GPR is a non-parametric, Bayesian method for interpolating
 scattered data. In our approach, we assume a stochastic prior model
@@
 transient displacements to covary spatially and temporally. A
-posterior estimate of transient strain rates is obtained by
-differentiating the posterior displacements, which are formed by
+posterior estimate of transient strain is obtained by differentiating
+the posterior transient displacements, which are formed by
 conditioning the prior with the GNSS data. As a demonstration, we use
@@
 strain rates have been used to constrain seismic hazard models such as
-UCERF3 \citep{Field2014}. Transient crustal strain, which is caused by
-geophysical processes such as slow slip events (SSEs) or postseismic
-deformation, is also relevant for assessing seismic hazard. While
-transient strain itself is not damaging, there is a possibility that it can
-trigger major earthquakes \citep{Roeloffs2006,Freed2001}.
+UCERF3 \citep{Field2014}. Transient crustal strain, which is anomalous
+strain caused by geophysical processes such as slow slip events (SSEs)
+or postseismic deformation, is also relevant for assessing seismic
+hazard. While transient strain itself is not damaging, there is a
+possibility that it can trigger major earthquakes
+\citep{Roeloffs2006,Freed2001}.
 
@@
 subnetwork are then determined from the least squares fit to the
-observations. The assumption that deformation gradients are spatially
+velocities. The assumption that deformation gradients are spatially
 uniform is not appropriate when subnetworks span a very large area. To
@@
 inverse distance weighting scheme, in which the estimated secular
-strain rate at a point is primarily controlled by observations at
+strain rate at any point is primarily controlled by the velocities at
 nearby stations. The method of \citet{Shen1996,Shen2015} can be viewed
@@
 deformation with a physically motivated model of interseismic
-deformation \citep[e.g.,][]{Meade2005, McCaffrey2007}. In such models
+deformation \citep[e.g.,][]{Meade2005, McCaffrey2007}. In such models,
 the lithospheric rheology and fault geometries are assumed to be
@@
 method described in \citet{Holt2013} was also tested in the SCEC
-transient detection exercise. \citet{Holt2013} detect transient
-signal from an estimate of strain rates, and they estimate strain
-rates using a bi-cubic spatial parameterization of displacements
-between time epochs. \citet{Holt2013} define a detection threshold
-based on the strain rate magnitude, and we demonstrate that this is
-indeed an effective criterion for identifying a geophysical signal.
-For the same reasons described above, the transient deformation and
-corresponding uncertainties estimated by these methods can be biased
-by the chosen spatial parameterization. It is then difficult to
-distinguish signal from noise with these methods, which limits their
-utility for transient detection.
+transient detection exercise. \citet{Holt2013} detect transient signal
+from an estimate of transient strain rates, and they estimate
+transient strain rates using a bi-cubic spatial parameterization of
+displacements between time epochs. \citet{Holt2013} define a detection
+threshold based on the transient strain rate magnitude, and we
+demonstrate that this is indeed an effective criterion for identifying
+a geophysical signal. For the same reasons described above, the
+transient deformation and corresponding uncertainties estimated by
+these methods can be biased by the chosen spatial parameterization. It
+is then difficult to distinguish signal from noise with these methods,
+which limits their utility for transient detection.
 
@@
 seasonal terms in $\G$. We let the temporal covariance of $v$ be a
-squared exponential, and we let $v$ be spatially uncorrelated so that,
+squared exponential, and we let $v$ be spatially uncorrelated so that
 \begin{equation}

### 2017-11-25 ResponseToReviewers/ResponseToReviewers.tex
made erics edits

@@
 not think this would be an appropriate place to discuss a scenario in
-which we do not know u or eta. I also think that it should go with
-saying that the validity of our results are contingent on having a
-good model for u and eta.
+which we do not know u or eta. I also think that it is clear that our
+results are contingent on having a good model for u and eta. Asking
+whether the strain signal can be resolved without a prior statistical
+model for u or eta is asking whether the signal can be separated from
+the noise if one does not know anything about the noise or the signal.
+I think it goes without stating that signal can not be extracted from
+observations without a model of the noise or an assumption of the
+signal sought.
 
@@
 from eta is that u is negligible for inland stations. This allowed us
-to isolate and constrain eta first, and then we constrained u. If we
-allowed eta to be spatially correlated (which we did not), then we
+to isolate and characterize eta first, and then we characterized u. If
+we allowed eta to be spatially correlated (which we did not), then we
 would have conceivably been able to describe the spatially and
@@
 \textbf{Figure 6: I like that the strain uncertainties can be shown, but is
-there a reason not to plot principle strains using arrows, showing
+there a reason not to plot principal strains using arrows, showing
 errors by an ellipse around the tips?}
@@
 I have put a lot of thought into representing strain, and I recognize
-that strain is conventionally represented by showing the principle
+that strain is conventionally represented by showing the principal
 strains axes. However, there is no good way of showing the
-uncertainties for the principle strain axes. For example, consider a
+uncertainties for the principal strain axes. For example, consider a
 perfectly dilational strain that is perturbed by a small amount of
-noise. The orientation of principle strains will be highly uncertain,
-but the magnitudes of the principle strains will be well constrained.
+noise. The orientation of principal strains will be highly uncertain,
+but the magnitudes of the principal strains will be well constrained.
 If we were to draw some sort of confidence interval around the tips of
-the principle strains, then they would have a crescent shape (not an
+the principal strains, then they would have a crescent shape (not an
 ellipse). It is unclear how I would write a program to draw these
@@
 
-Instead of showing the principle strains, I am showing the normal
+Instead of showing the principal strains, I am showing the normal
 strains as a function of azimuth. For a given azimuth, the uncertainty

### 2017-11-26 ResponseToReviewers/ResponseToReviewers.tex
finished revisions

@@
 and uncorrelated noise. I then break $\eta$ down into correlated and
-uncorrelated components in Section 3.1 (eq. 25). The covariance matrix
-for $\eta$ in eq. 26 is more consistent with the literature.
+uncorrelated components in Section 3.1 (eq. 24). The covariance matrix
+for $\eta$ in eq. 25 is more consistent with the literature.
 
@@
 noise model for Section 3. This is clarified in Section 2 of the
-revised draft when I say "The appropriate noise model may vary
+revised draft when I say ``The appropriate noise model may vary
 depending on the application, and we hold off on specifying the
@@
 
-By "formal data uncertainties" I meant "the derived uncertainties that
+By ``formal data uncertainties" I meant ``the derived uncertainties that
 accompany the GNSS displacement solutions". The revised text has been
@@
 on Cascadia SSEs. In the revised draft, we refer the reader to the
-review by Schartz and Rokosky 2007.
+review by Schwartz and Rokosky 2007.
 
@@
 In Section 3.1 of the revised draft, where we elaborate on the noise
-model, we describe common mode error by saying "Another significant
+model, we describe common mode error by saying ``Another significant
 source of noise in GNSS data is common mode error (e.g., Wdowinski et
@@
 However, Wendland covariance functions are typically described as
-"compact" in the literature, and I think it is a key adjective that
+``compact" in the literature, and I think it is a key adjective that
 should not be dropped out. I have left the wording as is.
@@
 statistical properties are invariant to translations. We have replace
-"isotropic" with "stationary" in the text because "stationary" is a
+``isotropic" with ``stationary" in the text because ``stationary" is a
 more appropriate, and perhaps more familiar, term.
@@
 reasoning and implications for this assumption. The paragraph starts
-with "It should be noted that we have ignored any covariances ...".
+with ``It should be noted that we have ignored any covariances ...".
 
@@
 At this point in the text, it is assumed that the statistical
-properties of u and eta are known, although not yet specified. So I do
+properties of u and $\eta$ are known, although not yet specified. So I do
 not think this would be an appropriate place to discuss a scenario in
-which we do not know u or eta. I also think that it is clear that our
-results are contingent on having a good model for u and eta. Asking
+which we do not know u or $\eta$. I also think that it is clear that our
+results are contingent on having a good model for u and $\eta$. Asking
 whether the strain signal can be resolved without a prior statistical
-model for u or eta is asking whether the signal can be separated from
+model for u or $\eta$ is asking whether the signal can be separated from
 the noise if one does not know anything about the noise or the signal.
@@
 In Section 3.1 and 3.2, we discuss how we choose a model for u and
-eta. The key piece of "prior knowledge" that allows us to discern u
-from eta is that u is negligible for inland stations. This allowed us
-to isolate and characterize eta first, and then we characterized u. If
-we allowed eta to be spatially correlated (which we did not), then we
+$\eta$. The key piece of ``prior knowledge" that allows us to discern u
+from $\eta$ is that u is negligible for inland stations. This allowed us
+to isolate and characterize $\eta$ first, and then we characterized u. If
+we allowed $\eta$ to be spatially correlated (which we did not), then we
 would have conceivably been able to describe the spatially and
@@
 representative of the entire region. We added the disclaimer saying
-"We assume the noise at these inland stations is representative of the
+``We assume the noise at these inland stations is representative of the
 noise at all the stations considered in this study, which is probably
@@
   8) I expanded the explanation of the REML method, which is used to
-     constrain u and eta in Section 3.1 and 3.2.
+     constrain u and $\eta$ in Section 3.1 and 3.2.
 
-  9) I attempt to avoid esoteric jargon such as "hyperparameter". In
-     the revised draft, "hyperparameter" has been replaced with
-     "parameter", which does not degrade the precision of our
+  9) I attempt to avoid esoteric jargon such as ``hyperparameter". In
+     the revised draft, ``hyperparameter" has been replaced with
+     ``parameter", which does not degrade the precision of our
      description of the method.
@@
 
-The text has been clarified to say "The method described in Holt 2013
+The text has been clarified to say ``The method described in Holt 2013
 was also tested in the SCEC transient detection exercise. Holt 2013
@@
 
+\vspace{1cm}
 \closing{Thank you for your consideration,}

### 2017-11-26 revised_source_files/hines_hetland_2017_gji.tex
ready to resubmit

--- 2017-11-12 RevisedSubmission/hines_hetland_2017_gji.tex
@@
 strain rates have been used to constrain seismic hazard models such as
-UCERF3 \citep{Field2014}. Transient crustal strain, which is anomalous
-strain caused by geophysical processes such as slow slip events (SSEs)
-or postseismic deformation, is also relevant for assessing seismic
+UCERF3 \citep{Field2014}. Transient strain, which is anomalous strain
+caused by geophysical processes such as slow slip events (SSEs) or
+postseismic deformation, is also relevant for assessing seismic
 hazard. While transient strain itself is not damaging, there is a
@@
 compute transient strain rates (See \citet[sec. 2.2]{Adler1981} or
-\citet[sec. 10A]{Papoulis1991} for a definition of stochastic
+\citet[appx. 10A]{Papoulis1991} for a definition of stochastic
 differentiation and the conditions for differentiability).
@@
 We provide an example to give the prior on transient displacements a
-more tangible meaning. we can use a squared exponential function for
+more tangible meaning. We can use a squared exponential function for
 $X_i$ and $T_i$,
@@
 covariance function is shown in Figure \ref{fig:PriorDemo}B (See
-\citet[sec. A3]{Rasmussen2006} for details on drawing realization from
+\citet[appx. A.2]{Rasmussen2006} for details on drawing realization from
 Gaussian processes). While the squared exponential function is a
@@
 It should be noted that we have ignored any covariances between the
-easting and northing components of $\vec{u}$ and $\vec{\data}$.
-This simplification reduces the computational complexity of evaluating
-the posterior transient displacements because each component can be
+easting and northing components of $\vec{u}$ and $\vec{\data}$. This
+simplification reduces the computational complexity of evaluating the
+posterior transient displacements because each component can be
 evaluated independently. However, we are inherently assuming that the
-principle axes describing the distribution of $\vec{u}(p)$ and
-$\vec{d}^{(jk)}$ are aligned with the cardinal directions, which is
-admittedly an arbitrary assumption.
+principal components of $\vec{u}(p)$ and $\vec{d}^{(jk)}$ are aligned
+with the cardinal directions, which is admittedly an arbitrary
+assumption.
 
@@
 processes $\post_i$ and $\post_j$, we know that $\strain_{ij}$ is also
-a Gaussian process. From \citet[sec. 10]{Papoulis1991}, we find that
+a Gaussian process. From \citet[ch. 10]{Papoulis1991}, we find that
 the mean and covariance functions for $\dot{\varepsilon}_{ij}$ are
@@
 signal can be easily identified by visually inspecting the solution
-for $\strain$ from eqs. (\ref{eq:StrainMean}) and
-(\ref{eq:StrainCov}). However, if we want to detect geophysical signal
+for $\strain$. However, if we want to detect geophysical signal
 automatically, then we need to define a detection criterion. We use a
@@
 \begin{align}
-\mu_{||\strain||_\mathrm{F}} &\approx ||\strain||_F \big|_{\strain = \mu_{\strain}}\nonumber \\
+\mu_{||\strain||_\mathrm{F}} &\approx ||\strain||_F \big|_{\strain = \mu_{\strain}} \nonumber \\
                              &= \left(\mu_{\strain_{\e\e}}^2 + 
@@
 \begin{align}\label{eq:SNR}
-\mathrm{SNR}(p) &= \frac{\mu_{||\strain||_\mathrm{F}}(p)}{\sigma_{||\strain||_\mathrm{F}}(p)} \\
+\mathrm{SNR}(p) &= \frac{\mu_{||\strain||_\mathrm{F}}(p)}{\sigma_{||\strain||_\mathrm{F}}(p)} \nonumber \\
                 &= \frac{\mu_{\strain_{\e\e}}(p)^2 +
@@
 strain resulting from SSEs. SSEs in the Pacific Northwest were first
-discovered by \citep[e.g.,][]{Dragert2001} and we refer the reader to
-\citet{Schwartz2007} for a review of the subsequent research. Before
-estimating transient strain rates, we establish a noise model for GNSS
-stations in this region, and we establish a prior Gaussian process to
-describe displacements from SSEs. SSEs in the Pacific Northwest can be
-detected by monitoring for associated seismic tremor
-\citep{Rogers2003}, which is actively being done by the Pacific
-Northwest Seismic Network \citep{Wech2010}. We can then compare the
-tremor records to our estimated transient strain rates to verify that
-we are indeed identifying strain from SSEs.
+discovered by \citet{Dragert2001} and we refer the reader to
+\citet{Schwartz2007} for a review of the subsequent research. SSEs in
+the Pacific Northwest can be detected by monitoring for associated
+seismic tremor \citep{Rogers2003}, which is actively being done by the
+Pacific Northwest Seismic Network \citep{Wech2010}. We can then
+compare the tremor records to our estimated transient strain rates to
+verify that we are indeed identifying strain from SSEs. Before
+estimating transient strain rates, we first establish a noise model
+for GNSS stations in this region, and then we establish a prior
+Gaussian process to describe displacements from SSEs.
 
@@
 henceforth ignore temporally correlated noise in the easting and
-northing component because of its low amplitude. We also do not use
+northing components because of its low amplitude. We also do not use
 vertical displacements because of the presumably low signal-to-noise
@@
 Note that $T_i$ includes the parameter $\phi$, which serves to scale
-the covariance function $C_{u_i}$. The second option is a member of
+the covariance function $C_{u_i}$. A second option is a member of
 the Wendland class of covariance functions \citep{Wendland2005}.
-Wendland covariance functions have compact support and hence their
+Wendland covariance functions have compact support, and thus their
 corresponding covariance matrices are sparse. We exploit this sparsity
@@
 \end{equation} 
-where $(t)_+ = \mathrm{max}(0,t)$. The third option for $T_i$ is the
+where $(t)_+ = \mathrm{max}(0,t)$. A third option for $T_i$ is the
 covariance function for integrated Brownian motion (IBM). IBM is a
@@
 \begin{align}\label{eq:IBM}
-T_i(t,t') &= \int_0^t \int_0^{t'} \phi^2 \min(s,s') \,ds'\,ds \\
+T_i(t,t') &= \int_0^t \int_0^{t'} \phi^2 \min(s,s') \,ds'\,ds \nonumber \\
           &= \frac{\phi^2}{2}\min(t,t')^2 \left(\max(t,t') - \frac{1}{3}\min(t,t')\right), \ \ \ t,t' \geq 0.
@@
 have this reference time. For example, if we are trying to identify
-postseismic deformation then $t$ should be zero at the time of the
+postseismic deformation, then $t$ should be zero at the time of the
 earthquake.  However, if we are interested in detecting transient
@@
 Vancouver Island, coinciding with where fault slip tends to be
-inferred \citep[e.g.,][]{Dragert2001, Wech2009, Schmidt2010}. Thus,
+inferred \citep[e.g.,][]{Dragert2001, Wech2009, Schmidt2010}. Hence,
 this pattern of strain is to be expected. Over decadal time-scales,
@@
 \caption{
-Transient strain rates during the Winter 2015-2016 SSE. The glyphs
+Transient strain rates during the winter 2015-2016 SSE. The glyphs
 show the normal strain rates as a function of azimuth, where orange
 indicates compression and blue indicates extension. The shaded regions
-indicate the one standard deviation uncertainties for the normal
-strain rates.
+show the one standard deviation uncertainties for the normal strain
+rates.
 }   
@@
 exploration. However, a discussion on features in $\strain$ that have
-unknown origin would be outside the scope of this paper.
+unknown origin is beyond the scope of this paper.
 
@@
 sparsity decreases when the Wendland covariance function has a larger
-time-scale parameter. The Wendland covariance then offers less of a
-computational advantage when we were interested in geophysical
-processes that occur over longer times-scales. An alternative approach
-to dealing with large datasets would be to explore approximation
-methods for GPR \citep[sec. 8]{Rasmussen2006}.
+time-scale parameter. The Wendland covariance function then offers
+less of a computational advantage when we are interested in
+geophysical processes that occur over longer times-scales. An
+alternative approach to dealing with large datasets would be to
+explore approximation methods for GPR \citep[ch. 8]{Rasmussen2006}.
 
@@
 $\mitbf{w}$ is a vector of uncorrelated Gaussian noise with known
-standard deviations $\mitbf{\sigma} = [\sigma_1, \sigma_2, ...,
-\sigma_n]$. The Gaussian process $v$ is intended to describe transient
-features in the data that cannot be explained by the linear trend or
-seasonal terms in $\G$. We let the temporal covariance of $v$ be a
-squared exponential, and we let $v$ be spatially uncorrelated so that
+standard deviations $\mitbf{\sigma}$. The Gaussian process $v$ is
+intended to describe transient features in the data that cannot be
+explained by the linear trend or seasonal terms in $\G$. We let the
+temporal covariance of $v$ be a squared exponential, and we let $v$ be
+spatially uncorrelated so that
 \begin{equation}
@@
 consider outliers to be data that are poorly explained by the model
-$\G\mitbf{m} + v(\points)$, which is determine by the residual vector
+$\G\mitbf{m} + v(\points)$, which are determined by the residual vector
 \begin{align}\label{eq:Residual} 

### 2017-11-26 revised_source_files/compile.sh

@@ -1,0 +1,6 @@
+#!/bin/bash
+
+pdflatex -synctex=1 -interaction=nonstopmode hines_hetland_2017_gji.tex
+bibtex hines_hetland_2017_gji.aux
+pdflatex -synctex=1 -interaction=nonstopmode hines_hetland_2017_gji.tex
+pdflatex -synctex=1 -interaction=nonstopmode hines_hetland_2017_gji.tex
